\begin{college}[1.1]{binder}{Binding and Animating Magics}{BI}

\textbf{This College is a probationary College. This means that this
College is under trial for two years (as of April 1999). All characters
that join this College do so under the understanding that if this
College is deemed unplayable, it may be withdrawn or radically
changed. At that point, all such characters may have the EP and time
spent in this College refunded, to be re-spent freely in another
College as if that College had been their original choice.}

Members of the College of Binding and Animating Magics specialise in
the binding, manipulation and animation of non-living matter; they are
commonly known as Binders. Binders tend to be less concerned with the
theory and philosophy of magic than members of the other
Thaumaturgical Colleges, and are usually more inclined to tinker
around until something works. Binders are often accused of lacking
empathy, because of the amount of time they spend associating with
inanimate objects. However, many Binders cultivate an impish sense of
humour, which their College gives ample opportunity to develop. The
College has been cloistered until relatively recently, and it is not
well known. Much of the Binders' knowledge was lost during the Fall.
They are found almost exclusively in highly developed and civilised
areas.

Most Colleges are scornful of Binders, due to their reputations as
tricksters and their lack of direct combat magics.  They are also
mocked for their strong association with the artisan trades, such as
Carpentry, Sculpting and Smithing. Many Binders possess at least one
such skill, Mechanician, or Philosopher specialising in engineering,
architecture and the like, as all of these abilities complement the
College's mechanistic style of magic

\subsubsection{Traditional Colours}

Binders do not have traditional colours, but tend to wear practical
work clothes, usually with a leather apron anchored in place, and
bedecked with tools.

\subsubsection{Traditional Symbols}

The College's traditional symbol is that of many cogwheels, each
turning others, in an endless chain. However, the populace generally
associates the College with the Rag and String Golems that many
Binders use as followers and helpers.

\subsection{Restrictions}

A Magical Aptitude of 17 is required to join the College.  Members of
the College of Binding and Animating Magics may operate without
restriction.

\subsection{Base Chance Modifiers}

The following modifiers are cumulative with all other modifiers
(including those specified in \S\ref{magic:modifiers}).

\begin{tabularx}{\linewidth}{Xr}
Rituals: Per 10\% extra spent on Ritual materials & +1 \\
Spells: Per hour of Ritual Spell Preparation    & +1 \\
Per Rank in Mechanician or Philosopher    & +1 \\
\end{tabularx}

Note: Rank in Mechanician OR Philosopher may be applied, not both.

\subsection{Definitions}

\begin{Description}
\item[Assistants] An assistant may be used to provide skills or
abilities which the Binder does not possess. The assistant must be
present throughout the entire ritual. If using an assistant's skill to
craft a golem, the assistant's rank may affect the PB of the golem but
not the base chance of performing the ritual.

\item[Construction Time] Each golem type has a base construction
time. This is multiplied by the height of the golem in feet (round
up).  This assumes that the adept has the requisite quantity of
materials on hand and the tools or ability to shape the material.

\item[Crafting Golems] To perform a Shaping Golem ritual the binder
must have all of the materials on hand. They must also have the tools
and skill to shape the material, or an assistant with the appropriate
tools and skills.

\item[Magical Materials] The following statistics are for golems made
of non-magical materials. Golems made of magical, enchanted or
formerly enchanted materials are less predictable. A Binder attempting
to use enchanted materials should exercise the utmost caution. Some
golems may turn out fine, possibly even with beneficial side-effects,
others may be actively inimical.

\item[Material Costs] This is the cost for magical materials to perform
the ritual. The cost of materials for the golem itself is additional
to this.

\item[Upgrades] When an Adept increases their rank in a golem ritual
they may upgrade existing golems to their new rank. This requires the
performance of the ritual. The time required is the base time for the
golem (no matter how big it is).

\item[Weathering / Deterioration] When a golem is crafted or animated,
the magic involved does not provide any sort of protection from normal
weathering (\ie Clay will dry and crumble, cloth will get torn, wood
will rot, iron will rust, etc). Separate preservation magics may be
cast on a golem to reduce or prevent normal weathering.

\end{Description}

\subsection{Attuning Golems}

When constructing a golem, it may be attuned to certain skills or
magics. The adept may attune the golem with 1 ability (+1/Rank). Each
skill, language, weapon, talent (racial or college), spell, or ritual
counts as 1 ability. What a golem has been attuned to cannot be
changed but it can be increased when a golem is upgraded.

\subsection{Golems in Combat}
\begin{Description}
\item[Bleeding] Golems don't bleed.

\item[Damaged Golems] Golems cannot be healed, instead they must be
repaired.  A golem must have positive Endurance to be activated.

\item[Dead Golems] A golem which has been killed cannot be re-activated
until the damage is repaired and the appropriate shaping ritual is
performed, as if doing an upgrade.

\item[Fatigue Recovery] Golems always have full fatigue on activation.
They also recover 1 FT per hour while activated.

\item[Sleep and Charm] Golems do not sleep and cannot be charmed.
However, if the Binder is possessing a golem, they can be slept or
charmed by targeting the golem.

\item[Stun] Golems do not stun.

\item[Unconsciousness] If a golem's Endurance is reduced below 1 then
it deactivates.

\item[Weapon Ranks] A golem may wield weapons.  If the golem has been
attuned to a weapon then it may use that weapon at Rank 0.  If that
weapon is also built-in to the golem (or unarmed) then it will wield
the weapon at Rank (Lower of: (Rank in ritual / 2) or maximum Rank).
All other weapons will be wielded as if unranked.  Note that a golem
will not evade or attempt any special manoeuvre (\eg trip, disarm)
unless possessed by an Adept.
\end{Description}

\subsection{Golem Intellience}

Golems have an animal level intelligence. Each golem has an animal
equivalent to use as a guideline for their intelligence and
temperament when their masters' instructions are unclear, incomplete
or no longer applicable.

Golems are somewhat different from their animal equivalents.  They
have a weak survival instinct, and their strongest drive is to follow
their masters' instructions.

Golem Animal Equivalents: Clay = Ox; Rag \& String = Monkey; Iron =
Tiger; Stone = Elephant; Wood = Dog.

\subsection{Instructing Golems}

Golems must be verbally instructed.  Instructing a golem requires
active concentration.  A golem will only take instructions from its
master.  A golem will attempt to follow the nature and intent of the
Adept's instructions.

\subsection{Talents}

\begin{talent}[T-1]{Activate Golem}

\begin{effects}
This talent requires 1 pulse of active concentration.  It costs 2 FT
to attempt this talent which activates an inactive golem.  If the
golem is one which the Adept has constructed then the talent
automatically works; otherwise the chance of success is the Adept's
chance of creating a golem of that type.  This talent may not be
ranked.  The adept who activates a golem is considered its master.
\end{effects}
\end{talent}

\begin{talent}[T-2]{Detect Enchantment}

\range{1. 30 feet (+ 5 / Rank)  2. Touch}
\multiple{50}
\basechance{1. PC + 3\% / Rank  2. PC + 5\% / Rank}
\begin{effects}
This talent determines whether an item, person, or area in
line-of-sight and within range is currently under an enchantment or
magical effect. The Adept can tell whether the enchantment is a
current spell or ritual, a warded, invested, or permanent effect, is
contained in a trap or is a curse. A double effect will reveal the
general nature of the spell (eg; defensive, summoning, damaging) and
the effective Rank or remaining duration of the enchantment. In
addition, a triple effect will tell the Adept the exact name of the
magic (eg; Wall of Bones, Hellfire), or the College of the spell.

Once the initial, most recent, magical effect has been successfully
detected, older enchantments on a target with multiple layers of magic
may also be detected. This may be continued while the Adept continues
to succeed in detecting Enchantments. However, only one attempt per
quarter may be made to detect any given enchantment. If a new
enchantment occurs, the Adept may attempt to detect it, though this in
no way affects the status of the old layer. If an old, unsuccessfully
detected enchantment expires, the Adept may attempt to detect any
newly revealed magic beneath it.

If the Adept is in contact with the target then the base chance of
this talent is higher.
\end{effects}
\end{talent}

\subsection{General Knowledge Spells}


\begin{spell}[G-1]{Adhesion}

\range{10 feet + 10 / Rank}
\duration{5 minutes + 5 / Rank}
\multiple{175}
\basechance{35\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Volume}
\begin{effects}
The surface of a non-living solid, up to 1 cubic foot (+ 2 / Rank), is
magically enchanted to adhere to any objects coming into contact with
it. Once stuck, an object is released when the spell's duration
expires, or the applied PS + D10 exceeds the spell's PS of 10 (+ 2 /
Rank), which tears the object free (the durability of some objects may
be less than the force required to tear them free from the
spell). Several individuals may combine their PS to free an
object. Being broken free of the area of the Adhesion, or resisting a
particular contact, in no way protects the object from becoming stuck
if brought into contact with the affected area again; nor is the spell
in any way broken by having an object torn away; the area remains as
adhesive as before. Except for the crowding of the area, there is no
limit to the number of objects that may be stuck with this spell. The
chance of a person coming into contact with an adhesive portion of a
hex is 10\% / Rank applied to that hex.
\end{effects}
\end{spell}

\begin{spell}[G-2]{Animating Objects}

\range{10 feet + 5 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{300}
\basechance{20\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Object}
\begin{effects}
This spell may be used to animate any one object, of up to 10 pounds
(+ 10 / Rank) in weight. By taking a pass action the Adept may control
the actions of 1 (+ 1 / Rank) previously controlled animates within
range. Once set in motion, the animates will attempt to carry out the
action until ordered otherwise. The animates will move about in a
manner applicable to their shape. Their TMR will not exceed 4.

Animates have a nominal PS value of 5 (+ 1 / Rank) though this will
have limited effect on objects made of flimsy materials.  Their strike
chance will be no more than 20\% (+2\% / Rank), with a maximum of D+2
damage. The animate will cease to function if the object is destroyed,
or a Binder General Knowledge Counterspell is cast on it. An animate
is an object, but may also be targeted as an Entity. If an animate is
created through the use of a Ward or Magical Trap, it will receive one
command determined when first cast.
\end{effects}
\end{spell}

\begin{spell}[G-3]{Bound Speech}

\range{1 foot + 1 / Rank}
\duration{1 day + 1 day / Rank}
\multiple{200}
\basechance{40\%}
\resist{Passive}
\storage{None}
\target{Object, Area of Object}
\begin{effects}
This spell allows the Adept to record a verbal message in an object,
and defines the conditions under which the message will be
replayed. This spell operates in most respects similarly to the Ward
ritual, except that the range is the range of this spell or touch, and
a message is stored instead of a spell. The message is replayed
exactly as the Adept recorded it, and may contain any verbalisations
that the Adept is capable of. The message may not exceed 5 words (+ 3
words / Rank). Unlike the Ward ritual, the message may be triggered
one additional time per two full Ranks.
\end{effects}
\end{spell}

\begin{spell}[G-4]{Durability}

\range{10 feet + 5 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{250}
\basechance{25\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Object}
\begin{effects}
One object weighing up to 2 pounds (+ 2 / Rank) may be made more
resilient and less susceptible to damage.  The item becomes almost as
strong as steel of the same thickness, without losing any flexibility.
The item cannot be broken unless exposed to stresses beyond that which
steel could withstand, given the object's size and shape.  The
strength is improved with rank to a maximum of slightly stronger than
steel.

This spell does not protect against soiling, corrosion or fire or any
forms of damage other than physical stress.

A weapon treated with this spell allows the wielder to add 2\% per
Rank of the spell to any roll to save the weapon from breaking.
Armour protected by this spell will have 2 extra Protection ( + 1 / 10
full ranks) to a maximum of the equivalent Steel armour.  At Rank 20,
1 may be added to this maximum.

Note that the added Protection replaces (rather than adds to) any
other Protection bonuses due to a material's strength (eg Armourer
bonuses).  If the item is broken, or armour suffers damage from a
Specific Grievous, the magic is dispelled.
\end{effects}
\end{spell}

\begin{spell}[G-5]{Mending}

\range{Touch}
\duration{Immediate}
\multiple{150}
\basechance{40\%}
\resist{None}
\storage{Investment}
\target{Parts of an Object, Golem}
\begin{effects}
Any single object weighing up to 10 pounds (+ 10 / Rank), or one
Golem, can be mended. A mended object becomes exactly as it was before
it was broken or deformed. Any pieces missing when the spell is cast
will remain missing when the object is mended. Mending used to fix
objects with an effective Artisan Rank greater than that the Adept
possesses will degrade the object's effective rank. Magical items made
mundane through breaking will remain mundane even after the use of a
Mending. The Spell of Mending may be used on a living creature that
has been transformed into stone, and subsequently broken. This
requires that the Binder be a Sculptor of at least Rank 8. A Spell of
Mending may be used to repair Golems. It may repair either a Specific
Grievous injury, or all general Endurance damage.
\end{effects}
\end{spell}

\begin{spell}[G-6]{Modify Aura}

\range{5 feet + 5 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{100}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Object, Entity}
\begin{effects}
This spell allows the Adept to able to modify the aura strength of any
one object or entity. The target object may be up to 20 cubic feet (+
20 / Rank).The strength of the aura increases or decreases by up to 1
+ (1 / 4 Ranks) on the table below. This spell does not alter the
target's aura in any other way.

\begin{tabular}{cl}
\textbf{Aura} & \textbf{Equivalent} \\
0    & No Aura \\
1    & Magic \\
2    & Formerly living as part of an object \\
3    & Formerly living \\
4    & Living plants \\
5    & Non-sentient Undead / Construct \\
6    & Living animals \\
7    & Sentient plant \\
8    & Sentient Undead / Construct \\
9    & Short-living Sentient \\
10   & Long-living Sentient \\
15   & Avatar \\
\end{tabular}
\end{effects}
\end{spell}

\begin{spell}[G-7]{Minor Creation}

\range{Touch}
\duration{15 minutes + 15 / Rank}
\multiple{250}
\basechance{40\%}
\resist{None}
\storage{Investment}
\target{Object}
\begin{effects}
The Adept may create a simple, common object from a larger source of
its constituent substances by reaching into the source and withdrawing
the object. For example, the Adept may reach into a tree and produce a
staff. One (+ 1 / 5 full Ranks) different substances may be combined
into a single finished object with this spell. The substance sources
are in no way damaged or reduced by the spell. The created object will
be a common example of its type and may not exceed 1 cubic foot (+ 1
per Rank) in volume, nor 1 lb (+ 1 lb / Rank) in weight. The object is
physically real with all normal attributes. The object may not be
created enclosed by or enclosing anything. Complicated or fine quality
objects may not be created without the appropriate artisan skill. No
alchemical, herbal or other quasi-magical objects may be
created. Created food provides no sustenance.  At the end of its
duration the item vanishes.  The object is a magical construct and
will have a magical aura, regardless of its constituent materials.
\end{effects}
\end{spell}

\begin{spell}[G-8]{Moulding Elements}

\range{30 feet + 10 / Rank}
\duration{5 minutes + 5 / Rank}
\multiple{250}
\basechance{20\%}
\resist{None}
\storage{None}
\target{Volume}
\begin{effects}
A volume of up to 2 cubic feet (+ 2 / Rank) containing one of the four
material Elements (air, water, earth, fire) can be moulded by the
Adept into any form desired, and will retain that form for the
duration of the spell. The Adept must immediately mould the element
into the correct shape with their hands. During this time the Adept is
protected from the effects of the element, which is as malleable as
putty. The time taken to mould the element is dependent on the size
and complexity of the object desired, but will be at least a
pulse. After the initial moulding, the element will retain its shape,
unless something disrupts it, when it will immediately attempt to
reform, possibly around or on top of the impedance. The shaped element
will have a defined boundary and shape, but will not be rigid
(excluding shaped earth). The spell cannot be cast over any
entity. The Adept may use applicable artisan skills to shape an object
of greater than Rank 0 quality. Once the spell expires the element
will act normally according to its substance, shape, and natural laws.
\end{effects}
\end{spell}

\begin{spell}[G-9]{Preservation}

\range{5 feet + 1 / Rank}
\duration{Special}
\multiple{100}
\basechance{40\%}
\resist{Passive}
\storage{Ward, Investment, Magical Trap}
\target{Volume, a dead / undead Entity}
\begin{effects}
This spell preserves and protects one animate, dead or undead entity,
of up to 100 pounds (+ 100 / Rank) or a collection of small objects of
a total weight not greater than 1 pound (+ 1 / Rank) against the
effects of time, decay, rust, erosion or wave action. It does not
confer any protection against magical attacks. It will not suspend
time with regard to resurrection, poisons, curses, etc. Duration is 4
(+ Rank squared) days, but is permanent at Rank 20.
\end{effects}
\end{spell}

\begin{spell}[G-10]{Transparency}

\range{10 feet + 5 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{175}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Volume}
\begin{effects}
This spell causes a volume of non-living solid material to become as
transparent as high quality glass, but to otherwise retain its
original characteristics. One (+ 1 / Rank) adjacent 1 foot cubes may
be affected. The Adept may cause the volume to be transparent from
only a single direction, by reducing the BC by 10.
\end{effects}
\end{spell}

\subsection{General Knowledge Rituals}


\begin{ritual}[Q-1]{Linking Lifeforce}

\duration{Permanent}
\multiple{300}
\basechance{40\% + 4\% / Rank}
\resist{Passive}
\target{Object, entity}
\casttime{1 hour}
\material{Object}
\concentration{Standard}
\begin{effects}
This ritual allows the Adept to bind an object to the life-force of an
entity. The object will reflect the entity's physical condition. If
the entity is well, the object will be in perfect order; if the entity
is sick or wounded, the object will appear appropriately damaged; and
if the entity dies, the object will seem ruined. There is no
limitation on the size or type of object, but it must remain the
entity's possession during the entire ritual. Traditionally, apples,
roses, statues, paintings or diamonds are used. The Life-force Link
may be broken by destroying the object, or by the object being beyond
100 miles (+ 100 / Rank) from the entity. The death of the entity will
not break the Link. If 10,000 (-500 / Rank) is spent on ingredients,
the range is unlimited and the entity and object may occupy different
planes without breaking the Link.
\end{effects}
\end{ritual}

\begin{ritual}[Q-2]{Petrifaction}

\duration{Permanent}
\multiple{200}
\basechance{55\% + 3\% / Rank}
\resist{Active \& Passive}
\target{Entity}
\casttime{1 hour}
\concentration{Standard}
\begin{effects}
This ritual allows the Adept to either change one entity to
marble-like stone, or turn one magically petrified entity back to
flesh. The entity must be present for the entirety of the Ritual. Any
possessions of an entity are (un)petrified with the entity. A
petrified entity is not aware of their surroundings, and has time
stopped for the purposes of poison, resurrection, curses and
ageing. Petrifaction is not fatal, although a dead entity may still be
petrified. If an entity is damaged after petrifaction, they may be
repaired with the Mending Spell if the Adept is a Rank 8 Sculptor. Any
damage will be applied when the entity becomes flesh. A petrified
entity weighs 3 times their normal weight.
\end{effects}
\end{ritual}

\begin{ritual}[Q-3]{Shaping Clay Golems}

\duration{3 hours + 3 / Rank}
\multiple{300}
\basechance{10\% + 4\% / Rank}
\resist{None}
\target{Object}
\basetime{2 hours}
\material{Clay \& Rare Earths}
\materialcost{100 sp}
\actions{Sculpting a statue}
\concentration{None}
\begin{effects}
Turns a quantity of earth into a golem.  Clay golems are susceptible
to fire and immersion; exposure to either will cause D10 FT damage to
the golem every minute. Fire and water based attacks do double damage.
A clay golem will gradually dry out and become immobile if not kept
moist.  In a temperate climate a golem will lose 1 AG per day if no
action is taken to prevent this.  A golem which has completely dried
out cannot be reactivated.  During construction a clay golem may be
designed to resemble a humanoid; this requires that the Adept or an
assistant has the sculptor skill and is familiar with the humanoid
race.  The chance of the golem passing visual inspection is 40\%
(+4/Rank sculptor, +2 / Rank spy).  Note that the golem will not pass
close inspection --- cold, hard skin, lifeless face.  Base

Base Materials: Clay or any soft earth. However clay tends to be the
material of preference as golems made of more crumbly earth fall apart
easily.
\end{effects}
\end{ritual}

\begin{ritual}[Q-4]{Shaping Rag \& String Golems}

\duration{2 hours + 2 / Rank}
\multiple{250}
\basechance{40\% + 3\% / Rank}
\resist{None}
\target{Object}
\basetime{3 hours}
\material{Cloth, straw \& herbs}
\materialcost{50 sp}
\actions{Building a golem}
\concentration{None}
\begin{effects}
Turns a small mannequin into a golem.  Rag \& string golems are
susceptible to fire; exposure to fire will cause [D - 4] FT damage to
the golem every second pulse. Fire based attacks do double damage.

Base Materials: Cloth, string, rope, straw. Flexible, easily
manipulable materials.
\end{effects}
\end{ritual}

\subsection{Special Knowledge Spells}


\begin{spell}[S-1]{Bubble of Force}

\range{5 feet + 1 / Rank}
\duration{20 minutes + 20 / Rank}
\multiple{450}
\basechance{15\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
The spell causes an iridescent bubble of force to appear. The
substance of the bubble prevents the occupant(s) from interacting with
the world outside the bubble other than by sight and sound. The bubble
is transparent to low levels of light and sound, but it prevents
damaging levels of either passing through it. The bubble is impervious
from both sides. It also absorbs physical damage, including that
caused by falling. The bubble is impenetrable to magic - it may not be
targeted through, and area effects will not pass through the
bubble. Inside, it provides a stable, warm and dry environment, with
fresh air to breathe. The surface of the bubble is hard to the
touch. The bubble may be rolled on a solid surface by an entity within
it, by entities outside, or by a very strong wind or similar physical
force. It will float on water or mud, lava and similar semi-solid
substances, and may be carried by wind or current. At Ranks 0-10, the
bubble is 5' in diameter (one hex), at Ranks 11-15, it is 10' across
(three hexes), and at Ranks 16-20, it is 15' (seven hexes). All the
bubble must appear within range. The bubble may enclose any entities
and objects completely within the target hexes, if they fail to
resist. Successful resistance will cause the entity to be gently
pushed aside by the bubble. If the spell is cast in a gas, the gas is
displaced outside the bubble. If cast in a liquid, the liquid is
displaced, and the bubble will bob to the surface. The spell may not
be cast inside a solid substance, or where there is insufficient room
for the bubble to form without having to displace an immovable
solid. Thus the whole sphere will always appear above ground. The
bubble may be dissipated by anyone casting a Binder Special Knowledge
Counterspell of at least equal Rank.
\end{effects}
\end{spell}

\begin{spell}[S-2]{Disintegration}

\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{300}
\basechance{20\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Object}
\begin{effects}
The Adept may disintegrate up to 2 cubic feet (+ 2 / Rank) of a
non-magical object. If the object has any magical attributes, the
spell will fail.
\end{effects}
\end{spell}

\begin{spell}[S-3]{Frictionless Floor}

\range{40 feet + 15 / Rank}
\duration{30 seconds + 10 / Rank}
\multiple{300}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
Up to 20 (+ 10 / Rank) square feet, or 1 (+ 1 / 2 Ranks) hexes, of
contiguous surface becomes nearly frictionless, somewhat like wet
ice. Footing becomes treacherous, and handholds non-existent. Any
entity attempting to move on a frictionless surface must make a 1 \x
AG check each pulse, or fall prone. If they do fall, a successful AG
check is required to regain their feet; and while prone, they will
continue in the direction of their last movement until they clear the
frictionless surface. The effect is not normally visible.
\end{effects}
\end{spell}

\begin{spell}[S-4]{Instant Golem}

\range{10 feet + 5 / Rank}
\duration{5 minutes + 5 / Rank}
\multiple{200}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward}
\target{Object(s)}
\begin{effects}
This spell is cast to produce a specific type of golem with which the
adept is familiar (must be greater than Rank 0 in the appropriate
ritual) from materials within their range.

If a sufficient quantity of materials is not within the Adept's range
the spell will automatically fail.

When cast the materials within range will animate and draw together at
the point where the majority of the materials are coming from.  A
golem will form in 20 seconds (-1 / Rank).  In the pulse after it has
finished forming the golem will be ready to carry out its master's
commands.

The effective rank of the golem is the rank which the adept has in the
appropriate ritual.

The duration of the golem is the duration of this spell.  At the end
of the duration the golem will collapse and its component materials
will re-animate and attempt to return to their previous location and
state.

Note for Investment: When Invested, the effective rank of the golem is
the rank that the investing adept had in the appropriate ritual at the
time of investment.  The type of golem is set at the time of
investment.  The created golem will consider the triggerer to be its
master.

Note for Ward: The type and effective rank of the golem are set at the
time of warding.  The Adept may also instill a basic instinct into the
golem at the time of warding (\eg kill, serve, assist, guard, dig,
eat, etc.); the actions of the golem will be governed by this
instinct.  The instinct is always simple and indiscriminate (\eg kill:
the golem will attempt to kill anyone or anything that it perceives).
The golem's master is the warding adept.

\end{effects}
\end{spell}

\begin{spell}[S-5]{Instant Petrifaction}

\range{10 feet + 5 / Rank}
\duration{Special}
\multiple{500}
\basechance{1\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
This spell immediately starts to petrify the target. For the next 5
seconds, all the target's percentage chances and D10 rolls are
halved. At the end of this time, they must make their resistance
check. If they succeed, they may resume normal activities, otherwise
the petrifaction runs its course, leaving the target as marble-like
stone.

As with Ritual Petrifaction, any possessions of the target are
petrified with the target. A petrified entity is not aware of their
surroundings, and has time stopped for the purposes of poison,
resurrection, curses and ageing. Petrifaction is not fatal, although a
dead entity may still be petrified.

\end{effects}
\end{spell}

\begin{spell}[S-6]{Instilling Flight}

\range{Touch}
\duration{Concentration: Maximum 30 minutes + 30 / Rank}
\multiple{350}
\basechance{20\%}
\resist{None}
\storage{None}
\target{Object}
\begin{effects}
This spell enables the Adept to instil a possession of up to 5 lbs (+5
/ Rank) with the power of flight. The spell will dissipate if the
object stops being a possession of the Adept, the Adept loses
concentration, or if the object is broken. The Adept may cause the
object to fly at 20 miles per hour (+ 2 / Rank). It will take off and
accelerate up to full speed, or halt and land, in a single pulse. The
object may support 150 lbs (+ 50 / Rank) in addition to its own
weight. Naturally flexible or fragile items gain sufficient strength
and rigidity to support the load. Any object or entity that falls from
the flying object will move off in a random direction. If the object
is about to crash into a surface, it will attempt to land, although
some surfaces may be inappropriate for this (lava, sheer walls, etc.).
\end{effects}
\end{spell}

\begin{spell}[S-7]{Itemisation}

\range{Touch}
\duration{1 day + 1 / Rank}
\multiple{300}
\basechance{25\%}
\resist{None}
\storage{None}
\target{Object}
\begin{effects}
One object is transformed into a representative figurine of itself.
This spell will fail if the target is not free-standing or if the
target does not fit within a 5 foot (+ 1 / Rank) cube.  The figurine
will look like the original item but have the structural strength of
soft wood and will have no moving parts.  The figurine will be 1/12th
of the size (min 1 inch in its longest dimension) and 1/10th the
weight of the original.  Any items contained within the target will
not be affected, and may destroy the item as it shrinks.  The figurine
will revert to normal when the spell expires or the figurine is
broken.

\end{effects}
\end{spell}

\begin{spell}[S-8]{Making}

\range{Self}
\duration{Concentration: maximum 1 hour + 1 / Rank}
\multiple{200}
\basechance{20\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
This spell enables the Adept to use their hands as if they were common
tools. The hands do not change appearance. The Adept may freely change
from the mimicry of one tool to another without re-casting the
spell. This spell does not affect the damage done by Unarmed
Combat. This spell does not enhance the Adept's craftsmanship, nor
protect them while they work (except for the protective properties of
the tools they are emulating).  The tools that may be mimicked must be
reasonably simple, and include: adze, auger, chisel, crowbar, hammer,
hatchet, level, pickaxe, plane, pliers, plumb-bob, rock drill, saw,
shovel, sickle, spokeshave, square, tongs and mallet.

\end{effects}
\end{spell}

\begin{spell}[S-9]{Matter Transmutation}

\range{2 feet + 2 / Rank}
\duration{30 seconds + 30 / Rank}
\multiple{400}
\basechance{20\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Volume}
\begin{effects}
The Adept may transmute a volume of 2 cubic feet (+ 2 / Rank) of
non-living matter into any other matter. The matter must retain its
elemental state --- solid to solid, liquid to liquid, gas to gas. The
transmuted matter will retain its original size and shape, but assumes
all other physical characteristics of the resulting matter. The Adept
must have some familiarity with the resulting matter.

\end{effects}
\end{spell}

\begin{spell}[S-10]{Possess Golem}

\range{10 feet}
\duration{10 minutes + 10 / Rank}
\multiple{250}
\basechance{35\%}
\resist{Passive}
\storage{Potion}
\target{Self and Controlled Golem}
\begin{effects}
The Adept's body goes into a coma and the Adept's mind goes into the
body of the target golem.  The Willpower, Magical Aptitude, Perception
and Magic Resistance of the golem are replaced by those of the Adept.
The Adept may only use ranks in abilities which they have, and the
golem has been attuned to.

During the possession the Adept uses the golem's senses.  The Adept
may also speak through the golem's mouth.

The Adept may cast while possessing any non-iron golem, however: The
FT cost of the magic is applied to the Adept's body and the host
golem; The Adept cannot cast if the host golem has no FT; Backfires
affect the Adept's body and the host golem (if applicable).

If the host golem takes Endurance damage then the adept's body takes
half of that amount as FT damage.  If the host golem is knocked
unconscious or killed then the adept must make a Willpower check or
fall unconscious for D10 minutes; the multiple is \x 2 for
unconsciousness or \x 1 for death.
\end{effects}
\end{spell}

\begin{spell}[S-11]{Unfastening}

\range{10 feet + 10 / Rank}
\duration{Immediate}
\multiple{250}
\basechance{30\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity or Object}
\begin{effects}
This spell unfastens, opens or unties all closures, buckles, straps,
ties, knots, locks and other fastenings on the target (causing most
armour, clothes, and packs to fall off).  This will not undo stitching
or weaving, and it will not cause any damage to the target.  If
targeted on an entity then all of their possessions will be affected
by the spell.
\end{effects}
\end{spell}

\begin{spell}[S-12]{Wall of Dust and Sand}

\range{20 feet + 10 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{150}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This spell summons and binds together particles of dust and sand to
create a wall with the strength and solidity of sandstone. The wall
may be 15' tall 20' long 1' thick; or a ring 10' high with a 10'
diameter; or a pillar 15' high with a 4' diameter. The Adept may
increase any dimension by 1' / Rank. The Adept determines the position
and orientation of the Wall, but at least one edge must be affixed to
a solid surface. The fixed edge of the wall will bind fast to the
adjacent surface, and can support the rest of the wall. Any entity
that is in the area of the wall when it is cast will be ejected to the
closest point outside the wall - if this is impossible, the wall will
not appear. The wall may be destroyed by inflicting 100 points of
damage, or a 5' \x 5' \x 5' hole may be caused by inflicting 50 points
of damage.
\end{effects}
\end{spell}

\begin{spell}[S-13]{Wall Walking}

\range{Self}
\duration{10 seconds + 10 / Rank}
\multiple{450}
\basechance{10\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
This spell allows the Adept to pass through solid matter by becoming
insubstantial. However, they may not pass through cold iron. The Adept
has a TMR of 1 while within solid matter, and may move in any
direction, including up and down. The Adept gains no ability to see
through solid matter, but may breathe in it. While under the effects
of this spell, the Adept is virtually immune to physical damage except
that inflicted by cold iron, but conversely cannot harm anyone in
melee unless they use cold iron in return. If the Adept is caught in a
solid object when the spell expires, they lose the ability to breathe
in solid matter, and are trapped.
\end{effects}
\end{spell}

\subsection{Special Knowledge Rituals}


\begin{ritual}[R-1]{Binding Permanency}
\duration{Permanent}
\multiple{400}
\basechance{20\% + 4\% / Rank}
\resist{None}
\target{Spell}
\casttime{1 hour}
\material{Special}
\concentration{Standard}
\begin{effects}
This ritual enhances the duration of one of the Adept's spells or
rituals belonging to the College of Binding and Animating Magics.  The
magic must have been cast previously, and have sufficient duration to
last through the performance of this ritual. This ritual may not be
used with any magics with concentration-based or indefinite,
condition-based duration. A spell that has been enhanced by this
ritual may not be counter-spelled, but may still be dissipated.
Material costs for this ritual are 10,000 sp (- 400 / Rank) for
General Knowledge Spells, 20,000 sp (- 800 / Rank) for Special
Knowledge Spells, and 30,000 sp (- 1200 / Rank) for Rituals.
\end{effects}
\end{ritual}

\begin{ritual}[R-2]{Investment}

\begin{effects}
Except as noted below, this ritual is identical to the ritual of the
same name in the College Magic - Investment section
(\S\ref{investment}). Adepts of this College learn this Ritual of
Investment instead of the standard version. The Adept may invest
either a spell of their own or that of a willing Mage who participates
in the full ritual. If the spell being invested is contributed by
another Mage, the maximum number of charges stored is half the Rank in
the ritual. The maximum investable rank of the spell is Spell Rank - [
(20 -Ritual Rank) / 2 ]. A spell whose maximum investable rank is
reduced below Rank 0 may not be invested. The Adept may not have any
assistance in creating scrolls.
\end{effects}
\end{ritual}

\begin{ritual}[R-3]{Item Divination}

\duration{Immediate}
\multiple{150}
\basechance{2 \x MA + 3\% / Rank}
\resist{None}
\target{Object}
\casttime{1 hour}
\material{500sp (-25 / Rank)}
\concentration{Standard}
\begin{effects}
Through this ritual, the Adept discovers the exact nature of all
enchantments, mechanisms, curses, side-effects, etc.\ placed on an
item.  If an item has been imbued with an Individual True Name, the
name will not be revealed, but its existence will be.  This ritual
cannot backfire.
\end{effects}
\end{ritual}

\begin{ritual}[R-4]{Major Creation}

\duration{1 day + 1 / Rank}
\multiple{250}
\basechance{40 + 3\% / Rank}
\resist{None}
\target{Object}
\casttime{1 hour}
\material{None}
\concentration{Standard}
\begin{effects}
The Adept may create an object from a larger source of its constituent
substances by drawing forth the materials from the source.  All
constituent materials must be within 5 feet (+ 5 / Rank).  Any number
of different substances may be combined into a single finished object
with this ritual. The substance sources are in no way damaged or
reduced by the ritual. The created object may be any object which the
adept can see or has studied and may not exceed 10 cubic feet (+ 10
per Rank) in volume, nor 50 lbs (+ 50 / Rank) in weight. The object is
physically real with all normal attributes. The object may not be
created enclosed by or enclosing anything. Complicated or fine quality
objects may not be created without the appropriate artisan skill. No
alchemical, herbal or other quasi-magical objects may be
created. Created food provides no sustenance.  At the end of its
duration the item vanishes.  The object is a magical construct and
will have a magical aura, regardless of its constituent materials.
\end{effects}
\end{ritual}

\begin{ritual}[R-5]{Shaping Iron Golems}

\duration{6 hours + 6 / Rank}
\multiple{350}
\basechance{20\% + 4\% / Rank}
\resist{None}
\target{Object}
\basetime{4 days}
\material{Forgeable metal \& Rare Earths}
\materialcost{1,000 sp}
\actions{Forging a golem}
\concentration{None}
\begin{effects}
Turns a quantity of metal into a golem.  When activated an iron golem
radiates a faint glow, especially its eyes, and is warm to the touch.

Base Materials: All metals which are normally solid at room
temperature.
\end{effects}
\end{ritual}

\begin{ritual}[R-6]{Shaping Stone Golems}

\duration{5 hours + 5 / Rank}
\multiple{350}
\basechance{20\% + 4\% / Rank}
\resist{None}
\target{Object}
\basetime{4 days}
\material{Stone \& Rare Earths}
\materialcost{500 sp}
\actions{Carving a golem}
\concentration{None}
\begin{effects}
Turns a quantity of stone into a golem.  Stone golems are the heaviest
of all golems.  Stone golems cannot be stealthy, they grind as they
move.  A Stone Golem can usually be heard coming from 100 yards away.

Base Materials: All types of stone or rock which are normally solid.
\end{effects}
\end{ritual}

\begin{ritual}[R-7]{Shaping Wood Golems}

\duration{4 hours + 4 / Rank}
\multiple{250}
\basechance{40\% + 3\% / Rank}
\resist{None}
\target{Object}
\basetime{2 days}
\materialcost{100 sp}
\actions{Carving a Golem}
\concentration{None}
\begin{effects}
Turns a quantity of wood into a golem.  Wood golems are susceptible to
fire; exposure to fire will cause D10 FT Damage to the Golem every 30
seconds.  Fire based attacks do double damage.  During construction a
wood golem may be designed to resemble a tree, this requires that the
Adept or an assistant has the Herbalist skill.  The chance of the
golem passing casual inspection is 40\% (+ 5 / Rank Herbalist).  Note
that the disguise will only work if the golem is not moving.

Base Materials: All wood types, including wicker, cane and bamboo, but
excluding fossilised woods and grasses.  GMs may rule that unusual
wood types have different properties (\eg An iron wood golem might
have +1 NA, +1 PS, + 2 EN, -2 TMR, -4 AG).
\end{effects}
\end{ritual}

\begin{table*}[h]
\subsection{Golem Statistics}

\begin{tabular}{llllll}
\textbf{Golem}  & \textbf{MA} & \textbf{WP} & \textbf{PC}  & \textbf{PB} & \textbf{NA} \\
Clay   &  0 &  8 &  8  &  5 + 2 / Rk (Sculptor) &  2 \\
Iron   &  0 & 18 & 15  &  8 + Rk (Blacksmith)	&  8 \\
R\&S   &  0 & 10 & 12  & 10 + Rk (Tailor)	&  0 \\
Stone  &  0 & 20 & 10  &  6 +3 / 2 Rk (Sculptor)&  6 \\
Wood   &  0 & 12 & 15  & 10 +3 / 2 Rk (Carving)	&  4 \\
\end{tabular}

\bigskip

\begin{tabular}{llllllllll}
\textbf{Golem Type} & \textbf{Size} & \textbf{PS} & \textbf{MD} & \textbf{AG} & \textbf{EN} & \textbf{FT} & \textbf{TMR} & \textbf{Mod} \\ \hline
\multicolumn{8}{l}{Clay} \\ \hline
 & 3' & 8 & 17 & 14 & 8 & 12 & -1 \\
 & 4' & 10 & 16 & 13 & 10 & 15 & -1 \\
 & 5' & 12 & 15 & 12 & 12 & 18 \\
 & 6' & 14 & 14 & 11 & 14 & 21 \\
 & 7' & 16 & 13 & 10 & 16 & 24 \\
 & 8' & 18 & 12 & 9 & 18 & 27 & 1 \\
 & 9' & 20 & 11 & 8 & 20 & 30 & 1 \\
 & Rk & 1 & 1 & 1 \\ \hline
Iron \\ \hline
 & 4' & 10 & 19 & 15 & 20 & 20 & -1 \\
 & 5' & 12 & 18 & 14 & 22 & 22 \\
 & 6' & 14 & 17 & 13 & 24 & 24 \\
 & 7' & 16 & 16 & 12 & 26 & 26 \\
 & 8' & 18 & 15 & 11 & 28 & 28 & 1 \\
 & 9' & 20 & 14 & 10 & 30 & 30 & 1 \\
 & 10' & 22 & 13 & 9 & 32 & 32 & 1 \\
 & Rk & 1 & 0.5 & 0.5 & 1 & 0.5 \\
\hline
Rag \& String \\ \hline
 & 6" & 3 & 22 & 20 & 1 & 3 & \x 0.25 \\
 & 1' & 4 & 21 & 19 & 2 & 4 & \x 0.33 \\
 & 1'6" & 5 & 20 & 18 & 3 & 5 & \x 0.50 \\
 & 2' & 6 & 19 & 17 & 4 & 6 & \x 0.66 \\
 & 2'6" & 7 & 18 & 16 & 5 & 7 & \x 0.75 \\
 & Rk & 0.5 & 0.5 & 0.5 & 0.5 \\
 \hline
Stone \\ \hline
 & 5' & 15 & 14 & 12 & 14 & 12 \\
 & 6' & 18 & 13 & 11 & 16 & 15 \\
 & 7' & 21 & 12 & 10 & 18 & 18 \\
 & 8' & 24 & 11 & 9 & 20 & 21 & 1 \\
 & 9' & 27 & 10 & 8 & 22 & 24 & 1 \\
 & 10' & 30 & 9 & 7 & 24 & 27 & 1 \\
 & 11' & 33 & 8 & 6 & 26 & 30 & 2 \\
 & 12' & 36 & 7 & 5 & 28 & 30 & 2 \\
 & Rk & 1.5 & 1 & 1 \\
 \hline
Wood \\  \hline
 & 2' & 5 & 16 & 14 & 8 & 12 & -1 \\
 & 3' & 6 & 15 & 13 & 10 & 15 & -1 \\
 & 4' & 8 & 14 & 13 & 10 & 15 & -1 \\
 & 5' & 10 & 13 & 12 & 12 & 18 \\
 & 6' & 12 & 12 & 11 & 14 & 21 \\
 & 7' & 14 & 11 & 10 & 16 & 24 \\
 & 8' & 16 & 10 & 9 & 18 & 27 & 1 \\
 & 9' & 18 & 9 & 8 & 20 & 30 & 1 \\
 & Rk & 0.5 & 0.5 & 0.5 & 0.5 & 0.5 \\
\end{tabular}

*Rk - Bonus to statistic per rank in the ritual. Always truncate fractions.
\end{table*}
\end{college}
