\chapter{Magic}

\begin{multicols*}{3}

% \section{General Magic}
\label{magic}
\index{magic|(}

\subsection{Introduction to Magic}
\label{magic:intro}

Magic represents the effects of the unknown forces that shape and
control the worlds. Those who have talent and knowledge can tap these
energies (known as mana) and shape them to their own ends. These
people are known as Mages. They are usually either revered or reviled
by the normal population.

There are three types of Magic: Talent Magic operates more or less
immediately, while Spell and Ritual Magic require preparation before
taking effect. Spells may be prepared in seconds or minutes, but
Rituals take hours (and sometimes many weeks) to perform.

There are a number of separate Colleges of Magic. Each represents a
specific type of magic, and each has a list of Spells, Rituals and
Talents available only to Adepts of that College. Most of the magic
detailed within these rules is Collegiate magic.

\subsubsection{Definitions:}

\begin{Description}

\item[Active Resistance] A special type of Magic Resistance, where the
entity can choose to concentrate their attention on resisting a
magical effect, and thus reduce its Cast Chance by their Magic
Resistance. Only some magic is actively resistible.
\index{resistance!active}

\item[Adept] A member of a College of Magic is known as an Adept.

\item[Backfire] If a spell or ritual is particularly incompetently
cast, unpredictable and often dangerous effects can occur. This is
colloquially known as a backfire.
\index{backfires!definition}

\item[Branches of Magic] There are 3 branches of Magic: the
Thaumaturgies, the theoretical branch of Magic including the Bardic, E
\& E, Mind, Naming, Illusion and Binding Colleges; the Elementals, the
naturalistic branch of Magic that includes the Earth, Water, Fire,
Air, Ice and Celestial Colleges; and the Entities, the old, ``dark''
branch of Magic that includes the Necromantic, Rune, Summoning and
Witchcraft Colleges. The Thaumaturgies and Entities are opposed to
each other.  \index{magic!branches}

\item[Cast Chance] The modified Base Chance of effectively casting a
spell or performing a ritual.

\item[Cast Check] The game mechanic whereby a Mage's player determines
the result of an attempted spell or ritual.

\item[Cold Iron] All solid metals that are primarily composed of iron
ore are termed Cold Iron. This includes both Iron and Steel. Such
metals in a liquid state are not ``cold''. Cold Iron inhibits the
ability of Mages to use mana.
\index{iron!cold}

\item[College] Most magic is divided up into numerous Colleges, each
of which specialise in a type of magic (\eg Fire, Necromancy). A Mage
who has joined a particular College is known as an Adept of that
College, and may not belong to another College without first forsaking
all knowledge of their previous College.
\index{colleges!definition}

\item[Concentration] If a spell has a concentration component in its
duration, then the Adept must concentrate in order to maintain the
spell. A Mage may only have one concentration spell in effect at any
time.
\index{concentration!definition}

\item[Consecrated Ground] Any ground that has been consecrated to the
``Powers of Light'' affects the Magic Resistance of all within
it. There is no College specifically dedicated to the Powers of Light,
because they are, in effect, opposed to the use of magic. Most temples
and monasteries and some graveyards will be consecrated
ground. Barrows, pagan temples and the dwellings of magical beings can
never be consecrated ground. Undead and Necromancers suffer special
penalties on consecrated ground.

\item[Counterspell] A type of spell which helps to protect
individuals and areas against the effects of a particular College of
Magic.

\item[Fatigue Cost] The amount of energy, in the form of Fatigue, that
a Mage must expend in order to cast a spell.

\item[General Knowledge] All Colleges of Magic have a body of Spells,
Talents and Rituals which are classified as General Knowledge. These
magics are taught to all Adepts of the College during their initial
training.

\item[High Mana] An area that is rich in mana is referred to as a high
mana area. Such areas are rare, and include locations where human
sacrifice is practised or where the inter-planar boundaries are weak,
and mana leaks through. Often mountain tops or clearings in jungles
will contain such areas. They are likely to be well guarded by beasts
and individuals attracted by the mana, including a larger than usual
proportion of fantastical beasts. Magic is easier to perform in these
areas.
\index{mana!high}

\item[Low Mana] An area with depleted mana is known as a low mana
area. Most densely populated or civilised parts of the world are Low
Mana, as are some battle-scarred areas. Magic is harder to cast in low
mana areas.
\index{mana!low}

\item[Mage] Any sentient being who can manipulate mana to produce
(often fantastic) results (excluding racial Talents). A Mage must have
a Magical Aptitude characteristic.

\item[Magic Resistance] All targets with a Willpower value have the
capacity to resist some magics directed against them. This ability is
their Magic Resistance, and is a function of their Willpower. Not all
magic is resistible.

\item[Magical Animates] Anything that has been animated, except
undead, is a magical animate. Some Magical animates gain a magic
resistance.  Only those animates that have a Magical Aptitude or
Willpower gain a Magic Resistance.  Those animates that have neither
MA nor WP have no resistance to magic, and in addition, may be
affected by spells that affect Entities and those that affect
objects.

\item[Mana] The type of energy that is used in all magic. A Mage must
draw upon mana to perform any magic. If there is no mana present, a
Mage cannot perform any magic.
\index{mana}

\item[Object] An item wholly composed of never living or formerly
living matter, or some combination thereof.  Objects do not have a
Magic Resistance except when they are Possessions or Magical Animates.

\item[Passive Resistance] This is the default Magic Resistance made by
all targets with willpower and operates automatically against all
spells that may be passively resisted. It is possible to stop
passively resisting temporarily.
\index{resistance!passive}

\item[Place of Power] Certain places aid the practise of magic. The
most well known places are Earth places of power, but they exist for
all the Elemental and Entity Colleges (excluding Rune). Such places
are rare, and often co-exist with High Mana areas.

\item[Possessions] Possessions are those objects held, carried or
otherwise within the personal area of an Entity.  They are affected by
those spells that affect the Entity, and are entitled to the Entity's
Magic Resistance.

\item[Resistance Check] The game mechanic which determines whether a
resisting entity is fully affected by a magical effect.

\item[Ritual Magic] Complex procedures and techniques that require the
Mage to spend large amounts of preparation time (and often
ingredients) to complete successfully.

\item[Special Knowledge] All Colleges of Magic have a body of complex
or specialised spells and rituals which are not taught to mere
apprentices, but which are gained with time and effort after the
Adepts prove themselves worthy. These magics are termed Special
Knowledge.

\item[Spell Magic] Codified magical formulae that take anywhere from a
few seconds to a minute to perform, require energy from the Mage, and
which result in specific alterations to Natural Law.

\item[Talent Magic] Magical abilities that require mana, but no energy
and minimal time from the user. Many species have racial Talents.

\end{Description}

\subsection{How Magic Works}
\label{magic:howmagicworks}

There are three types of Magic: Talent, Spell and Ritual Magic.

Talent Magic is broken into Racial and Collegiate Talents.  Talents
are common to all members of a Race or College of which they are a
characteristic part and may never be learned or forgotten, though they
often may be ``ranked''. Talents require no preparation, take a
maximum of 5 seconds to utilise, and require no expenditure of
energy. All Talents can be classified as either active or
passive. Passive Talents are always in effect. Active Talents require
a Pass Action to utilise, and often require rolls to see if they
succeed. Racial Talents are described in Character
Generation. Collegiate Talents are discussed in the individual
Colleges.

Spell Magic constitutes the great majority of the magic utilised by
Mages. Unless otherwise stated, all magic mentioned in these rules is
Spell Magic. All Spell Magic has the following characteristics in
common:

Each individual Spell has a defined range, duration, base chance and
effect. Spells must usually be prepared by the Mage through a process
of incantation to draw mana to activate the Spell. Spells are unstable
in their workings, and if cast ineptly, may fail entirely or have
unexpected effects on the vicinity. The casting of a Spell drains
energy from the caster in the form of tiredness Fatigue.

Ritual Magic requires the expenditure of large blocks of time (usually
hours) and usually certain conditions must be fulfilled while
performing the Ritual. Ritual Magic occasionally requires a large
number of special tools and substances and may be restricted to
particular times or places. Magical effects from Ritual Magic tend to
be more powerful, prolonged or delayed than those of Spells. Most
rituals require a Cast Check to determine whether the ritual was
successful. If not otherwise stated in the specific ritual
description, a ritual may backfire (roll greater than Base Change +
30) with similar consequences to a spell.  Rituals may also multiple
effect similar to spells.

\subsubsection{Extended Rituals}
\label{magic:extendedrituals}
\index{rituals!extended rituals}

Some rituals require a far greater time to perform than the standard
one hour, possibly requiring weeks or even months. During these
rituals the Adept is not involved in constant concentration.  The
Adept may eat, sleep (8 hours a day) and perform other activities
requiring less than 2 hours a day while engaged in a lengthy ritual.
During the extended ritual the Adept can utilise only stored magic, and
that inherent in the ritual.

\subsection{How to Cast Spells}
\label{magic:casting}
\index{spells!casting}

Casting a Spell is a complex process.

\begin{Description}

\item[Preparing] Firstly, the spell must be prepared and mana gathered
for the spell. This does not require any fatigue, and normally carries
no risk. However, it does involve gesticulations and
conversation-level speech, which will be obvious to observers. The
spell may be prepared in 5 seconds, 1 minute, or multiples of an hour
(using Ritual Spell preparation). The length of time taken to prepare
the spell is proportional to the resulting safety of the Mage. The
length of time spent preparing a spell must be decided upon in
advance. Preparing a spell is subject to the restrictions mentioned in
\S\ref{magic:restrictions}.
\index{magic!preparing a spell}

\item[Casting] Once prepared, the spell is Cast by an expenditure of
Energy in the form of tiredness Fatigue, used to shape and direct the
magic. This takes 5 seconds. Once cast, a spell will either impact
upon its target or fail. If the spell impacts on its target, it may be
partially or wholly resisted. If the spell fails, it may backfire (see
\S\ref{magic:backfire}). If the spell is cast particularly competently,
it may be especially effective.
\index{magic!casting a spell}

\end{Description}

\subsubsection{Casting Mechanics}

\begin{Description}

\item[Preparation] The Mage's player announces the spell and length of
preparation (either 5 seconds, 1 minute, or a number of hours). They
may break off their preparation at any time and abort the casting
process. A spell must be used immediately upon being prepared or it is
dissipated and the preparation must be restarted before it can be
Cast. Only one spell can be prepared at any one time. At the end of
the preparation, the Mage is aware of the state of the surrounding
mana. During combat, Preparing is a Pass Action.

\item[Casting] The Mage's player announces the spell, its target, and
any additional options desired (such as lowering the Rank of some
attributes). During combat, Casting is a Fire action.

\item[Cast Check] The player then modifies the Base Chance of the
spell due to current conditions to produce a Cast Chance as a
percentage. This Cast Chance is then compared with a D100 roll. If the
die roll is less than or equal to the Cast Chance, the spell works. If
the die roll is less than 5\% of the Cast Chance, the spell succeeds
with a ``triple effect''. If the die roll is between 6\% and 15\% of
the Cast Chance, the spell succeeds with a ``double effect''. If the
die roll is more than 30 higher than the Cast Chance (with 5 second
preparation), or 40 higher than the Cast Chance (with longer
preparation), the spell has not only failed, but Backfired, as per \S
\ref{magic:backfire}.

\item[Fatigue] A Mage may not cast a spell unless they have sufficient
Fatigue to pay the expected Fatigue cost. At the end of the
preparation, the state of the Mana (none, Low, normal or High) is
known, and the Mage may abandon the attempt at that point, before
losing the Fatigue. Whether the spell succeeds, fails or backfires,
the Mage must pay the Fatigue cost. It usually costs 1 Fatigue point
to cast a General Knowledge Spell or 2 Fatigue points to cast a
Special Knowledge Spell. In a High Mana area, these Fatigue costs are
reduced by one. In a Low Mana area, the Fatigue Cost is doubled.

\item[Success]
If the Cast Check is a double or triple effect, the element to be
doubled or tripled must be decided before anything else is
resolved. If the Cast Check is a success or better, the target(s) may
resist the spell, if it is passively resistible (see
\S\ref{magic:resistance} for details). This will reduce or nullify the
effects of the spell, as defined in the individual spell
description. See \S\ref{magic:effects} to resolve the spell
effects. Note that some backfires will result in partial success.

\item[Failure]
Nothing occurs (except Fatigue loss). If the result is a backfire,
consult the Backfire Table \S\ref{tables:backfire}.

\end{Description}

\subsection{Cast Check Modifiers}
\label{magic:modifiers}
\index{spells!cast changes modifiers}

In addition to the individual College modifiers, all Mages receive the
following modifiers whenever engaging in Spell casting:

{\small \begin{tabularx}{\linewidth}{Xr}
Each point the Mage's MA is greater than 15 & +1 \\
Each point the Mage's MA is less than 15 & -1 \\
Each Rank the Mage has with the spell they are casting & +3 \\
Each hour the Mage engages in Ritual Spell preparation & +3 \\
\end{tabularx}}

\subsection{Spell Effects}
\label{magic:effects}

Spells which are successfully cast on valid targets immediately take
effect on those targets (unless explicitly stated within the spell
description). If the spell is Passively Resistible, as stated in the
individual spell, all targets with a willpower value may resist at
their current Magic Resistance. A successful resistance may reduce the
spell effect for the target, or even nullify the effects altogether.
In some cases, the duration, damage, or other aspect, may be random,
and will need to be determined. If the Cast Check is a double or
triple effect, the element to be doubled or tripled must be decided
before anything else is resolved.

All magic works in quanta. Any attribute of any magic may be performed
at any Rank up to the Mage's maximum level of skill, but only at a
whole number of Ranks. It is assumed that all attributes are being
cast at maximum, unless it is otherwise stated before the Cast Check
is made.

\subsubsection{Damage}
\index{spells!damage}

Damage due to magic ignores armour and does not cause stunning, unless
a strike check against the target's defence is required as part of the
spell. Damage is done to Fatigue, and to Endurance once Fatigue is
exhausted. A single Damage Check will not ``wrap'' from Fatigue to
Endurance, unless the total damage exceeds the target's combined full
Fatigue and Endurance.

\subsubsection{Doubles and Triples}
\index{spells!multiple effects}

There are three characteristics of a spell which can be increased by
the Mage as a result of a spell causing a double or triple effect:
Range, Duration and Damage.

Whenever a spell is cast for double effect, the Mage has the option to
double one of Range, Duration or Damage. Some spells may not have one
or more of these attributes. Such attributes may not be affected.

Whenever a spell is cast for triple effect, the Mage has the option of
either tripling one of Range, Duration or Damage; doubling any two of
these attributes; or decreasing the target's Magic Resistance by 20.

A Mage may attempt to cast a spell at a target which is not within
range in the hope of achieving a double or triple effect.

\subsection{Restrictions on Magic}
\label{magic:restrictions}

Mages are restricted as to where and when they can employ magic.
General restrictions that apply to all Mages are covered in this
section. Specific restrictions applying to Adepts of particular
Colleges are covered in the opening sections of those Colleges.

\subsubsection{Cold Iron}
\index{cold iron}

A Mage may never prepare or cast a spell or engage in Ritual Magic
while in physical contact with Cold Iron. They may exercise any Racial
Talent Magic, but no learnt Talent Magic. Wearing armour made of Cold
Iron, or holding weapons or tools made of Cold Iron is regarded as
being in physical contact. GM discretion covers all other
cases. Several ounces of Cold Iron is required to cut off the mana
flow.

There are several possible means of circumventing the effects of cold
iron:

\begin{Itemize}
\item The Mage may drop all iron items prior to performing any
magic. Note that donning and doffing armour is very time consuming.

\item The Mage may employ weapons, tools and armour that are not
metallic (\eg quarterstaff, leather armour).  Weapons that are
normally metallic can be made out of wood, bone or stone, but their
Base Chance is reduced by 10, their Damage reduced by 2; and their
Fumble chance increases (see \S\ref{combat:fumbles}).  A similar loss of
effectiveness will be experienced with other tools that are normally
iron.

\item The Mage may use metallic items that have no iron content, such
as copper, tin and bronze. Such items cost the same as equivalent iron
items, but they are less effective: weapons do one less point of
damage, and fumble on a 98 to 00 roll (instead of just a 00); and
armour provides two less points of protection.

\item The effects of cold iron can be neutralised by combining it with
precious metal (silver or true-silver). Such items are as effective as
Iron items. Silvered items costs at least 10 times the standard price;
and true-silvered items 180 times the cost. A Mage's Cast Chances are
reduced by 10 if carrying silvered items.  Wearing Cold Iron does not
protect from the effects of magic.
\end{Itemize}

\subsubsection{Confinement}

A Mage must have the freedom to make the necessary gestures and sounds
in order to cast a spell or perform a ritual. Mute, bound, paralysed,
unconscious, stunned, prone or restrained Mages may not use Spell or
Ritual Magic, though Talent Magic is usually possible.  In addition,
an Adept must have at least one hand free and be able to speak clearly
to prepare and cast a spell.

\subsubsection{Proper Procedure}

A Mage may never employ a type of Magic, whether a Spell, Ritual or
Talent, which they have not learnt. The Mage also must have whatever
equipment or working materials are specified in the Spell or Ritual
description.

\subsubsection{Concentration}
\index{concentration!casting}
\label{magic:concentration}

A Mage may not cast a spell or perform a ritual if their concentration
is broken. This usually occurs by being engaged in Melee or Close
Combat. Other types of attack or distraction may also suffice. If an
event is deemed distracting, the Mage's player must roll a 4 times
Willpower check to maintain concentration, or the spell or ritual is
disrupted.

The concentration required to control spells already cast, or the
concentration required to control an entity, will not be broken by
entering combat or being attacked. It will only be broken if they are
stunned, knocked out or killed.

\subsubsection{Queueing}
\index{spells!queueing}

Spells that have the same effects are not cumulative. If a spell is
cast on a target that is already under the effect of a spell which has
the same effect, then the spell ``queues''. This means that, although
the spell is in effect on the target, it has no effect until the
earlier spell is gone.

Spells with any overlapping effects are affected by this rule. Note
that it is the spells' effects that are important, not the spell
itself. If the spells both affect one attribute (\eg defence), they
queue. If two spells affect different attributes (\eg PS \& FT), they
stack. This rule also applies to items. Unless stated otherwise, any
item that contains a magical effect that can be caused by a spell,
cannot be affected by that spell (\eg a weapon with a magical bonus to
hit or damage may not stack with Weapon of Flames).

The duration of the second, queueing, spell is measured from casting,
but it only takes effect when the first spell wears off.

\subsection{Backfires}
\label{magic:backfire}
\index{backfires}

Particularly inept Spell casts or Ritual performances may cause
backfires. If the Mage's Cast Check fails by more than 30 for a 5
second Spell preparation, or 40 for a longer Spell preparation or
Ritual performance, the Magic backfires. The Magic will always
backfire on a natural roll of 100, unless the Cast Chance is over
100\%, in which case the Magic fails. The GM rolls on the Backfire
Table (\S\ref{tables:backfire}) and applies the result. Effects
include extra Fatigue loss, partial or awry magical effects, and
curses and afflictions on the caster. The Backfire Table
(\S\ref{tables:backfire}) is just one of many possible tables, and the
GM may make free of any other table they possess.

The general structure of the table should be retained, with about 25\%
being Fatigue Loss, 35\% of the spell going awry, and 40\%
afflictions.  High mana and / or Special Knowledge backfiring could
make backfire effects worse, whereas low mana and / or General
Knowledge could make them not as bad.

\subsubsection{Backfire Interpretation}

The effects are to be interpreted as widely as desired by the GM. All
curses and afflictions are resistible, and partial or awry effects may
be also, depending on the magic. All backfire effects are
cumulative. It may be impossible to apply a specific backfire effect
in certain situations. This is generally described by ``No apparent
effect'', along with most subtle afflictions. The effects of a
backfire should be kept secret for as long as possible.

In most cases, specific reductions in numerical ratings are given when
a Mage is cursed as a result of a backfire. However, ancillary effects
of the curse must be determined within the guidelines of the curse
description.

\subsection{Magic Resistance}
\label{magic:resistance}
\index{resistance!magic}
\index{magic!resistance}

An entity who is the target of a spell may resist the spell if it is
resistible. There are two types of resistance: Active Resistance
reduces the Cast Chance of a spell; and Passive Resistance avoids or
reduces the effects of a spell. Magic Resistance is used for both
Active and Passive Resistance. It is equal to an entity's Willpower,
modified as follows:

{\small \begin{tabularx}{\linewidth}{Xr}
Target and caster are of the same Branch of Magic & +5 \\
Target and caster are of opposed Branches of Magic & -5 \\
Target is sentient but not a Mage & +20 \\
Target has on, or is in the area of, the appropriate counterspell & +30\dag \\
Spell had triple effect and the caster chose to affect Magic Resistance & -20 \\
Target is in a consecrated area & +50 \\
\end{tabularx}
\dag plus 3 per Rank}

Branches of Magic are covered in \S\ref{magic:colleges}.  Sentient
entities are those with an MA of 0 or greater.  Counterspells are
covered in \S\ref{magic:counterspells}.  Triple Effects are covered in
\S\ref{magic:effects}.  Consecration is not covered in these rules.

Purification, certain other spells, and many items also affect Magic
Resistance.

If an entity successfully resists a targeted spell, then they will be
aware that they have resisted some Magic. While the spell did not
directly affect the entity, their aura is marked sufficiently that the
last spell to impact is determinable. Note that only spells which
target possessions or entities are noticeable in this way.  Spells
which affect an area cannot be detected when a character resists them.

Objects cannot normally resist Magic. However, if the object is an
entity's possession, and the spell can be passively resisted, then the
entity may apply their normal resistance.

\subsubsection{Passive Resistance}
\index{resistance!passive}

When a spell that is passively resistible impacts on an entity, the
entity may attempt to resist the effects of the spell. This is known
as Passive Resistance. The player must roll D100. If the die roll is
less than or equal to the entity's Magic Resistance, the spell's
effects are reduced or nullified. Passive Resistance is an automatic
bodily function which occurs regardless of whether an entity is
conscious or not. If an entity re-encounters an area effect magic,
they must re-resist, whether they successfully resisted last time or
not. At the start of a pulse, an entity may choose to not resist. For
the remainder of the pulse, the entity may not passively resist any
spells, unless they become stunned or unconscious. While choosing to
not resist, an entity may only perform a pass action.

\subsubsection{Active Resistance}
\index{resistance!active}

An entity can choose to actively resist another entity. When an entity
attempts to cast or trigger a spell which is actively resistible, then
the highest Magic Resistance of any target who is actively resisting
the entity is subtracted from the Cast Chance. Note that Active
Resistance is only effective if the entity who is actively resisting
is a target, or in the area of effect of the spell. In combat, Active
Resistance is a Pass Action. An entity must concentrate on the caster
in order to actively resist them. Anything that can disrupt Spell
preparation can also disrupt Active Resistance. Active Resistance
against a spell that is not able to be actively resisted, or against a
spell which is not targeted at the entity, has no effect.

\subsection{Incorporating Magic into Combat}
\label{magic:combat}
\index{combat!magic}

Ritual Magic is difficult to employ during combat. During each pulse
that noisy or dangerous events take place, the Mage may need to make a
concentration check (as per \S\ref{magic:restrictions}). Talent Magic
may be used during combat.  Passive talents operate normally and do
not require any actions to employ. Active talents (\eg Detect Aura)
require a pulse to implement. Spell Magic may be employed, usually
with 5 second preparation.

A Mage must perform a magical pass action to prepare and a cast action
to cast a spell.  See \S\ref{combat:engaged} and
\S\ref{combat:unengaged} for what can be done during pass and cast
actions in combat.

An entity may actively resist a spell during combat by implementing a
pass action. They may lower passive resistance during any pass action.
Triggering an invested item takes one full pulse irrespective of the
number of actions that can be performed.

\subsection{The Colleges of Magic}
\label{magic:colleges}
\index{magic!colleges}

Most Magic is divided into 16 Colleges representing specific types of
magic. A Mage may only employ the powers and spells of one College. If
a Mage belongs to a College, they are known as an Adept.

\subsubsection{Branches of Magic}
\index{magic!branches}

The Colleges are divided into three Branches of Magic, as follows:

\begin{flushleft} {\bf The Thaumaturgies:} \\

The College of Bardic Magics \\
The College of Binding and Animating \\
The College of Ensorcelments and Enchantments \\
The College of Illusions \\
The College of Sorceries of the Mind \\
The College of Naming Incantations \\

{\bf The Elementals:} \\
The College of Air Magics \\
The College of Celestial Magics \\
The College of Earth Magics \\
The College of Fire Magics \\
The College of Ice Magics \\
The College of Water Magics \\

{\bf The Entities:} \\
The College of Greater Summonings \\
The College of Necromantic Conjurations \\
The College of Rune Magics \\
The College of Witchcraft \\
\end{flushleft}

An Adept's Magic Resistance is affected by their Branch of
Magic. Their resistance against spells of their own Branch is
increased by 5, while their resistance against magic of the opposed
branch is reduced by 5. Thaumaturgies and Entities are opposed
Branches. The Elemental Branch is not opposed to any group of Magics.

{\small \begin{tabular}{l@{\hspace{0.5em}}l@{\hspace{0.5em}}l@{\hspace{0.5em}}l}
		& {\em Thaum}	& {\em Elemental} & {\em Entity} \\
{\em Thaum}	& Same		& Neutral	& Opposed \\
{\em Elemental}	& Neutral	& Same		& Neutral \\
{\em Entity}	& Opposed	& Neutral	& Same \\
\end{tabular}}

\subsubsection{Restrictions and Modifications}

Each College of Magic has its own individual minimum Magical Aptitude
requirement.  This must be met at the time any entity becomes a Adept
of the college.  Many Colleges have restrictions on casting magic
further to those in \S\ref{magic:restrictions}.  These are specified
in the first sub-section of each College.

Most Colleges are subject to certain modifications to Cast Chances in
addition to \S\ref{magic:modifiers}.  These are specified in the
second sub-section of each College.

\subsubsection{Learning College Magic}
\index{magic!learning}

An Adept is assumed to have mastered all of the General Knowledge
magic of their College upon the completion of their training. This
mastery is at Rank 0. Special Knowledge magic is not taught to
novices, and can only be acquired by expending time (and usually
money) to learn it to Rank 0. Most Special Knowledge magic is
available at the Guild, at fixed prices. All General and Special
Knowledge magic may be ranked to Rank 20 by the expenditure of time
and experience.

An Adept must have ranked a spell or ritual to Rank 6 before they can
teach it. They must have ranked all talents, and general knowledge
spells and rituals to Rank 6 before they can teach a novice their
College.

In general, an Adept may never use spells, talents, or rituals of more
than one College of Magic at one time (except Counterspells). An Adept
may change College but loses all General and Special Knowledge magics
from their old College, and must spend 6 months (and 6,500 ep) learning
the ways of the new College (including characters learning a college
for the first time).  Once an Adept has renounced a College, they may
never return to it.

\subsubsection{Knowledge Limitations}
\index{spells!magical aptitude limitation}

An Adept may only employ Talents, Spells and Rituals that they
know. They may know any number of talents, but may not know more
spells and rituals below Rank 6 than their Magical Aptitude. They may
know an unlimited number of Spells and Rituals of Rank 6 or
higher. All General and Special Knowledge Spells and Rituals of the
Adept's College, plus whatever non-Colleged magics known, apply to
this limit (except Ritual Spell Preparation, and the General Knowledge
spells of Namers). This includes Counterspells of other Colleges.

An entity may not become an Adept of a College of Magic unless they
have the Magical Aptitude to account for mastery of the General
Knowledge spells and Rituals of that College. This is enumerated in
the restrictions of the College. An Adept may not attempt to acquire a
spell or ritual if they already know a number of these equal to their
Magical Aptitude below Rank 6.

If, as a result of a decrease of Magical Aptitude or spell ranks, the
Adept knows more spells and rituals below Rank 6 than their Magical
Aptitude, then they will permanently lose sufficient knowledge to
satisfy this rule, losing the lowest ranked magics first. They will
still remain Adepts of their College, while sentient.

\subsection{Magic Descriptions}

The description of all the College Magics work under certain
conventions. The more important of them follow:

\subsubsection{Sub-Sections}

Each College description is broken up into
\begin{Itemize}

\item An Introduction
\item Restrictions
\item Cast Chance Modifications
\item Talents (coded T-\#)
\item General Knowledge Spells (coded G-\#)
\item General Knowledge Rituals (coded Q-\#)
\item Special Knowledge Spells (coded S-\#)
\item Special Knowledge Rituals (coded R-\#)

\end{Itemize}


\subsection{Spell Descriptions}

The description of each spell lists its specific effects, range,
duration, and other attributes. Each spell is fully described under
the College to which it belongs. The following information is
included:

\begin{Description}

\item[Rank Modifications]
Often range, duration and other effects will be given as ``x + y /
Rank''. This means that the characteristic is equal to x, with an
additional y for each Rank attained in the magic. Unless otherwise
noted, the unit of measurement is the same for x \& y. If an increase
of y is noted for each n Ranks, then partial multiples of n do not
count unless specifically stated.

\item[Range]
The maximum radius (in feet) within which the Mage can make the spell
take effect.  This is always the distance from the Adept. It can be a
linear measurement between Adept and target, or a radius of
effect. Unless explicitly stated, magical effects will not occur
beyond the range of the magic. In combat, measurements are taken from
the middle of a hex, and rounded upwards.

\item[Duration] The maximum length of time that the spell remains in
effect. Spells with a concentration component will stop as soon as
concentration is broken. Spells that do not require concentration will
persist regardless of the suffering of the Mage; even unto death.  If
a spell is cast in the middle of a pulse, that pulse counts towards
its duration.

\item[Experience Multiple] The multiple used in conjunction with the
rank to be achieved to determine the experience cost of increasing a
Mage's Rank with a particular spell.

\item[Base Chance] The base percentage chance of succeeding in casting
a spell.  This may be equal to some multiple of a characteristic of
the Adept. The characteristic is taken at its current value,
multiplied appropriately, and then other modifiers are added.

\item[Resistance] The Magic Resistances (Passive and Active) which may
be applied against the spell by its targets.

\item[Storage] The valid ways that the spell may be stored, for
example, investment, potions or warding.

\item[Target]
Spells and Rituals are targeted at either an Area, Object or
Entity. Some spells can be cast at more than one target type. If a
multi-target or area effect spell is actively resisted, the highest
Active Resistance of those to be affected is applied. Area effect
spells are resisted each time that an entity encounters them.
Animated objects count as objects and entities.

\item[Effects] The general purpose and consequences of the
spell. Includes potential damage, effects of resistance, special
effects, and any exceptions to the normal workings of magic.

\item[Difficulty Factor]
A difficulty factor will sometimes be given to avoid a spell. This is
always a number by which the stated characteristic of the target is
multiplied, before modifiers are added.

\end{Description}

\subsubsection{Interpretations}

Most of the magic in DQ is designed to be flexible in application, and
up to the interpretation of the GM within the guidelines laid down by
the Gods. The effects and procedures are meant to apply to humanoid
entities of human size.  An Incinerate Spell that would fry a human
would do little more than discomfort a Dragon. To close every loophole
and explain every application would be impossible. Therefore, these
matters of interpretation have been left to your GM, in the context of
their game and the atmosphere that they are trying to promote.

\subsection{Storage and Entrapment of Magic}
\index{spell storage}

There are various methods of storing magical effects.  Each has
different properties and can store different types of spells.

\begin{Description}

\item[Potion] Spells, Talents and some quasi-magical skills (\eg
Healer) can be potioned by an Alchemist.  For Talents, the imbiber
receives the usage of the talent for a duration dependent on the Rank
(see \S\ref{alchemist}).  For skills, see the skill involved.  For
spells, imbibing a potioned spell is equivalent to being the Adept and
casting the spell on oneself.  This is normally the only way for self
only spells to be stored.  Target area and target object spells cannot
be potioned, but spells that affect entities may (possibly) be
potionable, for example, Necrosis could be potioned, but drinking the
potion would only cause the imbiber to have their internal organs
ruptured.  Potions always work, but they cannot double or triple
effect.

\item[Investment] Spells effects can often be stored in items and at
a later stage, be \emph{triggered}.  When a spell is triggered, it is
as if the Adept was there and had just cast the spell (except the
spell modifiers such as base chance, range, etc, are fixed at the
time of investment).  The entity triggering the spell gets to choose
the target(s) of the spell at the time of triggering and maintain
concentration spells.

\item[Ward] A ward is a way of storing a spell within an object, area
or volume so that when some simple condition is met, the ward is
triggered, and the entites or objects that fulfilled the condition
becomes the target for the spell.  When a ward is triggered, it is
always successful, but cannot multiple effect.  For a spell to be
wardable, it must have a range or area component (the range may not be
touch, nor self).  Wards cannot maintain concentration spells.

\item[Magical Trap] Magical effects can be stored in mechanician
traps.  Unlike wards, traps have to be physically triggered.  The
spell effect, unless an area effect, will only be targeted on the
triggerer.  Similar to wards, magical traps can only store spells that
have a non self range.  The spell in a trap can only target an area or
the triggerer.  Spells in traps will always work, but cannot multiple
effect.

\item[Shaped magic] Magical items beyond that of containing simple
invested spells are known as shaped items.  Shaped items come in two
flavours, charged and permanent.  Charged items have a number of
charges which diminishes with use.  Certain items are said to have
``bound charges'', which means they behave as an invested item except
they can be recharged.  Charged items which cannot be recharged lose
their magical status once all charges have been expended.  Permanent
shaped items can come in any shape or form and can defy many of the
usual ``rules'' of the magical universe.

\end{Description}

\index{magic|)}
\end{multicols*}
