\begin{college}[1.6]{mind}{Sorceries of the Mind}{MI}

\textbf{This version of the College is probationary, and is under
trial for two years (as of September 2001).  All characters that join
this College do so under the understanding that if this version is
deemed unplayable, it may be withdrawn or radically changed.  At that
point, all such characters may revert to the previous version or have
the EP and time spent in this College refunded, to be re-spent freely
in another College as if that College had been their original choice.}

\subsection{Introduction}

The College of Sorceries of the Mind deals primarily with controlling
or influencing the minds of others. Adepts of this College are
variously known as Mind Mages, or Sorcerers. It is widely believed
that Sorcerers can peer into the darkest nooks and crannies of the
soul, and are near-universally feared.

\textbf{Traditional Colours \& Symbols:} This college has no
traditional colours or symbols as Sorcerers prefer to blend into
society.

\textbf{Restrictions:} Adepts of the College of Sorceries of the Mind
may practice their arts without restriction.

The Magical Aptitude requirement for becoming a member of this College
is 11.

\subsubsection{Modifications}

The Base Chance of performing any talent, spell or ritual of the
College of Sorceries of the Mind against a single entity target is
affected by relative Willpower:

\begin{inline}
\begin{tabularx}{\linewidth}{Xr}
For each point the Adept's Willpower is above the target's Willpower	& +1 \\
For each point the Adept's Willpower is below the target's Willpower	& -2 \\
\end{tabularx}
\end{inline}

\subsection{Talents}

\begin{talent}[T-1]{Resist Pain}

\range{Self}
\multiple{300}
\storage{Potion}
\begin{effects}
The Adept is largely immune to pain. They cannot be tortured or
stunned by physical means. In addition, if there is a chance that
their concentration may have been broken, 5 (+ 1 / Rank) is always
added to their Concentration Check. No matter what the circumstances
the Adept will always have a chance of maintaining concentration. Like
all magic, this talent is affected by cold iron. However, for a weapon
to affect the talent, the weapon must be made of cold iron, must have
done endurance damage, and must remain in contact with the body. If
cold iron is used in the torture of a mind mage, the Adept adds their
concentration bonus to the base chance of resisting the torture
attempt.
\end{effects}
\end{talent}

\begin{talent}[T-2]{Resist Temperature}

\range{Self}
\multiple{250}
\storage{Potion}
\begin{effects}
The Adept is immune to climatic extremes of temperature from arctic
cold to desert heat whether generated naturally or by magic. They
therefore do not suffer from exposure, frostbite, wind-chill, heat
stress, hypothermia or hyperthermia, and their fatigue loss is
unaffected by temperature. If the Adept is attacked by fire, ice, heat
or cold (whether magical or not) then the damage points are reduced by
1 + 1 per 5 full ranks.
\end{effects}
\end{talent}

\begin{talent}[T-3]{Sense Danger}

\basechance{PC + 3 / Rank}
\multiple{300}
\storage{Potion}
\begin{effects}
The Adept may sense the presence of a hazard to the Adept's life or
wellbeing. Generally this may be interpreted as a ``bad feeling''
about a person or an object, or a sense of foreboding about a
situation.  This talent operates continuously and unconsciously. The
Adept may also add 2 + 2 / Rank to the chance of detecting an ambush
(see Ranger) and to the chance of detecting a trap (see Spy/Thief).
\end{effects}
\end{talent}

\subsection{General Knowledge Spells}

\begin{spell}[G-1]{Control Animal}

\range{30 feet + 15 / Rank}
\duration{Concentration: no maximum}
\multiple{100}
\basechance{40\%}
\resist{Passive}
\target{Non-sentient entity}
\storage{Investment}
\begin{effects}
The Adept gains control over one non-sentient entity within
range. Control is lost if the entity leaves the range but recommences
if range is re-entered. The Adept does not receive any sensations from
the animal. If the Adept releases the animal or their concentration is
broken the animal may attack the Adept. The Adept controls the
animal's mind and therefore need not know how to make a bird fly: the
Adept commands the animal to do something and the animal knows how to
do it. The animal will comply within the spirit of the command.
\end{effects}
\end{spell}

\begin{spell}[G-2]{Control Person}

\range{30 feet + 15 / Rank}
\duration{Concentration: no maximum}
\multiple{650}
\basechance{30\%}
\resist{Active, Passive}
\storage{Investment}
\target{Sentient Entity}
\begin{effects}
The Adept gains absolute control over every action of one sentient
entity within range. The Adept does not receive any sensations from
the target. The Adept does not need to know the target's language to
control them. Control is lost if the target leaves the range but
recommences if range is re-entered. They may control the physical
activities of that target, and use the target's physical skills where
known, but may not cause the target to cast spells.

The Adept may not use their own skills through the target. The target
always acts as if they were also affected by a Spell of Slowness (E\&E
S-11) unless they choose not to resist the spell, or subsequently
decide to co-operate with the Adept's commands. The spell continues in
effect until the Adept's concentration is broken or the Adept chooses
to release the target. The target is aware of the control, and
although the target is unable to do anything about it at the time,
they will remember being controlled. The Adept may force the target to
act in direct opposition to the target's own wishes. Suicidal
instructions however, grant the target another magic resistance which,
if successful, dissipates the spell.
\end{effects}
\end{spell}

\begin{spell}[G-3]{Empathy}

\range{Touch until Rank 10}
\duration{Immediate}
\multiple{200}
\basechance{20\%}
\resist{None}
\storage{Investment}
\target{Entity}
\begin{effects}
The spell allows the Adept to feel the emotions and physical
sensations which the target of the spell is currently experiencing. It
also allows the Adept to absorb wounds from Endurance and Fatigue at a
rate of 2 points cured for every 1 which the Adept agrees to subtract
from their own Fatigue (never Endurance). The additional fatigue
subtracted by the Adept is damage fatigue, not spell fatigue. This
spell may not be used to regenerate spell or tiredness fatigue, but
only reduces/transfers damage. At Rank 10 or higher, the Adept can
cast this spell on a target who is within 15 feet (+ 15 / Rank over
10) or less from them. This spell may be cast on self, but only to
cure endurance damage. On a `Double' effect, 3 Damage Points are cured
for each point inflicted on the Adept; a `Triple' effect allows for 4
points per point inflicted on the Adept.
\end{effects}
\end{spell}

\begin{spell}[G-4]{ESP}

\range{30 feet + 15 / Rank}
\duration{30 seconds + 10 / Rank}
\multiple{100}
\basechance{40\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept can sense the presence and general mood (but not the exact
nature) of all entities within range of the spell. The Adept has a
general idea of where each entity is, say to within 90 or 60 degrees,
the accuracy of which may increase with rank. ESP will detect the
presence of an entity even if that entity is mind cloaked. Spells
cannot be targeted through ESP.
\end{effects}
\end{spell}

\begin{spell}[G-5]{Hypnotism}

\range{15 feet + 15 / Rank}
\duration{Concentration: no maximum}
\multiple{200}
\basechance{40\%}
\resist{Active, Passive}
\storage{Investment}
\target{Entity}
\begin{effects}
The spell causes an entity to accept suggestions from the Adept.  The
spell may only be cast over a target with whom the caster is normally
able to communicate verbally. It can never be cast over a totally
hostile target. The target may be enabled to remember otherwise
forgotten details through appropriate questioning.

Once the target has been hypnotised, the Adept can make suggestions
which the target will readily accept unless they conflict directly
with their best interests. The target will never remember where these
suggestions came from.

The Adept may choose up to 1 (+ 1 / 5 Ranks) of the above suggestions
to be active in a post-hypnotic manner, which the target will continue
to implement for 3 (+ 3 / Rank) hours after the spell ceases. The
target will stop following all other suggestions once the spell
ceases.
\end{effects}
\end{spell}

\begin{spell}[G-6]{Limited Precognition}

\range{Special}
\duration{Immediate}
\multiple{150}
\basechance{20\%}
\resist{May not be resisted}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept may see (unclearly) up to 1 (+ 1 / Rank) hours into their
own future and may foresee possible events.  The Adept's player should
describe a plan of action, and the GM will describe a vision based on
the consequences of those actions, which can be as literal or
metaphorical as the GM wishes. The clarity and detail of the vision
will be dependant on the Adept's rank. Should the Adept foresee their
own death, a fright check may be appropriate. This spell cannot be
recast immediately with a different plan; the Adept must wait until
they have taken some action to change their situation. This spell
works at any range.
\end{effects}
\end{spell}

\begin{spell}[G-7]{Mind Shield}

\range{Self}
\duration{1 hour + 2 / Rank}
\multiple{250}
\basechance{30\%}
\resist{May not be resisted}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept cloaks their own mind so that their thoughts cannot be
detected or ``read'', \eg by Telepathy or other mind reading talents,
spells or rituals. The Adept's Resistance versus Mental Attack is
increased by 10 (+ 2 / Rank) while the spell is in effect. Mind Cloak
does not block ESP or Empathy. The Adept can cast Mind Speech as
normal, and may choose to decline Mind Speech from an external source.
\end{effects}
\end{spell}

\subsection{General Knowledge Rituals}

\begin{ritual}[Q-1]{Binding Will}
\range{10 feet}
\duration{1 day + 1 / Rank}
\multiple{500}
\basechance{10\% + 5\% / Rank}
\resist{Passive}
\casttime{1 hour}
\begin{effects}
The Adept may bind the will of an entity so that they become the loyal
retainer of the Adept and serve them in all things.  The target must
be in range during the ritual but may be unconscious or otherwise
restrained. The target gets an additional resistance check at the
beginning of each following day and if required to do anything
suicidal. If they successfully resist, the binding is broken and they
are free. The target is then aware that they have been under an
external influence. The Adept may never release the binding
voluntarily, and until the duration expires the binding continues,
even if the Adept dies (the target will do everything possible to get
the Adept resurrected).  If the ritual backfires, the Adept loses D10
from their Willpower in addition to any other backfire effect;
Hypnotism (G-5) cures this loss.
\end{effects}
\end{ritual}

\subsection{Special Knowledge Spells}

\begin{spell}[S-1]{Disruption}

\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{400}
\basechance{25\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Object or Entity}
\begin{effects}
The Adept causes an object or corporeal entity (substantial rather
than insubstantial) to pulsate, doing [D + 1] (+ 1 / Rank) points of
damage if the target fails to resist, and half damage to any target
that successfully resists. At Rank 10 or above, the target's internal
structure and surface will be partially disrupted, requiring the arts
of a healer or artisan of rank equal to (spell Rank / 3 , or Rank / 6
if the target resisted the spell) to prevent or remove the scaring
(damage can be cured normally).
\end{effects}
\end{spell}

\begin{spell}[S-2]{Force Shield}

\range{Self}
\duration{10 minutes + 10 / Rank}
\multiple{250}
\basechance{40\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
This spell creates an invisible protection around the entire body
which increases the Defence of the Adept by 5 (+ 1 / Rank) against all
attacks at any range.
\end{effects}
\end{spell}

\begin{spell}[S-3]{Healing}

\range{Touch}
\duration{Immediate}
\multiple{375}
\basechance{40\%}
\resist{None}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
The Adept may heal the target of 2 (+ 1 / Rank) Damage Points that
have been removed from the target's Fatigue or Endurance. Endurance
damage is healed first, wrapping to Fatigue damage. On a multiple
effect, the Damage Points restored may be multiplied.
\end{effects}
\end{spell}

\begin{spell}[S-4]{Mental Attack}

\range{15 feet + 15 / Rank}
\duration{10 seconds + 10 / Rank}
\multiple{350}
\basechance{25\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
The Adept may cast this spell at any entity within range who can be
seen or whose position has been pinpointed via Telepathy. If the
target fails to resist, they fall unconscious and at Rank 15 and above
the Target loses [D - 5] from Willpower.  The Willpower loss is
recoverable via Hypnotism, Remove Minor Curse or naturally at one
point every 3 days. Repeated casts may further reduce the targets
Willpower to a minimum of 1.
\end{effects}
\end{spell}

\begin{spell}[S-5]{Mind Speech}

\range{30 feet + 30 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{400}
\basechance{20\%}
\resist{May not be resisted}
\storage{Investment}
\target{Entities}
\begin{effects}
Allows the Adept to choose a Principal Target, plus up to 1 + 1 / Rank
other targets, who may then communicate mentally to each other as if
talking aloud in a group.  The entities may be targeted via the Spell
of Telepathy (at the time of casting). If an entity subsequently goes
further than 30 feet + 30 / Rank from the Principal Target, their
participation ceases until they re-enter that range. A language in
common is required for normal speech. At Rank 10 and above visual
images may be transmitted. All communication requires the same level
of concentration as for normal speech. The spell dissipates if the
Principal Target is killed.

Multiple casts are a special case. If one or more entities are part of
two or more spells, and are within range of all of their Principal
Targets, then every target of those spells may speak to each other as
if in one group. This can effectively increase the working range of
the effect.
\end{effects}
\end{spell}

\begin{spell}[S-6]{Phantasm}

\range{Special}
\duration{Concentration: no maximum}
\multiple{400}
\basechance{1\%}
\resist{May not be resisted}
\storage{Investment}
\target{Entity}
\begin{effects}
This spell creates an insubstantial, magical beast that must be
targeted at a particular victim which is either visible to the Adept
or located through the spell of Telepathy. The target may interpret
the beast as being a phantasm from their own worst nightmare. Lack of
imagination, however, does not offer any protection or immunity. The
Adept has no control over the shape of the creature.

The phantasm appears beside the Adept, and then travels straight
towards the target at a TMR equal to 1 + 1 / Rank, passing through
walls, air etc. When in melee range of the target the Phantasm always
does [D - 4] (+ 1 / Rank) damage each pulse, including the pulse when
it arrives at the target.

The phantasm is only substantial to, and visible to, the target and
thus can be directly affected by the target (using spells or magical
weapons). Anyone may cast a Mind College Special Counterspell over
the area currently occupied by the phantasm to dissipate it. The
phantasm is immune to all other spells and weapons.

The Phantasm's magic resistance, defence and initiative are all zero
and it always acts last in the pulse. The phantasm has a single
combined Endurance and Fatigue of 20 (+ 5 / 3 Ranks).

The phantasm dissipates when: its endurance drops to zero; it is
dispelled by counterspell; its target dies or leaves the plane; the
Adept drops or loses concentration. The Adept has some awareness of
the phantasm's location, and is aware of its dissipation.
\end{effects}
\end{spell}

\begin{spell}[S-7]{Telekinesis}

\range{15 feet + 15 / Rank}
\duration{Concentration: maximum 10 seconds + 10 / Rank}
\multiple{250}
\basechance{25\%}
\resist{None}
\storage{Investment}
\target{Entity or object}
\begin{effects}
Allows the Adept to lift and manoeuvre a target weighting up to 2
pounds and move it at the rate of 2 TMR. The Adept may move an
additional 5 pounds or move the chosen target at an additional 3 TMR
per Rank. The Adept may increase both mass and speed by applying
separate ranks to each. Gross movement is automatic, although fine
movements require an MD check and should attract negative modifiers
for range.
\end{effects}
\end{spell}

\begin{spell}[S-8]{Telekinetic Rage}

\range{30 feet + 5 / Rank}
\duration{Instantaneous}
\multiple{750}
\basechance{1\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
The Adept creates an instantaneous, magical storm that fills the area
within range with roiling, body-wrenching forces. All objects and
entities that fail to resist, are hurled away from the Adept until
they are out of the area of effect, or they strike a solid barrier.

The storm inflicts [D - 5] (+ 1 / foot travelled) of magical damage
and is inflicted once only, at the time of casting. The damage is not
contingent on striking an object. Additional damage may be done on
hitting a wall, entity or other substantial object.
\end{effects}
\end{spell}


\begin{spell}[S-9]{Telepathy}

\range{30 feet + 30 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{550}
\basechance{10\%}
\resist{Passive}
\storage{Potion}
\target{Self}
\begin{effects}
The spell gives the Adept the ability to read the surface thoughts of
any entity within range who fails to resist. Each target may only
resist each Telepathy once, regardless of whether they move in or out
of the range of the spell.

Each pulse, the Adept may use their telepathy to either locate the
minds that are within range, or to read the thoughts of a particular
mind. The Adept need not know the language of the entity whose
thoughts are being read, but will have difficulty comprehending the
thoughts of alien minds. Animal thoughts are usually very primitive
and can summed up in a single word, \eg ``food'', ``danger'', ``sex.''

It is not possible to utilise a target's senses and the Adept may only
gain an impression of physical sensations if the entity is
concentrating on them, \eg savouring a meal.

Telepathy may be used to target the Mind Speech, Mental Attack and
Phantasm spells.

Targets are not aware that their thoughts are being read.
Concentration checks should be required of targets who try to control
their thoughts by reciting poetry etc.
\end{effects}
\end{spell}

\begin{spell}[S-10]{Transmutation}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{1000}
\basechance{1\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity or Object}
\begin{effects}
The Adept may transmute and rearrange the elemental components of any
target that fails to resist, transforming it into any form of the same
mass that the Adept desires. A living entity can only be rearranged
into another living form, and objects into objects. This spell cannot
confer skills or magical abilities, but the target will acquire the
non-magical features of the new form such as wings, gills etc, and may
use them. If the new form is maintained without disruption for the
duration of the spell, the target will revert to its original shape.
\end{effects}
\end{spell}

\begin{spell}[S-11]{Undetectability}

\range{Touch}
\duration{10 minutes + 10 / Rank}
\multiple{450}
\basechance{15\%}
\resist{May not be resisted}
\storage{Potion, Investment}
\target{Entity}
\begin{effects}
The target is totally undetectable by physical senses. Changes to the
physical environment which are made by the target and are entirely
within 5 feet (+ 1 / Rank) (\eg leaves tracks, opens doors, picks up
an object) will be unnoticed by normal observers. Other entities will
automatically take such changes in their stride (\eg will just walk
through the previously shut door, will not sit down on the chair that
the target whipped out from under them but go and get another drink
instead).

Magical observation (Wizards Eye, Witchsight, Clair-audience etc) will
not detect the target. The spell does not however, affect indirect
observation, such as Locate, Telepathy or ESP.

The Adept may nominate, at the time of casting, up to 1 (+ 1 / 3 full
Ranks) entities who will ignore the workings of the spell.

The spell will dissipate if the target is in physical contact with
another entity for more than 5 seconds (other than the nominated
entities) or makes a strike check for a non-ranged attack (regardless
of its success). The target may choose to end the spell at any time.
\end{effects}
\end{spell}
\end{college}
