\begin{multicols}{3}

\section{Non-College Magic}
\label{rituals:noncollege}
\index{rituals!non-college}

\subsection{Introduction}

All spells and rituals given in a listing in this section can be
learnt by Adepts of any college.

\subsection{Special Knowledge Spells}

\begin{spell}[]{Geas}

\range{The Adept must be able to see and communicate with
the target}
\duration{Until  removed, fulfilled or target dies}
\multiple{250}
\basechance{Always successful (see below)}
\resist{Special}
\target{Sentient entity}
\begin{effects}
A geas is an obligation to complete a quest, an injunction against the
performance of a particular action, or a requirement to respond in the
same fashion to particular stimuli.

A being to be compelled must acknowledge their acceptance of the geas.
Furthermore, the Adept must believe that the being deserves the geas,
or the being must truly wish (not forced by physical or magical means)
to have an unmerited geas placed upon them.  The Adept specifies the
nature of the geas in 25 words or less, and the GM will use the most
liberal interpretation of that wording to the benefit of the compelled
being.  Rank with the geas spell does not affect the chance of casting
the spell as it is always automatically successful.  The Rank equals
the effectiveness of the geas, expressed in percentage terms.  If a
geased being directly contravenes the letter of a geas, they have a
chance of dying equal to the Rank of the geas.  A compelled being will
begin to feel weak or ill when they first take an action counter to
the restriction of the geas, and will become increasingly afflicted
until they once more comply with the geas.  There is no limit to the
Rank of a geas.

If a quest geas is fulfilled by the compelled being, they are no
longer subject to that geas.  The other two types of geas (for and
against a given action) last indefinitely.  A geas can be removed
automatically by the one who placed it upon the compelled being.  A
geased person will not attempt to free themself from the compulsion.
An Adept with Rank greater than that of the geas may try to do so.
The Adept must inscribe a triangle about the compelled, and perform
the ritual of geas removal for 12 consecutive hours.  If the triangle
is silver or True-silver, the compelled being does not suffer the
penalties for ignoring the geas during the ritual.

The Adept attempting to remove the geas has a success chance equal to
five times the difference between their Rank with geas and the Rank of
the geas being removed.  The GM rolls percentile dice: if the roll is
less than or equal to the success percentage, the geas is removed. If
the roll is greater than the success percentage, the Rank of the geas
is increased by one.

\textbf{Full Geas} A being with Rank greater than 50 with the geas
spell has the power of full geas. A full geas can be placed upon a
being without their consent, though it can be passively resisted.
Additionally, one with the power of full geas may automatically remove
(without the support of a triangle and 12 hours of ritual) a geas
which is at least 5 Ranks less than their Rank with the spell.
\end{effects}
\end{spell}

\begin{spell}[]{Major Curse}
\label{noncollege:majorcurse}
\index{curse!major}

\range{20 feet + 15 / Rank}
\duration{Until removed or target dies}
\multiple{750}
\basechance{15\%}
\resist{Passive (unless a Death-curse)}
\target{Entity or Object}
\begin{effects}
An Adept's Endurance value is decreased by one whenever they inflict a
major curse upon a being.  Note that when casting a Death-curse this
Endurance point loss is in addition to any possible Endurance point
loss due to resurrection.  There are several types of major curses:
\begin{Description}
\item[Affliction] The Adept may choose to torment or kill their
target. If the effects of the affliction curse are intended to be
deadly, the target may not die as a direct result of the curse before
(24 - Rank) hours have passed.  The following list of sample
affliction curses is provided to give the GM a guideline as to what
major curses should be allowed in their campaign.
\begin{Enumerate}
\setlength\itemsep{0pt}
\item
Target becomes totally blind, deaf or mute.

\item
Target becomes senile. 

\item
Target suffers from a contagious disease (for example open running sores). 

\item
Target is transformed into a frog or other small creature.

\item
Target becomes weakened and enfeebled and must be helped with any
physical action.
\item
Target falls into century-long sleep.
\end{Enumerate}

\item[Ill  Luck]
Add two times the Rank of the major curse spell to any percentile roll
involving the target or the use any of their abilities.  This may not
be applied favourably.

\item[Doom]
A doom is a pronouncement, by the Adept, upon an event that will occur
in the target's future (Example: ``You will die by the hand of a loved
one.'').  The statement which must be indefinite, will be true unless
removed.  The GM should be careful as to what to allow for dooms.

\item[Death-curse]
At the moment of their death, an Adept may automatically cast a major
curse (unless backfire occurs).  The being at which it is cast may not
resist the curse.  A death-curse must be an affliction, ill luck or
doom.  If a doom, it will be gasped out with the Adept's final breath.

\item[Note]
Lycanthropy is considered a major curse.
\end{Description}
\end{effects}
\end{spell}

\subsection{Special Knowledge Rituals}

\begin{ritual}[]{Remove Curse}
\label{ritual:removecurse}
\index{curse!remove curse ritual}
\duration{Immediate}
\multiple{500}
\resist{None}
\target{A curse}
\concentration{Standard}
\begin{effects}
Every curse is rated by the Magical Aptitude (MA) of the Adept who
cast it.  If the curse is natural (such as Lycanthropy) it usually has
an MA of 20.

There are two types of curses, minor ones and major ones.  A minor
curse causes its victim to suffer from a non-fatal malediction.  Minor
curses come from various sources, for example the spells Evil Eye (G-9
of the College of Ensorcelments and Enchantments), the Damnum Minatum
(G-1 of the Witchcraft College) and certain backfires.  Major curses
normally comes from the Major Curse spell (\S \ref{noncollege:majorcurse}).

When a ritual of curse removal has been completed, the GM rolls
percentile dice.  If the roll is less than or equal to the success
percentage the curse is removed. If the roll is between one and two
times the success percentage, the curse remains in effect.  If the
roll is equal to or greater than twice the success percentage, the MA
of the curse is increased by one.  This ritual does not backfire in
the normal fashion.

\textbf{Minor Curse}

\basechance{(15 - MA of curse + 5 \x Rank )\%}
\casttime{6 hours}
\actions{Inscribe symbol of power}
The Adept must inscribe a triangle or symbol of power about the cursed
being, and perform this ritual for six consecutive hours.

\textbf{Major Curse}

\basechance{(Adept's MA - MA of curse + 2 \x Rank)\%}
\casttime{18 hours}
\actions{Inscribe symbol of power}
The Adept must have a Magical Aptitude greater than that of the curse.
They must inscribe a triangle or symbol of power about the cursed
being, and perform this ritual for eighteen consecutive hours.  If the
major curse is a death-curse, add only one times the Rank.

\textbf{Precious Metals}

The use of triangles or symbols of power fashioned of varying amounts
of precious metals causes an addition to the success percentage, per
the following schedule:

\begin{tabularx}{\linewidth}{Xcr}
\textbf{Metal} & \textbf{Add} & \textbf{Cost} \\
Silver      & +  3 &  1,000 sp \\
Gold        & +  7 & 10,000 sp \\
Platinum    & + 10 & 15,000 sp \\
Truesilver  & + 15 & 20,000 sp \\
\end{tabularx}

The symbol necessary for this ritual is large enough for the target to
sit in, and is inscribed in the ground.  This symbol may be portable.
\end{effects}
\end{ritual}

\raggedcolumns
\clearemptydoublepage
\end{multicols}
