\section*{THE ADVENTURERS GUILD: ADVENTURING AGREEMENT}

\fontspec{TeX Gyre Termes}\fontsize{8}{9pt}\selectfont

{\bfseries
THIS AGREEMENT made \hspace{15em} this day of  \hspace{15em}  A.P.

BETWEEN the undersigned, hereinafter called collectively ``The Party'', and severally ``Party Member''.

PARTICULARS:}

\begin{enumerate}
\item
Definitions
\begin{enumerate}
\item
``Treasure'' means any monies or goods gained while actively pursuing the ``the Mission''.
\item
``The Mission'' means the activities performed under Guild auspices.
\end{enumerate}

\item
Each Party Member is entitled to one full share of any treasure gained
by the Party.

\item
Each Party Member who dies, is captured, or is incapacitated but who
is subsequently restored is entitled to a full share.

\item
The heirs of a Party Member who dies, is captured, or is incapacitated
and is not restored are entitled to shares of any treasure gained
before the relevant misfortune.

\item
Each party Member may, at the beginning of the adventure, state one
item desired should the Party find it, and will received priority of
choice in the event of a dispute. {\tiny Please list overleaf and
inital}.

\item
Treasure must be converted to currency or specie and be divided evenly
amoungst all shares.  Any Party Member may choose to buy a particular
item at value rather than have it sold to an outside purchaser.  If no
general agreement can be reached as to which member may buy a given
item then a Guild approved random method will be implemented subject
to special claims as per cluase 5. {\item See beolow for a description
of the approved random method.}

\item
Any of the Party Leader, Military Scientist, or Scribe may be removed
from that post by a majority vote of the rest of the party.

\item
The Party may never be without a Scribe.
\end{enumerate}

The Leader of the Party is (required):

The Military Scientist of the Party is (optional):

The Scribe of the Party is (required by the Guild):

{\bfseries
DECLARATION:

I/We agree to abide by the terms and conditons specified herin, and
agree that in all cases the Adventurers Guild of Seagate are the final
arbiters.}

Signatures:

\vspace{14em}

Witnessed by:

\vspace{4em}

Signed for the Guild:

\vspace{4em}

Mission concluded on:

{\tiny Approved Random Method: each candidate rolls percentile dice
--- the order of choice is dterminated by this roll, lowest to
highest, until every player who wishes to choose, has chosen.
The sequence reverses for the next choice.  This process continues
until all items are gone, or there are no more purchasers.

Modification/Extension: This document may be modified or extended
prior to signing by a majority agreement (other than Particular 8).
Pleas write modifications/extensions overleaft and initial.}
