\section{THE ADVENTURERS GUILD: MEMBERSHIP AGREEMENT}

\bigskip

{\fontspec{TeX Gyre Termes}\fontsize{8}{9pt}\selectfont

{\bfseries
THIS AGREEMENT made this \hspace{15em} day of \hspace{16em} A.P.

BETWEEN the Adventures Guild of Seagate, hereinafter called "The Guild"

\vspace{5mm}

\hspace{5mm}   AND  \hspace{5mm}        (name)

\vspace{5mm}

\hspace{5mm}     (sex)
\hspace{5mm}     (status)  \hspace{5mm}   (college)

\vspace{5mm}

\hspace{5mm}   hereinafter called "The Member".

\vspace{5mm}


PARTICULARS:}

\begin{enumerate}

\item    Definitions.
\begin{enumerate}
\item "Treasure" means any monies or goods gained while actively pursuing "the Mission".
\item "The Mission" means the activities performed under Guild auspices.
\item "The Party" means a group with whom the Member is on a Mission.
\end{enumerate}

\item \begin{enumerate}
\item  The Guild does not condone activities such as killing, theft,
extortion, etc. These activities (and others) are illegal in most
cultures we interact with. Any crimes committed against a society will
be answerable to that society. The Guild will not stand in the way of
the normal course of justice.
\item The Guild cannot supplant or override the laws of the land.
\item  All members are subject to the laws of the Duchy of Carzarla.
\end{enumerate}

\item    THE GUILD RULES:

The Member is subject to Guild justice if they:

\begin{enumerate}
\item  attack another member without reasonable provocation,
\item  kill another member,
\item  steal from another member, including the withholding of treasure,
\item  otherwise harm another member, including the use of magical coercion or control,
\item fail to comply with an "Adventuring Agreement" provided that such is
currently in effect and binding on the Member,
\item fail to make all reasonable attempts to regain, if necessary, and
return another Party Member who is dead, incapacitated and/or
captured, to the Guild Headquarters in Seagate,
\item  desert the Party voluntarily; if accidentally separated from the
Party the Member is obliged to make all reasonable attempts to rejoin
the Party,
\item  fail to follow instructions given by a duly appointed Leader of
the Party of which the Member is part.
\end{enumerate}
\item Guild justice may include: fines, forfeiture of part or whole of
shares of Treasure, suspension, expulsion, offering of a reward for
capture, turning over to Civil Authorities and/or the bringing of a
civil case.
\item 10\% of all Treasure gained by the Member or 200sp per annum,
whichever is greater, is due to the Guild.
\end{enumerate}
\vspace{5mm}

{\bfseries
DECLARATION:
\vspace{5mm}

I agree to abide by the rules of the Guild as above, and as may be,
from time to time, laid down by the council, and to accept their
rulings in matters of arbitration and justice.}

\vspace{5mm}

Signed

\vspace{15mm}

(by the Member)


\vspace{5mm}

(for and on behalf of the Guild)
}

\pagebreak
