\documentclass[a4paper]{article}
\usepackage{dq}
\begin{document}
\section{Arcane Magics (Version 2.0)}
\begin{multicols}{2}

\subsection{Introduction}

Arcane Magics are concerned with extending magic beyond the current
limits.  It does not contain any spells and rituals itself, but
enables to enchance the magic the Adept current has and gain new
talents, spells or rituals from other colleges in due time.  Arcane
Magics are orthogonal to the current colleges of magic and
complementary to them.

\subsubsection{Power}

Adepts of this college can gain Power through the course of
adventuring and spent Power as needed.  The Adept gains a a pool of
Power which can be used at any time.  The pool can be expanded by one
point per three month adventure (in the same way as characteristic
points) at the cost of 4000 EP per point.

Power can be used to tempararily enchance abilities or avoid magic,
damage, etc.\ or it can be permanantly spend to gain new abilities.
Currently the college is fairly limited in the range of things Power
can be spend on, but this will grow as Adepts advance.

\subsection{Restrictions}

This college \emph{cannot} be taught any anybody who was a pool of
less than 100 points, and then only if the Adept has the permission of
their teacher.

\subsection{Spending and recovering points}

Points can either be spend temporarily for limited effects and
permanently for more powerful effects.  Permanently spend points
can be bought back for 2000 EP each with temporarily lost points
recovering at one point per sunrise.

\subsection{Abilities}

\subsubsection{Lesser abilities (temporary loss)}

\begin{Description}

\item[Avoid damage] Each point of power can be used to ignore D0
points of damage, or to instantly heal a character of D0 points of
damage (EN then FT - a 10 is a roll-up).  Can be done anytime as a
free action.

\item[Bodily Sacrifice] By maneuvering in combat, a character may
alter the position they take a Specific Grievous taken by moving the
dice roll up or down by a number of points equal to the power they
spend (they cannot move the value outside the range of the weapon).

\item[Multiple Rolls] The character may spend a number of power in
order to roll multiple times and choose the best one.  It requires
\emph{n} power for the \emph{n}th extra roll cumulatively (i.e.\ 1
extra roll costs 1, 2 extra rolls cost 3 etc.) --- must be decided
before the dice is rolled.

\item[Self Sacrifice] A character may choose to dive into an adjacent
hex, pushing aside the individual residing in that hex, and taking any
damage (magical or mundane) that that individual would have taken upon
themselves.  Costs 2 power.

\item[Burst of Speed] The character may move 1 additional hex of
movement for each power spent in that action.

\item[Combined action] For each power the character spends, they make
make a pass action for free, except that the pass action may not
involve movement.  If the character wishes to move, they should use
the ``Burst of Speed'' ability listed above.  (e.g.\ 2 points allow a
character to get a potion out and drink it in addition to their normal
actions).

\item[Enhance Quality] A character may subtract 1 from the dice roll
for each point of power spent.  This may be done after the dice is
rolled.

\item[Enhance Ability] For every 2 points spent a character may
increase the rank of a spell, weapon or ability by one for a single
action.

\end{Description}

The above abilities may be used at any time and may be used
together. Example: A character could be on no fatigue, then get out
and drink a potion (2 points), then upon seeing another party member
getting attacked with Dragon Flames could move 4 hexes (4 points) and
dive in the way taking the damage themselves (2 points).  This could
all be done as well as their normal action.  Total cost is 8
power. (i.e.\ likely to occur once per combat)


\subsubsection{Greater Abilities (permanant loss)}

\begin{Description}

\item[Avoid magic] A character may avoid the effect of any magic.  It
costs 2 points if they failed to save vs the effect, but only 1 point
if they already saved (i.e.\ 1 point to avoid a 1/2 effect).

\item[Regrowth] The character may ignore the effect of 1 Specific
Grievous injury.  Cost 1 point.

\item[Rebirth] The character is fully healed to maximum Endurance and
Fatigue.  Cost: 1 point, requires a pass action to do --- calling upon
inner strength etc.  Does not cure Specific Grievous injuries.

\item[Miraculous Success] The character may choose to make an
exceptional success at any time.  Cost 1 point to double effect
(endurance), 2 points for doubling different component (i.e.\ can
double range and double damage but not triple damage).

\item[Second Chance]  The character may reroll a single die roll if they
choose to do so.  Cost 1 point.

\end{Description}

\raggedcolumns
\newpage
\end{multicols}
\end{document}
