\documentclass{article}

\usepackage{ep}

\newcommand{\signature}{
\tcblower\centering{Five Woks: Clay  --- 1/11/803 WK -- 1/12/803 WK --- Ross Alexander}
}

\begin{document}

\begin{ep-box}{Small Box}

This box is $1' \times 0.5' \times 3''$ in dimension and has been
enchanted with a minor preservation.  Any formally living material
(e.g.\ a book) placed in the box will not decay.  It will not preserve
any possible magical or quasi magical properties of the object.  The
box itself is made from teak and is in a plain leather binding.

\emph{Value: 1500 silver pennies}
\signature
\end{ep-box}

\begin{ep-box}{Manual of Good Health Through Suffering}

This is a diatripe of fantical exercise, body nazism, puritanical
living and regimes of binge and purge is of no apparent use at all,
except possibly for up and coming paladins.  The paper and the binding
are its only real value.

\emph{Value: 20 silver pennies}
\signature
\end{ep-box}

\begin{ep-box}{Players of Games}

These two figures of unidentifiable composition are antimates with the
generic true name of ``Player of Games''.  They are also highly
magical and interact with their surrounding, changing their shape,
colour and aspect in apparently inexplicable ways.  They live off
magic and will obsorb any magic cast at them.  Their texture is of
hard rubber coated in talcum powder.  Sufficient magic cast at them
will cause them to disintegrate, but the exact amount and result is
unknown.  They do not seem to serve any useful purpose.

\emph{Value: 1500 silver pennies}
\signature
\end{ep-box}


\begin{ep-box}{Wakizashi}

This weapon acts as if it is an shortsword for ranking etc.  It is not
particularly ornate but has been weaponsmithed to give it a strike
chance of 50\% and 4 damage.  The steel has been finely tempered and
can be overstrengthed up to PS 20 without increasing the break chance.

\emph{Value: 150 silver pennies}
\signature
\end{ep-box}


\begin{ep-box}{Tripod mounted crossbow}

This is a semi portable field weapon. It has a strike chance of 45\%
and does [D + 9] damage, but is treated as a heavy crossbow to
calculate total BC. Loading the weapon requires two people of at least
20 PS each five pulses to crank.

It weighs 30lbs and 5' wide and 5' long. A giant could wield it
without the tripod, but the strike chance would be reduced to 30\%.

\emph{Value: 400 silver pennies}
\end{ep-box}


\begin{ep-box}{Ash staff}

This is a plain ash quarterstaff.

\emph{Value: 20 silver pennies}
\signature
\end{ep-box}


\begin{ep-box}{Light teak staff}

This ash quarterstaff has been carefully selected from woods and is
much lighter than an ordiary staff.  It weights 1lb and requires 8 PS
to wield, with a strike chance of 60\% doing +3 damage.

\emph{Value: 600 silver pennies}
\signature
\end{ep-box}


\begin{ep-box}{Ceramic staff}

This quarterstaff is made from some type of ceramic and has some minor
magical binding to make it as strong as steel.  Otherwise it just an
ordinary quarterstaff.

\emph{Value: 150 silver pennies}
\signature
\end{ep-box}


\begin{ep-box}{Dwarven Battleaxe}

This battleaxe is finely crafted and is enruned with banes against
goblins, orcs, drow, elves, dragons and other noted enemies of the
dwarven race.  It is enchanted with a permanant enchant weapon spell
(Rank 5) giving a strike chance of 70\% and +6 damage.

\emph{Value: 1300 silver pennies}
\signature
\end{ep-box}


\begin{ep-box}{Wyvern Skin}

This skin is not in good quality giving a base protection of 4 points.
There is enough for a set of armour for a medium sized figure and
could be armourered up to 6 points of protection.

\emph{Value: 200 silver pennies}
\signature
\end{ep-box}
\end{document}
