\documentclass[a4paper]{article}

\usepackage[libertine]{dq}

\begin{document}

\begin{skill}{Celestial Astrology}{1.0}{astrologer}

The celestial bodies have a definite, if not entirely understood,
effect upon the lives of the inhabitants of a DragonQuest world.
These Great Powers seem to impose predestination upon all but the
strong-willed, and determine the aspect of each being.  The Sun, the
Moon, and the Planets regularly cause perturbations in the flow of
mana; the mighty Stars affect a world across the vast reaches of space
by their positions relative to it.  The study of the purpose and
method of the Powers is the science of astrology.

An astrologer's main talent is a limited ability to predict and shape
the future.  An astrologer will be able to make clear, general
assertions, but will only be able to give obscure clues when asked for
specific details.

The Astrology does not need to make their prediction immediately after
studying the stars but must do so before next the midday.

All Astrology base chances gain +1/Rank in astronomy and +1/Rank in
navigator.

\subsection{Restrictions}

An astrologer must be able to read and write in one language at Rank
8.

\subsubsection{An astrologer may only try once to answer a particular
question or to forecast the outcome of an event.}

Once an astrologer has made a reading (\ie a determination about the
future), they may not seek to change or influence the reading through
their art. Other astrologers who attempt to read the same future will
receive the same information that the first astrologer did.  A second
astrologer may, however, receive some clarification about the first's
reading.

\subsubsection{The results of a reading will affect the pertinent course
of events.}

The GM is expected to modify the outcome of an adventure or happening
in their world to conform with a determination made by an astrologer or
by an astrologer at the behest of a character.  The determination does
not preclude the characters' actions from affecting the outcome of the
adventure or event: to the contrary, the GM must interpret the reading
as they see fit, and alter a few of the random dice-rolls engendered
by the characters' actions accordingly.

A prophecy cannot be avoided by the affected character(s) changing
they plans.  The doom (which may be good) will follow them to the
undertaking they substitute for that which was predicted. However, if
a character asks a specific question predicated upon a given action,
the prophecy will not come to pass unless and until that action is
taken.

\subsection{General Predictions}

The astrologer spends an hour observing the night's sky and another
hour determining the future from their star charts.  This reading will
give the subject some insight into a future event to their benefit.
This insight cannot never be subject to prediction (i.e.\ the player
cannot know beforehand whether it is accurate or not).

The subject (or possibly the GM) will at some point then declare the
insight refers to a current action.  The GM will then roll percentile
dice.  If the roll is less than or equal to (1 \x PC + 5 \x Rank) then
the astrologer can reduce their dice roll by the Astrologer's Rank.
Should the roll be greater than (1 \x PC + 5 \x Rank + 30) then the
insight has mislead them and they must add the Astrologer's Rank to
the roll.

An astrologer can only make a one general prediction on an individual
per season and a maximum of Rank general perdictions in total per
season.

\subsubsection{Specific Questions}

The Astrologer may ask the stars a specific question, including
possible future events, current happenings, magical attributes, or
past events.  This requires an hour studying the stars and a later
hour interpreting the results.

When a being poses a specific question to an astrologer willing to
attempt an answer, the GM rolls percentile dice.  If the roll is equal
to or less than (1 \x PC + 6 \x Rank), the astrologer is able to give
a correct answer. If the roll is greater than the success percentage,
they mutter meaningless gibberish.

All answers given to specific questions must be, at the very least,
obscure.  The GM may respond with cryptic poetry, much like the Oracle
at Delphi, or may choose to have the astrologer supply a riddle
(though the Player of the astrologer does not know the answer
themself).

An astrologer may seek to answer up to Rank specific questions per
month.

\begin{example}
\begin{Itemize}
\item The location of an individual.
\item If an individual has had a particular spell cast upon them with the last month.
\item The primary magical effect of an item.
\item Where an individual will be in a week's time.
\end{Itemize}
\end{example}

\subsection{Divine Aspect}

The Astrologer can read the stars and divine the aspect of an entity
they have encountered.  The stars will reveil their star sign and
aspect.  This requires the Astrologer to study the stars for one hour.

\end{skill}
\end{document}
