\begin{multicols}{2}

\setlength\columnseprule{0.2mm}

\section{Aquatics}
Acquatics include only species described in this Section, though other
species may have the ability to swim or breathe water. All aquatics
except Dolphins and Whales can breathe water indefinitely.  Dolphins
and Whales can remain submerged for 5 to 20 minutes only before having
to rise to the surface for air.  Fish must move at least once per
minute or take [D - 3] damage to Fatigue due to suffocation.  All
aquatics move within the water the same way that avians move through
the air and the rules for range and combat applying to avians also
apply to combat involving aquatics or others in an acquatic
environment.

\subsection{Fish}
Fish include barracuda, manta rays, pike, piranha and sharks. All fish
except piranha are found in an oceanic environment.  Piranha tend to
be found in streams, rivers, marshes and lakes. Fish must keep a
steady flow of water over their gills in order to breathe. They are
non-intelligent, but very hostile.

\subsubsection{Barracuda}

\begin{description}
\item[Natural Habitat] Ocean Uncommon

\item[Number] 1-6 (1) Barracuda are predatory fish about 6 feet long.

\item[Talents, Skills and Magic] Barracuda possess no magical abilities or other special
skills or talents.

\item[Weapons] Barracuda attack with their bite (Base Chance of 50\%,
[D + 1] Damage).

\item[Movement Rates] Swimming: 350

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  12-14
& 
\textbf{MD:}  None
& 
\textbf{AG:}  17-20
& 
\textbf{MA:}  None
\\
\textbf{EN:}  6-8
& 
\textbf{FT:}  11-14
& 
\textbf{WP:}  8-10
& 
\textbf{PC:}  14-16
\\
\textbf{PB:}  4-7
& 
\textbf{TMR:}  7
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Barracuda will not generally attack man unless threatened
or particularly hungry. They might, however, mistake a person carrying
jewelry or making quick movements for another fish and attack for that
reason.

\end{description}

\subsubsection{Manta ray}

\begin{description}
\item[Natural Habitat]  Ocean Uncommon

\item[Number] 1-100 (15)
Manta rays are large, flat ocean creatures. Their bellies are white,
while the top is dark blue. They are triangular in shape, and reach
sizes more than 20 feet across, weighing up to 3000 pounds. The eyes
are set forward in front of the body on two foot long protrubcrances,
between which rests the mouth. They have a long, very thin tail.

\item[Talents, Skills and Magic] Manta Rays possess no special talents, skills or magic.

\item[Weapons] Manta Rays can deliver killing blows with their massive fins
in Melee Combat, crushing the ribcages of swimmers, and breaking in
the sides of boats. Base Chance for the Manta: 65\%, [D + 14]
Damage.

\item[Movement Rates]  Swimming: 200

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  35-40
& 
\textbf{MD:}  None
& 
\textbf{AG:}  13-15
& 
\textbf{MA:}  None
\\
\textbf{EN:}  16-19
& 
\textbf{FT:}  24-28
& 
\textbf{WP:}  10-12
& 
\textbf{PC:}  12-14
\\
\textbf{PB:}  5-8
& 
\textbf{TMR:}  4
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin abosrbs 4 DP
}\\
\end{tabularx}

\subsubsection{Pike}

\begin{description}
\item[Natural Habitat] Ponds, Lakes, Rivers, Streams, Marsh Common

\item[Number] 1-30 (1)
 Pike are 2 to 4 foot long fish which specialize in the
solitary hunting of other fish. They have needle-like teeth and spiky
fins and are covered with tough scales. Though usually not a danger to
humans, they can be called in large numbers by nixies and will attack
at the nixie's command.

\item[Talents, Skills and Magic] Pike have no skills, magic, or talents.

\item[Weapons] Pike attack in Close Combat with a Base Chance of 30\%of doing [D - 4] Damage. In schools, each group of 5 attacks with a
Base Chance of 70\% of doing [D + 2] Damage.

\item[Movement Rates]  Swimming: 300

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  8-11
& 
\textbf{MD:}  None
& 
\textbf{AG:}  20-25
& 
\textbf{MA:}  None
\\
\textbf{EN:}  5-10
& 
\textbf{FT:}  10-15
& 
\textbf{WP:}  6-8
& 
\textbf{PC:}  14-16
\\
\textbf{PB:}  5-6
& 
\textbf{TMR:}  6
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Scales absorb 1 DP 
}\\
\end{tabularx}

\subsubsection{Piranha}

\begin{description}
\item[Natural Habitat] River Rare

\item[Number] 1-200 (50)
 Piranha are small rish with very sharp teeth and a
vicious temper.

\item[Talents, Skills and Magic] Piranha have no magic abilities, talents or skills.

\item[Weapons] Piranha attack in schools. Every group of ten gets one
attack (Base Chance of 80\%, [D + 4] Damage). They can only
attack in Close Combat.

\item[Movement Rates] Swimming: 300

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  1-2
& 
\textbf{MD:}  None
& 
\textbf{AG:}  20-23
& 
\textbf{MA:}  None
\\
\textbf{EN:}  1
& 
\textbf{FT:}  1-2 
& 
\textbf{WP:}  4-5
& 
\textbf{PC:}  15-16
\\
\textbf{PB:}  5-6
& 
\textbf{TMR:}  6
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Piranha are voracious. They can devour an elephant in
minutes, and will immediately attack anything edible (primarily
meat-bearing) which falls into the water with them.

\end{description}

\subsubsection{Shark}

\begin{description}
\item[Natural Habitat] Ocean Uncommon

\item[Number]  1-50 (5)
 Sharks are long fish with very sharp teeth and tough skin. A
shark's colour varies from light blue to white to orange, depending on
the waters that they are swimming in.  Sharks reach sizes in excess of
40 feet, at which point they are threats to boats as well as to
swimmers.  Sharks are normally two-hex monsters.

\item[Talents, Skills and Magic] Sharks can detect the presence of blood in the water from
extremely long distances, and they will gather around the blood,
working themselves up into a killing frenzy. They can also sense
motion in water, especially in the form of the types of disturbances
caused by swimmers.

\item[Weapons] Sharks attack via their bite (Base Chance of 40\%, [D +
8] Damage).

\item[Movement Rates]  Swimming: 350

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  35-45
& 
\textbf{MD:}  None
& 
\textbf{AG:}  12-16
& 
\textbf{MA:}  None
\\
\textbf{EN:}  25-30
& 
\textbf{FT:}  40-50  
& 
\textbf{WP:}  8-10
& 
\textbf{PC:}  9-12
\\
\textbf{PB:}  6-8
& 
\textbf{TMR:}  7
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 7 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] As soon as blood is drawn in waters containing sharks (most
waters do) there is a 60\% chance that 10-60 sharks will be
attracted. Sharks teeth are fairly valuable (a set might go for 100
Silver Pennies) as charms.

\end{description}

\subsection{Aquatic Mammals}
Aquatic mammals include dolphins, great white whales killer whales,
and merfolk.  They are generally moderately intelligent and neutral,
or even friendly, to adventurers.  Except for merfolk, they cannot
remain submerged indefinitely.  Merfolk must keep a steady flow of
water over their gills in the same manner as fish.

\subsubsection{Dolphin}

\begin{description}
\item[Natural Habitat] Ocean Common

\item[Number]  1-20 (4)
 Dolphins are mammals of the sea. They grow to 12 feet in length
and are basically fish-like in form (although their bodies are more
cylindrical than most fish). They have long, thin snouts, and are
generally greyish in color. Dolphins are two-hex monsters.

\item[Talents, Skills and Magic] Dolphins have no talents, skills or magic.

\item[Weapons] Dolphins attack in Melee Combat by ramming with their snouts
with a Base Chance of 40\% and [D - 1] Damage.

\item[Movement Rates]  Swimming: 400

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  25-30
& 
\textbf{MD:}  None
& 
\textbf{AG:}  24-26
& 
\textbf{MA:}  None
\\
\textbf{EN:}  12-15
& 
\textbf{FT:}  20-25
& 
\textbf{WP:}  15-17
& 
\textbf{PC:}  14-17
\\
\textbf{PB:}  12-15
& 
\textbf{TMR:}  8
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Dolphins will sometimes aid humans in need, and will
remember with a vengeance those humans who have harmed them in the
past. They will never attack humans unless provoked.

\end{description}

\subsubsection{Great white whale}

\begin{description}
\item[Natural Habitat]Ocean Very Rare

\item[Number] 1
 These creatures are the largest mammals found on sea or
on land. They reach lengths of 200-300 feet, and are a truly
spectacular sight when they rise three-quarters of the way out of
the water, sending spray hundreds of yards in all directions. These
whales are pure white in color.

\item[Talents, Skills and Magic] Great white whales cannot use magic and have no special
talents or skills.

\item[Weapons] Rather than doing actual damage to most creatures, the great
white whale simply swallows them alive (Base Chance 50\%). If a
creature is swallowed alive, it is presumed dead unless it is somehow
magically saved. If a creature is large enough so that it cannot be
swallowed by the creature's 30 foot wide gullet, then the whale does
[D + 10] Damage. The great white whale can only attack in Melee
Combat.


\item[Movement Rates]  Swimming: 700

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  700-1000
& 
\textbf{MD:}  None
& 
\textbf{AG:}  8-9
& 
\textbf{MA:}  None
\\
\textbf{EN:}  300-400
& 
\textbf{FT:}  500-600
& 
\textbf{WP:}  8-10
& 
\textbf{PC:}  12-15
\\
\textbf{PB:}  4-6
& 
\textbf{TMR:}  14
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 8 DP
}\\
\end{tabularx}

\subsubsection{Killer whale}

\begin{description}
\item[Natural Habitat] Ocean (generally arctic waters) Rare

\item[Number] 1-25 (20)
 Killer whales are black on top and white on their
bellies. They have fins on either side of their body, as well as one
fin which sticks up from their back much like the fin of a
shark. Large males reach 30 feet in length. Killer whales are 14-hex
monsters.

\item[Talents, Skills and Magic] Killer whales have no special talents, skills, or magic.

\item[Weapons] Killer whales can attack either in Melee or Close Combat
with their bite (Base Chance of 50\%, [D + 10] Damage).

\item[Movement Rates] Swimming: 400

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  40-45
& 
\textbf{MD:}  None
& 
\textbf{AG:}  13-15
& 
\textbf{MA:}  None   
\\
\textbf{EN:}  35-40
& 
\textbf{FT:}  60-70  
& 
\textbf{WP:}  12-13
& 
\textbf{PC:}  15-17
\\
\textbf{PB:}  10-12
& 
\textbf{TMR:}  8
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 3 DP
}\\
\end{tabularx}

\subsubsection{Merfolk (sea people)}

\begin{description}
\item[Natural Habitat]  Ocean Rare

\item[Number] 1-100 (1)
 Merfolk are half human, half fish, the top half being
the human half. Merfolk lose 3 points of Fatigue (and Endurance when
Fatigue is exhausted) for each hour they remain out of water.

\item[Talents, Skills and Magic] Merfolk get along very well with the indiginous life of the
sea, and they can summon fish of various types, although they will
only rarely summon the larger predators. Some merfolk will have
powerful magic, while others will be ignorant of magic. All breathe
both air and water.

\item[Weapons] The merfolk ride seahorses and can use weapons underwater.
They prefer pole-type weapons (especially tridents) and stabbing
knives and swords. They will wear fish-skin armor (similar to leather
armour) or laminated seashell armour (similar to lamillar armour) when
going to war.

\item[Movement Rates] Swimming: 350

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  10-25
& 
\textbf{MD:}  14-25
& 
\textbf{AG:}  16-21
& 
\textbf{MA:}  13-23
\\
\textbf{EN:}  12-14
& 
\textbf{FT:}  20-24
& 
\textbf{WP:}  13-18
& 
\textbf{PC:}  14-20
\\
\textbf{PB:}  18-30
& 
\textbf{TMR:}  7
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Merfolk will often sit on a beach, waiting to ensnare a
human of the opposite sex, whom they will then bring down into the
deeps with them as lovers. They give the human a cap woven of gold
which has magical properties which allow the human to live and breath
under water. Down under the sea, the merfolk will have cities and
societies much like those on the surface, with wars, diplomacy, trade,
etc. Great treasure will often be found in the sea-peoples' underwater
lairs, often including gold from sunken wrecks. The merfolk are not
usually hostile to man, but they will occasionally force a boat to
turn back if it is tresspassing on water they consider sacred or
otherwise not appropriate for man.

\end{description}

\subsection{Others}
This section includes eels, kraken, octopi, and squids. These
creatures are often intelligent and usually hostile to adventurers.

\subsubsection{Eel}

\begin{description}
\item[Natural Habitat] Ocean Uncommon

\item[Number] 1-4 (1) Eels are long, thin, grayish fish up to 7 feet in
length.

\item[Talents, Skills and Magic] Eeels have no special talents, skills, or magic.

\item[Weapons] An eel attacks with its bite (Base Chance of 50\%, [D +
2] Damage), which it can only use in Close Combat.


\item[Movement Rates]  Swimming: 300

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  18-20
& 
\textbf{MD:}  None
& 
\textbf{AG:}  14-17
& 
\textbf{MA:}  None
\\
\textbf{EN:}  8-10
& 
\textbf{FT:}  14-16  
& 
\textbf{WP:}  10-11
& 
\textbf{PC:}  17-20
\\
\textbf{PB:}  6-8
& 
\textbf{TMR:}  6
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 3 DP
}\\
\end{tabularx}

\subsubsection{Kraken}

\begin{description}
\item[Natural Habitat] Ocean Very Rare

\item[Number] 1-2 (1)
 Kraken are huge, black, fierce octupoids. They have a pointed
beak and large, red eyes. A kraken's body is 30-40 feet long, and its
8 tentacles are about as long as its body.  Their hide is thick and
horny and they are 14-hex monsters.

\item[Talents, Skills and Magic] The kraken's most fearsome attribute is its ability to rip
ships apart with its hideously strong tentacles. It can usually handle
ships up to 40 feet in length, shredding and then entirely consuming
them and their contents within (D + 2) minutes after it emerged from
the depths. In addition, a kraken can use the maelstrom spell from the
College of Water Magics at rank 8-10.

\item[Weapons] In any given Pulse a kraken can attack with up to 5 of its 8
tentacles. It has two options as to what it can do with these
tentacles: it can either try to damage a character with the tentacles
themselves at a Base Chance of 70\%, and Damage of [D + 4], or it
can attempt to draw the adversary towards its beak at a Base Chance of
30\%, and [D + 15] Damage. Either option can be taken with any of
the tentacles.


\item[Movement Rates]  Swimming: 600

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  100-200
& 
\textbf{MD:}  15-20
& 
\textbf{AG:}  12-14
& 
\textbf{MA:}  12-15
\\
\textbf{EN:}  60-80
& 
\textbf{FT:}  80-100
& 
\textbf{WP:}  25-30
& 
\textbf{PC:}  17-20
\\
\textbf{PB:}  1-3
& 
\textbf{TMR:}  12
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 6 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Krakens typically live by the shores of large, saltwater
bodies, pulling swimmers and small vessels down to their demise. If a
kraken's lair is found, it will frequently (80\%) have a large
treasure (25,000-35,000 Silver Pennies in value) stockpiled. Krakens
tend to be unintelligent, but grow larger and more intelligent in cold
water.

\end{description}

\subsubsection{Octopus}

\begin{description}
\item[Natural Habitat] Ocean Uncommon

\item[Number]  1-6 (1)
 Octupi are small bags of flesh with eight long thin
tentacles attached. They rarely grow to larger than 6 feet, from
tentacle-tip to tentacle-tip. Their color shifts with their
surroundings to provide the maximum camouflage. The body is highly
maleable, and can usually fit into small cracks in coral beds. Rare
species of giant octopi may grow to 3 times normal size.

\item[Talents, Skills and Magic] When frightened, octopi can squirt a jet of dark substance
into the water, obscuring vision while they make their escape.

\item[Weapons] In Close Combat, an octopus can attack once using its
tentacles (Base Chance of 80\%, - 2 Damage).

\item[Movement Rates] Swimming: 150; Climbing on Rocks: 75

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  8-12
& 
\textbf{MD:}  15-17
& 
\textbf{AG:}  14-16
& 
\textbf{MA:}  None
\\
\textbf{EN:}  5-6
& 
\textbf{FT:}  10-11
& 
\textbf{WP:}  6-8
& 
\textbf{PC:}  17-19
\\
\textbf{PB:}  5-7
& 
\textbf{TMR:}  3/1
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 2 DP
}\\
\end{tabularx}

\subsubsection{Squid}

\begin{description}
\item[Natural Habitat] Ocean Rare

\item[Number]1-10 (1)
 These creatures are bullet-shaped, with 10 tentacles streaming
from the base of the bullet. They grow to 50 feet in length, and their
color changes according to the waters that they inhabit. The Squid
also has a beak and large, round eyes. Squids are 14-hex monsters.

\item[Talents, Skills and Magic] The squid has no magic or special skills or talents.

\item[Weapons] Squids attack with their tentacles, drawing their prey in
toward their beak. A squid can make up to 5 attacks each pulse using
its tentacles, but the first time that it hits something, it must
cease attacks for that pulse as it can only bite one thing at a
time. The Base Chance for each attack is 25\%, with [D + 12]
Damage. The squid can also choose not to attempt to bite and merely
squeeze with its tentacles, in which case the Base Chance is
50\%, and Damage [D - 1]. It can score up to 5 hits in this way.

\item[Movement Rates]  Swimming: 600

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  60-70
& 
\textbf{MD:}  25-30
& 
\textbf{AG:}  10-13
& 
\textbf{MA:}  None
\\
\textbf{EN:}  50-60
& 
\textbf{FT:}  65-70
& 
\textbf{WP:}  10-12
& 
\textbf{PC:}  10-12
\\
\textbf{PB:}  6-8
& 
\textbf{TMR:}  12
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 5 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Large squid generally prefer very deep water, and will
rarely be found at the shallower depths. Squid are carnivorous, and
will eat almost any meat (including man) that they can find.

\end{description}
\end{multicols}
