\begin{multicols*}{2}

\setlength\columnseprule{0.2mm}

\section{Avians}

\subsection{Common Avians}
 Common avians include buzzards, eagles, goshawks, owls.

\subsubsection{Buzzard}

\begin{description}
\item[Natural Habitat] Plains, waste Common

\item[Number] 1-60 (10)

\item[Description] These birds have dark brown plummage and thin,
unfeathered necks of a tannish color. Their beaks are black and
snub-tipped. They prey on the dead and near-dead (espcially if
unconscious).

\item[Talents, Skills and Magic] Buzzards can see each other miles away, and thus when one
buzzard descends for a meal, others soon follow. Except for their keen
eyesight, buzzards possess no special talents, skills or magic. They
are not tool users.

\item[Weapons] A buzzard attacks in Close Combat with its claws and
beak. The claws' Base Chance is 45\%, for [D - 5] Damage, while
the bite's Base Chance is 40\%, for [D - 6] Damage. A buzzard
cannot attack in Melee Combat.

\item[Movement Rates]  Flying: 600

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  3-5
& 
\textbf{MD:}  15-17
& 
\textbf{AG:}  14-16
& 
\textbf{MA:}  None
\\
\textbf{EN:}  6-7
& 
\textbf{FT:}  9-11
& 
\textbf{WP:}  8-10
& 
\textbf{PC:}  17-20
\\
\textbf{PB:}  8-10
& 
\textbf{TMR:}  12
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 1 DP
}\\
\end{tabularx}

\subsubsection{Eagle}

\begin{description}
\item[Natural Habitat] Rough (usually mountainous) uncommon

\item[Number] 1-20 (1)

\item[Description] Great Eagles are the largest of the birds of prey.  They
have a wingspan of 7 feet when full grown. Their feathers are gray
except around neck and head (which is white).

\item[Talents, Skills and Magic] Eagles have no special talents or skills. They are not magic
users and do not use tools. They do have keen eyesight.

\item[Weapons] Eagles can attack only in Close Combat with either beak
(Base Chance of 50\% of doing [D - 1] Damage) or 2 talons (Base
Chance of 70\% of doing [D + 1] Damage) without penalty in the
same Pulse. An eagle may achieve Ranks 1-3 with beak and 1-2 with
talons.

\item[Movement Rates]  Flying: 1200

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  12-16
& 
\textbf{MD:}  22-24
& 
\textbf{AG:}  19-23
& 
\textbf{MA:}  None
\\
\textbf{EN:}  9-12   
& 
\textbf{FT:}  17-20
& 
\textbf{WP:}  11-13
& 
\textbf{PC:}  20-24
\\
\textbf{PB:}  10-12
& 
\textbf{TMR:}  24
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]  Eagles avoid humans except to occasionally steal
goats or other small animals from human flocks. They will
sometimes attack a party which ventures too near their nest.

\end{description}

\subsubsection{Goshawk}

\begin{description}
\item[Natural Habitat]  Woods Rare

\item[Number] 1-2 (1)

\item[Description] Goshawks are the largest of the hawks. They are almost
20 feet in length. The feathers of a goshawk are primarily brown,
although the belly is pure white and some of the tail feathers are
black. The sharp beak and claws are black.

\item[Talents, Skills and Magic] Other than their excellent eyesight, goshawks have no
special talents, skills, or magical abilities and are not tool users.

\item[Weapons] In Close Combat, a Goshawk can attack with two claws and its
beak in the same Pulse without penalty. The claw attacks have a Base
Chance of 80\% and do [D - 3] Damage, while the beak attack has a Base
Chance of 65\% and does [D - 4] Damage. The goshawk cannot attack in Melee
Combat.

\item[Movement Rates] Flying: 1100

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  4-6
& 
\textbf{MD:}  18-20
& 
\textbf{AG:}  20-22
& 
\textbf{MA:}  None
\\
\textbf{EN:}  6-7
& 
\textbf{FT:}  9-11
& 
\textbf{WP:}  8-10
& 
\textbf{PC:}  20-21
\\
\textbf{PB:}  10-12
& 
\textbf{TMR:}  22
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 1 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Like the other birds of prey, the goshawk can be trained to
serve man in captivity. It will bring 700-1200 Silver Pennies as a
Trained hunter. Other types of hunting birds (including boshawks and
gyrfalcons) will have characteristics 1-2 less than the goshawk, will
fly at speeds of 100-200 yards per minute less, and will have a value
of 200-300 fewer Silver Pennies. They will otherwise be similar.

\end{description}

\subsubsection{Owl}

\begin{description}
\item[Natural Habitat] Woods Uncommon

\item[Number]  1-6 (1)

\item[Description]  Owls can be of several types, but in general they
are between one and two feet in length. Coloring varies, but
brown feathers speckled with grey is a common coloration.
Owls' eyes are both on the front of the head, unlike inost birds,
and they tend to have large, dark rings around the eyes.

\item[Talents, Skills and Magic]  Owls see well in the dark and have
excellent hearing. They have no magical abilities, skills or talents
and are not tool users.

\item[Weapons] In Close Combat, Owls can attack with onc bite (Base Chance
of 55\% and Damage [D - 6]) and two claws (Base Chance of
50\% and [D - 7] Damage) in the same Pulse without penalty. Owls
cannot attack in Melee Combat.

\item[Movement Rates]  Flying: 600

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  3-4
& 
\textbf{MD:}  18-20
& 
\textbf{AG:}  22-24
& 
\textbf{MA:}  None
\\
\textbf{EN:}  3-4
& 
\textbf{FT:}  5-7 
& 
\textbf{WP:}  7-9
& 
\textbf{PC:}  20-22
\\
\textbf{PB:}  9-11
& 
\textbf{TMR:}  12
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 1 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Owls are primarily nocturnal predators, hunting small
mammals as well as birds. They are said to understand human speech and
to be friendly to man, often providing travellers with valuable
information.

\end{description}

\subsection{Fantastical Avians}

\subsubsection{Gargoyle}

\begin{description}
\item[Natural Habitat]  Ruins, Caverns, Crypts Rare

\item[Number] Rare 1-10 (2)

\item[Description] Gargoyles are humanoid in form, but they have leathery
wings, sharp claws, a hom, a tail, and spiked elbows and knees. Their
skin is stony in appearance, and upon contact it feels very
rough. Gargoyles have deep-set eyes, and sharp, small teeth.

\item[Talents, Skills and Magic] Gargoyles possess no special skills or talents. They have
only limited magical abilities in general but may use tools and even
weapons.

\item[Weapons] A gargoyle can use his horn, his claws, and his bite in
Melee and Close Combat. In any one Pulse, he may make up to two strike
attempts with any combinations of these weapons.  Base Chance for the
bite is 45\%, [D + 2] Damage; Base Chance for the horn is
35\%, [D + 1] Damage; Base Chance for the claws is 50\%, [D
- 3] Damage. Gargoyles sometimes (but not often) will use weapons
(instead of claws).

\item[Movement Rates]  Flying:400; Running: 200

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  14-20
& 
\textbf{MD:}  13-18
& 
\textbf{AG:}  12-16
& 
\textbf{MA:}  10-12
\\
\textbf{EN:}  7-10
& 
\textbf{FT:}  18-20
& 
\textbf{WP:}  9-13 
& 
\textbf{PC:}  12-16
\\
\textbf{PB:}  2-5
& 
\textbf{TMR:}  8
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}   Hide absorbs 8 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Gargoyles are hopelessly evil, and will attack immediately
any party that they think they have a fair chance of handling. They
can sometimes be found in the service of a highly evil character, in
which case they will usually show more restraint.

\end{description}

\subsubsection{Gryphon}

\begin{description}
\item[Natural Habitat] Rough (especially mountainous) Very Rare

\item[Number] 1-7 (1)

\item[Description] The gryphon is a large beast with the body of a lion, the head
and wings of an eagle and back covered with feathers. The gryphon's
claws are so large that they can be made into serviceable drinking
horns. The gryphon is about 5 feet long and stands man-high when full
grown. Gryphon eggs (2-5 possible per nest) are solid agate and quite
valuable (4000 - 6000 Silver Pennies apiece).

\item[Talents, Skills and Magic] Gryphons have no special skills or magical abilities, but
their special talent is locating buried treasure and their nest will
quite often be built near such treasure which they will actively
guard. A ``domesticated'' gryphon has a Base Chance of 60\% of
locating any buried treasure within a radius of 250 feet of
it. Gryphons make valuable steeds, once tamed.

\item[Weapons] The gryphon is not a toot user, but has natural weapons in
the form of beak (useable in Melee and Close Combat with a Base Chance
of 65\%, [D + 5] Damage, possible Rank of 1-3) and claws (useable
in Melee and Close Combat with a Base Chance of 60\%, Damage of
[D + 5], possible Rank of 1-4). The gryphon may attack twice with
claws and once with beak in the same pulse.


\item[Movement Rates] Running: 200; Flying: 500

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  30-36
& 
\textbf{MD:}  15-20
& 
\textbf{AG:}  15-20
& 
\textbf{MA:}  6-9
\\
\textbf{EN:}  25-30
& 
\textbf{FT:}  30-35
& 
\textbf{WP:}  12-16
& 
\textbf{PC:}  16-22
\\
\textbf{PB:}  0-5
& 
\textbf{TMR:}  4/10
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 6 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Gryphons prefer solitude and will fly away when approached
if possible, but will always chase and attack horses (whose flesh they
prize). They are extremely valuable for use as mounts and cost
accordingly (16,000-20,000 Silver Pennies).  They nest in cliffsides
and build their nests of spun-gold (value 8,000-10,000 Silver
Pennies).

\end{description}

\subsubsection{Harpy}

\begin{description}
\item[Natural Habitat] Rough (lair in caverns or cliffs near water) Rare

\item[Number] 1-20 (6)

\item[Description] Harpies are buzzard-like birds with the upper bodies of
women. They have long claws, and pale, thin faces.  They make a
tremendous amount of noise flying, and have a horrible smell.

\item[Talents, Skills and Magic] Harpies can speak, but have limited magical abilities and no
special skills or talents. They are occasionally prophetic, although
they will only speak a prophecy if it is evil and they wish to torment
the listener.

\item[Weapons] Unless cornered, harpies do not fight man. If they do fight,
they have their long talons (Base Chance of 70\%, [D + 5] Damage)
to strike with, both of which can strike twice in Melee or Close
Combat in the same pulse at no penalty to the harpy.


\item[Movement Rates] Flying: 500; Running: 50

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  14-20
& 
\textbf{MD:}  22-26
& 
\textbf{AG:}  26-30
& 
\textbf{MA:}  13-17
\\
\textbf{EN:}  8-10
& 
\textbf{FT:}  14-19
& 
\textbf{WP:}  17-20
& 
\textbf{PC:}  10-16
\\
\textbf{PB:}  2-6
& 
\textbf{TMR:}  10/1
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 7 DP
}\\
\end{tabularx}

\subsubsection{Hippogriff}

\begin{description}
\item[Natural Habitat] Rough (especially mountains) Very Rare

\item[Number] 1-12 (2)

\item[Description] Hippogriffs are similar to Pegasi in that they are
basically winged horses, but their head is that of some great bird of
prey, with a fearsome beak. The head differs from that of an eagle in
that it has large, feathered ears, almost like wings, at the back of
the head. The front legs of the hippogriff also resemble those of a
bird of prey, with huge talons and feathers.

\item[Talents, Skills and Magic] Hippogriffs have no magical or other extra-ordinary
abilities, talents or skills, and are not tool users.

\item[Weapons] In Melee Combat, the hippogriff can attack with two claws
and one bite in the same pulse with no penalty. The bite has a Base
Chance of 65\% and does [D + 4] Damage, while the claws have a
Base Chance of 55\% and do [D + 2] Damage. In Close Combat, the
hippogriff can still attack with its claws normally, but the Base
Chance on its bite is reduced to 35\%.

\item[Movement Rates]  Running: 200; Flying: 400

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  18-20
& 
\textbf{MD:}  19-22
& 
\textbf{AG:}  16-19
& 
\textbf{MA:}  None
\\
\textbf{EN:}  15-18
& 
\textbf{FT:}  25-30  
& 
\textbf{WP:}  12-17
& 
\textbf{PC:}  17-20
\\
\textbf{PB:}  8-11
& 
\textbf{TMR:}  4/8
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Hippogriff do not like people, and generally shun any area
that is habitated. They are trainable with difficulty.  Their value is
3,000-4,000 Silver Pennies untrained and 3 times that price
trained. Hippogriff eggs are valuable (circa 800 Silve Pennies each)
and very young hippogriffs are even more so (1400-2000 Silver
Pennies).

\end{description}

\subsubsection{Pegasus}

\begin{description}
\item[Natural Habitat] Rough (primarily mountaintops) Very Rare

\item[Number] 1-6 (1)

\item[Description] Pegasi are winged horses with pure white coats tinged
with silver around mane and flowing tail. They are intelligent for
equines and can sometimes speak the common tongue.

\item[Talents, Skills and Magic] Pegasi have no special talents or skills except for their
flying ability. They are not magic or tool users.

\item[Weapons] In battle, a pegasus attacks as a heavy warhorse.

\item[Movement Rates] Running: 300; Flying: 500

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  55-60
& 
\textbf{MD:}  15-17
& 
\textbf{AG:}  15-18
& 
\textbf{MA:}  None
\\
\textbf{EN:}  25-30
& 
\textbf{FT:}  50-60  
& 
\textbf{WP:}  19-23
& 
\textbf{PC:}  16-20
\\
\textbf{PB:}  16-20
& 
\textbf{TMR:}  6/10
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}   Hide absorbs 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Wild pegasi are tameable with difficulty. They will only
allow themselves to be bridled with gold, however. They can carry up
to 300 pounds flying. They are worth, 4,000 to 5,000 Silver Pennies
untamed and 5 times that amount tamed.

\end{description}

\subsubsection{Phoenix}

\begin{description}
\item[Natural Habitat] Woods, Rough, Plains Very Rare

\item[Number] 1

\item[Description] The Phoenix appears to be much like an eagle in profile but is
three times the size of an eagle.  The bird's plummage is partly red
and partly gold, and is dazzlingly beautiful. The Phoenix is a
four-hex monster.

\item[Talents, Skills and Magic] The Phoenix is highly intelligent, although it cannot
speak. It does not have any magical abilities.  The distinguishing
ability of the Phoenix is its ability to reproduce itself (see
``Comments''). Other than that it has no special talents or skills.

\item[Weapons] Phoenix cannot attack except in Close Combat when they can
make three attacks in one Pulse without penalty: one with beak (Base
Chance of 60\%, [D + 8] Damage) and two with talons (Base Chance
of 65\%, [D + 6] Damage).


\item[Movement Rates]  Flying: 2000

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  17-20
& 
\textbf{MD:}  20-24
& 
\textbf{AG:}  17-20
& 
\textbf{MA:}  None
\\
\textbf{EN:}  18-24
& 
\textbf{FT:}  30-35
& 
\textbf{WP:}  20-23
& 
\textbf{PC:}  16-20   
\\
\textbf{PB:}  20-24
& 
\textbf{TMR:}  40
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 6 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Phoenix are famed for their longevity and their method of
reproduction. When an old bird is nearing the end of its days (500
years) it builds a nest at the top of an oak or palm tree, and makes a
pile of cinnamon and spikenard in the nest upon which it lays itself
down to die. A worm springs out of the body of the dead phoenix, and
after the worm has matured, it becomes a phoenix itself. It then bears
the parent bird to the Altar of the Sun, where it is consumed in
flames.

\end{description}

\subsubsection{Roc}

\begin{description}
\item[Natural Habitat] Mountain peaks Rare

\item[Number]  1-2 (1)

\item[Description] Rocs are huge birds with wingspans of up to 250
feet. Except for their size, they much resemble eagles. Their body
size is as a four-hex monster.

\item[Talents, Skills and Magic] Rocs have no special talents, skills or magic, and are not
tool users.

\item[Weapons] Rocs are fearsome creatures in battle. They can engage in
Melee Combat from a range of 3 hexes, and they can take three attacks
in one Pulse without penalty. They attack once with a bite (Base
Chance of 75\%,[D + 20] Damage) and twice with claws (Base Chance
70\%, [D + 25] Damage).


\item[Movement Rates]  Flying: 2000

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  200-300
& 
\textbf{MD:}  20-25
& 
\textbf{AG:}  16-20
& 
\textbf{MA:}  None
\\
\textbf{EN:}  150-200
& 
\textbf{FT:}  250-300  
& 
\textbf{WP:}  16-20
& 
\textbf{PC:}  17-21
\\
\textbf{PB:}  3-7
& 
\textbf{TMR:}  40
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Feathers absorb 8 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Rocs often swoop down from their nests to pick up
elephants, etc.\ to take back to their lair and eat. A roc will
sometimes attempt to steal a party's cart animals or riding animals,
but will rarely carry off the people themselves.

\end{description}
\end{multicols*}
