\begin{multicols}{2}

\setlength\columnseprule{0.2mm}

\section{Demons}
Demons are possessed of certain qualities not listed in their
descriptions.  All demons exercise their skills at Rank 15 and magic
powers at Rank 20.  Demons are not subject to specific Grievous
Injuries.  Instead, they suffer Endurance damage.  They can only be
harmed by magic or by silvered weapons.  Being creatures of mana, they
are not prevented from performing magic by the presence of cold iron.
A demon automatically recovers from all injury by spending one full
day in its own dimension.  Demons may choose to bring along one or
more hellhounds when summoned.

\subsection{Succubi and Incubi}
It is important to understand why Incubi and Succubi appear on this
plane at all in order to successfully GM them in a campaign.  They are
not intended to be the ``companions'' of any Adept.  Their sole
purpose is to increase the number of followers of the Powers of
Darkness on this plane.  They are representatives of this Power, and
as such, they cannot by themselves create anything.  Thus, they must
use mankind (members of both sexes) to aid in their mission.  By their
sexual activity, they seek to create children who are, in reality and
spirit, children of Darkness.

Because of their purpose, an Incubi will never kill the human he has
just mated with.  He will endeavour to keep the nature of his true
being from this woman so the child which he is trying to create will
be born and not aborted.  It is to be assumed that any mating between
Incubi and a human female will automatically result in a pregnancy
unless aborted.  For more detailed information on this subject, a good
reference is The Encyclopedia of Witchcraft and Demonology by Russell
Hope Robbins (New York Crown Publishers, 1959).

\subsection{Dukes}

\subsubsection{AGARES (The Duke of Changes)}

\begin{description}

\item[Base Chance]50%

\item[Lesser Spirits]31%
Agares appears as an aged and enfeebled man riding on a giant land
turtle. A crow perches on his right wrist.  He speaks with a voice
that is brittle with age.

\item[Talents, Skills and Magic]Agares is an adept of the College of Illusions. He possesses the
following skills: Alchemist, Astrologer, Beast Master, Healer,
Merchant, Navigator, Spy. He inspires terror in the enemies (and
sometimes the friends) of the summoner (roll on Fright Table). He is a
linguist and a masterful teacher of languages. He may also cause
earthquakes at will.  He has the power to find those who hide
themselves.

\item[Movement Rates]Running: 50 yards per minute.

\item[Weapons]Agares has no natural weapons. He is treated as a human for
purposes of unarmed combat. He may use weapons and wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 9	
& 
\textbf{MD:} 12	
& 
\textbf{AG:} 11	
& 
\textbf{MA:} 33
\\
\textbf{EN:} 14	
& 
\textbf{FT:} 26	
& 
\textbf{WP:} 35	
& 
\textbf{PC:} 36
\\
\textbf{PB:} 10	
& 
\textbf{TMR:} 1	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]Agares has a 25% chance of locating any individual of
the summoner's choice (regardless of how far away the individual may
be). He cannot, however, locate individuals on other planes of
existence. The individual need not be hiding (he may simply be lost,
for example) in order for Agares to exercise this power.

\end{description}

\subsubsection{AIM (The Fire Duke)}

\begin{description}

\item[Base Chance]63%

\item[Lesser Spirits]26%
 Aim appears as a man with three heads.  One head is
human, the other is that of a serpent, and the third is that of a
calf.  He bears two stars on the forehead of his human head.  In one
hand he carries a ball of fire that is eternally blazing.  He rides a
large lizard with scales of midnight blue.  Wherever he goes, Aim is
surrounded by billowing clouds of red-tinged smoke.

\item[Talents, Skills and Magic] Aim is a master of the College of Fire. He possesses the
following Skills: Alchemist, Military Scientist.  He can set fire to
any combustible object by touching it with the hand wherein resides
his fireball.  The fireball may not be thrown.

\item[Movement Rates] Running: 250 yards per minute.

\item[Weapons] Aim may bite in Close Combat with his non-human heads.  He
has a BC of 45% with his serpent head and a BC of 30%
with his calf head.  The serpent does + 2 damage (plus D-3 per Pulse
additional damage for poison) and the calf head does + 3 damage.  The
serpent's head is Rank 4 and the calf's head is Rank 2.  Aim may not
wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 22
& 
\textbf{MD:} 24
& 
\textbf{AG:} 23
& 
\textbf{MA:} 30
\\
\textbf{EN:} 25
& 
\textbf{FT:} 35
& 
\textbf{WP:} 34
& 
\textbf{PC:} 26
\\
\textbf{PB:} 3
& 
\textbf{TMR:} 5
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Aim delights in all things having to do with fire and will
gladly start a blazing inferno for the joy of watching the fire burn,
even when bound.  He must be watched carefully for this reason.

\end{description}

\subsubsection{ALLOCES (The Warrior Duke)}

\begin{description}

\item[Base Chance]74%

\item[Lesser Spirits] 36%
 Alloces appears as a man with a lion's face and eves made of
hot coals.  He wears burnished armour and rides a great war
horse. Alloces' flesh is like red gold.  His voice is harsh and
booming.  Those who look directly into his eyes see their own death
and are blinded for D + 5 days afterward.

\item[Talents, Skills and Magic] Alloces is a practitioner of the arts of ihe college of
Celestial Magics.  He possesses the following Skills: Assassin,
Navigator, Spy, Thief, Military Scientist.

\item[Movement Rates]Running: 350 yards per minute.

\item[Weapons] Alloces bite has a BC of 45% of doing + 3 damage in Close
Combat.  It is Rank percent Alloces is also armed as a knight with
broadsword and buckler and ax, mace and knife.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 30 
& 
\textbf{MD:} 28 
& 
\textbf{AG:} 32 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 26 
& 
\textbf{FT:} 38 
& 
\textbf{WP:} 32 
& 
\textbf{PC:} 29
\\
\textbf{PB:} 10 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Alloces will gladly serve a master who promises bloodshed,
but will turn on the summoner if there is no fighting for a prolonged
period and will attempt to drag the summoner into his mouth and devour
him.  Alloces may not be bound.

\end{description}

\subsubsection{AMDUSIAS (The Unicorn Duke)}

\begin{description}

\item[Base Chance] 80%

\item[Lesser Spirits] 29%
 Amdusias appears as a silver unicorn.  He will shed this
insubstantial form and take on the shape of a tall, thin man with a
white beard after a few moments.  His transformation is accompanied by
the sound of trumpets.

\item[Talents, Skills and Magic] Amdusias is a practitioner of the arts of the College of
Celestial Magics.  He possesses the following Skills: Beast Master,
Healer, Ranger, Spy.  He has the power to command entities to do his
bidding in the same manner as a member of the College of Naming
Incantations.  He knows all Generic and Individual True Names, but
will tell these only if his summoner is in dire need.

\item[Movement Rates] Running: 250 yards per minute.

\item[Weapons] Amdusias has no natural weapons and is treated as a man for
purposes of unarmed combat.  He will refuse to wear armor, but may
employ all weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23 
& 
\textbf{MD:} 26 
& 
\textbf{AG:} 24 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 22 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 32 
& 
\textbf{PC:} 31
\\
\textbf{PB:} 16 
& 
\textbf{TMR:} 5 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{ASTAROTH (The Terrible Duke)}

\begin{description}

\item[Base Chance] 66%

\item[Lesser Spirits] 40%
 Astaroth appears in the form of a dark aneel all black
and with a bloody mouth.  He carries a viper coiled about his right
hand and he rides an infernal dragon.  He speaks gaily of horrible
things and laughs readily at pain and disease.  He is exceedingly
cruel.

\item[Talents, Skills and Magic] Astaroth is a practitioner of the arts of the College of
Necromantic Conjurations.  He posesses the following Skills:
Alchemist, Assassin, Military Scientist, Navigator.  He knows the
Generic and Individual True Names of all things, but will only tell
these if it pleases him and he thinks that suffering may be increased
thereby.  He knows men's secrets and will tell them.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Astaroth's breath is poisonous to the same degree as
chlorine and extends outward from his mouth 24 inches in a cone 12
inches wide at the base.  He has no other natural weapons, but will
usually carry a heavy mace which he uses in shattering the bones of
his victims so that they will be crippled and take a long time dying.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 27 
& 
\textbf{MD:} 26 
& 
\textbf{AG:} 28 
& 
\textbf{MA:} 32
\\
\textbf{EN:} 26 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 35 
& 
\textbf{PC:} 26
\\
\textbf{PB:} 8 
& 
\textbf{TMR:} 9 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Astaroth will readily serve the summoner, but only so long
as he is pleased to do so.  He will usually serve so long as the
summoner is involved in much death and cruelty and will depart when
these cease.  Astaroth may not be bound.  If the summoner attempts to
bind him, Astaroth will turn on him and rend him and may claim his
soul.

\end{description}

\subsubsection{BARBATOS (Duke of Virtues)}

\begin{description}

\item[Base Chance] 53%

\item[Lesser Spirits] 30%
 Barbatos appears as a hunter in the company of an
insubstantial host of soldiers led by four kings.  He wears a green
hood frineed in scarlet and a gray cloak.

\item[Talents, Skills and Magic] Barbatos is a member of the College of Naming Incantations.
He possesses the following Skills: Assassin, Beast Master, Healer,
Ranger and Spy.  He is a linguist and understands the speech of birds
and beasts as well as men.  He can break an enchantment used to seal
a place or treasure with his touch. He knows the Generic and
Individual True Names of all things on earth.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Barbatos has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  He refuses to wear armour, but can
employ any weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23 
& 
\textbf{MD:} 27 
& 
\textbf{AG:} 29 
& 
\textbf{MA:} 32
\\
\textbf{EN:} 25 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 34 
& 
\textbf{PC:} 34
\\
\textbf{PB:} 16 
& 
\textbf{TMR:} 8 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Barbatos may not be compelled to tell the True Name of
anything, but he may choose to do so if the need of the summoner is
urgent.

\end{description}

\subsubsection{BATHIN (The Pale Duke)}

\begin{description}

\item[Base Chance] 62%

\item[Lesser Spirits] 30%
 Bathin appears as a pale and cadavertus man riding a
pale horse.  His voice is soft and sad and can compel others to love
him.

\item[Talents, Skills and Magic]Bathin is a practitioner of the arts of the College of the
Mind.  He possesses the following Skills: Alchemist, Beast Master,
Healer, Troubadour.  He can transport men from place to place through
the air and knows all of the virtues of herbs and stones. He is a
patient teacher.  The EP cost to learn a skill is halved if the skill
is learned with Bathin's aid.

\item[Movement Rates] Running: 350yards per minute.

\item[Weapons] Bathin has no natural weapons.  He is treated as a human for
purposes of unarmed combat.  He may wear armor and employ all types of
weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 26 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 22 
& 
\textbf{FT:} 30 
& 
\textbf{WP:} 34 
& 
\textbf{PC:} 26
\\
\textbf{PB:} 13 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\subsubsection{BERITH (The Savage Duke)}

\begin{description}

\item[Base Chance] 65%

\item[Lesser Spirits] 26%
 Berith appears as a soldier dressed in red and riding a
red horse. His face is seamed with what appear to be old scars and he
has a black, bushy beard.  His voice is subtle except when he is
enraged.  On his head is a crown of gold which appears to flame when
the demon is aroused.

\item[Talents, Skills and Magic] Berith is a practitioner of the arts of the College of the
Mind.  He possesses the following Skills: Alchemist, Assassin,
Merchant, Military Scientist, Spy, Thief.  He can turn all metals to
gold by touch.  He also has knowledge of the past, present and future,
but will only trade this knowledge for a life of man or woman.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Berith has no natural weapons. He is treated as a human
for purposes of unarmed combat.  Berith often carries a rapier and a
main-gauche, both of which he keeps envenomed with a nerve poison. He
will refuse to wear armor since this would cover his beautiful
clothes, of which he is inordinately proud.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 26 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 245 
& 
\textbf{FT:} 32 
& 
\textbf{WP:} 31 
& 
\textbf{PC:} 24
\\
\textbf{PB:} 11 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Berith delights in torture and will usually agree to serve
only upon being guaranteed that he can torture any prisoners the party
may take. These he will transport to another plane once they are
captured and bound over to him.  There, he will torture them at his
leisure.  If anvone attempts to deny him this boon (including the
summoner), he will attack them. He may not be bound.

\end{description}

\subsubsection{BUNE (The Dragon Duke)}

\begin{description}

\item[Base Chance] 64%

\item[Lesser Spirits] 30%
 Bune appears as a giant green dragon with three heads:
one that is seemingly human, one that is like a dingo and one that is
like a gryphon.  He speaks with a high-pitched, but comely voice. His
scales are jade and his talons are silver.

\item[Talents, Skills and Magic] Bune is a practitioner of the College of Necromantic
Conjurations.  He possesses the following Skills: Alchemist, Beast
Master, Merchant, Military Scientist.  He is also a linguist and knows
the Generic True Name of all things. He will often bring his summoner
presents in the form of gems or silver in exchange for an early
dismissal from this plane.

\item[Movement Rates] (yards per minute): Run: 500; Fly: 800.

\item[Weapons] Bune may roll over one hex characters as described in the
multi-hex character rules. In addition, he may use his gryphon head to
peck and his dingo head to bite while in Close or Mclee Combat.  Both
have a BC of 35% and do + 2 damage.  They are Rank 3.  Bune may
use his talons in either Closc Combat or Melee Combat with a BC of
50% and damage of + 6.  His talons are Rank 10.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 180 
& 
\textbf{MD:} 9 
& 
\textbf{AG:} 12 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 90 
& 
\textbf{FT:} 100 
& 
\textbf{WP:} 32 
& 
\textbf{PC:} 28
\\
\textbf{PB:} 2 
& 
\textbf{TMR:} 10/16 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 11 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Bunc rcquires an offering of human life beforc he will
serve. Each head must have a life (3 total) and only lives which aec
pleasing to the demon will be accepted.  If disatisfied with an
offering.  Bune can break any binding and will attack and devour the
summoner.  He prefers newborn babies, but will accept maidens when in
a jovial mood.

\end{description}

\subsubsection{CROCELL (Duke of the Bath)}

\begin{description}

\item[Base Chance]71%

\item[Lesser Spirits] 48%
 Crocell appears in the form of a dark angel with flowing
silver hair and yellow cat's eves.  His voice is great and
all-prevading and he ever speaks of things mystical and hidden.

\item[Talents, Skills and Magic] Crocell is a member of the College of Water Magics.  He
possesses the following Skills: Alchemist, Assassin, Astrologer,
Healer, Merchant, Navigator, Spy, Troubadour.  Hc has the power to
warm or cool waters.  He can discover water even in the desert and
can create the illusion of waters to fool the unwary.

\item[Movement Rates] (yards per minute): Run: 250; Fly: 500.

\item[Weapons] Crocell has no natural Aeapons.  Hc is treated as a human
for purposes of unarmed combat.  He may, however, carry a sword
whose blade is of infernal ice that can never melt and which inflicts
burning wounds that later turn white and waxy and eventually turn
black and rot unless cured by magic.  He may also use any other
weapons.  He will not wear armor.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 30 
& 
\textbf{MD:} 28 
& 
\textbf{AG:} 26 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 32 
& 
\textbf{FT:} 40 
& 
\textbf{WP:} 33 
& 
\textbf{PC:} 25
\\
\textbf{PB:} 11 
& 
\textbf{TMR:} 5/10 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\subsubsection{DANTALION (The Duke of Faces)}

\begin{description}

\item[Base Chance] 82%

\item[Lesser Spirits]36%
 Dantalion appears in whatever human form he desires to
put on and will wear the face of any man or woman in an ever changing
pattern or he may be compelled by the summoner to wear only one
face. In his right hand he carries a heavy book in which are written
all the thoughts of all men or women who have ever been and only
Dantalion can read the book.

\item[Talents, Skills and Magic] Dantalion is a mighty mage of the College of the Mind.  He
possesses the following Skills: Assassin, Beast Master, Courtesan,
Healer, Spy.  He knows the Generic and Individual True Names of all
things, the thoughts of all beings and all that has been or will be.
He cannot speak of these thines, however, and any attempt at
compelling him will lead to his return to his own plane.

\item[Movement Rates]Running: 400 yards per minute.

\item[Weapons] Dantalion has no natural weapons and is treated as a human
for purposes of unarmed combat. He may bear any weapons or wear
armour, but prefers not to do so since this makes it harder for him to
alter his shape at will.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 20 
& 
\textbf{MD:} 22 
& 
\textbf{AG:} 21 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 20 
& 
\textbf{FT:} 30 
& 
\textbf{WP:} 35 
& 
\textbf{PC:} 26
\\
\textbf{PB:} 23 
& 
\textbf{TMR:} 8 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Dantalion has great power, but there are limits placed on
its use.  He can employ his powers, but never pass them on. He can
reveal the things he discovers as a result of his powers, but the
knowledge which enables him to exercise those powers is forever
hidden.  He may never teach magic or Skills.  He can never prophesy
for others.

\end{description}

\subsubsection{ELIGOS (The Knightly Duke)}

\begin{description}

\item[Base Chance] 55%

\item[Lesser Spirits]60%
 Eligos appears in the form of a goodly knight in black
armor and riding a dark warhorse.  In his right hand he carries a
banner on a lance. It bears a legend which cannot be read except from
another dimension. In his left hand he holds a curling serpent with
ruby scales whose venom is acid.

\item[Talents, Skills and Magic] Eligos is a practitioner of the arts of the College of
Necromantic Conjurations.  He possesses the following Skills:
Assassin, Beast Master, Mechanician, Military Scientist, Ranger and
Spy. He has the ability to predict with 25% accuracy when and
how wars will come and battles will be lost and won. he can discover
all things hidden within 25 feet of him.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Eligos has no natural weapons.  He is armed with a lance,
sword, mace, battle axe and poinard.  The poinard is envenomed.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28 
& 
\textbf{MD:} 24 
& 
\textbf{AG:} 25 
& 
\textbf{MA:} 20
\\
\textbf{EN:} 26 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 31 
& 
\textbf{PC:} 28
\\
\textbf{PB:} 13 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Eligos is a great leader of men and inspires his allies
whenever he enters battle. No one allied with Eligos will ever flee
from combat so long as he is present.

\end{description}

\subsubsection{FURCALOR (Duke of the Waters)}

\begin{description}

\item[Base Chance] 6%

\item[Lesser Spirits]  30%
 Furcalor appears as a humanoid with gryphon's wings and
green, scaly skin. His hair is like seaweed and he smells of death and
the sea. His voice is a whisper and is compolling when he suggests
action to men.

\item[Talents, Skills and Magic] Furcalor is a practitioner of the College of Water Magics.
He possesses the following Skills: Astrologer, Beast Master,
Merchant, Navigator.  He has the power to command the beasts of the
sea and to gather great storms at sea, but he has no power away from
water. He breathes both water and air.

\item[Movement Rates]Fly 600; Swim 200

\item[Weapons]Furcalor has no natural weapons.  He is treated as a human for the
purposes of unarmed combat.  He may wear specially tailored armor (to
fit around his wings) and may employ any weapon.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 27 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 24 
& 
\textbf{MA:} 33
\\
\textbf{EN:} 27 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 35 
& 
\textbf{PC:} 28
\\
\textbf{PB:} 11 
& 
\textbf{TMR:} 12/4 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Furcalor is obsessed with slaying men by drowning them and
he will pursue his calling ruthlessly at every opportunity unless
bound by his summoner.  He may only be summoned when the summoner is
at sea or near the ocean.

\end{description}

\subsubsection{FURCAS (The Reaping Duke)}

\begin{description}

\item[Base Chance]72%

\item[Lesser Spirits] 20%
 Furcas appears in the guise of a cruel old man with a
long, plaited beard, dressed in pale armour and riding a pale
horse. In his hand he carries a scythe with which he cuts down those
he wishes to torment, for those who fall to his scythe become his
slaves and serve him in his own dimension.  He speaks harshly and is
ever preceeded by a sound like the rattling of old bones.

\item[Talents, Skills and Magic] Furcas is a master of the College of Necromantic
Conjurations.  He possesses the following Skills: Alchemist Assassin,
Astrologer, Merchant, Military Scientist, Navigator, Spy.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Furcas has no natural weapons, but can call his scythe to
him from other dimensions at will.  The Scythe has a BC of
40%. It does + 8 damage and is Rank 10. It is a two-handed
weapon.  Only Furcas can wield the Scythe.  He may use other weapons
as well.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 24 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 24 
& 
\textbf{FT:} 36 
& 
\textbf{WP:} 35 
& 
\textbf{PC:} 24
\\
\textbf{PB:} 12 
& 
\textbf{TMR:} 6 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\subsubsection{GREMORY (The Duke of Songs)}

\begin{description}

\item[Base Chance]76%

\item[Lesser Spirits] 26%
 Gremory appears as a beautiful woman with flowing auburn
hair riding on a camel. He wears a fillet of white gold and robes of
black velvet trimmed in threads of white gold.  His voice is rich and
melodic.

\item[Talents, Skills and Magic] Grcmory is a master of the College of Illusions. He
possesses the following Skills: Assassin, Assayer, Courtesan, Healer,
Spy, Thief, Troubadour.  He can procure the love of women and can
discover the location of treasures that are hidden within 50 feet of
him.  He knows the events of the past, present and future to a small
degree and will prophesy for gain (15% accuracy).

\item[Movement Rates] Running: 300 vards per minute.

\item[Weapons] Gremory has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  He always carries envenomed knives on
his person and rings full of poison on his fingers. His poisons are
extremely swift-acting (10 hits per pulse on the average). He will not
wear armour, but may agree to carry gilded weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 20 
& 
\textbf{MD:} 28 
& 
\textbf{AG:} 29 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 19 
& 
\textbf{FT:} 30 
& 
\textbf{WP:} 36 
& 
\textbf{PC:} 31
\\
\textbf{PB:} 28 
& 
\textbf{TMR:} 6 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Gremory is a lover of beautiful women and will prophesy,
speak of the past, seek treasure or any other thing desired by the
summoner in exchange for the love of a comely young woman. Gremory
usually tires of such liasons after a few months, however, and then
may release his lover and return her to her own plane or may devour
her.

\end{description}

\subsubsection{GUSION (The Wise Duke)}

\begin{description}

\item[Base Chance] 54%

\item[Lesser Spirits] 40%
 Gusion appears in the guise of a mature and strong man
in a saffron robe.  He speaks softly and his advice always seems good
in consequence.

\item[Talents, Skills and Magic] Gusion is a master of the College of the Mind. He possesses
the following Skills: Alchemist, Astrologer, Healer, Mechanician,
Merchant, Military Scientist.  He has the power to reconcile all but
sworn enemies.  He has knowledge of past events and can say what is
happening at any moment in places that are far away.  He can also see
dimly into the future.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Gusion has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  He will not wear armour, but will
employ any weapons provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 24 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 21 
& 
\textbf{FT:} 30 
& 
\textbf{WP:} 32 
& 
\textbf{PC:} 21
\\
\textbf{PB:} 18 
& 
\textbf{TMR:} 6 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Gusion's knowledge is never perfect.  He only catches
glimpses of events past, present and future and one can easily be
misled by him, but he will always tell the truth as he knows it.

\end{description}

\subsubsection{HAVRES (The Leopard Duke)}

\begin{description}

\item[Base Chance]78%

\item[Lesser Spirits] 36%
 Havres first appears as a leopard with blazing eyes, but
this is an insubstantial form.  At the command of the summoner, he
will adopt the form of a dark man dressed in the skin of a leopard and
with a terrible face and eyes which blaze with tongues of fire.

\item[Talents, Skills and Magic] Havres is a master of the College of Fire Magics. He
possesses the following Skills: Assassin, Beast Master, Ranger.
Havres has the power to create illusory fires which will destroy those
who believe them to be real, but will do no harm to those who
successfully disbelieve them.  He knows something of the past, present
and future and can see into places where there is/was/will be fire.

\item[Movement Rates] Running: 400 yards per minute.

\item[Weapons] Havres has no natural weapons and is treated as a human for
purposes of unarmed combat. He may carry weapons at the summoner's
request.  He will not wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 30 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 32 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 28 
& 
\textbf{FT:} 36 
& 
\textbf{WP:} 33 
& 
\textbf{PC:} 34
\\
\textbf{PB:} 8 
& 
\textbf{TMR:} 8 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Havres normally has only a 10% chance of accurately
telling of past or future events or events which are happening far
away, but if these events are happening close to fire, his chances
improve greatly. The larger the fire and the closer it is to the
event, the better Havres chances of seeing the event.

\end{description}

\subsubsection{MURMUR (Duke of Thrones)}

\begin{description}

\item[Base Chance]75%

\item[Lesser Spirits] 30%
 Murmur appears as a warrior in green armor, riding on a
gryphon and wearing a ducal crown.  He is preceded by shades blowing
trumpets.

\item[Talents, Skills and Magic] Murmur is a practitioner of the arts of the College of
Necromantic Conjurations.  He possesses the folloA,ing Skills:
Alchemist, Assassin, Astrologer, Beast Master, Healer, Mechanician,
Merchant, Military Scientist, Navigator.  He also possesses the power
to make the dead appear before the summoner and speak to him of their
lives and deaths and to answer the questions he may put to them.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Murmur has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  He usually carries a heavy war
hammer, a shield and a brace of knives.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 27 
& 
\textbf{MD:} 23 
& 
\textbf{AG:} 25 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 23 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 33 
& 
\textbf{PC:} 31
\\
\textbf{PB:} 15 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{SALLOS (The Mighty Duke)}

\begin{description}

\item[Base Chance] 57%

\item[Lesser Spirits] 30%
 Sallos appears as a mighty knight in silver armor
mounted on a crocodile and wearing a ducal crown on his head.  He
comes in peace, though, and will fight only with reluctance.

\item[Talents, Skills and Magic] Sallos is a practitioner of the arts of the College of
Ensorcelments and Enchantments. He possesses the following Skills:
Beast Master, Healer, Military Scientist, Ranger, Troubador.  He has
the talent to inspire love between men and women.

\item[Movement Rates] Running: 300yards per minute.

\item[Weapons] Sallos has no natural weapon.  He is armed with a sword
and a main gauche.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 24 
& 
\textbf{MD:} 25 
& 
\textbf{AG:} 25 
& 
\textbf{MA:} 29
\\
\textbf{EN:} 26 
& 
\textbf{FT:} 34 
& 
\textbf{WP:} 32 
& 
\textbf{PC:} 31
\\
\textbf{PB:} 21 
& 
\textbf{TMR:} 6 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  3 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Sallos would rather drink and wench than fight and kill and
is, perhaps, something of a coward.  Once in combat, though, he will
not flee until all his enemies are dead.

\end{description}

\subsubsection{UVALL (Duke of Sands and Wastes)}

\begin{description}

\item[Base Chance]70%

\item[Lesser Spirits] 37%
 Uvall appears first as a large black dromedary.  At the
command of the summoner, he will leave this insubstantial form and
adopt the form of a man garbed in flowing robes and burned dark, like
one of the wild dwellers in the wastelands.  His voice is always
hushed and he speaks with menace.

\item[Talents, Skills and Magic] Uvall is a practitioner of the arts of the College of
Illusions. He possesses the following Skills: Alchemist, Assassin,
Beast Master, Merchant, Ranger, Spy, Thief.  He has the power to
procure for the summoner the love of women.  He also can cause others
to love the summoner and become devoted to him.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Uvall has no natural weapons. He is treated as a human for
purposes of unarmed combat. He usually comes armed with a scimitar,
spear and a brace of sharp throwing knives which break when they enter
the victim's body and leave splinters that must be removed or they
will burrow their way to the heart and kill the victim in D10 + 3 days.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25 
& 
\textbf{MD:} 26 
& 
\textbf{AG:} 27 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 24 
& 
\textbf{FT:} 38 
& 
\textbf{WP:} 35 
& 
\textbf{PC:} 34
\\
\textbf{PB:} 13 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Uvall has a small knowledge of the past, present and future
of things and can be persuaded to speak (with 20% accuracy) by
a gift of a young boy to serve him on his own plane. He is cunning and
evil and will attempt to betray the summoner even when bound, though
he will not directly attack him.

\end{description}

\subsubsection{VALEFOR (The Duke of Thieves)}

\begin{description}

\item[Base Chance] 52%

\item[Lesser Spirits] 10%
 Valefor appears as a mountain lion with the head of an
ass.  He has a sharp, braying voice and an unpleasant manner.

\item[Talents, Skills and Magic] Valefor is a practitioner of the College of Ensorcelments
and Enchantments.  He possesses the following Skills: Assassin,
Ranger, Spy and Thief. He sometimes represents himself as the god of
thieves and the EP cost to learn the thief skill is halved if the
skill is learned from Valefor.

\item[Movement Rates] Running: 400 yards per minute.

\item[Weapons] Valfor may bite in Close Combat with a BC of 25% of
doing D + 1 damage.  His bite is Rank 3.  He may also use his claws in
Melee or Close Combat with a BC of 45%. He does + 2 damage with
claws in Melee and + 4 in Close Combat.  His claws are Rank 5.  He may
not wear armor or employ man-made weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 30 
& 
\textbf{MD:} 4 
& 
\textbf{AG:} 32 
& 
\textbf{MA:} 33
\\
\textbf{EN:} 29 
& 
\textbf{FT:} 36 
& 
\textbf{WP:} 35 
& 
\textbf{PC:} 35
\\
\textbf{PB:} 4 
& 
\textbf{TMR:} 8 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Valefor will always counsel theft even where it involves
stealing from companions and friends. He cannot be broken of that
habit, no matter how he is bound.  His advice will tend to be
persuasive.  Further, he will steal objects from castles or inns where
his summoner might stay and secret them in the summoner's belongings.

\end{description}

\subsubsection{VAPULA (The Lion Duke)}

\begin{description}

\item[Base Chance]77%

\item[Lesser Spirits] 36%
 Vapula appears as a tawny lion with the wings of a
gryphon and a red mane.  He has great green eyes which seem full of
wisdom and his voice is a quiet rumbling.

\item[Talents, Skills and Magic] Vapula is a member of the College of Earth Magics. He
possesses the following Skills: Beast Master, Healer, Ranger.

\item[Movement Rates] Running: 400 yards per minute.

\item[Weapons] Vapula may bite in Close Combat with a BC of 40% of
doing + 4 damage.  His bite is Rank 5.  Vapula may also use claws in
Melee and Close Combat.  He has a BC of 45% with claws in Melee
and a BC of 55% with claws in Close Combat.  His claws do + 6
damage and are Rank 8.  He mav not wear armour or employ weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 35 
& 
\textbf{MD:} 4 
& 
\textbf{AG:} 30 
& 
\textbf{MA:} 32
\\
\textbf{EN:} 28 
& 
\textbf{FT:} 36 
& 
\textbf{WP:} 34 
& 
\textbf{PC:} 31
\\
\textbf{PB:} 18 
& 
\textbf{TMR:} 8 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\subsubsection{VEPHAR (The Sea Duke)}

\begin{description}

\item[Base Chance] 69%

\item[Lesser Spirits] 29%
 Vephar appears as a mermaid with emerald scales trimmed
in silver and seaweed tangled in his hair.  He may also appear at
times to be a beautiful woman, but this shape is insubstantial and
is used only to lure sailors to their deaths.  His fingers are webbed
and he has small gills behind his cars.

\item[Talents, Skills and Magic] Vephar is a master of the College of Illusions.  He
possesses the following Skills: Beast Master, Courtesan, Navigator.
He can raise great storms at sea by willing it.  Vephar also has
the ability to cause horrible wounds by looking upon a man.  These
will appear as gashes, but will putrefy and fill with worms and
within three days the victim will die unless cured by magic.

\item[Movement Rates] Swimming: 600 yards per minute.

\item[Weapons] Vephar has no natural weapons and is human for purposes of
unarmed combat.  He may not wear armour, but may employ weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 12 
& 
\textbf{MD:} 19 
& 
\textbf{AG:} 18 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 18 
& 
\textbf{FT:} 30 
& 
\textbf{WP:} 34 
& 
\textbf{PC:} 26
\\
\textbf{PB:} 28 
& 
\textbf{TMR:} 12 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Vephar cannot exit the sea.  he becomes totally
insubstantial and melts into his own plane if he attempts to do so.
He can guide ships through any waters (even in the absence of sun
and stars) and will always do so at the summoner's request.

\end{description}

\subsubsection{ZEPAR (The Red Duke)}

\begin{description}

\item[Base Chance] 60%

\item[Lesser Spirits] 26%
 Zepar appears as a slight warrior with a clubfoot
dressed all iri red armor.  He is of a dark mcin and speaks in a
grating voice filled with contempt for all who inhabit this plane.

\item[Talents, Skills and Magic] Zepar is a practitioner of the arts of the College of
Illusions.  He possesses the following Skills: Alchemy, Assassin,
Merchant, Military Scientist, Spy and Troubador. He is able to cause
women to fall in love with men and so allow themselves to be seduced.
His touch, however, makes women barren.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Zepar has no natural weapons.  He is treated as a human for
purposes of unarmed combat.  He may carry a variety of weapons, but
all will be smeared with foul venom which inflicts horrible scarring
injuries that may only be cured by magic.  Zepar's envenomed weapons
do 2 points of damage per hour to Fatigue or Endurance as a result of
their poisonous coating.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 18 
& 
\textbf{MD:} 29 
& 
\textbf{AG:} 28 
& 
\textbf{MA:} 30
\\
\textbf{EN:} 26 
& 
\textbf{FT:} 35 
& 
\textbf{WP:} 32 
& 
\textbf{PC:} 28
\\
\textbf{PB:} 13 
& 
\textbf{TMR:} 7 
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

      
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Zepar is an unfaithful sevant who always resents his
summoner's ability to bring and hold him in this plane.  He will
make common cause with his summoner's enemies if they promise to
return him to his own dimension, even if he is bound.  However, when
bound, he cannot directly attack his summoner and must act with
guile.

\end{description}

\subsection{Princes}

\subsubsection{GAAP (The Servant Prince)}

\begin{description}

\item[Base Chance] 52%

\item[Lesser Spirits] 66%
 Gasp appears as a man dressed in animal skins and
accoutered for war.  He is bronzed (as if by the sun) and looks to be a
man of the wilderness.  Behind him come the insubstantial shades of
four princes whom he appears always to serve and guide.

\item[Talents, Skills and Magic] Gasp is a master of the COllege Of Ensorcelments and
Enchantments.  He possesses the following Skills: Alchemist, Assassin,
Beas, Master, Healer, Mechanician, Navigator, Ranger, Spy, Thief.  He
has the power to break the ties between a magician and his familiar
and to send the familiar back to where he came.  He can teleport
beings within their plane by willing it.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Gasp has no natural weapons.  He is treated as human for
purposes of unarmed combat.  However, see below.  Gasp may wear no
armor.  He carries any weapons the summoner provides.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28		
& 
\textbf{MD:} 26		
& 
\textbf{AG:} 27		
& 
\textbf{MA:} 31
\\
\textbf{EN:} 28		
& 
\textbf{FT:} 36		
& 
\textbf{WP:} 33		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 12		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Those eho touch Gasp will have their senses stolen for a
number of minutes equal to D100.  Thereafter, they will awaken to find
that they love what which they formerly hated and hate that which they
once loved.  The effect of Gaap's touch will last until dispelled bv
magic.

\end{description}

\subsubsection{IPOS (The Prince of Fools)}

\begin{description}

\item[Base Chance] 49%percnt;

\item[Lesser Spirits] 36%
 Ipos appears as an angel with a lion's head, the feet of
a goose and the tail of a hare.  He has a sweet voice and is an
accomplished musician.

\item[Talents, Skills and Magic] Ipos is a master of the College of Illusions.  He possesses
the following Skills: Healer, Spy, Thief and Troubadour.  He can speak
truely of what has happened or is about to happen in a place he
occupies.  He has the talent of making men bold and will use this
talent no matter how he is bound.  Those affected by the talent add 10
to their Willpower in the presence of lpos.

\item[Movement Rates] (yards per minute): Run: 200; Fly: 400

\item[Weapons]Ipos bites for +2 damage in close combat.  He has a BC of
45% with his bite.  It is Rank 4.  Ipos will not wear armour.
He weill carry any weapoons provided by the summer.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 24		
& 
\textbf{MD:} 21		
& 
\textbf{AG:} 18		
& 
\textbf{MA:} 34
\\
\textbf{EN:} 26		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 38		
& 
\textbf{PC:} 29
\\
\textbf{PB:} 11		
& 
\textbf{TMR:} 4/8	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]Ipos makes all men in his presence bold (not merely the
summoner and friends) when he uses his talent.  Further, men are more
likely to fight than talk when they are in the presence of Ipos and
come among strangers.  Those affected by Ipos talent will never flee
danger and will fight to the death.

\end{description}

\subsubsection{OROBAS (The Equine Prince)}

\begin{description}

\item[Base Chance] 65%

\item[Lesser Spirits] 20%
 Orobas appears in the form of a great black stallion
with a carmine main and flowing carmine tail. His hooves are silver
and sharpened like those of a warhorse.  He speaks in the squealing
voice of a horse.

\item[Talents, Skills and Magic] Orobas is a practitioner of the arts of the College of
Naming Incantations.  He possesses the following Skills: Beast Master,
Healer, Navigator, Ranger, Spy, Thief.  Orobas knows the Generic and
Individual True Names of all things of this plane.  He will gladly
permit the summoner to acquire these names, but will provide them to
no one else.

\item[Movement Rates] Running: 500 yards per minute.

\item[Weapons] Orobas may bite for + 2 damage in Close Combat.  His BC for
biting is 30% and his bite is Rank 2.  He may use his hooves to
strike in Melee Combat or to trample in Close Combat.  They have a BC
of 40% in Melee and do + 3 damage.  In Close Combat they have a
BC of 55% and do + 6 damage.  They are Rank 6.  Orobas will wear
the barding of a warhorse if the summoner provides it. He uses no
weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 45		
& 
\textbf{MD:} 20		
& 
\textbf{AG:} 20		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 30		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 24		
& 
\textbf{TMR:} 10		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Orobas is a faithful retainer and, once bound, will look
after the summoner's well being as if it were his own.  He will be
especially careful to protect the summoner from the powers and whiles
of other demons or spirits the summoner may encounter.  He will even
permit the summoner to ride him as a horse.

\end{description}

\subsubsection{SEIR (The Willing Prince)}

\begin{description}

\item[Base Chance]75%

\item[Lesser Spirits] 26%
 Seir appears as a handsome man with hair the colour of
corn and eyes like ice riding a silver stallion with gryphon wings. He
speaks with a soft and cheerful voice and is of a good nature and
always amenable to the desires of the summoner.

\item[Talents, Skills and Magic] Seir is a practitioner of the College of Air Magics. He
possesses the following Skills: Astrologer, Healer, Merchant,
Navigator, Spy, Troubador.  He has knowledge of all theives and can
sometimes tell where the loot is hidden.  He can pass over the entire
earth at the winking of an eye and can carry others with him.  He
brings abundance of things in his wake.

\item[Movement Rates] Runninz: 350 yards per minute.

\item[Weapons] Seir has no natural weapons. He is treated as a human for
purposes of unarmed combat.  Seir will not wear armour or carry any
weapons except a quarterstaff.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25		
& 
\textbf{MD:} 22		
& 
\textbf{AG:} 26		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 24		
& 
\textbf{FT:} 32		
& 
\textbf{WP:} 31		
& 
\textbf{PC:} 32
\\
\textbf{PB:} 23		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Characters in the company of Seir and his summoner will
always experience a flood of good luck so far as physical objects are
concerned.  The average value of treasure they may acquire will be
10% greater than what they may acquire when not in company with
Seir.

\end{description}

\subsubsection{SITRI (The Beautiful Prince)}

\begin{description}

\item[Base Chance] 47%

\item[Lesser Spirits] 60% Sitri appears as a large, muscular man with the head of
a leopard and the wings of a gryphon.  He may also appear in
insubstantial form as a man who is unbearably beautiful and
well-formed or as a lovely woman of sensual mein.

\item[Talents, Skills and Magic] Sitri is a master of the College of Illusions.  He possesses
the following Skills: Assassin, Beast Master, Courtesan, Spy, Thief,
Troubadour.  Sitfi has the power to inflame men and women with the
love of each other and to cause them to show themselves naked and
frolic with one another in unseemly ways.

\item[Movement Rates] (yards per minute): Run: 350; Fly: 400.

\item[Weapons] Sitri may bite during Close Combat with a BC of 45%and +2 damage.  His bite is Rank 4.  Sitri will wear armour if the
summoner has it specially made for him and will use any weapons
provided.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 30		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 26		
& 
\textbf{MA:} 33
\\
\textbf{EN:} 28		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 25
\\
\textbf{PB:} 35		
& 
\textbf{TMR:} 7/8	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Sitri is a lover of all things carnal and will not be
prevented from encouraging carnal activity or using his special
talents, no matter how he is bound.

\end{description}

\subsubsection{STOLAS (The Raven Prince)}

\begin{description}

\item[Base Chance] 54%

\item[Lesser Spirits] 26% Stolas appears as a massive raven with talons of silver
and eyes rimmed with red.  He speaks with a cracked and horrible
voice.

\item[Talents, Skills and Magic] Stolas is a mage of the College of Air Magics.  He possesses
the following Skills: Assassin, Beast Master, Navigator, Spy, Thief.
He knows the magical powers and virtues of all herbs and stones and
will speak the truth about these when requested to do so.

\item[Movement Rates] Flying: 500 yards per minute.

\item[Weapons] Stolas has a BC of 50% of doing D - 1 damage in Close
Combat with his beak. It is Rank 4. He has a BC of 45% of doing
D + 3 damaze in Close Combat with his talons.  They are Rank 6.
Stolas cannot wear armor or use weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 14		
& 
\textbf{MD:} 6		
& 
\textbf{AG:} 12		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 21		
& 
\textbf{FT:} 28		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 34
\\
\textbf{PB:} 13		
& 
\textbf{TMR:} 10		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{VASSAGO (Prince of Prophecy)}

\begin{description}

\item[Base Chance] 45%

\item[Lesser Spirits] 26% Vassago appears as an aged man riding a crocodile and
carrying on his right wrist a Goshawk.  His eyes are hollow and see
into other dimensions, but he is blind in this dimension.

\item[Talents, Skills and Magic] N@Vassago is a mage of the College of Naming Incantations.
He possesses the following Skills: Alchemist, Healer, Troubadour.
Vassago has the ability to locate things which have been hidden or
lost if he is within 100 feet of them.  He also knows all Generic and
Individual True Names, but is reluctant to tell these names except to
save the life of the summoner.

\item[Movement Rates] Running: 150 yards per minute.

\item[Weapons] Vassago has no natual weapons.  He is treated as a human for
purposes of unarmed combat.  Vassago cannot wear armour or use
weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 12		
& 
\textbf{MD:} 13		
& 
\textbf{AG:} 13		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 17		
& 
\textbf{FT:} 28		
& 
\textbf{WP:} 37		
& 
\textbf{PC:} 19
\\
\textbf{PB:} 12		
& 
\textbf{TMR:} 3		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] There is only a 40% chance what Vassago will reveal a
True Name in the absence of physical danger to the summoner. There is
only a 50% chance that he will willingly tell the location of a
hidden or lost object.  He may be compelled to speak of these things
through the performance of a Ritual of True Speaking.

\end{description}

\subsection{Presidents}

\subsubsection{AVNAS (President of Fire)}

\begin{description}

\item[Base Chance] 50%

\item[Lesser Spirits] 36% Avnas appears as a pillar of fire and smoke.  Those who
look into the flames can see the panorama of the history of other
dimensions and can sometimes read the events of their own plane and
even their own future.  At the command of the summoner, Avnas will
abandon this insubstantial form and take on the form of a small
withered brown man.

\item[Talents, Skills and Magic] Avnas is a mage of the College of Fire.  He possesses the
following Skills: Alchemist, Astrologer, Healer, Mechanician,
Merchant, Navigator.  Avnas can locate any treasure that is guarded
by spirits, ghosts or other undead and that is within 500 feet of him.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Avnas has no natural weapons. He is treated as a human for
purposes of unarmed combat.  He may not wear armour or use weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 14		
& 
\textbf{MD:} 26		
& 
\textbf{AG:} 21		
& 
\textbf{MA:} 38
\\
\textbf{EN:} 19		
& 
\textbf{FT:} 30		
& 
\textbf{WP:} 40		
& 
\textbf{PC:} 33
\\
\textbf{PB:} 12		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Avnas may be ordered to take form as a pillar of fire at
any time.  However, in that state, he cannot act on this plane and his
fire will not burn those who touch it.  It will instead feel cold and
dead and those who touch it will see their own death.

\end{description}

\subsubsection{BUER (The Star President)}

\begin{description}

\item[Base Chance] 37%

\item[Lesser Spirits] 50% Buer appears always as a Centaur, armed with a great
bow.  His skin is silver and shimmers like starlight.  His hair is all
white and his eyes are the color of Mars.

\item[Talents, Skills and Magic] Buer is a master of the College of Celestial Magics. He
possesses the following Skills: Healer, Navigator, Ranger, Troubador.
Buer's touch can heal all hurts and diseases, especially those which
have been caused by magic.  He is a great teacher of languages and
will freely pass on his knowledge to others.

\item[Movement Rates] Running: 500 yards per minute.

\item[Weapons] Buer uses weapons as a man. In addition, his hooves do + 4
damage in Close Combat due to trampling.  The BC for use of hooves is
60%. They are Rank 8. Buer may use any weapons, but may not wear
armor.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 38		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 33		
& 
\textbf{PC:} 32
\\
\textbf{PB:} 19		
& 
\textbf{TMR:} 10		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\subsubsection{CAMIO (The Thrush President)}

\begin{description}

\item[Base Chance] 47%

\item[Lesser Spirits] 30% Camio first appears as a small thrush, but then leaves this
insubstantial form and takes on that of a yellow-skinned man. He
carries in his hand a sharp sword smeared with venom. He does not
speak, but is, nonetheless, a great disputer.  His words appear before
his hearers as letters written in the air with fiery coals.

\item[Talents, Skills and Magic] Camio is a member of the College of Fire.  He possesses the
following Skills: Beast Master, Healer, Ranger, Spy, Thief.  Camio is
a great linguist and will willingly teach his summoner the languages
of birds and mammals and also teach him to understand what has passed
in a place by reading the speech of running waters.

\item[Movement Rates]   Running:   400   yards   per    minute.

\item[Weapons] Camio has no natural weapons.  He is treated as a human for
purposes of unarmed combat.  Camio will wear any armor provided by the
summoner and will carry other weapons, but will not use them unless
his sword is damaged.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 18		
& 
\textbf{MD:} 23		
& 
\textbf{AG:} 26		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 24		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 38		
& 
\textbf{PC:} 36
\\
\textbf{PB:} 14		
& 
\textbf{TMR:} 8		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Camio will always give true answers to all things.

\end{description}

\subsubsection{FORAS (President and Lord of Seekers)}

\begin{description}

\item[Base Chance] 42%

\item[Lesser Spirits] 29% Foras appears in the form of a strong man, unarmed and
mild seeming.

\item[Talents, Skills and Magic] Foras is a master of the College of Illusion.  He possesses
the following Skills: Astrologer, Healer, Mechanician, Merchant,
Navigator, Troubadour. He has the power to make men temporarily
invisible.  He is a linguist and a teacher of language.  He also has
the ability to discover things hidden (especially treasure) if he is
within 50 feet of them.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Foras has no natural weapons. He is treated as a human for
purposes of unarmed combat.  Foras will wear any armor provided by the
summoner and will carry and use weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 33		
& 
\textbf{MD:} 28		
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 32
\\
\textbf{EN:} 27		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 27
\\
\textbf{PB:} 23		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Foras will grant long life (triple life span) to those who
will worship him and know no other gods or converse with other demons.
As a teacher he may pass on Skills at the rate of one hour being equal
to one month of training under anv human teacher and at half the
normal EP cost.

\end{description}

\subsubsection{HAAGENTI (The Winged President)}

\begin{description}

\item[Base Chance] 45%

\item[Lesser Spirits] 33% Haggenti at first appears as a Red bull with a Gryphon's
wings and horns tipped in gold, but this form is insubstantial and at
the command of the summoner he will put on the shape of a red man with
black hair.

\item[Talents, Skills and Magic] Haagenti is a mighty member of the College of Ensorcelments
and Enchantments.  He possesses the following Skills: Alchemist,
Healer, Merchant, Navigator.  Haagenti can change wine into water and
water into wine with a touch.  He can also transmute all metals into
gold with a touch.  One hour spent learning alchemy under the tutelage
of Haagenti is the equivalent of two months learning from a human.
The EP cost to achieve Rank with this skill is halved (round down)
when taught by Haagenti.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Hagenti has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  Haagenti will wear armour and use
weapons provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 40		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 23		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 28		
& 
\textbf{FT:} 40		
& 
\textbf{WP:} 38		
& 
\textbf{PC:} 24
\\
\textbf{PB:} 13		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Haagenti's office is to make men wise and he will instruct
anyone (not merely the summoner) and demand no fee if the lesson is
well-learned, but will become enraged if he feels that the pupil has
been lax and will turn on him and rend him with his powerful hands.

\end{description}

\subsubsection{LABOLAS (President and Master of Murderers)}

\begin{description}

\item[Base Chance] 40%

\item[Lesser Spirits] 36% Labolas always appears in the form of a winged mastiff.
His wings are always those of a Griffon and bear him at great
speeds. His mouth is covered in froth from his slavering and gnashing
of teeth.

\item[Talents, Skills and Magic] Labolas is a member of the College of Illusions.  He
possesses the following Skills: Assassin, Beast Master, Military
Scientist, Ranger, Spy, Thief.  He can teach any Skill of which he is
possessed with an hour at no EP cost, but demands as his reward a
life.  If the offering is not pleasing to him, he will take the life
of the summoner instead.  He prefers human lives to all others.

\item[Movement Rates] Running: 350 yards per minute; Flying: 500 yards per minute.

\item[Weapons] Labolas may bite in Close Combat with a BC of 50%.  He
does + 2 damage and his bite is Rank 8.  The bite of Labolas always
becomes infected.  Labolas may never wear armour or carry or use
weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 18		
& 
\textbf{MD:} 3		
& 
\textbf{AG:} 15		
& 
\textbf{MA:} 32
\\
\textbf{EN:} 19		
& 
\textbf{FT:} 30		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 2		
& 
\textbf{TMR:} 7/10	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Labolas is capable of reconciling enemies when he so
desires, but prefers strife and bloodshed.  His psychic life requires
manslaughter and he will seek to promote it at every opportunitv,
even among the summoner's family and companions.

\end{description}

\subsubsection{MALPHAS (President of Deceivers)}

\begin{description}

\item[Base Chance] 45%

\item[Lesser Spirits] 40% Malphas appears as a large crow, but this form is
insubstantial and will melt away at the summoner's request to reveal
the form of a dark man dressed in black velvet studded with precious
gems. His voice will be rough, but deep and compelling.

\item[Talents, Skills and Magic] Malphas is a member of the College of Celestial Magics. He
possesses the following Skills: Alchemist, Assassin, Beast Master,
Merchant, Ranger, Spy, Thief. He is also a master of spoken and
written language.  He knows the art of magically raising houses and
high towers.  He also knows the secret desires of the summoner's
enemies and will reveal those thoughts upon request.

\item[Movement Rates] Running: 350 yards per minu[e.

\item[Weapons] Malphas has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  Malphas will refuse to wear
armor. However, he will use any weapons provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 26		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 27		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 30
\\
\textbf{PB:} 18		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Malphas will deceive the summoner knowingly and willfully
unless compelled to tell the truth by a Ritual of True Speaking.

\end{description}

\subsubsection{MARBAS (President and Master of the Seal)}

\begin{description}

\item[Base Chance] 35%

\item[Lesser Spirits] 36% Marbas first appears as a tawny lion with a black mane
and bloody muzzle, but this form is insubstantial.  Upon the order of
the summoner, he will adopt the guise of a golden-skinned man with
a shock of black hair.

\item[Talents, Skills and Magic] Marbas is a master of the College of Illusions. He possesses
the following Skills: Alchemist, Mechanician, Merchant, Military
Scientist.  He has the power to inflict on men rotting diseases by the
merest touch.  He also may change the shape of men by his touch so
that they become as wolves or bears. He has knowledge of many of the
secrets of the universe.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Marbas has no natural weapons.  He is treated as a human for
purposes of unarmed combat.  Marbas will wear any armor and use any
weapons provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 26		
& 
\textbf{MD:} 29		
& 
\textbf{AG:} 24		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 26		
& 
\textbf{FT:} 38		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 25
\\
\textbf{PB:} 16		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Marbas is a speaker of truth and will never knowingly
deceive his summoner.  He knows the Generic True Names of all things
and knows the Individual True Names of many beings who are wealthy and
famous.

\end{description}

\subsubsection{VOLAC (The Dragon President)}

\begin{description}

\item[Base Chance] 52%

\item[Lesser Spirits] 38% Volac appears as a child with the wings of an angel
riding on the back of an infernal Lion-headed red dragon.  Hc speaks
with a high pitched and silly voice, but his words are always true, if
not wise.

\item[Talents, Skills and Magic] Volac is a practitioner of the College of Naming
Incantations.  He possesses the following Skills: Beast Master,
Ranger, Spy, Thief, Troubador. He can locate treasure hidden anywhere
within 25 feet of him.  Volac can also call serpents and cause them to
obey him once they have arrived. Volac knows and will tell the Generic
True Names of all things.

\item[Movement Rates] (yards per minute): Run: 250; Fly: 500.

\item[Weapons] Volac has no natural weapons. He is treated as a human for
pur-poses of unarmed combat.  He may use weapons, but will not wear
armor.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 11		
& 
\textbf{MD:} 18		
& 
\textbf{AG:} 17		
& 
\textbf{MA:} 36
\\
\textbf{EN:} 18		
& 
\textbf{FT:} 30		
& 
\textbf{WP:} 39		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 26		
& 
\textbf{TMR:} 5/10	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\subsubsection{VOSO (The Leopard President)}

\begin{description}

\item[Base Chance] 49%

\item[Lesser Spirits] 30% Voso appears as a mighty leopard with terrible green
eyes and spots tinged with carmine.  He speaks with a cunning voice,
the better to fool his victims. He will aerempt to devour the
summoner unless bound.

\item[Talents, Skills and Magic] Voso is a great member of the College of Ensorcelments and
Enchantments.  He possesses the following Skills: Alchemist, Healer,
Mechanician and Navigator.  He has the power to change a man's form by
touch into another form so that his victim does not know that he has
been ensorcelled and believes himself always to have had the form
given him by Voso. He knows many secrets.

\item[Movement Rates] Running: 500 yards per minute.


\item[Weapons] Voso's bite does + 2 damage in Close Combat. It has a BC of
45% and is Rank 6. Voso may use his claws in either Melee or
Close Combat.  They have a BC of 40% in Melee and of 60%in Close Combat and do + 4 damage. They are Rank 8. Voso will not use
weapons or wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 35		
& 
\textbf{MD:} 4		
& 
\textbf{AG:} 32		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 29		
& 
\textbf{FT:} 36		
& 
\textbf{WP:} 37		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 20		
& 
\textbf{TMR:} 37		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Voso knows the Generic and Individual True Names of all
things and beings.  He will, however, reveal a True Name only if he is
provided with a human victim (preferably a maiden or child since their
flesh is the most tender).  A new victim must be provided for each
True Name the summoner wishes to acquire.

\end{description}

\subsection{Earls}

\subsubsection{ANDROMALIUS (Earl of Justice)}

\begin{description}

\item[Base Chance] 45% 

\item[Lesser Spirits] 36% Andromalius appears as a large man of fearsome mean
holding a viper in one hand.  He dresses all in cloth of gold.

\item[Talents, Skills and Magic] Andromalius is a member of the Colleize of Naming
Incantations.  He possesses the following Skills: Alchemist,
Astrologer, Beast Master, Healer, Mechanician, Merchant, Navigator,
Ranger, Spy.  He also has the ability to discover any hidden treasure
within 200 feet of where he is standing.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Andromalius has no natural weapons.  He is treated as a
human for purposes of unarmed combat.  He will refuse to wear armour,
but will use any weapons provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 26		
& 
\textbf{MD:} 23		
& 
\textbf{AG:} 24		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 33
\\
\textbf{PB:} 13		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Andromalius is dedicated to the punishment of thieves
wherever he finds it and to the return of the victims valuables.  He
will turn on any summoner who attempts to steal while in his company.
He may also react in the same manner to underhanded dealing even with
non-humanoid beings.

\end{description}

\subsubsection{BIFRONS (Earl of the Dead)}

\begin{description}

\item[Base Chance] 38%

\item[Lesser Spirits] 60% Bifrons may appear in insubstantial form as a horned
monster, but always appears as a handsome young man when fully
material.

\item[Talents, Skills and Magic] Bifrons is a member of the College of Necromantic
Conjurations.  He possesses the following Skills: Alchemist, Assassin,
Beast Master, Merchant Ranger, Spy, Thief. He is also a great
Navigator and an hour spent learning navigation from Bifrons is equal
to a month spent learning the art elsewhere and costs no EP's.
Whenever he passes a place of the dead at night candles will appear
lit on the graves.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Bifrons has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  Bifrons will wear any armour and use
any weapons provided by the summoner.  He prefers envenomed weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23		
& 
\textbf{MD:} 26		
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 32
\\
\textbf{EN:} 24		
& 
\textbf{FT:} 31		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 32
\\
\textbf{PB:} 22		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Bifrons has the power to make dead bodies appear as they
did when they were alive so that it cannot be told that they are dead
and corrupted.

\end{description}

\subsubsection{BOTIS (The Ugly Earl)}

\begin{description}

\item[Base Chance] 25%

\item[Lesser Spirits] 60% Botis will first appear as a laree, horrible snake of
insubstantial form. At the command of the summoner, he will
take on human form, but with great teeth and sharp horns. He
will be ca@ing an envenomed sword.

\item[Talents, Skills and Magic] Botis is a member of the College of Ensorcelments and
Enchantments. He possesses the following Skills: Alchemist,
Astrologer, Beast Master, Navigator, Ranger, Spy. He has the power to
make all but bitter sworn enemies friendly to each other while in his
presence.  He also has the ability to tell what has transpired in any
place he occupies and part of what will (may) transpire there.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Botis may do D + 1 damage with his bite in Close Combat with a
BC of 35% of hitting his target. His bite is Rank 4. He may make a
single horn attack in Close Combat with a BC of 30% of doing + 3
damage.  His horns are Rank 6. The venom of his sword does 6 damage
points per Pulse. He will wear any armour provided and employ other
weapons if the summoner desires, but will prefer to use his sword.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 27		
& 
\textbf{MD:} 24		
& 
\textbf{AG:} 25		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 11		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{FURFUR (Earl of Storms and Lightning)}

\begin{description}

\item[Base Chance] 32%

\item[Lesser Spirits] 26% Furfur appears as a hart with a fiery tail. He may also
take on an insubstantial form as an angel, but only when forced
within a triangle and compelled by a Ritual of True Speaking to utter
truth instead of the falsehoods he prefers.  Furfur speaks with a
rough and grating voice that is extremely unpleasant.

\item[Talents, Skills and Magic] Furfur is a master of the College of Air Magics.  He
possesses the following Skills: Alchemist, Mechanician, Navigator,
Ranger, Spy.  Furfur can raise thunder and lightnings and cause great
storms to form.  He also is able to influence men and women to make
love and participate in perversions by the sound of his voice.

\item[Movement Rates] Running: 400 yards per minute.

\item[Weapons] Furfur may use his horns in Melee or Close Combat and his
hooves in Close Combat.  His horns have a BC of 45% and do + 4
damage.  His hooves have a BC of 50% and do + 3 damage.  They
are both Rank 8.  Furfur will not use weapons or wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 32		
& 
\textbf{MD:} 4		
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 31
\\
\textbf{EN:} 28		
& 
\textbf{FT:} 34		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 33
\\
\textbf{PB:} 18		
& 
\textbf{TMR:} 8		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Furfur knows many secrets of things that have happened or
will happen on the earth, but he will not speak truthfully unless
compelled by a Ritual of True Speaking.

\end{description}

\subsubsection{MALTHUS (Earl of Death and Havoc)}

\begin{description}

\item[Base Chance] 34%

\item[Lesser Spirits] 26% Malthus appears in the form of a large, filthy buzzard.
His feathers are black as night and his eves are the color of new
blood. Hc carries about him the smell of carrion.

\item[Talents, Skills and Magic] Malthus is a Mage of the College of Necromantic
Conjurations.  He possesses Lhe following Skills: Alchemist, Assassin,
Beast Master, Mechanician, Military Scientist, Navigator, Spy,
Thief.  He is a great general and can build strong forts and organize
warriors to march and to slaughter and can keep them supplied by his
arts when then are in far places.

\item[Movement Rates] (yards per minute): Run: 100; Fly: 500.

\item[Weapons] Malthus may attack with beak and talons in Close Combat. The
BC to Strike with either is 40%.  His beak does + 1 damage and
his talons do + 3.  both are Rank 5.  Malthus may never wear armour
or use weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25		
& 
\textbf{MD:} 10		
& 
\textbf{AG:} 14		
& 
\textbf{MA:} 32
\\
\textbf{EN:} 22		
& 
\textbf{FT:} 33		
& 
\textbf{WP:} 33		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 4		
& 
\textbf{TMR:} 2/10	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Maithus will never be a willing servant.  If he does agree
to serve, he will do so only under the condition that he be allotted a
daily portion of human flesh and will be made free to eat the flesh of
the summoner if the required diet is not forthcoming. He may not be
bound.

\end{description}

\subsubsection{MARAX (The Learned Earl)}

\begin{description}

\item[Base Chance] 27%

\item[Lesser Spirits] 30% Marax appears as a bull with a man's face and is
exceedingly ugly.

\item[Talents, Skills and Magic] Marax is a mwee of the College of Celestial Magics.  He
possesses the following Skills: Alchemist, Beast Master, Healer,
Merchant. Navigator, Ranger. He is especially learned in the use of
herbs and stones to produce magic.

\item[Movement Rates] Runnings: 400 yards per minute.

\item[Weapons] Marax may charge in the same manner as a multi-hex character
in combat.  His horns have a BC of 50% of doing + 6 damage in Close
Combat and are Rank 5.  His hooves have a BC of 50% of doing + 5
damage in Close Combat and are Rank 4. Marax will not use weapons or
wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 60		
& 
\textbf{MD:} 3		
& 
\textbf{AG:} 20		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 34		
& 
\textbf{FT:} 42		
& 
\textbf{WP:} 29		
& 
\textbf{PC:} 24
\\
\textbf{PB:} 3		
& 
\textbf{TMR:} 8		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 6 DP
}\\
\end{tabularx}

\subsubsection{RAUM (Earl of Theft and Destruction)}

\begin{description}

\item[Base Chance] 36%

\item[Lesser Spirits] 30% Raum first appears as a crow with bloody talons.  Upon
the command of the summoner he divests himself of this insubstantial
form and adopts that of a tall, gaunt man.

\item[Talents, Skills and Magic] Raum is a Magician of the College of Earth Magics.  He
possessed the following Skills: Assassin, Beast Master, Healer,
Merchant Ranger, Spy, Thief, Troubadour.  He has the power to cause
the earth to quake and to make fall the walls of cities.  He can make
all but the summoner's sworn enemies love him and be willing to lay
down their lives in the service of the summoner.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Raum has no natural weapons.  He is treated as a human for
purposes of unarmed combat.  Raum will wear any armour and use any
weapons the summoner provides.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 21		
& 
\textbf{MD:} 28		
& 
\textbf{AG:} 26		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 26		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 33		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 14		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Raum has some knowledge of what has been and will be and
will speak truthfully of what he knows about a place he occupies.

\end{description}

\subsubsection{RENOVE (The Beautiful Earl)}

\begin{description}

\item[Base Chance] 30%

\item[Lesser Spirits] 19% Renove first appears as a cloud of red mist, but will
adopt human form at the summoner's behest. He then becomes a beautiful
young boy with a melodious voice which is unbearably sweet to the
listener.

\item[Talents, Skills and Magic] Renove is a mage of the College of Illusions.  He possesses
the following Skills: Assassin, Beast Master, Courtesan, Healer,
Merchant, Navigator, Ranger, Spy, Thief, Troubador.  He also has a
deep knowledge of rhetoric and is a master of spoken and written
tongues.  He has the ability to reconcile all but the bitterest of
enemies and is a great teacher of the young.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Renove has no natural weapons.  He is treated as a human for
purposes of unarmed combat.  Renove will wear any armour and use any
weapons the summoner provides.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 13		
& 
\textbf{MD:} 26		
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 19		
& 
\textbf{FT:} 30		
& 
\textbf{WP:} 35		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 30		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] An hour spent learning any of his Skills in the company of
Renove is equal to a month spent learning the same Skill from any
human teacher and costs half (round down) the normal number of
EP's. However, Renove's tastes are not natural, and he will demand
that the summoner either lie with him or procure him others more to
his taste in return for any lessons.

\end{description}

\subsection{Marquis}

\subsubsection{AMON (The Marquis of Fire)}

\begin{description}

\item[Base Chance] 17%

\item[Lesser Spirits]40% Amon appears as a wolf with a serpent's tail.  In this
form he has the capacity to spit fire balls to a distance of 20 feet.
He may also appear as a raven-headed man (often with a dog's teeth).

\item[Talents, Skills and Magic] Amon is a mage of the College of Fire Magics.  He has no
Skills, but possesses the following powers: He may give a true account
of events that have passed in the place he is standing and may guess
accurately at what is to come in that place.  He also has the power to
cause all but sworn enemies to look upon each other as true friends
and to desire each other's company above others.

\item[Movement Rates] Running: 400 yards per minute.

\item[Weapons] When in wolf form, Amon may use his claws to rend in Close
Combat with a BC of 50% and damage of + 2.  Amon's claws are
Rank 8.  In both wolf and man-raven form, Amon may bite or peck in
Close Combat with a BC of 45% of doing + 3 damage.  Amon's bite
is Rank 8.  Amon's fireballs are useable in Ranped or Melee Combat,
have a BC of 40% do + 6 damage, and are Rank 9.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 26		
& 
\textbf{MD:} 5/20	
& 
\textbf{AG:} 29/23	
& 
\textbf{MA:} 30
\\
\textbf{EN:} 23		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 28		
& 
\textbf{PC:} 31
\\
\textbf{PB:} 6		
& 
\textbf{TMR:} 8		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Where two numbers are given for one of Amon's
characteristics, the rirst applies to him in wolf form, the second in
raven-man form.

\end{description}

\subsubsection{ANDRAS (Marquis of Discord)}

\begin{description}

\item[Base Chance] 35%

\item[Lesser Spirits]  30% Andras appears as an angel with the head of a raven.  He
rides a dire wolf and carries a magical sword aloft in his richt hand.
The sword burns as with fire and drains 2 Fatigue Points when it hits
in addition to any other damage it may In all other respects, Andras'
sword is treated as a normal broadsword.

\item[Talents, Skills and Magic] Andras is a mage of the College of the Mind.  He possesses
the following Skills: Assassin, Beast Master, Military Scientist,
Spy, Thief.  His presence in a party will be an automatic cause of
disruption and discord since he delights in such things and may not be
made to cease encouraging bad feelings even when bound.

\item[Movement Rates] Run: 350yard per minute; Fly: 450  yards per minute.

\item[Weapons] Andras possesses no as a human in unarmed combat.  Andras
may wear specially manufactured armour and carry weapons provided by
the summoner.  He will always use his sword, however, in preference
to other weapons unless it is broken.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28		
& 
\textbf{MD:} 30		
& 
\textbf{AG:} 29		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 27		
& 
\textbf{FT:} 36		
& 
\textbf{WP:} 25		
& 
\textbf{PC:} 29
\\
\textbf{PB:} 31		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

}\\
\end{tabularx}

\subsubsection{ANDREALPHUS (The Beautiful Marquis)}

\begin{description}

\item[Base Chance] 37%

\item[Lesser Spirits]  30% Andrealphus appears initially as a great and beautiful
Peacock, but this is an insubstantial form, and he is without power in
that guise.  At the summoner's command, he will be forced to adopt
the form of a man.

\item[Talents, Skills and Magic] Andrealphus is a member of the College of Illusions. He
possesses the following Skills: Alchemist, Mechanician, Merchant,
Navigator.  He is a great teacher of all forms of measurement and
celestial studies.  An hour in his teaching will increase the pupil's
abilities in counting and measuring fourfold.  Any Skill dependent
upon such talents will also be increased thereby and subsequent
increase in Rank will cost only half EP's (round down).

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Andrealphus has no natural weapons.  He is treated as a
human in unarmed combat.  Andrealphus may wear armour and use weapons
provided by the summoner, but only if they are silvered or gilded.  He
will not use any other kind of armour or weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 26		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 25		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 24		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 25
\\
\textbf{PB:} 17		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] For a payment of the life of a human being, Andrealphus
will turn the summoner into a bird for a period of one hour.  Each
individual transformation or hour of transformation must be purchased
by the sacrifice of an additional human being.  Andrealphus eats the
souls of those sacrificed in this manner. He must be present during
the sacrifice.

\end{description}

\subsubsection{CIMEJES (Marquis of the Dark Continent)}

\begin{description}

\item[Base Chance] 38%

\item[Lesser Spirits]20%
 Cimejes appears as a powerful black man riding upon a
black horse.

\item[Talents, Skills and Magic] Cimejes is a master of the College of Ensorcelements and
Enchantments.  He Possesses the following skills: Alchemist, Beast
Master, Healer, Merchant, Military Scientist, Spy, Thief. He is a
great linguist.  He also has the ability to find things that have
been lost or hidden (especially treasure) anywhere within 100 feet of
him.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Cimejes has no natural weapons.  He is treated as a human in
unarmed combat.  He will wear any armour and any weapons provided by
the summoner, but will first insist that anr armour be lacquered black
to match his skin.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 26		
& 
\textbf{MD:} 24		
& 
\textbf{AG:} 24		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 27		
& 
\textbf{FT:} 38		
& 
\textbf{WP:} 31		
& 
\textbf{PC:} 32
\\
\textbf{PB:} 18		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP

}\\
\end{tabularx}

\subsubsection{DECARABIA (The Marquis in the Pentacle)}

\begin{description}

\item[Base Chance] 39%

\item[Lesser Spirits] 30% Decarbia will first appear as a bright star burning
inside a pentacle of fire and will remain in this insubstantial and
powerless form until ordered to adopt human form by the summoner.
Upon adopting human form, Decarabia will appew naked and remain so
unless clothed by the summoner. He will always be surrounded by a
barely-detectable blue aura whcre his skin is exposed.

\item[Talents, Skills and Magic] Decarabia is a member of the College of The Mind.  He
possesses the following Skills: Alchemist, Beast Master, Merchant,
Ranger and Troubadour.  He can command all types of birds by their
Generic and Individual True Names in the same manner as a Namer,
despite the fact that he is not a member of the College of Naming
Incantations.

\item[Movement Rates] Running: 350yards per minute.

\item[Weapons] Decarabia has no natural weapons.  He is treated as a human
in unarmed combat.  He will wear any armour and use any weapons
provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 24		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 25		
& 
\textbf{MA:} 31
\\
\textbf{EN:} 23		
& 
\textbf{FT:} 34		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 37
\\
\textbf{PB:} 26		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{FORNEUS (Marquis of the Waters)}

\begin{description}

\item[Base Chance] 23%

\item[Lesser Spirits] 29% Forneus appears as a great monster of the sea 30 feet
long and 6 feet in diameter.  His scales are precious stones, his eves
are like fire and the flesh of drowned sailors clinlzs to his teeth.

\item[Talents, Skills and Magic] Forneus is the greatest mage of the College of Water Magics.
He possesses the following skills: Beast Master, Military Scientist
and Navigator. He also is a great linguist and may make men knowing in
tongues.  An hour spent in his company in the learning of a language
is equal to two menths of learning from any non-demonic teacher.

\item[Movement Rates] Swimming: 600 yards per minute.

\item[Weapons] Forneus may roll over one hex characters, using his entire
body as a weapon (but with no chance of falling) as described in the
multi-hex character rules.  In addition, his bite may be used in Melee
or Close Combat to do 2D + 5 damage (total possible damage of 25) with
a BC of 50% and a Rank of 10. Forneus may not wear armour. Since
he has no gripping members, he may not use weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 150		
& 
\textbf{MD:} None	
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 38
\\
\textbf{EN:} 60		
& 
\textbf{FT:} 80		
& 
\textbf{WP:} 35		
& 
\textbf{PC:} 25
\\
\textbf{PB:} 2		
& 
\textbf{TMR:} 12		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 10 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] The Ritual of Summoning Forneus may only be performed while
at sea or on a coastline where Forneus can remain in the water.  He
may never appear on land.

\end{description}

\subsubsection{LERAJE (The Marquis of Archers)}

\begin{description}

\item[Base Chance] 19%

\item[Lesser Spirits]   30% Leraje appears as a man clothed in forest green and
carnling a great bow and quiver.

\item[Talents, Skills and Magic] Leraje is a member of the College of Earth Magics.  He
possesses the foliowing skills: Alchemist, Assassin, Beast Master,
Miliiarv Scientist, NaN,igator, Ranger, Spy, Thief and Troubador,
Hewlso possesses the power to make wounds that have been inflicted bN,
arrows or quarrels putrify.  Add 60 to the Infection Chance \% henever
Leraje or anyone assisted by him inflicts damage 'with arro\"s.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Leraje possesses no natural weapons.  He is treated as
manlike in that respect.  However, the Greatbow he carries may only
be pulled by him.  It has a Range of 450 hexes (2250 feet) and a BC
of 75%. It does + 6 damage.  He will not wear armor, but may carry
other weapons.  Leraje will always use his bow in preference to all
other weapons, however.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28		
& 
\textbf{MD:} 30		
& 
\textbf{AG:} 27		
& 
\textbf{MA:} 28
\\
\textbf{EN:} 24		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 26		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 32		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Leraje is a lover of strife and will attempt to provide
battle whenever possible.  Chance encounters be'ween normally neutral
parties will usually end in a bloody combat to the death if Leraje is
present and this should be taken into account by the GM.

\end{description}

\subsubsection{MARCHOSIAS (Marquis of the Seventh Throne)}

\begin{description}

\item[Base Chance] 25%

\item[Lesser Spirits] 30% Marchosias appears as a great Ox with the wings of a
Gryphon, the tail of a serpent and a breath of fire shaped like a cone
6 feet long and 4 feet wide at the base.  He may take the shape of a
man, but will be insubstantial and powerless in this form.

\item[Talents, Skills and Magic] Marchosias is a practicioner of the College of the Mind.  He
possesses the following skills: Beast Master, Military Scientist and
Ranger.

\item[Movement Rates] (yards per minute): Run: 400; Fly: 300.

\item[Weapons] Marchosias may charge in the same manner as a multi-hex
character in combat.  His horns have a BC of 40% of doing + 4
damage and are Rank 10.  His hooves have a BC of 45% of doing +
5 damage in Close Combat.  Marchosias breath has a BC of 70% in
Melee Combat, but may not be used in Close Combat. It does + 10 damage
and is Rank 3.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 65		
& 
\textbf{MD:} 5		
& 
\textbf{AG:} 18		
& 
\textbf{MA:} 28
\\
\textbf{EN:} 35		
& 
\textbf{FT:} 45		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 25
\\
\textbf{PB:} 6		
& 
\textbf{TMR:} 8/6	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 6 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Marchosias loves a good fight and will always counsel
attack in any situation, often disobeying his summoner in his desire
to charge and rend the enemy.

\end{description}

\subsubsection{NABERIUS (The Valiant Marquis)}

\begin{description}

\item[Base Chance] 21%

\item[Lesser Spirits] 19% Naberius appears as a large black crane which flutters
around the summoner and speaks in a rough and grating voice.

\item[Talents, Skills and Magic] Naberius is a master of the College of the Mind. He
possesses the following skills: Alchemist, Assassin, Healer,
Mechanician, Merchant, Military Scientist, Navigator, Spy, Thief and
Troubadour. Naberius is a great teacher and an hour spent in his
compan), learning a skill is equal to two weeks attempting to learn
the skill with non-demonic aid and allows the student to progress at
half the normal EP cost.

\item[Movement Rates] (yards per minute): Run: 250; Fly: 350.

\item[Weapons] Naberius beak docs + 3 damaze in either Melee or Close
Combatwith a 50% BC. It is Rank 8. Naberius claws may be used to grip
weapons. In addition, they have a BC of 45% of doing + 1 damage in
Close Combat. They are Rank 6. Naberius may not wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 20		
& 
\textbf{MD:} 16		
& 
\textbf{AG:} 18		
& 
\textbf{MA:} 36
\\
\textbf{EN:} 20		
& 
\textbf{FT:} 33		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 27
\\
\textbf{PB:} 22		
& 
\textbf{TMR:} 5/7	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Naberius has the special power to restore lost dignities by
manipulating events to the summoner's advantage.  A summoner who has
lost favor at court might use the power of Naberius to help him regain
his station. The GM must take this special power into account.

\end{description}

\subsubsection{ORIAS (The Celestial Marquis)}

\begin{description}

\item[Base Chance] 33%

\item[Lesser Spirits] 30% Orias appears in the form of a lion, but with the tail
of a serpent. He bears in his front claws a pair of great hissing
serpents and rides a mightily thewed warhorse.

\item[Talents, Skills and Magic] Orias is a mage of the College of Celestial Magics. He
possesses the following Skills: Beast Master, Navigator. He can
reconcile all but the most bitter enemies and any summoner who finds
his favour will be subject to assistance in acquiring honors and
dignities.

\item[Movement Rates] Running: 500 yards per minute.

\item[Weapons] The demon's claws do +4 damage in Melee and Close Combat and
have a BC of 40%. They are Rank 6. The demon's bite does + 3 damape in
Close Combat and has a BC of 45%. It is Rank 4. Orias may not use
weapons or wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 35		
& 
\textbf{MD:} 9		
& 
\textbf{AG:} 30		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 34		
& 
\textbf{FT:} 42		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 34
\\
\textbf{PB:} 18		
& 
\textbf{TMR:} 10		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\subsubsection{PHENEX (The Immortal Marquis)}

\begin{description}

\item[Base Chance] 27%

\item[Lesser Spirits]20%
 Phenex always appears as the great bird, the Phoenix. He
speaks with the voice of a small child and sings sweetly and
hypnotically to any who will listen.

\item[Talents, Skills and Magic] Phenex is a practitioner of the College of the Mind. He
possesses the following Skills: Alchemist, Beast Master, Healer,
Mechanician, Nlerchant Military Scientist, Navigator, Troubadour. The
songs of Phenex will hypnotise his victims (including (he summoner)
who will then willingly enter the demon's mouth to be devoured. Care
must always be taken when dealing with him.

\item[Movement Rates] (yards per minute): Run: 200; Fly: 400.

\item[Weapons] The demon's beak does + 6 damage in Melee or Close Combat
and has a BC of 60%.  It is Rank 5.  The demon's claws do + 4 damage in
Melee or Close Combat, have a BC of 35% and are Rank 3. Phenex
may not wear armour or use weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 75		
& 
\textbf{MD:} 16		
& 
\textbf{AG:} 18		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 26		
& 
\textbf{FT:} 32		
& 
\textbf{WP:} 33		
& 
\textbf{PC:} 27
\\
\textbf{PB:} 34		
& 
\textbf{TMR:} 4/8	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 5 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Phenex is ever hungry for human flesh and will attempt to
sing except when bound so as to draw victims close to his mouth where
they can be devoured.

\end{description}

\subsubsection{SAMIGINA (Marquis of Dead Souls)}

\begin{description}

\item[Base Chance] 15%

\item[Lesser Spirits] 30%
 Samagina appears as a small ass though he may take on
human form if the summoner so orders.

\item[Talents, Skills and Magic] Samagina is a master of the College of Necromantic
Conjurations.  He possesses the following skills: Alchemist Healer,
Mechanician, Merchant, Military Scientist, Navigator, Troubador.  He
has the power to converse with the dead to give an account of all
beings who have died except for servant of the Forces of Liehi.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] When in the form of an ass, Samigina may bite in Close
Combat with a BC of 50% of doing + 2 with a Rank of 10.  He may
also trample with his hooves in Close Combat with a BC of 50% or
use his hooves in Melee Combat with a BC of 40%.  His hooves are Rank
10 and do + 4 damage.  Samigina will never agree to use weapons or
wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25/23	
& 
\textbf{MD:} 5/20	
& 
\textbf{AG:} 20/25	
& 
\textbf{MA:} 30
\\
\textbf{EN:} 30		
& 
\textbf{FT:} 40		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 32
\\
\textbf{PB:} 10		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Where two numbers are given for one of Sanigina's
characteristics, the first is for his ass form, the second for his
human form.

\end{description}

\subsubsection{SAVNOK (Marquis of Corruption)}

\begin{description}

\item[Base Chance] 301%

\item[Lesser Spirits] 50% Savnok appears as a great warrior with the head of a
lion riding a pale horse.

\item[Talents, Skills and Magic] Savnok is a member of the College of Earth Magics.  He
possesses the following Skills: Beast Master and Military
Scientist. In addition, he possesses the power to raise high towers,
castles, and cities from out of earth and stone and to shape for them
furniture and armour. Any wounds he inflicts will automatically become
infected and will not be cured except by magic.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Savnok's bite inflicts + 4 damage in Close Combat and has a
BC of 50%. It is Rank 4.  Savnok may appear wearing any type of
armour.  He prefers axes, spears, and daggers, but will use other
weapons if the summoner provides them.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28		
& 
\textbf{MD:} 30		
& 
\textbf{AG:} 29		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 26		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 33		
& 
\textbf{PC:} 29
\\
\textbf{PB:} 24		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Savnok may start horrible plagues which cause their victims
to break out in rotten sores full of worms and which will ultimately
lead to death unless cured by magic.

\end{description}

\subsubsection{SHAZ (The Thieving Marquis)}

\begin{description}

\item[Base Chance] 31%

\item[Lesser Spirits] 30% Shaz appears as a great bird speaking with a subtle
voice.

\item[Talents, Skills and Magic] Shaz is a member of the College of Illusions.  He possesses
the following Skills: Assassin, MerchanE, Ranger, Spy, Thief.  He
may will an individual to lose his sight or hearing or ability to
comprehend the spoken word if the summoner so desires.  He is an
accomplished thief and a hoarder of useful things (especially fine
horses) which he keeps in his dimension.

\item[Movement Rates] (yards per minute): Run: 150; Fly: 450.

\item[Weapons] Shaz's beak does + 2 damage in Melee or Close Combat, has a
BC of 50%, and is Rank 8.  His talons do + 1 damage and have a
BC of 40%.  They are Rank 5 and may be used in Melee or Close
Combat. Shaz may not use weapons or wear armour.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23		
& 
\textbf{MD:} 18		
& 
\textbf{AG:} 26		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 38		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 18		
& 
\textbf{TMR:} 3/9	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Shaz is an inveterate liar and deceiver.  If commanded to
reveal information or fetch items from his hoard, there is only a
50% chance that he will speak entirely true or fetch the right items
unless he is placed in a triangle and compelled by a Ritual of True
Speaking.

\end{description}

\subsection{Kings}

\subsubsection{ASMODAY (The Infernal King)}

\begin{description}

\item[Base Chance] 13%

\item[Lesser Spirits] 72%
 Asmoday appears as a man with three heads (that of a
man, a bull, and a ram). He has a serpent's tail and breathes a cone
of fire 24 inches long and 12 inches wide at the base.  His feet are
webbed like those of a goose, and he comes riding upon a great
infernal dragon and carries a heavy spear with a banner. The blade of
his spear is poisoned.

\item[Talents, Skills and Magic] Asmoday is a master of the College of the Mind.  He
possesses the following Skills: Alchemist, Beast Master, Military
Scientist, Mechanician, Merchant, Navieator, Healer.  In addition,
Asmoday has the power to locate any treasure hidden within 50 feet of
him.  He also has the power to turn those who see him into beasts for
short periods if they so desire.

\item[Movement Rates] Running: 350 yards per minute.

\item[Weapons] Asmoday may make two horn attacks in Close Combat in
addition to any other attacks.  Each has a BC of 35% and does D + 2
damage.  They are both Rank 10.  In addition, Asmoday's breath of fire
has a BC of 50% in Melee and of 80% in Close Combat and does D + 4
damage.  It is Rank 5.  He may not wear armour, but will accept other
weapons for use in addition to his spear if they are gilded.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25		
& 
\textbf{MD:} 30		
& 
\textbf{AG:} 28		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 36		
& 
\textbf{PC:} 29
\\
\textbf{PB:} 4		
& 
\textbf{TMR:} 7		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Asmoday will not serve the summoner unless he is provided
with a maiden upon whom he may visit his carnal desires once his
service is complete. He will immediately transport her to his own
dimension upon agreeing to serve the summoner.  He may not be bound.

\end{description}

\subsubsection{BAEL (The King in the East)}

\begin{description}

\item[Base Chance] 5%

\item[Lesser Spirits] 66% Bael may appear as either a man, a toad, or a cat, or he
may appear as a man with the tail and claws of a cat and the head of a
toad. He is recognizable by his hoarse sinister voice.  The form he
most often adopts and that he uses in combat is that of the
toad-headed man-cat.  He may adopt a new form at will.

\item[Talents, Skills and Magic] Bael is a master of the College of Ensorcelments and
Enchantments.  He has the power to make himself totally invisible at
will.

\item[Movement Rates] Running: 300 yards PeT minute.

\item[Weapons] Bite has a BC of 40%, does D + 1 damage, and is always
Rank 10.  Claw's have a BC of 45%, do D + 2 damage, and are always
Rank 10.  Bite may only be used in Close Combat.  Claws may be used in
either Melee or Close Combat.  Bael will not wear armour, but use
gilded weapons if the summoner provides them.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25		
& 
\textbf{MD:} 23		
& 
\textbf{AG:} 25		
& 
\textbf{MA:} 30
\\
\textbf{EN:} 23		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 35		
& 
\textbf{PC:} 30
\\
\textbf{PB:} 4		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{BALAM (The Terrible King)}

\begin{description}

\item[Base Chance] 15%

\item[Lesser Spirits] 40% Balam appears as a man with three heads (one of a man,
one of a bull, and one of a ram), the tail of a serpent, and bright
flaming red eyes that can be seen at a great distance.  He appears
riding upon a greate bear, and carries a Goshawk on one fist. He
speaks with a hoarse voice.

\item[Talents, Skills and Magic] Balam is a mage of the College of the Mind.  He possesses
the following Skills: Alchemist, Merchant, Spy, Troubadour.  He also
has the power to make men witty and each hour spent in the company of
Balam counts as one month of training in the Troubadour Skill and no
EP's are expended to advance in this skill as a result of his
teaching.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Balam will make two horn attacks in Close Combat in addition
to other attacks.  Each has a BC of 30% and does D + 2 damage.  They
are both Rank 10.  Balam will refuse to wear armour, but will accept
and use gilded weapons if the summoner provides them.  He will not
always return these weapons.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25		
& 
\textbf{MD:} 23		
& 
\textbf{AG:} 21		
& 
\textbf{MA:} 34
\\
\textbf{EN:} 22		
& 
\textbf{FT:} 32		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 30
\\
\textbf{PB:} 3		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{BELETH (The Mad King)}

\begin{description}

\item[Base Chance] 9%

\item[Lesser Spirits] 85% Beleth always appears in the shape of a large man riding
a pale horse.  He is preceded by music and appears in a rage where at
he will attempt to kill anything not within the summoner's circle
of protection. He breathes a cone of fire 24 inches long and 12 inches
wide at the extreme end.

\item[Talents, Skills and Magic] Beleth is a master of the College of Fire Magics.  In
addition, he is possessed of the talent of causing any individuals of
the summoner's choice (not necessarily humanoids) to submit to the
sexual desires of the summoner until he (the summoner) is sated.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Beleth's breath has a BC of 50% in Melee Combat and 80% in
Close Combat and does D + 4 damage.  It is Rank 3.  Beleth will wear
silver or gilded armour (he much prefers the latter) if provided by
the summoner and will use gilded weapons.  He will claim these as
gifts before departing this plane.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 22		
& 
\textbf{MD:} 23		
& 
\textbf{AG:} 22		
& 
\textbf{MA:} 29
\\
\textbf{EN:} 24		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 26
\\
\textbf{PB:} 10		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{BELIAL (King of Fire)}

\begin{description}

\item[Base Chance] 19%

\item[Lesser Spirits] 50% Belial appears as a two-headed angel and may ride a
chariot of fire. He speaks with a comely voice.

\item[Talents, Skills and Magic] Belial is a mage of the College of Fire. He also possesses
the following Skills: Merchant, Beast Master.  He will cause all but
sworn enemies of the summoner to look upon the summoner as their true
friend as described in E \& E Charm.

\item[Movement Rates] (yards per minute): Run: 300; Fly: 400

\item[Weapons] Belial has no natural weapons.  Belial will demand gifts of
gilded armour and weapons before he can be bound to service. He will
never return these to the summoner. He is treated as a human for
purposes of unarmed combat.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 28		
& 
\textbf{MD:} 30		
& 
\textbf{AG:} 23		
& 
\textbf{MA:} 31
\\
\textbf{EN:} 25		
& 
\textbf{FT:} 37		
& 
\textbf{WP:} 30		
& 
\textbf{PC:} 30
\\
\textbf{PB:} 28		
& 
\textbf{TMR:} 6/8	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] The services of Belial must be paid for in advance in the
only coin he will accept: human suffering.  As part of the Ritual of
Summoning Belial, the summoner must torture to death a human being
(preferably an innocent child or a maiden since Belial prefers the
suffering of innocents and may refuse an offering of a warrior). He
may not be bound.

\end{description}

\subsubsection{PALMON (King of Dominations)}

\begin{description}

\item[Base Chance] 7%

\item[Lesser Spirits]100%
 Palmon appears as a man with a crown of light on his's
head and is often seen riding a dromedary.  His appearance will always
be accompanied by ,he sound of trumpets and symbals played by an
invisible band of demonic musicians.  His voice is heard as a roaring
at first, but he can be compelled to speak understandably as a man.

\item[Talents, Skills and Magic] Palmon is a master of the College of the Mind.  His Skills
include: Alchemist, Astrologer, Beast Master, Healer, Mechanician,
Merchant, Military Scientist, Navigator, Ranger, Troubadour.

\item[Movement Rates] Running: 300 yards per minute.

\item[Weapons] Palmon has no natural weapons.  He is treated as a human
for purposes of unarmed combat.  He wiil use weapons and armour that
are gilded if these are provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 22		
& 
\textbf{MD:} 24		
& 
\textbf{AG:} 24		
& 
\textbf{MA:} 34
\\
\textbf{EN:} 23		
& 
\textbf{FT:} 34		
& 
\textbf{WP:} 35		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 25		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Palmon may be accompanied by two kings (Abalim and
Labal) who assist and protect him. Each will possess
characteristics 2 to 3 points lower than Palmon's.  Thev will possess
the same Skills and Magics, but only at Rank 10.  Neither Palmon nor
the two Kings may be bound.

\end{description}

\subsubsection{PURSON (The Comely King)}

\begin{description}

\item[Base Chance] 11%

\item[Lesser Spirits]22%
 Purson appears as a muscular and well proportioned man,
but with the face of a lion and carries a viper in his hand.  He is
preceded by trumpets and often appears riding upon a bear.  He may
also appear in insubstantial form as a fairy.

\item[Talents, Skills and Magic] Purson is a member of the College of Namers and knows the
Generic True Name of all things on the Mundane Plane.  He has an 80%
chance of knowing the Individual True Name of any beinp he encounters.
He possesses the following Skills: Alchemist, Astrologer, Healer,
Mechanician, Merchant, Military Scientist, Navigator.

\item[Movement Rates] (yards per minute): Run: 300; Fly: 500.

\item[Weapons] Purson's bite has a BC of 40% of doing + 2 damage in Close
Combat and is Rank 6-8.  Purson will wear gilded armour and use
gilded weapons if they are provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 21		
& 
\textbf{MD:} 23		
& 
\textbf{AG:} 23		
& 
\textbf{MA:} 33
\\
\textbf{EN:} 21		
& 
\textbf{FT:} 35		
& 
\textbf{WP:} 34		
& 
\textbf{PC:} 35
\\
\textbf{PB:} 12		
& 
\textbf{TMR:} 6/10	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{VINE (The Lion-Headed King)}

\begin{description}

\item[Base Chance] 12%

\item[Lesser Spirits] 36% Vine always appears as a lion-headed man, sometimes
riding a great black horse and bearing a viper in his hand.

\item[Talents, Skills and Magic] He is a mage of the College of Naming Incantations. He also
possesses the following Skills: Alchemist, Merchant, Military
Scientist, Navigator, Mechanician. Vine has the ability to control
weather, create earthquakes, detect the presence of magic, and
transport and bind stone into magical dwellings.

\item[Movement Rates] Running: 300yardsper minute.

\item[Weapons] Vine's bite has a BC of 40% of doing + 3 damage in Close
Combat and is always Rank 10. Vine will wear gilded armour and use
gilded weapons if these are provided by the summoner.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 23		
& 
\textbf{MD:} 25		
& 
\textbf{AG:} 25		
& 
\textbf{MA:} 35
\\
\textbf{EN:} 22		
& 
\textbf{FT:} 33		
& 
\textbf{WP:} 32		
& 
\textbf{PC:} 36
\\
\textbf{PB:} 8		
& 
\textbf{TMR:} 6		
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\subsubsection{ZAGAN (The Winged King)}

\begin{description}

\item[Base Chance] 17%

\item[Lesser Spirits] 33% Zagan appears as a great bull with Gryphon's wings, but
may choose to adopt human form at will.

\item[Talents, Skills and Magic] Zagan is a master of the College of Illusions. He possesses
the following Skills: Alchemist, Assayer. He can change blood or water
to wine and wine to blood at will. He can shape all metals into coins
of the realm. He makes men witty and one hour in his company is equal
to 3 weeks experience in learning the Troubadour Skill.  Progress in
this Skill costs only half the normal EP cost if the summoner is taught
by Zagan. He makes fools wise (see Comments).

\item[Movement Rates] (yards per minute): Run: 400; Fly: 300.

\item[Weapons] Zagan may charge in the same manner as a multi-hex
character in combat. His horns have a BC of 40% of doing D + 4
damage and are Ranked at 10.  His hooves have a BC of 45% of doing
D + 5 damage in Close Combat and are Ranked 8.  Horns may be used in
Melee or Close Combat.  Zagan wail agree to wear gilded armour while
in his human form and will employ gilded weapons in that form.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 65/30	
& 
\textbf{MD:} 5/25	
& 
\textbf{AG:} 20/23	
& 
\textbf{MA:} 30
\\
\textbf{EN:} 35		
& 
\textbf{FT:} 40		
& 
\textbf{WP:} 28		
& 
\textbf{PC:} 25
\\
\textbf{PB:} 6		
& 
\textbf{TMR:} 8/6	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 6 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Any character in the presence of Zagan may have his
Willpower and Magical Aptitude increased by 10 and his Perception
increased by 5 at the discretion of the summoner.  This effect lasts
as long as the character remains in Zagan's presence. Where two
numbers are given for a characteristic, the first is for the bull
form.

\end{description}
\end{multicols}
