\begin{multicols*}{2}

\setlength\columnseprule{0.2mm}

\section{Creatures of night and shadow}
The species protrayed herein include those primarily connected with
night and shadow and the Powers of Darkness.  They include Weres,
Dopplegangers, Bats, and Dire Wolves.  Other species which may be
nocturnal hunters or may be at their most powerful at night are not
included in this heading and are not af- fected by magic designed to
call, communicate with, or control Creatures of Night and Shadow.
Only the species listed herein are affected by these types of magic.

\subsubsection{Bat}

\begin{description}
\item[Natural Habitat] Caverns Uncommon

\item[Number]  1-400 (100)

\item[Description] Bats are rodents with leathery wings. They are
greyish-black in color. They are found, in dark places only as they
are afraid of light and fire.

\item[Talents, Skills and Magic] Bats can determine directions in the dark by emitting
high-freq "pips" which reflect off of the surrounding walls. They
have no magical abilities or other special talents and are not tool
users.

\item[Weapons] Bats can only attack in Close Combat, where their bite has a
Base Chance of 40\% and does [D - 7] damage. A character bitten by a
bat has a 10\% chance of contracting rabies or some other loathsome
disease in addition to the possibility of infection.

\item[Movement Rates]  Flying: 500

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  2-3
& 
\textbf{MD:}  None
& 
\textbf{AG:}  18-21
& 
\textbf{MA:}  None
\\
\textbf{EN:}  1-2
& 
\textbf{FT:}  2-3  
& 
\textbf{WP:}  6-8
& 
\textbf{PC:}  14-17
\\
\textbf{PB:}  7-9
& 
\textbf{TMR:}  10
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\subsubsection{Dire wolf}

\begin{description}
\item[Natural Habitat] Woods, Rough, Plains, Field, Caverns Rare

\item[Number] 1-10 (3)

\item[Description] Dire wolves are wolves the size of ponies. Dire wolves
will be in league with the powers of darkness and will have almost
human intelligence. Their corpses disappear back into hell (from which
they are said to have sprung originally) if they are killed.

\item[Talents, Skills and Magic] Dire wolves have no special abilities but they are
exceptional trackers. They also never forget a smell. They may possess
knowledge of a limited number of spells of the College of
Ensorcelments and Enchantments or the College of the Sorceries of the
Mind.

\item[Weapons] Dire wolves can attack, either in Melee Combat or in Close
Combat with their huge teeth (Base Chance of 65\% [D + 6] damage).

\item[Movement Rates]  Running: 400

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  22-25
& 
\textbf{MD:}  24-26
& 
\textbf{AG:}  18-22
& 
\textbf{MA:}  8-12
\\
\textbf{EN:}  20-23
& 
\textbf{FT:}  25-30
& 
\textbf{WP:}  8-12
& 
\textbf{PC:}  19-23
\\
\textbf{PB:}  4-8
& 
\textbf{TMR:}  8
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Fur absorbs 5 DP
}\\
\end{tabularx}

\subsubsection{Doppelganger}

\begin{description}
\item[Natural Habitat] Caverns, Rough, Crypts, Woods Very Rare

\item[Number] 1-8 (1)

\item[Description] In their natural form, doppelgangers appear humanoid,
with thin arms, and sharp teeth and claws. They have thick, rubbery,
grey skin, and glowing eyes.

\item[Talents, Skills and Magic] Doppelgangers have the ability to form themselves to look
and act like any humanoid creature that is approximately their size
and body weight (100-250 lbs.).  Even the clothes and equipment of a
person can be imitated, although magical properties will not, of
course, adhere to the duplicates. Doppelgangers read the minds of the
people they will later imitate so as to learn things they will need to
know to duplicate the person properly. There is a 10\% chance that a
doppelganger's imitation will be detectable. See the rules governing
multi-sense illusions in the College of Illusions.

\item[Weapons] In their natural form, doppelgangers can make one attack
with their bite (Base Chance of 50\% [D + 3] damage). Once having
imitated something, they right as it does / did, using the same
weapons at a Rank two less than that achieved by the character being
imitated.

\item[Movement Rates] Running: 250

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  12-16
& 
\textbf{MD:}  14-17
& 
\textbf{AG:}  16-18
& 
\textbf{MA:}  20-25
\\
\textbf{EN:}  8-10
& 
\textbf{FT:}  15-20
& 
\textbf{WP:}  22-25
& 
\textbf{PC:}  15-17
\\
\textbf{PB:}  4-8
& 
\textbf{TMR:}  5
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 4 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Once a doppelganger has imitated someone, they will try to
kill that individual, attacking by surprise. They will then replace
that individual until they try the same maneuver with another member
of the party.

\end{description}

\subsubsection{Weres}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Weres are humans or humanoids who are afflicted with a
specific disease: Lycanthropy. They are adversely affected by the full
moon and on nights when Luna (the primary moon in worlds with more
than one) is full they change shape, becoming one of 5 types of
creature. The were may become a Bear, Boar, Tiger, Snake or wolf. The
actual type of beast the character becomes will depend on the type of
beast from which he contracted his illness. The only way that a
character can become a lycanthrope is if he is bitten for at least I
point of effective damage by a character who is already a lycanthrope
while that character is in his beast form. The victim will then during
the next full moon (and all succeeding full moons until he is killed)
assume the shape of the beast that bit him (wolf, snake, etc.). If a
character is killed by a lycanthrope, he does not become a lycanthrope
himself; instead, he usually becomes dinner.  A lycanthrope in his
were forin is always a beast. He has the characteristics of the
strongest specimen of the species whose form he has assumed. A
werewolf would have the highest possible characteristics for a Dire
Wolf, for example. A weresnake assumes the form of a Python, but with
the bile (and venom) of the King Cobra. Werebears, wereboars and
weretigers will have the highest possible characteristics for their
type of beast (+2 to PS, EN and FT, usually): when in their beast
form, Weres are semi-intelligent. Their human side is suppressed and
the character is played bN, the GM. Weres are primarily concerned
with feeding and simple survival. They will spend the entire period
when the moon is full either laid up in a lair or hunting. They will
attack the character's friends as readily as any other prey.  Once a
were returns to his normal human shape, he will be enfeebled for [D10 +
4] hours, desiring only to sleep. Upon waking, he will remember A,hat
he did in his beast form (usuallv with regret).  A character can be
cured of lycanthropy in exactly the same manner as he is cured of a
major curse.  When in beast form, a were can only be harmed or killed
by magic or by silvered weapons.  Pre-pubescent and post-menopausal
individuals never suffer the effects of lycanthropy (though they may
contract the disease).

\end{description}
\end{multicols*}
