\begin{multicols}{2}

\setlength\columnseprule{0.2mm}

\section{Summonables}
This section includes those entities which normally do not appear on
this plane, but are summonable from their own dimension.  These
entities include: demons, devils, djinni, efreeti, elementals (earth,
air, fire and water), hellhounds, imps, incubi, and succubi.  Devils,
demons, imps, incubi, and succubi are described in the College of
Greater Summonings which is concerned exclusively with the summoning
and control of these entities.  The other summonable entities are
described in this section.

Generally, an entity summoned from another dimension will require from
1 to 10 Pulses (5 to 50 seconds) to materialise once summoned.  In
some cases, entities appear in a non-corporeal form and must be
commanded to take on substantial form.  An insubstantial entity has no
power on this plane, but cannot be harmed. An entity in substantial
form, however, can be harmed and can harm others.


\subsubsection{Djinn}

\begin{description}
\item[Natural Habitat]  Other Planes Very Rare

\item[Number] 1

\item[Description] A djinn is a male humanoid about ten feet tall, with
darkly tanned skin and a black beard.

\item[Talents, Skills and Magic] A djinn has all the abilities of an air elemental summoned
by a magician with Rank 8 in the Ritual of Summoning Air Elementals.
In addition, a djinn can create matter weighing up to 200 pounds.
This ability can be used once per day.  The material created will have
varying permanence, depending upon its solidity. Wood, cloth and other
light materials will last for as long as a week, while steel, gold, or
other hard substances will last only for hours. Djinni cannot be
harmed by non-magical weapons.

\item[Weapons] Djinni prefer large, curved weapons, with a double-sized
scimitar being the weapon most commonly used. Add 10 to the Base
Chance of any weapon a djinn uses, and double damage done because of
its size. A djinn will have Rank 4 or better with a scimitar.

\item[Movement Rates]  Flying: 700; Running: 350

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  35-40
& 
\textbf{MD:}  20-24
& 
\textbf{AG:}  17-21
& 
\textbf{MA:}  12-18
\\
\textbf{EN:}  20-25
& 
\textbf{FT:}  30-38
& 
\textbf{WP:}  20-25
& 
\textbf{PC:}  16-21
\\
\textbf{PB:}  11-14
& 
\textbf{TMR:}  14/7
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 6 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Like efreet, if a djinn is discovered on this plane it will
either be in the service of some character or else be trapped in some
sealed container. If the container is opened, the djinn will service
whoever released it as a slave, provided that it is not asked to do
anything suicidal. Having a djinn in a party can occasionally cause
problems because other creatures, such as efreeti (who are deadly
enemies of the djinni) will be attracted to the party.

\end{description}

\subsubsection{Efreet}

\begin{description}
\item[Natural Habitat]  Other Planes Very Rare

\item[Number] 1

\item[Description] Efreeti are creatures of flame that have been trapped on
this plane. They are humanoid in form, about 12 feet tall, and their
dark-brown skin flickers as if it were about to burst into flame at
any moment.

\item[Talents, Skills and Magic] An efreet has all the talents and abilities of a fire
elemental summoned at Rank 12. In addition an efreet can grant three
wishes under certain conditions (see below). Efreeti can also become
invisible at will. Efreeti cannot be harmed by non-magical attacks.

\item[Weapons]  Efreeti use edged weapons of ordinary shape but
twice normal size and weight. Improve the weapon's Base
Chance by 10, and double any damage done. An efreet will have
Rank 4 or higher with whatever weapon it uses. In Close
Combat an efreet can immolate an opponent. Such an attack
has a Base Chance of 65\%, and does [D + 10] damage.

\item[Movement Rates]  Flying: 750; Running: 350

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  40-45
& 
\textbf{MD:}  17-20
& 
\textbf{AG:}  15-18
& 
\textbf{MA:}  12-18
\\
\textbf{EN:}  28-32
& 
\textbf{FT:}  45-50  
& 
\textbf{WP:}  23-29
& 
\textbf{PC:}  16-21
\\
\textbf{PB:}  8-10
& 
\textbf{TMR:}  15/7
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Skin absorbs 6 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] If discovered on this plane, an efreet will either be in
the service of some character, or trapped in a sealable container,
such as a bottle. If the latter is the case and the finder of the
efreet releases it from its captivity, there is a 95\% chance that the
efreet will serve him willingly. The other 5\% of the time the efreet
will attack. If the efreet decides to serve, it will either grant
three wishes, or become the character's slave for a period of 1001
days (at the liberator's option). If three wishes are chosen, they
must be used by the end of the day on which the efreet was released.
If service is chosen, the efreet will perform any task asked of it
that is not obviously suicidal. If a request involves physical danger
to the efreet is made of the efreet, there is a 10\% chance (+10 for
each additional request) that the efreet will rebel against its master
and attack him.  Efreeti are creatures of fire, and thus they will not
go to or be found in frigid climates or underwater.

\end{description}

\subsubsection{Hellhound}

\begin{description}
\item[Natural Habitat]  (See Below) Very Rare

\item[Number]  1-10 (5)

\item[Description] Hellhounds have the outward appearance of a Dire Wolf
with reddish-brown fur and bright red eyes.


\item[Talents, Skills and Magic] Hellhounds, will have some ability in the powers of either
the College of Illusions or of the Sorceries of the Mind. They are
able to breath fire in a cone 20 feet long by 10 feet wide at the
base, doing D + 8 damage to all within Range.  The cone will also
ignite any inflammables. While on the Tactical Display it takes a
Pulse for a Hellhound to breath fire.

\item[Weapons] In Close Combat Hellhounds attack with their bite (Base
Chance of 65\%, [D + 6] damage). They cannot attack in Melee or
Ranged Combat.

\item[Movement Rates]  Running: 450

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  14-16
& 
\textbf{MD:}  18-21
& 
\textbf{AG:}  19-22
& 
\textbf{MA:}  12-20
\\
\textbf{EN:}  10-11
& 
\textbf{FT:}  20-22  
& 
\textbf{WP:}  17-19
& 
\textbf{PC:}  25-30
\\
\textbf{PB:}  4-6
& 
\textbf{TMR:}  9
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Fur absorbs 6 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Hellhounds do not normally appear on this plane.  They are
usually brought here by a demon to guard some treasure. It is in this
general capacity that hellhounds are normally found, although they are
occasionally release to wander the world and cause misery as they
may. Hellhounds are highly evil. They love to kill and maim, burn
crops, and ruin buildings.

\end{description}

\subsubsection{Imps}

\begin{description}
\item[Natural Habitat] Seventh Plane Very Rare

\item[Number]  1

\item[Description]Imps appear as three to four foot high humanoids, but with horns,
bat's wings and barbed tails.  They tend to be muscular, but
pot-bellied.  Their skin may be blue, green, or purple, but is usually
red.  They have yellow eyes, flecked with silver and their teeth
(including prominent incisors) are made of silver.  Imps may also
appear as goats, spiders, bats or large rats, but these are
insubstantial forms. The imp cannot use his powers in this plane
except in his natural and substantial form.

\item[Talents, Skills and Magic]Imps may be members of any Colleges of Magic except the Entities as
determined by the GM.  Imps can only be harmed by silvered weapons or
magic. Cold iron does not affect them.  They have no skills.  They
have Rank 5 with all magic.

\item[Weapons]Imps may make a horn attack during Close Combat with a Base Chance of
20\% and possible damage of [D - 1].  Imp horns may have a Rank
of between 1 and 10 (controlling Player should roll D10 to determine
Rank).  Imps may also make an attack in either Melee or Close Combat
with their long tail. Its tip contains a sting with a quick acting
nerve poison (2--4 Damage Points per Pulse). The sting itself is
treated like a dagger.  Base Chance with tail is 30\%.  Imps may
use other weapons, but may not wear armour.


\item[Movement Rates]Running: 350 --- Flying: 500

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 15--20	
& 
\textbf{MD:} 13--17	
& 
\textbf{AG:} 16--22	
& 
\textbf{MA:} 22--26
\\
\textbf{EN:} 14--16	
& 
\textbf{FT:} 22--28	
& 
\textbf{WP:} 28--32	
& 
\textbf{PC:} 16--25
\\
\textbf{PB:} 3--5	
& 
\textbf{TMR:} 7/10	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 3 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]Imps are dangerous servants in that they have a cruel sense
of humour which they are not averse to visiting upon their master and
friends.  They may not be bound or forced to speak true, so there is
no way to really control their actions.  They will lie, steal, sow
dissension, or play practical jokes (magical or non-magical) on the
average once per day (total).  They will not knowingly get their
master killed or maimed but are otherwise not overly considerate of
the master's wellbeing.  They will use their powers on the master's
behalf when it is demanded of them (though even here their sense of
humour may have an effect) and will fight the master's enemies.  Imps
may be called upon as needed and then returned to their own dimension.
Given their warped personalities, the wise master will keep them in
their own dimension as much as possible.  Imps have a 7\% Base
Chance of knowing past, present and future events, but will seldom
speak truthfully of what they know.

\end{description}

\subsubsection{Half-Devil}

\begin{description}
\item[Natural Habitat]Seventh PlaneVery Rare

\item[Number] 1

\item[Description]Half-Devils appear as small (around 3ft) muscular humanoids with
membranous wings, a small, hard, curved horn and vicious teeth.  They
tend to have dark colourations, murky and speckled browns and blacks
most common.  Their tails are sharp, barbed and poisonous.
Half-Devils, when on this plane, can appear to be an animal but cannot
use their spell magic or rituals while in this form.  They may,
however, use talents.  it takes 5 seconds for the change to take
place.

\item[Talents, Skills and Magic]Half-Devils may be members of any Thaumaturgy or Elemental
college.  Being creatures of mana they can only be harmed by silvered
or magical weapons, cold iron does not affect their casting, and they
know all spells, rituals and talents of their college at rank 5.  Note
that this does not include Ward, Investment, or other spells and
rituals not intrinsic to their college.

\item[Weapons]Half-Devils may make a Horn attack or a Bite as well as Sting with
their tail in Close Combat.  They may use their Tail in Melee but at
-20\%.  Their Bite has a Base Chance of 30\% and does +1
damage (A class Critical).  Their Horn has a Base Chance of 25\%and does +1 damage (A class criticals and +20\% if used in
Charge).  Both can be rank 1 to 5. Their Tail has a Base Chance of
50\% doing -2 damage. It has a quick-acting nerve poison in it
that does 2 damage points per pulse. (Poison effects are not
cumulative.)  The victim can resist the effects of the poison by
rolling under normal endurance when fatigue has been removed, or by
rolling under twice normal endurance when their endurance is reduced
to 3 or less.  A Half-Devil's tail is always rank 4.  A Half-Devil may
use weapons (unranked), but will never wear armour.


\item[Movement Rates] Running : 200  Flying : 400

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 12--16	
& 
\textbf{MD:} 7--11	
& 
\textbf{AG:} 9--13	
& 
\textbf{MA:} 14--18
\\
\textbf{EN:} 12--16	
& 
\textbf{FT:} 17--21	
& 
\textbf{WP:} 11--15	
& 
\textbf{PC:} 10--14
\\
\textbf{PB:} 4--6	
& 
\textbf{TMR:} 4/8	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]Upon making a Lesser Pact with a Demon, that Demon assigns a
familiar to the adept.  This familiar is a Half-Devil.  The GM should
randomly roll the values of the above stats and then roll on the
following table for the college.

1 E \& E
2 Mind
3 Illusion
4 Namer
5 Air
6 Water
7 Fire
8 Earth
9 Celestial
10 Demon's Choice

The adept may choose the location of the witchmark and a ritual
drawing blood binds the Half-Devil to that adept.  The Half- Devil
will then always aim to please his master (although his understanding
of the world may hamper this) but will severely dislike (and hope to
avoid) any period of separation from the adept.  Should the adept or
familiar die, the bond is broken and the Half-Devil will immediately
return to the seventh plane (where the adept may pick it up later if
it was s/he that died).  If the Half-Devil dies, a new one may be
issued 5 years after the previous one was issued.  Half-Devils, during
ritual blood-letting, immediately acquire all spoken languages that
the adept knows at one less rank.  Half-Devils cannot read or write.
Half-Devils have no other skills, although in their animal form they
may have abilities, eg.  a weasel can stare down basilisks.  Note that
the familiar loses one fatigue per hour that it doesn't feed
(irrelevant of form).  Hence, the adept must feed it at least once a
day (using four fatigue which the Half-Devil enjoys (obviously).  In
animal or the Half-Devil can understand the languages it understands
in natural form, but cannot speak back.

\end{description}

\subsubsection{Devils}

\begin{description}
\item[Natural Habitat]Seventh PlaneVery Rare

\item[Number] 1

\item[Description]Devils appear as large muscular humanoids with
massive bat wings, horns and fangs.  They tend to have carmine skin
and green eyes.  Their tails are sharp and barbed and contain a
poisonous sting.

\item[Talents, Skills and Magic]Devils, like imps, may be members of any College (except the
Entities) as determined by the GM. Devils can only be harmed by
silvered weapons or by magic.  Cold iron does not affect them.  They
have no skills, but possess Rank 10 with all magic of their College.
Devils know a little of the past, present, and future of things
(7\% chance of accuracy) and will speak truthfully to the best of
their knowledge in exchange for their freedom.  They may not be bound
or subjected to a Ritual of True Speaking.

\item[Weapons]Devils may make a horn attack or bite in Close Combat and may sting
with their tail in Melee or Close Combat.  Their bite has a Base
Chance of 30\% and does [D + 2] damage.  Their horns have a Base
Chance of 25\% and do [D + 2] damage.  Both are Rank 1--10
(controlling Player rolls D10 to determine exact Rank).  Their tail
has a Base Chance of 30\% in Melee and a Base Chance of 50\%Close Combat and does [D - 1] damage.  It has a quick-acting nerve
poison in it that does 3--5 Damage Points per Pulse.  A devil's tail
is always Rank 8.  A Devil may use weapons, but will not wear armour.

\item[Movement Rates] Running : 400  Flying : 500

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 25--30	
& 
\textbf{MD:} 15--20	
& 
\textbf{AG:} 18--24	
& 
\textbf{MA:} 28--34
\\
\textbf{EN:} 25--30	
& 
\textbf{FT:} 35--40	
& 
\textbf{WP:} 24--28	
& 
\textbf{PC:} 22--26
\\
\textbf{PB:} 2--3	
& 
\textbf{TMR:} 8/10	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 5 DP

}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]Devils are extremely powerful companions, but tend to serve
only with reluctance.  They will attempt to bargain for their freedom
in exchange for the use of their magic powers in any situation that
does not involve immediate danger of death or maiming of their master.
They are required to assist the master only in situations of imminent
danger and so will usually be able to bargain successfully early in
their service.  Their power to see things past, present, and future is
the same as that of imps (7\%).  They will never use this talent
except in exchange for their release from service and return to their
own plane.

\end{description}

\subsubsection{Succubi \& Incubi}

\begin{description}
\item[Natural Habitat]Seventh PlaneVery Rare

\item[Number] 1

\item[Description]Succubi are lesser spirits of the seventh plane.  They always appear
in the form of tall, beautiful women with flowing hair which hides
their small horns.  They also possess membranous wings shaped like
bat's wings that fold down so as to be all but invisible when not in
use.  They have retractable fangs which they sometimes use to open the
veins of those individuals whose blood they desire to drink.

\item[Talents, Skills and Magic]Succubi possess Rank 10 in all spells and talents of the College of
the Mind.  In addition they possess the following skills: assassin,
courtesan, spy, thief and troubadour. All skills of Succubi are Rank
10.  Succubi possess the following characteristics:

\item[Weapons]Succubi may attempt to bite during Close Combat with a Base Chance of
40\%.  Their bite does [D + 2] damage to Fatigue each Pulse.
This drain continues so long as they continue to bite (they need not
make another Strike Check).  The Succubi will cease biting only when
they have become stunned or unconscious or when their victim is dead.
They never inflict specific Grievous Injuries, but may do Endurance
damage.  Their bite may be Rank 1--10 (roll D10 to determine its exact
Rank).  Succubi do not wear armour and carry only small weapons that
are easily concealed.  They prefer envenomed weapons to all others.

\item[Movement Rates]Running 350 yards per minute; Flying 400 yards per minute

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:} 16--21	
& 
\textbf{MD:} 24--28	
& 
\textbf{AG:} 28--32	
& 
\textbf{MA:} 26--30
\\
\textbf{EN:} 12--16	
& 
\textbf{FT:} 24--28	
& 
\textbf{WP:} 26--30	
& 
\textbf{PC:} 24--28
\\
\textbf{PB:} 26--34	
& 
\textbf{TMR:} 7/8	
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:} 2 DP
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments]Succubi will, at the behest of their summoner, attempt to meet,
seduce, and then murder any individual of the summoner's choice.  They
will pursue their victim until returned to their own plane or until
they succeed in killing him. Their speciality is to gain the
confidence of the victim and then murder him in his bed.  They cannot
be harmed except by magic or by the touch of the Book of Toth, an
ancient tome of magical lore found only in older and seldom frequented
libraries.  This rare volume immediately and permanently dissipates
them, forcing them back to their own plane.

Incubi are exactly like Succubi, except that they appear only in the
male form, rather than as females.  They will tend to be 2--3 points
stronger and will have 1--2 points less Endurance than Succubi, but
will other wise be the same.  They also are immune to all but magic
and the Book of Toth.

\end{description}

\subsection{Elementals}
Elementals are the physical manifestations of the four primary I
elements: fire, earth, air, and water. They do not normally exist on
this plane, but are summoned by members of the Elemental
Colleges. They will always be hostile to their summoner, and will
attempt to kill him if they are released from his control.  Elementals
are impervious to attacks made with non-magical weapons. Magic does
affect them. Each is vulnerable to its opposite element and can be
damaged by attacks involving that opposite. Water and fire are
opposite members, as are earth and air.  An elemental's Endurance,
Fatigue, and Strength vary according to it summoner's Rank with the
Ritual of Summoning Elementals. Endurance and Fatigue vary as
described in 42.7, while an elemental's Physical Strength equals a
base number for each of the elementals plus five for each Rank the
summoner has attained.

\subsubsection{Air elemental}

\begin{description}
\item[Natural Habitat] Other Planes Very Rare

\item[Number] 1

\item[Description] Air elementals appear as a whirlwind of a height equal
to their Endurance. They have facial features, which will appear on
one side of the whirlwind.

\item[Talents, Skills and Magic] Air elementals can predict weather with complete accuracy.
They can increase the speed of winds up to the equal of their
endurance, blowing in a line ten times their endurance in width
(measured in feet). They can also create windstorms per the Windstorm
Spell (College of Air Magics) at a Rank equal to their summoner's Rank
plus 4. They do not need to cast the spell, but rather they can cause
the windstorm just by willing it.


\item[Weapons] Air elementals can materialise their fists to strike at
their opponents. They can take two attacks in the same Pulse without
penalty. The attacks have a Base Chance of 65\% and do D10 damage, +
1 per Rank of their summoner.

\item[Movement Rates] Flying: 500 + (50 x Summoner's Rank)

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  Base 15   
& 
\textbf{MD:}  25-30
& 
\textbf{AG:}  32-37
& 
\textbf{MA:}  None
\\
\textbf{EN:}  5-50
& 
\textbf{FT:}  10-85
& 
\textbf{WP:}  14-18
& 
\textbf{PC:}  19-23
\\
\textbf{PB:}  10-12
& 
\textbf{TMR:}  10 + (Adept's Rank)
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}   None
}\\
\end{tabularx}

\subsubsection{Earth elemental}

\begin{description}
\item[Natural Habitat]  Other Planes Very Rare

\item[Number] 1

\item[Description] Earth elementals appear as humanoids with dirt-coloured
skin and gravelly eyes. They will be as tall, in feet, as their
Endurance.

\item[Talents, Skills and Magic] Earth elementals can sink into any substance made of rock or
dirt, and remain hidden with only 5\% chance of detection. They have a
chance equal to their Endurance of knowing the location of any
intelligent creature standing on earth or rock. They can move through
the earth (without creating a tunnel or disturbing it in any way)
earth elementals can destroy stone structures by boring through stone
at the rate of two cubic feet per Round for each point of Physical
Strength available to the Elemental. Rock bored through in this way
will be turned to dust incapable of supporting anything.

\item[Weapons] Earth elementals pound their adversaries with their massive
fists. They can attack twice in the same Pulse without penalty, and do
D + 6 damage if they hit, plus one additional point for each Rank of
their summoner.

\item[Movement Rates] Running: 200; Tunnelling: 100

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  Base 30
& 
\textbf{MD:}  10-15  
& 
\textbf{AG:}  12-17
& 
\textbf{MA:}  None
\\
\textbf{EN:}  5-50
& 
\textbf{FT:}  10-85
& 
\textbf{WP:}  14-18 
& 
\textbf{PC:}  16-20
\\
\textbf{PB:}  5-8
& 
\textbf{TMR:}  4/2
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  8 DP
}\\
\end{tabularx}

\subsubsection{Fire elemental}

\begin{description}
\item[Natural Habitat]  Other Planes Very Rare

\item[Number]  1

\item[Description]  Fire elementals appear as leaping flames of a
height in feet equal to their endurance. The flames are
humanoid in form.

\item[Talents, Skills and Magic] Any flammable substances (exclusive of human-size or larger
beings) that a Fire Elemental touches bursts into flames, while
non-flammable substances will experience a rise in temperature equal
to the Elemental's Endurance for every 10 seconds the Elemental is in
contact with the substance. Elementals can instantly extinguish any
normal fire within 100 feet. They can create a wall of smoke, (as per
the spell from the College of Fire Magics) at a Rank equal to that
summoner's Rank plus ten. They can also produce a wall of fire (as per
the spell from the College of Fire Magics) at a Rank equal to their
summoner's Rank plus five. Both of the above two walls can be created
as talents, without the necessity of preparing or casting a
spell. Only one wall of each type can be maintained at any one time,
however.

\item[Weapons] Fire Elementals can attempt to immolate their adversaries.
The Base Chance for such an attack is 70\%, and if the attack
succeeds D + 2 damage is done, with an additional point added for each
Rank of their summoner. If the attack is made in Close Combat, double
the damage done.

\item[Movement Rates] Running: 400 + 10 x Summoner's Rank

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  Base 20  
& 
\textbf{MD:}  20-25
& 
\textbf{AG:}  25-30
& 
\textbf{MA:}  None
\\
\textbf{EN:}  5-50
& 
\textbf{FT:}  10-85 
& 
\textbf{WP:}  14-18
& 
\textbf{PC:}  14-17
\\
\textbf{PB:}  5-7
& 
\textbf{TMR:}  8 + (Adept's Rank/5)
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  3 DP
}\\
\end{tabularx}

\subsubsection{Water elemental}

\begin{description}
\item[Natural Habitat]  Other Planes Very Rare

\item[Number]  1

\item[Description] Water Elementals appear as waves, although the water
does take on a basically humanoid form. The waves will be one-half as
tall, in feet, as the elemental's endurance.

\item[Talents, Skills and Magic] Water Elementals can disappear into any body of water and
become completely undetectable.  They can increase the size of waves
and swells as per the Spell of Wave Making from the College of Water
Magics at a Rank equal to their summoner's Rank plus four. They can
also produce a whirlpool as per the Maelstrom spell from the same
College, at a Rank equal to their summoner's. Neither spell has to be
prepared or cast, but only one can be maintained at a time.

\item[Weapons] Water Elementals form huge fists out of the waves.  They can
attack twice in a Pulse without penalty, and do D + 4 damage, plus one
additional point of damage for each Rank of their summoner.

\item[Movement Rates]  Swimming: 500 + (25 x Summoner's Rank)

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  Base 25  
& 
\textbf{MD:}  17-23
& 
\textbf{AG:}  16-20
& 
\textbf{MA:}  None
\\
\textbf{EN:}  5-50
& 
\textbf{FT:}  10-85
& 
\textbf{WP:}  14-18
& 
\textbf{PC:}  14-17
\\
\textbf{PB:}  6-8
& 
\textbf{TMR:}  10 + (Adept's Rank/2)
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  4 DP
}\\
\end{tabularx}
\end{multicols}
