\begin{multicols}{2}

\setlength\columnseprule{0.2mm}

\section{Undead}
Undead are player characters or NPC's who are neither alive nor
dead. They cannot be killed by normal means since they are not
alive. Instead, they may be destroyed by magic and by silvered
weapons, in most cases.  Lesser undead include ghosts, ghouls,
revenants, skeletons and zombies.  Greater undead include night-
gaunts, spectres, vampires, wights, and wraiths.  Greater undead have
the power to drain life force (in the form of Endurance and Fatigue)
from living victims.  They cannot affect other undead in this manner,
nor can they affect extra-dimensional beings (demons, devils, etc.).

lt is possible for characters to become undead in certain cases.  When
this occurs, the character's abilities will be altered somewhat,
depending upon the type of undead he becomes.  His characteristics may
be decreased or increased to fall within the appropfiate range for the
class of undead to which he belongs, but should be altered by the GM
as little as possible.  For example, a character with a Physical
Strength of 10 who became a vampire would have his PS increased, but
probably not by more than 5 (so as to possess the minimum Physical
Strength for a vampire).  A character who becomes undead retains his
Rank in all skills, talents and magic which are useable to him in an
undead state.  They may continue to acquire Experience and Rank, but
all Ex- perience costs are doubled for undead characters.  Undead
beings recover automatically from being stunned at the beginning of
each Pulse, if they have not already recovered.

\subsection{Lesser Undead}
Lesser Undead include all undead who do not have the power to drain
life force from the living of this plane.

\subsubsection{Ghost}

\begin{description}
\item[Natural Habitat] Where human beings may be found. Uncommon

\item[Number] 1-36 (1)
 A ghost is the insubstantial form taken by the spirit of
a recently deceased individual. It may appear in exactly the form the
individual himself took in life except for a certain shimmering
quality and the fact that a ghost seen in direct sunlight will tend to
fade into the background. Ghosts are usually produced when an
individual dies in particularly traumatic circumstances.

\item[Talents, Skills and Magic] Ghosts have no skills or magical powers and cannot harm
living beings directly since they are insubstantial. However, they can
frighten living beings (especially the unwary) into flight. A
character facing a ghost must roll 3 times his willpower or less on D
I00 to keep from panicking and fleeing from the ghost's presence. Add
20 to the dice roll if the ghost was unexpected.

\item[Weapons] Ghosts have no natural weapons and may use no man-made
weapons. They cannot harm others directly and cannot be themselves
harmed by physical means.

\item[Movement Rates]  Flying: 250

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  None
& 
\textbf{MD:}  None
& 
\textbf{AG:}  None
& 
\textbf{MA:}  5-23
\\
\textbf{EN:}  None
& 
\textbf{FT:}  None
& 
\textbf{WP:}  5-23
& 
\textbf{PC:}  5-23
\\
\textbf{PB:}  5-23
& 
\textbf{TMR:}  5-7
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Ghosts can be sources of information which may be wholly or
partially false. They may aid characters who are willing to assist
them in finishing any business they may have left uncompleted in their
former life and which is in whole or in part the cause of their
inability to find peace. For example, they may seek revenge on the
individual(s) who brought about their end OT may desire the A,elfare
of a still living former lover to be attended to. Ghosts are usually
chained to a place or object connected with their death or the
fullfilment of their desires.  They may not leave that place until put
to rest inless wronged or cheated by a character. They will then
become tied to that character, following him about, haunting him and
warning his enemies of his approach until he rights the wrong he has
done to them. Ghosts have only one aim: to alter the circumstances
that have made them ghosts. They may temporarily assist characters in
their own goals as a means of getting the characters' assistance in
their quest, but only in a very limited sphere. Once their problem has
been soln,ed, they are laid to rest and immediately dissipate.

\end{description}

\subsubsection{Ghoul}

\begin{description}
\item[Natural Habitat] Crypts, Ruins, Fields Uncommon

\item[Number]   3-30 (6)
 Ghouls are a form of undead who specialize in devouring
the flesh of the living or of recently dead characters.  They are
physically humanoid in all respects except that they may be ravaged by
sickness or maimed by the wounds which killed them. Their undead
status has robbed them of most of their intelligence and agility, but
has also made them fearless.

\item[Talents, Skills and Magic] Ghouls possess no skills or magic.  They possess two special
characteristics: Any character bitten by a Ghoul will automatically
die if the wound becomes infected and will rise at dawn on the day
after his death as a Ghoul.  Ghouls can only be killed by destroying
their motor center.  Thus, they are unaffected by specific Grievous
Injuries.

\item[Weapons] Ghouls may use their hands in Melee or Close Combat as
unranked weapons. They may only Strike once per Pulse (i.e., no Double
Strikes). Ghouls may Strike twice per Pulse in Close Combat (once with
hands and once with teeth).  Their bite has a Base Chance of 40\%,
damage of [D - 4] (but with + 20 to Infection Checks in addition to
any modification for bites) and is never Ranked.

\item[Movement Rates]  Running: 150

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  14-18
& 
\textbf{MD:}  8-12  
& 
\textbf{AG:}  6-12
& 
\textbf{MA:}  None
\\
\textbf{EN:}  18-24
& 
\textbf{FT:}  25-30  
& 
\textbf{WP:}  18-25
& 
\textbf{PC:}  10-14
\\
\textbf{PB:}  1-2
& 
\textbf{TMR:}  3
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Ghouls, like Trolls hate fire and will immediately attack
any character holding a torch in preference to all other targets.

\end{description}

\subsubsection{Revenant}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] A character who dies in circumstances conducive to the
creation of a ghost may, at the GM's discretion, become a revenant.
Thereafter, the character takes on the form of a ghost or night-gaunt
(GM's discretion) chained to the place of death. The GM always plays
the character. The revenant retains his previous personality and those
characteristics appropriate to his new form and gains those
characteristics appropriate to his undead state.  Revenants can never
be returned to life.

\end{description}

\subsubsection{Skeleton}

\begin{description}
\item[Natural Habitat]Any habitat where humans are found.Rare

\item[Number]3-12 (6)
 Skeletons are animated figures formed magically from the
bones of the dead. They are controlled and animated by the will of an
Adept in the same manner as zombies are.  Often they will be clothed
and armed. Skeletons immediately fall apart and turn to dust whenever
they cease to be animated by an Adept. Skeletons may never be Stunned.

\item[Talents, Skills and Magic] Skeletons have no special talents, skills, or magic.
However, their special magical nature makes them immune to damage from
weapons rated for A type damage (arrows, thrusting swords, stabbing
weapons). They may use weapons, including Ranked weapons, shields and
armor.

\item[Weapons] Skeletons may attack using the unarmed combat rules, but
will probably be armed with a weapon of some sort.

\item[Movement Rates]Running: 150

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  10-16
& 
\textbf{MD:}  14-18  
& 
\textbf{AG:}  12-16
& 
\textbf{MA:}  None
\\
\textbf{EN:}  8-12
& 
\textbf{FT:}  5-10
& 
\textbf{WP:}  20-25
& 
\textbf{PC:}  14-18
\\
\textbf{PB:}  1-2
& 
\textbf{TMR:}  3
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Skeletons are often found around old battlefields and
graveyards. They need to be animated before they can attack,
though. The skeletons used by an Adept need not be Human. He could
animate any type of being which would then have attributes close to
those it originally possessed while alive.  Usually, however, Human
skeletons are used since they are more easily animated.

\end{description}

\subsubsection{Zombie}

\begin{description}
\item[Natural Habitat] Any habitat where humans are found. Rare

\item[Number] 6-24 (6)
 Zombies are another variety of living dead. In this
case, the dead are animated by a spell cast by a member of the College
of Necromancy. They may (when newly dead corpses are used) appear
nearly normal except for a slight earthen hue.  When old corpses are
used, however, the zombies may the no more than skeletons held
together by a few strands of sinew.

\item[Talents, Skills and Magic] Zombies have no skills or talents and are not magic users.
Like Ghouls, they are unaffected by Grievous Injury since their motor
center must be destroyed before they cease to be animated.

\item[Weapons] Zombies never use weapons. They may attack with hands using
the bare-handed attack rules. Their hands are never Ranked.

\item[Movement Rates]Running: 150

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  10-16   
& 
\textbf{MD:}  6-10
& 
\textbf{AG:}  4-8
& 
\textbf{MA:}  None
\\
\textbf{EN:}  16-22
& 
\textbf{FT:}  20-25
& 
\textbf{WP:}  18-25
& 
\textbf{PC:}  8-12
\\
\textbf{PB:}  1-2
& 
\textbf{TMR:}  3
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] A zombie loses 1 point per day from his Endurance for each
day that he is not reenchanted by his master.

\end{description}

\subsection{Greater Undead}
Greater undead include all those undead who have the ability to drain
life force from the living of this plane.


\subsubsection{Spectre}

\begin{description}
\item[Natural Habitat] Ruin, Cavern, Crypt (usually enclosed area) Very rare

\item[Number] 1-3
 Spectres are undead beings half of this world and half
of another plane. They enter this world most fully at night and may
disappear entirely during the day. If exposed to direct sunlight, they
permanently dissipate. Spectres are humanoid in shape, but tend to
show an easily discernible blue g!ow when viewed in the shadowed areas
they prefer.

\item[Talents, Skills and Magic] Spectres may possess the entire gamut of human talents and
skills. In addition, they are Dark Mages of the College of Celestial
Magics. Their touch is ice cold and any character coming in contact
with a spectre which is in full physical form suffers [D + 3] damage
drained directly from Endurance. A character whose last point of
Endurance is drained by a spectre becomes a half-strength spectre.

\item[Weapons] Spectres have no natural weapons other than their
touch. They also are not weapon users even in their most substantial
state. Damage from contact with spectres is never absorbed by armor.

\item[Movement Rates]  Running: 300

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  3-30
& 
\textbf{MD:}  3-25
& 
\textbf{AG:}  10-20
& 
\textbf{MA:}  17-30
\\
\textbf{EN:}  20-30
& 
\textbf{FT:}  25-35
& 
\textbf{WP:}  20-30
& 
\textbf{PC:}  20-30
\\
\textbf{PB:}  3-20
& 
\textbf{TMR:}  6
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] The characteristics and power of spectres varies in direct
proportion to their substantiality. The same spectre which has a PS of
3 at noon would have a PS of 25-30 at midnight. A spectre cannot
physically harm a character while insubstantial.  He can perform
magic. Except for the few hours when the sun is highest, the spectre
may choose to be either in noncorporeal (insubstantial) or corporeal
(substantial) form. It takes a spectre 12 Pulses (1 minute) to change
between these forms. Spectres can only be harmed by exposure to direct
sunlight or by magic.  Magic does not affect spectres while in their
noncorporeal form.

\end{description}

\subsubsection{Vampire}

\begin{description}
\item[Natural Habitat] Wherever men are found. They make their homc in Caverns,
Ruins and Crypts, however, and are most likely to be found there.
 Very Rare

\item[Number] 1-6
  Vampires are human (or humanoid) beings who
have entered the netherworld of "life-drinkers." They support
their existence by drinking the blood of the living (or recently
dead) members of their own species (but not necessarily their
own race). A character becomes a vampire only by performing a
Ritual of Becoming Undead or by being killed by a vampire who
drains the character's last point of Endurance. A character who
dies in this manner rises at sunset of the third day after his
death in the form of a vampire and must, thereafter, be treated
in all ways as a vampire. Once the character's last point of
endurance has been drained, the character may not normally be
resurrected or prevented from becoming a vampire. He may be
destroyed in the same way as a vampire once he rises from the
dead, but not before. Vampires always appear to be the age they
were at when they became vampires. All physical wounds or
disfigurements during life are, however, removed. A character
may sometimes be prevented from becoming a vampire by being
pinned in his coffin by a stake driven through the heart.


\item[Talents, Skills and Magic] Vampires retain all magical and racial skills acquired
during their life and may learn any number of new skills. They lose
all magical knowledge unless they were, while alive, members of the
College of the Mind. All vampires are automatically members of the
College of Sorceries of the Mind. Any character who was a member of
this College while alive retains any Rank he achieved with the
knowledge of that College.

Vampires have the following special skills. They can change shape
between the following forms: Bat, rat, wolf, cloud of mist or their
"natural" human shape. It takes one full pulse for the transformation
to occur. A vampire must implement only a pass action during a pulse
in which he attempts to change shape. Vampires are hypnotically
attractive to members of their own species (including other races).
They possess most of the characteristics that they did during their
life except that their Physical Strength and Willpower are both
increased by 50\% (round up). They gain the ability to summon control
and communicate with all common land mammals and are especially adept
at communicating with and summoning bats, rats and wolves. This
ability operates automatically as a talent of the same type as G-1,
G-3, and Q-1 of the College of Earth Magics, except that the Base
Chance is always increased by 20 when a vampire attempts to summon,
control or communicate with an animal.  Vampires lose 1 point from
Fatigue (and from Endurance when Fatigue is exhausted) each hour
during which they do not feed.  When a vampire feeds, they add 1 point
per pulse (for each 10 seconds when not in combat) to either Fatigue
or Endurance (their choice) for each point they drain from their
victim. The vampire's Fatigue and Endurance may never be raised above
their normal maximum level by this method. A vampire whose last point
of Endurance is removed because they have failed to feed is
permanently destroyed. A vampire feeds by entering Close Combat with a
character and drinking that character's blood or by hypnotizing the
character into allowing them to drink the character's blood (using a
Spell of Hypnosis). Characters hypnotized by a vampire will not resist
a suggestion that they allow the vampire to feed on them.

Vampires do not cast a shadow. Their reflection does not show up in a
silvered mirror, but does sho-A, up in a mirror not made with
silver. They may not cross running water (streams, rivers, and the
like). They must sleep from sunrise to sunset in a bed of earth taken
from their grave. Vampires may not enter an inhabited dwelling unless
invited in by its inhabitants.  Relics and holy items associated with
the Powers of Light are repellant to vampires. So is garlic.  Thev
will not willingly come in physical contact with such items and will
tend to flee them or at least stay 10 feet or more away from them.
Vampires fear dogs with angel eyes (dark circles around the eyes) and
will generally flee such animals.  Vampires cannot die.  They can be
destroyed in the following manner:
 By direct exposure to sunlight.
 By being staked through the heart while in their coffin.
 By hearing the crow of a cock at dawn while out of their coffin.
 By having their heart pierced by a silvered weapon.
 By failing to replenish Endurance through feeding.


Vampires suffer [D - 5] damage from the touch of consecrated objects,
but will become infuriated by their touch so that they attack the
wielder of such objects instead of fleeing their presence.

\item[Weapons] Vampires are treated in all ways as human beings for
purposes of combat except that they cannot be harmed by normal
weapons. They have no special weapons except their ability to drink
blood as described above. This ability may not be used against
characters wearing armour other than cloth, leather or wicker unless
the victim is first hypnotised and opens his armor so that the vampire
can reach his neck.

\item[Movement Rates] Running while in human form: 300. The Movement Rates for
other forms (bat, rat and wolf) are as per the creatures of the same
name.

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  15-30
& 
\textbf{MD:}  15-25
& 
\textbf{AG:}  15-25
& 
\textbf{MA:}  15-25
\\
\textbf{EN:}  15-25
& 
\textbf{FT:}  20-30
& 
\textbf{WP:}  25-30
& 
\textbf{PC:}  20-30
\\
\textbf{PB:}  12-25
& 
\textbf{TMR:}  6
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Vampires are not harmed by weapons other than silvered
weapons. They may be harmed (but not destroyed) by magic.  A humanoid
character bitten by a vampire becomes a under the control of the
vampire that bit him if that vampire has any remaining capacity to
control his victims. Otherwise, he becomes a vampire under no one's
control. A vampire may control a number of his victims who have become
vampires equal to half his Willpower (rounded down). A vampire's
control over his undead victims is not total and the victim has a 10\%Base Chance of successfully disobeying any order of the controlling
Vampire which would lead to his injury or destruction or which would
prevent him from feeding. This Base Chance is increased by 5 for each
additional victim under the vampire's control. Control requires no
concentration.

A vampires may create a telepathic bond between himself and a victim
who is not undead by performing the Kiss of the Undead.  The vampire
opens a small wound over his heart with any handy edged weapon and
forces the victim to drink the vampire's blood from this open
wound. Thereafter, the vampire and the victim will have a limited
knowledge of each other's thoughts and location until one or both are
destroyed.  Note: A character who becomes undead can still participate
in the game, but would operate under severe limits. He would, for
example, be unable to travel during the day except while locked in his
coffin. He would be prevented from entering certain dwellings,
travelling over consecrated ground, etc. The character is always
played by his original player, not by the GM.

\end{description}

\subsubsection{Wight (barrow wight)}

\begin{description}
\item[Natural Habitat]  Crypts, Ruins, Graveyards, Fields \& Moors Rare

\item[Number] 1-6
 Wights are a form of ghost. They normally appear in
their human form, but vary in substantiality in direct proportion to
the time of day. Bright sunlight makes them fade into the spirit world
while moonlight and starlight increases their corporeality. When in a
corporeal state (usually just after moonrise), they are capable of
harming humans just as if they were, themselves, alive.

\item[Talents, Skills and Magic] Wights may exhibit any and all human skills and talents.
They may use magic, but are limited to the Celestial Magics
(Conjurations of Night and Stars) at Rank 10 or above due to their
unique status as beings who themselves wax and wane substantial in
direct proportion to the power of celestial bodies. They can use
weapons, wear armor and carry shields just as they did when they were
alive.

\item[Weapons] Wights are treated in all ways as human when in a
substantial state except that they cannot be harmed by normal
weapons. In addition, a wight may drain the life force of any
character with whom he comes into physical contact. Whenever a
character strikes or is struck by a wight, the contact does [D + 2]
damage. Damage inflicted on a character in this manner is never
absorbed by normal armor.

\item[Movement Rates]  Running: 250

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  15-30
& 
\textbf{MD:}  15-25
& 
\textbf{AG:}  10-20
& 
\textbf{MA:}  15-30
\\
\textbf{EN:}  10-25
& 
\textbf{FT:}  15-30
& 
\textbf{WP:}  20-30
& 
\textbf{PC:}  20-30
\\
\textbf{PB:}  3-7
& 
\textbf{TMR:}  5
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] Wights are individuals who have died under a geas or oath
which they have been unable to fulfill in life. Often, they are bound
to a specific place by an oath to protect that place. They are often
found guarding burial sites, especially where treasure is buried along
with the body of the master the wight has been set to guard. Only
magic or silvered weapons can harm a wight and these things can
destroy it. An individual whose last point of Endurance is drained by
a wight becomes a night gaunt upon the next moonrise, but will have
all his normal characteristics halved (round down) and will exist
under the control of the wight who killed them unless that wight has
been destroyed. When a Wight is destroyed (by magic or silver), all
night gaunts under the wight's control are also destroyed.

\end{description}

\subsubsection{Wraiths}

\begin{description}
\item[Natural Habitat] Fields, Crypts, Graveyards, Ruins, Caverns Uncommon

\item[Number] 1-10 (l)

 A wraith is much like a wight in that he appears human
and tends to wax and wane insubstantial. However, a wraith dissipates
entirely in full sunlight and may never reform Consequently, wraiths
stay in dark areas where the sun does no't penetrate and only come out
at night. The natural pallor of a wraith is replaced by a silvery aura
in bright moonlight when its power is at its height.

\item[Talents, Skills and Magic] Wraiths cannot harm living beings by physical means, nor can
they engage in any physical skills.  However, a wraith can, when its
power is high, perform Celestial Magic at Rank 8 or above. The touch
of a wraith operates in the same manner as the touch of a wight except
that +4 is added to the damage die roll. Even at their most
substantial, they are not otherwise physical beings.

\item[Weapons] Wraiths have no natural weapons and cannot use physical
objects. They can use magic.

\item[Movement Rates]  Running: 250

\end{description}
\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  2-5
& 
\textbf{MD:}  2-5
& 
\textbf{AG:}  25-35
& 
\textbf{MA:}  15-30
\\
\textbf{EN:}  15-30
& 
\textbf{FT:}  20-35  
& 
\textbf{WP:}  20-30
& 
\textbf{PC:}  20-30
\\
\textbf{PB:}  5-20
& 
\textbf{TMR:}  5
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  None
}\\
\end{tabularx}

\begin{description}
\setlength\itemsep{0pt}

\item[Comments] A wraith may not be harmed by weapons (including enchanted
weapons). It may only be destroyed by magic spells or by exposure to
direct sunlight.

\end{description}
\end{multicols}
