\documentclass[a4paper]{article}

\usepackage[libertine]{dq}

\begin{document}

\section*{Knight of the Golden Order (1.1)}

\begin{multicols*}{3}

\subsection*{Notes}

\begin{description}
\item[11 February 2009] Added Anywhere Is Home and Home Ground
  Advantage.  Changed armour to be explicitly against Endurance and
  then Magical damage.  Reduced damage negative in Close Combat.
  Soldiers now get Amulet.

\item[6 February 2009] Move items to own section and add mounts.

\item[31 January 2009] Revise bits.

\item[10 August 2007] Add limits to how many abilities a person can
  learn based on their membership.

\item[19 May 2007] Revising the armour protection and making sure
  costs are cumalative.
\end{description}

\subsection*{Introduction}

The Golden Order is a secular order originally sponsered by the
ancient golden dragon Orthosopheles.  It was formed to confront the
armies of the evil dragons and their wormkin servants.  A few of the
wormkin could fly but most where infantry, with the a number of them
having riders.

The wormkins' strong resistant to magic and the extreme scarity of
talented Adepts lead the Order to focus on elite officer corp which
would be in the vanguard of the defending armies.

To counteract the flying dragonkin it was necessary to have the
vanguard mounted, with the most experienced members able to fly with
their mounts.  The ability to rapidly respond to dragonkin incursions
and not have their mounts killed lead to development of super heavy
warhorses, able to stand up in melee against all but the greater
dragonkin.

After the Order's original fortress was destroyed the Order realised
that they had become a focus for aggression, both internally from
political enemies and externally from dragon host raids.  With many of
its members spending much of their time holding the line the focus
turned to improving the defense of frontier keeps.

The war ended in the defeat of the chromatic dragons and their kin but
continuing hostility to all their kind lead Orthosopheles and a number
of his kin to go into exile, leaving the Order to be disolved into the
mainstream political structure of the time.

Recent events have lead Orphostopheles to revive the Order and invited
a number of worthy candidates to join.  Members of the order are
expected to perform the following actions.
\begin{enumerate}
\item Always be in the vanguard of any action against invading forces.

\item To resist tyranny and oppression and champion justice.

\item To not involve the Order in any political or religious dispute.

\end{enumerate}

The Order's training focused on four seperate area of expertise:
Attack, Defense, Mount and Keep.  Abilities costs are expressed in
Points, with non rankable abilities costing one Point and rankable
abilties Rank Points (so to get an ability to five costs a total of 15
Points).  Points cost 1000 EP but may be given as reward for
completing tasks assigned by the Order.  All training and attunement
counts as weapons training for the purposes of ranking.

\subsection{Attack}

All abilities only work with the Attuned weapon unless otherwise
specified.

\begin{description}

\item[Attune Weapon] A specific weapon is attuned for use with the
  Order Abilties.  This takes one week of training and only one weapon
  can be attuned at a time.  The attuned weapon can always be prepared
  as a free action.

\item[Weapon Focus] Gains 4\% / Rank Strike Chance with attuned weapon
  (or shield).  [maximum 5 Ranks]

\item[Improved Rank] Gains additional effective rank with the attuned
  weapon (can exceed existing maximum).  [requires Rank 5 Weapon Focus]

\item[Instant Response] The Knight has a (IV - 50 + 10 / Rank)\%
  chance to simultanously retaliating against a surprise attack.  This
  can be done with any available weapon.  [maximum 5 Ranks]

\item[Preemptive Retaliation] The Knight has a IV\% chance of
  retaliating to a surprise attack before attacker actually strikes.
  This can be done with any available weapon.  [requires Rank 5 Instant
    Response]

\item[Power Attack] The Knight can increase damage of their weapon by
  up to Rank points of damage by reducing their strike chance by -15\%
  / Rank. This costs 1 FT to use this ability.  [maximum Rank 5]

\item[Weapon Throw] The Knight may throw their weapon up to 10 feet /
  Rank as a melee attack.  [maximum Rank 5]

\item[Close Attack] The Knight can fight in close with their weapon at
  a reduce strike chance of [SC - 100 + 15 / Rank] with damage reduced
  by [5 - Rank].  [maximum Rank 5]

\item[Ambidexterous] Attuned weapon (or shield) can be wielded in
  either hand without the offhand penalty.  [requires 10 Points in
    Attack]

\item[Dying Strike] The Knight has an IV\% chance of retaliating to a
  blow which kills them.  [requires 15 Points in Attack]

\item[Melee Freedom] The Knight may bypass a threatening opponent but
  may suffer a strike from the opponent.

\item[Extended Movement] The Knight may spent 1 FT to move up to Rank
  additional hexes (maximum TMR) in melee at still make a full attack.
  [requires Melee Freedom, maximum 5 Ranks]

\item[Critical Strike] The Knight may spent up to Rank FT to reduce
  the dice roll on their attack by FT\%.  [requires 25 Points in
    Attack, maximum 5 Ranks]

\end{description}

\subsection{Defense}

All abilities only work with attuned armour or shields unless specified.

\begin{description}

\item[Attune Armour] The Knight attunes a specific set of armour and
  shield to use their Order Abilities.  This will take a week of
  training and only one set of armour and one shield can be attuned at
  a time.  The attuned shielded can always be readied as a free
  action.

\item[Endurance Armour] The attuned armour will block Rank points of
  Endurance damage.  This does not stack with any existing Endurance
  protection.  [maximum 5 Ranks]

\item[Magic Armour] The attuned armour will block Rank Points of
  magical damage.  This does not stack with existing permanant magic
  protection.  [requires Rank 5 Improved Armour, maximum 5 Ranks]

\item[Extended Fatigue] The Knight must suffer 4 / Rank EN damage
  before becoming fatigued for the purposes of combat
  modifiers.  [maximum 5 Ranks]

\item[Extended Endurance] This ability lowers the point at which it is
  required to make willpower checks to stay conscious by 1 / Rank (eg
  at Rank 3 they must have 3 x WP check at 0 EN).  [requires Rank 5
    Extended Fatigue, maximum 5 Ranks]

\item[Last Man Standing] The Knight will not become unconscious due to
  damage until are below -EN/3.  If they are not dead at this point
  (eg under Trollskin or similar) they will not gain consciousness
  until they have positive endurance.  [requires Rank 5 Extended
    Endurance]

\item[Stun Resistance] The amount of damage a Knight must suffer
  before being stunned is increased by +1 / Rank (ie stun on > EN/3 +
  Rank).  [requires 10 Points in Defense, maximum 5 Ranks]

\item[Dodge] A Knight has [1/2 AG + opponent hex size * Rank]\% chance
  (halved if from rear hex) of moving one hex to try to avoid being
  trampled (or being closed with).  The trample or close is resolved
  after the one hex move has been taken.  This can be done at any time
  except when surprised.  If the Knight fails to dodge they may not
  use their weapon to try to keep the opponent out of close.

\item[Instant Stand] A Knight can stand from prone as a free action.
  [requires Rank 5 Dodge]

\item[Quickened Equip] A Knight takes armour protection value full
  actions to equip their armour.  [requires 20 Points in Defense]
 
\item[Guard] As a full action the Knight may chose to force up to Rank
  attacks against an adjacent figure to be directed at them, provided
  they are also adjacent to the attacker.  [maximum 5 Ranks]

\end{description}

\subsection{Mount}

A knight of the Order designates a single mount their mount of
preference.  This requires them to train with the mount for one week.
The mount must be trained if an animal.

\begin{description}

\item[Bound Mount] The Knight can verbally communicate with their
  mount up to 20 feet / Rank away.  [maximum Rank 5]

\item[Improved Handling] The Knight adds +1 / Rank to their
  horsemanship (or flying) for purposes of controlling their mount
  (allows > Rank 10).  [maximum 5 Rank]

\item[Instant Mount] The Knight can mount as free action provided the
  mount is in an adjacent.  [requires Rank 5 Improved Handling]

\item[Improved Natural Armour I] The mount's natural armour improved
  by +1 / Rank.  [maximum 5 Ranks]

\item[Improved Natural Armour II] The mount's natural armour counts
  against all damage.  [requires Rank 5 Natural Armour I]

\item[Extended Charge] The mounts TMR improves by 2 TMR / Rank when
  charging.  [requires 10 points in Mount, maximum 5 Ranks]

\item[Flying Mount] A mount gains Rank in the Flying Skill.  This does
  not automatically allow the mount to fly but can control their
  flight if magically empowered.  [requires 5 points in Mount]

\item[Ranger] The mount gains its rider's ranger tracking, direction
  sense, detect ambush and detect traps and snares with a (50 - 10 /
  Rank) penalty to base chance.  [requires 15 points in Mount, maximum
    5 Ranks]

\item[Dodge] The knight use their dodge skill while mounted.

\item[Natural Weapons] The Mount's natural weapon attacks have a
  minimum weapon rank of Rank 0.

\item[Kick I] The mount's kick (or claw etc) attack Rank increases by
  +2 / Rank.  [maximum 5 Ranks]

\item[Kick II] The mount can kick (or claw) with its second hoof (claw
  etc) with both suffering a -20\% Strike Chance penalty.  [requires
    Rank 5 Kick I]

\item[Bite I] The mount's bite attack Rank increases by +2 / Rank.
  [maximum Rank 5]

\item[Bite II] The mount may bite as a secondary attack at -20\%
  SC. [requires Rank 5 Bite I]

\item[Trample] The mount's trample attack Rank chance increases by +10
  / Rank and improved damage of +1 / Rank. [maximum 5 Ranks]

\item[Endurance] The mount's overland movement rate improves by 20\% /
  Rank.  [maximum 5 Ranks]

\end{description}

\subsection{Keep}

A member can declare any solid structure a temporary keep after
sending a week establishing the environment.  The knight can only ever
have one temporary and one permanant keep attuned at any one time.
Only a Knight can learn keep abilities.

\begin{description}

\item[Improved Willpower] All within the Keep gain +1 / Rank Willpower
  bonus (doubled if permanant keep).  [maximum 5 Ranks]

\item[Anywhere Is Home] The Knight gets their permanant keep bonus for
  Improved Willpower in a temporary keep.  [requires Rank 5 Improved
    Willpower]

\item[Improved Fortitude] The Knight gains +1 / Rank FT and +2 / Rank
  EN while in the Keep (double if in permanant keep).  [maximum 5
    Ranks]

\item[Home Ground Advantage] The Knight gains additional effective
  ranks in Improved Willpower and Improved Fortitude will in their
  permanant keep.  [requires Rank 5 Improved Will, Rank 5 Improved
    Fortitude, maximum Rank 5]

\item[Antiscry] The Knight gains their MR + 5\% / Rank chance of been
  unscryable while in their Keep (rolled when ever somebody tries to
    scry them).  [requires 5 points in Keep, maximum Rank 5]

\item[Home Ground I] If the Knight is magically transported anywhere
  with a random destination and the Keep a possible destination then
  there is a 20\% / Rank chance they will appear in the Keep.
  [maximum 5 Ranks]

\item[Home Ground II] If the Knight is magically transported against
  their will they have a WP + 10\% / Rank chance of return to their
  Keep instead, if they will it and it a possible destination.
  [requires Rank 5 Home Ground I, maximum 5 Ranks]

\item[Beyond Life] If the Knight is slain within the Keep their spirit
  is beyond all Necromancy or Witchcraft.  In addition they gain +5\/%
  / Rank improvement to their chance of ressurection.  [requires 15
    points in Keep, maximum 5 Ranks]

\end{description}

\subsection{Affliation}

\begin{description}

\item[Fraternal] For those people who are leaders of an established
  organisation (eg.\ church prelate, royal retainer, landed duke etc)
  or are pacted to a power may only become Fraternal Members.  The
  Order recognises the person as worthy and will provide simple
  assistance if requested, such as personal lodging for a short
  period, provisions for themselves and direct staff etc.  A Fraternal
  member may only spend a maximum of 15 points in any one area, and a
  total of 25 points.

\item[Initiate] Initiate members can gain lodging and board at any
  Order lodge or Keep for extended periods of time.  They are expected
  to assist the local commander if requested.  Initiates may only
  spend a maximum of 20 point in any one area and 30 points total.
  The order will provide a blessed (counts as magical, can take
  weapons spells as per normal) master worked weapon (+5\% SC, +1 DM)
  and artisan worked silvered chain mail (+1 AR).

\item[Soldier] Soldiers must tithe 5\% of their income to the Order
  and spend atleast one week a season aiding the Order and have spent
  atleast 25 Points on abilities and can spent at most 70.  Soldiers
  receive an Amulet of the Golden Order.  They are provided with
  access to good riding horses (which they are expected to maintain
  are not lose too frequently).

\item[Knight] Knights must tithe 10\% of their income to the Order and
  spend atleast two weeks a season aiding the Order and have spent
  atleast 50 Points on abilities and a maximum of 100. They receive a
  Tabbard of the Golden Order.  They are provided with access to war
  horses if requested (which they are expected to maintain are not
  lose too frequently).

\item[Knight Commander] Knight Commanders must tithe 10\% of their
  income to the Order and spend atleast two weeks a season aiding the
  Order.  They will receive the Barding of the Golden Order and may
  request a flying mount (where available).

\item[Keep Commander] Most Keep Commanders are also local nobles (the
  Order will assist in worthy Knights gaining the necessary status).
  The Order will finance the construction of the keep but the local
  commander is expected to maintain and staff the keep.  Keep
  commanders will recieve the Standard of the Golden Order once the
  keep has been completed.

\end{description}

Members must be authorised to advance rank (ie.\ cleared by Ross
Alexander).

\end{multicols*}

\begin{multicols*}{3}
\section*{Items}

\subsection{Amulet of the Golden Order}

They gain the Amulet of the Order.  This amulet gives the wearer +5\%,
Magic Resistance, +5\% Defense, +5\% to all Strike Chances, an
additional 1 $\times$ WP to fear checks and -10 on all fear table
rolls.  It does not stack with amulets of Aquilegius, Diamonds or Luck
(or equivalent).  This item is attuned to the wearer and will not work
for anybody else.  It will also crumble to dust if not worn for more
than a week.

\subsection{Tabbard of the Golden Order}

This tabbard is magical and is impervious to dirt and soiling.  It
provides 2 points of armour against all damage and has the following
spell like abilities (each can be used once per 24 hours).  The spell
like abilities require a Trigger Action to use.

\begin{description}
\item[Light] This spell causes the cloak to glow like a torch,
  shedding bright light in a 20-foot radius (and dim light for an
  additional 20 feet) from the point you touch. The effect is
  immobile, but it can be cast on a movable object. Light taken into
  an area of magical darkness does not function.  The duration is 30
  minutes.

\item[Cure Moderate Wounds] Cures 2d10 + 10 damage to wearer.

\item[Bear's Endurance] The wearer gains +4 EN bonus for 5 minutes.
\end{description}

The tabbard will preserve the body of the wearer if the die while
wearing it and it is not removed.  This item is attuned to the wearer
and will not work for anybody else.  It will also crumble to dust if
not worn by for over a week.


\subsection{Barding of the Order}

\begin{enumerate}
\item Weights 60 lbs.
\item Has 14 points of normal protection and 7 points against
endurance, specific grevious and magical damage.
\item Additional 30\% Magic resistance and can be the target of one counterspell.
\item The mount can ride over water (but not mist etc) and up mountain
sides (but not vertical surfaces).
\item Mount gains an additional 50\% PS, FT \& EN.
\end{enumerate}

\end{multicols*}

\section*{Mounts}

\begin{multicols*}{2}

\subsection{Quarterhorse}
\begin{description}

\item[Description] These horses grow to five feet at the shoulders,
  and weigh about half a ton.

\item[Talents, Skills and Magic] Quarterhorses possess no special
  talents, skills, or magic.

\item[Movement Rates]  Running: 600
\end{description}

\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  40-45
& 
\textbf{MD:}  None
& 
\textbf{AG:}  17-20
& 
\textbf{MA:}  None
\\
\textbf{EN:}  18-23
& 
\textbf{FT:}  35-40
& 
\textbf{WP:}  9-10
& 
\textbf{PC:}  17-19
\\
\textbf{PB:}  10-12
& 
\textbf{TMR:}  12
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 3 DP
}\\
\end{tabularx}

\begin{description}

\item[Weapons] Kick: BC = 45\%, [D + 5] damage.  Bite: BC = 20\%, [D -
  1] damage.

\item[Comments] Quarterhorses are very useful as riding animals.  They
  can climb steep banks and ford water. They are extremely responsive
  to their rider if well trained and are able to stop or turn very
  quickly, suiting them for such tasks as rounding up herds. They do
  not have great endurance, however, and they can only maintain their
  maximum speed for periods of a few miles. They can carry loads of up
  to 350 pounds, but they are not intended to be beasts of burden, and
  so their spirit will be quickly broken if they are used as such. If
  a quarterhorse's spirit is broken, its speed will be decreased by
  30\%.
\end{description}

\subsection{Warhorse}

\begin{description}

\item[Description] Warhorses are heavily built, thickly muscled and
  limber animals, bred to carry large burdens for extended
  periods. They are often five and a half feet at the shoulder, and
  can weigh more than a ton.

\item[Talents, Skills and Magic] Warhorses are specially trained for
  combat. They panic less frequently than other horses A,hen faced
  with fire, sudden movements, or loud noises. A rider can also
  control a warhorse's actions in combat, making the horse kick, bite,
  or attempt to trample an adversary.  See the Rules on Mounted
  Combat. Other horses can be used in combat, but they cannot be used
  to attack.

\item[Movement Rates] Running: 500
\end{description}

\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  55-60
& 
\textbf{MD:}  None
& 
\textbf{AG:}  16-19
& 
\textbf{MA:}  None
\\
\textbf{EN:}  25-30
& 
\textbf{FT:}  50-60
& 
\textbf{WP:}  9-11
& 
\textbf{PC:}  16-18
\\
\textbf{PB:}  9-11
& 
\textbf{TMR:}  10
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 3 DP
}\\
\end{tabularx}

\begin{description}

\item[Weapons] In addition to the kick (Base Chance of 45\%, [D + 7]
  damage) and bite (Base Chance of 25\%, [D + 2] damage) of normal
  horses, when in Close Combat a warhorse can also attempt to trample
  an adversary (Base Chance of 25\%, [D + 8] damage).  Warhorses will
  willingly enter Close Combat if their rider directs them to do so.

\item[Comments] These are the most intelligent and (if properly
  treated) the most loyal of the horses. They are responsive to a
  rider, but can also be used to carry heavy burdens (up to 500
  pounds). Using a warhorse as a draft horse will generally lower the
  quality of its training for combat, however.
\end{description}

\subsection{Gryphon}
\begin{description}

\item[Description] The gryphon is a large beast with the body of a
  lion, the head and wings of an eagle and back covered with
  feathers. The gryphon's claws are so large that they can be made
  into serviceable drinking horns. The gryphon is about 5 feet long
  and stands man-high when full grown.

\item[Talents, Skills and Magic] Gryphons have no special skills or
  magical abilities.

\item[Movement Rates] Running: 200; Flying: 500
\end{description}

\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  30-36
& 
\textbf{MD:}  15-20
& 
\textbf{AG:}  15-20
& 
\textbf{MA:}  6-9
\\
\textbf{EN:}  25-30
& 
\textbf{WP:}  12-16
& 
\textbf{FT:}  30-35
& 
\textbf{PC:}  16-22
\\
\textbf{PB:}  0-5
& 
\textbf{TMR:}  4/10
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Hide absorbs 6 DP
}\\
\end{tabularx}

\begin{description}

\item[Weapons] The gryphon is not a toot user, but has natural weapons
  in the form of beak (useable in Melee and Close Combat with a Base
  Chance of 65\%, [D + 5] Damage, possible Rank of 1-3) and claws
  (useable in Melee and Close Combat with a Base Chance of 60\%,
  Damage of [D + 5], possible Rank of 1-4). The gryphon may attack
  twice with claws and once with beak in the same pulse.

\item[Comments] Gryphons are capable of flying a single rider (maximum
  400lb total weight) over long distances.  As carivours they do require
  considerable amounts of food and will raid livestock if hungry.
\end{description}

\subsection{Sand Drakes}
\begin{description}

\item[Description] Sand Drakes are lesser spawn of bronze dragons and
  are normally dark in colour.

\item[Talents, Skills and Magic] Drakes are intelligent and sentient
  but are normally not spell casters as such but it can be assumed
  they have the talents and general spells and rituals of the College
  of Earth magics at rank 10 as spell like abilities.

\item[Movement Rates]  Flying: 500; Running: 350
\end{description}

\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  120-160
& 
\textbf{MD:}  22-24
& 
\textbf{AG:}  20-25
& 
\textbf{MA:}  15-20
\\
\textbf{EN:}  120-140
& 
\textbf{FT:}  120-150
& 
\textbf{WP:}  27-33
& 
\textbf{PC:}  15-20
\\
\textbf{PB:}  6
& 
\textbf{TMR:}  14/5
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Top scales absorb 14 DP
}\\
\end{tabularx}

\begin{description}

\item[Weapons] The Base Chance for a drake's bite is 40\%, and damage
  is [D + 10]. The two claws have a Base Chance of 60\%, with [D + 8]
  damage, while the tail's Base Chance is 70\%, with damage [D + 4].
  These attacks count as Rank 0 Unarmed for the purposes of adding PS
  and MD to strike chances.

\item[Comments] Sand drakes prefer deserts or other hot environments.
  They are happy on the ground and are willing to run long distances
  overland rather than fly.  They are sociable creatures and are
  willing to work with others whom they agree with.
\end{description}

\subsection{Cloud Drakes}
\begin{description}

\item[Description] Cloud Drakes are lesser spawn of silver dragons and
  are normally white or blue in colour.

\item[Talents, Skills and Magic] Drakes are intelligent and sentient
  but are normally not spell casters as such but it can be assumed
  they have the talents and general spells and rituals of the College
  of Air magics at rank 10 as spell like abilities.

\item[Movement Rates]  Flying: 700; Running: 250
\end{description}

\begin{tabularx}{\linewidth}{@{}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{\hspace{0.5em}}X@{}}
\textbf{PS:}  100-150
& 
\textbf{MD:}  19-21
& 
\textbf{AG:}  12-15
& 
\textbf{MA:}  20-25
\\
\textbf{EN:}  80-100
& 
\textbf{FT:}  110-130
& 
\textbf{WP:}  27-33
& 
\textbf{PC:}  15-20
\\
\textbf{PB:}  8
& 
\textbf{TMR:}  14/5
& 
\multicolumn{2}{@{}l@{}}{\textbf{NA:}  Top scales absorb 10 DP
}\\
\end{tabularx}

\begin{description}

\item[Weapons] The Base Chance for a drake's bite is 60\%, and damage
  is [D + 12]. The two claws have a Base Chance of 50\%, with [D + 6]
  damage, while the tail's Base Chance is 70\%, with damage [D + 4].
  These attacks count as Rank 0 Unarmed for the purposes of adding PS
  and MD to strike chances.

\item[Comments] Cloud drakes prefer high, cold environments and can
  flying for extremely long distances without stop.  They are
  generally alouf of the happenings on the ground, living on high
  mountain eyries.

\end{description}

\end{multicols*}

\end{document}
