\section{Miscellaneous Tables}

\subsection{Fatigue, Encumbrance and Movement Charts}
\label{tables:encumbrance}
\label{tables:tmr}

\begin{tabular}[t]{cc}
{\fontsize{8}{11pt}\selectfont
\begin{tabular}[t]{l@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c@{\hspace{1.8em}}c}
\textbf{PS} & \multicolumn{8}{c}{\textbf{Weight of Load (lbs)}} & \textbf{Max} \\ \hline
3-5	&0	&0	&5	&14	&21	&30	&37	&45	&50 \\ \hline
6-8	&0	&5	&12	&17	&25	&40	&55	&67	&75 \\ \hline
9-12	&5	&12	&17	&25	&40	&60	&75	&90	&100 \\ \hline
13-17	&12	&17	&25	&40	&60	&80	&95	&112	&125 \\ \hline
18-20	&17	&25	&35	&50	&75	&105	&125	&140	&150 \\ \hline
21-23	&25	&40	&55	&70	&100	&140	&165	&185	&200 \\ \hline
24-27	&35	&50	&65	&85	&120	&160	&185	&202	&225 \\ \hline
28-32	&45	&65	&85	&105	&140	&180	&205	&230	&250 \\ \hline
33-36	&55	&80	&110	&140	&180	&220	&245	&262	&275 \\ \hline
37-40	&65	&85	&135	&170	&207	&247	&280	&307	&325 \\ \hline
\multicolumn{10}{l}{\textbf{Rate of Exercise}} \\
Light	&0	&0	&0	&$\frac{1}{2}$&$\frac{1}{2}$&1	&2	&3	&5 \\ \hline
Medium	&0	&0	&$\frac{1}{2}$&$\frac{1}{2}$&1	&1	&3	&4	&6 \\ \hline
Hard	&$\frac{1}{2}$&$\frac{1}{2}$&1	&1	&2	&3	&5	&6	&8 \\ \hline
Strenuous&2	&2	&3	&3	&4	&5	&6	&7	&9 \\ \hline
\multicolumn{10}{l}{\textbf{Agility Loss in Combat}} \\
Loss	 &0	&1	&2	&3	&5	&7	&9	&10	&12 \\ \hline
\end{tabular}}

&
\begin{minipage}[t]{2.5in}
\begin{tabularx}{2.5in}[t]{cXc}
& \textbf{Tactical Movement Rate} & \\
\textbf{Modified} & \\
\textbf{Agility} & & \textbf{TMR} \\ \hline
 $<$ 1		& & 0 \\ \hline
  1 -- 2	& &  1 \\ \hline
  3 -- 4	& & 2 \\ \hline
  5 -- 8	& & 3 \\ \hline
  9 -- 12	& & 4 \\ \hline
 13 -- 17	& & 5 \\ \hline
 18 -- 21	& & 6 \\ \hline
 22 -- 25	& & 7 \\ \hline
 26 -- 27	& & 8 \\ \hline
$>$ 27		& & \dag \\ \hline
\end{tabularx}

\dag TMR = 9 + 1 for every two points of AG over 28, for example AG 32 gives 11 TMR.
\end{minipage} \\
\end{tabular}

\begin{multicols}{2}

{\setlength\leftmargini{0pt}
\begin{description}
\setlength\itemsep{0pt}
\item[Weight of Load (lbs)] The weight, in pounds, that a character is
carrying, where each value is the category maximum.  Note: A mount can
carry weight for a character while they are riding.

\item[Max] The maximum load, in pounds, that a character can carry for
a sustained period of time.

\item[Rate of Exercise] See \S\ref{adventure:exercise}.

\item[Agility Points Lost] The temporary Agility Point loss suffered
by a character toting the given weight in combat.  Use the procedure
in rule \S\ref{adventure:fatigue} to use this chart.
\end{description}}

\end{multicols}

\rule{\linewidth}{0.6mm}

\begin{multicols}{2}
\subsection{Overland Movement Rate}
\label{tables:overland}

\begin{tabularx}{0.8\linewidth}{Xcccc}
& \multicolumn{4}{c}{\tiny\textsc{RATE OF EXERCISE}} \\
{\tiny\textsc{TERRAIN}} &
{\tiny\textsc{LIGHT}} &
{\tiny\textsc{MEDIUM}} &
{\tiny\textsc{HEAVY}} &
{\tiny\textsc{STRENUOUS}} \\
\hline
Cavern	& 5/-	& 10/-	& 15/-	& 20/- \\
\hline
Field	& 15/15	& 25/25	& 30/40* & 35/50* \\
\hline
Marsh	& -/-	& 5/5	& 10/10* & 15/15* \\
\hline
Plain	& 15/15 & 25/25	& 30/40* & 40/50* \\
\hline
Rough	& 10/5	& 15/10	& 20/15* & 25/-	\\
\hline
Waste	& 10/5	& 15/10	& 20/10* & -/-	\\
\hline
Woods	& 10/5	& 15/10 & 20/15* & 25/- \\
\hline
\end{tabularx}

\smallskip

The number before the slash indicates movement in miles per day on
foot; the number following the slash indicates mounted movement
(assuming horses).  Rates for other animals types must be adjusted by
the GM.  The day assumes a total of 8 hours marching.  Effects of
adverse weather must be adjudicated by the GM. Any paths or roads
negative the effect of other terrain, and the Plain movement rates are
used. (-): Movement type impossible at this exercise rate.

\begin{enumerate}
\item
Movement rates assume some sort of passable terrain; total swamp is
impassable except (sometimes) by canoes.

\item
If terrain is mountainous, halve rates on foot and prohibit mounted
movement.

\item
If terrain is desert, the GM should apply strictures regarding the
effects of heat, dehydration, etc.

\item
If terrain is deep woods, halve movement rates.
\end{enumerate}

* In these exercise rate categories, horses' maximum rates will
deteriorate 33\% per day.  They can travel at these rates for
approximately 4 consecutive days and then they will die.

\end{multicols}
