\begin{multicols}{3}

\section{Ranking}
\index{ranking|(}

Experience points are required to advance in anything.  Time spent
training is required to increase proficiency in spells, skills and
weapons.  Adventuring time is required to advance in characteristics
and talents.

EP is spent as per below but note the following.
\begin{itemize}

\item
Spells must be cast at least five times before the next rank can be
achieved.  Casting chambers are available in the Guild, and monitored
on request.

\item
Talents may be ranked only once per game adventuring week.

\item
Weapon skills take 1 week of training to reach Rank 0, and 2 weeks \x
(rank to be achieved) to improve.

\item
Skill ranks 8, 9, and 10 must be ratified by a GM and in general
required significant use of the skill.

\item
The character may rank any combination of two things at the same time,
providing the character does not rank magic (\ie spells or rituals)
at the same time as non-magic (\ie weapons or skills).

\item
Namers may rank 1 name in addition to other forms of ranking.  They
may also substitute ranking names for any other form of ranking.
Hence it is possible for a Namer to be ranking up to three names at
any one time.

\end{itemize}

All adventurers can learn any skill or weapon, and any magic within
their college. Learning costs EP, time and money.  The following can
be ranked: Skills, Weapons, Spells, and Languages, Names, Rituals,
Talents, Adventuring Skills and Characteristics.

\subsection{Skills}

All skills are assumed to be unranked (\ie unknown) initially. The
first level of competence is Rank 0, and will take eight weeks to
learn. Each subsequent rank will take that number of weeks to reach
(eg; to get to Rank 8 from Rank 7 will take 8 weeks). The EP cost for
ranking Skills is listed in the rulebook.  Some skills require minimum
Characteristic requirements to Rank, or impose EP penalties (or
discounts) for exceptional Characteristics.

If the character is taught by someone of greater Rank in the skill,
decrease any Experience Point cost by 10\%.  If the character learns
from a book (the availability of which is up to the GM), verbal
descriptions or practices with someone of equal or lesser Rank in the
skill, any Experience Point cost is unmodified.  If the character
practices with no useful outside assistance, any Experience Point cost
is increased by 25\%.  If training is done at the Guild, it costs
150sp \x (Rank to be achieved minimum 1). Achieving Ranks 8, 9 and 10
is difficult.  You must find and complete a special task relating to
your skill, with the assistance of a GM, for each of these Ranks. Rank
10 is the maximum achievable Rank in all Skills.

\subsection{Weapons}

All weapons are assumed to be unranked initially.  Rank 0 in a weapon
takes 1 week. All higher ranks take 2 \x (Rank to be achieved)
weeks. Weapons have individual maximum Ranks.  EP costs are detailed
in \S\ref{tables:ep}.  All Weapons require minimum PS and MD
Characteristics. If you do not fulfil both requirements, you may not
rank a weapon.  You may not get an EP discount for training, but if no
trainer is available, you may not increase in Rank.  The cost of a
trainer is 10 \x (Rank to be achieved squared) silver pennies.

\subsection{Spells}

If you are an Adept (\ie cast magic), you can rank your spells.
Each spell has an EM, or Experience Multiplier. This is multiplied by
the Rank that you wish to achieve, to give a total EP cost. If the
Adept has MA $>$ 15, (MA - 15) \x 5\% of the EP cost of General Spells
may be discounted. Training time for spells is (Rank to be achieved)
days.  All spells must be cast five times before each new rank is
achieved. As these spells may backfire, these rolls will need to be
made at some stage.  Learning a new spell to Rank 0 takes (EM / 100
rounded up + 1) weeks, but no experience points. See Handbook for
availability of special knowledge spells.  You can not have more
spells and rituals below rank 6 than your MA characteristic.  Rank 20
is the Maximum Rank achievable with any Spell (except Geas).

\subsection{Languages}

Languages have the same time requirement as normal skills, except that
the time for Rank 0 in a language is only 1 week.

The undiscounted EP costs are set out in Table \S\ref{epcosts:skills}.
Note that knowing the Philosopher Skill may grant an EP discount. The
maximum total EP discount applicable is 50\%, regardless of how many
individual discounts are available to the character.

If appropriate, the masterwork to gain Rank 8, 9, or 10 may be the
same masterwork as one presented for a skill, \eg Philosopher,
Troubadour, Courtier, etc.  Learning magic counts as a masterwork.

\subsection{Names}

Anyone can rank Names, but only Namers and E \& Es gain any normal
advantage from this.  Learning a Generic Name takes one day; learning
an Individual Name takes one week. Ranking a Generic Name takes (Rank
to be achieved) weeks; Ranking an Individual Name takes the same
number of months. No EP is required. Namers get special advantages in
Ranking (\S\ref{college:namer}).

\subsection{Rituals}

Rituals are learnt and ranked just like spells, except that Ranking
time is (Rank to be achieved) weeks, rather than days.  MA discount
applies to General Knowledge Rituals.

\subsection{Talents}

For each week of actual, out in the field, adventuring, you can rank
each of your talents once.  No training time is required to rank
Talents. Like Spells and Rituals, each Talent has an EM. No MA
discounts apply to any Talents.

\subsection{Adventuring Skills}

Adventuring Skills are skills used every day by adventurers to
survive, and thus are continually honed. These skills include
Knowledge, Horsemanship, Swimming, Flying, Stealth and Climbing.

If you have extensively used an adventuring skill while on adventure,
you may rank this skill without any time requirements. Otherwise,
Ranking time is as per normal skills.

Adventurers are assumed to start off with Rank 0 or more in all these
skills, unless specifically told otherwise. Experience Point costs are
set out in the rulebook.  The Maximum Rank in all these Skills is 10.

\subsection{Characteristics}

A Characterisitic may only be raised by five points over its
\emph{starting value} to a maximum of 25 (modified by racial bonuses /
penalties), except Fatigue, and Physical Beauty (which cannot be
raised).  The exception to this is Perception, which can be raised to
racial maximum.

To calculate the maximum for Fatigue, take the racial maximum for
endurance and find the maximum Fatigue from the chart (see character
generation).  For example, dwarves have +2 to EN so their maximum EN
is 27, hence their maximum FT is 24 from the chart.  Then apply any
additional racial modifiers to FT.  For example, orcs gain +2 to FT so
their maximum is $24 + 2 = 26$.

A Characteristic point may only be increased once per adventuring
session (if a session took more than the normal session length, then
this rule should be applied appropriately).  If a character did not
adventure during a session then they cannot raise any characteristic
points, and if they participated on more than one adventure during the
session, then they can still only raise a characteristic point by one.
Any or all of the stat points may be raised simultaneously if
permissable.

If a character has lost characteristic points for any reason, they may
buy back as many points as they wish in addition to any normal
increase.  The cost of buying a characteristic point back is the same
as buying an extra point (see table \S\ref{epcosts:stats}).

\index{ranking|)}
\raggedcolumns\newpage\end{multicols}
