\section{The Elohim}

\subsection{History \& Background}

It was in the last years of the Elven Empire, as the power of the Drow
was rising, and elven culture was stagnating in decadent luxury, that
the group called Noldanor was formed. In elven their name meant
Knowledge of Light or Illuminated, and they appointed themselves the
task of stopping the spread of the decay that threatened their
civilisation. They believed in purity of spirit, and discipline of
mind and body. They saw the casual use of magic as one of the prime
causes for the decadence of the elven people, and counselled against
it. In the end, the task proved too large, and their numbers too
few. As the fall began, and the war spread, they fought to the last
with a bravery that has seldom since been seen in all of the ages of
the world.

Although all of the Noldanor perished in the fall, their wisdom and
beliefs did not completely leave the world, for they had planned and
trained for this outcome. The spirits of the Noldanor passed beyond,
where, because of their dedication and training, they retained traces
of their former existence, and eventually formed immortal beings, the
Powers of Light, composites of the mortal beings that once were.

In the later ages of the world these Powers have sought to guide
mortals, to prevent another fall, and to protect them against the
designs of the Powers of Darkness, Demons who seek to seduce and use
mortals for their own ends.

Whilst the Powers of Light are technically Demons themselves, the term
is never used by them to refer to themselves, but rather used as a
generic term for the Powers of Darkness and any other Powers opposed
to their cause. It is mortals who have given to the Powers of Light
the name by which they are now known: Elohim, the Lordly ones.

\subsection{Organisation}

The Elohim possess a structured order and hierarchy comprising of
seven tiers. There are four great Powers of Light, called Archangels,
each with a broad sphere of influence. Each Archangel leads a host of
lesser Powers, the Bene Elim, who fill subordinate offices within the
Archangel's sphere. These are in turn served by Elohim of decreasing
power, the Malakim and Aishim. The lowest three tiers form the
servants (or minions) of the Elohim.

The seven tiers, their titles and meanings are:

\begin{description}
\item[Archangels]
\item[Bene Elim] Mighty Ones
\item[Malakim] Kings
\item[Aishim] Flames
\item[Erelim] Valiant Servants
\item[Kerubim] Servants of Knowledge
\item[Seraphim] Servants of Light
\end{description}

\subsection{Agents}

The Elohim seek mortal Agents to spread their teachings, and to work
against the plans of the Demons. The Elohim prefer elven or human
Agents, but will accept those of any race provided they meet their
stringent criteria. Each of the Archangels looks for different
abilities in their Agents, but all four only accept mortals who are
pure of spirit, brave, dedicated, and completely committed to their
cause.

Because the Elohim see the irresponsible use of magic as one of the
leading causes for the fall of the ancient elven world, they are
deeply concerned for those that delve in the magical arts. They
counsel the limiting of magical training to only those who show
integrity, strong morality, and personal responsibility.

Most of the Elohim's mortal Agents are followers of one of the four
Archangels. Relations between the Archangels are usually cordial,
although they do not always agree on matters affecting the
Elohim. Occasionally, Agents of one of the four may find that they are
working for goals opposed to the wishes of the other Archangels, and
may find themselves in conflict with other Powers of Light Agents.
This conflict of interests seldom leads to violence, and Agents will
usually strive for some compromise.

\subsection{The Archangels}

The four great Archangels are Michael, Gabriel, Raphael, and
Uriel. Their spheres, personalities, Avatar abilities, and other
information about them is listed below.


\subsubsection{Michael: "The Sword, Sword of Light"}

\begin{Description}
\item[Type] Demon

\item[Group] Elohim

\item[Response] 15\% / -

\item[Description] Michael always appears as a stunningly handsome
young, blond, human or elven warrior, dressed all in red, and armoured
for war, wielding his great two-handed sword, Ira (Wrath), that blazes
with the fire of his righteous anger. Michael may appear with great
feathered wings, or riding a sentient pegasus.

\item[Sphere] Michael's office is that of warrior and he leads the
Elohim's fight against the Powers of Darkness. He destroys and
confounds the works of evil and the Demons whenever possible, and
charges his Agents, the Michaelines, to do the same.

\item[Personality] Michael is rash and bloodthirsty and prefers to
charge into battle, so as to defeat the unholy by strength of arms.
He impatiently awaits Nightfall, (the resumption of the War of the
Powers), and has been known to be involved in many plans to hasten its
arrival.

\item[Greater minions] Erelim 

\item[Lesser minions] Kerubim 

\item[Minor minions] Seraphim 

\item[Beast minions] Sentient Pegasi (MA 0)

\item[Animal minions] Boar

AGENTS: Michael is extremely suspicious of Mages and will very seldom accept them as Agents. He expects all his Agents to be warriors, and to fight courageously against the Powers of Darkness and their Agents and followers. Michaelines may be found investigating rumours of Demonic cults, and the doings of "Black Mages". Once discovered they will attempt to convince civil authorities to deal with cults and covens, and will lend their martial prowess in eliminating them. They also root out heresy and corruption in established churches, and are often found acting as the martial arm of the Gabrielite inquisition. They tend to view any non-Elohim Agents with the gravest suspicion.  Michaelines have a well founded reputation for being rather blood-thirsty and over zealous in the pursuit of their goals.

AREAS: Michael's Agents may be found in all areas, though they tend to be concentrated in borderlands and trouble spots. In territories where the Elohim and their Agents are officially recognised, the Michaeline Order constructs chapter houses that serve as barracks and training facilities, and are usually fortified strongholds. These chapter houses recruit "Knights of Michael", fighting lay members dedicated to the cause of the Elohim. Agents of Michael will always be able to get free board and weapons training at these chapter houses.

COLOURS \& SYMBOLS: Michaelines often wear red garments, or white robes edged with red, and their symbol is a flaming sword (usually shown point downwards). They are seldom seen not wearing armour.

AVATAR ABILITIES: Michael is a mighty Warrior, and a superb Military Scientist (even though he will tend to counsel attack in most situations). Michael is very beautiful to look at and may inspire Awe. Michael cannot cast magic but may initiate various magical talents upon himself as a Pass action. These include: Quickness, Coruscate, and ESP. Michael's Magic Resistance is special, in that he has 99% MR normally, and in addition has an automatic 52% Active Resistance that is always operating even if he cannot see the Mage that has targeted him.

Movement Rate: (yards per minute): Run: 350; Fly: 600

PS: 30	MD: 28	AG: 32	MA: 10
EN: 26	FT: 38	WP: 32	PC: 30
PB: 30	TMR: 8	NA: 3 DP (+10 pt. Plate Armour)
MR: 99/52 (Special)

Weapons: Michael always wields his two-handed sword, which is of demonic quality, and with which he is Rank 9. His sword may be considered magical and can strike any target, even those usually immune to material weapons (equivalent to Spectral Weapon Rank 20). He has Rank 10 Unarmed Combat, and may use most other weapons at maximum Rank. He wears demon crafted red plate armour of some light-weight material. Whenever he enters combat Michael becomes surrounded by a coruscating aura (equivalent to Coruscate Rank 20).

AVATAR SUMMONING: Michael will seldom manifest except in the thick of battle, where he will immediately plunge into the combat wielding his great sword.
\end{Description}

\subsubsection{Gabriel: "The Herald, The Messenger"}

\begin{Description}
\item[Type] Demon
GROUP: Elohim
RESPONSE: 15%
DESCRIPTION: Gabriel manifests as either a dazzling human youth, with almost feminine features, or as a young elven woman of flawless beauty. Gabriel is dressed in blue robes, and bears a sword and an heraldic trumpet, and may in rare circumstances appear in armour.  The trumpet has a banner attached, on which its name, Veritas (Truth), is embroidered in gold. Gabriel may appear with great feathered wings, or riding a sentient pegasus. Gabriel's voice is melodious and mellifluous.
SPHERE: Gabriel's office is that of herald, messenger, evangelist, and inquisitor.  Gabriel's agents, the Gabrielites, are charged to spread the word of the Elohim, to convert entities to the faith, and to root out heresy and evil-doing. 
PERSONALITY: Gabriel is impulsive and often makes rash decisions.  Gabriel is closely associated with Michael, and often defers to him on matters of policy. Followers of the Elohim state the Gabriel will one day signal the resumption of the War of the Powers, and the onset of Nightfall, with a mighty trumpet blast, that will be heard on all planes and dimensions. Other observers have commented that Gabriel seems all too eager to discharge that particular duty.
GREATER MINIONS: Erelim 
LESSER MINIONS: Kerubim 
MINOR MINIONS: Seraphim 
BEAST MINIONS: Sentient Pegasi (MA 0)
ANIMAL MINIONS: Hawk
AGENTS: Gabriel is suspicious of Mages, but accepts their worth  and will accept as Agents those that have unequivocally proved their austerity and self discipline. Gabrielites are often trained as scribes in their churches and monasteries, and then sent out as missionaries to preach to the heathen (those who do not follow the Elohim's teachings). They are also taught the arts of war, so as to aid in getting their point across. Gabrielites check on the purity of the doctrine in established churches, and work along side Michaelines as Inquisitors. 
AREAS: Gabrielites are found both in civilised areas, where they serve in the churches as clerks and preachers, and also in the most untamed wilderness, where they seek out new converts, and look for cults and other signs of corruption. Because of their close association with the Michaelines, Agents of Gabriel can always get free board and weapons training at Michaeline chapter houses.
COLOURS \& SYMBOLS: Gabrielites usually wears robes of blue, or white edged with blue, and their symbol is an heraldic trumpet. They usually only wear armour when dealing with the heathen.
AVATAR ABILITIES: Gabriel is a mighty Adept of the College of Bardic Magics, and possesses  the skills of Courtier, Spy, and Troubadour. Gabriel is stunningly beautiful to look at and may inspire Awe. Gabriel may Charm as a Talent.

Movement Rate: (yards per minute): Run: 350; Fly: 600

PS: 24	MD: 24	AG: 30	MA: 30
EN: 22	FT: 32	WP: 26	PC: 26
PB: 35	TMR: 8	NA: 2 DP (+10 pt. Plate Armour)
MR: 60

Weapons: Gabriel is not a great fighter, but will do so if pressed, using a hand \& a half sword at Rank 7. This sword is of demonic quality, and may be considered magical, able to strike any target, even those usually immune to material weapons (equivalent to Spectral Weapon Rank 20). Gabriel may wear demon crafted blue plate armour of some light-weight material, although this is rare. Upon entering combat Gabriel becomes surrounded by a coruscating aura (equivalent to Coruscate Rank 20).
AVATAR SUMMONING: Gabriel may be summoned into any place of learning, or anywhere that heresy has been discovered.
\end{Description}

\subsubsection{Raphael: "The Shield, Defender of the Faithful"}

\begin{Description}

\item[Type] Demon
GROUP: Elohim
RESPONSE: 15%
DESCRIPTION: Raphael usually appears as a handsome mature man, dressed in green and outfitted for war, either with large, feathered wings, or more often, seated upon a pegasus. He bears a lance and his shield Fidei (Faith).  His shield is green and carries in gold the image of a crook, or crosier, reflecting his role as a shepherd and protector. He will often bring with him a pack of hunting or war dogs.
SPHERE: Raphael's office is that of the defender of the faithful, teacher and guide.  He charges his Agents, the Raphaelites, with the protection and education of their followers.
PERSONALITY: Raphael is careful and cautious, and will seldom counsel attack, but rather a brave and stalwart defence. He often finds himself in opposition to the more insane plans of both Michael and Gabriel, and co-operating with Uriel in matters of protection.
GREATER MINIONS: Erelim 
LESSER MINIONS: Kerubim 
MINOR MINIONS: Seraphim 
BEAST MINIONS: Sentient Pegasi (MA 0)
ANIMAL MINIONS: Dog
AGENTS: Raphael will accept Mages provided they can prove their bravery, loyalty, and discipline. His Agents may be found as guards, generals, or teachers, or as questing knights who seek to rid the world of dangers to their people, such as Dragons. Raphaelites tend to be the least impetuous of the Agents of the Powers of Light, and more reserved and conservative, as befits those whose office is defence.
AREAS: Raphaelites are usually found in settled and civilised areas where they act in roles that help to protect their people. When found in wilderness areas Raphaelites will usually be on a mission to destroy a particular threat to the safety of their "flock".
COLOURS \& SYMBOLS: Raphaelites often wear robes of green, or white edged with green. Their symbol is a shield or a crook. Raphaelites often wear armour.
AVATAR ABILITIES: Raphael is a great Armourer, Mechanician, Military Scientist, Ranger, and Warrior, and possess lesser abilities as a Beast Master, Healer, and Weaponsmith. Raphael is very handsome to look at and may inspire Awe. Raphael cannot cast magic but may initiate various magical talents upon himself as a Pass action. These include: Quickness, Coruscate, and ESP. Raphael's Magic Resistance is special, in that he has 95% MR normally, and in addition he has his Active Resistance chance of 54% of "parrying" an incoming spell with his shield, so that it rebounds and targets the caster instead. He can only do this if he can see the Mage that has targeted him.

Movement Rate: (yards per minute): Run: 350; Fly: 600

PS: 28	MD: 26	AG: 28	MA: 10
EN: 32	FT: 38	WP: 34	PC: 36
PB: 27	TMR: 8	NA: 5 DP (+10 pt. Plate Armour)
MR: 95/54 (Special)

Weapons: Raphael may use most weapons at maximum Rank, and prefers knightly weapons. He carries a lance and hand \& a half sword, which are of demonic quality, and with which he is Rank 6 \& 8 respectively. These weapons may be considered magical and can strike any target, even those usually immune to material weapons (equivalent to Spectral Weapon Rank 20). He wears demon crafted green plate armour of some light-weight material. Whenever he enters combat Raphael becomes surrounded by a coruscating aura (equivalent to Coruscate Rank 20).
AVATAR SUMMONING: Raphael can be summoned whenever or wherever defence is needed. 
\end{Description}


\subsubsection{Uriel: "The Judge, Lady Justice"}

\begin{Description}

\item[Type] Demon
GROUP: Elohim
RESPONSE: 15%
DESCRIPTION: Uriel usually manifests as a beautiful, dark-haired human or elven woman, dressed in purple robes, edged in gold, and bearing a sword and her balances or scales, Jus (Law). Uriel may appear with great feathered wings, or riding a sentient pegasus. She may appear wearing purple plate armour if she feels it warranted.
SPHERE: Uriel's office is the gathering of souls, the judging of their behaviour, and the choosing of the final rewards accorded to the followers of the Elohim in the afterlife. She charges her followers, the Urielites, with the dispensing of justice, and the punishment of criminal and unethical behaviour.
PERSONALITY: Uriel is stern but fair, brooking no argument with her judgements. She does know compassion, however, and may reduce the severity of her judgement if she feels that the subject had cause for their actions.
GREATER MINIONS: Erelim 
LESSER MINIONS: Kerubim 
MINOR MINIONS: Seraphim 
BEAST MINIONS: Sentient Pegasi (MA 0)
ANIMAL MINIONS: Owl
AGENTS: Uriel will accept Mages as Agents, favouring Thaumaturges. Those of her Agents that are Mages tend to be even more fervent than usual in finding and bringing to justice criminals who use magic. Her Agents are often judges and advocates, but also jailers, guards and even executioners.
AREAS: Urielites are usually found in civilised areas. In areas where the Powers of Light are very strong, the Agents and followers of Uriel are often accorded the legal right to try and punish criminals. Urielites found in wilderness areas will almost certainly be hunting escaped criminals, and may be accompanied by Michaelines or Raphaelites.
COLOURS \& SYMBOLS: Urielites often dress in the manner of their patron, in gold edged purple robes, or in white robes edged with purple, and sometimes have their armour embossed with their symbol, a pair of scales.  Urielites wear armour when it is fitting for their job or station, or when hunting criminals.
AVATAR ABILITIES: Uriel is a practitioner of the College of Sorceries of the Mind. She is a great debater and Philosopher, Astrologer and Courtier. Uriel is dazzlingly beautiful and may inspire Awe. Uriel can always detect if someone is knowingly lying to her, and can cause a mortal to feel great pain every time that they lie. Uriel's Magic Resistance is special in that she will always resist magic that is cast at her unjustly, or with criminal intent. Otherwise her MR is 66.

Movement Rate: (yards per minute): Run: 350; Fly: 600

PS: 26	MD: 22	AG: 26	MA: 28
EN: 22	FT: 30	WP: 36	PC: 32
PB: 32	TMR: 7	NA: 2 DP (+10 pt. Plate Armour)
MR: Special

Weapons: Uriel is not a great fighter preferring to work through legal and diplomatic channels rather than open combat. If forced into combat she uses her Hand \& a half sword, which is of demonic quality, and with which she is Rank 7. Her sword may be considered magical and can strike any target, even those usually immune to material weapons (equivalent to Spectral Weapon Rank 20). She may wear demon crafted purple plate armour of some light-weight material. If she enters combat Uriel becomes surrounded by a coruscating aura (equivalent to Coruscate Rank 20).
AVATAR SUMMONING: Uriel can only be summoned into a place of justice (court, prison, etc.), or to where a crime has recently been committed. 
\end{Description}

\subsection{Subordinate Offices}

The offices of the other Elohim are specific sub-sets of the spheres of the Archangels.  For example, Azrael, "The Executioner", is a Bene Elim (or Angel) of Uriel's host, whose office is the execution or slaying of those condemned for their heinous crimes. Agents of the Powers of Light who are executioners or scourges, may have Azrael as their Patron, but still consider themselves Urielites. This "tree" of offices extends down through the Malakim to the Aishim. Thus an Agent may have as their Patron an Aishim whose office is simply the defence of a particular sacred well, but may still be considered a Raphaelite.

When creating a new Elohim, GMs should decide what office they hold.  From this it may be ascertained which host they will belong to, and from the scope of the office, what degree in the hierarchical order they hold. 

\subsection{Elohim Names}
The names chosen by the Elohim follow a somewhat standard format, offered here as an aid to GMs.  The Elohim invariably use the suffix -el with their common names.  This suffix may be roughly translated as Lordly, and is used even by the servant Elohim.  To generate an Elohim's name, choose one syllable from the Primary list, append one or two syllables from the Secondary list and add el.

PRIMARY:
A, Ab, Ak, Ap, As, At, Az, Ba, Bel, Ca, Cher, Ches, Dan, E, Ez, Ga, Gab, I, Is, Jet, Ka, Ky, Lo, Mic, Na, O, Ra, Rach, Raz, Re, Sag, Ser, U, Ve, Yah, Zaph, Zek, Zu.

SECONDARY:
a, aph, as, da, di, ek, et, f, fa, fi, g, ga, gi, ha, hi, hu, i, ki, ma, mi, na, ni, pha, qu, ra, re, ri, ru, sa, si, ub, ubi, za, zi. 

\subsection{Agents Names}

Followers of the Elohim often choose a new name upon achieving Agent
status. Most often these names follow the same format as the names of
the Elohim themselves, save that they omit the -el, or lordly suffix,
and often substitute -im, meaning servant.

\subsection{Minions of the Elohim}

\subsubsection{Erelim "Valiant Servants"}

\begin{Description}

\item[Class] Greater Minion
Description: Erelim (also sometimes known as Thrones) appear as mighty humanoids, with the height and build of hill giants. They are however extremely good-looking, and appear more like giant-sized elven and human men and women, with immense feathered wings. Erelim radiate a dazzling inner light, and glow strongly in all lighting. They may choose to assume a less obvious form, concealing their wings, shrinking to 6 foot or so, and reducing their glow. They may appear armoured for war, or wearing  short robes (well tailored and cut to reveal their powerful bodies). Their armour or clothes will tend to be in the colour of their faction. Erelim usually appear armed with a giant bow, a giant spear, and a large sword and shield. They may use these even if they assume human size.
Talents, Skills and Magic: Erelim of Uriel are almost all Mages. Some of those of Raphael's and Gabriel's factions are Mages, but Erelim of Michael are not. Erelim that are Mages all belong to one to the Thaumaturgical Colleges. They will have Ranks 12-15 with the General and Special Knowledge of their College. All have 1 or 2 skills at Rank 10, these skills varying depending on their faction. Non-Mage Erelim may also have the Warrior skill, usually also at Rank 10. The presence of an Erelim may cause Awe.
Movement Rates: (yards per minute): Run: 450; Fly: 500
PS: 35-40		MD: 18-25	AG: 20-26	MA: 5-30
EN: 30-40	FT: 45-50	WP: 24-30	PC: 25-30
PB: 27-28	TMR: 9/10	NA: 3 DP (+10 pt. Plate Armour)
MR: 24-50
Weapons: Erelim usually have giant bow, giant spear, and 1 sword and shield at maximum ranks, and Warrior Erelim will know a large number of other weapons, at lower ranks. They may choose to appear in well-crafted plate armour of some light-weight material, and carrying a giant bow, giant spear, and sword and shield of their choice. These weapons will be demon crafted and have bonuses to BC and/or damage. When Erelim enter combat they become surrounded by a coruscating aura (equivalent to Coruscate Rank 20).
\end{Description}

\subsubsection{Kerubim "Servant of Knowledge"}

\begin{Description}

\item[Class] Lesser Minion
Description: Kerubim usually appear as extremely tall (7' - 7'6"), good-looking elven and human men and women, of Herculean or amazonian stature. All have large feathered wings, although they may choose to conceal these at will. Kerubim radiate a strong inner light, and glow faintly in all lighting. They wear short robes (cut to reveal their muscled physique) or armour, in the colour of their faction, and usually appear armed with spear,  sword and shield.
Talents, Skills and Magic: Kerubim of Raphael and Uriel tend to be Mages. Some of those of Gabriel's faction are Mages, but most Kerubim of Michael are not. Kerubim that are Mages tend to belong to one to the Thaumaturgical Colleges, but only a rare few are Elementalists. Kerubim of the Entities Colleges are almost unknown. They will have Ranks 10-12 with the General and Special Knowledge of their College. Most have 1 skill at Rank 9, this skill varying depending on their faction. Non-Mage Kerubim may also have the Warrior skill, usually also at Rank 9.
Movement Rates: (yards per minute): Run: 400; Fly: 500
PS: 28-30		MD: 18-25	AG: 22-28	MA: 5-30
EN: 28-34	FT: 35-40	WP: 22-28	PC: 20-25
PB: 20-26	TMR: 8-10/10	NA: 2 DP (+9 pt. Plate Armour)
MR: 22-48
Weapons: Kerubim usually have spear, and 1 sword and shield at maximum ranks, and Warrior Kerubim will know a large number of other weapons, at lower ranks. They may choose to appear in well-crafted plate armour of some light-weight material, and carrying a spear, sword and shield of their choice. When Kerubim enter combat they become surrounded by a coruscating aura (equivalent to Coruscate Rank 15).
\end{Description}

\subsubsection{Seraphim "Servant of Light"}

\begin{Description}

\item[Class] Minor Minion
Description: There are many more Seraphim than any other type of Elohim, for they are the manifestations of the individual souls of the original Noldanor, who died many aeons ago, and also of the many Agents of the Powers of Light who have fought for their cause throughout the later ages of the world. Seraphim usually appear as tall and handsome elven and human men and women, although some are of other races. All have large feathered wings, radiate an inner light and well-being, and glow faintly in dim lighting. They wear robes or armour in the colour of their faction and usually appear armed with sword and shield.
Talents, Skills and Magic: Many Seraphim are members of one of the Colleges of Magic, but as many are not Mages. Those that are tend to belong to one to the Thaumaturgical Colleges, but some are Elementalists. Seraphim of the Entities Colleges are very rare, but not unknown. They will have Ranks 6-12 with the General and Special Knowledge of their College. Most have 1 skill at Rank 7, this skill varying depending on their faction. Non-Mage Seraphim may also have the Warrior skill, usually at Rank 8.
Movement Rates: (yards per minute): Run: 350; Fly: 500
PS: 18-26		MD: 18-25	AG: 22-27	MA: 5-27
EN: 20-26	FT: 22-24	WP: 22-27	PC: 15-25
PB: 18-25	TMR: 7-10/10	NA: 1 DP (+8 pt. Plate Armour)
MR: 22-48
Weapons: Seraphim usually have 1 sword and 1 shield at maximum ranks, and Warrior Seraphim will know a large number of other weapons, at lower ranks. They may choose to appear in well-crafted plate armour of some light-weight material, and carrying a sword and shield of their choice. When Seraphim enter combat they become surrounded by a coruscating aura (equivalent to Coruscate Rank 12).
\end{Description}

