\section{Invocations (Version 1.1)}

\subsection{Power's Names}

Powers may be aware when their names are spoken.  Those Powers that
have a second "Response" percentage listed after their name are aware
when their listed common name or Individual True Name is spoken aloud.
Most Powers will not respond in any manner if their name is merely
used in casual conversation, or the like, but will do so if their name
is being spoken as part of an Invocation.  The Powers are not aware
when their titles, sobriquets, or nicknames are spoken.  Those Powers
that have a second Response percentage listed have that chance of
responding to the use of their name in casual conversation.  If their
Individual True Name is spoken, increase the chance by 25\%.  Those
Powers that have 0\% listed as their second Response percentage are
aware that their name has been used but never respond to that casual
usage.  Note that most Powers are not aware when their names are used
casually.

\subsection{Simple Invocations}

The "Response" percentage listed amongst the Power's attributes gives
the chance on D100 that the Power will respond to their Common Name
spoken as part of an Invocation.  If it is the Power's Individual True
Name that is being spoken, the percentage chance that they will
respond should be raised by 25.  The Power's response may take one of
three forms:
\begin{Enumerate}

\item
The Power may choose to despatch a Minor minion to spy on the Invoker,
ask them the purpose of their invocation, or even to slay them.  This
minion will be of the common Minor type for that Power and if
intending to speak with or slay the invoker, they will appear within
D-5 minutes.  Their appearance is almost invariably "natural"; flying
creatures will fly through a window, creeping ones may crawl from
behind furniture, etc.

\item
The Power may despatch an animal Minion to serve the same purpose as
in 1.  This animal will be of a type commonly associated with the
Power, it will be sentient, but otherwise unexceptional.

\item
The Power may choose to project an insubstantial image of themself.
This image will usually be of their common form, and has no powers
other than to communicate with the invoker or to confer upon the
invoker the status of Agent, along with the Call Patron Talent; this
latter ability only being available if the Invoker used an Ancient
Invocation.
\end{Enumerate}

\subsection{Ancient Invocations}

Powers will more often respond to Ancient Invocations.  In its
simplest form, an Invocation is merely the use of a Power's name in a
formal sense, along with the willingness for the Power to respond.  A
simple "I call upon the power of ...", or the use of the Power's name
as a chant or mantra is just as effective as the most elaborate ritual
that the Invoker can conceive.  Certain Ancient Invocations may
however have gained great power through their extreme age, and thus
have a greater chance of causing a Power to respond to their use.
These Ancient Invocations may be found in rare books in arcane
libraries, and usually only through a great deal of searching.  The GM
must determine the additional chance that an individual Invocation has
of causing the Power to respond, along with any special conditions or
ingredients that are required to perform it as well as the amount of
time required.

EX. "Ritual for Invoking Aim, (The Fire Duke), +10\%, requires a very
large bonfire, 10 minutes".
