\section{Presence (Version 1.0)}

An extremely high or low value for Physical Beauty may have
unpredictable results.

\subsection{Fright}
Whenever characters encounter a creature whose Physical Beauty is less
than 5, and whose description states a Fear causing ability, they must
make a Willpower check to determine if they are frightened. The
difficulty factor for this test is equal to the creatures
PB. Creatures whose PB is 0 cause a 0.5 difficulty factor.  If this
test is failed, the character must then roll on the Fright Table and
apply any results before they take another action.  They may attempt
to recover every pulse, by trying to succeed at the same Willpower
check.  Until that time they will act as the Fright Table indicates.

\subsection{Awe}
Whenever characters encounter a creature whose Physical Beauty is
above 26, and whose description states an Awe causing ability, they
must make a Willpower check to determine if they are awed. The
difficulty factor is dependent on the creatures PB:

\begin{tabular}{ll}
	& DIFFICULTY \\
PB	& FACTOR \\
27--28	& 4.0 \\
29	& 3.5 \\
30	& 3.0 \\
31	& 2.5 \\
32	& 2.0 \\
33	& 1.5 \\
34	& 1.0 \\
35+	& 0.5 \\
\end{tabular}

If this Willpower check is failed the character must then roll on the
Awe Table (see \S\ref{tables:awe}) and apply any results before they
take another action.  They may attempt to recover from being over-awed
as indicated on the Awe Table.

\subsection{Familiarity}
In both the cases of the extremely low and high beauty, the effects
are reduced with familiarity with the creature causing the fear or
awe.  Once a character has successfully made their Willpower check, or
has recovered from a failed check, the), need not check again (for an
effect from the same creature) for the remainder of the encounter.  If
the same creature is encountered again another Willpower check must be
made, but the GM may add 1 or more to the difficulty factor, thus
making it easier to succeed.  This bonus should not be applied to
encounters with other similar creatures, only to the same ones. (It is
not true to say that when you have seen one Troll, you have seen them
all).

\subsection{Magic Resistance}
These Fear and Awe effects are not magical, and so no magical
resistance check is applicable.

