\documentclass[a4paper]{article}

\usepackage[times]{dq}

\begin{document}

\begin{college}[1.1]{enchanter}{Ensorcelments and Enchantments}{EN}


\begin{spell}[S-20]{Enchant Aura}

\range{5 feet + 1 / Rank}
\duration{1 minute + 1 / Rank}
\multiple{500}
\basechance{15\%}
\resist{Passive}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
This spell will change one aspect of the target's aura.  This change
will show up with Detect Aura (so a Namer will see both the enchanted
and true auras).  The enchantment does not change the true aspect of
the target, so for example, the enchantment to make somebody have the
aura of death aspected will not gain death aspect.  The enchantment
will affect any magic which relies on aspects or auras, sure as wards
or various element spells.
\end{effects}
\end{spell}

\begin{spell}[S-21]{Install Flight}

\range{Self only}
\duration{5 minutes + 1 / Rank}
\multiple{500}
\basechance{15\%}
\resist{None}
\storage{Investment}
\target{Object}
\begin{effects}
This spell enables the Adept to instil a possession of up to 5 lbs (+5
/ Rank) with the power of flight. The spell will dissipate if the
object stops being a possession of the Adept, the Adept loses
concentration, or if the object is broken. The Adept may cause the
object to fly at 20 miles per hour (+ 2 / Rank). It will take off and
accelerate up to full speed, or halt and land, in a single pulse. The
object will support the weight of the Adept (+ 5 / Rank) in addition
to its own weight. Naturally flexible or fragile items gain sufficient
strength and rigidity to support the load. Any object or entity that
falls from the flying object will move off in a random direction.

At Rank 20 acceleration and deceleration becomes instantanous.
\end{effects}
\end{spell}

\end{college}

\begin{college}[2.0]{rune}{Rune Magics}{RU}
\begin{spell}[S-20]{Shapechange}

\range{Self only}
\duration{10 mins + 10 / Rank}
\multiple{300}
\basechance{15\%}
\resist{None}
\storage{None}
\target{Adept}
\begin{effects}
The Adept can transform themselves into any common animal (includes
aquatics and avians), gaining their physical stats (PS, MD, AG, EN \&
PB) and characteristics.  The creature however keeps the Adepts mental
statistics and health (MA, WP, FT \& PC).  All possessions are
transformed along with the Adept.  Transformations take a full pulse
to happen.  The Adept may drop the spell and will transform back the
following pulse.  At Rank 10 the Adept can transform into a two hex
creature and at Rank 20 can transform into a three hex creature.

At Rank 10 the Adept may transform between their animal and original
form without terminating the spell.  At Rank 20 the Adept may alter
their animal form at will, changing shape the following pulse.

The base chance can be modified as follows:
\begin{itemize}
\item +10\% for knowing the creature's runic name (ie GTN).
\end{itemize}

\end{effects}
\end{spell}
\end{college}
\end{document}

