\documentclass[a4paper]{article}

\usepackage{ep}

\usepackage{multicol}

\title{Items salvaged from the Pyramids of Blood}
\author{Ross Alexander}
\date{29 January 2004}



\begin{document}

\maketitle

\begin{ep-box}{Potion of Control Person}

This potion can be used to control a sentient entity (multihex
creatures make a size $\times$ EN save to resist).  The potion is
activated by controller placing a drop of their blood into the potion.
The potion must be used within 24 hours or it will become useless.

The potion can be either used by making the victim drink it or it can
be injected into the victim but it cannot be used as a venom.  The victim
is then controlled as per the Mind college ritual Bind Will and controlled
until the next full moon.

This control can be removed as if it was a MA 30 major curse.
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Potion of Contagion}

This potion contains a viralent strain of bubonic plague.  The potion
is potent enough to contaminate the water supply of a medium sized
town (approximately 2000 people).  Any victim must make a 1 $\times$
EN save or become infected.  The infection is 50\% contagious though
air and 100\% contagious through blood, saliva or other bodily fluids.
Infected people will suffer legions, boils, open running sores and
internal hemoraging including major bleeding in the stomach and colon.
The victim must make another 1 $\times$ EN save 72 hours after first
infection.  If they fail they will die within another 24 hours,
otherwise their will remain infected for another 72 hours before
recovering.  A victim cannot be infected more than once.

This plague will run indefinately until the population has either died
or recovered.  The plague can be cured by using the Healer ability of
cure disease provided the Healer is rank 5 or greater.  The base
chance of curing this disease is 5 $\times$ Rank + $2 \times$ victim's
EN.  Cure disease may be performed multiply times but each attempt
will take five fatigue which cannot be recovered magically.
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Lotion of Preserve Flesh}

This lotion can applied to any recently deceased body (within 24 hours)
and it will halt all natural decay indefinately.  The magic is
equivilent to a 30 MA major curse.  Should the body be destroyed the
magic will be dispelled but if the body is disected the parts will
remain preserved.
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Venom of Necrotising Flesh}

This high grade blade venom will cause any flesh which comes in
contact with it to immediately become necrotic, causing an immediate
2d10 damage (with 10s rolling up).  This damage cannot be cured until
the treated by a rank 7 or greater Healer using Cure Poison.  This
venom will continue to cause damage every hour until the victim is 
cured or dead.  This venom is enough for one application can will
be transferred to the victim if a sucessful strike causing actual
damage is make.
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Scarab of Purification}

This hand sized scarab is a weird science device with can filter blood
of all toxins, poisons, diseases or curses with directly affect blood.
To treat a patient the device this primed with 1,000sp worth of
crushed gems.  It is then placed on the patient's chest and activated.
The scarab will insert two narrow (quarter inch) tubes into the
patient's ventricules.  Over the period of the next hour it will
filter the patient's blood though the gem fragments, imbueing them
with any toxins found.

This operation is not without hazard or pain and the patient must make
a 1 $\times$ EN roll or go into anaflatic shock (Mind mages or people
immuned to pain may add 5 $\times$ rank of their ability to the base
chance).  A critical failure (00 on the dice) will cause the patient
to immediately die.  Patients in shock will not recover without the
help of a rank 4 Healer or 24 hours recovery.  During the process the
patient will suffer intense pain (Mind mages et al suffer intense
discomfort) unless the Healer sooth pain is used.  Any movement by the
patient will force another shock check so it is recommended the
patient is either sedated or confined.

The gem fragments can then be processed by an alchemist with a 10
$\times$ rank chance of extracting and seperating any toxins with the
gems.  The gems will be destroyed regardless.  Should the alchemist
make a critial failure of (90 + rank)\% or higher then their will
become infected with whatever toxins are in the gems.
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Dagger of Death Storage}

This dagger can capture the moment of an entity's death.  If
a ruby of value of atleast 1,000 silver pennies is inserted
into a pommel recepticle and a victim is then killed using
the dagger then their ``death buzz'' is drawn into the ruby.

The victim will suffer a cumulative 15\% penalty to their resurrection
chance when killed by the dagger so for example that if they are
killed four times then they will have a -60\% on the chance of been
resurrected afterward (this does not affect other resurrection
chances).

The gem can either be crushed, releasing the death, or disolved in
alcohol to form a potion.  The potion can then be drunk and the
imbiber will then be as if their were Death Aspected and the victim
had just died (effect is self only).

However a Life Aspected person will go immediately must make a 1
$\times$ EN roll or go into death shock for 2d10 pulses, reducing
their fatigue to zero.  A roll of (70 + EN) or greater will reduce
their endurance to -1.  The fatigue damage cannot be cured and the
endurance damage can only be cured by the Healer Cure Endurance
ability (or similar affect).

\textbf{Value: 15000 sp}
\begin{quote}
\emph{Ross Alexander (evilross@yahoo.co.uk) 29/1/2004}
\end{quote}
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Immortal bane sword}

This single edge sword (treat as scimitar, tulwar, sabre or similar
for melee) is a pure spectral construct with a material hilt bound to
it.  It is powered from the death of sentients and will evaporate if
not used to slay a victim every month.  People slain with this weapon
have -30\% to their resurrection chance and immortals slain must make
a magic resistance check verses Necromancy or have their lifespan cut
to 99 years, becoming short lived sentients (if they are resurrected).
The sword itself is a product of a rank 25 Spectral Weapon spell with
a base strike chance of 80\% and damage of [D + 12].

Creatures slain with the sword will not cause a ``death buzz'' as the
sword feeds on it instead.  The sword is virtually indestructable
against material damage but has only 20 points of life against Holy or
Divine damage.  In addition magic which would destroy a spector will
destroy it (it has a flat 80\% magic resistance).

\textbf{Value: 20000 sp}
\begin{quote}
\emph{Ross Alexander (evilross@yahoo.co.uk) 29/1/2004}
\end{quote}
\end{ep-box}

%% ----------------------------------------------------------------------


\begin{ep-box}{Scarab of Protection verses Poison}

This inch long scarab can be attached to a living creature and protect
the creature will be protected against poison.  While wearing the
scarab the wearer will have their maximum fatigue reduced by two
points.  Any poison which would have affected the wearer will be
absorbed by scarab, destroying it.
\end{ep-box}


\begin{ep-box}{Scarab of Protection verses Disease}

This inch long scarab can be attached to a living creature and protect
the creature will be protected against disease.  While wearing the
scarab the wearer will have their maximum fatigue reduced by two
points.  Any disease which would have affected the wearer will be
absorbed by scarab, destroying it.
\end{ep-box}

\begin{ep-box}{Scarab of Protection verses Draining}

This inch long scarab can be attached to a living creature and protect
the creature will be protected against draining.  While wearing the
scarab the wearer will have their maximum fatigue reduced by one
point.  Any draining which would have affected the wearer will be
absorbed by scarab, destroying it.
\end{ep-box}

\begin{ep-box}{Powder of Stopping Bleeding}

This small (1oz) bag of powder will immediate stop any bleeding if its
contents are poured onto a wound, but will cause [D-5] burns damage as
a result.
\end{ep-box}

\begin{ep-box}{Potion of Binding Damage}

This potion will cause a victim to be used to absorb damage for
somebody else.  The perpetrator must activate the potion by placing a
drop of their blood into the potion.  The potion must be used within
24 hours of activation.  The victim must imbibe the potion (or have it
forcably injected into them).  The victim is then bound to the
perpetrator, who may shift any damage (including specific grevious
wounds) onto the victim.  In addition the victim can be drained to
provide fatigue for spell casting.  Every point of fatigue spent on
the spell will require two points of fatigue (or endurance if the
victim has no fatigue) from the victim.

The perpetrator may drop the binding by casting a necro special
counterspell on themselves.  The victim must have an 30 MA major curse
removed to be freed of the binding.  The binding requires the victims
life force to operate and will drain a point of fatigue (reducing
their maximum fatigue) every day until they have no fatigue.  It will
then drain their endurance until the victim is reduced to negative
half their endurance, when their will die.  If the perpetrator is a
necromancer or undead then the dead victim will become a revanant
under their control.
\end{ep-box}

% ----------------------------------------------------------------------

\begin{ep-box}{Ring of Enchant Aura}

This ring contains 10 charges (95\%) of Enchant Aura.
\end{ep-box}

% ----------------------------------------------------------------------

\begin{ep-box}{Rod of Enchantments}

This powerful rod can enchant spells and items in a number of ways.
\begin{itemize}

\item
Any item or entity can be enchanted so that the next spell cast apon
them will stack, taking the maximum of any duplicate effects.  For
example with this a person could have both Armour of Earth and
Waithcloak on at the same item.

\item
Can be used to enhance a single charge in an invested item with 10
Ranks of any aspect.

\item
Can be used in conjunction with the Investment ritual to enchance the
any aspect up to 10 Ranks (up to a maximum of Rank 25) when creating
an invested item.  All changes will be affected.

\end{itemize}

Once all charges have been expended the rod remains magical and can be
used as a mace (strike chance 55\%, [D + 8] damage).  In addition it
can hold up to four (possibly different) charges of any E\&E spell.
Each charge must be triggered individually (unless the rod is broken,
in which case they are all triggered simultaneously).  Each charge
must be stored in the rod as per the investment ritual but does
require any material component (i.e.\ no cost is involved but the
required time must be spent).  The rod currently contains 15 charges.
\end{ep-box}

\begin{ep-box}{Wand of Necrosis}

This heavy malachite rod weights 2lb and is 2ft in length.  It has
invested in it 100 charges of Necrosis (each charge does Rank 10
effective damage). It can be tiggered on trigger on up to eight
targets, each of whom can be a single hex entity or 50 foot radius
area.  If the spell is targetted on an area it will affect any living
plants in that area.  All damage is resist for half.

Up to two charges can be spent per target causing damage of $[D + 5] +
20$ per charge.  A double effect can either double the number of
possible targets (which can double stack), increase the damage to $[D
+ 5] + 30$ per charge or reduce the target's Magic Resistance by 20.
A triple effect will allow the triggered to apply two different double
effects.

The rod has a base chance of 115\% with a maximum range of 200 feet.
Should the rod be broken (having more than 15 points of subjected on
it) all renaming charges will be expended immediately causing $$(([D +
5] + 20 \times \mbox{remaining charges})) / 50$$ points of damage to
every living creature (including plants) within a 200 foot radius.
\end{ep-box}

\end{document}
