%% 2025-04-05

\documentclass{report}

\RequirePackage{utopia-magic}

\begin{document}

\chapter{Introduction}

\section{Changes to core rules}

\begin{enumerate}
\item Initiative
\item Quickness \& Slowness
\item Damage \& FT
\end{enumerate}

\chapter{Utopia III}

\section{The Five Duchies}

\subsection{Fenwich}

The Duchy of A Hundred Lakes covers a wide area of lakes surrounded by
rich countryside.  Many of the lakes are connected by canals.

\subsection{Aquaterra}

Occupying the coastal region east of the Shifting Mire is consists of
many small villages scattered along the coast, each with their minor
hinterland.

\subsection{Aborea}

\begin{College}{Cardinals}

\section{Restrictions}

This magic is restricted to members of the Secretariat and can
be cast anywhere.

\section{Talents}

\begin{talent}[T-1]{Concealed Casting}
\multiple{150}
\begin{effects}
This Talent allows the Adept to cast without any apparent movement or
noise.  The Adept may not speak or make meaningful or extravagant
actions. They may not cast without a hand free or if mute.  If an
observer is intent on the Adept (e.g. in conversation with), the
observer may make a (2 × PC) - Rank check to notice. All aspects of
the Spell (including Base Chance) must be cast at the lower of the
Talent’s and Spell’s ranks, and the Base Chance is further reduced by
20\%. The Talent only functions with spells of the College of
Illusions.
\end{effects}
\end{talent}

\begin{talent}[T-2]{Detect Enchantment}

\range{30 feet (+ 5 / Rank)}
\multiple{50}
\basechance{PC + 3\% / Rank}
\begin{effects}
This talent determines whether an item, person, or area in
line-of-sight and within range is currently under an enchantment or
magical effect.  The Adept can tell whether the enchantment is a
current spell or ritual, a warded, invested, or permanent effect, is
contained in a trap or is a curse.  A double effect will reveal the
general nature of the spell (eg. defensive, summoning, damaging) and
the effective Rank or remaining duration of the enchantment.  In
addition, a triple effect will tell the Adept the exact name of the
magic (eg. Wall of Bones, Hellfire), or the College of the spell.

Once the initial, most recent, magical effect has been successfully
detected, older enchantments on a target with multiple layers of magic
may also be detected.  This may be continued while the Adept continues
to succeed in detecting Enchantments.  However, only one attempt per
quarter may be made to detect any given enchantment.  If a new
enchantment occurs, the Adept may attempt to detect it, though this in
no way affects the status of the old layer.  If an old, unsuccessfully
detected enchantment expires, the Adept may attempt to detect any
newly revealed magic beneath it.

If the Adept is in contact with the target then the base chance of
this talent is improved to PC + 5\% / Rank.
\end{effects}
\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Evil Eye}

\range{Self}
\duration{5 minutes + 5 / Rank}
\multiple{300}
\basechance{30\%}
\resist{None}
\storage{Potion, Magical Trap, Investment}
\target{Entity }
\begin{effects}
When the adept casts Evil Eye on themself, a third eye appears in
their forehead, which is normally invisible but is able to be seen by
Witchsight and similar effects.  When the adept is under the effect of
the Evil Eye all spells they cast on others are resisted at a penalty
of Rank unless the target is wearing an Amulet of Elder Flowers.
\end{effects}
\end{spell}


\begin{spell}[G-2]{Telekinesis}
\range{15 feet + 15 / Rank}
\duration{10 seconds + 10 / Rank}
\multiple{300}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward}
\target{Entity or object}
\begin{effects}
Allows the Adept to lift a target of weight up to 2 pounds and move it
at the rate of 2 TMR.
\end{effects}
\end{spell}

\begin{spell}[G-3]{ESP}
\range{30 feet + 15 / Rank}
\duration{30 seconds + 10 / Rank}
\multiple{100}
\basechance{40\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept can sense the presence and general mood (but not the exact
nature) of all entities within range of the spell.  The Adept has a
general idea of where each entity is, say to within 90 or 60 degrees,
the accuracy of which may increase with rank.  ESP will detect the
presence of an entity even if that entity is mind cloaked. Spells
cannot be targeted through ESP.
\end{effects}
\end{spell}
\end{College}


\begin{College}{Elements}

The world consists of various elements which together form the
physical world.  In some instances their negative counterparts
also exist.

\begin{uttblr}{colspec={XX}}
Element		& Negative \\
Air		& Vacuum \\
Earth		& \\
Fire		& Cold \\
Water		& \\
Light		& Dark \\
\end{uttblr}

Any two elements can be combined although not all combinations are
possible.

\begin{uttblr}{colspec={XXX}}
Air		& Earth		& Dust \\
Air		& Water		& Mist \\
Air		& Fire		& Lightning \\
Earth		& Water		& Silt \\
Earth		& Fire		& Magma \\
Water		& Fire		& Steam \\
Water		& Cold		& Ice \\
\end{uttblr}

\subsection{Elements}

\begin{description}
\item[Air]
\item[Earth]
\item[Fire]
\item[Water]
\item[Dust]
\item[Mist]
\item[Lightning]
\item[Silt]
\item[Magma]
\item[Steam]
\item[ice]
\end{description}

\subsection{Spells}

\begin{itemize}
\item Bolt
\item Ball
\item Wall
\item Shield
\item Shape
\item Transform
\end{itemize}

\begin{spell}[G-1]{Bolt of Elements}
\range{30 feet + 15 / Rank}
\duration{Instance}
\multiple{100}
\basechance{40\%}
\resist{Passive,Active}
\target{Entity, Object, Area}
\elements{Lightning, Magma, Steam, Ice, Fire, Water, Earth}
\begin{effects}
The Adept causes a bolt of Elements to streak from their hand towards
anywhere in range.  The first entity or object the bolt hits in its
path must resist or suffer D (+ 1 / Rank) damage.  If the bolt does
not hit anything it will dissipate at the end of its range.
\end{effects}
\end{spell}

\begin{spell}[S-4]{Fireball}

\range{60 feet + 10 / Rank}
\duration{Immediate}
\multiple{550}
\basechance{25\%}
\resist{Active, Passive}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
The Adept conjures a ball of Element 1 foot in diameter which rushes
to a point in line of sight of the Adept, and explodes in a radius of
5 foot (+ 5 / 5 or fraction Ranks). Everything in this radius must
resist or suffer D10 (+ 1 / Rank) damage.  If an entity resists the
damage is halved, unless they are more than half immersed in water
(e.g. swimming) where they take no damage.  The spell may set
flammable items afire when it bursts.  Hard cover (e.g. walls,
parapets, but not shields) reduces damage to half before resistance.
In some instances, e.g. a character wading through water, resistance
may negate all damage. The fireball is analogous to an instantaneous
flash of heat and there are no over-pressure or explosive effects.  A
fireball may be detonated prematurely by the imposition of barriers in
its line of flight.

At Rank 20 this spell may be cast with a detonation delay of up to 10
pulses.  This extends the casting time 4 pulses.  A delayed blast
fireball manifests itself as a 1 foot wide floating sphere of fire
until detonation occurs. If a second fireball is cast within the
proposed volume of effect of the first, both are detonated.
\end{effects}
\end{spell}



\end{College}

\begin{College}{Atavists}
\end{College}

\chapter{Skills}


\end{document}
