\documentclass[a4paper]{article}

\usepackage{dq}

\begin{document}

\begin{multicols*}{3}

  %% ----------------------------------------------------------------------

\section{Halfling Racial Modifications}

All halflings start with the following abilities, talents and stat
modifiers.  These cost 45,000 EP which must be paid back at a
rate of 10\% of earned experience.
  
\begin{Enumerate}
\item
A halfling has infravision, which allows them to see faint red shapes
where living beings are located in the dark. Their range of vision is
100 feet.

\item
A halfling adds 20\% to their chance to perform any activity requiring
stealth.

\item
If a halfling takes the thief skill, they pay half the EP cost to
advance ranks.

\item
A halfling may drop jewellery down active volcanos without anyone
thinking the worse of them.
\end{Enumerate}

\begin{tabularx}{\linewidth}{Xc}
\textbf{Characteristic} & \textbf{Modifier} \\
Physical Strength	& - 3 \\
Manual Dexterity	& + 3 \\
Agility			& + 1 \\
Endurance		& - 2 \\
Magical Aptitude	& - 1 \\
Willpower		& + 1 \\
Physical Beauty		& - 1 \\
\end{tabularx}

Once the initial EP has been paid down any new experience can be spend
on the following abilities.

\begin{Enumerate}
\item
Infravision can be extended by 10 feet / Rank (100 EM) to a maximum rank of 20.

\item
Stealth +2\% / Rank (100 EM) to a maximum of rank 20.
\end{Enumerate}


%% ----------------------------------------------------------------------

\section{Orc Racial Modifications}

All halflings start with the following abilities, talents and stat
modifiers.  These cost 45,000 EP which must be paid back at a
rate of 10\% of earned experience.
  
\begin{Enumerate}
\item
An orc's eyes are highly light-sensitive. They must decrease their
chance of hitting a target with Ranged Combat by 10\% in daylight.

\item
An orc has infravision, which allows them to see faint red shapes
where living beings are located in the dark. Their range of vision is
150 feet.

\item
Orcs are either back-stabbing scum or brutal bully-boys. An orc may
take one of either Assassin Skill or Warrior Skill and pay
three-quarters the EP to advance in Ranks.

\item
An orc's seed is highly fertile. The orc and hybrid orc population
increase mitigates against the high orc fatality rate.
\end{Enumerate}

\begin{tabularx}{\linewidth}{Xc}
\textbf{Characteristic} & \textbf{Modifier} \\
Physical Strength	& + 2 \\
Endurance		& + 1 \\
Magical Aptitude	& - 2 \\
Willpower		& - 2 \\
Fatigue			& + 2 \\
Physical Beauty		& - 4 \\
Natural Armour		& + 1 \\
\end{tabularx}

Once the initial EP has been paid down any new experience can be spend
on the following abilities.

\begin{Enumerate}
\item
  Infravision can be extended by 10 feet / Rank (100 EM) to a maximum rank of 20.

\item
  Natural Armour can be increased by 1 point at a loss of 1 PB (5000 EP, maximum 5 NA).

\item
  Both Assassin and Warrior cost three-quarters of the EP to advance.
\end{Enumerate}


%% ----------------------------------------------------------------------

\section{Shape Changer Racial Modifications}

All shapechangers start with the following abilities, talents and stat
modifiers.  These cost 45,000 EP which must be paid back at a
rate of 30\% of earned experience.

\subsection{Talent of Shape Changing}

\begin{talent}{Shape Changing}
  \multiple{500}
  \casttime{5 mins - 15 seconds / Rank}
\begin{effects}
  This Talent will transform the shapechanger from their human to
  animal form (or back).  Once used it cannot be used again for 1 hour
  - 5 minutes / Rank.  They may chose of increase their size by 5\% /
  Rank (with PS \& EN increasing 1 / 2 Ranks).  This talent will also
  increase the rank of their attacks by Rank / 2.
\end{effects}
\end{talent}


\subsection{Special Abilities}

\begin{Enumerate}
\item The shape changer gains the Talent of Shape Changing at Rank 0.
  
\item
  A shape changer possesses a dual nature. While in animal form they
  cannot cast or use magic and will have an MA of 0.
  
\item
A shape changer cannot be harmed while in animal form, unless struck
by a silvered weapon, magic or by a being with a Physical Strength
greater than 25. Five Damage Points are automatically absorbed in the
latter case.

\item
A shape changer will regenerate 1 Endurance Point every 60 seconds
while in animal form.

\item
The player must devise a set of characteristics for their animal form.
This will be same total as their human form.

\item
A shape changer's magic resistance is increased by 5\%.

\item
If a shape changer takes the courtier skill they pay three-quarters
the Experience Points necessary to advance ranks.

\item
  In animal form claws and teeth can be the target of weapon spells
  and their hide the target of any armour spells.
  
\end{Enumerate}

A new set of characteristics must be generated (see Ability 5).

\begin{tabularx}{\linewidth}{Xc}
\textbf{Characteristic} & \textbf{Modifier} \\
Physical Beauty		& + 1 \\
\end{tabularx}

Once the initial EP has been paid down any new experience can be spend
on ranking the Shapechange talent.

%% ----------------------------------------------------------------------
  
\section{Elf Racial Modifications}

All elves start with the following abilities, talents and stat
modifiers.  This talent cost 45,000 EP which must be paid back at a
rate of 15\% of earned experience.
				   
\begin{enumerate}
\item
Elves have superior vision especially over long distances or in poor
lighting.  An elf can see in the dark as a human does on a cloudy
day. Their effective range of vision in the dark is 150 feet under the
open sky, and half this elsewhere.

\item
If an elf character is a ranger specialising in woods, they pay
one-half the EP to advance ranks.

\item
\begin{talent}{Witchsight}
\index{elf!witchcraft}
\multiple{200}
\begin{effects}
The Adept may see objects or entities which are invisible and they
appear to have a slight blue sheen around them. If the invisibility
effect (excluding Walking Unseen) is of a higher Rank than the
Witchsight, the object or entity may not be clearly identified or
directly magically targeted.
\end{effects}
\end{talent}

\item
An elf makes little or no noise while walking and adds 10\% to their
chance to perform any activity requiring stealth.

\item
If an elf character takes the healer skill, the elf pays
three-quarters the EP to advance ranks, though they cannot resurrect
the dead.

\item
An elf is impervious to the special abilities of the lesser undead.

\item
If an elf character takes the courtier skill, the elf pays one-half
the EP to advance ranks.
\end{enumerate}

\begin{tabularx}{\linewidth}{Xc}
\textbf{Characteristic} & \textbf{Modifier} \\
Physical Strength	& - 1 \\
Agility			& + 1 \\
Endurance		& - 1 \\
Magical Aptitude	& + 1 \\
Willpower		& + 1 \\
Perception		& + 1 \\
Physical Beauty		& + 2 \\
Tactical Movement Rate	& + 1 \\
\end{tabularx}

Once the initial EP has been paid down any new experience can be spend
on the following abilities.

\begin{enumerate}
\item
Night vision can be extended by 10 feet / Rank (100 EM)

\item
Stealth +2\% / Rank (100 EM)

\item
\begin{talent}{Resist Undead}
\multiple{200}
\begin{effects}
\begin{Description}

\item[Fear Effects] The Adept gains a 5\% (+3 / Rank) bonus to any
Willpower check to resist fear effects caused by Undead.

\item[Infection] The Adept reduces their chance of becoming infected
by 5\% (+ 1 / Rank).

\item[Undead Draining] Damage done to the Adept due to the touch of an
Undead is reduced by 1 (+ 1 per 4 or fraction Ranks).

\item[Resistance] Gains an additional bonus of 5\% (+ 1 / Rank) to
resist any magical effects or spells cast by undead.

\end{Description}
\end{effects}
\end{talent}

\item
\begin{talent}{Resist Temperature}

\multiple{200}
\begin{effects}
The Adept is most immune to climatic extremes of temperature from
arctic cold to desert heat whether generated naturally or by magic.
They apparent temperature will be 2\textdegree{} / Rank less extreme,
therefore depending on rank they are less likely to suffer exposure,
frostbite, wind-chill, heat stress, hypothermia or hyperthermia, and
their fatigue loss is unaffected by temperature.  If the Adept is
attacked by fire, ice, heat or cold (whether magical or not) then the
damage points are reduced by 1 (+ 1 / 5 Ranks).
\end{effects}
\end{talent}

\item Bow specialisation (can rank short bow, long bow or composite
bow to Rank 10).  Cost 2000EP for Rank 9 and 4000EP for Rank 10.

\item All healer abilities take 2\% / Rank less time (200 EM).

\item If ranger then gains woodland specialisation automatically and
can choose additional specialisation (1000 EP).

\item Courtier abilities improved by one effective rank.  Must be
purchased for each category at 500 EP.

\item Base stat modifier +1 MA (7500 EP).

\item Base stat modifier +1 WP (7500 EP).

\item Base stat modifier +1 AG (7500 EP).

\end{enumerate}

\end{multicols*}
\end{document}
