\documentclass[a4paper]{article}

\usepackage{dq}

\begin{document}

\begin{multicols*}{3}

Updated 3/2/2011 to clarify Elf super racial stat increased can only
be applied once.

\section{Elf Racial Talent}

All elves start with the following abilities, talents and stat
modifiers.  This talent cost 45,000 EP which must be paid back at a
rate of 15\% of earned experience.
				   
\begin{enumerate}
\item
Elves have superior vision especially over long distances or in poor
lighting.  An elf can see in the dark as a human does on a cloudy
day. Their effective range of vision in the dark is 150 feet under the
open sky, and half this elsewhere.

\item
If an elf character is a ranger specialising in woods, they pay
one-half the EP to advance ranks.

\item
\begin{talent}{Witchsight}
\index{elf!witchcraft}
\multiple{200}
\begin{effects}
The Adept may see objects or entities which are invisible and they
appear to have a slight blue sheen around them. If the invisibility
effect (excluding Walking Unseen) is of a higher Rank than the
Witchsight, the object or entity may not be clearly identified or
directly magically targeted.
\end{effects}
\end{talent}

\item
An elf makes little or no noise while walking and adds 10\% to their
chance to perform any activity requiring stealth.

\item
If an elf character takes the healer skill, the elf pays
three-quarters the EP to advance ranks, though they cannot resurrect
the dead.

\item
An elf is impervious to the special abilities of the lesser undead.

\item
If an elf character takes the courtier skill, the elf pays one-half
the EP to advance ranks.
\end{enumerate}

\begin{tabularx}{\linewidth}{Xc}
\textbf{Characteristic} & \textbf{Modifier} \\
Physical Strength	& - 1 \\
Agility			& + 1 \\
Endurance		& - 1 \\
Magical Aptitude	& + 1 \\
Willpower		& + 1 \\
Perception		& + 1 \\
Physical Beauty		& + 2 \\
Tactical Movement Rate	& + 1 \\
\end{tabularx}

Once the initial EP has been paid down any new experience can be spend
on the following abilities.

\begin{enumerate}
\item
Night vision can be extended by 10 feet / Rank (100 EM)

\item
Stealth +2\% / Rank (100 EM)

\item
\begin{talent}{Resist Undead}
\multiple{200}
\begin{effects}
\begin{Description}

\item[Fear Effects] The Adept gains a 5\% (+3 / Rank) bonus to any
Willpower check to resist fear effects caused by Undead.

\item[Infection] The Adept reduces their chance of becoming infected
by 5\% (+ 1 / Rank).

\item[Undead Draining] Damage done to the Adept due to the touch of an
Undead is reduced by 1 (+ 1 per 4 or fraction Ranks).

\item[Resistance] Gains an additional bonus of 5\% (+ 1 / Rank) to
resist any magical effects or spells cast by undead.

\end{Description}
\end{effects}
\end{talent}

\item
\begin{talent}{Resist Temperature}

\multiple{200}
\begin{effects}
The Adept is most immune to climatic extremes of temperature from
arctic cold to desert heat whether generated naturally or by magic.
They apparent temperature will be 2\textdegree{} / Rank less extreme,
therefore depending on rank they are less likely to suffer exposure,
frostbite, wind-chill, heat stress, hypothermia or hyperthermia, and
their fatigue loss is unaffected by temperature.  If the Adept is
attacked by fire, ice, heat or cold (whether magical or not) then the
damage points are reduced by 1 (+ 1 / 5 Ranks).
\end{effects}
\end{talent}

\item Bow specialisation (can rank short bow, long bow or composite
bow to Rank 10).  Cost 2000EP for Rank 9 and 4000EP for Rank 10.

\item All healer abilities take 2\% / Rank less time (200 EM).

\item If ranger then gains woodland specialisation automatically and
can choose additional specialisation (1000 EP).

\item Courtier abilities improved by one effective rank.  Must be
purchased for each category at 500 EP.

\item One off base stat modifier +1 MA (7500 EP).

\item One off base stat modifier +1 WP (7500 EP).

\item One off base stat modifier +1 AG (7500 EP).

\end{enumerate}

\newpage

\section{Valkyrie Racial Talent}

Valkyries are the heralds and messengers of Odin.  Anybody becoming a
valkyrie changes race and generic true name.  Only females may be
valkyries.

All valkyries start with the following abilities, talents and stat
modifiers.  This talent costs 25,000 EP which must be paid back at a
rate of 15\% of earned experience.

\begin{enumerate}
\item
\begin{tabularx}{\linewidth}{Xc}
\textbf{Characteristic} & \textbf{Modifier} \\
Physical Strength	& + 2 \\
Endurance		& + 2 \\
\end{tabularx}

\item Herald of Odin. As a herald they can project their voice so that
anybody within 300 feet can clearly hear them.  In addition they gain
the standard of Odin.  While holding the standard with both hands any
entity must make a once WP to attack them, and to do so may call down
the wrath of Odin upon the assailant.

\item Call the Slain.  Should the valkyrie see a warrior fall in
battle then if she deems their actions worthy then she may call their
spirit to join the chosen of Odin for Ragnarok.  Any spirit called may
refuse such a call.

\item Heroic Return.  Should the valkyrie fall in battle her spirit
may call her sisters to ride forth and claim her corpse.  In such a
case a dozen valkyrie will appear riding down a rainbow bridge on
flying horses and claim the corpse, taking it back to Asgard.  They
will never interfere in the battle but will defend the corpse and use
all possible means to claim it.

\item Pegasus Mount.  A valkyrie with gain a heavy war pegasus as a
mount which may be summoned and banished to Asgard at will.

\item Warrior of Odin.  A Norse Warrior does not care if their
opponent is down or indisposed since to slay or be slain is all that
counts so warrior skill will count in all situations, period.

\item Hunter of Odin.  All great Norse Warriors are also great
Hunters.  The valkyrie gains all their warrior abilities with any
weapon or device a hunter would use (basically anything except sap).

\end{enumerate}

Once the initial EP has been paid down any new experience can be
spent on the following abilities.

\begin{enumerate}
\item
\begin{spell}{Ride the Lightning}
\multiple{200}
\range{10 feet + 10 feet / Rank}
\begin{effects}
This special knowledge non college spells allows the valkyrie mount
their pegasus as standard action provided their mount is within range.
They move within the pulse as bolt of lightning, so their must be a
path the lightning to travel for this ability to work.
\end{effects}
\end{spell}

\item 
\begin{talent}{Aurora Armour}
\multiple{300}
\begin{effects}
The valkyrie can chose to cloak themselves in an Aurora of brightly
simmering light.  This Aurora will reduce all damage (magical or
otherwise) by 1 point / 4 Ranks.  This will cancel any Shadow or
Illusion magic on the valkyrie and counts as Rank magical light in the
surrounding megahex (except does not have special Rank 20 light
effects).
\end{effects}
\end{talent}

\item
\begin{spell}{The Ravens of Odin}
\range{1 mile + 1 / Rank}
\duration{5 minutes + 1 / Rank}
\multiple{100}
\begin{effects}
A valkyrie can summon two of Odin's ravens to command.  These are
under the control of the summoner and will do the summoner's bidding.
The ravens are natural creatures with 2 EN and 3 FT.  The summoner can
control and scry with them as long as they are in range.  Summoning
them takes 10 minutes and costs 5 FT.  In addition the summoner can
temporarily transfer up to 4 EN and 6 FT to each raven (reducing their
own respectively).  The ravens have a TMR of 10, and a flight speed of
20 + Rank miles per hour.

At rank 20 the duration and range double.
\end{effects}
\end{spell}

\item
\begin{spell}{Storm Calling}
\range{1 mile + 1 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{200}
\basechance{40\%}
\resist{None}
\target{Special}
\begin{effects}
The Adept may summon a storm front to them which will travel with the
Adept within the range of the spell.  The storm will create a downpour
(snow, rain, hail, sleet, or whatever else that the GM feels the
clouds may contain). Generally a storm front can be seen for 20 to 30
miles.  The storm front will take D10 minutes to arrive.  Once the
duration has expired, the weather will gradually return to normal over
a similar amount of time.
\end{effects}
\end{spell}

\item
\begin{ritual}{Summon Rainbow Bridge}
\duration{1 minute + 1 / Rank}
\multiple{200}
\basechance{20\% + 4 / Rank}
\resist{None}
\target{Special}
...
\end{ritual}
\end{enumerate}

\newpage

\section{Knight of the Golden Order}

The Golden Order is a secular order originally sponsered by the
ancient golden dragon Orthosopheles.  It was formed to confront the
armies of the evil dragons and their wormkin servants.  A few of the
wormkin could fly but most where infantry, with the a number of them
having riders.

The wormkins' strong resistant to magic and the extreme scarity of
talented Adepts lead the Order to focus on elite officer corp which
would be in the vanguard of the defending armies.

To counteract the flying dragonkin it was necessary to have the
vanguard mounted, with the most experienced members able to fly with
their mounts.  The ability to rapidly respond to dragonkin incursions
and not have their mounts killed lead to development of super heavy
warhorses, able to stand up in melee against all but the greater
dragonkin.

After the Order's original fortress was destroyed the Order realised
that any become a focus for aggression, both internally from political
enemies and externally from dragon host raids.  With many of its
members spending much of their time holding the line the focus turned
to improving the defense of frontier keeps.

The war ended in the defeat of the chromatic dragons and their kin but
continuing hostility to all their kind lead Orthosopheles and a number
of his kin to go into exile, leaving the Order to be disolved into the
mainstream political structure of the time.

Recent events have lead Orphostopheles to revive the Order and invited
a number of worthy candidates to join.  Members of the order are
expected to perform the following actions.
\begin{enumerate}
\item Always be in the vanguard of any action against invading forces.

\item To resist tyranny and oppression and champion justice.

\item To not involve the Order in any political or religious dispute.

\end{enumerate}

The Order's training focused on four seperate area of expertise:
Attack, Defense, Mount and Keep.  Abilities costs are expressed in
Points, with non rankable abilities costing one Point and rankable
abilties Rank Points.  Points cost 1000 EP but may be given as reward
for completing tasks assigned by the Order.

\subsection{Attack}

A knight of the Order designates a single weapon as their weapon of
preference.  This requires them to train continuously for one week
with that weapon.  The weapon itself may be any weapon but the Knight
must have maximum Rank in it.


\begin{description}

\item[Attune Weapon] The Knight attunes a specific weapon to their
Order Abilties.  This takes one week of training and only one weapon
can be attuned at a time.  The attuned weapon can always be prepared
as a free action.

\item[Weapon Focus] The Knight gains 4\% / Rank Strike Chance (maximum
5 Ranks).

\item[Improved Rank] The Knight effective rank with the weapon of
maximum + 1 (requires Rank 5 Weapon Focus).

\item[Instant Response] The Knight has a (IV - 50 + 10 / Rank)\%
chance to simultanously retaliating against a surprise attack (maximum
5 Ranks).  This can be done with any available weapon.

\item[Preemptive Retaliation] The Knight has a IV\% chance of
retaliating to a surprise attack before attacker actually strikes
(requires Rank 5 Instant Response).  This can be done with any
available weapon.

\item[Power Attack] The Knight can overstrength their weapon by up to
Rank points of damage without penalty (maximum Rank 5).

\item[Weapon Throw] The Knight may throw their weapon up to 10 feet /
Rank as a melee attack (maximum Rank 5).

\item[Close Attack] The Knight can fight in close with their weapon at
a reduce strike chance of [SC - 100 + 20 / Rank] (maximum Rank 5).

\item[Ambidexterous] The Knight can use their weapon equally with
either hand (requires 10 Points in Attack).

\item[Dying Strike] The Knight has an IV\% chance of retaliating to a
blow which kills them (requires 15 Points in Attack).

\item[Melee Freedom] The Knight may bypass a threatening opponent but
may suffer a strike against them.

\item[Extended Movement] The Knight may move up to Rank hexes in melee
at still make a full attack (maximum Rank 5, 12 pulse cooldown,
requires Melee Freedom).

\item[Critical Strike] The Knight can strike with critical accuracy
[1\% / Rank off dice roll] (maximum Rank 5, requires 25 Points in
Attack, 30 minute cooldown).

\end{description}

\subsection{Defense}

A knight of the Order designates a single set of armour and a shield
as their armour of preference.  This requires them to wear
continuously for one week the armour and train with the shield.

\begin{description}

\item[Attune Armour] The Knight must attune a set of armour and shield
to use their Order Abilities.  This will take a week of training and
only one set of armour and one shield can be attuned at a time.  The
attuned shielded can always be readied as a free action.

\item[Improved Armour I] The Knight's armour gains (10\% / Rank)
additional armour points (maximum Rank 5).

\item[Improved Armour II] The Knight's armour gains (10\% / Rank) of
their improved armour verses all melee damage (maximum Rank 5,
requires Rank 5 Improved Armour I).

\item[Improved Armour III] The Knight's armour will reduce all damage
by 50\% of its improved protection (requires Rank 5 Improved Armour II).

\item[Extended Fatigue] The Knight must suffer 2 / Rank EN damage
before becoming fatigued (maximum Rank 5).

\item[Extended Endurance] The Knight will not become unconscious until
their EN is below -1 / Rank (maximum Rank 5, requires Rank 5 Extended
Fatigue).

\item[Last Man Standing] The Knight will not become unconscious due to
damage until are below -EN/2.  If they are not dead at this point (eg
under Trollskin or similar) they will not gain consciousness until
they have positive endurance (requires Rank 5 Extended Endurance).

\item[Stun Resistance] The amount of damage a Knight must suffer
before being stunned is improved by +1 / Rank (maximum Rank 5,
requires 10 Points in Defense).

\item[Dodge] A Knight has [AG + 10 / Rank]\% chance of stepping out of
the way if an opponent tries to come into close with them (maximum
Rank 5).  This can be done at any time except when surprised.

\item[Instant Stand] A Knight can stand from prone as a free action
(requires Rank 5 Dodge).

\item[Quickened Equip] The Knight takes 15\% / Rank less time to put
on their armour (maximum Rank 5).

\item[Instant Armour] The Knight may put their armour in a single
pulse (requires Rank 5 Quickened Equip, 20 Points in Defense).

\item[Guard I] The Knight can chose to talk another target's blow.  If
the target is prone and the Knight is in their hex then they take Rank
(per pulse) blows otherwise directed at the target.  The blow is taken
once it is determined if it has struck.  In addition they may use an
action to Guard a target in an adjacent hex, as acts as equivilent to
an evade (maximum Rank 5).

\item[Guard II] The Knight may block a blow into an adjacent hex at
any time, losing their next action (requires Rank 5 Guard I).

\end{description}

\subsection{Mount}

A knight of the Order designates a single mount their mount of
preference.  This requires them to train with the mount for one month.
The mount must be trained if an animal.

\begin{description}

\item[Bound Mount] The Knight can communicate perfectly with their
mount up to 20 feet / Rank away (maximum Rank 10).

\item[Improved Riding] The Knight gains +1 / Rank to their riding
skill (maximum Rank 5).

\item[Instant Mount] The Knight can mount as free action provided the
mount is in an adjacent (requires Rank 5 Improved Riding).

\item[Improved Natural Armour I] The mount's natural armour improved by
+1 / Rank (maximum Rank 5).

\item[Improved Natural Armour II] The mount's natural armour counts
against all damage.

\item[Extended Charge] The mounts TMR improves by 2 TMR / Rank when
charging (maximum Rank 5, requires 10 Points in Mount).

\item[Flying Mount] A mount gains the Flying Skill at Rank 0.  This
does not automatically allow the mount to fly but can control their
flight if magically empowered (requires 5 Points in Mount).

\item[Improved Flying] The mount's Flying Skill is increased by 2 /
Rank (maximum Rank 5, requires Flying Mount).

\item[Ranger] The mount gains its rider's ranger tracking, direction
sense, detect ambush and detect traps and snares with a (50 - 10 /
Rank) penalty to base chance (maximum Rank 5, requires 15 in Mount).

\item[Dodge] The mount has an AG + 10 / Rank chance of sidestepping an
opponent trying to enter close (maximum Rank 5).

\item[Improved Dodge] If the mount has a rider it's Dodge Rank is
improved by the rider's Rank for calculating dodge chance (requires
Rank 5 Dodge).

\item[Natural Weapons] The Mount's natural weapon attacks have a
minimum weapon rank of Rank 0.

\item[Kick I] The mount's kick (or claw etc) attack Rank increases by +1
/ Rank (maximum Rank 10).

\item[Kick II] The mount can kick (or claw) with its second hoof (claw
etc) with both suffering a -20\% Strike Chance penalty (requires Rank
5 Kick I).

\item[Bite I] The mount's bite attack Rank increases by +1 / Rank
(maximum Rank 10).

\item[Bite II] The mount may bite as a secondary attack (requires Rank
5 Bite I) at -20\% SC.

\item[Trample I] The mount's trample attack Rank increases by +1 /
Rank (maximum Rank 10).

\item[Trample II] The mount can trample an opponent as a secondary
attack at -20\% SC (require Rank 5 Trample I).

\item[Endurance] The mount's overland movement rate improves by 20\% /
Rank (maximum Rank 5).

\end{description}

\subsection{Keep}

\begin{description}

\item[Improved Willpower] All within the Keep gain +1 / Rank Willpower
bonus.  This is doubled if the Knight is present (maximum Rank 10).

\item[Improved Fortitude] The Knight gains +10\% / Rank increase in FT
\& EN while in the Keep (maximum Rank 10).

\item[Antiscry] The Knight gains their MR + 5\% / Rank chance of been
unscryable while in their Keep [rolled when ever somebody tries to
scry them] (maximum Rank 5, requires 5 Points in Keep).

\item[Home Ground I] If the Knight is magically transported anywhere
with a random destination and the Keep a possible destination then
there is a 20\% / Rank chance they will appear in the Keep (maximum
Rank 5).

\item[Home Ground II] If the Knight is magically transported against
their will they have a WP + 10\% / Rank chance of return to their Keep
instead, if they will it (maximum Rank 10, require Rank 5 Home Ground
I).

\item[Beyond Life] If the Knight is slain within the Keep their spirit
is beyond all Necromancy or Witchcraft.  In addition they gain +5\/% /
Rank improvement to their chance of ressurection (maximum Rank 5,
requires 15 Points in Keep).

\end{description}

\subsection{Affliation}

The Order has a number of levels of affiliation.  Each level gains
additional benefits but requires additional commitment.

\begin{description}

\item[Fraternal] The Order recognises the person as worthy and will
provide simple assistance if requested, such as personal lodging for a
short period, provisions for themselves and direct staff etc.
Fraternal members are general people who are possibly sworn to another
order or patron and are not expected to become members of the Order.

\item[Initiate] Initiate members can gain lodging and board at any
Order lodge or Keep for extended periods of time.  They are expected
to assist the local commander if requested.  They gain the Amulet of
the Order.  This amulet gives the wearer +5\%, Magic Resistance, +5\%
Defense, +5\% to all Strike Chances, an additional 1 $\times$ WP to
fear checks and -10 on all fear table rolls.  It does not stack with
amulets of Aquilegius, Diamonds or Luck (or equivalent).  An Initiate
must have 10 Points of Order Abilities.  This item is attuned to
the Initiate and will not work for anybody else.  It will also crumble
to dust if not worn by the Initiate for over a week.

\item[Soldier] Soldiers must tithe 5\% of their income to the Order
and spend atleast one week a season aiding the Order and have spent
atleast 25 Points on abilities.  Soldiers gain the Tabbard of the
Order.  This tabbard is magical and is impervious to dirt and soiling.
It provides 2 points of armour against all damage and has the
following spell like abilities (each can be used once per 24 hours).
The spell like abilities require a Trigger Action to use.

\begin{description}
\item[Light] This spell causes the cloak to glow like a torch,
shedding bright light in a 20-foot radius (and dim light for an
additional 20 feet) from the point you touch. The effect is immobile,
but it can be cast on a movable object. Light taken into an area of
magical darkness does not function.  The duration is 30 minutes.

\item[Cure Moderate Wounds]
Cures 2d10 + 10 damage to wearer.

\item[Bear's Endurance]
The wearer gains +4 EN bonus for 5 minutes.
\end{description}

In addition the tabbard will preserve the body of the wearer if the
die while wearing it and it is not removed.  This item is attuned to
the Initiate and will not work for anybody else.  It will also crumble
to dust if not worn by the Initiate for over a week.

\item[Knight] Knights must tithe 10\% of their income to the Order and
spend atleast two weeks a season aiding the Order and have spent
atleast 50 Points on abilities.  Knights gain the Barding of the Order.
\begin{enumerate}
\item Weights 60 lbs.
\item Has 14 points of normal protection and 7 points against
endurance, specific grevious and magical damage.
\item 30\% Magic resistance and can be the target of one counterspell.
\item The mount can ride over water (but not mist etc) and up mountain
sides (but not vertical surfaces).
\item Mount an additional 50\% PS, FT \& EN.
\end{enumerate}

\item[Knight Commander]
Will get flying mount.

\item[Keep Commander]
Will get castle.

\end{description}

\end{multicols*}



\begin{multicols*}{3}

\section{Turf}

\section{Sabrina}

\section{Ty}

\begin{spell}{Call of Hero}
\range{Special}
\duration{Immediate}
\multiple{200}
\basechance{20\%}
\resist{Special}
\storage{None}
\target{Entity}
\begin{effects}
The Adept may summon any entity they can see or know the True Name of.
The target must accept the summons.  The target will immediately
appear next the caster.  The range of the spell is 20 feet + 5 / spell
Rank + 20 / Name Rank.

This spell can be cast as a ritual, in which case the range is
extended from feet to miles.  All fatigue costs are the same but it
cannot backfire as a ritual.
\end{effects}
\end{spell}

\section{Tulip}

\begin{spell}[S-1]{Entangle}
\range{20 feet + 5 / Rank}
\area{Plants in a 40 ft radius}
\duration{5 seconds + 5 / Rank}
\multiple{150}
\basechance{25\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity, Area}
\begin{effects}
Grasses, weeds, bushes, and even trees wrap, twist, and entwine about
creatures in the area or those that enter the area, holding them fast
and causing them to become entangled. The creature can break free and
move half its normal speed by using a full-round action to make a
twice PS check or a once AG check. A creature that succeeds on a Magic
Resistance is not entangled but can still move at only half speed
through the area.  Each pulse on your turn, the plants once again
attempt to entangle all creatures that have avoided or escaped
entanglement.
\end{effects}
\end{spell}

\end{multicols*}

\end{document}
