\begin{Chapter}{The College of Air Magics (Ver 2.1)}

The College of Air Magics is concerned with the environment of air and
the weather.  It should be noted that there used to be a third aspect
to this college, but the Ice Mages’ increasing interest in ice and
cold (rather than air) was considered totally unacceptable and they
were expelled (to form the independent College of Ice Magics).

When trying to ascertain the effect of spells which change the
weather, especially wind, please read the notes accompanying the
Weather Table.  An area may be affected by more than one Air College
spell or ritual at the same time providing that each change affects a
different aspect of the weather.

2.  They  will  not  suffer  from  the  effects  of  high 
altitude.  This  talent  only  works  up  to  a  height  of 
2,000 feet per rank. 

Example 
sickness, altitude sickness. 

Hypothermia  (at  high  altitudes),  flying 

3. General knowledge relating to being airborne. 

Example 
hazards. 

Safe velocities, altitude, down drafts, flight 

4.  Due  to  their  close  association  with  the  environ-
ment, the  Adept  can  modify  the  Force  downwards 
and Gauge upwards on the Weather Scale Table by 
1 per 4 ranks, for themselves. 

Traditional Colours 

Detect Fumes (T-2) 

Air  Mages  traditionally  dress  in  blues  and  greys. 
Due to their habit of being outdoors in all weathers, 
Adepts of this college tend to wear practical cloth-
ing and avoid long trailing cloaks or skirts. 

Traditional Symbols 

Given  the  ephemeral  nature  of  their  element,  Air 
Mages have little use for symbols, usually allowing 
their  traditional  colours  to  advertise  their  profes-
sion. 

18.1 Glossary 
Totally Enclosed is defined as being surrounded by 
walls or earth in a windowless area of less than 100 
feet in any dimension where there is  no direct and 
immediate communication with the air outside. 

Partially Enclosed is defined as being in a cave or 
building or similar walled or earth enclosed area of 
greater  than  100  feet  in  its  smallest  dimension,  or 
in  a  smaller  area,  but  having  means  of  direct  con-
tact  with  the  air  outside,  such  as  through  an  open 
window or portal. 

On top of a Mountain is defined as being within ten 
minutes  walking  distance  of  the  peak.  Mountains 
have an enormous impact on the weather patterns. 

18.2 Restrictions 
Adepts  of  the  College  of  Air  Magics  may  only 
practise their arts if they are in contact with the air. 
They  may  never  practise  Air  magics  while  under-
water  or  in  a  vacuum.  They  may  never  summon 
avians  into  an  environment  where  they  could  not 
survive. 

The MA requirement for this college is 13. 

18.3 Base Chance Modifiers 
The  following  modifiers  are  added  to  the  Base 
Chance  of  performing  any  talent,  spell  or  ritual  of 
the College of Air Magics. Only one modifier may 
be applied: 

Caster is underground or totally enclosed 
Caster is only partially enclosed 
Caster is flying, or otherwise not in contact 
with any liquid or solid surface 
Caster is standing in natural winds of 40 mph 
or greater 
Caster is more than 100’ above ground (fly-
ing or falling etc.) 
Caster is above all landmarks within 10 miles 
and at least 500’ above the ground (flying or 
falling etc.) 
Caster is on top of a mountain 

-15 
-5 
+5 

+10 

+10 

+20 

+20 

18.4 Talents 

Aerial Affinity (T-1) 

Experience Multiple: 75 
Effects:  Due  to  their  close  association  with  the 
environment  of  air,  Air  Mages  have  the  following 
abilities: 

1.  The  Adept  can  modify  any  flying  or  landing 
rolls by 1 step better per 5 full ranks. 

Example 
The  Air  Mage  comes  in  for  a  landing  and 
the  GM  tells  the  player  that  they  need  to  make  a  3  ×  AG 
landing  roll,  but  because  the  Air  Mage  has  this  talent  at 
rank 12 they need only make a 5 × AG landing roll. 

Range: 10 feet + 10 / Rank 
Experience Multiple: 100 
Effects:  The  Adept  may  detect  the  presence  of 
fumes  and  vapours  and  has  a  chance  equal  to  the 
Adept’s modified perception (+ 3 / Rank) of accu-
rately  identifying  each  smell  present.  The  Adept 
must  spend a pulse  sniffing the  air  while  applying 
this  talent.  If  the  Adept’s  sense  of  smell  is  not 
operating, for whatever reason, then this talent may 
not  be  used.  This  talent  may  only  be  attempted 
once per set of smells. When new smells are intro-
duced, the Adept may be able to identify these new 
smells  by  attempting  this  talent  again,  but  any 
previously unidentified smells will remain so. 

Predict Weather (T-3) 

Experience Multiple: 50 
Base Chance: 30% + 4% / Rank 
Effects:  The  Adept  may  predict  the  local  weather 
over the next day (+ 1 / 3 Ranks). The Adept must 
be outside or able to see the sky to use this ability. 
If  the  Adept  has  resided  in an area  for  a  length  of 
time  they  will  be  more  familiar  with  the  local 
weather. If they have lived in the area for over one 
month  they  will  receive  a  bonus  for  predicting 
weather in the season they are familiar with. If they 
have  lived  in  the  area  for  over  one  year  they  will 
receive a bonus in any season. 

Time 

Familiarity 

Bonus 

1 month  Current season  + 5% † 
1 year 
+10% † 
†only one modifier applies. 

Complete 

18.5 General Knowledge Spells 

Calm (G-1) 

Range: 50 feet + 50 / Rank 
Duration: 15 minutes + 15 / Rank 
Experience Multiple: 100 
Base Chance: 50% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Volume 
Effects:  The  Adept  halts  all  gross or  extreme  non-
magical  air  movement  within  a  volume  of  10  foot 
(+  10  /  Rank)  cubed.  The  entire  volume  must  be 
within  the  Adept’s  range.  The  air  within  the  vol-
ume  will  stay  fresh  and  will  slowly  intermingle 
with air outside the volume. Once cast the volume 
may not be moved. 

Feather Falling (G-2) 

Range: 10 feet + 10 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 200 
Base Chance: 40% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  This  spell  causes  the  target  to  drift  gently 
downwards if they fall more than 5 feet (should the 
target  be  falling  faster  than  5  feet  per  pulse  they 
will  undergo  a  magical  deceleration  over  one 
pulse).  The  target  will  be  subject  to  wind  effects 
while drifting downwards. 

Mage Wind (G-3) 

Range: 10 feet + 10 / Rank 

57 

18 COLLEGE OF AIR MAGICS 

Duration: 1 hour + 1 / Rank 
Experience Multiple: 125 
Base Chance: 30% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Object 
Effects:  The  Adept  causes  1  wind-driven  object to 
be  affected  by  a  magical  wind  with  an  effective 
speed of up to 5 miles per hour (+ 1 / Rank). Only 
the object targeted is affected by the magical wind. 
By concentrating the Adept may alter the direction 
and strength of the Mage wind. 

Mist (G-4) 

Range: 20 feet + 20 / Rank 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 100 
Base Chance: 20% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Volume 
Effects: The Adept conjures 1000 cubic feet of mist 
(+  500  /  additional  Rank)  of  any  shape  entirely 
within the Adept’s range. The mist must be at least 
1  foot  in  any  dimension  and  may  have  detail  no 
finer than 1 foot in size. The visibility in this mist 
is  normally  30  feet  (6  hexes).  If  the  wind  is 
stronger  than  a  light  breeze,  the  Adept  must  ac-
tively concentrate to maintain the mist. Otherwise, 
while  the  Adept  actively  concentrates,  the  mist 
may be changed in one of the following ways: 

1. The Adept may move the mist up to 1 hex/pulse 
so long as it remains within range. 

2.  The  Adept  may  change  the  visibility  within  the 
mist down to a minimum visibility of 6 hexes (-1 / 
4 full Ranks). 

Speak to Avians (G-5) 

Range: 10 feet + 10 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 75 
Base Chance: 50% 
Resist: None 
Storage: Investment, Magical Trap, Potion 
Target: Self 
Effects: The spell allows the target to communicate 
with any one type of aerial life within range of the 
spell.  This  communication  usually  consists  of 
sound and gesture. If the Adept wishes to converse 
with  several  different  types  of  aerial  life  concur-
rently then they may cast this spell multiple times. 

Storm Calling (G-6) 

Range: Special 
Duration: 60 minutes + 30 / Rank 
Experience Multiple: 200 
Base Chance: 40% 
Resist: None 
Storage: Investment, Magical Trap 
Target: Special 
Effects:  The  Adept  may  summon  any  storm  front 
which may exist anywhere in sight. If no front can 
be  seen,  the  spell  can  still  be  cast  but  the  Base 
Chance  is  modified  by  -20.  Generally,  a  storm 
front can be seen for 20 to 30 miles. Upon reaching 
the spot occupied by the Adept at the time of cast-
ing,  the  storm  front  will  slow  and  finally  cease 
moving  and  begin  to  downpour  (snow,  rain,  hail, 
sleet or whatever else the GM feels the clouds may 
contain).  The  storm  front  will  take  D10  ×  3  min-
utes  (1  /  Rank)  to  arrive,  minimum  1.  Once  the 
duration  has  lapsed  the  weather  will  gradually 
return to normal over a similar amount of time. 

Summon Avians (G-7) 

Range: 5 miles 
Duration: Immediate 
Experience Multiple: 125 
Base Chance: 30% 
Resist: Active, Passive 
Storage: Investment, Magical Trap 
Target: Avians 

18 COLLEGE OF AIR MAGICS 

Effects:  The  Adept  may  summon  one  avian  per 
Rank  (minimum  1)  to  their  current  location.  The 
avians must be native to or present in the area and 
the  Adept  must  specify  the  species  (one  per  spell) 
that is being summoned. The Adept may summon a 
specific  avian  providing  it  is  within  line  of  sight 
(the species need not been known in this case). The 
avian  will  arrive  by  the  shortest  flight  path  and 
their initial reaction will be wary. The avians must 
have a clear flight path to the Adept and will spend 
up  to  5  minutes  attempting  to  reach  the  Adept. 
Note that not all avians will have a movement rate 
which  will  allow  them  to  reach  the  Adept’s  loca-
tion within this time. This spell may  not be cast if 
the  Adept  is  totally  enclosed.  If  the  avian  being 
summoned is fantastical it gains a +20% to its base 
chance to resist this spell. 

Vapour Breathing (G-8) 

Range: 10 feet + 10 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 150 
Base Chance: 35% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  The  target  can  breathe  any  gas  including 
otherwise  poisonous  vapours  (e.g.  Knockout  Gas, 
Noxious  Vapours  etc).  This  spell  will  not  enable 
the  target  to  breathe  in  a  vacuum  or  underwater. 
Also, this spell will not protect the target from any 
non-gaseous contents of the atmosphere, e.g. Sleep 
Dust. 

Wind Whistle (G-9) 

Range: Special 
Duration: D10 hours + 1 / Rank 
Experience Multiple: 125 
Base Chance: 40% 
Resist: None 
Storage: Investment, Potion 
Target: Special 
Effects: The Adept is able to create a wind over an 
open space of up to 100 feet (+ 100 / Rank) diame-
ter  centred  on  the  Adept.  Outside  of  this  area,  the 
wind  will  fade  back  to  the  prevailing  wind  over 
half again the distance. The wind will build up over 
D  +  5  minutes  (1  /  Rank,  minimum  1)  and  the 
Mage  must  choose  at  the  time  of  casting  which 
direction  the  wind  will  blow.  The  speed  of  the 
wind  is  determined  by  rolling  a  D100  roll  as  fol-
lows: 

Dice 

Force  Speed (mph) 

5 
4 
3 
2 
3 
4 
5 
6 

19–24 
13–18 
8–12 
4–7 
8–12 
13–18 
19–24 
25–31 

01–05 
06–15 
16–30 
31–50 
51–70 
71–85 
86–95 
96–100 
The Adept can choose to modify the dice roll up or 
down  by  up  to  2  ×  Rank  points  to  determine  the 
actual  wind  speed  once  the  dice  have  been  rolled. 
The effects of this spell override any existing wind 
effects  including  already  existing  Wind  Whistle 
spells. 

If the resulting wind is Force 5 or over, missile fire 
will  be  affected,  reducing  base  chances  by  (wind 
speed/2)%. 

18.6 General Knowledge Rituals 

Windspeak (Q-1) 

Range: 400 feet + 400 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 150 
Base Chance: 2 × MA + 3% / Rank 
Resist: None 
Storage: None 
Target: Area 
Cast Time: 1 hour 
Material: None 
Actions: None 
Concentration Check: Standard 

Effects:  The  Adept  can  speak  with  the  whispering 
spirits of the wind, learning what they have seen or 
heard  and  even  soliciting  their  aid.  All  winds 
within a range of 300 feet (+ 300 per rank) can be 
communed  with in this  manner.  There  is  no  back-
fire. 

18.7 Special Knowledge Spells 

Air Blast (S-1) 

Range: 10 feet + 10 / Rank 
Duration: Immediate 
Experience Multiple: 200 
Base Chance: 30% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity, Object 
Effects: This spell projects a narrow and extremely 
strong  blast  of  air  at  the  target.  The  blast  will  im-
pact  either  on the  target  or  on the  first  obstruction 
blocking  the  path  from  the  Adept  to  the  target.  If 
the  target  weighs  less  than  100  pounds  (+  20  / 
Rank)  they  will  be  thrown  back  10  feet  (+5  per  3 
full rank). On landing the target suffers damage of 
[D  -  5]  +  1  per  3  full  ranks  and  must  make  a  1 
times  PS  +  AG  check  to  remain  standing.  The 
target is thrown in an arc away from the Adept and 
the highest point of the arc equals (Rank + 1) feet. 

Arrow Flight (S-2) 

Range: 5 feet 
Duration: 20 minutes + 20 / Rank 
Experience Multiple: 200 
Base Chance: 35% 
Resist: None 
Storage: Investment 
Target: Arrows or Quarrels 
Effects:  The  Adept  can  temporarily  improve  the 
flying quality of 2 (+ 1 / 2 Ranks) arrows or quar-
rels.  Anyone  firing  arrows  or  quarrels  affected  by 
this  spell  has  their Base  Chance  modified by  +2% 
(+ 1 / Rank). At rank 16+ the arrows are counted as 
magical  for  determining  which  entities  may  be 
harmed. 

Avian Control (S-3) 

Range: 100 feet + 100 / Rank 
Duration:  Concentration:  maximum  10  minutes  + 
10 / Rank 
Experience Multiple: 200 
Base Chance: 20% 
Resist: Active, Passive 
Storage: Investment, Potion 
Target: Avian 
Effects:  The  Adept  may  control  1  (+  1  /  Rank) 
avian  within  range  unless  the  avian  successfully 
resists. If the avian is fantastical, it gains +20% on 
its base chance to resist this spell. If the Adept fails 
to  gain  control  or  loses  control,  the  avian  will 
immediately  attack  the  Adept.  An  avian  which  is 
still controlled when the spell duration expires, will 
flee  the  Adept’s  presence.  The  Adept  may  recast 
this  spell  before  its  duration  has  expired  without 
breaking  concentration.  When  the  duration  of  the 
first  casting  expires  the  target  gets  another  resis-
tance check. The Adept may also choose to release 
an avian from their control before the spell’s dura-
tion  is  up,  whereupon  it  will  flee  as  above.  If  the 
avian is sentient and is made to take an action that 
would  endanger  itself,  another  resistance  check  is 
made. 

Barrier of Wind (S-4) 

Range: 5 feet + 1 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 150 
Base Chance: 30% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects: This spell forms a swirling pattern of wind 
around  the  entity  in  all  directions.  Thrown  and 
missile weapons passing through the barrier have a 
chance  of  being  deflected  from  their  course,  to 
impact  harmlessly  elsewhere.  The  deflection  adds 
5  (+  2  /  Rank)  to  defence  against  missiles.  This 

58 

spell provides a bonus of 5 (+ 1 / Rank) in melee or 
close combat. 

Conjuring Air (S-5) 

Range: 10 feet + 10 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 150 
Base Chance: 25% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Volume 
Effects:  The  Adept  may  conjure  anywhere  within 
their  range  5  cubic  feet  (+  5  /  Rank)  of  sweet 
smelling  breathable  air  in  any  shape  or  shapes  of 
their  choice.  No  dimension  of  the  shape  may  be 
less  than  1  foot.  The  volume  of  air  will  hold  its 
shape  versus  any  non  solid  intrusions  for  the  full 
duration. At the end of the duration the volume will 
rapidly  disperse  into  the  surrounding  area,  mixing 
with whatever atmosphere was already there. 

Flying (S-6) 

Range: 5 feet + 5 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 250 
Base Chance: 30% 
Resist: Active, Passive 
Storage: Investment, Potion 
Target: Entity 
Effects:  This  spell  enables  the  target  to  fly  at  a 
speed  of  30  mph  (+  1  /  Rank)  by  walking  winds 
and  air  currents.  Initially  it  will  take  [D  +  10  - 
Rank]  minutes  for  the  wind  currents  to  arrive. 
After the wind currents have arrived the target can 
take off and land as many times as they desire, but 
to  recall  the  winds  to  take  off  after  landing  will 
take  [D  + 2]  pulses.  When  taking  off,  it takes  one 
pulse  to  accelerate  to  full  speed.  If  the  target  tries 
to land, it will take one pulse to slow to a standstill. 
The  target  can  only  travel  into  places  where  air 
currents  or  winds  are  possible.  This  will  not  nor-
mally occur inside buildings or tunnels. 

Gaseous Form (S-7) 

Range: Self, touch at Rank 11+ 
Duration: 5 seconds + 5 / 2 Ranks 
Experience Multiple: 300 
Base Chance: 20% 
Resist: Active, Passive 
Storage: Potion, Invested, Magical Trap 
Target: Entity 
Effects:  When  this  spell  is  cast  the  target  (and 
possessions)  turn  into  a  cloud  of  vapour,  that  ap-
pears  as  a  heavy  mist.  The  target  may  be  affected 
by  magic  spells.  Normal  or  silvered  weapons  do 
not harm the target, but magical weapons may. The 
target  may  not  use  any  possessions  such  as  weap-
ons,  nor  may  the  target  cast  or  trigger  magic. 
Magical  or  Racial  Talents  may  be  used  however. 
The target may move up to TMR 2 in any direction 
and  pass  through  any  opening  that  is  not  airtight. 
When  the  duration  of  the  spell  expires,  the  target 
will  reform  in  the  nearest  hollow  space  large 
enough to accommodate the target’s body. 

Gliding (S-8) 

Range: 5 feet + 5 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 200 
Base Chance: 30% 
Resist: None 
Storage: Investment, Potion 
Target: Sentient Entity 
Effects: This spell  enables the target to glide at an 
angle  of  approximately  1  + Rank  feet  horizontally 
for  every  foot  of  descent.  The  target’s  maximum 
forward speed will be approximately 80 feet / pulse 
(+  1  /  Rank)  and  their  minimum  vertical  speed 
(relative  to  the  air  around  them)  will  be  approxi-
mately  40  feet  /  pulse  (-2  /  Rank,  minimum  1) 
downwards.  If  the  target  is  gliding  in  thermals  or 
down-drafts, their actual rate of descent (or ascent) 
may  vary.  The  actual  rate  of  descent  is  up  to  the 
GM.  The  target  must  keep  their  arms  (or  equiva-
lent)  outstretched  until  landing  or  they  will  fall 
downwards,  although  the  spell  will  resume  if  the 

target returns their arms to an outstretched position. 
The target flies with their body horizontal and may 
turn  at  a  rate  of  30  degrees  /  pulse  (+  5  /  Rank) 
while  gliding.  The  spell  will  cause  the  target  to 
automatically  become  vertical  when  within  5  feet 
of a horizontal surface. The target must then make 
a flying roll when landing. 

Knockout Gas (S-9) 

Range: 30 feet + 10 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 450 
Base Chance: 15% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Area 
Effects: This spell creates a heavy cloud of knock-
out  gas  that  rises  10  (+  1  /  Rank)  feet  above  the 
ground.  The  area  has  a  diameter  of  15  feet  (+  5  / 
Rank).  At  ranks:  0  to  9  the  cloud  appears  as  a 
translucent mist (i.e. does not inhibit vision) and is 
pungent smelling, 10 to 15 the gas is invisible and 
pungent smelling, 16 and above the cloud becomes 
invisible  and  odourless.  Any  entity  (including  the 
Adept)  within  the  gas  must  resist.  If  they  fail  to 
resist,  they  must  roll  under  2  ×  EN  each  pulse  or 
fall  unconscious  (not  asleep)  while  they  remain 
within  the  gas.  Any  wind  over  15  mph  will  dis-
perse  the  gas  in  D-Force  pulses.  Once  the  gas  is 
dispersed,  the  duration  expires,  or  a  victim  is  re-
moved from the area, any unconscious entities will 
recover in 1 pulse (+ 1 / 5 full Ranks). 

Lightning Bolt (S-10) 

Range: 60 feet 
Duration: Immediate 
Experience Multiple: 225 
Base Chance: 30% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity, Object 
Effects:  The  Adept  may  throw  a  single  bolt  of 
lightning  60  feet  long  from  their  fingertips.  The 
bolt must extend the entire 60 feet and will deflect 
off stone until it reaches its full extent. All targets 
that are in the path of the bolt must resist or suffer 
[D + 5] (+ 1 / 3 Ranks) damage (save for half dam-
age).  In  addition,  any  target  who  fails  to  resist  is 
automatically stunned. 

Lightning Strike (S-11) 

Range: 15 feet + 5 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 250 
Base Chance: 35% 
Resist: Active, Passive 
Storage: Potion 
Target: Entity 
Effects:  This  spell  creates  sheets  of  lightning  that 
slowly arc around the entity’s body doing no dam-
age.  This  is  extremely  obvious.  The  first  time  the 
entity takes damage from a blow in combat and the 
attacker is within range, a bolt will strike out at the 
entity’s  attacker  doing  [D  -  4]  (+  1  /  2  Ranks) 
damage  (save  for  half  damage).  The  spell  then 
dissipates.  The  Range  is  5  feet  (+  5  /  Rank).  At 
rank  10  or  greater,  any  attacker  who  fails  to  resist 
and  is  capable  of  being  stunned  is  automatically 
stunned.  This  spell  cannot  be  cast  on  a  target  if 
they  are  already  under  the  effect  of  a  Lightning 
Strike spell. 

Resist Cold (S-12) 

Range: Touch 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 100 
Base Chance: 40% 
Resist: None 
Storage: Investment, Ward, Potion 
Target: Entity 
Effects:  This  spell  protects  the  target  from  the 
effects  of  cold  temperature  by  increasing  the 
Gauge by 1 (+ 1 / 4 full ranks) up to a maximum of 
Gauge  7  (Comfortable).  It  will  totally  protect  the 
target from the effects of Hypothermia at rank 11+. 
In addition, the target suffers 1 (+ 1 / 4 full ranks) 

less  damage  due  to  magical  or  non-magical  cold 
based attacks. 

Shaping Cloud (S-13) 

Range: 5 miles + 1 / Rank 
Duration: Concentration: maximum 1 minute + 1 / 
Rank 
Experience Multiple: 150 
Base Chance: 40% 
Resist: None 
Storage: Investment, Potion, Ward 
Target: Cloud band 
Effects: The Adept can cause some of a cloud bank 
(cloud degree 5+) within range and line of sight to 
slowly  shape  a  different  image  in  it  each  minute. 
Each new shape will be gradually formed from the 
old over the course of the minute. The shape in the 
clouds is recognisable up to a range of 5 miles (+ 1 
/  Rank).  Measurements  for  this  spell  should  be 
taken from the ground (i.e. disregarding the height 
of the cloud bank). 

Whirlwind Vortex (S-14) 

Range: 15 feet + 15 / Rank 
Duration: Immediate 
Experience Multiple: 650 
Base Chance: 1% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  This  spell  creates  a  tornado  around  one 
human  sized  target  for  every  3  or  fraction  ranks 
(minimum 1). If the target fails to resist they are so 
tossed  and  torn  by  the  winds  that  muscles  and 
ligaments tear, joints dislocate, bones break, organs 
rupture,  and  they  perish.  If  the  target  successfully 
resists they suffer [D - 4] (+ 1 / Rank) damage due 
to  excessive  forces  (instead  of  perishing).  Targets 
that fail to resist may be resurrected. 

Whispering Wind (S-15) 

Range: 100 miles + 100 / Rank 
Duration: Special 
Experience Multiple: 150 
Base Chance: 40% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  This  spell  sends  a  message  to  an  entity 
using the winds. The Adept must know the name of 
the  entity  and  be  able  to  pick  them  out  from  a 
crowd.  The  maximum  length  of  the  message  is  5 
words per rank. The messages will be in the voice 
of  the  Adept  at the  same  volume  as  when  spoken, 
and  can  only  be  heard  by  the  target.  There  is  a 
noticeable but minor effect of swirling air when the 
message  is  sent  and  received.  The  time  taken  for 
the  message  to  reach  the  entity  (the  duration)  is  1 
hour (+ 2 minutes per mile) - rank hours (minimum 
of 1 hour). 

Windstorm (S-16) 

Range: 30 feet + 30 / Rank 
Duration: 10 seconds + 10 / Rank 
Experience Multiple: 200 
Base Chance: 40% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Area 
Effects: The Adept creates a windstorm of Force 9 
centred on the Adept and extending out to the full 
range  of  the  spell.  Once  cast  the  Windstorm  will 
not move. The winds in the area of a windstorm are 
random  and  violent,  they  do  not  prevail  in  a  par-
ticular direction. All entities within this area except 
the  Adept  and  those  in  the  same  hex  must  check 
against  2  ×  (Physical  Strength  +  Agility)  -  2  × 
Rank,  every  pulse,  to  regain  their  feet  and/or  re-
main upright. Every  time  an  entity  within the area 
falls prone, they take [D - 2] damage. This damage 
is physical. For the duration of this spell they have 
their  TMR  halved.  All  entities  attempting  to  use 
missile  or  thrown  weapons  through  or  inside  the 
area of effect have their Base Chance reduced by 5 
per every 2 hexes of Windstorm travelled through. 

59 

18 COLLEGE OF AIR MAGICS 

Wind Walking (S-17) 

Range: 10 feet + 10 / Rank 
Duration: 30 seconds + 30 / Rank 
Experience Multiple: 350 
Base Chance: 25% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  The  target  of  this  spell  (and  possessions) 
turns into wind, keeping their approximate size and 
appearance. In the first pulse, the spell’s effect is to 
accelerate  the  target  to  full  speed  and  in  the  last 
pulse,  the  target  will  gradually  slow.  The  distance 
travelled over the first and last pulse is half that of 
normal. The target moves through the air at 50 mph 
(+  5  /  Rank).  The  target  can  only  pass  through  an 
area  that  they  could  pass  through  normally.  The 
target  chooses  the  direction  of  the  flight  but  the 
speed  may  not  be  altered.  The  target  will  not  col-
lide with stationary objects during the flight, but if 
rushing  directly  towards  an  object  or  wall,  will 
brush  along  or  past  it  as  a  wind  would.  If  blown 
down a corridor to a dead end or similar situation, 
the  target  will  not  stop  but  double  back  in  a  tight 
loop.  Anything  dropped  by  the  target  will  go  in  a 
random direction. 

Ball of Lightning (S-18) 

Range: 35 feet + 10 / Rank 
Duration: Immediate 
Experience Multiple: 350 
Base Chance: 30% 
Resist: Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity or Object 
Effects: This spell creates a ball of lightning which 
shoots  from  the  caster  to  the  target.  The  ball  is 
utterly silent and moves in a straight line. Anything 
standing  between  the  caster  and  the  target  will  be 
struck  instead.  Upon  striking  anything  the  ball 
explodes, in a bright flash, causing [D - 1] (+ 1 per 
Rank)  electrical  damage.  If  the  target  successfully 
resists  the  damage  is halved  (round up),  otherwise 
they  are  also  blinded  for  Rank  /  4  pulses  (round 
down). 

Thunderclap (S-19) 

Range: 20 feet + 20 / Rank 
Duration: Immediate 
Experience Multiple: 325 
Base Chance: 30% 
Resist: Passive 
Storage: Investment, Ward, Magical Trap 
Target: Area 
Effects:  The  Adept  causes  the  air  in  the  targeted 
area  to  violently  compress  with  a  loud  crash.  The 
target area has a diameter of 5’ at Ranks 0–5, 15’ at 
Ranks 6–12, 25’ at Ranks 13–19, and 35’ at Rank 
20. The entirety of the affected area must be within 
the caster’s spell range for the spell to be effective. 
All those within the area suffer [D + 1] (+ 1 per 2 
full  Ranks)  concussive  damage  (resist  for  half  — 
round up). Those failing to resist can hear nothing 
except a loud ringing for Rank pulses. On a Double 
or Triple effect any entities which fail to resist are 
also  stunned  (normal  stun  recovery  applies).  Note 
that  this  spell  can  be  heard  from a  distance  as  per 
normal thunder. 

18.8 Special Knowledge Rituals 

Air Spring (R-1) 

Duration: Rank × Rank hours (minimum 1) 
Experience Multiple: 150 
Base Chance: MA + 5% / Rank 
Resist: None 
Storage: None 
Target: Area 
Cast Time: 1 hour 
Material: None 
Actions: None 
Concentration Check: Standard 
Effects:  The  Adept  can  open  a  one  way  portal  2 
feet  in  diameter  from  the  elemental  plane  of  Air 
from which fresh clean air flows. The rate of flow 
is  5  cubic  feet  (+  5  /  Rank)  per  second.  Addition-

18 COLLEGE OF AIR MAGICS 

ally any modifier for being enclosed is reduced by 
5. 

Conjuring Air Elemental (R-2) 

Range: 20 feet 
Duration: Concentration: No maximum 
Experience Multiple: 200 
Base Chance: MA + 5% / Rank 
Resist: None 
Storage: None 
Target: Air Elemental 
Cast Time: 1 hour 
Material: None 
Actions: None 
Concentration Check: Standard 
Effects: The Adept may summon an Air Elemental 
and bind it to temporary service by performing this 
ritual. At the end of the ritual a Cast Check is per-
formed.  If  the  ritual  is  successful  the  Elemental  is 
summoned  and  controlled.  If  the  ritual  backfires 
then the Elemental is summoned but no controlled 
and will attack the summoner and companions. 

The  Air  elemental  will  always  appear  within  20 
feet  of  the  summoner.  If  has  a  combined  Endur-
ance  and  Fatigue  equal  to  15  (+  15  /  Rank).  The 
Elemental  will  remain  until  it  is  sent  back  to  its 
home plane by the Adept (with Special Knowledge 
Counterspell of the college of Air Magics) or ban-
ished.  If  it  is  controlled  by  the  summoner  it  will 
remain controlled until the summoner’s concentra-
tion is broken. 

Control Weather (R-3) 

Duration: 8 hours + 8 hours / Rank 
Experience Multiple: 300 
Base Chance: 2 × MA + 3% / Rank 
Resist: None 
Storage: None 
Target: Area 
Cast Time: 1 hour 

Actions: Dance (2 FT when unencumbered) 
Concentration Check: None 
Effects: The Adept may change one or more of the 
three  components  which  make  up  the  weather  by 
performing a ritual dance. The three components of 
weather are:  

• Precipitation, Degree  

• Temperature, Gauge  

• Wind, Force 

The  GM  should  consult  the  weather  table  and 
advise  the  Adept  of  the  current  level  of  each  of 
these  three  components  before  they  start  dancing. 
The Adept may change the current components by 
a total of 1 +1 per 2 full ranks. 

At rank 12, the Adept can change a Force 3 
Example 
wind  into  a  Force  10  wind  or  Force  3  into  Force  7  and 
Degree 5 into Degree 2 cloud cover. 

All  the  changes  may  be  in  any  direction  on  the 
table. 

This  ritual  will  not  cause  weather  effects  outside 
the  normal  climatic  range  of  the  area  (as  deter-
mined  by  the  GM).  The  weather  will  change 
gradually  over  (30  -  1  /  Rank)  minutes  per  level 
shifted  on the  table  and  the  three  components  will 
change simultaneously. The area of effect is circu-
lar  with  a  radius  of  1  mile  /  Rank  (minimum  1). 
Casting this ritual counts as strenuous activity and 
the Adept will lose fatigue. This ritual cannot back-
fire. 

Summon and Bind Cloud (R-4) 

Range: 5 miles + 5 / Rank 
Duration: 5 hours + 5 / Rank 
Experience Multiple: 300 
Base Chance: 2 × MA + 3% / Rank 
Resist: None 
Storage: None 

Target: Clouds 
Cast Time: 1 hour 
Material: None 
Actions: None 
Concentration Check: Standard 
Effects:  This  ritual  will  summon  a  quantity  of 
cirro-cumulus  cloud  and  change  its  consistency  to 
that of packed cotton wool to provide a method of 
transport.  Items  will  pass  through  the  cloud  after 
30  seconds  unless  supported  by  an  Entity.  The 
cloud  arrives  during  the  last  half  hour  of  casting 
and  spends  the  first  and  last  half  hour  of  travel 
rising to and descending from its natural travelling 
height.  The  natural height  of  cirro-cumulus clouds 
is between 2 and 4 miles above sea level (10,000 to 
20,000 feet). The cloud may support 1 entity (+ 1 / 
Rank)  and  transports  them  in  a  comfortable  and 
oxygenated  environment,  although  it  may  be 
slightly  cold  if  improperly  clothed.  Flying  entities 
may  take  off  or  land  on  the  cloud  as  long  as  the 
above limit is maintained. If more entities land on a 
cloud than it can carry, it will immediately start to 
descend  taking  half  an  hour  and  dissipate  upon 
reaching  the  ground.  The  cloud  appears  to  be  a 
normal  cloud  but  is  sufficiently  soft  to  prevent 
injury  to  anything  that  impacts  it  (it’s  also  non-
flammable). The clouds will move with the prevail-
ing  winds  or  can  be  moved  with  magical  winds 
such as Mage Wind. By actively concentrating the 
Adept  can  cause  the  clouds  to  move  at  a  different 
speed  or  in  a  different  direction  to  the  prevailing 
wind.  The  Adept  can  alter  the  movement  of  the 
clouds by 2 miles per hour (+ 2 / Rank) in any one 
direction.  The  altitude  of  the  clouds  may  not  be 
controlled.  While  in  contact  with  the  ground  the 
clouds will not move.  

 

 

60 

18.9 Weather Scale Table 

Force 

Wind 

Specification 

18 COLLEGE OF AIR MAGICS 

Speed (mph) 

0 

1 

2 

3 

4 

5 

6 

7 

8 

9 

10 

11 

12+ 
Gauge 

0 

1 

2 

3 

4 

5 

6 

7 

8 

9 

10 

11 

12+ 
Degree 

0 
1 

2 

3 

4 

5 

6 

7 

8 

9 

10 

11 

Calm 

Light air 

Light breeze 

Gentle breeze 

Smoke rises vertically 

Direction indicated by smoke only 

Wind felt on face, leaves rustle 

Leaves and twigs in constant motion, wind extends light flag 

Moderate breeze 

Wind raises dust and loose paper, small branches move 

Fresh breeze 

Strong breeze 

Near gale 

Gale 

Strong gale 

Storm 

Violent storm 

Hurricane 
Temperature 

Arctic 

Arctic 

Arctic 

Arctic-like 

Sub-arctic 

Sub-arctic 

Sub-tropical 

Sub-tropical 

Sub-tropical 

Tropical 

Tropical 

Desert-like 

Desert 
Cloud 

Clear 
Clear 

Clear 

Sparse 

Light 

Cloudy 

Overcast 

Heavy Cloud 

Dark Cloud 

Small trees in leaf start to sway, crested wavelets on inland waters 

Large branches in motion, whistling through trees 

Whole trees in motion, inconvenient to walk against wind 

Twigs break from trees, difficult to walk 

Slight structural damage occurs 

Trees uprooted, considerable structural damage occur 

Widespread damage 

Widespread damage 
Specification 

Dangerous 

Salt water freezes 

Branches become brittle 

Dangerously cold if not dressed in winter clothing 

Water freezes, numbness, precipitation becomes hail, snow, sleet 

Cold, uncomfortable if poorly clothed 

Cool 

Comfortable 

Comfortable, warm 

Hot, uncomfortable if poorly clothed 

Uncomfortably hot, sub-tropical plants wither 

Extreme heat, sub-tropical plants die, tropical plants wither 

Debilitating, tropical plants die 
Precipitation 

Dry, high fire danger 
Dry, fires easy to start 

Comfortable 

Humid, uncomfortable in high temperature 

Damp 

Drizzle, fog in cold conditions 

Showers 

Light rain, leaves move, fires difficult to start 

Average rain 

Low black cloud 

Heavy rain, small branches move, small fires doused 

Oppressive 

Oppressive 

Torrential rain, river rise, large fires doused 

Flood warning, rivers burst their banks 

< 1 

1–3 

4–7 

8–12 

13–18 

19–24 

25–31 

32–38 

39–46 

47–54 

55–63 

64–73 

74 and above 
Degrees Celsius 

-20 and below 

-15 

-10 

-5 

0 

5 

10 

15 

20 

25 

30 

35 

40 and above 
mm / Hour 

0 
0 

0 

1 

2 

3 

4 

5 

10 

15 

20 

25 

Oppressive 

12+ 
T-1 Aerial Affinity (Air Mage Talent) Due to their close association with the environment the Adept can modify the Force downwards, and Gauge upwards on 
the Weather Scale Table by 1 per 4 rank, for themselves. 

Flash floods 

35 

S-12 Resist Cold (Air Mage Spell) This spell protects the target from the effects of cold temperature by increasing the Gauge by 1 (+ 1 / 4 full ranks) up to a 
maximum of Gauge 7. 

R-3 Control Weather (Air Mage Ritual) The Adept may change the current components by 1 + 1 per 2 full ranks. The resulting weather effects cannot be outside 
the normal climatic range of the area. 

Example 
direction on the table. 

At rank 12, the Adept can change a Force 3 wind into a Force 10 wind or Force 3 into Force 7 and Degree 5 into Degree 2 cloud cover. Each of the changes may be in any 

\end{Chapter}
