\begin{Chapter}{Assassin (Ver 1.0)}

Assassin is not a skill which should be carelessly chosen, as the
skill is not looked upon with great favour by members of society (at
least until they need one).  Assassins will vary in philosophy and
methods; they may be cold-hearted but not necessarily evil.  The GM
must allow assassins to prac- tice their art as they wish, and this
may result in solo adventures.

\section{Restrictions}

An assassin must be able to use the sap or garotte at a minimum of
Rank 1 before advancing past Rank 2.

\section{Benefits}

\subsubsection{An assassin is trained in using envenomed weapons.}

An Assassin may envenom weapons and use them in combat without making
the 3 × MD check required by others (§6.13 Envenomed Weapons).

\subsubsection{An assassin increases their chance of causing a Grievous Injury as
their Rank increases.}

If the assassin is attacking in a surprise situation, their chance of
causing a Grievous Injury is increased by 2\% per Rank.  If an
assassin attacks a victim through a rear hexside during combat, their
chance of causing a Grievous Injury is increased by 1\% per Rank.  If
an assassin attacks a victim through a front hexside during combat,
their chance of causing a Grievous Injury is increased by 1\% per
three full Ranks. An assassin does not gain the above bonuses when
engaging in Ranged Combat.


\subsubsection{An assassin may gain information from a victim through torture.}

The assassin must torture their victim for a period equal to (Victim’s
Willpower / Assassin’s Rank) hours to try to gain an important
secret. The assassin’s chance of forcing the victim to reveal the
secret is ( 10 × Assassin’s Rank - 4 × Victim’s Willpower)\%.  If the
GM’s roll on percentile dice is equal to or less than the success
percentage, the assassin gains the exact information they require. If
the assassin fails to gain the desired information, they may try
another torture attempt. A victim can withstand a number of torture
attempts equal to one-fifth their Endurance (round down) before they
die.

If the information an assassin seeks is not of great importance to the
victim, the GM should decrease the time required to gain it and
increase the assas- sin’s success chance appropriately.

\subsubsection{An assassin is trained to improve their memory.}

Whenever an assassin wishes to recall the details of a place or
routine they have surveyed, the GM rolls percentile dice.  If the roll
is equal to or less than (5 × Perception + 2 × Rank), the assassin has
a perfect memory of the place or routine. If the roll is greater than
the assassin’s success percentage, the GM should include more and more
erroneous information into their description as the roll approaches
100.

The GM may decrease the success percentage for difficult feats of
memory.

\subsubsection{An assassin is able to buy poisons, distilled venoms and acids at cost
(i.e. no mark-up) from an alchemist.}

\subsubsection{An assassin inflicts increased damage on their target when attacking
through a rear hexside in Melee Combat.}

Increase the damage caused by a blow from an assassin by 1 per Rank
when they strike their target through a rear hex in Melee Combat.

\subsubsection{An assassin increases their chance of knocking out a target.}

An assassin may attempt to automatically knockout a target when using
a sap.  They must make a successful strike while attacking from behind
or surprise against an unhelmeted opponent.

In addition their chance of knocking out a target (see §6.9) with a
sap is increased by 2\% for each Rank they have achieved in assassin.

\subsubsection{An assassin increases their chance of performing any action involving
stealth.}

An assassin increases stealth (see §29.4) by 2\% per Rank.

\end{Chapter}
