\begin{Chapter}{Backfire Table}

Extra Fatigue The following effects result in the 
Adept losing extra Fatigue in order to contain the 
Backfire effect. If no Fatigue was expended during 
the cast then there will be no apparent effect. If the 
Adept has insufficient Fatigue, then they lose all 
remaining Fatigue and re-roll on this table for each 
Fatigue point they were unable to expend, with any 
further rolls of Fatigue loss taken as damage due to 
internal injuries. 

\begin{Description}

\item[01–09] Fatigue loss equal to that already expended.

\item[10–16] Fatigue loss equal to twice that already expended.

\item[17–21] Fatigue loss equal to 3 times that already expended.

\item[22–24]   Fatigue loss equal to 4 times that already 
expended. 

\item[25]  Fatigue loss equal to 5 times that already 
expended. 

\end{Description}

Spell Awry The following effects result in the spell working in a
manner that was not intended by the Adept. The spell may still be
resisted by the eventual target, if the spell is resistible. Note that
the Adept will not necessarily be aware of the outcome, since in many
cases there will be no apparent effect.

\begin{Description}

\item[26–30] The Adept becomes the target of the spell.

\item[31–32] The Adept becomes the target of the spell with some or
  all effects doubled.

\item[33] The Adept becomes the target of the spell with some or all
  effects tripled.

\item[34–35] The Adept becomes the target of the spell and the spell
  is delayed by D10 pulses.

\item[36] The Adept becomes the target of the spell and the spell is
  delayed by D100 pulses.

\item[37–39] The spell has opposite or different effect to that which
  it was designed.

\item[40–41] The spell’s effects are delayed by D10 pulses.

\item[42] The spell’s effects are delayed by D100 pulses.

\item[43–45] The spell’s effects are intermittent with D10 pulses or
  minutes on, followed by D10 pulses or minutes off.

\item[46–48] The spell affects a random target or area within range,
  or goes in a random direction.

\item[49] The spell affects a random target or area within twice
  range.

\item[50] The spell affects a random target or area within three times
  range.

\item[51] The spell is cast at random as though the caster is a random
  entity within D10 hexes.

\item[52–53] The spell affects a random target or area within range
  with some or all effects doubled.

\item[54] The spell’s effects are delayed by D10 pulses and affects a
  random target or area.

\item[55] The spell’s effects are delayed by D100 pulses and affects a
  random target or area.

\item[56–57] The spell works with some or all effects halved.

\item[58] The spell works as normal.

\item[59] The spell works with some or all effects doubled.

\item[60] The spell works with some or all effects tripled.

\end{Description}


Minor Curses The following effects result in the 
Adept being afflicted by a minor curse. The Adept 
may resist the curse by using their Magical Resis-
tance against non–College Magic. Some curses can 
be cured by healing skills, and all may be dispelled 
by a curse removal. Any result specifying a gradual 
loss of statistics will also cause an immediate loss. 

\begin{Description}

\item[61]  Blind for D10 pulses. 

\item[62] Blind for D10 × D10 minutes. 

\item[63] Blind for D10 × D10 hours. 

\item[64] Blind for D10 days. 

\item[65] Deaf for D10 pulses. 

\item[66] Deaf for D10 × D10 minutes. 

\item[67] Deaf for D10 × D10 hours. 

\item[68] Deaf for D10 days. 

\item[69] Mute for D10 pulses. 

\item[70] Mute for D10 × D10 minutes. 

\item[71] Mute for D10 × D10 hours. 

\item[72] Mute for D10 days. 

\item[73] Lose smell and taste for D10 days. 

\item[74] Lose smell and taste for D10 × D10 days. 

\item[75] Lose tactile sense for D10 days. 

\item[76] Lose tactile sense for D10 × D10 days. 

\item[77] Insomnia such that only 1 Fatigue is recovered for each hour
  of sleep for D10 days.

\item[78] Insomnia such that only 1 Fatigue is recovered for each hour
  of sleep for D10 × D10 days.

\item[79–80] A virulent skin disease halves Physical Beauty and causes
  intense itching which will increase the difficulty of concentration
  checks by 1, until stopped by Cure Disease.

\item[81] Wasting disease causes -1 Strength and -1 Endurance per day
  until stopped by Cure Disease. The Strength and Endurance lost will
  be recovered at 1 point per day, or by being treated by Repair
  Muscles.

\item[82–83] Periodic muscle spasms lasting D10 pulses cause a loss of
  1 Fatigue each pulse. There is D10 × D10 minutes between
  spasms. This can be cured by Repair Muscles.

\item[84] A deep sleep for D10 pulses.


\item[85] A deep sleep for D10 × D10 minutes.

\item[86–87] Recurring migraines cause a loss of 2 Magical Aptitude
  and 2 Willpower. Each minute of concentration requires a 4 ×
  Willpower concentration check. The effects can be treated by Soothe
  Pain and cured by Repair Vital Organs.

\item[88] Periodic hallucinations for D10 hours.  Each hallucination
  lasts D10 pulses and there is D10 × D10 minutes between them. Can be
  cured by Repair Vital Organs.


\item[89–90] Arthritis causes -4 Dexterity, -4 Agility and increases
  by 1 per hour the Fatigue loss due to exercise, until treated by
  Repair Tissues.

\item[91–92] Enfeeblement causes -4 Strength, -4 Endurance and doubles
  the Fatigue loss due to exercise, until treated by Repair Muscles.

\item[93] Asthma causes TMR to be halved, doubles the Fatigue loss due
  to exercise, and the Adept cannot perform strenuous exercise until
  treated by Repair Vital Organs.

\item[94–95] Creeping senility will cause a loss of 1 Magical Aptitude
  every two days until treated by Regenerate Vital Organs.

\item[96] Partial Amnesia causes the loss of all Magical abilities for
  D10 days.

\item[97] Partial Amnesia causes the loss of all Skills (excluding
  Magic and Weapons) for D10 days.

\item[98] Partial Amnesia causes the loss of all memories from the
  past 2D10 months.  The Adept will operate at lower ranks in the
  abilities that have been ranked during this period. The memories
  will return at a rate of 1 month each day.

\item[99] Total Amnesia causes the loss of all memories for D5 × D5
  days. All magic and skills other than the primary language will be
  lost, and all weapon ranks will be halved (round down) or lost if
  Rank 0.  The Adept’s original personality will come to the fore and
  they may need to make a reaction roll to determine their initial
  feelings towards each person.

\item[00] Roll two more times and apply both effects.

\end{Description}

\end{Chapter}
