\begin{Chapter}{The College of Bardic Magics (Ver 2.1)}

The College of Bardic Magic deals primarily with sound, language,
rhythm, and the power in music, particularly the power to soothe,
charm, and otherwise affect emotions.  Adepts of this College are
generally known as Bards.  Many Bards are employed as court musicians,
though a number exercise their art as sole practitioners, wandering
the highways and byways of the land.

Almost all Bards have performance skills, often highly ranked.
Alongside this they often have the Spy skill which is complemented by
certain of the magics of this College. It is not unknown for Bards to
pass themselves off as simple musicians or entertainers in order to
infiltrate unfriendly courts.  It is this fact, and the consequently
secretive nature of this College, which has lead to its relative
obscurity.

The principal difference between this college and all others is that
the spell’s verbal component is usually sung, instead of spoken.
Furthermore, because a musical instrument may be used to enhance the
effect, Bardic spells have little or no somatic component.  It is
almost impossible for a Bard to cast quietly.

Traditional Colours 

Bards usually wear an item of a particular shade of deep blue,
sometimes known as Harper Blue, somewhere on their person. Otherwise
they tend to dress to the occasion.

Traditional Symbols 

The symbol of the Bardic College is a golden harp.  This may be worn
as a brooch or amulet or embossed on the Bard’s instrument if
appropriate.

12.1 Restrictions 

Adepts  of  the  College  of  Bardic  Magics  can  only 
practise  their  art  in  a  region  where  sound  can  be 
heard. It is not possible to practise Bardic magic in 
a  location  where  sound  is  silenced,  magically  or 
otherwise.  Bardic  magic  may  be  practised  under-
water,  but  the  results  may  vary  somewhat  from 
those expected. 

Melodic Memory (T-3) 

Experience Multiple: 75 
Effects: The Adept can attempt to commit to mem-
ory any sounds that they can hear. The Adept must 
concentrate  to  activate  this  talent,  and  the  Adept’s 
player  should  represent  this  by  taking  down  key 
words  and  phrases.  By  means  of  this  talent  the 
Adept  can  memorise  music,  dramas,  dialogue  etc. 
even  if  they  cannot  understand  them.  The  Adept’s 
chance of success to recall the sounds is 2 × PC (+ 
5 / Rank) - 5 / week since the sounds were heard. If 
the Adept rolls above their Base Chance the higher 
the roll the greater the degree of error. 

Project Voice (T-4) 

Experience Multiple: 75 
Effects: The talent allows the Adept to project their 
voice  so  that  it  may  be  heard  clearly  everywhere 
within 25 feet (+ 25 / Rank). 

12.4 General Knowledge Spells 

Clairaudience (G-1) 

Range: 15 feet + 15 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 200 
Base Chance: 15\% 
Resist: None  
Storage: Potion 
Target: Self 
Effects:  The  Adept  creates  an  invisible,  intangible 
ear  that  can  be  moved  about  within  the  spell’s 
range. The ear appears in the same hex as the Bard, 
and  operates  as  a  normal  ear  except  that  it  is  not 
physically  attached  to  the  Bard.  The  Adept  may 
move  the  ear  at  a  TMR  of  up  to  the  Rank  of  the 
spell, taking pass actions to do so. The ear may be 
detected  by  Witchsight  or  similar  means  of  detec-
tion.  If  the  ear  takes  any  magical  damage  (it  may 
be struck by a magical weapon) it is destroyed and 
the  Adept  is  stunned.  It  cannot  be  used  to  target 
spells. In the area of a Confusion of Tongues spell 
the  Adept  will  only  hear  meaningless  jumbled 
words. Clairaudience will not penetrate the volume 
of a Shell of Silence spell. 

The MA requirement for this College is 16. 

Enchant Instrument (G-2) 

12.2 Base Chance Modifiers 
The  following  numbers  are  added  to  the  Base 
Chance  of  performing  any  talent,  spell  or  ritual  of 
the College of Bardic Magics: 

For each Rank of Troubadour (Singing) 
(This bonus is only applicable when spell is 
sung) 
Area is acoustically excellent 
Area deadens sound 
All modifiers are cumulative. 

+1 

+5 
-5 

12.3 Talents  

Concealed Casting (T-1)  

Experience Multiple: 150 
Effects: The Adept is able to conceal the casting of 
a  spell  within  the  words  of  a  mundane  song.  If  an 
observer  is  listening  intently  the  observer  may 
make a (2 × PC - Rank) check to notice. The bard 
may  not  make  any  movements  which  are  inappro-
priate  to  the  song  being  sung.  All  aspects  of  the 
spell  (including  Base  Chance)  must  be  cast  at  the 
lower  of  this  talent  and  the  spell’s  Rank  and  the 
Base Chance is further reduced by 20\%. This Tal-
ent  only  functions  with  spells  of  the  Bardic  Col-
lege. 

Enhanced Hearing (T-2) 

Experience Multiple: 75 
Effects: The Adept is able to hear sounds too faint 
to  be  heard  normally.  The  Rank  of  this  talent 
should  be  added  to  the  Adept’s  perception  for  the 
purposes  of  detecting  sounds  only.  This talent  can 
be activated and deactivated at will, and the Adept 
may be temporarily deafened by loud noises while 
it is activated. 

Range: Touch 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 100 
Base Chance: 25\% 
Resist: None 
Storage: Investment 
Target: Musical Instrument 
Effects:  With  this  spell  the  Adept  draws  on  the 
resonance left in an instrument from it having been 
played. The spell enchants one musical instruments 
so that a being may play it at an effective Trouba-
dour  Rank  equal  to  0  +  (1  /  3  or  fraction  ranks), 
whether or not they know how to play that form of 
instrument.  This  Rank  may  not  exceed  the  maxi-
mum  Rank  the  instrument  being  targeted  has  pre-
viously been played at. 

Ethereal Orchestra (G-3) 

Range: 25 feet + 25 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 100 
Base Chance: 35\% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  Creates  magical  accompaniment  for  the 
Adept’s performance. The accompaniment consists 
of the sounds of one instrument plus one per three 
full ranks. The Adept must be familiar with music 
made by their chosen instruments, but need not be 
able  to  play  the  instruments  themselves.  This  ac-
companiment  increases  the  effective  Rank  of  the 
performance by 1 (+ 1 / 5 Ranks) without bestow-
ing  new  skills,  or  affecting  quasi-magical  abilities 
or  base  chances.  The  maximum  volume  generated 
may  not  exceed  that  of  loud  chanting,  and  the 

33 

12 COLLEGE OF BARDIC MAGICS 

accompaniment  can  be  heard  clearly  everywhere 
within the range of the spell. 

Exhortation (G-4) 

Range: 25 feet + 25 / Rank 
Duration: Concentration: no maximum 
Experience Multiple: 200 
Base Chance: 15\% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  By  means  of  this  spell  the  Adept  may 
attempt  to  affect  the  mood  of  a  crowd,  inciting  a 
riot  or  calming  a  mob.  The  reaction  roll  of  the 
crowd is modified by +5\% ( + 1 / Rank). Once the 
Adept ceases to play and/or sing, the spell ceases to 
be in effect, but the effects may continue as deter-
mined by the GM. 

Quietness (G-5) 

Range: 1 foot + 1 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 100 
Base Chance: 30\% 
Resist: None 
Storage: Potion, Investment 
Target: Entity 
Effects:  The  sounds  of  the  target’s  movement  are 
partially deadened. The effect of this is to add 5\% 
(+  1  /  Rank)  to  Stealth.  Any  Entity  attempting  to 
cast  Bardic  magic  while  under  the  effect  of  this 
spell  suffers  a  5  penalty  to  their  Base  Chance  due 
to the deadening of sound. 

Mockery (G-6) 

Range: 15 feet + 15 / Rank 
Duration: 10 seconds + 10 / Rank 
Experience Multiple: 200 
Base Chance: 20\% 
Resist: Active, Passive 
Storage: Investment 
Target: Sentient Entity 
Effects: The Adept sings or orates a song or poem, 
detailing  the  target’s  shortcomings  and  inflicting 
general  abuse,  insults  and  mockery.  A  target  that 
fails  to  resist  may  be  embarrassed,  shocked,  hu-
miliated,  indignant  or  infuriated  as  appropriate  to 
their  personality.  Whatever  the  emotional  effect, 
the  character  is  distracted,  and  may  do  nothing 
other  than  attempt  to  silence  the  Adept  (by  what-
ever means they  wish) or take pass actions for the 
duration  of  the  spell.  A  check  of  1  ×  WP  may  be 
made at the end of the pulse following the pulse in 
which  the  spell  is  cast  and  every  pulse  thereafter. 
Once  the  target  successfully  makes  a  check  the 
spell  ends.  The  target  must  be  able  to  hear  and 
understand the Adept in order for the spell to have 
any effect. 

Shatter (G-7) 

Range: 5 feet + 5 / Rank 
Duration: Immediate 
Experience Multiple: 200 
Base Chance: 25\% 
Resist: Special 
Storage: Investment, Ward, Magical Trap 
Target: Object 
Effects:  The  Adept  shrieks  in  an  unnatural  and 
piercing  fashion,  creating  ethereal  dissonance  that 
can  destroy  objects.  The  Adept  may  affect  an  ob-
ject  of  no  more  than  0.5  pounds  (+  0.5  /  Rank). 
Additionally,  the  Rank  of  this  spell  determines 
what material may be destroyed: 

Rank  Material 

0–3 
4–6 
7–9 
10–12 
13–14 
15–16 
17–18 
19–20 

glass, mirror 
ceramics 
crystals 
stone 
gems 
bone or ivory 
hard metal (iron) 
soft metal (bronze) 

Note  that  the  base  MR  of  objects  is  0.  Crafted 
objects  and  those  made  of  precious  materials  re-
ceive  bonuses  to  their  resistance  rolls,  as  detailed 
below.  Shaped  magical  items  are  immune  to  the 
effects of this spell. Possessions have their owners 
MR. These bonuses are cumulative. Crafting modi-
fier:  2  ×  Artisan  Rank;  material  modifiers:  Silver 
+5\%,  Gold  +10\%,  Truesilver  +15\%,  Precious 
gems,  etc.  +20\%.  The  GM  should  only  apply  the 
modifier of the material that makes up the majority 
of the object. 

Silent Sounds (G-8) 

Range: Touch 
Duration: 10 minutes + 10 minutes / Rank 
Experience Multiple: 100 
Base Chance: 35\% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Object 
Effects:  The  target  object  generates  sounds  which 
are  almost  inaudible,  but  affect  living  entities 
within  5  feet  (+  5  /  Rank)  rendering  them  either 
edgy  and  more  susceptible  to  fear  and  awe,  or 
tranquil  and  less  prone  to  fear  and  awe.  Affected 
beings either add or subtract 1 (+ 1 / Rank) to rolls 
made on the fright or awe tables as appropriate. 

Soothe the Savage Beast (G-9) 

Range: 5 feet + 5 / Rank 
Duration: Concentration: no maximum 
Experience Multiple: 200 
Base Chance: 25\% 
Resist: Passive 
Storage: Potion 
Target: Self 
Effects:  The  Adept  plays  or  sings  soothing  music 
which  causes  any  living  non-sentient  entities 
within or entering the area of  effect to resist or be 
pacified.  Animals  that  do  not  resist  must  make  a 
check against 2 × WP - Rank of Spell to attack the 
Adept, and a check against 2 × WP to initiate any 
attack  whilst  in  the  area.  The  creature’s  reaction 
roll is increased by 10\% (+ 2 / Rank), but the reac-
tion  roll  may  not  be  made  to  exceed  95  by  the 
effects  of  this  spell.  If  any  soothed  entity  is  at-
tacked, or the Adept ceases to play and/or sing, the 
spell ceases to be in effect. 

Speaking Beasts (G-10) 

Range: 15 feet + 15 / Rank 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 200 
Base Chance: 35\% 
Resist: None 
Storage: Investment 
Target: Animal 
Effects: This spell confers any one language known 
to the Adept at Rank 6 or greater, upon an animal 
(beast,  avian,  or  aquatic;  any  non-sentient  able  to 
vocalise  sounds)  for  the  duration  of  the  spell,  at  a 
language Rank equal to 1 (+ 1 / 5 Ranks). 

Ventriloquism (G-10) 

Range: 10 feet + 10 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 150 
Base Chance: 40\% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  The  Adept  may  project their  voice  so  that 
it  appears  to  be  emanating  from  anywhere  within 
the  range  of  the  spell.  Also,  it  may  be  altered  so 
that  it  sounds  like  any  other  voice  or  voices  the 
Adept  has  heard  and  memorised  with  the  Melodic 
Memory  talent.  For  every  five  complete  ranks  the 
Adept can project an extra simultaneous and inde-
pendent voice. 

12.5 General Knowledge Rituals 

Implanting Sounds (Q-1) 

Range: 10 feet + 10 / Rank 
Duration:  1  week  +  1  /  Rank;  Permanent  at  Rank 
20 
Experience Multiple: 250 

Base Chance: 20\% + 4\% / Rank 
Resist: None 
Target: Object or Area 
Cast Time: 1 hour 
Actions: Perform song or sounds 
Concentration Check: None 
Effects: This ritual creates a type of ward, implant-
ing  a  sequence  of  sounds  (which  may  be  words, 
spoken or sung) of up to 10 seconds (+ 10 / Rank) 
duration  into  an  object  or  area.  The  Adept  may 
determine  the  triggering  conditions,  in  the  same 
way as a ward. Additionally, sounds implanted into 
an object may have a range of “Touch” thus being 
triggered by a tactile cue, usually simply the object 
being  touched.  The  sound  will  be  emitted  each 
time  the  triggering  conditions  are  met,  up  to  a 
maximum of 1 (+ 1 / Rank) times. The volume of 
the  sound  produced  may  not  exceed  that  of  loud 
ringing bells. The ritual dissipates at the end of its 
duration or if the Bardic General Knowledge coun-
terspell is cast on the target. If the ritual backfires, 
the  sounds  are  triggered  immediately  and  will 
repeat Rank times, or until counterspelled. 

Recitation (Q-2) 

Range: Touch 
Duration: Special 
Experience Multiple: 200 
Base Chance: 25\% + 4\% / Rank 
Resist: None 
Target: Object, Area or Entity 
Cast Time: 1 hour 
Actions: None 
Concentration Check: Standard 
Effects:  At  the  completion  of  this  ritual  the  Adept 
will enter a trance, during which they  will recite a 
legend or story relating to their current location, to 
an  object  held  in  their  hands,  or  to  an  entity  they 
are  touching.  If  no  such  story  exists,  or  the  ritual 
fails,  then  the  Adept  will  recite  some  amusing but 
trivial  song.  If  the  ritual  backfires  the  legend  will 
be  false.  The  amount  of  information  received  is 
related  to  the  Bard’s  Rank  with  this  ritual.  If  the 
GM  prefers  this  may  be  played  as  obtaining  an 
answer to 3 (+1/3 ranks) short questions related to 
the  location,  object  or  entity  which  will  be  an-
swered  in  “legendary”  terms.  This  ritual  may  not 
be  repeated  on  a  given  target  more  than  once  per 
season. 

12.6 Special Knowledge Spells 

Charming (S-1) 

Range: 50 feet + 5 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 400 
Base Chance: 25\% 
Resist: Active, Passive 
Storage: Investment 
Target: Entity 
Effects: The  Adept can influence the actions of an 
entity  who  fails  to  resist,  provided  that  the  target 
can hear and understand the Adept. The target will 
see  the  Adept  as  their  true  friend  and  will  readily 
accept  most  suggestions  emanating  from  them. 
Sworn enemies of the Adept (or of their race) will 
not be affected by this spell. Any suggestion that is 
not  directly  and  obviously  inimical  to  the  target’s 
interests (as defined by the GM, but usually limited 
to  actions that  would  be  injurious  or  fatal)  will  be 
acted upon  90\%  of  the  time.  The  Adept  may  only 
ensure 100\% compliance with a request by making 
it  an  order  in  which  case  the  target  makes  an  im-
mediate  Resistance  Check.  Any  suggestion  that 
would lead to the target’s injury or death results in 
an  immediate  Resistance  Check.  If  successful  the 
spell  is  broken.  Otherwise,  the  target  accepts  the 
order  and  will  carry  it  out  faithfully  for  the  dura-
tion of the spell. 

Compel Speech (S-2) 

Range: Touch 
Duration: 1 minute + 1 / Rank 
Experience Multiple: 200 
Base Chance: 25\% 
Resist: Active, Passive 

34 

Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  The  target  entity  must  resist,  or  be  com-
pelled  to  speak  continuously  on  random  subjects. 
Should  anyone  put  a  question  to  the  entity  they 
must make a 3 × WP  - Spell Rank check or speak 
to  the  question.  The  entity  is  not  compelled  to 
speak the truth by this spell, or any  particular lan-
guage, and may cast spells. 

Comprehending Tongues (S-3) 

Range: Self 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 300 
Base Chance: 15\% 
Resist: None 
Storage: None 
Target: Self 
Effects: Enables the Adept to speak and understand 
one language at a Rank equal to 1 (+ 1 / 4 Ranks). 
The Adept must have heard the language that they 
wish to speak. The language may be nominated by 
name, or by choosing to speak a language that the 
Adept  has  memorised  some  words  and  phrases 
from  using  the  Talent  of  Melodic  Memory.  This 
spell  allows  the  Adept  to  be  understood,  but  con-
fers no other bonuses. 

Confusion of Tongues (S-4) 

Range: 15 feet + 15 / Rank 
Duration: 10 seconds + 10 / Rank 
Experience Multiple: 350 
Base Chance: 10\% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Area 
Effects:  All  entities  within  the  radius  of  effect  of 
the spell  who fail to  resist become unable to com-
municate  in  or  understand  any  verbal  language 
whilst within the affected area. The affected area is 
a  circle  with  a  radius  of  5  feet  (+  5  /  Rank).  Af-
fected  entities  may  not  use  the  Military  Scientist 
“time out” ability. They can cast spells and trigger 
items,  but  if  casting,  the  backfire  threshold  (i.e. 
+30 or +40) is decreased by 5\% (+ 1 / Rank). 

Dance of Swords (S-5) 

Range: 5 feet + 5 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 250 
Base Chance: 15\% 
Resist: None 
Storage: Investment, Potion 
Target: Entity 
Effects:  While  under  the  effect  of  this  spell,  the 
target  may  cavort  and  leap  with  surpassing  grace 
and  extravagance  by  evoking  the  magic  of  the 
dance.  The  target  may  walk  and  act  in  all  ways 
normally,  however,  when  they  enter  combat  they 
may “dance” adding 2 (+2 / Rank) to their Defence 
provided they move at least 1 hex each pulse. If the 
target  is  unwillingly  confined  to  a  single  hex  (by 
the  effect  of  melee  zones  for  example)  the  target 
must  halve  the  defence  bonus  due  to  this  spell.  If 
the target is unable to move freely (in close combat 
or stunned) then no defence bonus is awarded. The 
target may also subtract the Rank of this spell from 
the  dice  roll  for  any  AG  Check  solely  involving 
Agility (e.g. avoid knockdown, leaping pits etc.). 

Enthralling Audience (S-6) 

Range: 15 feet + 15 / Rank 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 200 
Base Chance: 25\% 
Resist: Special 
Storage: Invested, Ward 
Target: Sentient living entities 
Effects: All entities who willingly listen to the song 
sung  by  the  Adept  will  perceive  a  succession  of 
evocative  images  which  illustrate  the  song  being 
sung.  These  images  are  hallucinatory  in  nature, 
having  no  external  reality,  but  are  seen  by  the 
audience. No resistance check needs to be made by 
those who do not submit willingly to the effects of 
the  spell,  but  they  will  not  perceive  the  images. 

Those  who submit to the effects will be enthralled 
by the images and may take no other action, unless 
and until their concentration is broken by physical 
contact with an external entity or force. The Adept 
must continue to sing throughout the duration. 

Shout of Thunder (S-7) 

Range: 15 feet + 15 / Rank 
Duration: Immediate 
Experience Multiple: 250 
Base Chance: 25\% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  The  Adept  projects  a  thunderous  shout  of 
rage at the target. If the target fails to resist it suf-
fers  [D  -  5]  (+  1 / Rank) damage  and  is automati-
cally stunned. The target will also be deafened for 
[D  -  5]  (+  1  /  2  Ranks)  minutes  and  will  have  a 
very  nasty  headache  which  will  make  concentra-
tion one level more difficult until it wears off or is 
cured by a healer. There is a chance equal to 5\% + 
1 / Rank that this deafness will be permanent until 
cured by the arts of a Healer of Rank 4 or greater, 
or  by  the  Ritual  of  Cure  Deafness  and  Muteness. 
The  sound  is  clearly  audible  in  the  area  surround-
ing the target. 

Inspirational Song (S-8) 

Range: 15 feet + 5 / Rank 
Duration: 20 seconds + 5 / Rank 
Experience Multiple: 300 
Base Chance: 15\% 
Resist: None 
Storage: None 
Target: Sentient Entities 
Effects:  The  Adept  sings  a  song  of  inspiration 
mentioning the name of each target. The spell will 
affect  1  target  (+  1/3  or  fraction  Ranks)  who  each 
gain 1 (+ 1/2 Ranks) to their Strike Chances, and 1 
(+  1  /  Rank)  to  Fear  Resistance  rolls  and  Rally 
Chances.  Targets  must  be  able  to  hear  and  under-
stand  the  Bard.  The  Adept  must  sing  or  orate  for 
the entire duration of the spell, or the effect ceases. 
The  Adept  may  perform  other  actions  as  long  as 
they have no vocal requirement. 

Satyr’s Dance (S-9) 

Range: 5 feet + 5 / Rank 
Duration:  Concentration:  Maximum  10  seconds  + 
10 / Rank 
Experience Multiple: 450 
Base Chance: 10\% 
Resist: Active, Passive 
Storage: Potion 
Target: Self 
Effects: The  Adept sings and dances in a hypnotic 
pattern,  entrancing  all  in  range  who  fail  to  resist, 
and  forcing  them to  join  in the  dance.  All  dancers 
(including  the  Adept)  may  move  at  1/2  TMR 
(round up), but are unable to take any other action. 
The  area  of  effect  moves  with  the  Adept  and  all 
victims  will  attempt  to  stay  within  range.  Any 
entity  that  enters  the  area  of  effect  must  resist  or 
suffer  the  same  fate.  The  Adept  must  be  seen  and 
heard  for  a  target  to  be  effected.  Victims  of  this 
song  get  an  additional  resistance  roll  every  pulse 
that  they  are  attacked  or  restrained,  or  somehow 
prevented from being within the area of effect. 

Shell of Silence (S-10) 

Range: 10 feet + 1 / Rank 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 150 
Base Chance: 15\% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Area 
Effects:  The  Adept  creates  an  invisible  spherical 
shell  with a diameter equal to 15 feet (+ 5 feet for 
every  5  Ranks).  Sound  created  within  this  shell  is 
totally  inaudible  to  entities  outside  the  shell  and 
vice  versa.  A  Clairaudience  spell  projected  from 
outside the shell cannot penetrate it. 

Silver Tongue (S-11) 

Range: Self 

Duration: 10 minutes + 10 / Rank 
Experience Multiple: 200 
Base Chance: 25\% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  Everything  that  the  Adept  says  becomes 
more  convincing  and  believable.  This  ability  does 
not  force  listeners  to  believe  the  Adept,  and  obvi-
ous lies where there is evidence to the contrary will 
be  quickly  dismissed.  But  in  the  absence  of  such 
evidence,  or  when  the  lie  is  not  blatant,  listeners 
will tend to take the Adept at their word. All natu-
ral  or  magical  abilities  that  are  normally  able  to 
determine whether the Adept is lying or telling the 
truth  suffer  a  negative  modifier  to  their  Base 
Chance  of  20\%  (+  5  /  Rank).  Those  abilities  that 
always  work  or  have  no  BC  may  be  presumed  to 
have a BC of 100\% for these purposes. In addition, 
no  magical  ability  is  able  to  coerce  the  Adept  to 
speak truthfully. 

Siren Song (S-12) 

Range: 5 feet + 5 / Rank 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 250 
Base Chance: 15\% 
Resist: Active, Passive 
Storage: None 
Target: Entity 
Effects:  All  entities  within  the  range  of  the  spell 
and  able  to  hear  the  Adept,  must  resist  or  feel  a 
sudden  affection  for  the  Adept.  They  will  then  be 
unable to carry out any action that might harm the 
Adept. However, if the  Adept carries out any hos-
tile  action  towards  an  entity  or  the  entity  is  ren-
dered  no  longer  able  to  hear,  that  entity  will  be 
released  from  the  spell.  The  Adept  must  continue 
to sing for the duration or the spell ceases to be in 
effect. 

Slumber Song (S-13) 

Range: 5 feet + 5 / Rank 
Duration: Until waking 
Experience Multiple: 250 
Base Chance: 30\% 
Resist: Special 
Storage: Investment 
Target: Entity 
Effects:  The  Adept  plays  a  song  that  affects  one 
willing  entity  (+  1  /  3  or  fraction  ranks)  causing 
them to  drift into  a  normal  sleep.  All  affected  tar-
gets  fall  asleep  in  4  minutes  (10  seconds  /  Rank, 
minimum  of  30  seconds).  The  targets  will  remain 
asleep  until  disturbed  or  they  awaken  normally. 
The resulting sleep is healing and refreshing allow-
ing the targets to regain 1 (+ 1 / 5 full Ranks) more 
FT per hour while under its effect. 

Whispering World (S-14) 

Range: 100 miles + 100 / Rank 
Duration: Immediate 
Experience Multiple: 250 
Base Chance: 20\% 
Resist: None 
Storage: potion 
Target: Self 
Effects:  The  Adept  whispers  a  message  consisting 
of no more than 5 words (+ 5 / Rank). The message 
travels to the recipient, who must be known to the 
Adept,  taking  D10  +  (distance  travelled  /  100) 
hours.  The  recipient’s  surroundings  (such  as trees, 
waves or  wind) whisper the message to the recipi-
ent.  There  is  a  once  times  perception  chance  that 
any Bard engaged in concentration will “overhear” 
a whisper which passes by their location. 

12.7 Special Knowledge Rituals  

Cure Deafness and Muteness (R-1)  

Range: Touch 
Duration: Immediate 
Experience Multiple: 200 
Base Chance: 25\% + 4\% / Rank 
Resist: None 
Target: Entity 
Cast Time: 1 hour 
Material: None 
Actions: Singing 
Concentration Check: Standard 
Effects:  The  target  is  cured  of  deafness  or  mute-
ness,  whether  of  natural  or  magical  origin.  Only 
those born deaf or mute cannot be affected by this 
ritual. 

Resounding Instrument (R-2) 

Range: Touch 
Duration: 4 weeks + 2 / Rank 
Experience Multiple: 250 
Base Chance: MA + 4\% / Rank 
Resist: None 
Target: Musical Instrument 
Cast Time: 6 Hours 
Material: Instrument 
Actions: Playing instrument  
Concentration Check: None 
Effects:  This  ritual  enchants  a  instrument,  which 
must  be  of  religious  or  martial  nature,  such  as 
trumpets,  horns,  bagpipes,  drums,  bells,  or  gongs. 
The  enchanted  instrument  can  be  heard  at  its  nor-
mal volume (usually loud) throughout an area with 
a  radius  equal  to  1  mile  (+  1  /  Rank).  The  effect 
may  be  made  permanent  if  the  Adept  chooses  to 
permanently  expend  a  point  of  Endurance.  If  the 
ritual backfires the instrument will be destroyed in 
addition to the normal backfire effect. 

Sound of Doom (R-3) 

Range: Sight 
Duration: Immediate 
Experience Multiple: 400 
Base Chance: MA + 3\% / Rank 
Resist: None 
Target: Structure 
Cast Time: Special 
Material: Musical Instrument 
Actions: Playing trumpet and walking 
Concentration Check: Standard 
Effects:  The  Adept  marches  around  the  structure, 
within  earshot  of  the  walls,  playing  a  musical  in-
strument.  The  music  must  be  able  to  be  clearly 
heard  at  the  structure.  The  size  of  structure  which 
may  be  encompassed  is  50  feet  (+  50  /  Rank)  in 
diameter.  The  structure  begins  to  shake  and  vi-
brate, and at the end of the ritual, if it is successful, 
the  structure  falls  apart.  The  Adept  must  walk 
slowly  (1  mph)  around  the  target  until  they  have 
completely  encircled  the  target  at  least  once,  and 
have marched for at least an hour. 

The Piper’s Song (R-4) 

Range: 30 feet + 30 / Rank 
Experience Multiple: 400 
Base Chance: MA + 3\% / Rank 
Resist: None 
Target: Self 
Cast Time: 1 hour 
Material: Instrument (usually pipe) 
Actions: Playing instrument and walking 
Concentration Check: Standard 
Effects:  The  Adept  nominates  one  type  of  non-
sentient  entity  which  normally  forms  swarms, 
packs  or  herds  (e.g.  rats,  locusts,  wolves,  ele-
phants)  at  the  start  of  this  ritual.  The  size  of  the 
entity which can be affected is dependent on Rank: 
Rank  0–5,  entities  less  than  1lb  in  weight  can  be 
affected;  Rank  6–10,  entities  less  than  10lb;  Rank 
11–15,  entities  less  than  100  lb;  Rank  16  and 
above,  entities  greater  than 100lb.  The  Adept  then 
begins to play a tune which has a compelling effect 
on all entities of the target type. The area of effect 
moves  with the Adept, and as they play and walk, 
all  of  these  entities  within  range  will  begin  to  fol-
low the Adept, growing into a horde. At the end of 
at  least  1  hour  the  Adept  gives  a  single  command 
to  the  horde.  The  wording  of  this  command  may 
not  exceed  1  word  (+  1/3  or  fraction  Ranks).  The 
command will be obeyed for a period of 1 hour (+1 
/ Rank). 

\end{Chapter}
