\begin{Chapter}{Beast Master (Ver 1.1)}

A  loyal  animal  or  monster  is  likely  to  serve  its 
master  far  better  than  adventuring  comrades  ever 
will.  A  beast  master  is  one  who  trains  these  crea-
tures  to  obedience.  They  take  a  wild  animal  and, 
from an adversarial relationship, develop a rapport 
with  it.  They  train  young  animals  from  birth  until 
they  heed  their  every  command.  A  beast  master 
will, in almost all cases, become very fond of ani-
mals.  They  will  defend  them  against  wanton  cru-
elty  and  slaughter,  and  will  treat  their  personal 
charges as family. 

A  beast  master  will  encounter  three  kinds  of  ani-
mals:  the  easily  domesticated  (such  as  the  horse), 
the  naturally  wild  (such  as  the  pegasus)  and  the 
intelligent  or  rebellious  creature  (such  as  the  uni-
corn).  The  latter  can  never  be  steadfastly  loyal  to 
the  beast  master;  such  creatures  always  have  at 
least  a  subconscious  desire  to  escape.  A  beast 
master can be a slaver if they specialise in training 
humanoids. 

35.1 Restrictions 
A beast master must have at least 15 Willpower. 

A beast master will normally use their skill to train 
or domesticate animals for their own use. Animals 
which spend their lives with a beast master and are 
trained  by  them  will  be  loyal  to  their  master  and 
serve  and  protect  them  as  much  as  possible.  If 
necessary, an animal can be trained to temporarily 
serve another master (if one week of mutual train-
ing is undergone), but the animal will always obey 
the original master before any new one. 

If  a  beast  master’s  Rank  is  5  or  greater,  they  may 
train  animals  for  other  people.  The  being  who  is 
acquiring  the  trained  creature  must  spend  (12  - 
Rank)  weeks  before  it  will  accept  them as  its new 
master,  during  which  the  beast  master  must  be 
present at least one day per week. The creature will 
heed the beast master’s commands before those of 
its new owner for as many years as the beast mas-
ter’s  Rank  at  the  time  the  creature’s  ownership  is 
transferred. 

A  beast  master  of  any  Rank  may  domesticate, 
rather  than  train,  animals.  Such  animals  can  be 
commanded  by  any  other  person,  but  will  tend  to 
wander off or revert to their wild state if not super-

vised,  tied  up,  or  stabled.  Note  that  horses  and 
dogs,  the  most  common  domestic  animals,  are 
governed by this rule. 

35.2 Benefits 

A beast master acquires the ability to train one 
type of animal or monster at Ranks 0, 5 and 10. 

A  beast  master  may  acquire  the  ability  to  train 
additional  types  of  creatures  without  increasing  in 
rank  by  spending  5,000  Experience  Points  and  4 
weeks training per type. These costs are discounted 
by  25%  if  the  beast  master  has  reached  rank 8,  or 
by 50% if they have reached rank 10. 

A  type  consists  of  all  creatures  listed  within  one 
subsection  of  the  beastiary  (e.g.  avians).  A  beast 
master may choose, instead, all creatures subsumed 
under a single animal family (e.g. canines). 

A  beast  master  must  spend  (12  -  Rank)  months  to 
train  an  animal  or  monster,  or  a  like  number  of 
weeks to domesticate one. 

Creature to be trained is 

Time 

Easily domesticated 
× 0.5 
Naturally wild 
× 1.0 
Intelligent or rebellious 
 × 3.0 
Raised by beast master from adolescence  × 0.5 
× 1.0 
Domesticated by another beast master 
Caught in wilderness 
× 1.5 
The  unmodified  number  of  months  required  is 
multiplied by all applicable modifiers. The time to 
train  a  monster  or  animal  is  always  dependent  on 
the beast master’s Rank when they begin the proc-
ess.  Any  increases  in  Rank  during  the  training  or 
domestication  period  have  no  effect  on  the  time 
required. 

Loyalty Checks 

A  trained  animal  or  monster  must  make  a  loyalty 
check  whenever  it  recognises  that  its  master  is 
endangering  it,  or  whenever  its  master  commands 
an  action  that  runs  counter  to  its  instincts.  When-
ever a loyalty check is required, the GM rolls per-
centile dice. The base chance is 2 × beast master’s 
WP  +  (4  ×  Rank  if  the  creature  is  intelligent  or 
rebellious,  6  ×  Rank  if  the  creature  is  naturally 
wild, and 8 × Rank if the creature is easily domes-
ticated). If the owner is not a beast master, use their 

35 BEAST MASTER 

WP  and  the  Rank  of  the  beast  master  when  they 
trained the creature. If the roll is less than or equal 
to this success percentage, the trained creature will 
do  as  its  master  commands.  If  the  roll  is  greater 
than  the  success  percentage,  the  creature’s  reac-
tions  will  range  from balking  to  fleeing  to  turning 
on  its  master,  as  the  roll  increases  (GM’s  discre-
tion). 

A domesticated creature must make a loyalty check 
if  the  circumstances  described  above  arise.  The 
GM rolls D100. If the resulting number is less than 
or  equal  to  current  master’s  WP  +  beast  master’s 
Rank,  the  domesticated  creature  will  perform  the 
action.  If  the  roll  is  greater  than  the  success  per-
centage,  but  less  than  or  equal  to  twice  that  per-
centage, the creature will balk. If the roll is greater 
than  two  times  the  success  percentage,  but  less 
than  three  times  that  percentage,  the  creature  will 
take flight. If the roll is greater than three times the 
success  percentage,  the  creature  will  turn  on  its 
master.  A  roll  of  100  always  indicates  that  a  do-
mesticated creature turns on its master. A roll of 96 
through 99 indicates that the creature takes flight if 
the success percentage is 47 or greater. 

A beast master who intimidates their animals adds 
one  to  their  Rank  when  calculating  training  or 
domestication  time,  but  the  GM  adds  10  to  any 
loyalty check dice-roll for one of their animals. 

A  beast  master  may  train  or  domesticate  as 
many creatures as their Rank at one time. 

All creatures being trained or domesticated concur-
rently must be of the same type. 

35.3 Cost 

A beast master must pay 100 Silver Pennies per 
creature trained and 25 Silver Pennies per crea-
ture domesticated. 

They may halve the cost for upkeep of creatures if 
they build a stable. A horse-sized stable costs (500 
+  150  ×  Stalls)  Silver  Pennies  to  construct,  and 
costs  (Stalls)  Silver  Pennies  for  repairs  after  the 
first year. 
\end{Chapter}
