\begin{Chapter}{Cantrips}

Cantrips  are  minor  magic  effects,  which  come  in 
two  types.  Glamour  cantrips  are  used  to  entertain 
and Household cantrips make life pleasanter. Most 
cantrip effects can also be reversed. 

\section{Restrictions}

The effects of cantrips are strictly minor in nature and cannot be
used during combat. Any mage can cast them and no preparation is
needed.  It costs 1 Fatigue to cast a cantrip.

\section{Modifiers \& Statistics}

All minor magics are non-college magic.  College bonuses and penalties
do not apply to them.  All Adepts will start off knowing the following
cantrips but they may not be ranked or improved, and do not affect the
number of low-ranked spells that may be learned.

Base Chance: 3 × MA 

Double or triple effects are not possible and should the roll exceed 4
× MA the cantrip will backfire.  Backfires result only in the effect
being distinctly different from the casters intentions.  Actively
resisting a cantrip results in it automatically failing.

Any stronger magic of the same type as a cantrip will immediately
replace it and any area counter- spell will dispel all cantrips within
range. Where an Adept is casting a cantrip that has effects similar or
related to their college then the effects will be stronger, or more
convincing than normal although a cantrip can never be as effective as
any spell.

\section{Glamour Cantrips}

Glamour cantrips normally last (Base Chance) minutes.  The Adept may
choose to reduce the duration.

Colour  The  surface  colour  of  one  object  may  be 
changed.  If  cast  on  an  entity  (e.g.  skin,  hair,  or 
eyes) the result is flat and not particularly life like. 

Perfume  A  faint  pleasant  (or  unpleasant)  scent 
permeates  one  hex.  Observers  will  not  notice  any 
specific  odour,  just  that  something  smells  nice  (or 
bad). 

Polish/Tarnish  One  entity  or  object is  spruced  up 
so  they  look  smartly  dressed  or  sparkling  clean. 
This doesn’t make something look more expensive 
merely newer and shinier. 

Fireworks  The  Adept  may  combine  any  two  of 
three  effects  in  a  1  foot  sphere  which  appears 
within  10  feet.  The  effects  are  coloured  smoke, 
coloured lights, and a small sharp noise (e.g. bang, 
tinkle, fizzle). 

8.4 Household Cantrips 
Household  cantrips  normally  have  an  immediate 
effect and no duration. 

Cool/Warm  Up  to  a  pint  of  liquid,  or  a  pound  of 
food,  may  be  cooled  or  warmed  a  few  degrees. 
This  might  cool  hot  food  to  an  edible  temperature 
or  warm  milk,  but  not  to  boiling  or  freezing  ex-
tremes. 

Clean/Dirty  Surface  dirt  and  grime  may  be  shed 
from  an  entity  or  object  to  fall  around  them.  If 
reversed  all  the  loose  dirt  within  one  hex  attaches 
to  the  target.  Several  casts  may  be  required  to 
removed  thick  mud  or  dirt  and  it  will  not  remove 
old stains. 

Dry/Dampen  The  cantrip  will  render  up  to  one 
item  of  clothing  (or  an  object  up  to  3  pounds) 
either dry to the touch or damp. Several casts may 
be required to dry sodden items or cause the target 
to drip water. 

Insect  Repellent  Small  insects  within  one  hex,  or 
on one entity leave as fast as they can. This effect 
lasts (Base Chance) minutes. 

Tie/Untie  causes  strands  to  knit  together,  or  fray 
apart. This can be used to knot string, repair minor 
rips  in  fabric,  repair  fraying  rope,  or  tangle  or 
untangle  hair  etc.  It  will  not  fill  in  holes  or  seam-
lessly join completely separate things together. 

Candle A small natural flame appears on an object 
and  lasts 5  seconds.  If  cast  on  something  flamma-
ble  the  object  may  catch  alight  (e.g.  wick,  kin-
dling). The flame will shed light about 2 feet. 

\end{Chapter}
