\begin{Tables}{Combat Tables}

\begin{multicols*}{2}

%% \raggedcolumns
  
\section{Strike Chance Modifiers Summary}
\label{table:strikechance}
\smallskip

\begin{tblr}{
    rowsep=0.0pt,
%%    stretch=0.9,
    rows={font=\fontsize{10}{10pt}\selectfont},
    colspec={rX},
    cell{1,7,19,25,38}{1}={c=2}{l,font=\bfseries},
    hlines={green7},
    hline{1-2,7-8,19-20,25-26,38-39}={1.2pt,green7}
    }
Close Combat Modifiers \\
+1 	& each point attacker’s PS is greater than target’s PS \\
+20	& target has no Fatigue \\
+20	& target is stunned \\
-1	& each point target’s PS is greater than attacker’s PS \\
-20	& attacker has no Fatigue \\
Melee Combat Modifiers \\
+10	& target has no Fatigue \\
+10	& target is being attacked though a flank hex \\
+15	& target is stunned \\
+20	& target is kneeling or prone \\
+20	& attacker is charging with pole weapon or shield \\
+30	& target is being attacked through rear hex \\
-4	& each Rank the target has with prepared weapon if evading \\
-10	& target is evading \\
-10	& attacker has no Fatigue \\
-15	& attacker is charging with non-pole weapon \\
-20	& attack is Melee attacking while withdrawing \\
Visibility Condition Modifiers \\
-10	& starry night or shadowy interior \\
-20	& cloudy night \\
-30	& cave or unlit interior \\
-40	& pitch blackness \\
-50	& target is invisible or undetectable \\
Ranged Combat Modifiers \\
+10	& target being attacked through a flank hex \\
+10	& target is stunned \\
+10	& attacker is kneeling \\
+20	& target is being attacked through rear hex \\
+20	& aim (also chances of Endurance or Specific Grievous damage are increased to 10\% and 20\%, see \S\ref{combat:attacking}) \\
-3	& every hex through which a thrown weapon travels \\
-3	& each 5 hexes (or fraction) after the first 5 through which a missile travels \\
-15	& snapshoot \\
-5	& target is currently moving \\
-10	& target is kneeling or prone \\
-20	& target is evading \\
-20	& target occupies a sheltered hex \\
Miscellaneous Modifiers \\
-20	& striking weapon held in attacker’s secondary hand \\
-10	& multiple strike; attack with weapon in primary hand \\
-30	& multiple strike; attack with weapon in secondary hand \\
-10	& multiple strike; attack with each weapon if ambidextrous \\
-20	& multiple strike with B class two-handed weapon \\
-20	& attacking an airborne figure \\
-15	& airborne figure attacking ground figure \\
-10	& airborne figure attacking another avian \\
\end{tblr}
\smallskip

Each modifier is added to the Strike Chance of the attacker in each
instance where it applies; all modifications are cumulative.

\vfill

\section{Combat Equation Summary}

\smallskip

\begin{tblr}{
    rowsep=0.0pt,
%%    stretch=0.9,
    rows={font=\fontsize{10}{10pt}\selectfont},
    colspec={m{8em}X},
    hlines={green7},
    }
Engaged Initiative			& PC + modified AG + weapon Rank + warrior Rank. \\
Unengaged Initiative			& D10 + (PC + 2 × Military Scientist). \\
Strike Chance with ranked weapon	& Weapon Base Chance + attacker’s modified Manual Dexterity + (4 × Rank with weapon) + Magic. \\
Modified Strike Chance			& Strike Chance + Modifiers (\S\ref{table:strikechance}) \mi opponent’s Defence. \\
Defence					& Modified AG + Shield defence + Modifiers (\S\ref{table:strikechance}) + Magic. \\
Repulse a charge attack			& D10 ≥ prepared weapon Rank.\\
Withdraw from close combat		& (D10 + total friendly Physical Strength \mi total hostile Physical Strength) ≥ 10. \\
Strike chance to trip			& 40\% + attacker’s modified Manual Dexterity + (4 × Rank with weapon) \mi opponent’s Defence (Damage: D10). \\
Restrain				& 3 × ((PS+AG of attacker) − (PS+AG of defender)). \\
Shield Rush				& 40\% + attacker’s modified MD + (4 × Rank with shield) − opponent’s Defence (Damage: [D − 2]). \\
Disarm					& −20 to strike chance. \\
Entangle				& Same as normal strike chance with weapon (Damage: [D \mi 4]). \\
Knockout				& Must roll under (15\% × Modified Strike Chance). \\
Fumble					& Reduce Initiative by 10. Roll under Initiative on D100 or roll on the Fumble Table (\S\ref{table:fumble}). \\
Avoid Disarm				& MD + Rank on D100. Chance doubled if two handed.\\
Stun recovery				& (2 × WP) + current Fatigue. \\
Parry result				& D10 + evader’s Rank − attacker’s Rank. \\
					& 3 or less: Successful parry; evader must pass next action. \\
					& 4--7: Disarm, 1 EN damage. \\
					& 8 or greater: Disarm plus a riposte; evader may melee attack, 1 EN damage. \\
\end{tblr}
\end{multicols*}

\begin{multicols}{3}

\section{Action Summary}
\label{table:actionsummary}
\subsubsection{Engaged figures}
\begin{Itemize}
\item Melee Attack  
\item Close and Grapple  
\item Evade  
\item Offensive Withdraw  
\item Defensive Withdraw  
\item Flee  
\item Pass  
\item Cast  
\item Throw  
\item Recover from Stun 
\end{Itemize}

\subsubsection{Close figures}

\begin{Itemize}
\item Grapple  
\item Withdraw  
\item Pass  
\item Recover from Stun 
\end{Itemize}

\subsubsection{Unengaged figures}

\begin{Itemize}
\item Move up to full TMR  
\item Step and Melee Attack  
\item Charge  
\item Charge with Polearm or Shield  
\item Charge and Close  
\item Evade  
\item Retreat  
\item Pass  
\item Cast  
\item Throw  
\item Fire  
\item Recover from Stun 
\end{Itemize}

\raggedcolumns

\section{Special Damage Chart}

\begin{dqtblr}{colspec={XXX}}
Success		& Triple	& Double \\
01 -- 09		& --		& 01 \\
10 -- 16		& 01		& 01 -- 02 \\
17 -- 23		& 01		& 01 -- 03 \\
24 -- 29		& 01		& 01 -- 04 \\
30 -- 36		& 01 -- 02	& 01 -- 05 \\
37 -- 43		& 01 -- 02	& 01 -- 06 \\
44 -- 49		& 01 -- 02	& 01 -- 07 \\
50 -- 56		& 01 -- 03	& 01 -- 08 \\
57 -- 63		& 01 -- 03	& 01 -- 09 \\
64 -- 69		& 01 -- 03	& 01 -- 10 \\
70 -- 76		& 01 -- 04	& 01 -- 11 \\
77 -- 83		& 01 -- 04	& 01 -- 12 \\
84 -- 89		& 01 -- 04	& 01 -- 13 \\
90 -- 96		& 01 -- 05	& 01 -- 14 \\
97 -- 103	& 01 -- 05	& 01 -- 15 \\
104 -- 109	& 01 -- 05	& 01 -- 16 \\
110 -- 116	& 01 -- 06	& 01 -- 17 \\
117 -- 123	& 01 -- 06	& 01 -- 18 \\
124 -- 129	& 01 -- 06	& 01 -- 19 \\
130+		& 01 -- 07	& 01 -- 20 \\
\end{dqtblr}

\end{multicols}
\end{Tables}
