\begin{Chapter}{The College of Earth Magics (Ver 1.2)}

The College of Earth Magics is concerned with the shaping of the
powers of the earth itself and of those entities and things that are
rooted in the earth or in contact with it.

There are two distinct branches of this College, and a practitioner of
Earth Magic college can be a member of only one branch at a time.
They may alter their affiliation, but are treated as if they had quit
their original branch and lost all knowledge of Earth Magic until such
time as they has spent six months in study and meditation to relearn
the General Knowledge of the College in a new form. The two divisions
of this College are:
\begin{Description}
\item[Pacifistic Earth Magic] Usually practiced by wild people who
  live in isolation in the wilderness, this type of Earth Magic is
  very powerful, but entirely defensive in nature. An adherent of this
  division of Earth Magic will never attack without provocation (i.e.
  unless attacked themselves or unless they see animals or plants
  attacked). They are unaffected by counterspells of their College
  cast over an area they which occupy while attempting to work the
  magic of their College.  In addition, the counterspells of their
  College give only half the normal benefit to characters attempting
  to resist their magic.  However, they may never participate in
  rituals of this College which involve human sacrifice.

\item[Druidic Earth Magic] This form of Earth Magic is practiced by
  strong-willed individuals who have no objection to the taking of
  human life to further their magic.  It is often practiced communally
  since it is in part ritual magic.  Many of the rituals involve
  courting the darker sides of earth by providing blood to quench its
  thirst.
\end{Description}

\section{Restrictions}

Practitioners of the College of Earth Magics must always be in contact
with the earth to perform magic of this College.

A practitioner of this College is considered in contact with the earth
if they are in contact with an item which is, itself, rooted in the
earth (tree, plant, foundation of a house, etc.). They would not be in
contact with earth if they were in the air or in water where they
could not touch bottom, or if they were atop a piece of furniture or
an animal as these are not rooted in the ground.  An Adept in a
person’s arms is not in contact with the earth.

The MA requirement of this College is 15. 


\section{Base Chance Modifiers}

The following numbers are added to the Base Chance of performing any
talent, spell or ritual of the College of Earth Magics.

\begin{tabularx}{\columnwidth}{Xr}
Adept is wearing a sprig of fresh mistletoe & +5 \\
Adept is in contact with earth, but beneath water† & -20 \\
Adept is in a manmade dwelling without an earthen floor & -5 \\
Adept occupies a place of power‡	& +20 \\
\end{tabularx}

†For example, beneath the ocean, if the Adept can find a way of
breathing in such a situation, the Adept’s BC would always be reduced
by 20.

‡These can be any places frequented by worshippers of earth. Examples
from mythology and litera- ture might include: Stonehenge, Finn
McCool’s Seat, The Hill of Tara, etc. If the Place of Power is used
for ritual sacrifice, the practitioners of Pacifistic Earth Magic
would receive no benefit.

All modifiers are cumulative. They are in addition to the modifiers
listed in §7.4.

\section{Talents}

\begin{talent}[T-1]{Detect Aura}

\range{Special}
\multiple{75}
\basechance{PC + 5\% / Rank}
\resist{Active}
\target{Entity, Object, Area, Volume}
\begin{effects}
The Base Chance is reduced by 1\% for every foot after the first five
feet the target is from the Adept. The results of this talent are the
same as given in §9.1.
\end{effects}
\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Converse with Animals}

\range{10 feet + 10 / Rank}
\duration{3 hours + 3 hours / Rank}
\multiple{50}
\basechance{45\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept may communicate with fauna (whether verbally or
symbolically, and to what extent, are left up to the GM’s discretion).
Physical contact between the animal and the Adept increases the Base
Chance of successfully casting this spell by 5.  This spell does not
include avians or fish and has to be recast for each type of animal,
for example talking to wolves does not allow the Adept to talk to
tigers.

\end{effects}
\end{spell}

\begin{spell}[G-2]{Converse with Plants}

\range{10 feet + 10 / Rank}
\duration{3 hours + 3 / Rank}
\multiple{50}
\basechance{25\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept can communicate with flora with which they are familiar. The
mode and extent of communication is up to the GM’s discretion.

\end{effects}
\end{spell}

\begin{spell}[G-3]{Controlling Animals}

\range{10 feet + 10 / Rank}
\duration{Concentration: no maximum}
\multiple{100}
\basechance{20\%}
\resist{Passive}
\storage{Investment}
\target{Animal}
\begin{effects}
The Adept controls the actions of any animal that does not
successfully resist.  It will serve them so long as they continue to
concentrate.  If they release the animal or their concentration is
broken, it may attack them or flee.  The chance to cast this spell is
reduced by 5\% if the Adept cannot speak to the animal. If the Adept
cannot make eye contact, the Base Chance is reduced another 5\%.

\end{effects}
\end{spell}

\begin{spell}[G-4]{Blending}

\range{15 feet + 1 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{50}
\basechance{60\%}
\resist{None}
\storage{Investment, Potion, Ward}
\target{Entity}
\begin{effects}
Once this spell is cast, the target must remain still in order for it
to have effect.  While remaining still, the target is not able to be
seen by non-magical means (i.e.  as for invisibility).  If the target
moves, the spell ceases to work. However, if the target becomes still
again during the duration of the spell, it will resume its effect.
The duration of the spell refers to the time since casting, not the
time that the spell is actually in effect (i.e.  while the target is
still).

Keeping still will require (as a minimum) a 4 × WP check every hour.
The target may be required to make additional willpower checks at the
GM’s discretion.

The spell only has effect while the target is “still”.  This means
that the target is unable to move any of its external surfaces, with
limited exceptions. Moving an external surface encompasses such
actions as moving a limb, or opening and closing the mouth.  Blinking
and normal bodily movement caused by normal breathing will not
constitute moving for the purposes of this spell.  The following
actions will automatically cause the spell to cease working: talking,
spellcasting, triggering (subject to any revision of the Investment
ritual), readying a weapon, altering facing in a hex, moving from the
hex, using a silent language, or indeed any Action other than a Pass
action (and Pass actions being restricted as outlined).  Note that it
is not relevant if an observer would see the move- ment for the spell
to cease working (e.g.  moving hands behind back, or talking behind
hand will both cause the spell to cease having an effect).
\end{effects}
\end{spell}

\begin{spell}[G-5]{Walking Unseen}

\range{1 foot + 1 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{100}
\basechance{50\%}
\resist{None}
\storage{Investment, Potion, Ward}
\target{Entity}
\begin{effects}
The target of this spell may move unnoticed, not invisible.  This
means that it will not transmit light. As a consequence the target
will cast a shadow (which may or may not be noticed depending on the
lighting etc — even if noticed may not be connected to the target) and
have a reflection in a mirror (or any reflective surface).  However
the target may not be noticed even if another entity is looking
directly at him/her.  It should be noted that a crystal of vision or
similar would count as looking directly at the target, not as a
reflection.  An entity will get a perception check if the target
becomes invasive on that entity’s senses (e.g. standing in a frontal
adjacent hex, or standing behind the entity with the target’s hands
over his/her eyes). Although the target is not invisible, it may be
detected using any magical means for detecting invisible entities
(e.g. witchsight).

If the target of the spell be touched by another entity, or that
entity’s possessions, then the spell is broken.  The target of the
spell may not break it voluntarily (other than by, for example,
touching another entity).  Once broken the spell must be recast.

\end{effects}
\end{spell}

\begin{spell}[G-6]{Healing}
\range{Touch}
\duration{Immediate}
\multiple{100}
\basechance{40\%}
\resist{None}
\storage{None}
\target{Entity}
\begin{effects}
Through a combination of magic and the application of healing herbs
and salves, the Adept can cure 3 ( + 1 / Rank) Damage Points suffered
due to disease or injury.  The spell takes about 5 minutes to cast.
\end{effects}
\end{spell}

\begin{spell}[G-7]{Detecting Traps and Snares}

\range{20 feet + 5 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{75}
\basechance{60\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
This spell increases the Base Chance of detecting traps or ambushes
while outdoors by 10 (+ 1 / Rank).
\end{effects}
\end{spell}

\begin{spell}[G-8]{Detecting Poisons}

\range{Touch}
\duration{5 minutes + 5 / Rank}
\multiple{75}
\basechance{50\%}
\resist{None}
\storage{Investment (in valid target only)}
\target{Object}
\begin{effects}
The spell imbues a piece (usually a wand) of ash wood, ivory, or
unicorn horn with the ability to detect poison.  When the wand is
touched to an object or substance for 5 seconds, the wand will
momentarily turn black if poison is present.  The wand may be used to
check up to 1 (+ 1 / Rank) objects or substances. By the shading and
patterns of the wand, further information may be determined about the
poison based on rank:

\begin{tabularx}{\linewidth}{lX}
Rank 8+ & Origin (Mineral, Plant, Animal, Magic) \\
Rank 10+ & Type (Alchemical or Natural) \\
Rank 16+ & Effects (Damage, Sleep, Unconsciousness, etc.) \\
Rank 20+ & Potency (Effective rank or damage of poison) \\
\end{tabularx}

\end{effects}
\end{spell}

\begin{spell}[G-9]{Lesser Enchantment}

\range{10 feet + 10 / Rank}
\duration{Special}
\multiple{125}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Sentient Entity}
\begin{effects}
The target of this spell is either blessed or cursed (Adept’s choice).
The spell increases either the character’s luck or misfortune
(depending on whether it operates as a blessing or as a curse) by 1 on
every percentile dice roll in which the character is directly
involved.  This spell cannot be cast over oneself.  The duration of
the enchantment is as follows: Ranks 1–10 = a fortnight, Ranks 11–19 =
3 months, Rank 20 permanent until dispelled.
\end{effects}
\end{spell}

\begin{spell}[G-10]{Herbal Lore}

\range{Self}
\duration{1 hour + 1 / Rank}
\multiple{75}
\basechance{25\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
This gives the Adept Rank 0 Herbalist for the duration of the spell.
It also increases their chance of finding herbs (as per the Herbalist
or Ranger skills) by 2\% (+ 2 / Rank). If the Adept is already ranked
in Herbalist or Ranger, then they gain an additional +10\%.
\end{effects}
\end{spell}

\begin{spell}[G-11]{Tracking}

\range{Self}
\duration{1 day + 1 / Rank}
\multiple{100}
\basechance{15\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept adds 10\% (+ 2 / Rank) to their chance of Tracking while outdoors.
\end{effects}
\end{spell}

\section{General Knowledge Rituals}

\begin{ritual}[Q-1]{Summoning Animals}

\duration{Immediate}
\multiple{150}
\basechance{MA + 3\% / Rank}
\resist{None}
\target{Animals}
\casttime{1 hour}
\material{None}
Actions: Concentration 
Concentration Check: Standard 
\begin{effects}
If the Ritual Check is successful then a number of small animals equal
to the Adept’s Rank (minimum 1) are summoned. The animal the Adept
attempts to summon must be a native of the area. If the Adept chooses
to vocalise their summons in the form of a loud shout or call the Base
Chance is increased by 25\%.
\end{effects}
\end{ritual}

\section{Special Knowledge Spells}

\begin{spell}[S-1]{Earth Hammer}

\range{25 feet + 10 / Rank}
\duration{Immediate}
\multiple{200}
\basechance{40\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
The Adept conjures a giant maul of stone which hurls itself at a
target of the Adept’s choosing. If the target fails to resist the
spell, it takes [D - 5] (+ 1 / Rank) damage.
\end{effects}
\end{spell}

\begin{spell}[S-2]{Hands of Earth}

\range{20 feet + 5 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{150}
\basechance{25\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity, Area}
\begin{effects}
The spell causes 1 + (Rank / 2) hands of stone to materialise out of
the ground within range.  Each hand is 7 feet tall and all must be
contiguous to each other. They may clutch anything that enters the hex
they occupy and will continue to do so until they have caught
something.  Any entity standing on a spot where a hand materialises
will be snatched up if they fail to resist, otherwise the hand appears
next to them within the hex.  Any entity caught will still be able to
wield a weapon or cast if they roll under 2 × MD. Entities (or
combination of entities) with a combined PS + target’s AG greater than
40 (+ 2 / Rank) may attempt a 1 × PS to escape from the Hands, and may
do so every pulse after the first.  Any hand which has been broken out
of will either dissipate or become hard cover (GM discretion).
\end{effects}
\end{spell}

\begin{spell}[S-3]{Strength of Stone}

\range{10 feet}
\duration{1 hour + 1 / Rank}
\multiple{200}
\basechance{20\%}
\resist{Passive}
\storage{Investment, Potion, Magical Trap}
\target{Entity}
\begin{effects}
The target of this spell has their Physical Strength or Endurance
(Adept’s choice) increased by Rank points (minimum 1).
\end{effects}
\end{spell}

\begin{spell}[S-4]{Armour of Earth}

\range{10 feet}
\duration{30 minutes + 30 / Rank}
\multiple{200}
\basechance{20\%}
\resist{None}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
This spell increases the defence from physical weapons of the target
by 2 + 2 / Rank.  At Rank 11 and above, it also absorbs 1 Damage Point
per strike.

\end{effects}
\end{spell}

\begin{spell}[S-5]{Diamond Weapon}

\range{5 feet + 5 / Rank}
\duration{5 minutes + 1 / Rank}
\multiple{250}
\basechance{30\%}
\resist{None}
\storage{Investment}
\target{Object}
\begin{effects}
This spell is cast over a weapon of the Adept’s choice, increasing its
strike chance by Rank\% (minimum of 1).  The chance of an attacker’s
weapon breaking increases by 5\% when striking a weapon under the
effect of this spell (i.e.  break chance becomes 94–99, instead of
99).
\end{effects}
\end{spell}

\begin{spell}[S-6]{Gem Creation}

\range{10 feet}
\duration{1 day + 1 / Rank}
\multiple{250}
\basechance{10\%}
\resist{None}
\storage{Investment}
\target{Area}
\begin{effects}
This spell creates one gemstone of random value for each 5 (or
fraction) Ranks.  The gems dissipate at the end of the spell.
\end{effects}
\end{spell}

\begin{spell}[S-7]{Animal Growth}

\range{10 feet + 10 / Rank}
\duration{1 day + 1 / Rank}
\multiple{225}
\basechance{15\%}
\resist{Passive}
\storage{Investment, Ward}
\target{Non-sentient mammal}
\begin{effects}
One mammal of the Adept’s choice is doubled in size.  The effects of
this radical change are determined by the GM.

\end{effects}
\end{spell}

\begin{spell}[S-8]{Enchanting Plants}

\range{10 feet + 10 / Rank}
\duration{1 day + 1 / Rank}
\multiple{225}
\basechance{15\%}
\resist{None}
\storage{Investment, Ward}
\target{Plant}
\begin{effects}
This spell may impart partial mobility to a number of plants
(including trees) equal to the Adept’s Rank.  The plants may not
uproot themselves, but may move their branches and leaves while
remaining in the same spot.  The plant’s actions are always under the
control of the Adept so long as they maintain their concentration.  If
their concentration is broken, voluntarily or otherwise, the plants
will be controlled by the GM until the Adept reestablishes control and
could, conceivably attack the Adept.

\end{effects}
\end{spell}

\begin{spell}[S-9]{Binding Animals}

\range{10 feet + 10 / Rank}
\duration{Until dispelled}
\multiple{250}
\basechance{10\%}
\resist{Passive}
\storage{Investment, Ward}
\target{Animal}
\begin{effects}
This spell is similar to the Spell of Controlling Animals except that
the Adept does not have to concentrate on it to maintain it.  The
spell will last until broken by the animal or dispelled by magic. Any
animal subject to this spell makes one check per week against its
Willpower to determine if the spell is broken.  This only works on
land animals, it will not work on avians or aquatics.

\end{effects}
\end{spell}

\begin{spell}[S-10]{Conjuring Earth Elemental}

\range{20 feet}
\duration{Immediate}
\multiple{225}
\basechance{25\%}
\resist{None}
\storage{Investment, Ward}
\target{Earth Elemental}
\begin{effects}
If the cast is successful, an Earth Elemental with a combined Fatigue
and Endurance of 15 (+ 5 / Rank) appears within 20 feet of the Adept.
The Adept automatically casts a Spell of Controlling Earth Elemental
to see if they control the Elemental. The Control spell is a separate
spell and requires additional fatigue to cast, but does not require
any preparation and is cast in conjunction with this spell.

\end{effects}
\end{spell}

\begin{spell}[S-11]{Controlling Earth Elemental}

\range{20 feet}
\duration{Concentration: no maximum}
\multiple{225}
\basechance{20\%}
\resist{None}
\storage{Investment}
\target{Earth Elemental}
\begin{effects}
The Adept may attempt to control an Elemental they have just summoned.
The Elemental does not get to resist.  If successful in establishing
control over the Elemental, the Adept maintains control until their
concentration is broken or they banish the Elemental with a
counterspell.  If they fail to gain control of the Elemental or gain
control, but have their concentration broken, the Elemental will
immediately attack them.  The Adept cannot banish an Elemental they do
not control.

\end{effects}
\end{spell}

\begin{spell}[S-12]{Sinking Doom}

\range{30 feet + 10 / Rank}
\duration{Immediate}
\multiple{650}
\basechance{2\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
This spell opens a circular pit under a single human sized target per
5 (or fraction) Ranks and sucks the unfortunate standing over it down
to be encased in rock 5 (+ 5 / Rank) feet underground.  This spell may
be used to affect multi-hex targets, each hex of the target counts as
one human-sized target. If the spell does not affect every hex which
the target occupies then the target will not be affected.
\end{effects}
\end{spell}

\begin{spell}[S-13]{Wall of Stone}
\range{20 feet + 10 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{150}
\basechance{10\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
Creates a 10 foot high × 20 foot long wall of granite or a 10 foot
high ring of stone with a 20 foot radius or a pillar of stone 15 feet
high and with a 2 foot radius. The Adept increases any dimension by 1
foot per Rank. They may not attempt to create a wall on top of an
entity.
\end{effects}
\end{spell}

\begin{spell}[S-14]{Wall of Iron}

\range{20 feet + 10 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{250}
\basechance{5\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
Same as for S-13 (Wall of Stone Spell) except that the Adept creates a
wall of cold iron.
\end{effects}
\end{spell}

\begin{spell}[S-15]{Tunnelling}

\range{5 feet + 1 / Rank}
\duration{30 seconds + 5 / Rank}
\multiple{200}
\basechance{10\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
The Adept creates a circular opening or tunnel 10 feet in diameter and
20 feet deep in a wall, ceiling, floor, ground surface, cliff face,
etc.  The Adept may add 1 foot to either depth or radius per Rank.
\end{effects}
\end{spell}

\begin{spell}[S-16]{Trollskin}

\range{10 feet + 10 / Rank}
\duration{3 hours + 1 / Rank}
\multiple{400}
\basechance{10\%}
\resist{Passive}
\storage{Investment, Ward}
\target{Volume}
\begin{effects}
The spell allows the subject to regenerate Endurance Points.  The
target begins to regenerate the pulse after the spell is cast and
continues to regenerate at the rate of 1 Endurance Point per pulse for
the duration of the spell.  The spell will not help regenerate wounds
inflicted by acid or fire.  An entity will not die from damage while
under the effects of a trollskin, even if they are below negative half
Endurance.
\end{effects}
\end{spell}

\begin{spell}[S-17]{Smoking Magma}

\range{25 feet + 5 / Rank}
\duration{10 seconds + 10 / Rank}
\multiple{300}
\basechance{7\%}
\resist{Passive}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
The Adept creates a pool of molten rock which wells up from
underground.  The pool has a radius of 5 feet (+1 / Rank).  Any entity
within the area covered by the pool suffers damage of [D - 5] (+ 1 /
Rank).  If the target successfully resists, this damage is halved.
Note that while the entity is within the area of effect the damage is
applied every pulse.

\end{effects}
\end{spell}

\begin{spell}[S-18]{Diamond Javelins}

\range{30 feet + 10 / Rank}
\duration{Immediate}
\multiple{300}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward}
\target{Entity}
\begin{effects}
The spell causes diamond-tipped javelins to fly from the earth at the
Adept’s feet and travel toward a target(s) of the Adept’s choice. The
number of javelins which appear is 1 (+ 1 / 2 (or frac- tion) Ranks).
The javelins have a Base Chance to hit equal to the Base Chance of
ordinary javelins + 3 / Rank.  Each javelin does [D - 4] (+ 1 / 2 (or
fraction) Ranks) damage.

\end{effects}
\end{spell}

\begin{spell}[S-19]{Earth Transformation}

\range{10 feet}
\duration{30 seconds + 5 / Rank}
\multiple{250}
\basechance{20\%}
\resist{None}
\storage{Investment, Potion, Ward}
\target{Entity}
\begin{effects}
This spell turns 3 (+ 1 / Rank) cubic feet of stone into mud or vice
versa. An entity standing on a spot containing a mud puddle that is
turned to stone has an opportunity to passively resist the effects of
the spell.  If they fail to resist, they will become trapped in the
stone that has taken the place of the mud they were standing in.  If
they resist, the mud turns to stone, but they are not trapped.
\end{effects}
\end{spell}

\section{Special Knowledge Rituals}


\begin{ritual}[R-1]{Binding Earth}

\range{10 feet + 10 / Rank}
\duration{ Concentration:  maximum  1  hour  +  1  / Rank}
\multiple{500}
\basechance{10\%}
\resist{None}
\target{Earth}
\casttime{1 hour}
\material{A human (or humanoid)}
\actions{Sacrifice human}
\concentration{Standard}
\begin{effects}
If the is ritual successful, the Adept gains complete control over a
500 pound weight of earth and stone (plus an additional 500 pound
weight per Rank).  They can shape or move the earth, change its
consistency or instill intelligence in it as they choose.  The Base
Chance to successfully employ this ritual is 10\%.  It can affect any
earth or stone within 10 feet (+ 10 / Rank) of the Adept. If a ritual
sacrifice of a human (or humanoid) being is per- formed at the end of
the duration, the affected earth becomes permanently bound (that is,
it contains no life and blocks all earth shaping effects, for example,
Hands of Earth, Earth Elementals, Tunnelling).
\end{effects}
\end{ritual}

\end{Chapter}
