\begin{Chapter}{The College of Ensorcelments \& Enchantments (Ver 1.1)}

This College is concerned with general magic, but especially with
charming and enchanting individuals and objects.  Practitioners of
this college are often knows an Enchanters.


\section{Restrictions}

Adepts of the College of Ensorcelments and Enchantments may practice
their arts without restriction.

The MA requirement for this College is 16. 


\section{Base Chance Modifiers}

There are no modifiers to the Base Chance of performing any talent,
spell, or ritual of the College except as listed in magic modifiers or
under the descriptions of the specific spells, talents, and rituals of
the College.


\section{Talents}

\begin{talent}[T-1]{Wizardsight }

\multiple{150 }
Effects: The Adept may see objects or entities that 
are  invisible  —  they  appear  to  have  a  slight  blue 
sheen  around  them.  If  the  invisibility  effect  (ex-
cluding Walking Unseen) is of a higher Rank than 
the  Wizardsight,  the  object  or  entity  may  not  be 
clearly identified or directly magically targeted. 
\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Charming }

\range{15 feet + 15 / Rank }
\duration{1 hour + 1 / Rank }
\multiple{500 }
\basechance{15\% }
\resist{Active, Passive }
\storage{Investment, Ward }
\target{Entity }
Effects:  The  Adept  may  influence  the  actions  of 
any one entity by casting a Spell of Charming over 
the individual. If the Generic True Name is known, 
the Base Chance is increased by 15. If the Individ-
ual  True  Name  is  known  and  used,  the  Base 
Chance is increased by 25. 

The  victim  of  the  spell  will  then  see  the  caster  as 
their  true  friend  and  will  readily  accept  most  sug-
gestions  emanating  from  them.  Sworn  enemies  of 
the caster (or of their race) will not be affected by 
this spell unless the victim’s Individual True Name 
is  spoken  and,  even  then,  the  duration  of  the  spell 
is halved. 

The caster may either suggest actions to the victim 
or  may  order  them to  act in  a  certain  way  on  pain 
of losing the caster’s “friendship”. Any suggestion 
that  is  not  directly  and  obviously  inimical  to  the 
victim’s  interests  (as  defined  by  the  GM,  but usu-
ally  limited  to  actions  that  would  be  injurious  or 
fatal  to  the  victim)  will  be  acted  upon  90\%  of  the 
time.  The  caster  may  only  ensure  100\%  compli-
ance  with  a  request  by  making  it  an  order  based 
upon the friendship that binds the victim and caster 
together.  In  such  instances,  however,  the  victim 
immediately  makes  a  Resistance  Check.  If  they 
resist,  then  the  spell  is  broken.  Otherwise,  the 
victim  will  accept  the  order  and  be  100\%  faithful 
to it for the remainder of the spell. 

Whenever any suggestion is made that would lead 
to  the  victim’s  injury  or  death,  another  Resistance 
Check  is  made.  If  the  check  is  unsuccessful,  the 
victim  will  accept  the  suggestion,  otherwise  the 
spell will be broken. 

For each Rank of the spell the Adept may move an 
additional  5  pounds  or  increase  the  speed  by  an 
additional 3 TMR. 

\end{spell}

\begin{spell}[G-3]{Enchanted Sleep }

\range{15 feet + 15 / Rank }
\duration{1 hour + 1 / Rank }
\multiple{250 }
\basechance{15\% }
\resist{Active, Passive }
\storage{Investment, Ward, Magical Trap }
\target{Entity }
Effects: The Adept may send the target, provided it 
normally  spends  any  time  sleeping,  into  a  deep 
enchanted sleep which will last for the duration of 
the  spell  or  until  the  entity  is  awoken  by  another 
entity  (by  being  shaken,  etc.).  The  target  may  not 
be  wakened  if  the  spell  is  Rank  10  or  higher,  but 
must continue to sleep until the spell wears off. If a 
General  Knowledge  counterspell  of  this  college  is 
cast upon an affected target by any Adept then the 
spell  will  immediately  dissipate  and the  target  can 
then be woken normally. 

\end{spell}

\begin{spell}[G-4]{Walking Unseen }

\range{1 foot + 1 / Rank }
\duration{1 hour + 1 / Rank }
\multiple{100 }
\basechance{40\% }
\resist{None }
\storage{Investment, Potion, Ward, Magical Trap }
\target{Entity }
Effects:  The  target  of  this  spell  may  move  unno-
ticed,  not  invisible.  This  means  that  it  will  not 
transmit light. As a consequence the target will cast 
a  shadow  (which  may  or  may  not  be  noticed  de-
pending on the lighting etc — even if noticed may 
not  be  connected  to  the  target)  and  have  a  reflec-
tion  in  a  mirror  (or  any  reflective  surface).  How-
ever  the  target  may  not be noticed  even  if  another 
entity  is  looking  directly  at  him/her.  It  should  be 
noted  that  a  crystal  of  vision  or  similar  would 
count  as  looking  directly  at  the  target,  not  as  a 
reflection.  An  entity  will  get  a  perception check if 
the target becomes invasive on that entity’s senses 
(e.g. standing in a frontal adjacent hex, or standing 
behind  the  entity  with  the  target’s  hands  over 
his/her eyes). Although the target is not invisible, it 
may  be  detected  using  any  magical  means  for 
detecting invisible entities (e.g. witchsight). 

If  the  target  of  the  spell  is  touched  by  another 
entity, or that entity’s possessions, then the spell is 
broken.  The  target  of  the  spell  may  not  break  it 
voluntarily  (other  than  by,  for  example,  touching 
another  entity).  Once  broken  the  spell  must  be 
recast. 

\end{spell}

\begin{spell}[G-5]{Speaking to Enchanted Creatures }

\range{15 feet + 15 / Rank }
\duration{10 minutes + 10 / Rank }
\multiple{100 }
\basechance{40\% }
\resist{None }
\storage{Investment }
\target{Entity }
Effects:  The  spell  gives  the  Adept  the  ability  to 
speak  and  understand  the  language  of  all  magical 
creatures.  One  casting  will  allow  communication 
with any fantastical creature, but will not allow the 
Adept  to  comprehend  fantastical  creatures  talking 
amongst themselves or to other people affected by 
this spell. 

\end{spell}

\begin{spell}[G-2]{Telekinesis }

\end{spell}

\begin{spell}[G-6]{Location }

\range{15 feet + 15 / Rank }
\duration{10 seconds + 10 / Rank }
\multiple{300 }
\basechance{20\% }
\resist{None }
\storage{Investment, Ward }
\target{Entity or object }
Effects: Allows the Adept to lift a target of weight 
up  to 2 pounds  and  move  it  at the  rate  of  2  TMR. 

\range{10 miles + 5 / Rank }
\duration{1 hour + 1 / Rank }
\multiple{200 }
\basechance{15\% }
\resist{None }
\storage{Potion }
\target{Self }
Effects: The  Adept may determine the direction in 
which  they  will  find  any  person  or  object  of  their 

43 

desire  which  they  have  previously  encountered  or 
studied  and  which  is  within  range.  The  direction 
will  be  indicated  by  a  large  glowing  arrow,  only 
visible  to  the  Adept.  The  arrow  will  not  appear  if 
the target is not within range. If cast, and the target 
is not within range, then the spell is dissipated. The 
arrow  will  not  appear  should  the  target  come 
within  range  during  the  period  the  spell  would 
normally have been in effect. 

\end{spell}

\begin{spell}[G-7]{Mass Charming }

\range{15 feet + 15 / Rank }
\duration{Concentration: no maximum }
\multiple{850 }
\basechance{5\% }
\resist{Active, Passive }
\storage{Investment }
\target{Entity }
Effects:  Charms  1  entity  per  Rank  as  long  as  the 
Adept  maintains  their  concentration.  The  spell 
takes 3 minutes to take effect and the effects linger 
3 to 5 minutes after concentration is broken. 

The effects of the spell on the individuals are iden-
tical to the effects of Spell of Charming [G-1]. 

\end{spell}

\begin{spell}[G-8]{Invisibility }

\range{15 feet + 15 / Rank }
\duration{5 minutes + 5 / Rank }
\multiple{450 }
\basechance{45\% }
\resist{None }
\storage{Investment, Potion, Ward }
\target{Entity or object }
Effects:  The  target  becomes  invisible.  Unless  cast 
at rank 16 or higher, the spell ceases whenever the 
target  makes  a  strike  check  in  melee  or  close 
(whether successful or not). The target may choose 
to end the spell at any time. 

An  invisible  thing  does  not  have  a  shadow  or  re-
flection  but  is  still  affected  by  light  (e.g.  may  still 
see,  be  sunburned,  be  blinded  by  Flash  of  Light, 
damaged by Solar Flare etc.).  

The possessions of the target are also invisible. An 
object’s  possessions  are  anything  that  is  totally 
enclosed  by  that  object  (e.g.  coins  in  an  invisible 
chest  which  is  closed, but not  one  which  is  open). 
If  a  thing  ceases  to  be  a  possession,  then  it  is  no 
longer  invisible.  Similarly  if  an  object  becomes  a 
possession  then  the  spell  will  affect  it  (i.e.  an  ob-
ject  put  down  by  an  invisible  entity  will  become 
seen; a coin put in the invisible chest will become 
invisible). Note that light may not be a possession 
—  if  the  target  is  carrying  a  lantern  it  will  be  in-
visible, but the light it emits will not. 

\end{spell}

\begin{spell}[G-9]{Evil Eye }

\range{Self }
\duration{1 day + 1 / Rank }
\multiple{300 }
\basechance{30\% }
\resist{None }
\storage{Potion, Magical Trap, Investment }
\target{Entity  }
Effects:  When  the  adept  casts  Evil  Eye  on  them-
self, a third eye appears in their forehead, which is 
normally invisible but is able to be seen by Witch-
sight  and  similar  effects.  When  the  adept  is  under 
the  effect  of  the  Evil  Eye  all  spells  they  cast  on 
others  are  resisted  at  a  penalty  of  Rank  unless  the 
target is wearing an Amulet of Elder Flowers. 
\end{spell}


\section{General Knowledge Rituals}

\begin{ritual}[Q-1]{Greater Enchantment }

\duration{Special }
\multiple{125 }
\basechance{80\% + 1\% / Rank }
\target{Entity }
\casttime{1 hour }
\material{Black Myrrh (optional) }
Material Cost: 200 sp per oz used 
Effects:  The  ritual  takes  1  hour  and  requires  that 
the  Adept  first  draw  a  Pentacle  within  which  they 
and  the  target(s)  must  remain  during  the  entire 
ritual. The Adept may burn black myrrh during the 
ritual  to  enhance  the  effects.  The  target  of  this 
ritual  will  be  either  blessed  or  cursed  (Adept’s 
choice)  with  an  increase  or  decrease  in  their  Base 
Chance of doing anything or suffering any good or 
ill fortune by plus or minus 1 + (1 / Rank). 

The ritual will affect the target’s fortune in one of 
the following areas:  

•  Resistance  —  Magic  Resistance,  Fear,  Fright  \& 
Awe Checks.  

• Magic — Talent, Spell, and Ritual Base Chances.  

•  Combat  —  Weapon  Strike  Chances  and  Stun 
Recovery.  

• Skills and Statistic Checks — Skill Base Chances 
and  General  Statistic  Checks  (e.g.  1  ×  PC  or  4  × 
AG). 

The default duration is until the next end-of-season 
High  Holiday.  This  will  enhance  one  of  the  above 
areas. 

The  Adept  may  alternatively  cast  a  form  of  the 
ritual that will affect 5 targets for 1 day (plus 1 day 
OR  1  target  per  4  full  ranks).  This  will  enhance 
two of the above areas. 

To  increase  the  number  of  areas  affected,  the 
Adept may burn (1 oz × Ritual Rank) Black Myrhh 
per additional area. 

If the ritual is used to curse, the curse is minor. 
\end{ritual}

\begin{ritual}[Q-2]{Creating Crystal of Vision }

\multiple{200 }
\basechance{75\% + 1\% / Rank }
\casttime{1 hour }
Effects: The Adept can create a crystal that acts as 
a  viewing  crystal.  They  perform the  ritual  over  an 
available  piece  of  crystal  (the  bigger  the  crystal, 
the  better  the  image  will  ultimately  be)  in  their 
possession.  The  Adept  must  burn  1  ounce  of  am-
bergris  during  the  ritual  at  a  cost  of  1,000  Silver 
Pennies.  The  resulting  crystal  may  be  used  once 
per day for 10 minutes + 1 minute per Rank (GMs 
should carefully time consultations). 

They  may  then  view  visions  (usually  precognitive 
in  nature)  concocted  by  the  GM.  At  Rank  6  and 
above, they may use the crystal to spy into an area 
to  see  what  is  going  on  there.  The  maximum  dis-
tance  from  the  character  to  the  area  being  spied 
into is 5 miles (+ 15 / Rank). To do so, the Adept 
must  remain  in  the  same  place  and  take  no  other 
action. 

If a crystal ball is used to look at a location then the 
ball  will  only  look  at  the  location  specified.  It 
cannot then be commanded to move around. It will 
enable the observation of events in that location for 
its duration and will then clear. Once started, it will 
continue to look at that location for its entire dura-
tion.  If  the  user  cannot  form a  clear  mental  image 
for the ball to focus on, or command it to focus on 
an unambiguously defined point in space, then it is 
the GM’s discretion as to what the ball  will show. 
These are mystical devices, not telescopes or x-ray 
machines. It is not possible to use detection talents 
(such as Detect Aura) through a crystal of vision. 
\end{ritual}

\begin{ritual}[Q-3]{Creating Sleep Dust }

\multiple{250 }
\basechance{80\% + 1\% / Rank }
\casttime{3 hours }
Effects:  The  Adept  must  spend  3  hours  preparing 
and  implementing  this  ritual  and  must  expend 
10,000 Silver Pennies (500 / Rank) to purchase the 
necessary  ingredients  prior  to  making  the  attempt. 
If  the  attempt  fails,  the  ingredients  are  ruined  and 
may not be reused or resold. If the ritual succeeds, 
one ounce (a single dose) of sleep dust results. 

When  thrown  in  the  face  of  one  target  it  has  the 
effect of an instant Spell of Enchanted Sleep of the 
same  Rank  as  the  sleep  dust.  The  sleep  dust  will 
only  remain  fresh  for  three  weeks  after  manufac-
ture. 

The  effects  of  sleep  dust  can be  passively  resisted 
by  the  victim,  but  with  a  reduction  of  20  to  their 
Magic Resistance. 
\end{ritual}

\begin{ritual}[Q-4]{Manufacturing Poison Dust }

\multiple{250 }
\basechance{80\% + 1\% / Rank }
\casttime{3 hours }
Effects: This ritual works in the same manner as Q-
3 and employs the same numbers for all purposes. 
However,  instead  of  causing  the  victim  to  fall 
asleep,  it  inflicts  [D  -  5]  +  1  for  every  2  (or  frac-
tion)  ranks  damage  due  to poisoning,  on  individu-
als  who  fail  to  resist.  Like  sleep  dust,  poison  dust 
must be thrown in the face of the target. 
\end{ritual}

\section{Special Knowledge Spells}


\begin{spell}[S-1]{Ventriloquism }

\range{90 feet }
\duration{5 minutes + 3 / Rank }
\multiple{100 }
\basechance{60\% }
\resist{None }
\storage{Potion }
\target{Self }
Effects: The spell allows the Adept to project their 
voice  and  alter  it  so  that  it  sounds  like  any  other 
voice the Adept has heard. The voice may be pro-
jected  so  that  it  appears  to  be  emanating  from 
anywhere within the radius of the spell. 

\end{spell}

\begin{spell}[S-2]{Bolt of Energy }

\range{15 feet + 15 / Rank }
\duration{Immediate }
\multiple{200 }
\basechance{50\% }
\resist{Passive }
\storage{Investment, Ward, Magical Trap }
\target{Entity or object }
Effects:  The  Adept  may  cast  a  bolt  of  energy  at  a 
target  and  will,  as  a  result,  inflict  [D  -  5]  (+ 1 per 
Rank) damage on the first entity or object that the 
bolt encounters. 

\end{spell}

\begin{spell}[S-3]{Opening }

\range{15 feet + 15 / Ranks }
\duration{Immediate }
\multiple{150 }
\basechance{30\% }
\resist{Passive }
\storage{Investment }
\target{Object }
Effects:  Instantly  opens  any  one  object  or  portal, 
including  those  locked  by  Mage  Lock  (S-6).  A 
door  or  lid  will  have  all  locks  unlocked  and  bolts 
undone and will become immediately open (ajar). 

\end{spell}

\begin{spell}[S-4]{Enchant Weapon }

\range{5 feet + 5 / Rank }
\duration{5 minutes + 1 / Rank }
\multiple{250 }
\basechance{30\% }
\resist{None }
\storage{Investment }
\target{Object }
Effects:  Increases  the  Base  Chance  to  hit  with  the 
weapon over which it is cast by 1 (+ 1 / Rank) and 
increases the damage done by the weapon by 1 for 
every 3 (or fraction) Ranks. 

\end{spell}

\begin{spell}[S-5]{Web of Entanglement }

\range{15 feet + 15 / Rank }
\duration{ Concentration:  maximum  15  minutes  + }
15 / Rank 
\multiple{150 }
\basechance{35\% }
\resist{Passive }
\storage{Investment, Ward }
\target{Entity or object }
Effects:  This  spell  allows  the  Adept  to  project  a 
sticky web, similar to a spider web in construction, 
from  their  fingertips  at  a  target  hex,  object  or  en-
tity.  Any  objects  or  entities  standing  between  the 
Adept and the target are ensnared by the web along 
with the target. The web may only ensnare a num-
ber  of  human-sized  entities  equal  to  the  Adept’s 

44 

rank and so will stop at the hex at which this limit 
is reached (or maximum range). 

The  web  continues  until  cut  or  until  the  Adept 
ceases to concentrate on it. In order to cut the web, 
a character must successfully strike with a weapon 
that does type B damage and must inflict at least 10 
points  of  damage  as  a  result  of  that  single  strike. 
The web has no defence. 

Any  ensnared  entity  must  roll  1  ×  PS  (2  ×  PS  if 
they  successfully  resisted)  in  order  to  move  them-
selves to an adjacent hex (which may be free of the 
web),  or  to  perform  an  action  within  the  web.  A 
similar  check  is  required  for  any  entity  attempting 
to  remove  an  object  from  the  web.  If  an  entity 
receives aid in removing themselves from the web, 
the  PS  of  the  aiding  character  may  be  combined 
with  their  own.  Any  dropped  object  will  become 
ensnared by the web, as will any entity who comes 
into contact with it (up to the limit of the web). 

\end{spell}

\begin{spell}[S-6]{Mage Lock }

\range{15 feet + 15 / Rank }
\duration{1 hour + 1 / Rank }
\multiple{150 }
\basechance{30\% }
\resist{None }
\storage{Investment }
\target{Object }
Effects:  The  spell  may  be  cast  over  any  portal 
(door,  window,  etc.)  that  can  normally  be  opened 
or  closed.  It  effectively  locks  the  portal  with  an 
unpickable  lock.  The  portal  may  still  be  forced 
open by brute strength. The Physical Strength(s) of 
all characters attempting to force a portal locked in 
this  manner  are  added  together  and  multiplied  by 
the  Difficulty  Factor  of  the  task.  The  Difficulty 
Factor is a function of the Rank of the spell: 

Rank  Difficulty Factor 

1–5 
6–10 
11–20 

3 
2 
1.5 

\end{spell}

\begin{spell}[S-7]{Enhancing Enchantment }

\range{15 feet + 15 / Rank }
\duration{10 seconds + 5 / Rank }
\multiple{300 }
\basechance{25\% }
\resist{None }
\storage{Investment, Ward }
\target{Area }
Effects: The Rank of this spell is added to the Rank 
of one characteristic of any spell being cast within 
the area of effect. Note that the actual Rank of the 
affected  spell  is  unchanged  —  only  the  effect  of 
one  characteristic  is  enhanced  (as  per  double  and 
triple effects). 

While  casting  the  Spell  of  Enhancing  Enchant-
ment, the adept specifies the spell characteristic to 
be  affected.  Only  Range,  Duration,  Base  Chance, 
or  (where  appropriate)  Damage,  may  be  affected 
by this spell. 

A  spell  may  never  have  a  characteristic  enhanced 
by  Ranks  greater  than  its  own  rank  by  the  use  of 
this  spell  (e.g.  a  Rank  6  spell  that  is  cast  within  a 
Rank 8 Spell of Enhancing Enchantment will only 
gain  the  effects  of  6  extra  ranks  in  the  affected 
characteristic). 

Only  spells  that  are  cast  are  affected,  not  spells 
being  released  from  any  form  of  storage.  The 
caster  of  the  subject  spell  will  be  aware  that  their 
spell  is  being  affected  by  enhancement  during  the 
pulse  that  they  are  preparing  (as  per  high  or  low 
Mana). They  will not know which characteristic is 
being  affected  nor  the  amount  of  enhancement. 
They  may  not  restrict  the  effect  of  the  enhance-
ment,  but  may  reduce  the  spell  effects  in  the  nor-
mal manner. If an attempt is made to cast a Spell of 
Enhancing  Enchantment  on  an  area  that  has  some 
part  of  it  under  the  effects  of  another  Spell  of  En-
hancing  Enchantment,  then  the  second  spell  will 
fail. 

Special  Duration,  Range,  or  Damage  effects  may 
be caused by the use of this spell (e.g. Duration of 
Lesser Enchantment at Rank 20). 

\end{spell}

\begin{spell}[S-8]{Levitation }

\range{15 feet + 15 / Rank }
\duration{10 minutes × [D - 5] × [Rank] }
\multiple{125 }
\basechance{25\% }
\resist{Active, Passive }
\storage{Investment, Ward, Potion }
\target{Entity }
Effects:  Causes  the  target  of  the  spell  to  rise  into 
the  air  15  feet  (+  1  /  Rank)  at  the  rate  of  1  foot  / 
pulse.  The  spell  is  limited  to  vertical  movement 
and will in no way propel the target horizontally. 

\end{spell}

\begin{spell}[S-9]{Enchant Armour }

\range{15 feet + 15 / Rank }
\duration{30 minutes + 30 / Rank }
\multiple{200 }
\basechance{20\% }
\resist{None }
\storage{Ward, Investment }
\target{Entity }
Effects:  This  spell  adds  2  (+  2  /  Rank)  to  the  tar-
get’s defence versus physical weapons. At Rank 11 
and  above  it  also  permits  the  target’s  armour  to 
absorb  1  additional  point  of  damage.  The  target 
must be armoured. 

Wizard’s Eye (S-10) 

\range{15 feet + 15 / Rank }
\duration{1 minute + 1 / Rank }
\multiple{200 }
\basechance{20\% }
\resist{None }
\storage{Potion }
\target{Self }
Effects:  The  Adept  creates  an  invisible,  intangible 
eye  which  they  can  move  around  within  a  radius 
equal to the spell’s range. The eye originates in the 
same spot as the Adept and operates as would any 
normal  eye  except  that  it  is  not  attached  to  the 
Adept  physically.  Wizard’s  Eyes  have  a  TMR  of 
10. 

The  eye  can  move  through  solid  objects  but  will 
see only blackness while within an object. The eye 
can  be  seen  with  witchsight  or  other  means  of 
detecting invisible objects. It is possible to use the 
eye  to  target  spells,  but  the  range  is  determined 
from  the  Adept,  not  from  the  eye.  The  eye  (and 
hence the Adept) is susceptible to magical effects if 
the  effect  can  affect  eyes  (e.g.  Flash  of  Light, 
blindness). 

\end{spell}

\begin{spell}[S-11]{Slowness }

\range{15 feet + 15 / Rank }
\duration{10 seconds × [D - 5] × [Rank] }
\multiple{300 }
\basechance{20\% }
\resist{Active, Passive }
\storage{Potion, Investment, Ward, Magical Trap }
\target{Entity }
Effects:  The  spell  affects  1  (+  1  /  2,  or  fraction, 
Rank) targets. All entities subject to this spell have 
their running, crawling, flying, or swimming speed 
halved  and  have  the  time  it  takes  them to  do  any-
thing  on  the  Tactical  Display  doubled  (e.g.  they 
could only attack once every two pulses). 

\end{spell}

\begin{spell}[S-12]{Quickness }

\range{15 feet + 15 / Rank }
\duration{10 seconds × [D - 5] × [Rank] }
\multiple{300 }
\basechance{20\% }
\resist{Active, Passive }
\storage{Potion, Investment, Ward }
\target{Entity }
Effects:  The  spell  affects  1  (+  1  /  3,  or  fraction, 
Ranks)  targets.  The  targets’  Initiative  Value  is 
increased  by  10,  and  they  can  perform  Actions 
twice as often. 
\end{spell}


\section{Special Knowledge Rituals}

There are no Special Knowledge Rituals of the College of Ensorcelments
and Enchantments.

\end{Chapter}
