\begin{Chapter}{The College of Fire Magics (Ver 2.0)}

The College of Fire Magics is concerned exclusively with manipulating
the element of fire.  Adepts of this College are referred to as Fire
Mages or Pyromancers.

Within range low level heat sources (such as living 
bodies) can be seen. At double range medium level 
sources  (such  as  camp  fires)  are  visible  and  high 
level heat sources may be visible at any range. 

The  affinity  fire  mages  have  for  their  element  has 
often led to them being considered to be pyromani-
acs.  Most  fire  mages,  however,  have  a  healthy 
respect  for  fire,  and  rarely  will  they  wantonly  set 
people  or places alight. They do  enjoy being in or 
near  fire  and  are  generally  the  first  to  suggest  a 
nice  camp  fire  when  adventuring.  Fire  mages  are 
also  thought  to  be  hot-tempered,  to  enjoy  hot  and 
spicy  food,  and  to  dislike  getting  wet.  They  fre-
quently  have  red  hair.  Generally  fire  mages  are 
flamboyant  characters,  particularly  in  their  spell-
casting. The College of Fire Magics is anything but 
subtle,  and  its  spells  involve  more  than  the  usual 
amounts of shouting and arm waving. 

The  element  of  fire  is  unquestionably  destructive, 
and  as  a  consequence  the  majority  of  the  spells  of 
this college cause considerable amount of damage, 
if  only  as  a  side-effect.  The  Fire  College  is  not 
noted  for  being  harmless,  and  so  fire  mages  are 
even  less  popular  than  most  mages  among  the 
common folk. When it is revealed that a person is 
an Adept of the Fire College the usual reaction will 
generally  be  one  of  fear  and  trepidation.  Fire 
mages  are  quite  amicable  towards  the  Adepts  of 
other  colleges  with  the  notable  exception  of  those 
of the College of Water Magics. 

Unless  otherwise  specified,  Magical  Fire  will 
automatically ignite flammables. 

Traditional colours 

Fire mages generally prefer to wear red, the colour 
of their element, often trimmed with black, yellow 
or  orange.  For  jewellery  the  fire  mage  most  fre-
quently chooses garnets or rubies, set in gold. 

Traditional symbols 

The  symbol  of  the  Fire  College  is  the  flame,  usu-
ally depicted in red or gold. 

21.1 Restrictions 
Adepts  of  the  College  of  Fire  Magics  may  only 
practice their arts in an area where it is possible for 
fire  to  exist.  They  may  not  practice  fire  magic 
underwater or in a vacuum, for example. 

The MA requirement for this college is 12. 

21.2 Base Chance Modifiers 
The Base Chance of performing any talent, spell or 
ritual of the College of Fire Magics is modified by 
addition of the following numbers: 

-5 

-10 

Adept is in light mist, fog, or rain, or is 
soaking wet 
Adept is in heavy fog, or rain, or is par-
tially immersed in water 
Adept is almost totally immersed in water 
Adept is within 30 feet of large campfire 
Adept is within 30 feet of bonfire 
Adept is in contact with small campfire 
Adept is in contact with large campfire 
Adept is in contact with bonfire 
Adept is in a hot, dry region (e.g. desert) 
Only one modifier from each group may be applied 
together. These bonuses only apply to normal fire. 
Magical  fire  may  only  be  used  to  reduce  negative 
modifiers to zero. 

-15 
+5 
+10 
+5 
+10 
+15 
+5 

21.3 Talents 

Infravision (T-1) 

Range: 50 feet + 5 / Rank 
Experience Multiple: 75 
Resist: None 
Effects: The Adept is able to see heat sources as if 
they  emit  normal  light  and  target  them  (e.g.  with 
weapons or spells) in the absence of visible light. It 
works  best  in  relative  darkness,  since  it  is  easily 
over-powered by visible light. 

21 COLLEGE OF FIRE MAGICS 

Fire Armour (G-3) 

Range: Touch 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 200 
Base Chance: 25% 
Resist: None 
Storage: Potion 
Target: Object, Entity 
Effects:  This  spell  protects  the  target  against dam-
age by fire for 4 (+ 4 / Rank) points of  protection. 
Protection  is  ablative  and  when  the  damage  the 
spell  may  absorb  is  exceeded,  the  spell  is  dissi-
pated with any excess damage applied to the target. 
Double and triple effects may apply to duration or 
degree  of  protection.  At Rank  20  the  spell  confers 
100 points of protection. 

Firelight (G-4) 

Range: Touch 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 75 
Base Chance: 50% 
Resist: Passive 
Storage: Investment 
Target: Object, Point 
Effects:  The  Adept  creates  a  source  of  light  ema-
nating  in  all  directions  from  an  object  or  point 
touched by them. At ranks 0 to 5 the light emitted 
is equivalent to that of a torch, at ranks 6 to 10 that 
of a small campfire, at ranks 11 to 15 a large camp-
fire,  and  at  ranks  16  to  20  a  bonfire.  The  light 
emitted will have the appearance of firelight of the 
appropriate  strength.  It  is  magical  light,  and  will 
cast poor contrast shadows. 

Fireproofing (G-5) 

Range: Touch 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 150 
Base Chance: 30% 
Resist: None 
Storage: Potion, Investment 
Target: Entity, Object 
Effects:  The  spell  protects  the  target from all  non-
magical fire and heat effects up to the heat equiva-
lent  of  a  bonfire.  An  entity  or  object  is  also  pro-
tected  against  smoke  effects  (including  smoke 
inhalation), heatstroke and sunstroke. 

Increase Temperature (G-6) 

Range: Touch 
Duration: Special 
Experience Multiple: 100 
Base Chance: 35% 
Resist: Passive 
Storage: Investment 
Target: Volume, Object 
Effects: The Adept must remain in contact with the 
target for heating to occur. When the spell ends the 
target will cool normally. 

Gases 

Duration: 1 hour + 1 / Rank 
The  Adept  may  cause  the  temperature  of  a  closed 
contiguous volume of gas (such as air) to increase. 
The volume affected is equal to 125 (+ 125 / Rank) 
cubic feet. The temperature can be increased by up 
to  2(C  per  Rank.  At  Rank  20  the  volume  that  can 
be affected by Increase Temperature is doubled 

This  talent can  sometimes penetrate where  normal 
vision cannot. At half normal  range the Adept can 
see  heat  sources  which  are  warmer  than  the  ob-
scurement,  eg.  living  entities  through  mist  or  light 
bushes.  It  may  also  possible  to  detect  the  residual 
heat of a source which has recently been moved or 
extinguished. 

The Adept has a base chance of PC (+ 5 / Rank) - 1 
/  10  feet  between  the  Adept  and  the  heat  source 
(after the first 10) of gaining additional information 
about a particular heat source. They may:  

1.  Determine  the  relative  temperature  of  the  heat 
source. 

2.  Identify  the  general  type  of  heat  source.  Only 
those  heat  sources  that  the  Adept  has  previously 
encountered  may  be  identified  with  this  talent.  At 
higher ranks familiar individuals may be identified. 

3. Determine if the heat radiated by source is being 
generated  by  magic,  e.g.  the  Adept  is  able  to  dis-
tinguish  between  normal  and  magical  fire.  (Note 
only Fire College infravision has this ability). 

The  Adept  may  re-attempt  any  of  these  abilities 
after a period of 41 pulses (2 / Rank) has elapsed or 
if the Adept is 20 feet closer to the target. 

Pyrogenesis (T-2) 

Range: Sight 
Duration: Immediate 
Experience Multiple: 75 
Base Chance: MA + 5 / Rank - 1 / 5 feet separating 
the Adept from the target (after the first 5 feet) 
Resist: Passive 
Effects: The Adept may cause to burst into flame a 
single  mass  of  dry,  flammable  material  weighing 
up to 1oz (+ 1/ 

5 ranks). Once alight it will burn normally and may 
be  extinguished  by  either  mundane  or  magical 
means.  Flammable  materials  are  defined  as  wood, 
paper,  cloth  etc,  but  not  flesh,  except  that  Pyro-
genesis  may  be  used  to  cremate  insects  and  small 
creatures  within  the  maximum  weight  restrictions. 
This talent cannot be used on possessions. Utilising 
this talent requires a pass action, and is obvious to 
observers. 

21.4 General Knowledge Spells 

Bolt of Fire (G-1) 

Range: 25 feet + 25 / Rank 
Duration: Immediate 
Experience Multiple: 200 
Base Chance: 40% 
Resist: Passive 
Storage: Ward, Investment, Magical Trap 
Target: Entity, Object, Area 
Effects:  The  Adept  causes  a  bolt  of  Fire  to  streak 
from  their  hand  towards  anywhere  in  range.  The 
first  entity  or  object  the  bolt  hits  in  its  path  must 
resist  or  suffer  D  (+  1  / Rank)  damage.  If  the  bolt 
does not hit anything it will dissipate at the end of 
its range. At Rank 20 the Adept may delay releas-
ing the bolt for a pulse if they should choose. 

Extinguish Fires (G-2) 

Solids 

Range: 15 feet + 15 / Rank 
Duration: Immediate 
Experience Multiple: 100 
Base Chance: 50% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Volume 
Effects:  When  successfully  cast,  this  spell  will 
extinguish all fire in a 10 foot (+ 10 / Rank) radius 
sphere.  All  the  volume  affected  must  be  within 
range of the spell. If the range is doubled or tripled 
the volume may likewise be increased. 

Duration: 5 minutes + 5 / Rank 
The  Adept  may  heat  1  lb  (+  1  /  Rank)  of  solid 
material  up  to  50  (+  50  /  Rank)  degrees  C.  The 
Adept may choose any combination of temperature 
increase increment and mass increment to a total of 
the  Adept’s  rank.  The  temperature  increase  takes 
10 seconds / 100 degrees / pound mass. At Rank 20 
any metal item within the weight limit of the Adept 
can  be  reduced  to  molten  slag  in  as  little  as  10 
seconds (resistance roll is applicable). 

Liquids 

Duration: 5 minutes + 5 / Rank 

71 

21 COLLEGE OF FIRE MAGICS 

The Adept may heat 1 (+ 1 / Rank) pints of liquid 
up  to  20  (+  20  /  Rank)  degrees  C.  They  may 
choose  any  combination  of  temperature  increase 
increment  and  mass  increment  to  a  total  of  the 
Adept’s  rank.  The  temperature  increase  takes  10 
seconds / 10 degrees / pint. The volume of material 
that  may  be  affected  is  increased  by  6  times  at 
Rank 20. 

Slow Fire (G-7) 

Range: 5 feet + 5 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 100 
Base Chance: 20% 
Resist: None 
Storage: Investment 
Target: Object 
Effects:  This  spell  increases  the  duration  that  a 
wood,  lamp  or  oil  fire  will  burn  for  by  1  (+  1  / 
Rank) hours. Light remains constant while the spell 
is in effect but heat from the fire source is halved. 
This spell may be used to slow the effects of small 
fires  or  to  ensure  a  fire  lasts  for  a  considerable 
period of time. 

Smoke Creation (G-8) 

Range: 10 feet + 10 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 75 
Base Chance: 25% 
Resist: None 
Storage: Investment, Ward 
Target: Volume 
Effects: The Adept may create a volume of smoke 
equal to 1000 cubic feet (+ 500 / Rank). This vol-
ume  must  be  closed  and  contiguous,  and  has  a 
minimum  thickness  of  5  feet.  The  density  of  the 
smoke  may  vary  from  light  vapours  at  Rank  0  to 
thick,  roiling  smoke  at  Rank  20.  The  Adept  may 
choose the density of the smoke when casting. The 
smoke  produced  reduces  visual  perception  by  1 
multiplier per 5 full ranks (but may not reduce it to 
less  than  once  times  perception).  The  effective 
Rank of the spell will be decreased by 1 Rank per 
10  miles/hour  of  wind  (minimum  Rank  0)  in  the 
target volume. Below rank 10 the smoke created by 
this spell is completely transparent to infravision. 

21.5 General Knowledge Rituals 

Binding Fire (Q-1) 

Range: 10 feet + 10 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 500 
Base Chance: 10% 
Resist: None 
Target: Fire 
Material: Fire source + Endurance (optional) 
Actions: Chanting and dancing 
Concentration Check: Standard 
Effects:  The  Adept  gains  complete  control  over  a 
fire source as large as 10 (+ 10 / Rank) cubic feet. 
This  fire  may  then  be  moulded  and  shaped  as  de-
sired but with a minimum of one foot thickness in 
any  dimension.  Regardless  of  the  shape  created, 
the fire retains its normal heat and damaging prop-
erties. Damage is as per a large bonfire or [D - 3] if 
the  entity  is  within  one  hex.  Bound  fire  is  unaf-
fected by normal rains and winds short of hurricane 
conditions,  but  may  be  extinguished  by  a  dousing 
of water of similar or greater volume than the fire. 

If the Adept sacrifices 2 points of Endurance from 
a  sentient  as  part  of  the  ritual,  the  duration  of  the 
Ritual of Binding Fire is enhanced to 1 year (+ 1 / 
Rank). 

If the Adept sacrifices 4 points of Endurance from 
a  sentient  creature  the  duration  of  the  Ritual  of 
Binding Fire is permanent. This Fire  is bound to a 
specific  non-movable  item  from  which  the  fire 
appears to issue (e.g. a ruby set in a wall, a line of 
runes on a floor etc). Permanent Bound Fire can be 
suppressed for [21 - Rank] minutes by a dousing of 
water  of  similar  or  greater  volume  than  the  fire, 
attack with magical cold for more damage than the 
damage rating of the fire etc. To dissipate a perma-

nent  Bound  Fire  the  object  to  which  it  is  bound 
must be broken, destroyed or seriously defaced. 

The Adept may instil basic intelligence in a Bound 
Fire  if  desired  and  give  the  entity  thus  created 
simple commands. 

21.6 Special Knowledge Spells 

Cleansing Flame (S-1) 

Range: 10 feet + 10 / Rank above 10 
Duration: 125 seconds - 5 / Rank 
Experience Multiple: 300 
Base Chance: 20% 
Resist: Special 
Storage: Potion, Ward, Investment 
Target: Entity 
Effects:  The  target  is  wreathed  in  yellowy  green 
flames  causing  hideous  suffering,  yet  curing  the 
target  of  the  effects  of  natural  poisons  (acts  as 
Rank / 2 vs. synthetic poison), venoms, fevers and 
diseases. As the foulness is burnt away, the flames 
change  colour  until  they  become  silvery  white. 
While  cleansing  takes  place  any  ongoing  harm  is 
halted.  At  the  end  of  the  duration,  the  target  is 
cured  and  their  possessions  are  cleaned  to  a  high 
polish.  During  the  cleansing  all  the  target’s  Strike 
Chances,  Defence  and  Cast  Chances  are  reduced 
by  25  (-1  /  Rank).  This  spell  only  affects  willing 
targets. At Rank 20 this spell also cures the effects 
of Malignant Flames. This spell can also be used to 
cure  burns,  including  third  degree  burns  but  scar-
ring from burns are not affected. 

Dragon Flames (S-2) 

Range: Special 
Duration: Immediate 
Experience Multiple: 500 
Base Chance: 10% 
Resist: Passive 
Storage: Investment 
Target: Area 
Effects: The Adept may breathe magical fire like a 
Dragon  causing  D10  (+ 3 / Rank)  damage  to  enti-
ties within the area of effect. The area of effect is a 
cone issuing from the Adepts mouth and is 20 feet 
(+  5  /  Rank)  long,  and  5  (+  5  /  3  Ranks)  wide  at 
end farthest from the Adept. If an entity resists the 
damage  is  halved,  unless  they  are  more  than  half 
immersed in water (eg. swimming) where they take 
no  damage.  Dragon  Flames  are  analogous  to  an 
instantaneous  flash  of  heat,  but  there  are  no  over-
pressure or explosive effects. 

Fire Arc (S-3) 

Range: Touch 
Duration: Immediate 
Experience Multiple: 250 
Base Chance: 15% 
Resist: Passive 
Storage: Potion, Investment 
Target: Entity 
Effects: The Adept may transform 1 entity (+ 1 / 3 
or fraction Ranks) into a bolt of fire that flashes to 
any unobstructed point in line of sight up to 25 (+ 
25 / Rank) feet away. The bolt must land within 5 
feet of a potentially flammable substance, an entity 
or  a  solid  surface.  When  the  bolt  arrives,  it  bursts 
into  flame  causing  D10  damage  to  all  entities 
within  1  hex.  At  Rank  20  the  target  may  instanta-
neously return to the point of origin of the  spell at 
the  end  of  the  next  pulse  should  they  so  desire. 
Double and triple effects apply to range only. This 
spell affects only willing targets and will not work 
underwater. Any barriers, wards etc that are passed 
over affect the targets as if crossed normally. 

Fireball (S-4) 

Range: 60 feet + 10 / Rank 
Duration: Immediate 
Experience Multiple: 550 
Base Chance: 25% 
Resist: Active, Passive 
Storage: Investment, Ward 
Target: Area 
Effects: The Adept conjures a ball of fire 1 foot in 
diameter which rushes to a point in line of sight of 

72 

the Adept, and explodes in a radius of 5 foot (+ 5 / 
5 or fraction Ranks). Everything in this radius must 
resist  or  suffer  D10  (+  1  /  Rank)  damage.  If  an 
entity resists the damage is halved, unless they are 
more than half immersed in water (e.g. swimming) 
where  they  take  no  damage.  The  spell  may  set 
flammable  items  afire  when  it  bursts.  Hard  cover 
(e.g. walls, parapets, but not shields) reduces dam-
age  to  half  before  resistance.  In  some  instances, 
e.g.  a  character  wading  through  water,  resistance 
may negate all damage. The fireball is analogous to 
an  instantaneous  flash  of  heat  and  there  are  no 
over-pressure  or  explosive  effects.  A  fireball  may 
be  detonated  prematurely  by  the  imposition  of 
barriers in its line of flight. 

At Rank 20 this spell may be cast with a detonation 
delay  of  up  to  10  pulses.  This  extends  the  casting 
time  4  pulses.  A  delayed  blast  fireball  manifests 
itself  as  a  1  foot  wide  floating  sphere  of  fire  until 
detonation occurs. If a second fireball is cast within 
the proposed volume of effect of the first, both are 
detonated. 

Heat Shield (S-5) 

Range: Touch 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 200 
Base Chance: 25% 
Resist: None 
Storage: Potion 
Target: Object, Entity 
Effects:  This  spell  protects  the  target  against dam-
age  by  cold  or  ice  for  4  (+  4  /  Rank)  points  of 
protection.  Protection  is  ablative  and  when  the 
damage the spell may absorb is exceeded, the spell 
is dissipated with any excess damage applied to the 
target. Double and triple effects may apply to dura-
tion or degree of protection. 

Hellfire (S-6) 

Range: 10 feet + 5 / Rank 
Duration: Immediate 
Experience Multiple: 650 
Base Chance: 5% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  This  sulphurous  fire  attacks  1  target  for 
every  3  (or  fraction)  Ranks.  The  target’s  Magical 
Resistance is reduced by 5 (+ 1 / Rank). The spell 
does D10 (+ 2 / Rank) damage to each target.  If a 
target  successfully  resists,  they  suffer  only  half 
damage  (round  up).  Double  damage  adds  an  addi-
tional 1 / Rank damage and triple damage adds an 
additional 2 / Rank damage. 

Immolation (S-7) 

Range: Self 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 200 
Base Chance: 15% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  The  Adept  gains  the  grace  and  form  of  a 
dancing  flame,  without  in  anyway  altering  their 
physical  nature,  but  increasing  their  defence  by  2 
(+  3  /  2  Ranks).  Any  entity  that  is  within  5  feet 
must  resist  or  take  1  /  2  (or  fraction)  Ranks  Fire 
damage  per  pulse.  The  target  will  glow  with  the 
same  degree  of  brightness  as  a  Firelight  spell  5 
Ranks  lower,  (minimum  1).  The  target  appears  to 
be  a  humanoid  shape  composed  of  flame.  If  the 
target  hides  in  a  fire,  the  target  may  only  be  de-
tected by witchsight or by infravision.. 

Malignant Flames (S-8) 

Range: 10 feet + 10 / Rank 
Duration: Immediate 
Experience Multiple: 550 
Base Chance: 5% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects:  The  target  is  swathed  in  flames  that  are 
only visible to the victim causing D10 (+ 2 / Rank) 

damage. If the spell is at rank 5 or higher the target 
may  be  cursed  with  the  loss  of  [D  -  4]  (+1  /  5 
Ranks)  points  from  one  characteristic  or  statistic, 
as  chosen  by  the  Adept.  A  characteristic  may  not 
be  reduced  below  1  as  a  result  of  this  spell.  This 
minor curse will last 4 (+ 1 / Rank) days. 

Pyrotechnics (S-9) 

Range: 60 feet + 10 / Rank 
Duration: Immediate 
Experience Multiple: 200 
Base Chance: 20% 
Resist: Passive 
Storage: Investment, Magical Trap 
Target: Volume 
Effects:  This  spell  may  be  mistaken for  a  Fireball. 
The Adept conjures a ball of fire 1 foot in diameter 
which  rushes  to  a  point  in  line  of  sight  of  the 
Adept,  and  bursts in a  radius  of  5  foot  (+  5  /  5  or 
fraction Ranks) with an enormous flash of light and 
fireworks.  Any  entity  in  this  radius  who  can  see 
must resist or they are dazzled by the flash for [D - 
5] (+ 1 / 3 or fraction Ranks) pulses. Dazzled enti-
ties  have  their  strike  chances,  cast  chances  and 
perception checks reduced by 1 (+ 2 / Rank). Enti-
ties  without normal  vision  (e.g.  Undead)  are  unaf-
fected by Pyrotechnics. 

Speak to Fire Creatures (S-10) 

Range: 15 feet + 15 / Rank 
Duration: 20 minutes + 10 / Rank 
Experience Multiple: 75 
Base Chance: 40% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  This  spell  allows  the  Adept  to  communi-
cate  with  all  fire  creatures  within  range.  This 
communication is  at an  effective  language  rank  of 
1  (+  1  /  2  Ranks).  For  the  purposes  of  this  spell, 
fire  creatures  are:  Elementals,  Salamanders, 
Efreets,  desert  creatures  and  creatures  created 
using a Binding Fire Ritual. 

Summoning Salamander (S-11) 

Range: Unlimited 
Duration: Until dispelled 
Experience Multiple: 200 
Base Chance: 15% 
Resist: None 
Storage: Investment, Magical Trap 
Target: Salamander 
Effects:  The  Adept  may  summon  a  salamander 
which will then attempt to set afire anything flam-
mable that it can reach. A Salamander may only be 
summoned to an environment it can survive in (e.g. 
a large fire source, volcano, in a desert). The Adept 
has no control over the salamander. Dispelling the 
spell returns the salamander to its origin. 

Weapon of Flames (S-12) 

Range: 5 feet + 1 / Rank 
Duration: 5 minutes + 1 / Rank 
Experience Multiple: 250 
Base Chance: 30% 
Resist: None 
Storage: Investment 
Target: Weapon 
Effects: The Adept may cause any weapon to burst 
into  flame  but  without  causing  damage  to  the 
weapon  or  its  wielder.  The  weapon  has  its  Strike 
Chance  increased  by  1  (+  1  /  Rank),  damage  in-
creased  by  +  1  /  2  (or  fraction)  Ranks.  When  the 
weapon is used against a entity which are creatures 
of  cold  or  water  or  against  the  Undead,  damage 
increases  to  1  /  Rank.  A  missile  may  also  be  the 
target  (e.g.  arrow),  but  not  missile  weapons.  At 
rank  11  any  hand-held  weapon  may  be  created 
from  fire.  This  weapon  has  normal  characteristic 
requirements but no weight. 

Wall of Fire (S-13) 

Range: 10 feet + 10 / Rank 
Duration: 10 minutes + 10 / Rank 
Experience Multiple: 250 
Base Chance: 25% 
Resist: Passive 

Storage: Investment, Ward, Magical Trap 
Target: Area 
Effects:  The  Adept  may  create  a  5  foot  high  ×  20 
foot long × 1 foot thick wall of flames, or a 5 foot 
high  ×  1  foot  thick  circle  of  flames  with  a  5  foot 
radius, or a pillar of fire 15 feet high with a 2 foot 
radius.  The  Adept  can  increase  any  single  dimen-
sion by  1  foot  /  Rank.  The  wall  must  be  anchored 
to a surface and cannot easily be seen through. Any 
object  or  entity  that  passes  through  the  wall  must 
resist  or  suffer  D10  (  +  1  /  Rank)  damage.  If  an 
entity  resists  the  damage  is  halved.  The  wall  will 
provide  light  equivalent  to  a  large  campfire.  At 
Rank  20  the  Adept  may  double  the  damage  of  a 
Wall of Fire by halving its duration. 

Wildfires (S-14) 

Range: Touch 
Duration: 20 minutes + 20 / Rank 
Experience Multiple: 250 
Base Chance: 25% 
Resist: Active, Passive 
Storage: Potion, Investment 
Target: Entity 
Effects: This spell imbues an entity with the speed 
and  essence  of  a  forest  fire.  The  target  may  run 
without  tiring  at  a  speed  of  20  (+  2  /  Rank)  miles 
per  hour  over  any  solid  surface.  Additionally  they 
may  “run”  across  any  substance  that  would  ordi-
narily  support  fire  (e.g.  across  tree  tops,  up  a 
wooden wall, or across oil). Like a fire, momentum 
effects  do  not  apply  to  the  target:  they  may  take 
corners  at  impossible  angles,  climb  a  flammable 
surface  impossibly  quickly,  or  stop  instantly  in 
place. The target may  leap a gap (as a fire  leaps a 
firebreak) a horizontal distance of PS (+ 1 / Rank) 
feet  or  a  vertical  distance  of  1/2  PS  (+  1  /  Rank) 
feet. The target will leave charred footprints behind 
them  but  will  not  normally  ignite  surfaces  they 
cross. 

An entity under Wildfires is moving too quickly to 
undertake  normal  tactical  combat  or  to  otherwise 
interact  with  other  entities.  If  they  wish  to  engage 
in melee they must dissipate the spell. 

The target must keep moving at least two hexes in 
any  direction  per  pulse.  If  they  stop  the  spell  is 
dissipated.  When  the  spell  ends  the  target  stops 
instantly in place. 

At  Rank  20  the  target  may  extend  the  duration  of 
the spell by suffering 1 EN damage / 10 additional 
minutes. 

21.7 Special Knowledge Rituals 

Create Drought (R-1) 

Range: 2 miles + 2 / Rank 
Duration: 8 hours + 8 / Rank 
Experience Multiple: 200 
Base Chance: 30% + 3% / Rank 
Resist: None 
Target: Area 
Material: Fire source 
Actions: Chanting and dancing 
Concentration Check: Standard 
Effects:  This  ritual  increases  temperature  in  the 
target area by 2C per Rank, slowly building up at 2 
degrees C per hour. This may impede rainfall, kills 
plants,  dry  up  wells,  ponds,  etc  and  ruin  crops  if 
repeated  regularly.  In  cold  or  icy  areas  climes  it 
may  make  an  area  tolerable  for  life  for  a  time  or 
result  in  standing  fogs  if  the  difference  between 
climes is too extreme. 

Flame Sight (R-2) 

Range: 5 feet 
Duration: 1 minute + 1 / Rank 
Experience Multiple: 250 
Base Chance: 40% + 5% / Rank 
Resist: None 
Material: Small fire 
Actions: Staring into flames 
Concentration Check: Standard 
Effects: The Adept must sit next to a fire and stare 
into  the  flames  for  half  an  hour.  They  may  then 
attempt one of three visions:  

73 

21 COLLEGE OF FIRE MAGICS 

•  A view as if the Adept was looking out of  a fire 
within  1/2  a  mile  (+  1/4  mile  Rank).  The  Adept 
may choose to look out of a specific fire, otherwise 
the  fire  is  randomly  determined  from  all  those  in 
range.  If  no  fires  are  within  range  the  Adept  will 
see  nothing.  They  may  move  their  viewpoint 
within the fire so as to see in any direction. There 
is no way to perceive that a fire is looked through 
by this ritual.  

•  A  precognitive  vision  which is  controlled  by  the 
GM.  The  detail  of  this  vision  and  the  amount  of 
information obtained increases with rank.  

• If the Adept chooses to sacrifice an object (which 
is destroyed) in the fire, they gain a vision relevant 
to  that  object’s past.  This  version  of  the  ritual  can 
only be performed on flammable objects. 

Ritual of Summoning a Lesser Efreeti (R-3) 

Range: 5 feet 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 200 
Base Chance: 30% + 3% / Rank 
Resist: None 
Target: Entity 
Cast Time: 3 Hours - 10 minutes / Rank (minimum 
10 minutes) 
Material: none 
Concentration Check: Standard 
Effects:  The  Adept  may  summon  a  Lesser  Efreeti 
to act as a steed for themselves. The Lesser Efreeti 
will  obey  all  its  rider’s  commands  while  it  is 
mounted. If the rider is not mounted or is out of its 
presence  then  it  will  only  follow  simple  passive 
commands (e.g. wait here, go there). At Rank 14 or 
higher it will obey commands which move it away 
from the rider (e.g. go attack people behind the hill, 
go and pick up Bob). 

Its characteristics and statistics are based on either 
the Adept or their rank in the ritual. 

PS = 30 (+ 2 / Rank)  MD = Adept 
EN = 30 (+ 1 / Rank)  AG = Adept 
FT = 30 (+ 2 / Rank)  WP = Adept 
PC = Adept 
MA = Adept 
Natural Armour WP / 4 (+ 1 / 3 or fraction 
Ranks) 
DEF WP + Ritual BC / 2 
TMR Walking = 5 (+ Rank / 4); Flying 8 (+ Rank / 
3) 
Movement Rate = 20 (+ 2 / Rank) miles per hour 
Size 1 (+ 1 / 10 Ranks) hexes 
Weapons  The  Lesser  Efreeti  has  three  physical 
attacks,  two  claws  and  a  bite.  These  have  a  base 
chance of 40% and do D + 4 (+ 1 / 2 Ranks) dam-
age.  It  also  has  a  65%  chance  of  igniting  in  battle 
with  flames  extending  in  a  5  foot  radius.  Entities 
within  the  radius,  except  for  the  rider,  must  resist 
or  suffer  D  (+  1  /  2  Ranks)  damage.  Entities  that 
resist suffer half damage. 

Abilities  The  Lesser  Efreeti  is immune  to  damage 
from  fire  but  takes  double  damage  from  ice  or 
water  and  magic  resistance  to  these  attacks  is  re-
duced  by  10%.  It  cannot  be  banished  while  the 
rider is mounted. 

At Rank 20 the ritual has extra benefits:  

•  the  adept  may  nominate  another  entity  as  the 
rider.  

• the Lesser Efreeti has Strength and Fatigue of 80 
and Endurance of 60.  

• NA is 8 (+ WP / 4).  

• damage from claws and bite is [D + 16].  

• the ritual’s duration is 24 hours. 

This  ritual  may  double  or  triple  duration  and  may 
backfire. It may be invested with a material cost of 
50,000 silver pieces per charge. 

Summon Fire Elemental (R-4) 

Range: 20 feet 
Duration: Concentration, no maximum 
Experience Multiple: 450 
Base Chance: MA + 4% / Rank 

21 COLLEGE OF FIRE MAGICS 

Resist: None 
Target: Fire Elemental 
Material: Large fire 
Actions: Concentration 
Concentration Check: Standard 
Effects: The  Adept may summon a Fire Elemental 
with a combined endurance and fatigue of 15 (+ 5 / 
Rank)  which  appears  within  the  fire.  The  Elemen-
tal is under the Adept’s control but strongly resents 
being summoned. 

If  the  ritual backfires  the  elemental  arrives  uncon-
trolled  and  will  immediately  attempt  to  kill  the 
summoner 

The  Elemental  is  returned  to  its  own  dimension  if 
the  summoner’s  concentration is broken,  it  is ban-
ished,  or  the  Adept  casts  a  Fire  Special  Counter-
spell. 

Boiling Water 

21.9 Some Useful Temperatures 
100°C 
113°C  Molten Sulphur 
Burning paper and meths 
200°C 
Tin melts 
232°C 
Boiling oil 
300°C 
327°C 
Lead melts 
419°C  Molten zinc 
800°C  Molten salt 
1000°C  Silver melts 
1053°C  Gold melts 
1083°C  Bronze and copper armour and weapons 

melt 

1600°C  Molten iron 
1800°C  Sand melts 
4000°C  Molten graphite 
>4000°C  Sauron’s jewellery 
 

 

21.8 Damage by Burning 
Occasionally  a  character  will  suffer  damage  from 
falling in to, or being exposed to, fire. The follow-
ing table indicates the pulse by pulse damage taken 
by an entity being in contact with fire. While igni-
tion  of  flammables  on  a  person  will  not  always 
occur, if an entity is immersed in fire (e.g. passage 
through a Wall of Fire) flammable possessions will 
catch  alight.  Unless  otherwise  specified,  Magical 
Fire will automatically ignite flammables. 

If  the  pulse  after  ignition is  spent putting  flamma-
bles  out,  no  further  damage  occurs.  In  general 
exposure  to  a  fire  source  indicates  exposure  for  a 
full  pulse,  therefore  standing  in  a  bonfire  might 
hurt  a  character  but  simply  jumping  through  one 
might not. This table should be used as a guide to 
how  much  damage  exposure  to  a  particular  fire 
source might cause. 

Clothing alight after passage through 
a fire 
Standing in campfire 
Standing in a one Hex bonfire 
Very large bonfire, house fire 
Raging forest fire 
Immersion in boiling water 
Immersion in boiling oil 
Immersion in molten metal or lava 
Sauron’s forge 

D + 3 

D + 5 
D + 10 
D + 15 
D + 25 
D + 35 
D + 50 
D + 100 
D + 150 

\end{Chapter}
