\begin{Tables}{Fright \& Awe Tables}

\begin{multicols}{3}
    
Rolls against these tables can made as the result of magic, when
meeting entities who are extremely ugly or beautiful, and in other
surprising situations.

\subsubsection{Modifiers}

All modifiers are cumulative lasting 24 hours or until 8 hours sleep,
but each division only counts once.  For example, a target gets feared
3 times.  The first is a hysterical which adds 10\%; second is
catatonic which adds 10\%.  Total so far is 20\%.  The third, however,
is another hysterical but as a hysterical component is already
included in the 20\%, the total remains at 20\%.  A minor heart attack
within 24 hours would push the total to 30\%.

\subsubsection{Familiarity}

In both the cases of extremely low and high beauty, the effects are
reduced with familiarity with the creature causing the fear or awe.
Once a character has successfully made their Willpower check, or has
recovered from a failed check, they need not check again (for an
effect from the same creature) for the remainder of the encounter.  If
the same creature is encountered again another Willpower check must be
made, but the GM may add 1 or more to the difficulty factor, thus
making it easier to succeed.  This bonus should not be applied to
encounters with other similar creatures, only to the
same ones (it is not true to say that when you have 
seen one Troll, you have seen them all). 

\subsubsection{PB Fright Checks}

Whenever characters encounter a creature whose Physical Beauty is less
than 5, and whose description states a Fear causing ability, they must
make a Willpower check to determine if they are frightened.  The
difficulty factor for this test is equal to the creature’s PB (use a
factor of 1 if PB is 0).  If this test is failed, the character must
then roll on the Fright Table (see \S\ref{table:fright}) and apply any
results before they take another action. They may attempt to recover
every pulse, by trying to succeed at the same Willpower check.  Until
that time they will act as the Fright Table indicates.

\subsubsection{PB Awe Checks}

Whenever characters encounter a creature whose Physical Beauty is
above 26, and whose description states an Awe causing ability, they
must make a Willpower check to determine if they are awed.

If this Willpower check is failed the character must then roll on the
Awe Table (see \S\ref{table:awe}) and apply any results before they
take another action. Recovery from awe effects is as indicated on the
Awe Table.  Awe can be caused by a creature with a Physical Beauty
greater than 26.  The range of the effect is line of sight to the
creature. The character will only be affected if the creature is
within their line of sight, and they are facing towards it.  Once
seen, however, facing becomes irrelevant, i.e. if the Target turns
away, so as to run, they do not lose the Awe effect simply because
they are no longer facing the source.  The effect will remain until
the Target is out of line of sight of the source. Should the character
return into line of sight of the source, the same result as before
will be applied, unless their Willpower has increased, in which case,
they receive another Willpower check to resist the effect of the Awe.

The difficulty factor of the Willpower check is dependent on the
creature’s PB:

\begin{dqtblr}{colspec={lXlX}}
PB	& Difficulty Factor	& PB	& Difficulty  Factor \\
27–28	& 4.0			& 32	& 2.0 \\
29	& 3.5			& 33	& 1.5 \\
30	& 3.0			& 34	& 1.0 \\
31	& 2.0			& 35+	& 0.5 \\
\end{dqtblr}

\end{multicols}

\section{Fright Table (Ver 1.1)}
\label{table:fright}

\begin{multicols}{3}

\begin{Description}

\item[< 20] Wary The target will not voluntarily approach the source
  of their fear. If they are not aware of the source they will be very
  cautious and seek to optimise safety.

\item[21--25] Berserk They immediately charge to attack the object of
  their rage. If the source is not apparent they will charge about
  noisily looking for it.  Add +10 to Strike Chance and −10 to Defence.

\item[26--75] Panic They will attempt to maximise their safety in
  relation to the source of their fear.  This usually involves fleeing
  as rapidly as possible, but could also include cowering in the
  centre of the party, curling up in a small ball, hiding under a bed,
  etc.  While a state of panic prevails, some sanity is present and
  the target would not normally do anything suicidal (e.g. running
  over the edge of a cliff) but they might use abilities to increase
  their safety (e.g. flying away).  If the target wishes to use an
  ability, (e.g. casting a spell) the GM should give a suitable
  negative modifier to their base chance (e.g. −20).


\item[76--90] Frozen They may take no action until snapped out of it
  (e.g. slapped on the face, attacked, etc).  The target can attempt
  to break out of it themselves by making a 1 × WP check per pulse.
  On recovery, the target rolls again at −30 (with no other modifiers)
  to determine their next action.  Add +10 to subsequent rolls on the
  fright table.

\item[91--95] Hysterical They stand and scream and may take no other
  action until snapped out of it (as for 76--90).  On recovery, roll
  again at −20 (with no other modifiers).  Add +10 to subsequent rolls
  on the fright table.

\item[96--100] Catatonic Target becomes catatonic.  Their hair turns
  white and they may take no other action until snapped out of it (as
  for 76--90). On recovery, roll again at −20 (with no other
  modifiers).  Add +10 to subsequent rolls on the fright table.

\item[101--110] Faints The target faints into unconsciousness and
  loses 5 Fatigue.  At the end of each minute they roll 1 × WP in
  order to regain consciousness.  Add +10 to subsequent rolls on the
  fright table.


\item[111--115] Collapses The target collapses into unconsciousness
  and loses all of their Fatigue.  After (30 − Endurance) minutes, or
  being tended by a Healer, they will regain consciousness.  All their
  Characteristics and Ranks will be reduced by 2, and they will not be
  able to recover Fatigue, until they have had comfortable bed rest
  for (40 − Endurance − tending Healer Rank) hours. Add +10 to
  subsequent rolls on the fright table.

\item[116+] Heart Attack The target suffers a heart attack and must
  receive the attention of a Healer of at least Rank 3 within
  Endurance pulses or they are dead.  If they survive they will be on
  0 Endurance and 0 Fatigue, and will be unconsciousness for (30 −
  Endurance) minutes.  All their Characteristics and Ranks will be
  reduced by 5, and they will not be able to recover Fatigue or more
  than half their Endurance, until they have had comfortable bed rest
  for (60 − Endurance − tending Healer Rank) hours.  Add +10 to
  subsequent rolls on the fright table.

\end{Description} 

\end{multicols}


\section{Awe Table}
\label{table:awe}

\begin{multicols}{3}

  \begin{Description}
\item[01--20] Awe Target is slightly awed. Will not voluntarily
  approach the source of the Awe, unless requested to do so by the
  source. If not aware of the source, will be slightly cautious.

\item[21--25] Enamoured Target is completely enamoured of the
  source. Will do anything that the source requests, to the extent of
  attacking comrades and friends, but not to the extent of killing
  themselves if so requested. If the source is not apparent they will
  rush around noisily looking for it.

\item[26--76] Panic Target is panicked by the Awe. They will attempt to
  maximise their safety, as they perceive it. This may involve
  fleeing, cowering, hiding, pleading and whimpering. They will not
  usually further endanger themselves in their attempt to escape, by
  running off a cliff for example. Add +5 to subsequent Awe Table
  rolls.

\item[77--90] Humble Target is completely humbled and prostrates
  themselves before the source of the Awe. They may take no other
  actions until they are snapped out of their grovelling by an outside
  agency, or by rolling less than or equal to their Willpower on
  D100. They may attempt this roll every second Pulse after being
  affected. Add +10 to subsequent Awe Table rolls.

\item[91--95] Hysterical Target becomes hysterical and falls to the
  ground, weeping, laughing, singing and/or praying as appropriate
  until snapped out of it by an outside agency. They may attempt to
  break out of it by themselves, by rolling less than or equal to
  their Willpower on D100, at the end of each minute following the
  affect. On recovery, roll D100: 1–50 slightly awed (as for 1–20,
  this table), 51–55 enamoured (as for 21–25), 56–100 panic (as for
  26–76). Add +10 to subsequent Awe Table rolls.

\item[96--100] Catatonic Target collapses, becomes catatonic and may
  take no further action until snapped out of it by an outside
  agency. Upon recovery, roll D100: 1--26 slightly awed, 27--31
  enamoured, 32--95 panic, 96--100 hysterical. Add +15 to subsequent
  Awe Table rolls.

\item[101--106] Faints Target faints dead away and will remain
  unconscious for [D + 6] minutes. Add +15 to all subsequent Awe Table
  rolls.

\item[107--110] Mild Heart Attack Target suffer a mild heart
  seizure. The result is the same as for 101-–106 except that the
  Target may not move about under their own power for the remainder of
  the day and suffers a decrease of 2 to PS, MD, AG, EN, and FT, until
  either they spend one month resting in bed, or their heart is
  repaired using the Healer ability “Repair Tissues and Organs”. Add
  +15 to subsequent Awe Table tolls.

\item[111+] Severe Heart Attack Target suffers a severe heart attack
  and must have the attention of a Healer of at least Rank 2 within
  one minute (12 pulses) or they will die. Otherwise this result is
  the same as for 107-–110, except that +20 is added to subsequent Awe
  Table rolls.

\end{Description}

\end{multicols}


\section{Spell Effects}
\begin{multicols}{3}
\begin{Itemize}

\item Wall of Bones (Necromancy): A target is not affected by the wall
  until they touch it. If they fail to resist they are affected any
  time they are in line of sight of the wall but facing is not
  important (if a target turns to flee the wall then they do not lose
  the fear because they are no longer facing it). If the target
  touches the same wall again they make another resistance
  check. Duration is that of the spell; once the wall is gone, so is
  the fear.

\item Fear (Necromancy, Wicca, Celestial): single target spell.

If duration is immediate then after (Rank of spell / 5) pulses,
rounded up, the target may make a 3 × WP check each pulse to recover
from fear, otherwise the fear lasts the duration of the spell.  A
further check may be required to recover from the secondary effects of
76--90, 91--95 and 96--100.  Once recovered from, the spell has no
further effect.  The range of the spell only determines the possible
targets, so the target being more than this distance from the Adept
will not affect the duration.

\item Mass Fear (Necromancy, Wicca): has set range and duration. Range
  is a sphere around the Adept.  Targets going out of range of the
  spell are no longer affected. Each time they re-enter the area of
  effect of the spell they make another magic resistance (and roll on
  the Fright Table if they fail).
\end{Itemize}

\emph{Example:} A group of adventurers are in the middle of a
forest. Rolf the Harrier gets struck by a Fear spell from Nasty the
Necro, and fails to resist.  A roll of 45 indicates panic. Rolf’s
player thinks Rolf has 3 options:
\begin{Itemize}
\item Run back to a clearing just left by the party. 

\item Feeling no direction to be safe, Rolf clings to Hu’ug the 
giant’s leg and trembles. 

\item Being a “If I can’t see them they can’t see me” believer, 
Rolf puts a sack over his head. 

Rolf’s player and the GM decide on percentages for the options and one
is determined. In this case Rolf flees. The spell is Rank 7 so Rolf
gets to attempt to snap out of the fear each pulse after the first
two. Six pulses after the spell is cast, on his third try, Rolf
succeeds his 1 × WP check and may turn around and attack Nasty the
Necro.
\end{Itemize}

\end{multicols}

\end{Tables}
