\begin{Chapter}{Fumble Tables}

When an attacker fumbles, they lose 10 from their Initiative Value
until the end of the next pulse.  Then they make a totally unmodified
D100 roll.  If that roll is under their current Initiative Value, they
suffer no further penalty for their slight fumble; if it is not under
their current Initiative Value, apply the corresponding result from
the appropriate table below. See (§6.10).

\section{Consequences}

A broken weapon is useless until repaired; a shattered weapon is
useless until reforged. Any combat spell on a broken or shattered
weapon is dissipated.

A damaged weapon is bent, dented, nicked, or similarly flawed.  You
may still use the damaged weapon but it does 1–2 points less damage
and has 0–20 penalty to its strike chance (GM decides these figures),
until repaired.

A damaged magical weapon must be repaired by a Weaponsmith of at least
Rank 6.  A damaged weaponsmithed weapon loses all its weaponsmith
bonuses to Strike chance and/or Damage, until repaired.

Any self-inflicted damage ignores your armour (including magic); it
usually represents bruising, minor strains, etc.  Naturally, take EN
damage if you have no FT available.  If a specific injury is stated
(e.g. pulled groin muscle, or broken bones), then healing requires a
lot of time or the appropriate minimum rank of Healer.

\section{Special cases}

An innately magical weapon  

• ignores asterisked results (*).  
• does not break.  
• may shatter or be damaged, but less often.  
• does not include any non-magical weapon merely 
under the effects of magic. 

Unarmed Combat  

•  “Shattered” = broken bone(s); lose 2 EN; useless in combat until healed.  
•  “Broken” = Seriously bruised; lose 2 FT; may be used  in  combat,  but  (until  healed)  each  successful 
blow does 2 damage points less to opponent and 2 FT to yourself.  
•    “Damaged”  =  Ouch!  Lose  2  FT;  no  further  effect. 

Strike Chance over 100 

If the fumble indicates a broken or damaged weapon, but your modified
Strike Chance was over 100, you have also struck your opponent; roll
[Fatigue] damage as normal.

\section{Bows and Crossbows only}

\begin{Description}

\item[01–12] Bowstring snaps and lashes you; lose 2 EN.

\item[13] Bowstring snaps and lashes you in the eye; lose 2 EN; you
  are blinded in one eye for 3 weeks or until cured by a Rank 7
  Healer. A figure who is blind in one eye suffers the following
  subtractions: 1 from MD, 2 from PB, 4 from Perception and reduces
  their base chance with any missile or thrown weapon by 30.

\item[14–29] Bowstring snaps; no further penalty.

\item[30] Traditional Hunting accident. Clumsy release causes arrow /
  quarrel to fly towards a random “friendly” back in approximately the
  same direction as you were aiming: Strike Chance = Weapon BC +
  weapon bonuses + 30 - target’s defence.

\item[31–33]  Clumsy release; bolt/arrow flies wide missing friends and foes. 

\item[34–36] Brief twinge of pain in your arm or back; Lose 1 EN.

\item[37–39] as per 34-36, but lose 2 EN.

\item[40–59] Dropped bolt or quarrel.

\item[60–99] Bowstring snaps; no further damage.

\item[00] Bowstring snaps and lashes you; lose 2 EN.

\end{Description}


\section{All Other Weapons}

The following results are generalised.  Therefore the GM is free to
ignore or downgrade any result which is inapplicable to a specific
case.  Some outcomes are avoidable through a successful characteristic
check.

\begin{Description}
  
\item[01–09] Shattered weapon.

\item[10] Shattered weapon; some slivers fly at you, potentially
  causing you a grievous injury — roll on the Grievous Injury Table
  (§51), but ignore any result over 13.

\item[11–12] Shattered weapon, flying splinters; you and your
  opponent(s) lose 1 EN each.

\item[13] Playing the Roman fool? You just did yourself an Endurance
  blow; fortunately you rolled minimum damage (but don’t forget the
  extra damage from poison, magic, etc).

\item[14–16] Your wild swing possibly connects with someone other than
  your intended target or yourself – immediately make a strike check
  at your new victim, the nearest being in range other than you or
  your intended target.  Hope you weren’t mounted.

\item[17–18] Lose 1 EN. Feels as if you pulled something.


\item[19] Lose 2 EN. You really pulled something.

\item[20] Oops! You’ve flung your weapon in a high parabolic
  arc. Normally a flung or dropped weapon falls without hurting anyone
  — however, in this case, it falls on a random target, possibly even
  you, and maybe hurts them: Strike Chance = [Weapon’s BC] + [magical
    / weaponsmith bonuses] - [random target’s defence].

\item[21–26] Butterfingers! Make 3 × MD to avoid your weapon flying
  2–3 hexes in a random direction.

\item[27–28] Klutz! Make 3 × MD to avoid dropping your weapon in your
  hex.

\item[29] Whoops! You’ve caught your weapon in your own armour or
  gear. You may choose to automatically free it in the next pulse, in
  which case you may not attack or cast magic until after the end of
  the next pulse. Or else you may wish to prepare another weapon in
  your next action.

\item[30] Overly enthusiastic lunge. GM moves you to an unoccupied
  forward hex (make 3 × AG to choose your new facing) — but if no
  empty hex is available, you just tried to close on an opponent, who
  gets a free chance to keep you out of close. If you did close, you
  don’t have to drop any non-close weapon, but it may not be used to
  attack effectively.

\item[31–32] Poor balance; make 3 × AG or no offensive action until
  after the end of the next pulse.

\item[33] Stumble; make 3 × AG or fall prone.

\item[34–35] as per 31–32, but make 2 × AG.

\item[36] as per 33, but make 2 × AG.

\item[37–38] as per 31–32, but make 1 × AG.

\item[39] as per 33, but make 1 × AG.

\item[40] Broken weapon.

\item[41–49] Damaged weapon.

\item[50] Momentary dizziness; make 3 × EN or you may not attack or
  cast magic until after the end of next pulse.

\item[51–52] Your vigorous swing causes a slight twinge.  Make 3 × EN
  or lose 2 FT.

\item[53] Make 3 × EN or pull a groin muscle, lose 2 FT and have half
  Base TMR until healed.

\item[54–55] as per 51–52, but make 2 × EN.

\item[56] as per 53, but make 2 × EN.

\item[57–58] as per 51-52, but make 1 × EN.

\item[59] as per 53, but make 1 × EN.

\item[60] No effect unless you used a A or B-class melee weapon
  against an opponent with a non-magical shield. In which case, you
  have spectacularly wedged you weapon into their shield. Make 1 × PS
  to immediately wrench your weapon out, or it will be torn from your
  grasp in the fracas. Don’t worry if you fail — perhaps their shield
  is now useless?

\item[61–62] Your melee weapon is stuck, caught, or entangled in your
  opponent’s armour or gear (and you didn’t even hurt them). Make 3 ×
  PS to immediately disengage your weapon, or it will be yanked from
  your grasp in the fracas.

63   You palpably hit a tree, rock, wall-hanging, 
furniture, or some other adjacent “scenery”. 
Make 3 × PS to immediately disen-
gage/extract your weapon. You may try 
again, as a future action; but perhaps you 
should prepare a new weapon instead. 

\item[64–65] as per 61-62, but make 2 × PS.

\item[66] as per 63, but make 2 × PS.

\item[67–68] as per 61-62, but make 1 × PS.

\item[69] as per 63, but make 1 × PS.

\item[70*]   Shattered weapon, if it is not at least Rank 1 
weaponsmithed. 

\item[71–73*] Your weapon breaks unless you roll under its weaponsmith
  rank on D10. Indeed it shattered if you failed the roll by 5 or
  more.

\item[74–77] Twinge of pain. Take (D10 - rank in weapon) FT damage.

\item[78–79] as per 74–77, but also you may not attack or cast magic
  for remainder of the pulse.

\item[80] Your weapon flies from your grasp. You may choose to drop
  whatever is in your other hand; in which case, make 3 × MD to catch
  the weapon in that other hand.

\item[81–82] Butterfingers! Make 3 × MD to avoid your weapon flying
  2–3 hexes in a random direction.

\item[83] Klutz! Make 3 × MD to avoid dropping your weapon in your
  hex.

\item[84–85] as per 81-82, but make 2 × MD.

\item[86] as per 83, but make 2 × MD.

87–88  as per 81–82, but make 1 × MD. 

\item[89] as per 83, but make 1 × MD.

\item[90] Broken weapon.

\item[91–99] Damaged weapon.

\item[00] Your bizarre but highly spectacular fumble is mistaken for
  an obscure martial technique.  All engaged melee opponents hastily
  elect to neither attack or cast magic as their next action. If you
  have another action before they actually perform their next action,
  you may choose to run away (retreat up to full TMR) as your action
  without the need for a Withdrawal manoeuvre — you are no longer
  engaged with those particular opponents.

\end{Description}

\end{Chapter}
