\begin{Chapter}{Health and Fitness}

A character’s Fatigue will vary depending upon the 
amount of food and rest they get compared to their 
activities. 

A  character’s  Endurance  may  be  temporarily  re-
duced  by  lack  of  sustenance,  extreme  activities, 
damage, or illness. 

4.1 Eating and Drinking 
The  amount  of  food  and  water  required  per  day  is 
dependent  on  many  factors.  These  include  the 
person  (endurance,  weight,  build,  metabolic  rate 
and race) and the level of activity they are involved 
in (light, medium, hard or strenuous). 

On  average  1  lb  of  food  and  2  pints  of  water  per 
day is required. 

4.2 Starvation 
Starvation  occurs  when  a  character  does  not  have 
at least 1 nourishing meal a day. 

If a character is starved they will have their Fatigue 
maximum and Endurance temporarily reduced by 1 
each day. This decrease will last until the character 
starts receiving proper nourishment. 

A  starved  character’s  Fatigue  maximum  and  En-
durance will recover by 1 point each day, after the 
first, that they receive proper nourishment 

4.3 Dehydration 
Dehydration occurs when a character does not have 
at  least  2  pints  of  water  a  day.  This  amount  will 
increase  in  high  temperatures  by  1  pint  per  10 
degrees above 20. 

If  a  character  is  dehydrated  they  will  have  their 
Fatigue  maximum  and  Endurance  temporarily 
reduced  by  5  each  day.  If  the  character  receives 
part  of  their  water  requirement,  the  penalty  is  re-
duced.  For  every  20%  (or  fraction)  less  than  the 
daily  requirement  they  lose  1  from  FT  max  and 
EN. This decrease will last until the character starts 
receiving adequate quantities of water. 

A  dehydrated  character’s  Fatigue  and  Endurance 
maximums will recover by 5 points each day, after 
the  first,  that  they  receive  adequate  quantities  of 
water. 

4.4 Tiredness and Rest 
Characters  have  a  tendency  to  lose  Fatigue  points 
on  adventure.  A  fatigued  character  must  rest  to 
recover  Fatigue  points.  Sleep,  as  might  be  ex-
pected,  is  the  best  way  to  become  refreshed,  but 
food and rest will also help. 

The  Fatigue point loss and recovery  rates given in 
these  rules  assume  that  the  character  is  in  good 
health  and  is  well  fed.  If  the  character  is  not  in 
good  condition,  the  GM  may  adjust  the  effects  of 
activity,  the  effects  of  weight  carried  and  the  rate 
of recovery. 

Fatigue Loss 

A  character  can  lose  Fatigue  points  when  they 
engage  in  any  activity  more  stressful  than  a  lei-
surely walk. 

There are four classes of activity which can fatigue 
a character: 

1. Light Exercise includes moderate to brisk walk-
ing, riding slowly or at a moderate pace on a docile 
mount, etc. 

2.  Medium  Exercise  includes  jogging,  riding  on  a 
cantering  mount,  light  construction  or  precision 
work, etc. 

3. Hard Exercise includes paced running, riding at 
a gallop, hard manual labour, etc. 

4.  Strenuous  Exercise  includes  constant  sprinting, 
breakneck  riding,  and  generally  those  actions  with 
which the character pushes their body to its practi-
cal limits. 

It  is  possible  for  a  character’s  actions  to  be  more 
taxing  than  Strenuous  Exercise,  which  requires 

superhuman  exertion.  This  Fatigue  loss  from  this 
activity will be determined by the GM. 

A  character’s  degree  of  exertion  is  judged  each 
hour. 

The  GM  should  indicate  to  players  the  level  of 
exertion of their activities (averaging where neces-
sary).  If  the  GM  gives  consistent  guidelines  the 
players  will  be  able  to  keep  an  ongoing  track  of 
fatigue loss. 

Encumbrance 

A character is limited in the weight they can bear, 
and  may  become  fatigued  if  they  engage  in  exer-
cise. 

The  Fatigue  and  Encumbrance  Table  (§58.1)  lists 
the maximum weight a character may carry. 

A  player  must  determine  the  total  weight  their 
character is carrying if the character is to engage in 
light  or  more  stressful  exercise  for  a  significant 
length of time during a day. 

When  an  entity  has  a  Physical  Strength  value 
greater  than  40,  the  GM  divides  that  value  by  40. 
Multiply  the  quotient by  the  entry  for  40,  and  add 
the  entry  corresponding  to  the  remainder  to  deter-
mine that entity’s capabilities. 

Damage 

A  character  may  lose  Fatigue  by  being  damaged. 
This  may  be  recovered  naturally  or  by  being 
healed. 

Spell Casting 

A  character  may  lose  Fatigue  by  using  magical 
abilities. This may be recovered naturally but may 
not be healed. 

Calculating Current Fatigue 

The  Fatigue  status  of  a  character  only  needs  to  be 
calculated  before  they  enter  into  combat,  wish  to 
perform magic  or  if  they  perform  fatiguing  activi-
ties  for  long  periods.  To  calculate  current  Fatigue 
use the Fatigue and Encumbrance Table (§58.1): 

1.  Cross-reference 
Strength and the weight they are carrying. 

the 

character’s  Physical 

2.  Read  down  this  column  until  it  intersects  with 
the  row  corresponding  to  the  character’s  rate  of 
exercise. 

3.  Multiply  the  resulting  number  (Fatigue  points 
lost per hour) by the number of hours at this exer-
cise level. 

4.  Perform this  calculation  once  for each time  one 
(or more) of the three factors changes. 

5. Add each separate subtotal to determine the total 
Fatigue points expended by the character so far. 

Exhaustion 

If a character’s Fatigue point total is reduced below 
zero, they are exhausted. An exhausted character is 
limited in the activities they may choose to do and 
the  performance  of  their  abilities  is  adversely  ef-
fected.  Their  Fatigue  is  considered  zero  for  the 
purposes of combat or magic use. 

A  character  may  choose  to  exert  themselves  after 
their  Fatigue  points  are  reduced  to  zero  until  they 
have  expended  a  nominal  one-half  their  initial 
Fatigue points (round down). When they reach this 
limit  they  will  collapse  unless  they  succeed  a  1  × 
WP check every (2 × Endurance) minutes. 

An  exhausted  character  must  sleep  for  as  much 
time  as  they  were  performing  any  exercise  while 
exhausted  before  they  may  recover  any  Fatigue 
points. 

If  an  exhausted  character  wishes  to  engage  in 
Strenuous Exercise, they must succeed a separate 1 
× WP check. 

Exhaustion Modifier 

The  character  must  subtract 1  /  half  hour  (or  frac-
tion) of exhaustion to any base chance. 

13 

4 HEALTH AND FITNESS 

4.5 Fatigue Recovery 
A character may regain Fatigue points naturally by 
eating a hot meal or resting. 

A character may never have a Fatigue total greater 
than their Fatigue Characteristic. 

A  character  naturally  recovers  Fatigue  points  as 
follows: 

Activity 

Fatigue points / Hour 

2 
1 
2 
3 

Eat Hot Meal 
Relaxation 
Nap  
Sleep  
1.  A  character  may  benefit  from  a  hot  meal  no 
more than three times during a 24 hour period, and 
each meal must be separated by at least 4 hours. 

2. A character that does not get at least 6 hours of 
rest  and/or  sleep  per  day  will  have  their  Fatigue 
maximum  temporarily  reduced  by  1  FT  /  hour  (or 
fraction)  of  sleep  under  6  hours.  This  may  be  re-
covered at the rate of 1 FT / 4 hours sleep. 

3.  If  a  character’s  Endurance  is  less  than  10,  they 
recover  one-half  of  a  FT  point  less  per  hour  or 
meal,  and  if  their  Endurance  is  less  than  5,  they 
recover  one  less  FT  point.  However,  a  character 
always recovers a minimum of one-half a FT point 
when resting. 

4. If a character’s Endurance is from 21 to 30, they 
recover  an  additional  one-half  of  a  FT  point  per 
hour or meal. Each succeeding ten point Endurance 
bracket  carries  an  additional  one-half  FT  point 
bonus. 

5. Fatigue loss from damage may also be recovered 
by  magical  healing  (but  not  the  Healer  skill  Heal 
Endurance). 

4.6 Damage and Illness 

Effects of Low Endurance 

Unconsciousness  When  an  entity’s  Endurance 
reaches 3 or less, they must make a (current EN) × 
WP  check  or  fall  unconscious;  this  WP  check  is 
repeated every minute or if their EN changes. 

An  entity  on  0  Endurance  is  unconscious,  but  sta-
ble.  An  entity  with  a  full  Endurance  of  5  or  less 
does not make consciousness checks. They remain 
conscious until they fall to 0 or less Endurance. 

Below  Zero  Endurance  An  entity  on  negative 
Endurance  will  lose  one  point  of  Fatigue  (Endur-
ance when no Fatigue remains) until the bleeding is 
stanched  by  a  Healer,  or  until  dead.  They  will 
continue  to  take  damage  from  any  further  blows, 
spells,  grievous  wounds  which  are  bleeders,  burn-
ing, etc. 

When  an  entity  is  below  zero  Endurance  they  are 
on the very brink of death. It takes time and skill to 
tell the difference between this state and death (e.g. 
empathy,  DA).  GMs  should  not  let  players  take 
advantage  of  out  of  character  information  when 
another player’s character is below 0 Endurance. 

Death  When  an  entity’s  Endurance  falls  below 
negative  one-half  their  full  Endurance,  they  are 
dead.  Once  dead,  ongoing  damage  (e.g.  poison  or 
bleeding)  ceases  but  further  damage  may  be  in-
flicted on the body. 

Endurance Recovery 

There  are  many  causes  of  a  character  losing  En-
durance  points.  However,  once  lost  there  are  two 
primary methods of recovering them. 

Healers and Magical Healing 

Healers,  herbalists,  potions,  medicines  and  some 
magics  may  aid  the  recovery  of  Endurance.  The 
exact  effects  can  be  found  under  the  appropriate 
skill or magic. 

4 HEALTH AND FITNESS/ 5 EQUIPMENT AND MONEY 

Natural Healing 

Natural Healing of Grievous Injuries 

The  rate  at  which  Endurance  Points  recover  natu-
rally  primarily  depends  on  how  active  the  injured 
being is. 

If an entity is resting they regain 1 Endurance point 
every three full days. 

This rate is reduced to 1 / 4 days if the entity:  

• takes any further EN damage  

• uses more than half their FT  

• does not receive adequate nourishment 

If  an  entity 
is  given  ministrations  from  a 
physicker’s kit, their body requires one  less day to 
regain an Endurance Point. 

Injuries  which  are  not  quantified  as  Endurance 
point  losses  or  grievous  injuries  (e.g.  hamstrung 
muscles)  heal  at  the  same  rate  as  they  do  in  this 
world. 

These  healing  rates  are  based  on  average  Endur-
ance  value  of  15.  The  GM  may  chose  to  increase 
the healing rate if an entity’s full Endurance is very 
high or decrease it for a low Endurance entity. 

Potions \& Unconscious Patients 

An entity cannot drink a healing potion when they 
are  unconscious  or  below  zero  endurance  but  one 
can be massaged down their throat. The chance of 
doing  this  is  equal  to  the Manual  Dexterity  +  Per-
ception of the person administering the potion, or if 
a healer, 90 + Healer Rank. If successful then D10 
per 10 points of the healing potion’s curing (round 
down) will be received. If the person fails the roll, 
the  potion  is  wasted,  but  no  harmful  effects  occur 
to the patient. 

Grievous Injuries 

Endurance  loss  resulting  from  specific  grievous 
injuries  may  not  be  healed  separately  from  the 
underlying  specific  injury.  When  the  specific  in-
jury  is  fully  cured  the  related  endurance  is  recov-
ered automatically. 

Major  injuries  take  a  long  time  to  heal  and  some 
will  not  heal  naturally  but  require  a  healer.  Here 
are guidelines for the healing requirements of some 
common major injuries. 

Broken  bones  will  knit  in  4  weeks  for  a  simple 
fracture,  or  up  to  10  weeks  for  a  compound  frac-
ture.  A  bone  must  be  properly  set  before  the  bone 
may knit together. 

Internal  injuries  an  entity  will  usually  die  from 
internal  injuries.  If  the  patient  is  comfortable,  un-
moving,  and  kept  alive  by  a  healer  or  physician, 
internal  injuries  will  heal  1  Endurance  point  per 
week 

Open  wounds  will  heal  at  half  the  normal  rate, 
provided that they are kept free  of infection. Open 
wounds will leave scars. 

Removed  body  parts  will  not  regrow  naturally. 
However,  the  remaining  wound  will  heal  over  at 
quarter  the  normal  rate,  provided  it  is  kept  free 
from infection. 

Magical  healing of  Specific  Injuries  Healers  and 
certain magics may heal specific injuries. The time 
taken  and  effects  of  these  magics  may  be  found 
under the appropriate skill or magic. 

4.7 Infection 
If  a  character  is  wounded  there  is  the  possibility 
that they  have  become  infected  as  a result  of  their 
wounds. 

An  Infection  Check  must  be  performed  to  deter-
mine whether they are infected or not. 

Becoming Infected 

The  chance  of  becoming  infected  depends  on  the 
entity’s health, the type of injury, and the environ-
ment the entity is in. Modifiers are cumulative one 
is applied from each category: 

There is a wound which is ... 
Dirty 
+20%  
Heavily contaminated  +50% 
The environment is ... 

5 Equipment and Money 

-5%  
+20% 

Dry 
Humid 
The average temperature is ... 
Below 0 
1 – 5 
30 – 40 
Above 40 
Some  specific  grievous  injuries  also  increase  the 
chance of infection. 

+20% 
+10% 
+10%  
+20% 

Effects of Infection 

An entity with an infection will be slowly poisoned 
by the infection. The damage is [D - 5] Endurance 
per  day,  until  the  infection  is  cured.  An  infected 
wound will not heal until the infection is cured. 

Curing Infection 

There are two ways to recover from infection. The 
first  is  to  tough  it  out.  The  second  is  to  be  healed 
by a healer. 

Toughing it out An infected character may make a 
1  ×  Endurance  check  every  day  to  recover  natu-
rally. 

Healing An infected character may be cured by the 
arts of a Healer or by magic. The rank at which this 
is possible, and the chance of success can be found 
under the appropriate skill or spell. 

4.8 Conception 
The natural conception chances for character races 
are:  

3% 
1% 
4% 
2% 
6% 
10% 
5% 

Dwarf 
Elf 
Halﬂing 
Hill Giant 
Human 
Orc 
Shapechanger 
Checks against the relevant chance should be made 
no more often than once per 48 hours of appropri-
ate activity. 
\end{Chapter}
