\begin{Skill}[1.4]{healer}{Healer}

The healer skill is pseudo-magical and healers are able to cure all
physical ills and perform miracles.  It is a highly skilled profession
and healers are not common.  However their existence means the health
and life span of people in the DQ world are considerably better than
their medieval counterparts.

A healer’s empathy often gives them a distaste for causing pain to
others.

A healer will charge whatever their client can afford for their lower
Ranked abilities.  The charge for a miracle (the performance of an
ability Rank 8 or greater) will normally exceed 2000 Silver Pennies.

For the purposes of the Healer skill, an entity’s body parts are the
head, torso, and each limb.  Vital Organs are the heart, stomach,
viscera (liver, small and large intestines and kidneys), genitalia,
brain, and eyes.

A healer may also use their abilities upon animals which they have
Beastmaster (\S\ref{skill:beastmaster}) familiarity.  However
non-sentients cannot be resurrected.

The section \S\ref{health} Health and Fitness details the effects of
injury and illness on patients.

\subsection{Field Operations}

Once a healer begins work on curing a patient the condition of the
patient is “stabilised” while that curing is continuing. This means
that no Endurance or Fatigue loss will occur for the condition that is
being cured.  Other afflictions will be unaffected.  After each
healing attempt a pulse effectively passes prior to any other attempt
beginning.

\subsection{Potions \& Unconscious Patients}

An entity cannot drink a healing potion when they are unconscious or
below zero endurance but one can be massaged down their throat. The
chance of doing this is equal to the Manual Dexterity + Perception
of the person administering the potion, or if a healer, 90 + Healer
Rank. If successful then D10 per 10 points of the healing potion’s
curing (round down) will be received. If the person fails the roll,
the potion is wasted, but no harmful effects occur to the patient.

\section{Benefits}

A healer gains specific abilities at each Rank as per 
the following table:

\begin{dqtblr}{colspec={m{3em}X}}
0 & Empathy, Ranged Empathy (optional) \\
1 & Cure Infection, Disease, Headaches, Fever \\
2 & Soothe Pain, Prolong Life \\
3 & Heal Endurance, Transfer Fatigue \\
4 & Neutralise Poison, Cure Burns and Repair Scars \\
5 & Repair Muscle, Preserve Dead \\
6 & Repair Bones \\
7 & Repair Tissues and Organs \\
8 & Resurrect the Dead \\
9 & Regenerate Limbs and Joints \\
10 & Regenerate Trunk, Head and Vital Organs \\
\end{dqtblr}

NB.  A healer must choose at Rank 0 whether or not to learn Ranged
Empathy.

\section{Restrictions}

A healer must expend as many Fatigue Points as the Rank at which they
acquired the ability they are using (except Empathy, Ranged Empathy).

A healer may use any of their abilities (with the exception of
resurrection) upon themselves.

A healer must “lay hands” (place their hands) on an entity on whom
they are to use any of their abilities (except Ranged Empathy).

A healer has the following modifications to their combat strike
chances:

\subsubsection{Tactile Empathy (only):}

\begin{dqtblr}{colspec={lX}}
−1 / 2 Ranks	& Close Combat strike chance \\
−1 / 5 Ranks	& Melee Combat strike chance \\
\end{dqtblr}

\subsubsection{Ranged Empathy:}

\begin{dqtblr}{colspec={lX}}
−1 / Rank	& Close Combat strike chance \\
−1 / 2 Ranks	& Melee Combat strike chance \\
\end{dqtblr}

\section{Ability Descriptions}

\subsection{Empathy}

\basechance{automatic}
\timetaken{5 seconds}
When a healer lays on hands they immediately invoke empathy.

A healer uses empathy to identify which of the healing abilities is
required to heal the patient.

The healer automatically detects the surface emotions of the entity
being healed. An entity’s surface emotions are those which currently
occupy their conscious mind. The GM informs the healer of the general
feelings of the being with which they have empathy.

\subsection{Ranged Empathy}

\basechance{Perception + 10 / Rank}
\timetaken{5 seconds}
If the healer has learnt Ranged empathy they may attempt to detect the
surface emotions of an entity no more than (2 × Rank) feet away from
them at a cost of 1 Fatigue Point. If the entity actively resists then
subtract twice the target’s Willpower from this success chance.

\subsection{Cure Infection, Disease, Headaches and Fever}

\basechance{15 × Rank + Patient’s Endurance} \timetaken{30 minutes \mi
  2 / Rank} A healer neutralises poisons and cures fevers and diseases
in much the same manner that medicines and antidotes (see
\S\ref{alchemist:benefits}) do.

If the healing attempt is unsuccessful the patient subtracts 10 from
their next die roll to see if they naturally recover from their
affliction.

\subsection{Soothe Pain}

\basechance{90 + Rank}
\timetaken{60 seconds − 5 / Rank}
\duration{Rank squared hours}
When a healer uses their soothe pain ability, they numb their
patient’s nervous system so that it will not transmit pain sensations
to their brain.  The ability also has a soporific effect upon the
patient, so that they will not inadvertently injure themselves while
unable to distinguish hurtful actions.  The GM may, at their
discretion, permit the healer to use this ability as if they had fed
or injected their patient with a local or general anaesthetic,
tranquilliser, etc.

\subsection{Prolong Life}

\basechance{90 + Rank}
\timetaken{60 seconds − 5 / Rank}
When a healer uses the prolong life ability add D10 × (Healer’s Rank +
Patient’s Endurance) days to the life of their patient. A patient’s
life may not be prolonged to over three times their natural life.  An
entity with a prolonged life has a reduced chance of resurrection.

\subsection{Heal Endurance and Transfer Fatigue}

\basechance{90 + Rank}
\timetaken{11 minutes − 1 / Rank}
Heal Endurance will cure the patient of [D + Rank − 5] Endurance
points.  It will not heal damage associated with a specific grievous
injury.

When a healer transfers fatigue the patient gains one Fatigue point
for each point the healer expends (above the fatigue cost to use the
ability).

An entity may never have more Fatigue or Endurance Points than the
value of the relevant characteristic and excess points cured have no
effect upon the patient.

\subsection{Neutralise Poison}

\basechance{90 + Rank (natural) or 50 − 5 × difference in Rank (synthetic)}
\timetaken{5 seconds}
A healer may neutralise the effects of a natural venom or the effects
of a synthetic poison created by an alchemist of equal or lesser Rank.
If a synthetic poison is produced by an alchemist of greater Rank they
must roll the second Base Chance above.

\subsection{Cure Burns and Repair Scars}

\basechance{90 + Rank}
\timetaken{30 minutes − 2 / Rank}
A healer may immediately halt ongoing burning damage, and prevent
scarring from a fresh burn.  They may also reduce the effects of old
scars; minor scars can be completely removed, but major scars would
take (2 × PB Modifier) successful cures to be removed.

\subsection{Repair Damage}

\basechance{90 + Rank}
\timetaken{50 hours − 3 / Rank}
A healer may repair torn, damaged, or broken muscles, bone, tissues
and organs. Generally these abilities will be used to repair the
effects of Grievous Injuries.

A Healer may repair all injuries to one body part at the same time. To
use Repair, a body part must be mostly intact and no more than half
damaged.  Regeneration is required for more significant healing.

A healer can act as a cosmetic surgeon. First, they sedate their
patient with the soothe pain ability.  They then slice and reshape the
skin, muscles, and bones which are deemed unsightly, and make them
whole with the appropriate repair ability.  Unless the healer has also
learnt regeneration, it is best that they work with a partner.

\subsection{Preserve Dead}

\basechance{90 + Rank}
\timetaken{60 minutes − 5 / Rank}
A healer can suspend the time limit on resurrection by preserving the
dead body of a being. Each time a healer uses the preserve dead
ability, the body will not “age” for a number of days equal to the
healer’s Rank. This ability may be repeated by the same healer on the
same body.

\subsection{Reattach Body Parts}

\basechance{90 + Rank}
\timetaken{30 minutes − 2 / Rank}
A Healer able to Repair Muscle may reattach a severed body part.  The
reattached body part may be Repaired to full functionality provided
the wound meets the requirements of repair and it was severed for no
more than Rank minutes. Or it may be Regenerated to full functionality
provided it was severed for no more than Rank hours.

\subsection{Resurrection}

\timetaken{60 minutes − 5 / Rank}
\basechance{Patient’s Endurance + 8 / Rank, minimum = Rank \& maximum = 90 + Rank regardless of the total modifiers}

Base Chance Modifiers: 

\begin{dqtblr}{colspec={Xr}}
Healer is life aspected			& +5 \\
Patient is life aspected		& +5 \\
Healer is death aspected		& −5 \\
Patient is death aspected		& −5 \\
Per year (or fraction) the patient’s life has been prolonged & −1 \\
Per day of regeneration it would normally require to make the body whole & −1 \\
Body is whole but has suffered Damage points equal to or greater than twice its Endurance (including after death damage) & −10 \\
Per unsuccessful resurrection attempt since patient died  & −10 \\
\end{dqtblr}

A resurrection will cure the body of all ills and damage done to it
provided that Rank 8 healing or below would have been sufficient had
the patient been alive.  For example, poison and non-specific wounds
will be cured automatically.

If the resurrection is successful, the patient is resurrected with
their body whole.  Their Endurance characteristic is decreased by one,
although all of their other characteristics remain as before they
died.

After a resurrection the patient will have 1 Endurance point and 0
Fatigue.  The Endurance is considered to be grievous damage and the
Fatigue loss is deemed to be due to tiredness. This means that the
Fatigue loss may only be recovered by sleep, rest, hot meals or some
form of fatigue transfer and the Endurance loss may be cured by a
healer, magic, or by letting the body heal itself naturally.

If the resurrection is unsuccessful the patient is not resurrected and
their Endurance characteristic is decreased by one.  The body is
preserved for one full day after the attempt. When an entity’s
Endurance is reduced to zero or less, that entity may no longer be
resurrected.

If the roll for resurrection is equal to or greater than (90 + Rank),
the healer has summoned a malignant spirit, rather than the patient’s
life-force. The spirit will drain the healer’s Endurance
characteristic by [D − 5].  The spirit will then return to the
netherworld.

\subsubsection{Restrictions:}
\begin{Itemize}
\item A healer may attempt the resurrection of an entity who is less
  than 10 × Rank hours dead.

\item A healer must have a body part at least the size of a torso to
  attempt the resurrection of an entity.  A healer will not succeed if
  they attempt the resurrection of a living being from a severed body
  part (there is only one life force).  If a body is completely
  destroyed (perhaps burned), which prevents the resurrection of the
  entity, the entity may become a revenant.

\item If the patient has wounds that require regeneration (Rank 9 or
  10) healing, these need to be healed separately.

\item Most vital organs will need to be healed prior to the
  resurrection otherwise the body will die again immediately (the
  notable exception being the eyes).

\item The healer need not know what the patient looked like since the
  healing of the body is governed by its own characteristics.  Hence
  any changes that had been made to the body (for example facial
  changes or embedding of items) will be gone after the resurrection.

\item A player may take no action with their character’s dead body.

\item In rare instances a healer may be able to resurrect a life force
  into a different body.  The resurrected entity has the physical
  characteristics of the new body and the mental characteristics from
  the life force. It will take some months for the entity to get used
  to the new body and this will affect base chances of physical
  abilities.  The GM will advise the specifics.

\item A body that has been animated (e.g.  Zombie) may still be
  resurrected provided it is no longer animated and all the other
  conditions have been met (for example length of death, condition,
  etc.).  Note that the Healer Preserve Dead will not affect Zombies.
\end{Itemize}

\subsection{Regeneration}

\basechance{90 + Rank}
\timetaken{50 hours − 3 / Rank (refer below)}
A healer can regenerate every portion of an entity’s body including
vital organs and severely damaged or severed body parts.  A healer
must regenerate each vital organ or body part separately.

Regeneration will regenerate [Healer Rank] percent of a missing body
part or vital organ per day costing the target [Healer Rank] FT per
day.  A Healer must initiate the Regeneration of a body part or vital
organ, the regeneration will then continue unattended until the
regrowth is complete or the healing area sustains further significant
damage.

Aregenerated vital organ will immediately begin to function if enough
of the rest of the entity’s body is in working order.  Otherwise, the
vital organ will be dormant until the healer can repair or regenerate
the necessary body parts. The regeneration time does not need to be
consecutive, but the damaged part will not function nor continue
growing until the regeneration time is complete.

\section{Potions}

A healer can manufacture certain potions in conjunction with an
alchemist.  See Alchemist (\S\ref{alchemist:benefits}) for more
details.

\begin{table*}[t]
\section{Healer Summary Chart}

\medskip

\begin{dqtblr}{colspec={|X|X|X|},%
    vlines={green7},%
    hline{1-2,16}={1.2pt}}
Ability						& Base Chance			& Time \\
Empathy						& Automatic			& 5 seconds \\
Ranged Empathy					& PC + 10 / Rank		& 5 seconds \\
Cure Infection, Disease, Headaches \& Fever	& (15 × Rank) + Patient’s EN	& 30 minutes − 2 / Rank \\
Soothe Pain					& 90 + Rank			& 60 seconds − 5 / Rank \\
Prolong Life					& 90 + Rank			& 60 seconds − 5 / Rank \\
Heal Endurance					& 90 + Rank			& 11 minutes − 1 / Rank \\
Transfer Fatigue				& 90 + Rank			& 11 minutes − 1 / Rank \\
Neutralise Poison				& 90 + Rank or 50 − 5 × difference in Rank	& 5 seconds \\
Cure Burns and Repair Scars			& 90 + Rank			& 30 minutes − 2 / Rank \\
Repair Muscle, Bones, Tissues and Organs	& 90 + Rank			& 50 hours − 3 / Rank \\
Preserve Dead					& 90 + Rank			& 60 minutes − 5 / Rank \\
Reattach Body Parts				& 90 + Rank 			& 30 minutes − 2 / Rank \\
Resurrection					& special			& 60 minutes − 5 / Rank \\
Regeneration					& 90 + Rank			& 1 week \\
\end{dqtblr}

\end{table*}

\end{Skill}
