\begin{Chapter}{Herbalist (Ver 1.1)}

Herbalism is an old profession, and herbalists of varying quality can
be found throughout the land.  Whilst most of the magical power of
this skill is derived from the herbs themselves, herbalists will have
quasi-magical abilities to derive the greatest benefit from the
herb. To some extent the ranger’s healing herbs are a simple aspect of
the herbalist’s skill.  Through the use of special herbal mixtures the
herbalist can mimic some of the abilities of other skills, in
particular those of the healer and of the alchemist. In general a
herbalist cannot perform their skills at short notice since the number
of herbs and preparations required to meet all needs would be too
large to carry, and many preparations have a limited life. But given
enough time a com- petent herbalist can usually find and prepare a
herbal mixture to deal with most problems.

A herbalist may make some potions in advance but these potions will
only last for a limited time.  Potions are always roughly a quarter of
a pint per dose and weigh, with their container, ½ pound.  Salves with
their container weigh ¼ pound.  A herbalist may make up to Rank / 3
(round up) batches (a single batch may produce multiple doses if
enough herbs are used) of potion or salve simultaneously.  Should they
attempt to make more potions at once then there is a 50\% chance for
each extra batch that all potions being made will fail.  Should the
batches being made fail there is a further 30\% chance that some form
of accident will occur (GM’s discretion).  While the herbalist is
producing potions to their full capacity no other actions are
permitted, but if the herbalist is working at less than full capacity
then they can interrupt their work for training, meeting visitors
etc., providing they tend the potions at least every hour.

\section{Herbs}

\subsection{Locate and Identify Herbs}

A herbalist will learn how to locate and identify herbs and spices
growing in any terrain.  Their chance of finding fresh herbs suitable
for perform- ing a given skill is (Perception × 2 + Rank × 15 - 10 ×
the minimal rank at which they would be able to perform that skill) +
the following modifiers

\subsubsection{Rarity}

\begin{tabularx}{\columnwidth}{Xr}
common		& 0 \\
uncommon	& -20 \\
rare		& -70 \\
very rare	& -100 \\
\end{tabularx}

\subsubsection{Season}

\begin{tabularx}{\columnwidth}{Xr}
spring		& -20 \\
summer		& -5 \\
autumn		& 0 \\
winter		& -25 \\
\end{tabularx}

\subsubsection{Search Area}

\begin{tabularx}{\columnwidth}{Xr}
Less than ¼ of a square mile (roughly 1 hour of searching per ¼ of a square mile) & -30 \\
Between ¼ and ½ square miles	& -20 \\
Between ½ and 1 square mile	& -10 \\
Between 1 and 1½ square miles	& +0 \\
Over 1½ square miles		& +10 \\
each subsequent hour or ¼ square mile	& +5 \\
\end{tabularx}

\subsubsection{Other}

\begin{tabularx}{\columnwidth}{Xr}
For rare and very rare herbs, area is high mana		& +25 \\
Area is wilderness					& +10 \\
Area is settled						& -15 \\
Area is civilised					& -30 \\
For each rank of ranger skill				& +1 \\
or for each rank of  ranger skill if the environment is the ranger’s speciality as per rule §45.2 & +2 \\
\end{tabularx}

In some environments (for example off plane or tropical jungle) the GM
may choose to add extra modifiers.  Also some environments are simply
unsuited to plant growth and a herbalist will not be able to find
herbs.

The herbalist must state that they are looking for herbs required for
a particular ability before setting out to search.  If the herbalist
then rolls over their modified chance, they will find no useful
herbs. A herbalist may search the same area a number of times,
providing that they search for herbs required for a different ability
each time.

Common and uncommon herbs may be found in quantities larger than a
single dose depending on conditions, but it would be rare to find
sufficient for more than three or four doses at once.

In general, rare and very rare herbs are found in small quantities
sufficient for a single application, and these magical herbs are more
frequently found associated with places of mana.

After harvesting herbs there, the herbalist cannot return at a later
date to harvest more, since the original harvest will almost certainly
kill the plant.  If the herbalist does not harvest the herbs and
returns at a later time to harvest the herbs, there is a 5\%
cumulative chance per days delay that the herb will have died or been
consumed.

A herbalist may alway successfully identify herbs (in fresh or
prepared form) used in abilities which the herbalist can perform.  If
the herbalist does not have the ability for which the herbs are used
they still have a Perception + 5\% / Rank chance of knowing what the
herb is used for providing that the herb is the usual one used for
that ability, and not one of the rarer substitutes.

If the herbalist is so far from the environment in which they learned
their skill (for example off plane) that the plants and herbs are
unknown, then they will not know what abilities the unknown herbs
could be used for.

\subsection{Herbal Garden}

A herbalist may establish a garden where common and uncommon herbs and
spices are grown. Such a garden will require tending by a herbalist
for 1 week every month.  If left untended for 2 months uncommon plants
will die or be overgrown and if left untended for 6 or more months
such a garden will be only as fruitful as wilderness (see above) for
finding herbs.

\subsection{Preserving Herbs}

A herbalist will learn how to preserve herbs and spices and how to
prepare them for maximum efficacy.  Such preparation is simple and
will only take a few minutes.  Only common herbs will last for any
length of time.  For common herbs preser- vation is Rank / 2 months,
for uncommon herbs Rank /

3 weeks, for rare herbs Rank / 2 days and for very rare herbs Rank × 2
hours.  The herbs will last at full efficacy for the time indicated
but after that time will decay at a rate proportional to the time of
preservation.

\section{Benefits}

A herbalist gains specific abilities at each Rank as 
per the following table: 

\begin{Enumerate}\setcounter{enumi}{-1}
\item Herbs, Cooking

\item Restorative Meal; Cure Infections, Headaches and Fevers;
  Relaxation Tea

\item Cure warts, boils and pimples; Healing Sleep; Sleep Draught;
  Animal Perfume; Mild Poison

\item Enhance Endurance Healing; Fertility Brew

\item Cure Disease;
  Sentient Perfume; Strong Poison 

\item Enhance Organ Healing

\item Enhance Bone Healing; Hallucination Libation; Fantastical
  Perfume; Multi-part Deadly Poison

\item Enhance Characteristic Potion; Spirit Travel

\item   – 

\item Restore Life

\end{Enumerate}

\section{Cooking}

A herbalist, should they choose to take the artisan cooking skill,
gains a reduction of 25\% to their experience for ranking that artisan
skill.  In addition they will be able to use their knowledge of herbs
to enhance their cooking skill and prepare an invigorating broth.
As their rank increases a herbalist who is also an artisan Cook, or
acting in concert with an artisan Cook, will be able to prepare superb
food, providing there are sufficient ingredients, common herbs and
spices available.  Even with limited ingredients a herbalist will
enhance the art of preparing enjoyable food.

\subsection{Restorative Meal or Broth}

At Rank 1 and above the herbalist may include common herbs that will
turn a normal hot meal into an invigorating restorative meal (capable
of restoring fatigue loss due to damage or spell use) that will add 2
+ (Rank / 4, round up) lost fatigue points, or the herbalist may make
a broth or tea that will restore 1 + (Rank / 2, round up) fatigue
points.  The restorative meal or the broth/tea replace a normal hot
meal and if taken in addition to a hot meal no extra fatigue will be
restored.  Like a hot meal, the restorative meal or the broth / tea
can be taken once every 4 hours.

\section{Healing}

A herbalist will be able to prepare herbal mixtures, and salves that
will heal, or enhance the normal healing processes.

\subsection{Cure Infections, Headaches and Fevers}

At Rank 1 a herbalist may prepare, using common herbs, antiseptic
salves that will disinfect wounds and cure infections, and potions
that will cure headaches and fevers.  These salves and potions will
last Rank months, and it will take 24 hours to prepare Rank / 2 (round
up) doses.

\subsection{Cure Warts, Boils and Pimples}

At Rank 2 a herbalist may prepare, using common herbs, salves that
will cure warts, boils and pimples.  These salves will last Rank
months, and it will take 24 hours to prepare Rank / 2 (round up)
doses.

\subsection{Healing Sleep}

At Rank 2 a herbalist may prepare, using common herbs, a potion that
will cause the drinker to fall into a healing sleep that will heal
fatigue or endurance damage at the rate of 4 fatigue per hour or 1
endurance per hour if the sleeper has lost endurance.  Note that if
the sleeper is recovering endurance they will recover no fatigue.  The
sleeper can be woken at any time but must sleep at least 4 hours to
receive any benefit at all.  These potions will last Rank months, and
it will take 24 hours to prepare Rank / 3 (round up) doses.

\subsection{Enhance Endurance Healing}

At Rank 3 a herbalist may prepare, using uncommon herbs, healing
salves that will enhance the healing time of wounds. Divide the normal
rate by their rank.  For example a Rank 10 herbalist treating a deep
sword cut that would normally take 20 days to heal would cause it to
heal in 2 days.  This healing will include repair of muscle, ligament,
tendon and skin, all without scarring provided the wound is open and
the salve can be ap- plied directly to the wound. Note that one
preparation of salve will heal one serious wound or a number of minor
wounds.  If the herbalist uses uncommon herbs these salves will last
for Rank days, and it will take 6 hours to prepare Rank / 3 (round up)
doses.  If the herbalist uses rare herbs these salves will last for
Rank weeks, and it will takes 12 hours to prepare Rank / 4 (round up)
doses.

\subsection{Cure Disease}

At Rank 4 a herbalist may prepare, using uncommon herbs, potions that
will cure diseases (note a specific disease will require a specific
potion).  These potions will last for Rank weeks, and it will take 12
hours to prepare Rank / 3 (round up) doses.

\subsection{Enhance Organ Healing}

At Rank 5 a herbalist may prepare, using uncom- mon herbs, potions
that will enhance the healing of bruising and internal injuries to
Rank × the normal rate.  If the herbalist uses uncommon herbs these
potions will last for Rank days, and it will take 30 minutes to
prepare one dose.  If the herbalist uses rare herbs then these potions
will last for Rank weeks, and it will take 12 hours to prepare Rank /
4 (round up) doses.

\subsection{Enhance Bone Healing}

At Rank 6 a herbalist may prepare, using rare herbs, potions that will
enhance the healing of broken bones to Rank × the normal rate.  These
potions will last for Rank weeks, and it will take 12 hours to prepare
Rank / 5 (round up) doses.

\subsection{Restore Life Elixir}

At Rank 9 a herbalist may prepare, using the very rarest of herbs, a
potion that will restore life to a sentient being providing that the
being died not more than 12 hours previously or if the body was
preserved as per the healer skill within 12 hours of death, and the
body is capable of supporting life.  Thus the body must be
substantially whole (ie no more than an arm or a leg lost) and any
fatal wounds (for example sword thrust through the heart) must have
the appropriate healing salves or potions applied immediately prior to
the restoration of life.  After life has been restored, the being will
be too weak to move for a number of days equal to the number of hours
which they were dead after which they will regain one point of
endurance as per the healer skill. This potion will last Rank × 3
hours and it will take 30 minutes to prepare a single drop, which is
then placed on the tongue of the patient to restore life.

\section{Alter Natural Rhythms}

A herbalist may prepare herbal mixtures that alter the normal rhythms
of a sentient being’s body.

\subsection{Relaxation Tea}

At Rank 1 a herbalist may prepare, using common herbs, a potion that
will reduce tension and promote relaxation.  This effect will last for
Rank × 4 hours.  This potion will last for Rank weeks, and it will take
24 hours to prepare Rank / 2 (round up) doses.

\subsection{Sleep Draught}

At Rank 2 a herbalist may prepare, using uncommon herbs, a potion that
will induce sleep for up to Rank × 2 hours unless the drinker strongly
resists (rolls under 4 × willpower every 10 minutes).  This potion
will last for Rank weeks, and it will take 24 hours to prepare Rank /
3 (round up) doses.

\subsection{Fertility Brew}

At Rank 3 a herbalist may prepare, using common herbs, a potion that
will increase the likelihood of fertilisation by 1 + 1\% per Rank for
a period of Rank days, or reduce the likelihood of becoming pregnant
by 2 + 2\% per Rank, for a period of Rank weeks.  These potions will
last for Rank months, and it will take 24 hours to prepare Rank / 3
(round up) doses.  Note these potions will be gender and race
specific.

\subsection{Hallucination Libation}

At Rank 6 a herbalist may prepare, using uncommon herbs, a potion that
will cause the drinker to hallucinate, unless they roll less than (2 ×
Willpower - 3 × herbalist’s Rank), for up to Rank hours.  The
herbalist can usually (75\% chance) control the extent (i.e. subtle,
with only a slight dreamlike character to reality, mild, with reality
moderately distorted, or total, with no limits at all) and the general
nature (i.e. pleasant or unpleasant) of the hallucination.  If the
herbalist uses uncommon herbs these potions will last for Rank days,
and it will take 6 hours to prepare Rank / 3 (round up) doses.  If the
herbalist uses rare herbs these potions will last for Rank weeks, and
it will take 12 hours to prepare Rank / 4 (round up) doses.

\subsection{Enhance Characteristic Potion}

At Rank 7 a herbalist may prepare, using very rare herbs, a potion
that will allow the drinker to enhance any one characteristic by 2 ×
(Rank - 6) points. This has a duration of (Rank 6) hours. Only one
Herbalist characteristic increasing potion may be in effect on an
entity at any one time.  There is no reduction of characteristics or
sleep period required after the duration of the positive effects —
i.e. no down-side.  These potions will last for Rank × 2 weeks, and it
will take 48 hours to prepare Rank / 4 (round up) doses.

\section{Perfumes and Odours}

A herbalist may prepare perfumes that will attract or repel animals or
sentients.  Such perfumes will only mildly modify (plus or minus Rank
× 4\% to reaction roll) the reactions of the creature being affected
(except that in this case a low roll indi- cates the degree of
repulsion rather than hostility).  These perfumes are subtle and only
if the being makes the appropriate willpower roll will they
consciously notice the perfume.

The perfumes will not change the basic nature of the creature (i.e.
no matter how nice you smell a dragon will still kill and eat you,
although she may recommend the herbalist to other dragons as a
supplier of fine condiments). All perfumes will last (in the bottle)
for Rank weeks.  Thorough washing will reduce the duration of a
perfume by 10 fold. A clean cantrip will remove the perfume.

\subsection{Animal Perfume}

At Rank 2 a herbalist may prepare, using uncommon herbs, a perfume
that will attract or repel a given animal, the animal must roll under
(WP - herbalist’s Rank) to resist the effects of the perfume.  The
perfume will continue to act for Rank days once applied.  It will take
12 hours to prepare Rank / 3 (round up) doses.

\subsection{Sentient Perfume}

At Rank 4 a herbalist may prepare, using rare herbs, a perfume that
will attract or repel a given sentient creature (including character
races), the creature must roll under (2 × WP - herbalist’s Rank) to
resist the effects of the perfume.  The perfume will continue to act
for Rank × 3 hours once applied. It will take 24 hours to prepare Rank
/ 4 (round up) doses.

\subsection{Fantastical Perfume}

At Rank 6 a herbalist may prepare, using very rare herbs, a perfume
that will attract or repel fantastical or magical creatures, the
creature must roll under (3 × WP - herbalist’s Rank) to resist the
effects of the perfume.  The perfume will continue to act for Rank
hours once applied. it will take 48 hours to prepare Rank / 3 (round
up) doses.

\section{Poisons}

A herbalist may prepare toxic mixtures of herbs that may harm and
ultimately kill animals and sentients that ingest them. All poisons
will last (in the bottle) for Rank weeks.

\subsection{Mild Poison}

At Rank 2 a herbalist may prepare, using uncommon herbs, a poison that
will cause [D - 6] + (Rank / 4, round up) damage per minute for Rank ×
2 minutes (damage is applied at the end of each minute).  It will take
24 hours to prepare Rank / 3 (round up) doses.

\subsection{Strong Poison}

At Rank 4 a herbalist may prepare, using uncommon herbs, a poison that
will cause [D - 6] + (Rank / 3, round up) damage per pulse, for Rank ×
2 pulses. Or using rare herbs a poison that will cause [D - 3] + (Rank
/ 2, round up) damage per pulse, for Rank × 2 pulses.  If the
herbalist uses uncom- mon herbs it will take 24 hours to prepare Rank
/ 3 (round up) doses. If the herbalist uses rare herbs it will take 36
hours to prepare Rank / 4 (round up) doses.

\subsection{Multi-part Deadly Poison}

At Rank 6 a herbalist may prepare, using rare herbs, a poison in
multiple parts (2 or more) that will cause [D - 4] + Rank damage per
pulse, for Rank pulses. The time between application of first and last
part must be no more than 24 hours. Using very rare herbs, the
herbalist may prepare a poison that will kill a human sized creature
in 3 pulses.

Larger creatures have a percentage chance of dying in proportion to
their body size e.g. a creature four times the size of a human has a
25\% chance of dying. If the creature does not die then it takes [D -
  4] + Rank damage per pulse, for Rank pulses.  It will take 72 hours
to prepare Rank / 4 (round up) doses of multi-part poison. It will
take 48 hours to prepare Rank / 4 (round up) doses of instant kill
poison.

\section{Spirit Travel}

A herbalist can prepare herbal mixtures that can free the mind from
the body.

At Rank 7 a herbalist may prepare, using very rare herbs, a potion
that will allow them to separate their spirit from their body and
travel in spirit anywhere within 10 miles of the body for up to Rank
hours at a cost of 3 endurance points per hour.  After such spirit
travel the herbalist may regain lost endurance by healing or normal
recovery except for 1 point per hour travelled which can only be
regained by expenditure of experience (2500 EP per point lost).  While
in spirit form the herbalist may, move at normal walking or running
speed, pass through any solid objects, see as they would normally see
(i.e.  racial talents apply), hear as they would normally hear, use
any talent (including magical), cast any spell that affects only the
Adept, and fly at normal walking speed.  The spirit form cannot touch
nor move any object by non-magical means, nor can it speak.  The
spirit form is insubstantial and cannot be touched or harmed in any
way (magically or physically).  The spirit form is unseen but may be
detected by Witchsight as per the spell of Walking Unseen. An entity
that loses their last point of endurance as a result of this potion
will become a spectre.  This potion will last for Rank × 2 days, and
it will take 48 hours to prepare Rank / 4 (round up) doses.

\end{Chapter}
