\begin{Chapter}{Version History}

\subsection{History of Edition 2024e}

\begin{Description}

\item[March 2025] Work on replacing hypens with a minus and
  replace hardwired references with dynamic labels and
  references.  TeX engine moved from xe\TeX to luahb\TeX.
    
\item[January 2024] All tabular material converted to using
  tabularray tblr.  Find and remove remaining incorrect hypenation.

\subsection{History of Edition 2020e}

\item[September 2020] PDF of 2014 edition converted to plain
text then reformatted into \LaTeX.

\end{Description}

\subsection{History of Edition 2.0}

\begin{Description}

\item[September 2014] Blowgun damage added and note changed to similar
  to venomous snakes.  Counterspell of other college clarified as
  being special knowledge.  Spy stealth bonus changed back to 2\%/rank
  to match stealth skill.

\item[May 2014] Jono's version of Evil Eye (E\&E G-9).  Rename Wicca
  Evil Eye to Hex.  Remove requirement to cast 5 times per when
  ranking.  Remove the 100 charge rule from investment.  Warrior 3.0
  basic and special abilities.  Rewording of Spirit Vision (Rune T-2)
  re visibility restriction.  Change the Background Experience in
  Character Generation from 250 experience to 2,500.  Add decreasing
  Racial Modifiers to Character Gen: Race.  Interrupted Ranking:
  replace completed with resumed.  Reworded E\&E and Illusion
  invisibility.  Moved damage by burning to end of fire college.  Add
  adventuring skill combo summary.  In Heritage change Baronies to
  Western Kingdom and Cazarla.  Change to Armourer to V1.3.  Removed
  base chance from wizardsight.

\end{Description}

\subsection{History of Edition 1.8}

\begin{Description}

\item[January 2013] EPub Edition. Spelling and grammar edits.  College
  intro hierarchy rearrangement in Mind, Namer, Fire, Necro,
  Celestial. Removed first level heading as inconsistent and had
  orphans.  Removed cytogenesis.  Changed non-tactile empathy to
  ranged empathy.  Changed collegiate to college. Checked semi-colons.

Moved sap from hafted weapons to special weapons in 56.1.

Changed wording of spirit vision removing sentence "This vision is
blocked by material objects (even if invisible) and magical darkness
of Rank 20.", adding "as though they were normally visible.  Fixed
Detect Enchantment (T-2) binding college "Range: 1.  30 feet (+ 5 /
Rank) 2.  Touch "Base Chance: 1. PC + 3\% / Rank 2. PC + 5\% / Rank"
"If the Adept is in contact with the target then the base chance of
this talent is higher."

The Ice elemental has AG 15-2.  Assumed that it should be 15-20.

Removed non-magical from these two lines as otherwise it implies that
the it doesn't apply to the quasi-magical abilities mentioned just
below them.  “A character may attempt to employ a *non-magical* skill
any number of times during a day.”  “The use of a *non-magical* skill
is rarely automatic.”

6.10 Fumbles If the weapon is not Magical, or made of cold iron, or a
Bow, or a Crossbow, increase this chance of fumbling by: any silver or
truesilver alloy of iron 1\% etc changed to: This chance of fumbling
is increased if the weapon is made of a material other than cold iron,
as listed below, unless it is magical, or a Bow or Crossbow.  Removed
reference in Ice College spell Refrigeration (G-5) to fire college
temperature alteration spell.

Corrected Wiccan Damnum Minatum reference to Ice College from Air
College.
\end{Description}

\subsection{History of Edition 1.7}

\begin{Description}

\item[July 11, 2010] Large Print Edition. 

\item[June 10, 2010] Enhance Indexes.  Move Combat section to before
  Magic.  Reorder character-related sections.  Move change log to end
  of Rulebook.  Add ToC to Combat, Magic.

\item[June 9, 2010] Miscellaneous edits for greater clarity and
  readability.  Allow Courtier and Troubadour specialisation when
  taking subskill.  Adjust bonus for rank in Stone Golems. Re-order
  Binder Golem info.  Add back clothing exemption to Encumbrance.

\item[June 8, 2010] Changes to Healer Regeneration, Graft Skin to Cure
  Burns, adjust and correct summary table.  Update Languages.
  Waters of Healing neutralise venom if present. More info gained from
  Detect Poisons.  Standardise to ‘Shapechanger’.  Masterwork
  exceptions for Adventuring and Language Skills. Move Knowledge Skill
  from Adventuring Skills to Skill Intro.  Add Salve details to
  Alchemist, and time to make Medicines/Antidotes.  Clarify Combat
  Equation Summary.  Add Compose Music to Courtier.

\item[June 7, 2010] Repeat lost June 1-7, 2006 miscellaneous edits.

\end{Description}

\subsection{History of Edition 1.6}

\begin{Description}

\item[June 7, 2006] Miscellaneous edits for clarity and readability.

\item[May 25, 2006] Move Conception from Character Generation to
  Health and Fitness.  Added 6 month limit and Subskills to Ranking.
  Permanency changes to Binder and Illusion Colleges.  Weather table
  rationalised. Rune v2.2 in playtest and available for new
  characters.  Modified Sub-skill rules for all applicable skills.
  Remove Ropes cost and Orienteering from Adventuring skills. Move
  Healer potion costs to Alchemist.  Cost formula clarifications in
  Armourer.  Added new Alphabet and Languages to Languages.  Removed
  Spy discount from Thief.  Cost formula changes in Weaponsmith.
  Added EP costs for Lasso and removed Katana.  Rationalised Skill
  notes.

\item[January 11, 2006] Add Spy 2.0 

\item[December 6, 2005] Include Subskill costs in the EP table.

\end{Description}

\subsection{History of Edition 1.5}

\begin{Description}

\item[June 3, 2004] Miscellaneous edits for greater clarity and
  readability.  Acknowledge contributors for the August 13, 2001 Mind
  College revision.  Remove partial contributors list from the
  cover.

\item[May 26, 2004] College Magic: The Investment Ritual (Ver 1.2).
  New Version of the E\&E Ritual of Greater Enchantment.  Geas changed
  to max rank 20 and Full Geas at rank 15.  Mind College spell of
  Undetectability removed.  Necromancy spell of Necrosis: limitation
  of only affecting living restored. Remove Probation tags from the
  colleges of Binder, Namer, Mind, and Fire.  Add Shields to
  Weaponsmith.

\end{Description}

\subsection{History of Edition 1.4}

\begin{Description}

\item[December 19, 2001] All files revised to version 5.10 in RCS.

\item[September 28, 2001] Ranger version 2.1 added.  This is a
  complete replacement for the current ranger. Thanks to Rosemary
  Mansfield for this.

\item[September 17, 2001] More changes to stun in combat to fix
  editor’s misunderstanding about the free recovery at the end the
  pulse.  Fix typo in Namer probation paragraph and copy it to the
  Mind College.

\item[September 3, 2001] Change wording in restrictions on magic to
  enforce the limitation of cold iron on casting.  Fix wording in
  Unarmed combat for kick so that the wording is now Rank 3 or above.
  Fix reference to DA in Namer.  Fix wording under Encumbrance in
  Adventure.  Try to fix Recover from Stun so that figures get a
  Recover from Stun action at the end of the pulse they were stunned
  in.  Text of Mind Speech tightened up. The Ranking of Names modified
  to reflect new Namer. The section on Namer Ranking has been modified
  accordingly.  The cost of a trainer for Ranking a weapon to Rank 0
  has been included (10sp).

  The sections Auras and Names merged into a new section “Auras \&
  Names”.  Names wasn’t really significant to have its own section and
  adding it threw out all the section numbering.  Text of Scry Shield
  tightened up thanks to Terry Spencer. Minor corrections of Mind by
  Ian Wood.

\item[August 30, 2001] Minor corrections to combat suggested by Errol
  Cavit.  Order credits by alphabetical order and add Errol Cavit to
  it.  Celestial Light and Dark aspect rewritten by Errol Cavit.  This
  had previously been munged by Ross Alexander from the GM’s Guide.
  Fix logic error in Name ranking in Namer spotted by Terry Spencer
  and fixed by Martin Dickson.  Resistance on Wall of Darkness
  corrected from None to Passive.

\item[August 13, 2001] The Mind College revision 1.6 done by Struan
  Judd, Jacqui Smith and Ian Wood.  Detect Aura rewritten and a new
  section on True Names added.  The Counterspell subsection of general
  magic rewritten with respect to the changes to Namer.  The College
  of Naming Incantations completely rewritten by Martin Dickson and is
  version 2.0.

\item[July 11, 2001] Change EM of Fireball from 500 to 550.  Check
  duration of Weapon of Flames is 5 minutes + 1 / Rank. Add duration
  to Speak to Fire Creatures of 20 minutes + 10 / Rank.  Duration of
  Alchemist’s potion of Talents is now based on Rank of Talent, not
  Rank of Alchemist.  Change text of reference of Health and Fitness
  in Healer.  Significant changes to Recovery from Stun introduced
  into Combat.  Add Recover from Stun to combat actions table.

\item[June 13, 2001] Fix typo in history.  Fix problem with section
  headings in character generation and conception.  Add note to Aspect
  about Celestial college affecting light/dark aspect.

\end{Description}

\subsection{History of Edition 1.3}

\begin{Description}

\item[August 31, 2000] Looking at Fire, the EM of Fireball is 500,
  which is what was agreed by Paul Schmidt just be the final release.
  Duration of Weapon of Flames corrected to 5 minutes + 1 / Rank.
  Duration of potioned talents changed from Rank of alchemist to Rank
  of talent.

\item[June 8, 2000] Final checkin for version 1.3.  Last minute change
  to EM of Fireball from 350 to 550 as a preventive measure.
  Requested by Jim Arona and put in by the editor on the principle
  that it is easier to refund overspend EP then to increase the EM of
  a spell and try to sort out the deficit.

\item[June 6, 2000] Change warrior initiative to +1/Rank (except Rank
  0).  Healer and adventure retypeset.  Minor magic replaced with
  Cantrips.  All weapon spells (except Runeweapon) to have duration 5
  minutes + 1 / Rank.  Weapons of Flames toned down.

\item[May 23, 2000] Lots of minor corrections.  Conception added to
  race descriptions.

\item[May 19, 2000] Notes on grievous injury moved into adventure
  under the new subsection of Health and Fitness. The addendum of
  Healer has been merged into the text of healer where possible.  The
  notes about the Guild healing services have been removed and
  should be in the Players Guide.

\item[March 29, 2000] Illusion Talent of Enhanced Vision changed
  (Andrew Withy).  The description for Wizard Sight changed (Andrew
  Withy). Range and EM of Witchcraft Hellfire changed to match new
  Fire college.

\item[March 2, 2000] Language bonus for Namers changed to fit new
  language document.  Celestial Witchsight description changed.  E\&E
  Witchsight renamed to Wizard Sight.  Fire College rewritten (Paul
  Schmidt) and updated to version 2.0 and probationary notice added.
  Wiccan Hellfire and Instilling Flight changed to match Fire and
  Binder respectively.  Expert changed to Knowledge in EP table notes.
  Symbols of power for curse removal are now portable.  Version 2.0
  removed from title of Explanation of Characteristics.  Orienteering
  removed and artisan cartographer added instead (Andrew Withy).

\item[February 29, 2000] Languages skill updated to version 2.0.
  Change languages in ranking section of adventuring.  The Talent of
  Speaking to Creatures of Dark/Light, Bardic College and Philosopher
  changed to fit in with new Languages document.  Healers can now use
  limited abilities on non-sentient animals. Runeweapons made from
  yew are to be treated as natural poisons.  Mechanician changed to
  version 2.1 and spell containment rewritten (Martin Dickson).

  The Bardic spell Great Shout replaced with a new spell Shout of
  Thunder.  Bardic and Ice now made full colleges with Binder still
  probationary.

\item[October 20, 1999] Add new fumble rules to combat and fumble
  table.

\item[October 18, 1999] Fix rules where the term “round” appears and
  should be pulse.  Changes from Kelsie.

\item[October 8, 1999] Delete Spell of Echosense and cascade reference
  numbers down.

\item[October 7, 1999] Change to ritual description in general
  magic. Rituals now backfire on BC + 30\% and can double and triple
  effect. The description to Ritual of Water Elemental (R-1) and
  Ritual of Summoning Creatures of Light and Dark (Q-2) modified.  Add
  note that Purification cannot backfire. Add note of backfire of
  Investment ritual and add note that Remove Curse does not backfire
  normally. Probation note removed from Bardic and Ice colleges.
  Backfire description on Ritual of Resounding Instrument (R-2)
  changed.  Binding and Animating Ritual of Divination (R-3) changed
  so it no longer backfires.  Ritual of Fire Elemental (R-1)
  description changed. Description of Ritual of Illusory Fog (Q-2)
  changed.  Active Talents now require a pass action rather than one
  pulse to use.

\item[August 18, 1999] Banishing demons summoned by a Greater Summoner
  using Special Counterspell can now only be done by the Adept.
  Counterspelling multiple instances of a spell clarified.

\item[August 2, 1999] Conjuring the Controlling Air Element (R-2)
  modified.

\end{Description}

\subsection{History of Edition 1.2}

\begin{Description}

\item[June 4, 1998] Various editing changes before final print.

\item[May 11, 1998] Add aquatic adventuring changes from Keith Smith.
  Fix e.g. and i.e.  problems.  Change descriptions of Greater
  Enchantment, Sleep, Purification, Strength of Stone and Damnum
  Minatum.  Herbalist stat increase potions changed.  Namer encounter
  method removed and generic names now teachable.

\item[May 7, 1998] Alter navigator find landmark back to DQII original
  rule (Michael Parkinson).  Change Sinking Doom, Whitefire and
  Incinerate to no longer be irresurrectable. (Stephen Martin).

\item[May 6, 1998] New combat section by Andrew Withy.

\item[May 5, 1998] New character generation from Rosemary
  Mansfield. Major typographical changes in magic.

\item[April 15, 1998] Artisan and Mechanician completely rewritten
  (Martin Dickson).

\item[March 30, 1998] Expert Knowledge and Supervisor added to skills
  (Martin Dickson).  Courtier, Philosopher, Military Scientist and
  Troubadour completely rewritten (Martin Dickson).

\item[September 3, 1997] Add minor typographical changes to Ice, Water
  and corrections from Keith Smith.

\item[June 23, 1997] Typesetting changes to force Binder and Ice to
  put in an empty page if end on odd page.

\end{Description}

\subsection{History of Edition 1.1}

\begin{Description}

\item[6 June, 1997] Final print for 1997 rulebook. 

\item[5 June, 1997] Witchcraft control weather modified.  Initial stat
  generation modified so player can choose 90 points rather than
  rolling.  Bardic 1.1 added (Jacqui Smith and Martin Dickson).

\item[22 May, 1997] Parts of Stephen Martin’s document on eating,
  recovery and infection added.  Infection removed from combat.
  Falling removed from Thief and put into Adventure.  Binder 1.1 from
  Stephen Martin added.

\item[15 May, 1997] Aquatic Affinity talent added to Water (Keith
  Smith).  Paragraph about Bardic and Ice being probationary until
  June 1999.  Skills imply knowledge of the subject area added (Martin
  Dickson).  Skills do not necessarily imply traits added (Ian Wood).
  Notes on blood agents moved from infection in combat. Infection
  removed from combat.

\item[7 May, 1997] Walls of Darkness and Walls of Starlight changed so
  that BC/EM switched for Solar/Star (Andrew Withy).  Version changed
  to 1.3.  Necro and Witchcraft Darkness modified to Celestial
  Darkness.  Ice 1.5 from Carl Reynolds introduced (May Gods
  meeting).

\item[29 April, 1997] Numerous minor changes. Damage on heavy
  crossbows changed from +4 to +5.

\item[22 April, 1997] Weight of invested items changed to minimum of
  one ounce (April Gods meeting).

\item[9 April, 1997] Weight table modified so that values represent
  maximums rather than median values.  This should make it simpler to
  use.  Mages now cannot cast while prone.  Gilding removed from
  restrictions on magic and cold iron.

\item[6 April, 1997] Ice College 1.3 added by Carl Reynolds
  12/1/96. Bardic 1.0 by Jacqui Smith added.

\item[April 1, 1997] Typo in Waters of Vision and Crystal of Vision
  fixed.

\item[March 20, 1997] Notes on shaped items added to introduction to
  magic.

\item[March 19, 1997] Monsters introduction and aspect moved to GM’s
  Guide.

\item[[March 14, 1997] DQ Swimming Skill Ver 2.3 by Keith Smith
  added. Flying added and EP added to EP table (as 125 / Rank). Notes
  on Ranking added to adventuring section.

\item[March 13, 1997] Extended rituals and notes on possessions moved
  from Binder to general magic.  Counterspells made wardable and
  trappable.  Storages on Namer spells fixed.

\item[February 28, 1997] Guild banking removed from adventure section
  and appending to player’s guide.

\item[February 25, 1997] Spelling mistake in combat summary fixed.
  Invisibility added to modifiers (- 50). Water fixed by Clare West,
  with new version (1.3).

\item[February 11, 1997] Giants changed to weight multiple 9 (to be
  confirmed).  Notes of other creatures added. Minor fixes to Illusion
  1.4.

\item[February 7, 1997] Change maximum Rank of Silent Tongue to six.

\item[January 15, 1997] Spelling mistakes corrected in Namer Name
  List, Light and Dark Aspect and Minor Magics.  Additional indices
  added to Introduction to Magic.

\item[December 17, 1996] New version of the Investment ritual (v1.0)
  added.  Extra indices added to college and non college ritual
  section.

\item[November 5, 1996] Awe table appended to fright table. Fright
  from low PB now minimum 1 × WP.

\item[November 1, 1996] Additional indexes added and notes on
  orienteering changed in the skills cost tables.  Ordering of tables
  changed.  Daylight time and holiday times moved to Player’s Guide.

\item[October 8, 1996] Map reading added to ranger. Map reading and
  drawing maps changed in navigator.

\item[October 1, 1996] Tidy up files in general. Targeting in magic
  changed so that an Adept can attempt to cast at a target that is
  potentially out of range in the hope of a multiple effect. Notes on
  magical storage (not approved at this time) added to magic. Distract
  removed from minor magic.

\item[September 26, 1996] Greater Summonings altered.  The
  descriptions of imps, devils, succubi and incubi removed and put
  into the monster manual, while the descriptions of the special
  rituals (summoning Dukes, Princes, Presidents, Earls, Marquis and
  Kings) added.  Half devils also added to monster manual, under
  summonables.

\item[September 16, 1996] Add two new combat spells to Air (Jon
  McSpadden), and change version to 2.1.  Fix resistance in Damnum
  Magnatum.  Add notes on magical storage to introduction to magic.
  Add note that learning a college for the first time takes 6500 ep
  and six months.  Dragon Flames and Necrosis now actively resistible.

\item[September 12, 1996] Changes to Celestial Lighting modifiers and
  the spells of Light and Darkness added. Version updated to 1.2.

\item[August 12, 1996] Add new bow rules.  Missile weapons before
  unarmed combat in combat section and tables updated.

\item[August 9, 1996] Change the description of orcs and their racial
  multiplier to 1.1. Change hill giants size multiplier to 11 (rather
  than 15).  This has to be verified.  Be able to target outside
  normal range in the hope of a double or triple.

\item[July 11, 1996] Add Monsters section introduction and notes on
  aspects.

\item[March 27, 1996] Add Name list of Namer.  Add holidays and
  sunrise / sunset table.  Add combat equation summary to tables.

\item[March 25, 1996] Rework coinage table to make clearer. Make first
  reprint.

\item[February 22, 1996] Change weight of coins and remove personal
  names from colleges.

\item[February 17, 1996] Parameters tweaked for formatting, and tables
  moved to back of book, after skills.

\item[January 12, 1996] Initial release for players and GMs. Release
  document does not contain combat.

\item[January 5, 1996] E \& E and Mind revised by Ross Alexander and
  Brent Jackson.

\item[October 14, 1995] Air rewritten by Jono Bean, Carl Reynolds,
  Phil Judd and Rosemary Mansfield.  Version 2.0 released.

\item[October 14, 1995] Illusion rewritten by Andrew Withy to version
  1.3.

\item[October 1, 1995] General magic revised by Ross Alexander, Andrew
  Withy and Brent Jackson.

\item[June 8, 1995] Water revised by Clare West to version 1.2.

\item[June 6, 1995] First draft is released to GMs for comments and
  corrections.

\item[June 4, 1995] Earth revised by Ross Alexander to version 1.2.


\end{Description}

\end{Chapter}
