\begin{College}[1.5]{ice}{Ice Magics}{IC}

The College of Ice Magics is concerned with the shaping of ice and
snow, freezing and the manipulation of cold.

Ice Mages have been most valuable to those fragile communities living
in cold, arid and often dangerous places, where food is hard to grow
and neighbours can be separated by days of travel.  Many of the Ice
Mages’ Spells and Rituals are designed either to enhance living in
these inhospitable areas or to aid hunting and defending themselves
from arctic predators. Ice Mages are seldom seen in warm climates, and
have been unpopular with Philosophers of magic who cannot agree
whether to place Cold as an element in its own right.  The College of
Ice Magics has been likened to the Colleges of Water, due to its
connections with Ice and Snow, Fire, as an antithesis, Air, due to its
weather-like effects and Mind, due to the peculiar mental fortitude
its Adepts have demonstrated while living in cold climes.

\subsection{Philosophy}

Ice Mages tend to be of a solitary and quiet nature, often having
lived their lives isolated in areas of sparse population density.
Many are hunters or only part-time Mages.  They generally have fair
relations with most Colleges, but relations with the Fire and Water
Colleges are distinctly icy.

\subsection{Traditional Colours}

Ice Mages are not known to dress flamboyantly.  Whites and Greys are
popular colours when dressing in warmer climes, perhaps with a hint of
pale blue.  However in their own element they usually wear undyed
leathers and furs.

\subsection{Notes}

All the Ice and Snow generated by the spells of this college is real
unless otherwise specified. It is in no way magical, and does not have
a duration. External effects (e.g. heat) will begin to affect it
immediately upon its generation.

Ice Magics is considered “opposite” to Fire Magics.  Many Ice spells
and effects are uniquely vulnerable to the spells of the Fire College,
and to Fire. Additionally Ice Magics have some spells and effects that
are especially effective versus Fire and Fire creatures.  This in no
way affects the “opposite” relationship that the Colleges of Water and
Fire Magics have to each other. Water Magics and Ice Magics are not
considered related, rather they are generally regarded as antagonistic
rivals.


\section{Restrictions}

Adepts  of  the  College  of  Ice  Magics  may  practise 
their arts without restriction. 

The MA requirement for this College is 13. 


\section{Base Chance Modifiers}

All base chances of Ice Mages are affected by temperature. Only one of
the temperature modifiers may be applied at one time.  Consult the
Weather Scale Table for Weather Gauge details.

\begin{dqtblr}{colspec={XXXX}}
Temp		& Gauge		& Temp °C	& Mod. \\
Very Cold	& 0--3		& ≤ -5		& +10 \\
Cold 		& 4--5		& ≤ 5		& +5 \\
Average		& 6--8		& 6--24		& 0 \\
Hot		& 9--10		& ≥ 25		& -5 \\
Very Hot	& 11+		& ≥ 35		& - 10 \\
\end{dqtblr}

If the Adept is standing on Ice or Snow in an area where it is in
abundance they gain an additional +5\%.


\section{Talents}

\begin{talent}[T-1]{Cold Affinity}

\multiple{100}
\begin{effects}
This talent allows the Adept to ignore the deleterious effects of low
body temperature.  The adept is treated as if having a Resist Cold
spell of equal Rank to the Rank in this talent in effect at all
times. Should the Adept also be under the effect of a Resist Cold
spell, the higher of the two ranks is in effect.
\end{effects}
\end{talent}

\begin{talent}[T-2]{Endure Hardship}
\multiple{150}
\begin{effects}
This talent allows the Adept to function capably even in harsh and
forbidding environments.  The Adept may go without food (but not
water!) for 1 (+1 / Rank) days every three months with no ill
effects. These days may be taken singly or consecutively, and the
Adept need not consume extra food later to make up for this time spent
fasting.  The Adept may additionally increase the Base Chance of any
concentration checks made in hostile environments, when the
concentration checks are reduced below 4 × WP due to environmental
effects, by 1 × WP / 5 full Ranks the Adept has in this talent, up to
a maximum of 4 × WP.  Finally, the Adept may reduce high fatigue rates
due to environmental and weather effects, other than those relating to
heat or fire, by a 1 row shift per 10 full ranks the Adept has in this
talent on the Rate of Exercise Chart (see §58.1, under the Fatigue
and Encumbrance Chart), down to a minimum rate of medium fatigue, or
light fatigue at rank 20.  For example, at Rank 10 a Strenuous climb
up a mountain may be treated as if it is only a Hard climb.
\end{effects}
\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Extinguish Fires}

\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{100}
\basechance{50\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Volume}
\begin{effects}
When successfully cast, this spell will extinguish all fire in a 10
foot (+ 10 / Rank) radius sphere, by smothering them with ice
crystals.  All the volume affected must be within range of the
spell. If the range is doubled or tripled the volume may likewise be
increased.  Magical fires are not affected.  This spell is identical
to the Fire College “Extinguish Fires”, except that the effect is
achieved by physical means (this is merely a cosmetic difference).
\end{effects}
\end{spell}

\begin{spell}[G-2]{Freeze}
\range{5 feet + 5 / Rank}
\duration{1 day + 1 / Rank}
\multiple{50}
\basechance{40\%}
\resist{None}
\storage{Investment}
\target{Object}
\begin{effects}
The Adept may freeze one object of up to 5 lb.  (+ 5 / Rank).  This
freezing will protect the object from decay while the duration lasts.
While frozen the object will be as cold as ice to the touch and will
drip slightly from condensation. When the duration has expired the
object will defrost at the rate of 1 minute per pound of weight.
\end{effects}
\end{spell}

\begin{spell}[G-3]{Ice Creation}
\range{15 feet + 10 / Rank}
\duration{Immediate}
\multiple{100}
\basechance{25\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This spell creates a film of ice 1 inch thick in a single square of
dimensions 1 (+ 1 / Rank) foot each side or a single cube of ice of
dimensions 6 (+ 6 / Rank) inches cubed. The ice must be created on the
ground and not on top of an entity.  It is non-magical and will
persist until melted, etc.
\end{effects}
\end{spell}

\begin{spell}[G-4]{Ice Traversal}
\range{10 feet + 10 / Rank}
\duration{20 minutes + 20 / Rank}
\multiple{125}
\basechance{30\%}
\resist{None}
\storage{Investment}
\target{Entity}
\begin{effects}
This spell enables 1 target (+ 1 / 4 full Ranks) to travel over ice
and/or snow without slipping or sinking in, as if it were normal earth
and/or rock. For example, this would enable climbers to climb icy
slopes.  Quadrupeds are treated as two targets for the purposes of
this spell.  In addition, if the terrain travelled on is flat ice,
each target’s TMR is increased by 1 (+ 1 / 3 full Ranks) while on the
ice.  See Travel on Ice (22.8), for additional detail.
\end{effects}
\end{spell}

\begin{spell}[G-5]{Refrigeration}
\range{25 feet + 5 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{50}
\basechance{35\%}
\resist{None}
\storage{Investment}
\target{Volume}
\begin{effects}
The caster may cause the ambient temperature of one 15 × 15 × 15 foot
cube to lower by 2°C / Rank.
\end{effects}
\end{spell}

\begin{spell}[G-6]{Resist Cold}
\range{Touch}
\duration{1 hour + 1 / Rank}
\multiple{100}
\basechance{40\%}
\resist{None}
\storage{Investment, Ward, Potion}
\target{Entity}
\begin{effects}
This spell protects the target from the effects of cold temperature by
increasing the Gauge by 1 (+ 1 / 4 full Ranks) up to a maximum of
Gauge 7 (Comfortable). It will totally protect the target from the
effects of Hypothermia at Rank 11+.  In addition, the target suffers 1
(+ 1 / 4 or fraction Ranks) less damage due to magical or non-magical
cold based attacks.  This spell is identical to the special knowledge
Air college spell of the same name.
\end{effects}
\end{spell}

\begin{spell}[G-7]{Snow Shovel}
\range{Self}
\duration{Concentration: maximum of 15 minutes + 15 / Rank}
\multiple{125}
\basechance{20\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
This spell enables the Adept to clear a path along snow or ice
obstructed ground and/or to tunnel through snow and ice.  Any snow or
ice up to 2 feet in front of the Adept undergoes a change in density
to dry snow, and is pushed to either side, leaving a gap 2 feet wider,
and higher (if applicable), than the size of the Adept.  No more than
1 hex (+ 1 / 2 full Ranks), may be cleared per pulse in this manner.
This effect moves with the Adept.  The Adept may lean in order to
direct the path up or down.  Note that the walls and roof of a tunnel
through snow or ice are merely packed snow and do not confer any
particular structural support or stability. See Travel on or through
Snow (22.8) for additional detail.
\end{effects}
\end{spell}

\begin{spell}[G-8]{Water to Ice}
\range{10 feet + 10 / Rank}
\duration{Immediate}
\multiple{100}
\basechance{15\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Volume of Water}
\begin{effects}
The Adept can freeze up to 10 (+ 10 / Rank) cubic feet of existing
water based liquids into solid ice, or into snow of a density chosen
by the Adept. All the water to be transformed must be within the
Adept’s range at the time of casting.  This spell may not be cast on
or near entities or their possessions.
\end{effects}
\end{spell}

\section{General Knowledge Rituals}

\begin{ritual}[Q-1]{Create Igloo}
\duration{1 hour + 1 / Rank}
\multiple{100}
\basechance{MA + 4\% / Rank}
\begin{effects}
The Adept must spend one hour in ritual construction of a miniature
dome made out of snow or ice cubes. At the end of this time the Adept
must make a successful cast check.  If successful, the dome swells in
size to become an igloo of internal size 5 (+ 1 / Rank) feet radius
and 2 feet thick. This ritual cannot backfire.  The Igloo has a single
entrance which is chosen by the Adept to be up to half its internal
height in both height and length.  The inside temperature of the igloo
always counts as very cold (-10 degrees) and the following
enchantments apply to objects or entities while they remain inside the
igloo:
\begin{Itemize}
\item All entities and creatures are treated as having a resist cold
  spell upon them of equal rank to that of the Adepts Rank in this
  ritual

\item All organic objects are preserved from decay
\end{Itemize}

In addition the igloo counts as bound snow while the duration is in
effect and will not melt or break due to non-magical forces (although
magical attacks affect it as normal).  Once the duration runs out
the igloo reverts to a normal (non-magical) igloo and will thereafter
melt, collapse etc.  as normal due to external conditions.
\end{effects}
\end{ritual}

\begin{ritual}[Q-2]{Bind Ice and Snow}
\duration{ Concentration: Maximum 1 hour + 1 / Rank}
\multiple{750}
\basechance{MA + 4\% / Rank}
\begin{effects}
The Adept may bind all of the ice and snow within a 5 (+ 5 / Rank)
feet radius circle of the Adept.  The results of this ritual are
similar to those for the binding of other elements. The Adept gains
control of all of the facets of the element.  The Adept may move or
shape the ice and snow, change its consistency and instil intelligence
in it as desired.  Finally, the Adept may sacrifice a point of MA
(this may be bought back with EP) in order to make a part or all of
the bound ice and snow permanently bound. In this instance, the bound
ice and snow is non-intelligent but magical, and is enduring. Almost
no magical or physical force will affect it (e.g.  it resists Wizard’s
Eye and Telepathy), with the exception of magical heat and fire,
against which it has 100 times the resistance of ordinary ice and
snow, and if any part of it remains it will (slowly) regenerate.
\end{effects}
\end{ritual}


\section{Special Knowledge Spells}

\begin{spell}[S-1]{Armour of Ice}
\range{Touch}
\duration{30 minutes + 30 / Rank}
\multiple{250}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
The target of this spell is covered by a magical Armour of Ice, which
provides 5 points of Armour Protection (+ 1 / 4 full Ranks) which
fluctuates according to the current temperature: the Armour gains +2
points of Armour Protection when the temperature is very cold, +1 when
it’s cold, -1 when it’s hot, and -2 when it’s very hot (see 22.2). The
armour has a weight rating of 5 (see 56.3) and subtracts 2 from AG and
20 from stealth.  Ice Armour may not be cast on entities wearing
armour.  This spell will stack with other defensive spells.  The Ice
Armour is vulnerable to Fire, and has an ablative effect; it will
absorb up to half of any fire damage taken, but for every 10 points of
damage taken (before halving) one point of protection is removed from
the armour, and if the protection is reduced to zero the spell is
immediately dissipated.
\end{effects}
\end{spell}

\begin{spell}[S-2]{Icy Transformation}
\range{10 feet + 10 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{300}
\basechance{25\%}
\resist{Special}
\storage{Investment, Ward, Magical Trap}
\target{One (metal or mineral) Object}
\begin{effects}
This spell turns one metal or mineral object of up to 5 lbs (+ 5 /
Rank) entirely into ice (a wall of iron would have to be entirely
turned to ice but a single brick in a wall could be transformed).  The
object is then transparent and vulnerable to damage, heat etc.  At the
end of the duration, the object will revert back, but any damage will
not be repaired.
\end{effects}
\end{spell}

\begin{spell}[S-3]{Freezing Wind}
\range{5 feet + 5 / Rank}
\duration{5 seconds + 5 / Rank}
\multiple{225}
\basechance{30\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Volume}
\begin{effects}
This spell causes Arctic conditions to prevail in a 10 (+ 1 / Rank)
feet cube.  Any entity within this cube which fails to resist will
suffer [D - 4] (+ 1 / 2 full Ranks) points of magical cold damage per
pulse.  Creatures of fire (efreet, salamander or elemental) take half
damage even if they resist.
\end{effects}
\end{spell}

\begin{spell}[S-4]{Frostbite}
\range{50 feet + 25 / Rank}
\duration{Immediate}
\multiple{200}
\basechance{20\%}
\resist{Special}
\storage{Investment}
\target{Special}
\begin{effects}
 This spell may either be used to kill (5 × [Rank])\% of all
 non-sentient plants within the Adepts range, or the Adept may target
 and kill 1 individual plant (+ 1 / Rank).  There is no visible
 effect, but death due to frostbite occurs at the time of casting and
 no amount of non-magical effort will revive plants affected by this
 spell.  Sentient plants take 1 (+ 1 / Rank) points of damage (instead
 of dying), and may passively resist for no damage. Other plants that
 are magical or especially resistant to cold may also be entitled to a
 passive resistance roll versus the spell’s effects.
\end{effects}
\end{spell}

\begin{spell}[S-5]{Frozen Doom}
\range{10 feet + 10 / Rank}
\duration{Immediate}
\multiple{500}
\basechance{5\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
This spell uses magical cold to freeze solid the blood of one target
entity, which may be up to 1 hex (+ 1 / 2 full Ranks) in size, killing
it instantly if they fail to resist.
\end{effects}
\end{spell}

\begin{spell}[S-6]{Hibernation}
\range{5 feet}
\duration{Special}
\multiple{250}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
The target of this spell is placed in suspended animation for up to
([Rank × Rank]) days as specified by the Adept.  At Rank 20 there is
no maximum and the Adept may choose any duration.  All bodily
functions including ageing, are suspended for the duration of the
spell and the target feels cool to the touch.  The target is immune to
cold and suffocation, and takes no more damage from existing injuries
while this spell is in effect.

Additional injuries will still affect the target, but any damage that
would occur due to bleeding, poison etc. is ignored. When the spell
duration runs out, or the spell is dispelled, the target awakens with
physical strength reduced by 1 / full week hibernated and immediately
begins to suffer from any existing injuries and conditions (poison,
disease, shock, bleeding and suchlike).  Physical strength may be
regained at a rate of 1 point per day, and is not reduced below
1. Entities that naturally hibernate suffer -20 to their Magic
Resistance vs. this spell.
\end{effects}
\end{spell}

\begin{spell}[S-7]{Ice Bolt}
\range{20 feet + 10 / Rank}
\duration{Immediate}
\multiple{300}
\basechance{35\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Entity, Object}
\begin{effects}
The Adept creates a 2 foot long, 2 inch diameter non-magical bolt of
ice which is projected at a target with a Strike chance of 30\% (+ 3 /
Rank) + MD, modified by range as if it were a heavy crossbow.  The ice
bolt strikes as an A class weapon doing [D + 4] (+ 1 / Rank) damage
which can stun and inflict specific grievous injuries. Once the spell
is cast the target gets no magic resistance, but the target’s defence
is subtracted from the chance to hit, since the effect of the spell is
to create a physical bolt.  A double or triple effect cannot affect
damage but may add +10\% and +20\% to the Strike Chance respectively.
\end{effects}
\end{spell}

\begin{spell}[S-8]{Ice Construction}
\range{15 feet + 5 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{225}
\basechance{15\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Volume}
\begin{effects}
The Adept may conjure 30 (+ 30 / Rank) cubic feet ice in up to 1 +
Rank shapes of the Adept’s choice. The shapes always appear entirely
within the range of the Adept and may not appear above or inside
(partially or wholly) any entity.  Each shape must appear on the
ground in a stable fashion (not about to topple over) and must have a
minimum thickness of 6 inches in any part. When the spell duration
expires, the ice disappears (returns to whence it came).
\end{effects}
\end{spell}

\begin{spell}[S-9]{Iceberg}

\range{10 feet + 10 / Rank}
\duration{ Concentration:  Maximum  30  minutes  + 30 / Rank}
\multiple{150}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area of Water}
\begin{effects}
The Adept creates a polyhedral iceberg of dimensions 10 (+ 2 / Rank)
feet cubed. It may only be successfully created in an existing volume
of liquid sufficient to hold it without it touching the bottom. 8/9 of
the Iceberg will be submerged. The Iceberg will be flat topped and may
stably support up to 1 hex of entities (+ 1 / Rank) (Additional
entities will cause it to roll in the water). While the Adept is in
contact with the Iceberg the Adept may move the Iceberg at a speed of
5 (+ 1 / Rank) miles per hour.  The Iceberg spell also has a calming
effect on the water around it, reducing the size of all waves up to
Rank feet away by Rank feet.
\end{effects}
\end{spell}

\begin{spell}[S-10]{Ice Pack}
\range{10 feet +5 / Rank}
\duration{Immediate}
\multiple{150}
\basechance{15\%}
\resist{None}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
This spell immediately halts any damage being taken due to shock or
blood loss and the target gains 1 / Rank to their chance of recovering
from stun or unconsciousness (if applicable). It will restore a (live)
character on negative endurance to zero, although any further damage
will start the process of shock and bleeding as usual. In addition, if
the target is suffering from the adverse effects of a fright roll or
similar emotional effect, they get a second roll (in some cases a
second resistance roll) to recover from (“or snap out of”) it with
+1\% / Rank added to their dice roll.  The cause of the fright or
shock may have been magical or otherwise.  This spell may be recast
on each target as often as desired. This spell will not work on
regenerating entities (including those under the effects of a
trollskin spell).
\end{effects}
\end{spell}

\begin{spell}[S-11]{Ice Projectiles}

\range{20 feet + 5 / Rank}
\duration{Immediate}
\multiple{300}
\basechance{30\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{1 Entity + 1 / Rank}
\begin{effects}
Each entity targeted by this spell must resist or suffer [D - 4] (+ 1
/ Rank) points of magical damage due to being pierced by flying A
class shards of ice (armour does absorb damage but there is no AG
loss).  If this spell is doubled the adept may not double damage but
may choose to have the ice projectiles ignore armour instead.  If this
spell is tripled, the adept may roll for a possible A class specific
grievous injury for each target that failed to resist, in addition to
ignoring armour as above (and still doing fatigue damage except as
part of the grievous result), or reduce the targets’ resistances as
usual.
\end{effects}
\end{spell}

\begin{spell}[S-12]{Ray of Cold}
\range{30 feet + 15 / Rank}
\duration{Immediate}
\multiple{300}
\basechance{30\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity, Object}
\begin{effects}
This spell projects a blast of intense magical cold at the target.
The ray of cold will impact either on the target or on the first
obstruction blocking the path from the Adept to the target.  Anything
struck by the ray must either resist or suffer [D + 1] (+ 1 / Rank)
damage (resist for half damage).
\end{effects}
\end{spell}

\begin{spell}[S-13]{Snowball}
\range{10 feet +10 / Rank}
\duration{Concentration: Maximum of 10 minutes + 10 / Rank}
\multiple{200}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Adept, Special (chosen each pulse)}
\begin{effects}
This spell causes the ground within 5’ of the Adept to be instantly
covered with 5 inches of conjured snow (returns to whence it came when
the spell duration expires). Snowballs then form themselves out of the
snow and launch themselves at a target or targets of the Adepts
choice.  Up to 1 target (+ 1 / 4 full Ranks) can be pelted with a
flurry of snowballs each pulse.  A different set of targets may be
chosen by the Adept each pulse.  Each target must resist or suffer a
reduction of -1\% (1 / 2 full Ranks) to their Base Chance of doing
anything while being snowballed and in addition must make a 4 × WP
roll (3 × WP if Adept is above Rank 10 in this spell) to perform any
action that involves concentrating (e.g.  casting a spell).  Each
target need only resist this spell once.  If the Adept leaves the 5’
diameter circle of snow while this spell is in effect, the spell is
automatically dissipated.
\end{effects}
\end{spell}

\begin{spell}[S-14]{Snowfall}
\range{10 feet + 10 / Rank}
\duration{10 minutes + 5 / Rank}
\multiple{200}
\basechance{40\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
This spell causes snow to begin gently falling in an area of between 5
feet minimum and 40 feet (+ 5 / Rank) maximum diameter (chosen by the
adept).  The snow will form in the air at between 10 and 40 feet
above the ground, and will gently float to the ground.  The area
within this spell, if above zero degrees in temperature, will be
magically cooled to zero degrees for the spell duration (nb.  This
will not count as a positive modifier to Ice mage base chances as it
is magical, although it may reduce negative modifiers to zero).  For
each 5 minutes that this spell is in effect, one inch of powder will
cover the ground. At the end of the spell the snow will remain, but
the temperature will be restored to normal (and the snow may begin
melting).  Note that if this spell is, for example, cast inside a
house with a 20 foot ceiling, as the snow is formed in the air between
10 and 40 feet up, only a third of the snow will fall in the room.
The other two thirds will appear above the 20 foot line and will fall
on the roof, with the exception of a small amount that appears in the
attic.
\end{effects}
\end{spell}

\begin{spell}[S-15]{Wall of Ice}
\range{20 feet + 10 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{200}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
This spell conjures a wall of ice that is 10’ high, 20’ long and 2’
thick, or a pillar 15’ high with a 3’ diameter.  The adept may choose
to increase the height of the wall by 1’, the width by 2’ or the
thickness by 6", or the height or diameter of the pillar by 1’, for
each Rank the Adept possesses in the spell. The wall may be uniformly
curved up to a half circle. The wall may not be created on top of an
entity, and is subject to the usual restrictions on physical walls.
The wall is translucent but not transparent. When the spell duration
expires the ice returns to whence it came.
\end{effects}
\end{spell}

\begin{spell}[S-16]{Weapon of Cold}
\range{10 feet + 5 / Rank}
\duration{5 minutes + 1 / Rank}
\multiple{250}
\basechance{30\%}
\resist{None}
\storage{Investment}
\target{Weapon}
\begin{effects}
The weapon over which the spell is cast becomes infernally cold
without harm to either the weapon or the user of it. The base chance
of hitting with the weapon is increased by 1 (+ 1 / Rank) and the
damage done by the weapon is increased by 1 (+ 1 / 3 full Ranks).
This amount of damage is tripled if the object of the attack is a
creature of fire.
\end{effects}
\end{spell}

\begin{spell}[S-17]{Winter Garden}

\range{20 feet + 10 / Rank}
\duration{2 weeks + 2 / Rank}
\multiple{100}
\basechance{35\%}
\resist{None}
\storage{Investment, Ward}
\target{Plant}
\begin{effects}
This spell bestows resistance to cold on one plant (+ 1 / Rank) or 25
square feet (+ 25 / Rank) in one patch of the same species of plant of
the Adepts choice.  Plants with this resistance are immune to the
Frostbite spell and will thrive even in permafrost and arctic
temperatures.
\end{effects}
\end{spell}


\section{Special Knowledge Rituals}

\begin{ritual}[R-1]{Snow Simulacrum}
\duration{ Concentration: Maximum 1 hour + 1 / Rank}
\multiple{300}
\basechance{MA + 3\% / Rank}
\resist{Special}
\begin{effects}
The Adept spends an hour forming a human or animal figure, no larger
than 1 hex (+ 1 / 3 full Ranks) in size, out of snow, which must be
already present.

The animated sculpture will have the same characteristics as the
sculptured entity, except that all characteristics are reduced by
25\%. This will generally cause abilities to be lessened by a like
amount (e.g.  Flying speed is reduced by 25\% (if the entity can fly)
due to Agility being reduced, Damage from attacks is reduced by 25\%
due to Strength reduction, Base Chance to hit is reduced because of MD
reduction, etc.).

Magical abilities are not copied in the copied form, although
quasi-magical abilities may be. The simulacrum will not bear a close
resemblance to any other figure unless the figure chosen to be
duplicated is present during the ritual.  In the latter instance the
figure being duplicated may choose to actively resist the ritual;
otherwise the ritual is nonresistible.  The simulacrum has only animal
intelligence, and the Adept may give it simple instructions or
actively concentrate to control its movements (requires the Adept to
perform pass actions).

The simulacrum will have vague inclinations relating to its original
(borrowed) form, and although no longer comprised of snow (it takes on
a fleshy appearance, or whatever is appropriate for the original
entity) will have an adverse reaction to the presence of heat and
flame and will take an additional 2 points from heat and flame
attacks.  The simulacrum normally resembles a generic example of its
borrowed form except that it has a snowy sheen to it and is cool to
the touch.  Clothes, possessions and suchlike are not duplicated by
this ritual.

When the Adepts concentration ceases, the simulacrum will collapse
into a pile of ordinary non-magical snow.
\end{effects}
\end{ritual}


\begin{ritual}[R-2]{Summoning and Controlling Ice Elemental}
\duration{ Concentration: Maximum 1 hour + 1 / Rank}
\multiple{250}
\basechance{MA + 5\% / Rank}
\begin{effects}
The Adept summons an Ice Elemental from its home plane (conjectured to
be the elemental plane of Cold), which will appear within 20 feet.
This ritual takes 2 hours, and may only be performed in temperatures
of 0 degrees or less.  If the ritual backfires the Elemental arrives
uncontrolled and will attack the Adept (and others nearby); however,
if the ritual is successful the Elemental is controlled and must obey
the Adept’s every whim.  The Elemental remains until dispelled, which
the Adept may do by successfully casting a Special Knowledge
counterspell at it. 1\% is added to the base chance of success on this
ritual for each point of Willpower the Adept has over 15.  Ice
Elementals are similar to other Elementals in that they do not
normally exist on this plane, but are summoned by Ice magics. They
will always be hostile to their summoner and will attack if
uncontrolled.  Ice Elementals are impervious to attacks with
non-magical weapons.  Magic does affect them. Ice Elementals are
vulnerable to fire and can be damaged by attacks involving this
“opposite element”.  Ice Elementals have a combined endurance and
fatigue of 15 (+ 6 / Rank), which must be divided in such a way that
they fall into the ranges indicated below.  Any reference to Rank
below refers to the Adept’s Rank in this ritual.

Ice Elementals have the following characteristics: 
\begin{Description}
\item[Habitat] Other Planes  

\item[Frequency] Very Rare  

\item[Number] 1 

\item[Description] Ice Elementals appear as lean crystalline humanoids
  with frosty hair and silvery blue eyes.  They are half as tall, in
  feet, as their Endurance.

\item[Talents, Skills and Magic] Ice Elementals can disappear into Ice
  with only a 10\% chance of being detected. They can freeze water
  within line of sight at a rate of one cubic foot per pulse for every
  point of Physical strength available to the Elemental (entities
  within the area get a 3 × AG check to avoid being caught and trapped
  while Ice forms around them), and they can cast a Ice construction
  and Wall of Ice at a Rank equal to their summoner’s Rank + 4 in this
  ritual.  These are talents and cost no fatigue. They may expend 2
  fatigue to fire an Ice bolt (as per the Spell) striking as an A
  class weapon with a base chance equal to the Elemental’s combined
  maximum Endurance and Fatigue, and doing [D + 4] (+ 1 / Rank) points
  of damage.  Since this is a physical attack formed from the Ice
  Elemental, the Ice bolt remains after firing.

\item[Movement Rates] Running 200 + 10 × Summoner’s Rank

\item[Weapons] Ice Elementals strike their opponents with open hands
  and pierce them with their long icy fingers. They can attack twice
  in the same pulse without penalty doing [D + 3] (+ 1 / Rank) damage
  per strike.
\end{Description}

\begin{dqtblr}{colspec={XX},
    cell{5-7}{1}={c=2}{l}}
PS: 20 + 5 / Rank	& WP: 14 - 18 \\
MD: 20 - 25		& EN: 5 - 50 \\
AG: 15 - 20		& FT: 10 - 85 \\
MA: None		& PC: 15 - 20 \\
PB: 8 - 10		\\
TMR: 4 + 1 / 5 Ranks	\\
NA: Skin absorbs 5 DP	\\
\end{dqtblr}
\end{effects}
\end{ritual}

\begin{ritual}[R-3]{Ritual of Winter}

\duration{Concentration: Maximum 1 hour + 1 / Rank}
\multiple{350}
\basechance{MA + 3\% / Rank}
\begin{effects}
The Adept may change one or more of the three components which make up
the weather by performing a ritual dance. The three components of
weather are:
\begin{enumerate}
\item Precipitation, Degree 
\item Temperature, Gauge 
\item Wind, Force 
\end{enumerate}

The GM should consult the weather table and advise the player of the
current level of these three components before the Adept starts
performing the ritual.  The Adept may change the current components by
1 (+ 1 / 2 full Ranks).  All the changes may be in any direction on
the table with the proviso that the Adept may never raise the
temperature, and must lower it by at least one degree.  All weather
within 2 miles (+ 2 / Rank) of the Adept is affected by the
ritual. This ritual cannot backfire.
\end{effects}
\end{ritual}

\section{The Element of Ice}

\subsection{Ice}

Ice weighs 47.2 lb. per cubic foot and is translucent 
(but not transparent). 

\subsection{Breaking through Ice}

This applies particularly to the spells Ice Creation, Ice Construction
and Wall of Ice, but may also be used in natural settings (for example
a single inch of ice may cover a lake).

Ice is deemed to have 10 points of fatigue per inch, 120 points per
foot. It takes only half damage from being hit by B class weapons,
full damage from axes, fire and water based attacks, and double damage
from picks (note that for fire based attacks bonuses against cold
creatures apply, for those spells that have them).  Other weapons only
do a single point.  Endurance blows do double damage.  The exception
is when a blow by any weapon exceeds the remaining fatigue of the ice;
in this case all damage is applied (the ice breaks).  Successful
elimination of the ice’s fatigue makes a human sized hole in the ice,
big enough for a one hex entity to go through.  Multiple entities may
attempt to break through the same area of ice, within reason.  Smaller
holes may be made but for the purposes of simplicity are no quicker to
make (when digging far into the ice at least a human sized hole must
be made in order to keep working anyway).

\subsection{Travel on Ice}

Bipeds must travel at half TMR (round up) while on ice or make a 2 ×
AG roll each pulse or go prone.  Quadrupeds may make a 4 × AG roll to
travel at full speed, or go prone. Subtract 1 × AG if the ice is wet.
Entities going prone may also slide up to half their TMR along the ice
in the pulse in which they fell, depending on how much of their
movement was in one direction (GM’s discretion).

\subsection{Travel on or through Snow}

Travel through snow usually causes entities to lose 1/4 of their TMR
per foot of powder, down to a minimum TMR of 1 (unless the powder is
higher than the entity is tall).  In addition there is an encumbrance
shift of one column on the encumbrance table for each 1/4 TMR slowed.
Note the reference to powder — if the snow is denser than freshly
fallen powder the effects may be less.  These figures are based on a
human sized entity; a giant (for example) would only suffer the
effects of 4 foot of powder (TMR reduced to 1) when up to its chest;
that is, in about 13 feet for a Stone Giant (normally 20 feet
tall). Quadrupeds tend to manage snowy conditions well, and should be
treated as if they were standing erect for the purposes of height
calculation. Hobbits, and other creatures with large feet, may have
these penalties halved at the GM’s discretion.  GM’s may also allow
items such as snow shoes; a suggestion is a straight halving of TMR
but the entity does not sink into the snow.  Note that such items are
not known or common in warmer climes!

\end{College}
