\begin{Chapter}{The College of Illusions (Ver 1.5)}

The College of Illusions is concerned with truth, deception and
sensation.  Practitioners of the College (or Art) of Illusions are
called Illusionists.  This College is one of extremes, of subtlety and
flamboyance, and especially of deception and honesty.  Illusionists
are taught that they are privy to one of the most potent truths known
to the sentient races, namely that our senses are the only window to
the existence of reality.  Based on this doctrine, they change the
appearance of reality rather than the actual nature.  This often leads
to more impressive but superficial effects than other Colleges. These
effects have meant that Illusionists have historically been associated
with stage magicians and sleight-of-hand artists, which has degraded
their status in the eyes of other Adepts.  Most accept their stage
heritage, and revel in the layers of deception involved in revealing a
great truth to their audience.  Some, however, become enmeshed within
their own fantasy and fail to differentiate between this and reality.
Others deny this deception totally, and turn to death as the only
fixed point in a changing world. These Adepts use people’s dependency
on their senses to cause death and despair.

\subsubsection{Traditional Colours}

The more flamboyant of the Illusionists will often wear bright, even
garish clothing incorporating as many colours as possible.  These
colours will usu- ally not clash, but the Illusionist wearing them
will stand out in any crowd.  At the very least, most Illusionists
will wear extremes of colours: fire, blood, sable, snow, charcoal,
emerald, gold, etc.

\subsubsection{Traditional Symbols}

The College of Illusions is associated with a plethora of symbols
adopted from individual performers. Theatrical symbology is common.

\subsubsection{On Illusions}

Most Magics in this College are Illusions. Illusions cannot be
“disbelieved”; the images are there and will interact with light and
sound in the same manner as the objects they represent. Mirrors and
light sources may not be created.  Unless stated, the Illusions will
not be substantial, and any solid objects or spells will pass through
them.  Illusions will interact with each other as if real.  Illusion
spells may be dispelled by anyone with the appropriate counterspell,
although Illusions on an entity gains their passive magic resistance.

\section{Restrictions}

Adepts of the College of Illusions may practice their arts without
restrictions.

The MA requirement for becoming an Adept of the College of Illusions
is 13.

\section{Base Chance Modifiers}

The Base Chance of performing any talent, spell or ritual of the
College of Illusions may be modified by conditions.

As the successful performance of Illusion magic is as much finesse and
presentation as raw power, a knowledge of the Performing Arts is
considered essential by many Adepts.  Furthermore, willing a change in
reality to the level of detail necessary for deception, requires good
concentration and Willpower.  The following conditions modify the
base chances of the College of Illusions:

\begin{tabularx}{\columnwidth}{Xr}
For each point the Adept’s Willpower varies from 15	&  +/- 1 \\
For each two full Ranks in Troubadour			& + 1 \\
\end{tabularx}

\section{Talents}

\begin{talent}[T-1]{Concealed Casting}

\multiple{150}
\begin{effects}
This Talent allows the Adept to cast without any apparent movement or
noise.  The Adept may not speak or make meaningful or extravagant
actions. They may not cast without a hand free or if mute.  If an
observer is intent on the Adept (e.g. in conversation with), the
observer may make a (2 × PC) - Rank check to notice. All aspects of
the Spell (including Base Chance) must be cast at the lower of the
Talent’s and Spell’s ranks, and the Base Chance is further reduced by
20\%. The Talent only functions with spells of the College of
Illusions.
\end{effects}
\end{talent}

\begin{talent}[T-2]{Enhanced Vision}

\multiple{225}
\begin{effects}
The Adept may see through any obscurement such as smoke, dust, water,
fog, cloud or rain (though not darkness) to a distance of 10 feet (+
10 / Rank) as if unobscured.  Normal penalties then accrue from this
range. The Adept may see objects or entities that are invisible — they
appear to have a slight blue sheen around them.  If the invisibility
effect (excluding Walking Unseen) is of a higher Rank than the
Enhanced Vision, the object or entity may not be clearly identified or
directly magically targeted. The Adept has a Base Chance of PC (+ 5 /
Rank) of Detecting Illusions if they take a Pass Action --- each
Illusion may only be checked once.  Detecting an Illusion does not
nullify its effects.
\end{effects}
\end{talent}

\begin{talent}[T-3]{Project Image}

\multiple{200}
\begin{effects}
This Talent creates a visual Illusion of the Adept, who becomes
Invisible as per (G-4) except that attacking in melee will always
disrupt this Talent.  Initially, the image forms over the Adept.  The
image will reflect the current appearance of the Adept, in any desired
stance.  The image may only be changed or moved (at TMR) when the
Adept takes a pass action. The image must remain in line of sight and
within 15 feet (+ 15 / Rank), and may not pass through solid objects.
At Rank 5, the image may make appropriate environmental noises,
e.g. footsteps.  At Rank 10, the Adept may move while maintaining the
relative distance and direction to their image.  At Rank 15, the Adept
may project their voice through the image. At Rank 20, the Illusionist
may also see through the image.  The abilities gained at Ranks 15 and
20 require pass actions to use.
\end{effects}
\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Audile Illusion}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{125}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This Illusion allows the Adept to create up to Rank simple sounds,
each coming from a location fixed either absolutely or relative to an
object.  Whilst taking a pass action, the Adept may cause any
combination of these sounds to occur.  Sounds and locations are
defined as they are first used, not at time of casting.  No language
or musical instru- ment may be imitated.  The maximum volume at Ranks
0--4 is that of a stage whisper (usually clearly audible at 25’), at
Ranks 5--9, normal conversation (50’), at Ranks 10–14, shouting
(200’), at Ranks 15–19, screaming (500’), and at Rank 20, thunder (1
mile). The range at which the sound will be clearly heard is affected
by the ambient level of noise.
\end{effects}
\end{spell}

\begin{spell}[G-2]{Flash of Light}

\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{150}
\basechance{20\%}
\resist{Passive}
\storage{Potion}
\target{Self}
\begin{effects}
This spell creates a blinding flash of light emanating from the
Adept’s body.  This light is magical in nature, and may cut through
areas of magical darkness.  All entities not shielded by a solid
object and within range must either resist or be blinded for 5 seconds
(+ 5 / 2 full Ranks).  Blinded characters have their Strike Chance
reduced by 50, and they move at half their normal TMR.  The Adept is
normally not affected by this spell.
\end{effects}
\end{spell}

\begin{spell}[G-3]{Illusion of Food}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{100}
\basechance{50\%}
\resist{None}
\storage{Investment}
\target{Object}
\begin{effects}
The Adept may change the taste, appearance and smell of up to 1 cubic
foot of food or drink (+ 1 / Rank).  At Rank 11, the Adept may create
(Rank - 10) pounds of food or drink.  This magically created food will
recover fatigue for the duration of the spell.  The Adept’s Cooking
Rank will effectively be improved by 2, or to a minimum of Rank / 2.
No other properties (e.g.  poison, magic) may be changed or created
with this spell.
\end{effects}
\end{spell}

\begin{spell}[G-4]{Invisibility}

\range{15 feet + 15 / Rank}
\duration{5 minutes + 5 / Rank}
\multiple{450}
\basechance{30\%}
\resist{None}
\storage{Investment, Potion, Ward}
\target{Entity or Object}
\begin{effects}
The target becomes invisible.  Unless cast at rank 16 or higher, the
spell ceases whenever the target makes a strike check in melee or
close (whether successful or not). The target may choose to end the
spell at any time.

The effects of this spell are identical to the E \& E College spell of
the same name.
\end{effects}
\end{spell}

\begin{spell}[G-5]{Personalised Illusion}

\range{30 feet + 30 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{75}
\basechance{40\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This Illusion creates an image of a specific type peculiar to the
Adept.  At Rank 0, the image must be of a rock.  At each of Ranks 4,
8, 11, 14, 16, 18, 19 and 20, another image may be chosen.  The Adept
may select the colouring, size and conformation of the image at cast
time (e.g.  a closed iron chest, or an open wooden casket) up to a
maximum diameter of 5’ and height of 8’.  The Illusion is not affected
by physical laws except that it is opaque to light. The Adept may move
or rotate (though not manipulate) the Illusion by taking a pass
action.  The workmanship is as if the Adept were an Artisan of
one-third Rank.
\end{effects}
\end{spell}

\begin{spell}[G-6]{Illusion of Seeming}

\range{5 feet + 5 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{200}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward}
\target{Object or Illusion}
\begin{effects}
This Illusion creates a visual image of up to 2 cubic foot (+ 2 /
Rank) on an object, which will move with the object if the object
changes shape or position, and dissipate if the object is destroyed.
The Adept may move or change the image by taking a pass action to
concentrate on it.  The object has not been changed, and still has its
original physical properties.  The workmanship is as if the Adept were
an Artisan of one-third Rank.

If the target is an Illusion, the target may be changed to any other
form that it could normally take. Up to 1 cubic foot of the target (+
1 / Rank) may be changed during each pass action.  This change will
revert if the Seeming ceases before the target Illusion.  The Adept
may only affect their own Illusions.
\end{effects}
\end{spell}

\begin{spell}[G-7]{Static Illusion}

\range{15 feet + 15 / Rank}
\duration{30 minutes + 30 / Rank}
\multiple{200}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This spell creates a visual Illusion within a volume no larger than
(Rank + 3) feet by (Rank / 2 + 2) feet by (Rank / 3 + 1) feet.  The
Illusion is fixed at casting, and may not be changed or moved with
this spell. The workmanship is as if the Adept were an Artisan of
one-half Rank.
\end{effects}
\end{spell}

\begin{spell}[G-8]{Illusionary Wall}

\range{15 feet + 15 / Rank}
\duration{30 minutes + 30 / Rank}
\multiple{125}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This spell creates an Illusion of a non-solid, linear or smoothly
curved wall.  It may have all the characteristics of any other wall
spell (except solidity, damage or fear), or mimic an adjacent surface
instead. The wall is 20’ by 10’ by 12", with each Rank altering its
height or width by 2’ or thickness by 4".  The wall must be attached
along the entirety of one of its thinnest edges.
\end{effects}
\end{spell}

\section{General Knowledge Rituals}

\begin{ritual}[Q-1]{Illusionary Cloaking}

\range{5 feet + 5 / Rank}
\duration{6 hours + 6 / Rank}
\multiple{100}
\basechance{40\% + 4\% / Rank}
\target{Area}
\casttime{1 hour}
\material{None}
\concentration{Standard}
\begin{effects}
This ritual creates a subtle visual Illusion over all non-living
matter within range according to the Adept’s wishes.  Only surface
appearance is changed, and this ritual does not provide invisibility.
This means that objects will retain their silhouette, and the Illusion
will fill the same volume as the object.  The new appearance will fade
over the last hour of the duration. The workmanship is as if the Adept
were an Artisan of one-half Rank.
\end{effects}
\end{ritual}

\begin{ritual}[Q-2]{Illusory Fog}

\range{30 feet + 30 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{75}
\basechance{40\% + 4\% / Rank}
\target{Area}
\casttime{30 minutes}
\material{None}
\concentration{Standard}
\begin{effects}
This ritual creates Illusory fog to a height of 10 feet + 2 / Rank
within range.  The fog will rise within five minutes of the ritual’s
completion, centred on the spot where the Adept performed the ritual.
The fog has a visibility of 20 hexes (1 / Rank).
\end{effects}
\end{ritual}

\section{Special Knowledge Spells}

\begin{spell}[S-1]{Illusionary Animal}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{150}
\basechance{35\%}
\resist{None}
\storage{None}
\target{Area}
\begin{effects}
An Illusion of a non-magical, non-sentient creature of no larger than
10 lb + 10 lb × Rank squared is created.  The Illusion will have the
instincts of the creature it is based on, though it will obey most
simple mental instructions from the Adept.  This communication is
one-way, and may only occur while the animal is within range of the
spell. No other communication is possible — it has no mind.  The
animal has the minimum PS, MD, AG, TMR and PC for the selected
creature.  The animal is solid, and can bear weight.  It cannot attack
effectively, nor cause fear.  Only creatures previously observed by
the Adept may be created.  Smell, sound and feel are created as
appropriate. If damaged, the Illusion is dissipated. The Adept may
perceive through the animal’s senses as follows: Rank 3 taste, Rank 7
smell, Rank 11 touch, Rank 15 hearing, Rank 19 sight.
\end{effects}
\end{spell}

\begin{spell}[S-2]{Illusionary Bolt}

\range{15 feet + 5 / Rank}
\duration{5 seconds + 5 / Rank}
\multiple{200}
\basechance{30\%}
\resist{Active, Passive}
\storage{Investment}
\target{Entity, Object or Area}
\begin{effects}
An Illusion of any dangerous-looking object of “throwable” size is
created in the Adept’s hand. When thrown, it flies in a straight line,
doing [D-2] + 1 / 2 full Ranks magical damage to the first Entity or
object struck.  Illusions take double damage from this bolt.
\end{effects}
\end{spell}

\begin{spell}[S-3]{Illusory Creation}

\range{Touch}
\duration{10 minutes + 10 / Rank}
\multiple{250}
\basechance{20\%}
\resist{None}
\storage{Investment}
\target{Area}
\begin{effects}
This spell creates the Illusion of one small object. The object will
have the physical attributes of one of the following: cloth, leather,
wood, bronze to a maximum of 1 lb (+ 1 / 3 full Ranks).  The object
will be of a single form, with no moving or removable pieces, though
of any combination of colours.  The object will interact normally with
its surrounds.  The workmanship is as if the Adept were an Artisan of
one-third Rank.
\end{effects}
\end{spell}

\begin{spell}[S-4]{Illusion of Deep Pockets}

\range{Self}
\duration{1 hour + 1 / Rank}
\multiple{200}
\basechance{40\%}
\resist{None}
\storage{Potion}
\target{Entity}
\begin{effects}
This Illusion creates 1 magical pocket (+ 1 / Rank) about the Adept’s
clothing.  Each pocket can hold an object with a largest dimension of
no more than 1 inch (+ 1 / Rank).  No entity can be placed in a
pocket.  The total weight held may not exceed 1 lb (+ 1 / Rank), and
no individual object can weigh more than half this amount. The pockets
are not obvious, but thorough investigation may find them: searchers
get a (2 × PC) - Rank check to notice. Items within the pockets may
not be located by spells of a lesser rank. The pockets may only be
accessed while the Adept is clothed; however, a new set of clothes
will still contain the pockets. At the end of the spell, the pockets
expel their contents.
\end{effects}
\end{spell}

\begin{spell}[S-5]{Disguise Illusion}

\range{Touch}
\duration{1 hour + 1 / Rank}
\multiple{200}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment, Potion, Ward}
\target{Entity}
\begin{effects}
The Adept may change the target’s appearance.  The target may not be
located by location spells of a lesser rank. The target’s height may
vary by 1\% / Rank, and their weight by 2\% / Rank.  The target’s
voice or clothing may be changed at Rank 5 or both of these at Rank
10; an individual may be duplicated or gender or race changed at Rank
15, or two of these at Rank 20. If clothes are included in the spell,
they revert to normal if discarded. The disguised form has the PB of
the target +/2 / 3 full Ranks, within racial limitations.
\end{effects}
\end{spell}

\begin{spell}[S-6]{Illusionary Euphonia}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{150}
\basechance{25\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This Illusion creates the sound of harmonious voices or musical
instruments.  One instrument (+ 1 / 3 full Ranks) may be played, or
one voice (+ 1 / 6 full Ranks) may speak or sing. Each may come from a
different location fixed either absolutely or relative to an object.
The effective rank of the performance or conversation is Rank/2, to a
maximum of the Adept’s rank in the instrument or language.  Pass
actions are required for musical performance at a skill level above
Rank/5, or conversation. The maximum volume at Ranks 0--4 is that of a
stage whisper (usually clearly audible at 25’), at Ranks 5--9, normal
conversation (50’), at Ranks 10--14, shouting (200’), at Ranks 15--19,
screaming (500’), and at Rank 20, thunder (1 mile).  The range at
which the sound will be clearly heard is affected by the ambient level
of noise.
\end{effects}
\end{spell}

\begin{spell}[S-7]{Hallucination}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{225}
\basechance{25\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
This spell enables the Adept to select 1 entity (+ 1 / 3 full Ranks)
as an audience for their Illusions.  Whenever the Adept casts an
Illusion, they must decide whether everyone can perceive it, or just
the targets of this spell.  This spell does not circumvent any
resistance checks by the targets.  The Illusion must be cast during
the duration of this spell although the effects may last after the
Hallucination duration has expired.
\end{effects}
\end{spell}

\begin{spell}[S-8]{Heroism}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{200}
\basechance{30\%}
\resist{None}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
This spell charges the target with great might and heroism.  They seem
taller and stronger, and their actions gain a natural grace and power.
All allies (in line-of-sight) led by the target gain a bonus die
modifier of 1 (+ 1 / Rank) to all fear checks, while neutral observers
gain a bonus of 1 (+ 1 / Rank) to reaction rolls.  The target gains a
defence bonus of 1\% + 1\% / Rank.
\end{effects}
\end{spell}

\begin{spell}[S-9]{Illusion of Innocence}

\range{1 foot + 1 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{250}
\basechance{30\%}
\resist{Active}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
This Illusion changes the mien of 1 entity (+ 1 / 4 full Ranks).  This
influences the initial reactions of all entities who perceive the
target: they will react as if the target is an innocent (this does not
necessarily mean a favourable reaction).  The spell does not force
people to act stupidly --- if the target does not act appropriately,
their opinion may be revised rapidly.  Extreme behaviour terminates
these effects.  For example, should a target prepare a weapon, cast a
passively resistible spell, or attack, the spell will dissipate.
\end{effects}
\end{spell}

\begin{spell}[S-10]{Maze}

\range{15 feet + 15 / Rank}
\duration{5 seconds + 5 / Rank}
\multiple{450}
\basechance{1\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
The target is transported into a hedge-maze of perpetual fog, where
they may wander as they will.  The hex that they occupied is filled
with opaque mist that disperses in 5 seconds. At the end of the spell,
the target reappears in their original hex (they are displaced by
solids but displace fluids); any items dropped inside the maze
reappear along with the target.  Whilst in the maze, they must make a
2 × WP check each pulse to perform a non-pass action. Personal magical
effects continue, (e.g. while a Phantasm would follow the target into
the maze, Agony wouldn’t).  For all purposes, the Maze is deemed to be
on the same plane that the target just disappeared from.  Each target
appears within their own unique Maze. The Adept may still see (and
thus cast upon) the target, but not physically contact them.
\end{effects}
\end{spell}

\begin{spell}[S-11]{Illusion of Metamorphosis}

\range{1 foot + 1 / Rank}
\duration{5 minutes + 5 / Rank}
\multiple{225}
\basechance{15\%}
\resist{Active, Passive}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
The target’s appearance alters to that of a one hex creature of at
least (target’s PS + EN) pounds, over a period of 5 seconds.  Their
actual form and characteristics do not change.  They do not gain any
special abilities (e.g.  poison, magic, fear, flight) associated with
that animal. The Adept must be familiar with the desired animal. The
spell will cease when the target takes effective Endurance damage.
Creatures available by Rank: Rank 0 Small Land Animals; Rank 4
Felines, Apes \& Prehumans; Rank 8 Avians, Fish, Lizards \& Snakes;
Rank 12 Horses, Large Land Animals; Rank 16 Earth Dwellers, Fairy
Folk, Humans; Rank 20 Giants.
\end{effects}
\end{spell}

\begin{spell}[S-13]{Illusion of Multiple Images}

\range{1 foot + 1 / Rank}
\duration{1 minute + 1 / Rank}
\multiple{200}
\basechance{20\%}
\resist{None}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
One image of the target (+ 1 / 5 full Ranks) appears in the target’s
hex. Each image disappears upon receiving damage. All images are
identical to the target, and will imitate their actions faithfully.
The images have the same defence as the target, but no magic
resistance. Any targeted attack has an equal chance of hitting each
image and the target.  More than one image may be struck by a
multi-target or area effect attack.
\end{effects}
\end{spell}

\begin{spell}[S-14]{Nightmare Illusion}

\range{15 feet + 15 / Rank}
\duration{10 seconds + 10 / Rank}
\multiple{400}
\basechance{5\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
The Adept creates a terrifying, hostile and dangerous Illusion.  The
Nightmare is initially under the Adept’s control, and will obey their
commands until the Adept’s concentration is broken or the Nightmare is
slain or dissipated.  The Nightmare is totally insubstantial and
invisible to all who resist.  The Nightmare has a combined EN and FT
of 20 (+ 5 / 2 full Ranks), but no defence or armour.  It can
automatically hit up to two different targets a pulse for [D - 2] (+ 1
/ 2 full Ranks) magical damage.  It has a TMR of 10 and an Initiative
of 30 (+ 2 / Rank).  It may not be stunned.  At Rank 15, every target
must make a fear check upon first being struck.  The Nightmare may
move in any direction without restriction, including through air,
walls, water, etc.  except through the area of an Illusion Special
counterspell.  If concentration is lost, the Nightmare will attempt to
kill as many people as possible, and cannot be controlled by any
means.
\end{effects}
\end{spell}

\begin{spell}[S-12]{Illusion of Mist}

\range{20 feet + 20 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{100}
\basechance{30\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
The Adept creates an Illusion of 1000 cubic feet of mist (+ 500 /
Rank) of any shape entirely within the Adept’s range. The mist must be
least 1 foot in any dimension and may have detail no finer than one
foot in size. The visibility in this mist is 20 feet (1 / Rank). The
mist is not affected by wind.  By taking a pass action, the Adept may
move the mist up to 1 hex / pulse so long as it remains within range.
\end{effects}
\end{spell}


\section{Special Knowledge Rituals}

\begin{ritual}[R-1]{Illusionary Aura}

\range{5 feet}
\duration{1 day + 1 / Rank}
\multiple{200}
\basechance{30\% + 3\% per Rank}
\target{Entity}
\casttime{1 hour}
\material{None}
\concentration{Standard}
\begin{effects}
This ritual creates an Illusion to alter the Aura of the target.  The
target’s apparent Aura is altered by fooling the DA, DE or Divination
with a stereotype defined in Rank + 1 words, e.g. Rank 4 --- “Master
Thatcher and Evil Necromancer”.  The questioner then receives
appropriate answers based on the stereotype; if the question is not
covered by the stereotype, the real answer is returned.  The Illusory
Aura will not hide information, merely alter it. The Ritual of Warding
is not affected.
\end{effects}
\end{ritual}

\begin{ritual}[R-2]{Illusionary Terrain}

\range{5 feet + 5 / Rank}
\duration{2 hours + 2 / Rank}
\multiple{150}
\basechance{MA + 3\% per Rank}
\target{Area}
\casttime{1 hour}
\material{None}
\concentration{Standard}
\begin{effects}
This ritual creates an Illusion such that observers outside the range
cannot sense any sign of entities or their trappings within the area.
Temporary campsites will not be seen and permanent dwellings appear
deserted and dilapidated.  However, some effects may be seen, e.g. the
smoke and light of a campfire rising above the ritual’s range.  The
ritual will not hide entities from each other if both are outside the
effect, even if the area lies between them.  Magic that is not
targeted on entities or their possessions may not be hidden through
this ritual.  Any Location or Scrying magic lower than the Rank of
Illusionary Terrain may not detect into the ritual’s area.
\end{effects}
\end{ritual}

\begin{ritual}[R-3]{Permanency}

\range{5 feet}
\duration{1 Day (+1 / Rank) or Permanent}
\multiple{550}
\basechance{MA + 4\% / Rank}
\resist{None}
\target{Spell or Ritual}
\casttime{1 hour}
\material{Special}
\concentration{Standard}
\begin{effects}
This ritual enhances the duration of one of the Adept’s Illusion
spells or rituals.  The magic must have been cast previously, and have
sufficient duration to last throughout the ritual.  This ritual may
not be used with any magics with concentration-based or indefinite,
condition based duration.  Once the duration of an Illusion has been
enhanced, no changes are possible. The Illusion may be overridden by
any similar spell for the duration, rather than queuing.  This ritual
may be performed without material components to change the duration of
the target Spell or Ritual to 1 day (+1 per Rank).  Or the Adept may
use the material components to change the duration to permanent.
Material costs for this ritual are 5,000 sp (200 / Rank).  A spell
that has been enhanced by this ritual may only be removed by a Ritual
of Dissipation or the Adept’s own Special Knowledge counterspell.
\end{effects}
\end{ritual}
\end{Chapter}
