\begin{Chapter}{The College of Sorceries of the Mind (Ver 1.6)}

The College of Sorceries of the Mind deals primarily with controlling
or influencing the minds of others. Adepts of this College are
variously known as Mind Mages, or Sorcerers. It is widely believed
that Sorcerers can peer into the darkest nooks and crannies of the
soul, and are near universally feared.

Traditional Colours \& Symbols  

This  college  has  no  traditional  colours  or  symbols 
as Sorcerers prefer to blend into society. 

\section{Restrictions}

Adepts of the College of Sorceries of the Mind may practice their arts
without restriction.


16.2 Base Chance Modifiers 
The Base Chance of performing any talent, spell or 
ritual  of  the  College  of  Sorceries  of  the  Mind 
against a single entity target is affected by relative 
Willpower: 

For each point the Adept’s Willpower is 
above the target’s Willpower 
For each point the Adept’s Willpower is 
below the target’s Willpower 

+1 

-2 


\section{Talents }

\begin{talent}[T-1]{Resist Pain }

\range{Self }
\multiple{300 }
\storage{Potion }
Effects: The Adept is largely immune to pain. They 
cannot  be  tortured  or  stunned  by  physical  means. 
In addition, if there is a chance that their concentra-
tion  may  have  been  broken,  5  (+  1  /  Rank)  is  al-
ways  added  to  their  Concentration  Check.  No 
matter  what  the  circumstances  the  Adept  will  al-
ways  have  a  chance  of  maintaining  concentration. 
Like all magic, this talent is affected by cold iron. 
However,  for  a  weapon  to  affect  the  talent,  the 
weapon must be made of cold iron, must have done 
endurance  damage,  and  must  remain  in  contact 
with the body. If cold iron is used in the torture of 
a  mind  mage,  the  Adept  adds  their  concentration 
bonus  to  the  base  chance  of  resisting  the  torture 
attempt. 

\end{talent}

\begin{talent}[T-2]{Resist Temperature }

\range{Self }
\multiple{250 }
\storage{Potion }
Effects: The Adept is immune to climatic extremes 
of  temperature  from  arctic  cold  to  desert  heat 
whether  generated  naturally  or  by  magic.  They 
therefore  do  not  suffer  from  exposure,  frostbite, 
wind-chill,  heat  stress,  hypothermia  or  hyperther-
mia,  and  their  fatigue  loss  is  unaffected  by  tem-
perature.  If  the  Adept  is  attacked by  fire,  ice,  heat 
or  cold  (whether  magical  or  not)  then  the  damage 
points are reduced by 1 + 1 per 5 full ranks. 

\end{talent}

\begin{talent}[T-3]{Sense Danger }

\basechance{PC + 3 / Rank }
\multiple{300 }
\storage{Potion }
Effects:  The  Adept  may  sense  the  presence  of  a 
hazard  to  the  Adept’s  life  or  wellbeing.  Generally 
this  may  be  interpreted  as  a  “bad  feeling”  about  a 
person or an object, or a sense of foreboding about 
a  situation.  This  talent  operates  continuously  and 
unconsciously.  The  Adept  may  also  add  2  +  2  / 
Rank  to  the  chance  of  detecting  an  ambush  (see 
Ranger)  and  to  the  chance  of  detecting  a  trap  (see 
Spy/Thief). 

\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Control Animal }

\range{30 feet + 15 / Rank }
\duration{Concentration: no maximum }
\multiple{100 }
\basechance{40\% }
\resist{Passive }
\target{Non-sentient entity }
\storage{Investment }
Effects:  The  Adept  gains  control  over  one  non-
sentient  entity  within  range.  Control  is  lost  if  the 
entity leaves the range but recommences if range is 
re-entered.  The  Adept  does  not  receive  any  sensa-
tions  from  the  animal.  If  the  Adept  releases  the 
animal  or  their  concentration  is  broken  the  animal 
may  attack  the  Adept.  The  Adept  controls  the 
animal’s mind and therefore need not know how to 
make a bird fly: the Adept commands the animal to 
do  something  and  the  animal  knows  how  to  do  it. 
The  animal  will  comply  within  the  spirit  of  the 
command. 

The Magical Aptitude requirement for becoming a 
member of this College is 11. 

\end{spell}

\begin{spell}[G-2]{Control Person }

\range{30 feet + 15 / Rank }
\duration{Concentration: no maximum }
\multiple{650 }
\basechance{30\% }
\resist{Active, Passive }
\storage{Investment }
\target{Sentient Entity }
Effects:  The  Adept  gains  absolute  control  over 
every  action  of  one  sentient  entity  within  range. 
The  Adept  does  not  receive  any  sensations  from 
the  target.  The  Adept  does  not  need  to  know  the 
target’s language to control them. Control is lost if 
the  target  leaves  the  range  but  recommences  if 
range  is  reentered.  They  may  control  the  physical 
activities  of  that target, and use the target’s  physi-
cal  skills  where  known,  but  may  not  cause  the 
target to cast spells. 

The Adept may not use their own skills through the 
target.  The  target  always  acts  as  if  they  were  also 
affected by a Spell of Slowness (E\&E S-11) unless 
they choose not to resist the spell, or subsequently 
decide  to  co-operate  with  the  Adept’s  commands. 
The  spell  continues  in  effect  until  the  Adept’s 
concentration  is  broken  or  the  Adept  chooses  to 
release  the  target.  The  target  is  aware  of  the  con-
trol,  and  although  the  target  is  unable  to  do  any-
thing about it at the time, they will remember being 
controlled. The Adept may force the target to act in 
direct opposition to the target’s own wishes. Suici-
dal  instructions  however,  grant  the  target  another 
magic resistance which, if successful, dissipates the 
spell. 

\end{spell}

\begin{spell}[G-3]{Empathy }

\range{Touch until Rank 10 }
\duration{Immediate }
\multiple{200 }
\basechance{20\% }
\resist{None }
\storage{Investment }
\target{Entity }
Effects:  The  spell  allows  the  Adept  to  feel  the 
emotions  and  physical  sensations  which  the  target 
of the spell is currently experiencing. It also allows 
the  Adept  to  absorb  wounds  from  Endurance  and 
Fatigue  at  a  rate  of  2  points  cured  for  every  1 
which the Adept agrees to subtract from their own 
Fatigue  (never  Endurance).  The  additional  fatigue 
subtracted  by  the  Adept  is  damage  fatigue,  not 
spell fatigue. This spell may not be used to regen-
erate  spell  or  tiredness  fatigue,  but  only  re-
duces/transfers damage. At Rank 10 or higher, the 
Adept can cast this spell on a target who is within 
15  feet  (+  15  /  Rank  over  10)  or  less  from  them. 
This  spell  may  be  cast  on  self,  but  only  to  cure 
endurance  damage.  On  a  ‘Double’  effect,  3  Dam-
age Points are cured for each point inflicted on the 
Adept;  a  ‘Triple’  effect  allows  for  4  points  per 
point inflicted on the Adept. 

\end{spell}

\begin{spell}[G-4]{ESP }

\range{30 feet + 15 / Rank }
\duration{30 seconds + 10 / Rank }
\multiple{100 }
\basechance{40\% }
\resist{None }
\storage{Potion }
\target{Self }
Effects:  The  Adept  can  sense  the  presence  and 
general mood (but not the exact nature) of all enti-
ties  within  range  of  the  spell.  The  Adept  has  a 
general  idea  of  where  each  entity  is, say  to  within 
90  or  60  degrees,  the  accuracy  of  which  may  in-
crease  with  rank.  ESP  will  detect  the  presence  of 
an entity even if that entity is mind cloaked. Spells 
cannot be targeted through ESP. 

\end{spell}

\begin{spell}[G-5]{Hypnotism }

\range{15 feet + 15 / Rank }
\duration{Concentration: no maximum }
\multiple{200 }
\basechance{40\% }
\resist{Active, Passive }
\storage{Investment }
\target{Entity }
Effects:  The  spell  causes  an  entity  to  accept  sug-
gestions  from  the  Adept.  The  spell  may  only  be 
cast over a target with whom the caster is normally 
able to communicate verbally.  It can never be cast 
over  a  totally  hostile  target.  The  target  may  be 
enabled  to  remember  otherwise  forgotten  details 
through appropriate questioning. 

Once the target has been hypnotised, the Adept can 
make  suggestions  which  the  target  will  readily 
accept  unless  they  conflict  directly  with  their  best 
interests.  The  target  will  never  remember  where 
these suggestions came from. 

The  Adept  may  choose  up to  1  (+  1 / 5 Ranks)  of 
the  above  suggestions  to  be  active  in  a  post-
hypnotic manner, which the target will continue to 
implement  for  3  (+ 3  / Rank)  hours after  the  spell 
ceases.  The  target  will  stop  following  all  other 
suggestions once the spell ceases. 

\end{spell}

\begin{spell}[G-6]{Limited Precognition }

\range{Special }
\duration{Immediate }
\multiple{150 }
\basechance{20\% }
\resist{May not be resisted }
\storage{Potion }
\target{Self }
Effects: The Adept may see (unclearly) up to 1 (+ 1 
/  Rank)  hours into  their  own  future  and  may  fore-
see  possible  events.  The  Adept’s  player  should 
describe a plan of action, and the GM will describe 
a  vision  based  on  the  consequences  of  those  ac-
tions, which can be as literal or metaphorical as the 
GM wishes. The clarity and detail of the vision will 
be  dependent  on  the  Adept’s  rank.  Should  the 
Adept foresee their own death, a fright check may 
be appropriate. This spell cannot be recast immedi-
ately  with  a  different  plan;  the  Adept  must  wait 
until  they  have  taken  some  action  to  change  their 
situation. This spell works at any range. 

\end{spell}

\begin{spell}[G-7]{Mind Shield }

\range{Self }
\duration{1 hour + 2 / Rank }
\multiple{250 }
\basechance{30\% }
\resist{May not be resisted }
\storage{Potion }
\target{Self }
Effects:  The  Adept  cloaks  their  own  mind  so  that 
their thoughts cannot be detected or “read”, e.g. by 
Telepathy  or  other  mind  reading  talents,  spells  or 
rituals.  The  Adept’s  Resistance  versus  Mental 
Attack  is  increased  by  10  (+  2  /  Rank)  while  the 
spell  is  in  effect.  Mind  Cloak  does  not  block  ESP 
or  Empathy.  The  Adept  can  cast  Mind  Speech  as 
normal,  and  may  choose  to  decline  Mind  Speech 
from an external source. 

\end{spell}

\section{General Knowledge Rituals}

\begin{ritual}[Q-1]{Binding Will }

\range{10 feet }
\duration{1 day + 1 / Rank }
\multiple{500 }
\basechance{10\% + 5\% / Rank }
\resist{Passive }
\casttime{1 hour }
Effects:  The  Adept  may  bind  the  will  of  an  entity 
so that they become the loyal retainer of the Adept 
and serve them in all things. The target must be in 
range  during  the  ritual  but  may  be  unconscious  or 
otherwise  restrained.  The  target  gets  an  additional 
resistance check at the beginning of each following 
day and if required to do anything suicidal. If they 
successfully  resist,  the  binding  is  broken  and  they 
are  free.  The  target  is  then  aware  that  they  have 
been  under  an  external  influence.  The  Adept  may 
never release the binding voluntarily, and until the 
duration  expires the binding  continues,  even  if the 
Adept  dies  (the  target  will  do  everything  possible 
to get the Adept resurrected). If the ritual backfires, 
the Adept loses D10 from their Willpower in addi-
tion to any other backfire effect; Hypnotism (G-5) 
cures this loss. 

\end{ritual}

\section{Special Knowledge Spells}

\begin{spell}[S-1]{Disruption }

\range{15 feet + 15 / Rank }
\duration{Immediate }
\multiple{400 }
\basechance{25\% }
\resist{Passive }
\storage{Investment, Ward, Magical Trap }
\target{Object or Entity }
Effects:  The  Adept  causes  an  object  or  corporeal 
entity  (substantial  rather  than  insubstantial)  to 
pulsate, doing [D + 1] (+ 1 / Rank) points of dam-
age  if  the  target  fails  to  resist,  and half  damage  to 
any  target  that  successfully  resists.  At  Rank  10  or 
above,  the  target’s  internal  structure  and  surface 
will  be  partially  disrupted,  requiring  the  arts  of  a 
healer  or  artisan  of  rank  equal  to  (spell  Rank  /  3 , 
or  Rank  /  6  if  the  target  resisted  the  spell)  to  pre-
vent or remove the scarring (damage can be cured 
normally). 

\end{spell}

\begin{spell}[S-2]{Force Shield }

\range{Self }
\duration{10 minutes + 10 / Rank }
\multiple{250 }
\basechance{40\% }
\resist{None }
\storage{Potion }
\target{Self }
Effects:  This  spell  creates  an  invisible  protection 
around  the  entire  body  which  increases  the  De-
fence  of  the  Adept  by  5  (+  1  /  Rank)  against  all 
attacks at any range. 

\end{spell}

\begin{spell}[S-3]{Healing }

\range{Touch }
\duration{Immediate }
\multiple{375 }
\basechance{40\% }
\resist{None }
\storage{Investment, Potion }
\target{Entity }
Effects:  The  Adept  may  heal  the target  of  2  (+ 1 / 
Rank)  Damage  Points  that  have  been  removed 
from the target’s Fatigue or Endurance. Endurance 
damage  is  healed  first,  wrapping  to  Fatigue  dam-
age.  On  a  multiple  effect,  the  Damage  Points  re-
stored may be multiplied. 

\end{spell}

\begin{spell}[S-4]{Mental Attack }

\range{15 feet + 15 / Rank }
\duration{10 seconds + 10 / Rank }
\multiple{350 }
\basechance{25\% }
\resist{Active, Passive }
\storage{Investment, Ward, Magical Trap }
\target{Entity }
Effects: The Adept may cast this spell at any entity 
within  range  who  can  be  seen  or  whose  position 
has  been  pinpointed  via  Telepathy.  If  the  target 
fails to resist, they fall unconscious and at Rank 15 
and above the Target loses [D - 5] from Willpower. 
The Willpower  loss is  recoverable  via  Hypnotism, 
Remove  Minor  Curse  or  naturally  at  one  point 

every  3  days.  Repeated  casts  may  further  reduce 
the targets Willpower to a minimum of 1. 

\end{spell}

\begin{spell}[S-5]{Mind Speech }

\range{30 feet + 30 / Rank }
\duration{10 minutes + 10 / Rank }
\multiple{400 }
\basechance{20\% }
\resist{May not be resisted }
\storage{Investment }
\target{Entities }
Effects:  Allows  the  Adept  to  choose  a  Principal 
Target,  plus up  to  1  + 1  /  Rank  other  targets,  who 
may then communicate mentally to each other as if 
talking  aloud  in  a  group.  The  entities  may  be  tar-
geted  via  the  Spell  of  Telepathy  (at  the  time  of 
casting). If an entity subsequently goes further than 
30 feet + 30 / Rank from the Principal Target, their 
participation  ceases  until  they  re-enter  that  range. 
A  language  in  common  is  required  for  normal 
speech.  At  Rank  10  and  above  visual  images  may 
be  transmitted.  All  communication  requires  the 
same  level  of  concentration  as  for  normal  speech. 
The spell dissipates if the Principal Target is killed. 

Multiple  casts  are  a  special  case.  If  one  or  more 
entities  are  part  of  two  or  more  spells,  and  are 
within  range  of  all  of  their  Principal  Targets,  then 
every target of those spells may speak to each other 
as if in one group. This can effectively increase the 
working range of the effect. 

\end{spell}

\begin{spell}[S-6]{Phantasm }

\range{Special }
\duration{Concentration: no maximum }
\multiple{400 }
\basechance{1\% }
\resist{May not be resisted }
\storage{Investment }
\target{Entity }
Effects: This spell creates an insubstantial, magical 
beast  that  must  be  targeted  at  a  particular  victim 
which  is  either  visible  to  the  Adept  or  located 
through  the  spell  of  Telepathy.  The  target  may 
interpret  the  beast  as  being  a  phantasm  from  their 
own  worst  nightmare.  Lack  of  imagination,  how-
ever,  does  not  offer  any  protection  or  immunity. 
The  Adept  has  no  control  over  the  shape  of  the 
creature. 

The  phantasm  appears  beside  the  Adept,  and  then 
travels  straight  towards  the  target  at  a  TMR  equal 
to  1  +  1  /  Rank,  passing  through  walls,  air  etc. 
When  in  melee  range  of  the  target  the  Phantasm 
always  does  [D  -  4]  (+  1  /  Rank)  damage  each 
pulse,  including  the  pulse  when  it  arrives  at  the 
target. 

The phantasm is only substantial to, and visible to, 
the  target  and  thus  can  be  directly  affected  by  the 
target  (using  spells  or  magical  weapons).  Anyone 
may cast a Mind College Special Counterspell over 
the  area  currently  occupied  by  the  phantasm  to 
dissipate  it.  The  phantasm  is  immune  to  all  other 
spells and weapons. 

The  Phantasm’s  magic  resistance,  defence  and 
initiative  are  all  zero  and  it  always  acts  last in the 
pulse. The phantasm has a single combined Endur-
ance and Fatigue of 20 (+ 5 / 3 Ranks). 

The phantasm dissipates when: its endurance drops 
to  zero,  it  is  dispelled  by  counterspell,  its  target 
dies or leaves the plane or the Adept drops or loses 
concentration.  The  Adept  has  some  awareness  of 
the phantasm’s location, and is aware of its dissipa-
tion. 

\end{spell}

\begin{spell}[S-7]{Telekinesis }

\range{15 feet + 15 / Rank }
\duration{ Concentration:  maximum  10  seconds  + }
10 / Rank 
\multiple{250 }
\basechance{25\% }
\resist{None }
\storage{Investment }
\target{Entity or object }
Effects:  Allows  the  Adept  to  lift  and  manoeuvre  a 
target weighting up to 2 pounds and move it at the 

52 

rate of 2 TMR. The Adept may move an additional 
5 pounds or move the chosen target at an additional 
3  TMR  per  Rank.  The  Adept  may  increase  both 
mass and speed by applying separate ranks to each. 
Gross  movement 
is  automatic,  although  fine 
movements  require  an  MD  check  and  should  at-
tract negative modifiers for range. 

\end{spell}

\begin{spell}[S-8]{Telekinetic Rage }

\range{30 feet + 5 / Rank }
\duration{Instantaneous }
\multiple{750 }
\basechance{1\% }
\resist{Passive }
\storage{Investment, Ward, Magical Trap }
\target{Area }
Effects: The Adept creates an instantaneous, magi-
cal storm that fills the area  within range  with roil-
ing,  body-wrenching  forces.  All  objects  and  enti-
ties  that  fail  to  resist,  are  hurled  away  from  the 
Adept  until  they  are  out  of  the  area  of  effect,  or 
they strike a solid barrier. 

The  storm  inflicts  [D  -  5]  (+  1  /  foot  travelled)  of 
magical  damage  and  is  inflicted  once  only,  at  the 
time  of  casting.  The  damage  is  not  contingent  on 
striking an object. Additional damage may be done 
on hitting a wall, entity or other substantial object. 

\end{spell}

\begin{spell}[S-9]{Telepathy }

\range{30 feet + 30 / Rank }
\duration{10 minutes + 10 / Rank }
\multiple{550 }
\basechance{10\% }
\resist{Passive }
\storage{Potion }
\target{Self }
Effects:  The  spell  gives  the  Adept  the  ability  to 
read the surface thoughts of any entity within range 
who fails to resist. Each target may only resist each 
Telepathy  once,  regardless  of  whether  they  move 
in or out of the range of the spell. 

Each  pulse,  the  Adept  may  use  their  telepathy  to 
either locate the minds that are within range, or to 
read  the  thoughts  of  a  particular  mind.  The  Adept 
need  not  know  the  language  of  the  entity  whose 
thoughts  are  being  read,  but  will  have  difficulty 
comprehending  the  thoughts  of  alien  minds.  Ani-
mal  thoughts  are  usually  very  primitive  and  can 
summed up in a single word, e.g. “food”, “danger”, 
“sex.” 

It is not possible to utilise a target’s senses and the 
Adept  may  only  gain  an  impression  of  physical 
sensations  if  the  entity  is  concentrating  on  them, 
e.g. savouring a meal. 

Telepathy may be used to  target the Mind Speech, 
Mental Attack and Phantasm spells. 

Targets are not aware that their thoughts are being 
read.  Concentration  checks  should  be  required  of 
targets who try to control their thoughts by reciting 
poetry etc. 

\end{spell}

\begin{spell}[S-10]{Transmutation }

\range{15 feet + 15 / Rank }
\duration{10 minutes + 10 / Rank }
\multiple{1000 }
\basechance{1\% }
\resist{Active, Passive }
\storage{Investment, Ward, Magical Trap }
\target{Entity or Object }
Effects:  The  Adept  may  transmute  and  rearrange 
the elemental components of any target that fails to 
resist,  transforming  it  into  any  form  of  the  same 
mass  that  the  Adept  desires.  A  living  entity  can 
only  be  rearranged  into  another  living  form,  and 
objects into objects. This spell cannot confer skills 
or  magical  abilities,  but  the  target  will  acquire  the 
non-magical  features  of  the  new  form  such  as 
wings, gills etc, and may use them. If the new form 
is maintained without disruption for the duration of 
the spell, the target will revert to its original shape. 
\end{spell}

\end{Chapter}
