\begin{Chapter}{Names \& Auras}

All entities have an aura.  Those objects that were once alive retain
traces of their living auras.  The strength and composition of the
aura reflects the amount of life-force and magic that the entity or
object possesses and the other properties that are intrinsically part
of their being.

The base element of any aura is strength.  The strength of an aura is
always revealed by any magic that detects or reads auras. The
categories of strength and their relative strengths are:

\begin{dqtblr}{colspec={lX},hline{1-2}={0.8pt}}
Aura	& Equivalent\\ 
0	& No Aura \\
1	& Magic (magical wall, illusion) \\
2	& Formerly Living Composite (chair, stew) \\
3	& Formerly Living (dead orc, log) \\
4	& Non-Sentient Animates (stone golem) and Non-sentient Undead (skeleton) \\
5	& Living Plants (rose, oak) \\
6	& Living Animals (dog, cat) \\
8	& Sentient Animates (flesh golem) and Sentient Undead (vampire) \\
9	& Living Sentient (human, sphinx) \\
10	& Long Living Sentient (dragon, titan, elf) \\
15	& Avatar (material form of a Demon, etc.) \\
\end{dqtblr}

The strongest aura will be detected. Thus a human covered by an
illusion (without an aura component) will still be detectable as a
Living Sentient.

The rest of an aura consists of information intrinsic to the
possessor.  This information will vary depending upon the Generic type
of the target, but will either affect the life-force of the target, or
be magical.  Only effects that are still current or continuous in
nature will be detected. Information may be gained in descriptive
terms, values or even proportions as appropriate for the type of
information being read.

Information that may be gained from a living being includes: its
Generic True Name, its plane of origin, approximately how far it is
through its lifespan (e.g.  juvenile, 50\%, about 100 years old), its
general state of health (e.g.  healthy, diseased, 1/2 Endurance),
aptitude with a magical ability (e.g.  low overall, Rank of specific
ability) and to which College of magic (if any) it is attuned.  These
last two facts are discernible because the skills that they represent
have an effect on the level and type of magic in the entity’s aura.
Relatively little information can be divined regarding the non-magical
learned abilities of an entity.  It will be possible to learn what is
the being’s most intrinsic skill or ability, but lesser skills may not
be sufficiently intrinsic so as to have made an impression on the
being’s aura.

Magical auras will include information such as College, exact name of
a spell or other effect, level of magical ability (low, medium, high,
very high), approximate length of time that the magic has been in
effect (providing it is still present), and approximately how much
duration remains.

Any one part of an object will be representative of the entire object,
for example the aura of a toe sticking out from underneath a blanket
will reveal the same information as if the entire being were visible.
A detached thumb could reveal some information about its former
owner, up to the time it was detached, providing it is intrinsic to
the thumb, for example Generic True Name, plane of origin, or age —
when it was removed. The thumb would not reveal a magical college or a
skill, as these are properties of living beings only.

\section{Detect Aura Talent}

\begin{talent}{Detect Aura}
\range{Special}
\multiple{75}
\basechance{Perception (× 2 for Namers) + 5\% / 
Rank \mi 1\% for every foot after the first five feet the 
target is from Adept}
\resist{Active}
\target{Entity, Object, Area, Volume}
\begin{effects}
If the talent is successful the Adept learns which of the aura
categories they are seeing (with the strongest taking precedence), and
optionally learns the answer to one question of the Adept’s choice
about the target. The answer to a DA question will consist of a single
concept or "bit" of information. If the information sought is not
intrinsic to the target the Adept will receive no answer.  It is not
possible to determine the Individual True Name of an entity.  If the
Adept achieves a double or triple effect, the Adept may ask two or
three questions respectively.  The process of reading an aura and
asking a question entails concentration and requires a magical Pass
action.  Rereading a previously seen aura, or learning the category of
the aura without asking a question may be combined with other actions.

Only one attempt at Detect Aura may be made per object. An individual
object will change over time, however, and a fresh attempt may be made
when the aura has changed sufficiently to class it as a “new”
object. If an aura has been successfully read, the same information
will be available without a new Cast Check being made, until such time
that the object changes sufficiently to be considered a “new” object.

Aura require direct line of sight to be read. It is notpossible to use
Detect Aura through a mirror, crystal ball, Wizard’s Eye, or by any
other indirect means.
\end{effects}
\end{talent}

\section{Interpretation and Examples}

\begin{Enumerate}
\item In general the more specific the question the more specific the
  answer — the exact nature is left to the GM’s discretion. For
  example:
  \begin{Itemize}
  \item “What was the last type of magic to impact on this person?”
    might get “A spell”.

  \item “What is the nature of the most recent magical affect on this
    person”?" might get “fortitude”.

  \item “What was the last spell to impact on this person?” might
    get “Strength of Stone”.
  \end{Itemize}

\item All things change over time, even if outwardly they look the
  same. While the times may vary from object to object depending on
  circumstances, they tend to follow a seasonal cycle following the
  seasons.  An object whose aura was read in Spring will have changed
  sufficiently by Summer to be able to be DAed again.

\item The single attempt rule also applies to multi-hex objects.  Only
  one DA may be performed per item — one cannot DA a wooden floor
  three times just because it covers 3 hexes.  Attempts to do so will
  get the response (even before rolling the dice) --- “you see an
  aura, formerly living, and the answer is oak”, that is the same aura
  they had previously read until a change of season.  After the change
  of season a new attempt to detect the aura would have to be made.

\item Once an aura has been detected, it is available to the detector
  for the looking, as is the answer to the question/s asked, until the
  aura changes. For example, having detected the Aura of a Ward, the
  Adept is able to re-read the same information freely. If the Ward
  was then dissipated, the aura would vanish.

\item A DQ aura is located very close to the skin surface.  Thus a
  person in a full suit of plate armour, with the visor down, and no
  part of the body visible whatsoever would not be able to have their
  aura read. Cloth will hide an aura, but make up will not.

\item The size of the object will be determined by its utility.  Thus
  100’ of corridor could be a single object, while the next 5 feet,
  because it has a wooden floor, or is magical say, may be an object.
  The GM may deliberately break things up so as not to give away too
  much information from the groupings chosen.

\end{Enumerate}


\section{True Names}

\subsection{Generic True Names}

All living things have a Generic True Name.  This name is present in
their aura.  Formerly living things retain traces of their aura and
the name that they had when alive.  All True Names are in an ancient
language, believed by some to be the language that the gods used when
they made the world, and by others to be the original language of the
first mortal race, the Dragons. A translation of these names into the
common tongue yields such terms as human, elf, tiger, oak, bee,
rattlesnake, and rose.

A Generic True Name identifies the entity or object as being of a
distinct type. Very similar things with much the same form and
function, will probably have the same Generic True Name.  For example,
many small, harmless, plains-dwelling snakes will have the Generic
True Name “Grass Snake” even if they look somewhat different. A
venomous variety of similar nature will have a different Generic True
Name. Generic True Names may be learnt:

\begin{Itemize}

\item By means of a Detect Aura.  

\item From another being who has studied the name (e.g. knows the name
  at Rank 0 or higher).

\end{Itemize}

\subsection{Individual True Names}

All sentient entities (player character races, dragons, merpeople,
naga, etc.) have an Individual True Name. This becomes known to them
upon reaching maturity.  All sane sentient entities will know their
own Individual True Name and no force — physical or magical — can
coerce the entity to reveal it.  They may choose to reveal it,
however.  An entity will be called by a given (or use) name, often
given to them by parents or peers.  The entity will know their
Individual True Name in their native tongue and a Namer would have to
spend time translating the Individual Name into the Namer language
before it could be used.  Entities will protect their Individual True
Name vigorously as this knowledge can be used both defensively and
offensively.  Indeed, the very mention of an entity’s Individual True
Name would strike great fear into their heart.

Player characters and even members of the College of Naming
Incantations will know only their own Individual True Name upon
completing their education.  All other Individual True Names must be
acquired and learnt before they can be used.  Four methods exist for
acquiring an entity’s Individual True Name:

\begin{Itemize}

\item The entity may choose to reveal it.  

\item It may be obtained from a Namer Demon — when an entity is born,
  their Individual True Name becomes known to some of the Naming
  Demons (see the demon descriptions in the Bestiary).

\item It may be obtained from another other being who has studied the
  Individual True Name, if they choose to disclose it.

\item It may be found in written form — Adepts of various Colleges
  have been known to record important entities names in magic tomes.

\end{Itemize}

The aura of the entity contains both the Generic and Individual True
Names, and so the training of the College of Namers concentrates
greatly on the study and interpretation of auras, from all living
beings and formerly-living objects.  Whilst other Colleges use
abilities to detect auras, only the Namer is trained to make maximum
use of the information gained from perceiving auras.  The Generic Name
is instantly identified when a Namer perceives an aura, although this
Name must still be studied and Ranked to be of use.  The information
is coded into the aura in a form that Namers are trained to recognise,
but other Colleges, through use of the same Detect Aura spell /
talent, would need to inquire specifically to receive the same
information.

The Individual True Name is also coded into the aura, but is so
complex and varied that they cannot be deciphered and used by even a
Namer.  If the Namer is given the Entity’s True Name, then it is
possible for the Namer to identify the auric characteristics that make
up the Individual True Name.  The study of these components takes
considerable time due to their complexity.

\end{Chapter}
