\begin{College}[1.1]{necromancy}{Necromantic Conjurations}{NE}

The College of Necromantic Conjurations is concerned with death, decay
and especially with the Undead. Practitioners of the College of
Necromantic Conjurations are commonly known as Necromancers.  The
classic picture of a Necromancer is one of a black-robed figure, gaunt
and pale, leaning on a gnarled wooden or bone staff, and surrounded by
their slavering Undead minions. While it is true to say that many
Necromancers will tend towards this archetype, normally any peasant
upon seeing this apparition will arouse the local militia to lynch
said Necromancer as their very names are often uttered in the same
sentence as “Black Mage” and “Demon” with a like response being shown
to all.  As a result Necromancers will tend to neutrality in their
appearance and / or secrecy in their work, modified by their arrogance
and / or sense of style.  Others try to maintain a level of
cleanliness and hygiene not normally associated with a college whose
close ties with decay and corruption can make things very grubby and
the stench often involved is very hard to erase. Whilst it is true
that Necromancers have no direct affiliation to the Powers of Light or
Darkness, the consecration of a burial place severely limits the
ability of the Necromancer to tamper with the deceased in any way.

\subsubsection{Traditional Colours}

Necromancers usually tend towards one of the non-colours, black for
those that associate themselves more with their often night-time
activities or white for those who wish to maintain an aloofness from
their own profession which, by its very nature is dark and dirty.

\subsubsection{Traditional Symbols}

The College’s traditional symbol is the Grinning Skull.  Other common
symbols include any easily recognisable bone shape. Because of the
sometime secretive nature of Necromancers, given the fear and loathing
commonly shown them by the general populace, they will tend not to
openly tout their symbols and will quite possibly use symbols which do
not directly link them to their profession.

\section{Restrictions}

Adepts of the College of Necromantic Conjurations may practice their
arts without restriction.

The MA requirement for this College is 16. 

\subsubsection{Control Limitations}

An Adept is limited in the amount of Undead they can exert control
over at any one time. They have a maximum number of bound Lesser
Undead equal to (Willpower + Rank with the spell of Binding of Undead)
+ Military Scientist Rank.  Bound Lesser Undead include those made by
the spells of Animation of the Dead, and Animating Bodily Parts as
well as those bound by using the spell of Binding Lesser Undead.  Non
human-sized Undead will count as more than one Undead, or a fraction
of one Undead, for the purposes of this limitation, as per the spell
of Animation of the Dead.  An Adept may also never have more than one
bound Greater Undead at any one time.

\section{Base Chance Modifiers}

The Base Chance of performing any talent, spell or ritual of this
College is modified by the addition of the following numbers:

\begin{dqtblr}{colspec={Xr}}
The Adept occupies an unconsecrated burial place (e.g. pagan graveyard, barrow) & +5\%  \\
The Adept is standing on, or attempting to affect, consecrated ground & -20\% \\
During a Death Festival  & +15\% \\
During a Life Festival & -30\% \\
\end{dqtblr}

All modifiers are cumulative.  They are in addition to the modifiers
listed in §7.4.

\section{Talents}

\begin{talent}[T-1]{Ask the Dead}
\range{10 feet}
\multiple{150}
\basechance{20\% + 4\% / Rank}
\begin{effects}
The Adept may, whenever they occupy the place in which an entity has
died or has been buried, attempt to communicate with the spirit of the
entity.  Such communication is only possible if the Adept could have
communicated with the entity when it was alive (knew its language,
etc.).  The Adept must also be aware that the place is the site of an
entity’s death or burial. Questions put to the entity may only be
answered with a simple yes or no, and the dead may only provide
knowledge of events which transpired while they were alive.  Once the
dead have responded initially, they will continue to answer all
questions until dismissed, or until the Adept has asked 20 ( + 5 /
Rank) questions. Whether or not the Adept is successful, they may not
attempt to use this Talent again, on the same dead entity, until 24
hours have passed.
\end{effects}
\end{talent}

\begin{talent}[T-2]{Death Sense}
\range{1. 50 feet + 10 / Rank; 2. 10 feet}
\multiple{75}
\basechance{1. Automatic; 2. 40\% + 3\% / Rank}
\begin{effects}
This talent has two distinct uses: 
\begin{Enumerate}
\item The Adept will always be aware of the death of a Sentient Entity
  within range, provided they are awake and not engaged in an activity
  requiring active concentration.  The Adept will be aware of the
  approximate distance (within 20 feet) and approximate direction
  (within 45 degrees) of the demise.

\item The Adept may attempt to determine whether the location they
  occupy is the site of a death or burial.  If the Adept is
  unsuccessful they may not attempt to use this Talent on the same
  location until 24 hours have passed.
\end{Enumerate}
\end{effects}
\end{talent}

\begin{talent}[T-3]{Necrogeny}

\range{Self}
\multiple{150}
\basechance{Automatic}
\begin{effects}
Due to their close association with death, decay and the Undead,
Necromancers become somewhat resistant to:
\begin{Description}
\item[Fear Effects] The Adept gains a 5\% (+3 / Rank) bonus to any
  Willpower check to resist fear effects, caused by lesser Undead,
  hideous sights, etc.  This Talent does not aid in resisting magical
  fear (e.g.  Spell of Fear, Mass Fear Spell).

\item[Infection] The Adept reduces their chance of becoming infected
  by 5\% (+ 1 / Rank).

\item[Undead Draining] Damage done to the Adept due to the touch of a
Greater Undead is reduced by 1 (+ 1 per 4 or fraction Ranks).
\end{Description}
\end{effects}
\end{talent}

\section{General Knowledge Spells}

\begin{spell}[G-1]{Animating Bodily Parts}
\range{Touch}
\duration{60 minutes + 10 / Rank}
\multiple{175}
\basechance{30\%}
\resist{Passive}
\storage{Investment}
\target{Parts of Corpses}
\begin{effects}
The Adept may instill 1 humanoid bodily part (+ 1 / Rank) with a
gruesome semblance of life.  A single corpse consists of 6 bodily
parts: 2 legs, 2 arms, 1 head, 1 torso.  The Adept may join together
previously unconnected parts before animation.  The animated bodily
parts will be able to follow simple commands given by the Adept.  The
exact mode of locomotion of the parts, and their speed and other
abilities is left to the discretion of the GM.  They will be slower
and weaker than normal Skeletons or Zombies. The exact results of this
spell are hard to predict, and even two identically constructed
“animates” may not perform in precisely the same manner.
\end{effects}
\end{spell}

\begin{spell}[G-2]{Binding Lesser Undead}
\range{15 feet + 15 / Rank}
\duration{1 hour + 2 / Rank}
\multiple{200}
\basechance{35\%}
\resist{Passive}
\storage{Investment}
\target{Lesser Undead Entity}
\begin{effects}
The Adept may gain control of 1 (+ 1 per 2 or fraction Ranks) Lesser
Undead, that fail to resist.  If the Undead to be affected are already
controlled or bound, the rank of the Adept with this spell is compared
to the rank of the original control or binding.  If the original rank
is equal or higher, the Undead are unaffected by this spell, otherwise
the Undead must resist normally. The Undead will serve the Adept in
all ways, obeying simple commands communicated verbally by the Adept.
No target may resist an order once it has failed to initially resist
the spell.
\end{effects}
\end{spell}

\begin{spell}[G-3]{Conjuring Darkness}
\range{15 feet + 15 / Rank}
\duration{ 15  minutes × [D - 5] × Rank (× 1 if unranked)}
\multiple{100}
\basechance{50\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
The Adept creates a volume in which non-magical light is partially
suppressed.  The volume will be 1000 (+ 500 / Rank) cubic feet, and
may be in any one contiguous area the Adept desires, provided that no
dimension is smaller than one foot.  The entire volume must be visible
and within range at time of casting, and may not be moved.  For
visibility purposes, the Spell will increase Dark- ness levels within
the volume to 60\% + 2\% / Rank.  Rank 20 Darkness may not be seen
through. It will not aid in providing bonuses for casting purposes,
though it will neutralise penalties due to natural light, to a maximum
of 5\% + 1\% / Rank. The volume counts as direct shadow for Star \&
Shadow Mages.  If the lighting conditions are lower than that provided
by the spell, no effect will be apparent.  Note that because light is
only being suppressed, it may still pass through, and no shadows are
generated outside the volume. If it is possible to see through a
Darkness, everything beyond it is normally visible.  This spell an
engender silhouettes of lit objects against the darkness, though not
create shadows.  Any of this volume may be overridden by a higher
ranked Spell of Light, or neutralised (back to original conditions) by
an equal rank.
\end{effects}
\end{spell}

\begin{spell}[G-4]{Detecting Undead}
\range{50 feet + 50 / Rank}
\duration{Immediate}
\multiple{150}
\basechance{30\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept becomes immediately aware of the class (Lesser or Greater),
approximate number (to the nearest 5), and general location (to the
nearest 20 feet) of all Undead within range.  If the Undead are within
10 feet (+ 10 / Rank) the Adept is aware of their exact types,
numbers, and locations.
\end{effects}
\end{spell}

\begin{spell}[G-5]{Fear}
\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{250}
\basechance{25\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
The Adept instills in the target an uncontrollable fear. Unless the
target successfully resists they must roll on the Fright Table.  If a
double effect is achieved, the Adept may modify the Fright Table roll
up or down by an amount equal to the Rank of the spell. If a triple
effect is achieved, the Adept may modify the Fright Table roll up or
down by twice the Rank of the spell. See the Fright Table for the
exact results of the fear.
\end{effects}
\end{spell}

\begin{spell}[G-6]{Noxious Vapours}
\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{200}
\basechance{25\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
The entire area affected by the spell exudes a charnel stench, and all
entities within it, except the Adept, must make a Will Power check or
become nauseous.  The Difficulty Multiplier for the Check is dependent
on the Rank of the spell:

\begin{dqtblr}{colspec={m{3em}X}}
Rank	& Multiplier \\
0--7	& 4 \\
8--13	& 3 \\
14--18	& 2 \\
19--20	& 1 \\
\end{dqtblr}

Those entities who become nauseous have their Strike Chances reduced
by 10 (+ 1 / Rank), and must make a Willpower concentration check to
utilise Spell magic. The multiplier for this check is the same as that
for resisting the nausea.  In addition, this spell causes a thick,
roiling white mist to rise from the ground. The mist is 6 inches high
(+ 6 inches / Rank), and reduces the range of all forms of vision,
within the mist, to 20 feet (1 foot / Rank).
\end{effects}
\end{spell}

\begin{spell}[G-7]{Putrescence}
\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{100}
\basechance{50\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Volume of Food and / or Drink}
\begin{effects}
This spell causes up to 1 cubic foot (+ 1 / Rank) of food and/or drink
to putrefy, spoil, decay and rot.  The foodstuffs will thereafter be
unfit for consumption.  If a Double or Triple effect is achieved, the
amount of food and / or drink to be affected may be doubled or
tripled.
\end{effects}
\end{spell}

\begin{spell}[G-8]{Putrid Wound}
\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{250}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Living Entity}
\begin{effects}
The Adept may cause [D - 4] (+ 1 / Rank) damage in the form of a
putrid wound, to one entity within range, unless the target
successfully resists.  The wound is automatically infected.  This
spell will only affect living entities (i.e. it will not affect
Undead, Animates, Demons, etc.)
\end{effects}
\end{spell}

\begin{spell}[G-9]{Rigor Mortis}
\range{15 feet + 15 / Rank}
\duration{10 seconds + 10 / Rank}
\multiple{200}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
If the target entity fails to resist they are affected with a
corpse-like stiffening. Any attempt to move will result in pain and
accompanying cracking noises.  The victim suffers a penalty to their
Cast Chances and Strike Chances equal to 5 (+ 1 / Rank).  In addition,
their Agility is reduced by 1 (+ 1 per 3 or fraction Ranks).  This
Agility reduction will affect initiative and TMR.
\end{effects}
\end{spell}

\begin{spell}[G-10]{Spectral Hand}
\range{15 feet + 5 / Rank}
\duration{30 seconds + 10 / Rank}
\multiple{200}
\basechance{25\%}
\resist{None}
\storage{Investment}
\target{Special}
\begin{effects}
The Adept conjures a giant, invisible, spectral hand, which executes a
two word command e.g.  “Smash that, Lift that, etc.”  The hand may
move at TMR 4, and may not leave the Adept’s range.  The hand may
exert force in one direction equal to a Physical Strength of 15 (+ 2 /
Rank). Entities or Objects caught between the hand and an immovable
object suffer 1 (+ 1 per two full Ranks) physical damage per Pulse
(armour may reduce this).  Entities may break away from the hand by
executing a Withdraw action.  If seen by Witchsight, the hand appears
coldly blue and skeletal and is roughly 3 feet in diameter.
\end{effects}
\end{spell}

\begin{spell}[G-11]{Warping}
\range{Touch}
\duration{Immediate}
\multiple{175}
\basechance{30\%}
\resist{Passive}
\storage{Investment}
\target{Volume of Object}
\begin{effects}
The Adept may twist and warp up to 1 cubic foot (+ 1 / Rank) of any
formerly living matter, into the shape of their choosing.  The Volume
may contain varied items.

\begin{example} 
An Adept could warp a collection of flowers, bones, and sticks
together so as to form a funeral wreath.
\end{example}

The spell confers no special artisan abilities on the Adept.  Once
warped the object(s) will remain in their new shape unless remoulded.
\end{effects}
\end{spell}

\section{General Knowledge Rituals}

\begin{ritual}[Q-1]{Converse with the Dead}

\duration{Special}
\multiple{200}
\basechance{MA + 4\% / Rank}
\resist{None}
\target{Spirit}
\casttime{1 hour}
\material{Drawn Pentacle}
\actions{Concentration}
\concentration{Standard}
\begin{effects}
This Ritual summons a friendly Spirit, who will answer 3 questions (+
1 per two full Ranks) for the Adept. The Adept must first draw a
Pentacle, and remain within it throughout the Ritual.  Each question
will be answered with a riddle or puzzle.  The GM creates all riddles
and puzzles in answer to the Adept’s questions.  As the Adept’s Rank
with this ritual increases, the replies will become less cryptic and
confusing.  Undead may not be summoned or conversed with using this
ritual.
\end{effects}
\end{ritual}

\begin{ritual}[Q-2]{Summoning Lesser Undead}

\duration{Immediate}
\multiple{250}
\basechance{MA + 3\% / Rank}
\resist{None}
\target{Lesser Undead}
\casttime{2 hours}
\material{Drawn Pentacle}
\actions{Concentration} 
\concentration{Standard}
\begin{effects}
To use this Ritual the Adept must first draw a Pentacle in or near an
area where Lesser Undead may be found (near tombs, graveyards,
barrows, etc.).  The Adept must remain within the Pentacle during the
entire course of the ritual. The ritual summons 1 (+ 1 / Rank) Lesser
Undead.  Undead summoned successfully (i.e. not as a result of a
backfire) will appear within 20 feet of the Adept’s Pentacle, and will
be unable to cross into the pentacle.  The Undead are not bound or
controlled in any way. If the ritual backfires the Lesser Undead will
appear inside the pentacle and will attack the Adept.
\end{effects}
\end{ritual}

\section{Special Knowledge Spells}

\begin{spell}[S-1]{Agony}
\range{30 feet + 15 / Rank}
\duration{10 seconds + 10 / Rank}
\multiple{350}
\basechance{10\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
This spell causes all entities in the affected area, except the Adept,
to suffer extreme agony.  Entities who fail to resist may only take
Pass actions for the duration of the spell, or until such time as they
leave the area of effect.  Entities who successfully resist reduce all
Strike Chances by 30, and take twice as long to perform any action.
Note that Mind Mages gain a bonus to resist this spell equal to 2 ×
Rank with their Talent of Resisting Pain.
\end{effects}
\end{spell}

\begin{spell}[S-2]{Animation of the Dead}
\range{30 feet + 15 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{250}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward, Magical Trap}
\target{Corpses}
\begin{effects}
The Adept may fill 3 (+ 1 per Rank) human-sized corpses within range,
with the power of undeath, giving them a gruesome semblance of
life. The Undead so created will serve the Adept in all ways. The
Undead will obey simple commands communicated verbally by the
Adept. Corpses that possess most of their flesh will become Zombies,
those that are mostly devoid of flesh will become Skeletons.  If this
spell is used on corpses of other than human size, the following
strictures apply: Larger than human sized corpses count as 1 corpse
per hex. Smaller than human sized corpses count as fractions of a
corpse: Dog, 0.5, Cat, 0.2, Rat, 0.1.  No more than 10 small corpses,
even if smaller than rat sized, may be animated in the place of 1
human sized corpse.  All corpses animated by a single casting of this
spell must be of the same type. Note that if this spell is cast as the
result of a Ward or Magical Trap the reanimated dead will be
uncontrolled and will attack any living beings they can reach.
\end{effects}
\end{spell}

\begin{spell}[S-3]{Binding Greater Undead}
\range{15 feet + 15 / Rank}
\duration{30 minutes + 30 / Rank}
\multiple{300}
\basechance{20\%}
\resist{Active, Passive}
\storage{None}
\target{Greater Undead Entity}
\begin{effects}
The Adept may order one Greater Undead, that fails to resist, to do
anything that is within its physical capabilities, and is not
obviously suicidal.  No target may resist a valid order once it has
failed to initially resist the spell.  If the Undead to be affected is
already controlled or bound, the rank of the Adept with this spell is
compared to the rank of the original control or binding. If the
original rank is equal or higher, the Undead is unaffected by this
spell, otherwise the Undead must resist normally.  This spell does not
grant the Adept any particular ability to communicate with the target.
\end{effects}
\end{spell}

\begin{spell}[S-4]{Bone Construction}

\range{5 feet + 5 / Rank}
\duration{15 minutes + 15 / Rank}
\multiple{250}
\basechance{15\%}
\resist{None}
\storage{Investment}
\target{Area}
\begin{effects}
The Adept may create 25 cubic feet of interlocking bones (+ 25 / Rank)
in any shape or shapes of the Adept’s choosing.  Any dimension that is
less than 1 foot is considered to be 1 foot for the purposes of
computing volume.  The bones always appear entirely within range of
the Adept and may not appear on top of, or inside (partially or
wholly), any entity. The bones become increasingly strong with higher
Rank:

\begin{dqtblr}{colspec={m{3em}X}}
Rank	& Strength \\
0--5	& Bone \\
6--10	& Wood \\
11--15	& Bronze \\
16--20	& Iron \\
\end{dqtblr}
\end{effects}
\end{spell}

\begin{spell}[S-5]{Dark Vision}

\range{15 feet + 15 / Rank}
\duration{1 hour + 1 / Rank}
\multiple{100}
\basechance{60\%}
\resist{None}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
The Adept causes the target to develop excellent vision in the dark.
Everything will appear monochromatic (ie shades of grey) and it is
difficult to accurately estimate distance. The higher the Rank, the
less of a problem this will be.  Some amount of light must be present
for this vision to operate.  The range of the vision is 50 feet (+ 10
/ Rank).
\end{effects}
\end{spell}

\begin{spell}[S-6]{Hand of Death}

\range{15 feet + 15 / Rank}
\duration{5 seconds + 5 / Rank}
\multiple{250}
\basechance{20\%}
\resist{Active, Passive}
\storage{Investment}
\target{Entity}
\begin{effects}
The target suffers [D + 1] damage each pulse that the Adept takes a
Pass action and makes visible squeezing motions with their hand to
simulate the squeezing of the victim’s heart. Only at the time of
casting, and on subsequent pulses when the Adept takes a pass action,
must the target be in the Adept’s line of sight.  On Pulses that their
heart is squeezed the target suffers extreme pain, and may only
perform Pass actions.
\end{effects}
\end{spell}

\begin{spell}[S-7]{Life Draining}

\range{Touch}
\duration{Special}
\multiple{300}
\basechance{15\%}
\resist{Special}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept’s hand becomes charged for 5 seconds (+ 5 / 4 or fraction
Ranks) and will drain 1 (+ 1 / Rank) Fatigue from the next entity to
be touched, if the target fails to resist.  This thereby discharges
the spell.  The Fatigue may be used to repair the Adept’s own Fatigue
and/or Endurance.  If the target has no Fatigue remaining the Adept
may drain from Endurance instead. The Adept may not drain from both
Fatigue and Endurance with one cast of the spell. Note that only
living entities will be affected by this spell and the Adept may not
drain themselves.
\end{effects}
\end{spell}

\begin{spell}[S-8]{Mass Fear}

\range{30 feet + 15 / Rank}
\duration{30 seconds + 10 / Rank}
\multiple{400}
\basechance{10\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
The spell instills in all entities within range, other than the Adept
and those who successfully resist, an unreasoning and uncontrollable
fear.  All entities that fail to resist must roll on the Fright Table
(see 54.1).
\end{effects}
\end{spell}

\begin{spell}[S-9]{Necrosis}

\range{15 feet + 15 / Rank}
\duration{Immediate}
\multiple{450}
\basechance{5\%}
\resist{Active, Passive}
\storage{Investment, Ward, Magical Trap}
\target{Entity}
\begin{effects}
This spell causes 1 target for every 3 or fraction Ranks to suffer [D
  + 1] (+ 2 / Rank) damage in the form of internal haemorrhaging and
rotting.  If a target resists, they suffer only half damage (round
up). Wounds inflicted by this spell will automatically be infected.

Note that only living entities will be affected by this spell.
\end{effects}
\end{spell}

\begin{spell}[S-10]{Petit Mort}

\range{Touch}
\duration{Variable}
\multiple{250}
\basechance{30\%}
\resist{Active, Passive}
\storage{Investment, Potion}
\target{Entity}
\begin{effects}
By means of this spell, the Adept suspends all of the target’s bodily
functions, causing them to take on the semblance of death.  The target
will have no discernible signs of life, even so far as to appear dead
to a Healer using empathy.  The target’s body will slowly cool to room
temperature.  At Rank 10 or below, the target’s aura will still show
them to be living, but at Rank 11 or greater, they will detect as
“Formerly Living”.  A Healer will only be able to detect that the
target is not truly dead if they attempt to either Preserve Dead or
Resurrect.  The target’s body will require no sustenance of any sort,
nor will it decay or age. The duration of the spell must be decided by
the Adept at the time of casting, up to a maximum of:

\begin{dqtblr}{colspec={m{3em}X}}
Rank	& Duration \\
0--3	& 1 day \\
4--8	& 1 week \\
9--11	& 1 month \\
12--16	& 6 months \\
17--19	& 1 year \\
20	& Any duration \\
\end{dqtblr}

Note that if the spell is made into a Potion, the target of the spell
is the imbiber, and they may only passively resist the spell’s
workings.
\end{effects}
\end{spell}

\begin{spell}[S-11]{Scarring Terrain}

\range{50 feet + 25 / Rank}
\duration{1 month + 1 / Rank}
\multiple{225}
\basechance{20\%}
\resist{None}
\storage{Investment, Ward}
\target{Area}
\begin{effects}
The Adept causes terrible ruin to all ground within range.  The ground
will be so damaged that it will be unable to support any flora for the
duration of the spell (grass will turn to dust, small shrubs will
shrivel, trees will lose all foliage and slowly die).  Though the
spell will wither any flora in the area, fauna will be unaffected.
\end{effects}
\end{spell}

\begin{spell}[S-12]{Spectral Warrior}

\range{Sight}
\duration{Concentration: No maximum}
\multiple{400}
\basechance{5\%}
\resist{None}
\storage{Investment}
\target{Entity}
\begin{effects}
The Adept conjures to this plane a spectral warrior, and directs the
warrior to hunt down and slay one target, who must be within sight
when the spell is cast.  The warrior is completely insubstantial and
invisible except to its intended victim.  It may be seen by others
with Witchsight.  It appears as a glowing spectre in baroque armour,
with piercing red eyes.  The warrior will continue with its mission
until the Adept’s concentration is broken, its intended victim dies,
or it is dissipated or slain.  The warrior can unerringly locate its
intended victim and will always move towards them at its full movement
rate, and engage them in melee combat.

The warrior has a single, combined, Endurance and Fatigue
Characteristic with a value of 20 (+ 5 per 3 full Ranks).  The
spectral warrior automatically hits every Pulse for [D - 4] (+ 1 /
Rank) damage. Its Initiative is 30 (+ 2 / Rank). The warrior has no
defence or armour value. It may not be Stunned.  The warrior’s
movement rate is 650 yards per minute, and its TMR is 13.  The warrior
may move in any direction without restriction, including through the
air, walls, water, etc., except through the area of a Necromantic
Special Coun- terspell. The warrior may be dissipated by having a
Necromantic Special Counterspell cast on the area it occupies.
\end{effects}
\end{spell}

\begin{spell}[S-13]{Spectral Weapon}

\range{5 feet + 5 / Rank}
\duration{5 minutes + 1 / Rank}
\multiple{250}
\basechance{15\%}
\resist{None}
\storage{Investment}
\target{Object}
\begin{effects}
The Adept may increase the usefulness of any weapon within range. The
weapon begins to faintly glow with a cold, blue light.  The weapon has
its Base Chance increased by 1 ( + 1 / Rank), and the damage increased
by 1 per 3 or fraction Ranks.  In addition, the weapon may affect
targets that are insubstantial, such as Spectres, Spectral Warriors,
etc.
\end{effects}
\end{spell}

\begin{spell}[S-14]{Stream of Corruption}

\range{30 feet + 5 / Rank}
\duration{Immediate}
\multiple{350}
\basechance{5\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
From the Adept’s fingertips erupts a column of rotting blood,
lacerating bone shards, maggots, and corrosive pus, which travels to
the extent of the spell’s range, and is 5 feet wide. The Adept may
increase the width by 1 foot per Rank.  All entities occupying hexes
through which the stream passes must resist or suffer [D - 2] (+ 1 /
Rank) damage.  Living entities who are damaged by this spell have
their Base Chance of Infection increased by 20 (+ 4 / Rank).  The
putrid matter remains for about a minute after the spell is cast.
\end{effects}
\end{spell}

\begin{spell}[S-15]{Wall of Bones}

\range{15 feet + 15 / Rank}
\duration{10 minutes + 10 / Rank}
\multiple{150}
\basechance{30\%}
\resist{Passive}
\storage{Investment, Ward, Magical Trap}
\target{Area}
\begin{effects}
The Adept conjures a wall of interlaced bones, 1 foot thick, 10 feet
high, and either 20 feet long, or in a circle with a diameter of 10
feet. The Adept may increase height, length, or diameter, by 1 foot
per Rank.  Every time an entity comes in contact with the wall, they
must passively resist or suffer a roll on the Fright Table.  The wall
can withstand 100 points of damage before crumbling to dust.  “A”
Class weapons cannot damage the wall.
\end{effects}
\end{spell}

\begin{spell}[S-16]{Wraithcloak}
\range{Self}
\duration{30 minutes + 30 / Rank}
\multiple{200}
\basechance{15\%}
\resist{None}
\storage{Potion}
\target{Self}
\begin{effects}
The Adept becomes shadowy and less corporeal, becoming increasingly
ethereal with Rank. 2\% (+ 3 per two full Ranks) is added to their
defence, and if they are struck by weapon that is not magical or
silvered, 1 point per three full Ranks is subtracted from the damage.
In addition, the Adept gains a 1\% (+ 1 / Rank) bonus to their
Stealth.
\end{effects}
\end{spell}

\section{Special Knowledge Rituals}

\begin{ritual}[R-1]{Becoming Undead}

\duration{Permanent}
\multiple{250}
\basechance{10\% + 3\% / Rank}
\resist{Active, Passive}
\target{Entity}
\casttime{2 hours}
\material{Drawn Pentacle, Sacrifice}
\actions{Ritual Sacrifice} 
\concentration{None}
\begin{effects}
The Adept may transform a sentient entity into an Undead by employing
this ritual. Once the entity joins the ranks of the Undead the effects
may not be reversed.  The Adept must draw a Pentacle and remain within
it throughout the ritual, along with the entity to be transformed (who
may be the Adept themselves) and an entity to be sacrificed.  During
the course of the ritual, the Adept must sacrifice an entity of the
same race as the entity to be transformed.  The sacrifice may be
bound, but must be conscious during the ritual.  Upon the performance
of sacrifice, a Ghost appears over the corpse. This Ghost becomes
chained to the place in which the ritual was performed until freed by
the death of the Adept.  Once the sacrifice is performed, the Adept’s
Player checks to determine if the ritual was a success. If it is a
success, the entity to be transformed joins the ranks of the Undead as
the Greater Undead type of the Adept’s choice.  If the ritual
backfires, the Ghost is transformed into a Wraith, who may immediately
attack the Adept.  Greater Undead created by this ritual retain all of
their prior knowledge and magical abilities.  They will also retain
those physical abilities as may be used by their new form.
\end{effects}
\end{ritual}

\begin{ritual}[R-2]{Life Prolonging}

\duration{Special}
\multiple{350}
\basechance{5\% + 5\% / Rank}
\resist{None}
\target{Entity}
\casttime{8 hours}
\material{None}
\actions{Concentration}
\concentration{Standard}
\begin{effects}
The Adept may prolong an entity’s life, including their own, causing
them to remain unchanged and unaging for the duration of this ritual.
Once the effects of the ritual wear off, the target will begin to age
at the rate of 1 year per minute until they reach the proper
biological age to match their chronological age.  Normally, the
ritual’s effects last for 2 years (+ 2 / Rank), but at Rank 10 and
above, the Adept can immediately upon the ritual being completed
reduce the target’s biological age by up to 1 year (+1 / Rank) in
addition to halting aging.  Any backfire causes accelerated aging
(this effect may be passively resisted).
\end{effects}
\end{ritual}

\begin{ritual}[R-3]{Permanency}

\duration{Permanent}
\multiple{350}
\basechance{MA + 4\% / Rank}
\resist{None}
\target{Undead, Bones}
\casttime{1 hour}
\material{Precious Ingredients}
\actions{Anointing Undead or Bones}
\concentration{None}
\begin{effects}
The Adept may use this ritual to increase the duration of the
animation of any Lesser Undead, or any Bone Construction, created by
them to “Permanent”.  This can be applied to the spells of Animating
Bodily Parts, Animate Dead, and Bone Construction. The ritual may only
be performed by the same Adept who made the animate or bones to be
affected. The Adept must first cast the spell that they wish to make
permanent and then immediately begin this ritual.  A spell that has
been enhanced by the use of this ritual may not be Counterspelled,
but may still be Dissipated.  The Adept must expend precious
ingredients to perform this ritual.  The ingredients will cost 300sp
(5 / Rank), per cubic foot of bones or Undead to be affected.  Any
dimension that is less than 1 foot is considered to be 1 foot for the
purposes of computing volume.  A standard human-sized Undead will be
about 12 cubic feet.
\end{effects}
\end{ritual}

\begin{ritual}[R-4]{Shaping Flesh Golems}

\duration{Permanent}
\multiple{500}
\basechance{Special}
\resist{None}
\target{Parts of bodies}
\casttime{24 weeks}
\material{Parts of corpses \& (20,000 - 900 / Rank) Silver Pennies}
\actions{Constructing Golem}
\concentration{None}
\begin{effects}
This ritual is used to fashion a Golem from parts of different
corpses. All pieces that are to go into the Golem must be found before
the ritual can begin.  In addition to the pieces of corpses and
precious ingredients, the services of a Taxidermist or Undertaker of
at least Rank 6 are needed for this ritual. Once the ritual is
complete the Flesh Golem will become animate and will remain active
until it is killed. A Flesh Golem will always obey its creator and is
barely sentient, but the GM must determine the complexity of commands
it is capable of executing based on the intelligence of the brain used
to fashion the Golem. This ritual is automatic, but there is a 1\%
chance of an accident each week during a Flesh Golem’s creation.  It
must be noted that the creation of a Flesh Golem is an inexact
science, and no two Flesh Golems seem to turn out the same.  Even
building a second Golem using identical materials rarely results in
the same final creature.  To perform this ritual the Adept will
require access to an Alchemist’s laboratory or a Binder’s
workshop. Flesh Golems have the following characteristics:

\begin{Description}
\item[Description] A Flesh Golem can be put together from different
  creatures within a species, or from any number of different
  species. For example, a troll could be fashioned with the brains of
  a human, and the wings of a gargoyle, or an ogre could be made with
  the strength of one ogre (presumably a powerful one), the endurance
  of another hardy one, and the brain of a bright ogre. Seams will
  show wherever parts are sewn together, and at low Ranks the Golem
  will exude a horrible, rotting odour. By Rank 6, the odour will only
  be noticeable in very close Proximity to the Golem, and above Rank
  10the odour will only be noticed by canines and other creatures with
  an exceptional sense of smell.  Talents, Skills and Magic Flesh
  Golems never retain any of the skills or spell casting abilities
  (although they do retain some weapon skills) of any of the creatures
  from which they are made, but do retain any talents appropriate to
  the parts of a creature used. Thus, a Golem made with the eyes of a
  basilisk would be able to turn creatures to stone in the same manner
  as a basilisk.

\item[Movement Rate] (yards per minute): Variable, according to the
  Golem’s size and method(s) of locomotion.

\item[PS] Average the Physical Strength of the creatures 
from which the Golem’s muscle tissues were taken. 

\item[MD] Average the Manual Dexterities of the creatures from which
  the Golem’s muscles and brain were taken.
  
\item[AG] Average the Agilities of the creatures from 
which the Golem’s muscles and brain were taken. 

\item[MA] 0 

\item[EN] Average the Endurances of the creatures from which the
  Golem’s muscles and internal organs were taken.

\item[FT] As for Endurance, but the creature from which the Golem’s
  lungs were taken counts twice within the average.

\item[WP] Add 5 to the Willpower of the creature from which the
  Golem’s brain was taken. Maximum 30.

\item[PC] Variable, depending on the quality of the eyes, ears, and
  mental faculties of the creatures that were used, but never more
  than 10 (+ Rank / 2, rounded up).

\item[PB] Variable, depending on the creatures used, but never more
  than 6 (+ Rank).

\item[TMR] Variable depending on Agility and GM’s discretion.

\item[NA] Variable, depending on the creature from which the skin was
  taken.

\item[Weapons] A Flesh Golem can use any natural weapons (such as
  claws and fangs) built into it, as well as any weapons it can
  carry. It will have the same Ranks in weapons as the creature who
  contributed its brain divided by 2 and rounded down.  This is
  knowledge only, and there is no guarantee that the Golem will have
  the requisite PS or MD, or will even be able to manipulate weapons.
  Comments The GM will have to use discretion in allowing combinations
  of creatures.  It would be absurd to allow a Golem to be built with
  a dragon’s head on a human body – due to discrepancies in size, for
  instance. Flesh Golems can be harmed by weapons or magic as per
  normal.
\end{Description}

\end{effects}
\end{ritual}

\begin{ritual}[R-5]{Summoning Greater Undead}

\duration{Immediate}
\multiple{350}
\basechance{MA / 2 + 3\% / Rank}
\resist{Passive}
\target{Greater Undead}
\casttime{2 hours}
\material{Drawn Pentacle}
\actions{Concentration}
\concentration{Standard}
\begin{effects}
This ritual works in the same manner and under the same conditions as
the Ritual of Summoning Lesser Undead (Q-1) except that the Ritual
summons only one Greater Undead.
\end{effects}
\end{ritual}

\end{College}
