\begin{Chapter}{Non-College Magic}

11.1 Introduction 

All spells and rituals listed in this section can be learnt by Adepts
of any college.

11.2 Special Knowledge Spells 

consent, though it can be passively resisted. Addi-
tionally,  one  with  the  power  of  full  geas  may 
automatically  remove  (without  the  support  of  a 
triangle  and  12  hours  of  ritual)  a  geas  which  is  at 
least 5 Ranks less than their Rank with the spell. 

Geas 

Major Curse 

Range:  The  Adept  must  be  able  to  see  and  com-
municate with the target 
Duration: Until removed, fulfilled or target dies 
Experience Multiple: 250 
Base Chance: Always successful (see below) 
Resist: Special 
Target: Sentient entity 
Effects:  A  geas  is  an  obligation  to  complete  a 
quest,  an  injunction  against  the  performance  of  a 
particular action, or a requirement to respond in the 
same fashion to particular stimuli. 

The  target  must  acknowledge  their  acceptance  of 
the  geas.  Furthermore,  either  the  Adept  must  be-
lieve that the target deserves the geas, or the target 
must truly wish (not forced by physical or magical 
means)  to  have  an  unmerited  geas  placed  upon 
them. The Adept specifies the nature of the geas in 
25  words  or  less,  and  the  GM  will  use  the  most 
liberal interpretation of that wording to the benefit 
of  the  target.  Rank  with  the  geas  spell  does  not 
affect the chance of casting the spell as it is always 
automatically  successful.  The  Rank  equals  the 
effectiveness  of  the  geas,  expressed  in  percentage 
terms.  If  a  geased  entity  directly  contravenes  the 
letter of a geas, they have a chance of dying equal 
to the Rank of the geas. A geased entity will begin 
to  feel  weak  or  ill  when  they  first  take  an  action 
counter  to  the  restriction  of  the  geas,  and  will  be-
come  increasingly  afflicted  until  they  once  more 
comply with the geas. 

If a quest geas is fulfilled by the geased entity, they 
are  no  longer  subject  to  that  geas.  The  other  two 
types  of  geas  (for  and  against  a  given  action)  last 
indefinitely.  A  geas  can be  removed  automatically 
by the one  who placed it. A geased entity will not 
attempt  to  free  themselves  from  the  compulsion. 
An  Adept  may  attempt  to  remove  a  geas  if  they 
have  a  higher  rank  than  the  geas  in  effect.  The 
Adept  must  inscribe  a  triangle  about  the  geased 
entity,  and  perform  the  ritual  of  geas  removal  for 
12  consecutive  hours.  If  the  triangle  is  silver  or 
truesilver,  the  geased  entity  does  not  suffer  the 
penalties for ignoring the geas during the ritual. 

The  Adept  attempting  to  remove  the  geas  has  a 
success  chance  equal  to  five  times  the  difference 
between their Rank with geas and the Rank of the 
geas being removed. The GM rolls percentile dice: 
if  the  roll  is  less  than  or  equal  to  the  success  per-
centage,  the  geas  is  removed.  If  the  roll  is  greater 
than the success percentage, the Rank of the geas is 
increased by one. 

Range: 20 feet + 15 / Rank 
Duration: Until removed or target dies 
Experience Multiple: 750 
Base Chance: 15% 
Resist: Passive (unless a Death-curse) 
Target: Entity or Object 
Effects:  An  Adept’s  Endurance  value  is  decreased 
by one whenever they inflict a major curse upon a 
being.  Note  that  when  casting  a  Death-curse  this 
Endurance point loss is in addition to any possible 
Endurance point loss due to resurrection. There are 
several types of major curses: 

Affliction The Adept may choose to torment or kill 
their target. If the effects of the affliction curse are 
intended  to  be  deadly,  the  target  may  not  die  as  a 
direct  result  of  the  curse  before  (24  -  Rank)  hours 
have passed. The following list of sample affliction 
curses is provided to give the GM a guideline as to 
what major curses should be allowed in their cam-
paign. 

1. Target becomes totally blind, deaf or mute. 

2. Target becomes senile. 

3.  Target  suffers  from  a  contagious  disease  (for 
example open running sores). 

4.  Target  is  transformed  into a  frog  or  other  small 
creature. 

5.  Target  becomes  weakened  and  enfeebled  and 
must be helped with any physical action. 

6. Target falls into century-long sleep. 

Ill  Luck  Add  two  times  the  Rank  of  the  major 
curse  spell  to  any  percentile  roll  involving  the 
target or the use any of their abilities. This may not 
be applied favourably. 

Doom A doom is a pronouncement, by the Adept, 
upon an event that will occur in the target’s  future 
(e.g.  “You  will  die  by  the  hand  of  a  loved  one.”). 
The statement which must be indefinite will be true 
unless  removed.  The  GM  should  be  careful  as  to 
what to allow for dooms. 

Death-curse  At  the  moment  of  their  death,  an 
Adept may automatically cast a major curse (unless 
backfire occurs). The being at which it is cast may 
not  resist  the  curse.  A  Deathcurse  must  be  an  af-
fliction,  ill  luck  or  doom.  If  a  doom,  it  will  be 
gasped out with the Adept’s final breath. 

Note Lycanthropy is considered a major curse. 

11.3 Special Knowledge Rituals 

Full  Geas  An  Adept  with  Rank  greater  than  15 
with  the  geas  spell  has  the  power  of  full  geas.  A 
full geas can be placed upon an entity without their 

Remove Curse 

Duration: Immediate 
Experience Multiple: 500 

11 NON-COLLEGE MAGIC 

Resist: None 
Target: A curse 
Concentration Check: Standard 
Effects:  Every  curse  is  rated  by  the  Magical  Apti-
tude (MA) of the Adept who cast it. If the curse is 
natural (such as Lycanthropy) it usually has an MA 
of 20. 

There  are  two  types  of  curses,  minor  ones  and 
major  ones.  A  minor  curse  causes  its  victim  to 
suffer  from  a  non-fatal  malediction.  Minor  curses 
come from various sources, for example the spells 
Evil Eye (G-9 of the College of Ensorcelments and 
Enchantments), the Damnum Minatum (G-1 of the 
Witchcraft  College)  and  certain  backfires.  Major 
curses  normally  come  from  the  Major  Curse  spell 
(§11.2). 

When  a  ritual  of  curse  removal  has  been  com-
pleted,  the  GM  rolls  percentile  dice.  If  the  roll  is 
less  than  or  equal  to  the  success  percentage  the 
curse  is  removed.  If  the  roll  is  between  one  and 
two  times  the  success  percentage,  the  curse  re-
mains in effect. If the roll is equal to or greater than 
twice  the  success percentage,  the  MA  of  the  curse 
is increased by one. This ritual does not backfire in 
the normal fashion. 

Minor Curse 

Base Chance: (15 - MA of curse + 5 × Rank )% 
Cast Time: 6 hours 
Actions: Inscribe symbol of power 
The  Adept  must  inscribe  a  triangle  or  symbol  of 
power  about  the  cursed  being,  and  perform  this 
ritual for six consecutive hours. 

Major Curse 

Base  Chance:  (Adept’s  MA  -  MA  of  curse  +  2  × 
Rank)% 
Cast Time: 18 hours 
Actions: Inscribe symbol of power 
The  Adept  must  have  a  Magical  Aptitude  greater 
than that of the curse. They must inscribe a triangle 
or  symbol  of  power  about  the  cursed  being,  and 
perform  this  ritual  for  eighteen  consecutive  hours. 
If the major curse is a death-curse, Base Chance is 
(Adept’s MA - MA of curse + Rank)% 

Precious  Metals  The  use  of  triangles  or  symbols 
of power fashioned of varying amounts of precious 
metals  causes  an  addition  to  the  success  percent-
age, per the following schedule: 

Metal 

Add  Cost 

1,000 sp 
10,000 sp 
15,000 sp 
20,000 sp 

+ 3 
Silver 
+ 7 
Gold  
Platinum 
+ 10 
Truesilver  + 15 
The  symbol  necessary  for  this  ritual  is  large 
enough  for  the  target  to  sit  in,  and  is  inscribed  in 
the ground. This symbol may be portable. 

\end{Chapter}
