\begin{Chapter}{Spy (Ver 2.0)}

Amongst  the  many  professionals  encountered  in 
everyday  life,  there  will  be  a  scattering  of  those 
with  the  covert  skill  of  Spy.  This  is  a  profession 
dealing  with  obtaining  and  distributing  informa-
tion.  There  are  basic  information  gathering  and 
remembering  abilities  in  common  with  all  spies, 
but the methods of operation, and spheres of influ-
ence  vary  from  the  court  ambassador  to  the  mili-
tary scout to the pub minstrel. A spy may specialise 
in  a  particular  field  of  operation  and  if  they  have 
the  accompanying  skills  appropriate  for  their 
cover, will perform better, with less chance of their 
actions being discovered and their cover blown. 

46.1 Restrictions 
There  are  no  restrictions  on  learning  the  spy  abil-
ity, or to using it, except perhaps the fear of being 
discovered. Note that having learnt the Spy skill is 
not the same as actually spying. 

Discovery 

If  the  spy  catastrophically  fails  in  their  skill  by 
rolling  greater  than  30%  over  the  base  chance  for 
the ability, or 90% + Rank (whichever is the lesser 
BC),  then  they  may  have  been  discovered.  The 
repercussions  of  this  discovery  is  based  on  their 
situation:  what  exactly  they  were  attempting,  who 
the  discoverer  is,  the  difficulty  of  the  task  com-
pared to their rank in the skill, whether the spy has 
other  skills  to  back  up  their  cover  etc.  The  reper-
cussions  may  range  from  a  slap  in  the  face,  being 
given  disinformation,  expulsion  from  an  inn/town, 
a  beating  in  the  back  alley,  through  to  the  tradi-
tional  punishment  for  an  exposed  traitor:  to  be 
drawn  and  quartered,  (although  nobles  are  some-
times beheaded). Thus, a good spy is the one least 
likely to be discovered. 

46.2 Benefits 
All spies gain grounding in basic spy craft, includ-
ing memory enhancement, moving quietly, observ-
ing closely and communicating with their peers. 

Enhanced Memory  

•  A  spy  may  memorise  and  recall  visual  details, 
such  as  those  of  a  room,  a  person,  or  a  piece  of 
parchment, etc. Memorisation requires (120 - 10 × 
Rank) seconds of undisturbed concentration, study-
ing the object.  

•  A  spy’s  chance  to  recall  a  memorised  image 
accurately is (2 × Perception + 12 × Rank)% rolled 
by  the  GM.  If  the  failed  recall  attempt  occurs 
within  1  +  (1  ×  Rank)  days  of  the  memorisation, 
the spy merely cannot remember. After this period, 
their  effective  rank  for  recall  reduces  by  one  per 
subsequent  day,  and  with  a  failed  roll,  erroneous 
information  may  be  remembered  instead.  If  a  spy 
fails  to  recall  an  object  or  place,  they  may  not 
attempt to recall it again until they study it again.  

•  A  spy  may  also  use  this  ability  to  recall  spoken 
phrases  and  combination  of  sounds.  Even  if  a  spy 
does  not  know  the  language  used,  they  can  repro-
duce the phrases phonetically.  

•  A  spy’s  enhanced  memory  gives  them  an  effec-
tive  understanding  of  any  spoken  language  that 
they have at Ranks 0-3 as if one rank higher. 

Stealth 

A spy increases their chance of acting stealthily by 
2% per Rank. 

Fieldcraft 

A spy is trained to notice, recognise, and appropri-
ately interact with other spies. This includes hand-
ing off messages discretely, the use of dead-drops, 
safe-houses  and  identification  phrases,  as  well  as 
foiling  the  less  subtle  attempts  to  interfere  with 

such  exchanges.  Also,  they  may  identify  the  weak 
willed,  manipulable,  morally  suspect  or  gullible 
individuals that a spy prefers to associate with, and 
has a greater chance of enticing them to a course of 
action  by  coaxing,  flattery,  wheedling,  blackmail 
or other enticement. 

Optional Abilities 

In addition, a spy gains an ability chosen from the 
list  below  with  each  Rank  (including  Rank  0). 
Additional abilities may be gained without increas-
ing in rank by the expenditure of 2,500 Experience 
Points  and  4  weeks  of  training.  These  costs  are 
discounted  by  25%  if  the  Spy  has  reached  rank  8, 
or by 50% if they have reached rank 10. Individual 
Base Chances are provided for some of the various 
spy  abilities;  for  the  other  skills  there  is  a  generic 
Base Chance of 3 × appropriate characteristic (+ 5 / 
Rank), modified by difficulty. 

Assess:  a  spy  can  infer  some  information  from 
observing  the  grouping  and  activities  of  people. 
The level of information gained is logistical, rather 
than  the  in  depth  knowledge  that  knowing  the 
associated skill would give. Example uses include:  

•  Estimate  Entourage  ascertain  the  likely  social 
ranking  of  the  target  based  on  the  size,  quality, 
snootiness  etc  of  accompanying  servants.  Evalua-
tion  of  who  they  are  or  the  implications  of  the 
household makeup would require Courtier.  

•  Estimate  Goods  estimate  the  number  of  boxes, 
and  the  people,  wagons,  time  and  other  logistics 
required  to  move  them.  Evaluation  of  quality  or 
value would require Merchant.  

•  Troop  Estimation  estimate  size,  equipment  and 
quality  of  a  military  force  or  navy.  Inferring  the 
tactics  this  represents  would  require  Military  Sci-
entist. 

Befriend:  over  a  period  of  days  or  weeks,  a  spy 
may  target  individuals  to  engender  trust  in  them-
selves, and possibly distrust in others. The “friend” 
begins  to  willingly  and  unknowingly  reveal  infor-
mation to the spy. 

Bribery:  a  spy  may  recognise  appropriate  people 
and  which  “gifts”  they  prefer,  to  gain  information 
or access. They may also recognise situations when 
bribing will not work, before making the attempt. 

Codes:  a  spy  can  recognise  and  use  simple  codes. 
They  may  attempt  to  break  others’  codes  and  ci-
phers.  The  length  of  study  required  to  do  this  de-
pends on the familiarity of the code’s style and the 
difficulty set by the code maker. 

Counterspy:  a  spy  may  perceive  other  current 
spying activity and recognise the other spies. Also, 
a  spy  may  create  convincing  artifices  that  fit  with 
available  information,  to  spread  disinformation, 
misunderstanding or confusion. 

Disguise:  with  the  use  of  physical  props  and  re-
sources, the spy can apply makeup, dyes, false hair 
etc to convincingly alter the appearance of people. 
The success of disguising race and gender depends 
on  the  physical  similarity  of  the  person,  and  how 
closely the disguised entity is inspected. A spy may 
only  attempt  to imitate  a  specific  person’s  appear-
ance  after  prolonged  study  of  the  target.  Animals 
the spy is familiar with may also be disguised. 

Forgery:  With  the  right  materials  to  hand,  a  spy 
can  create  convincing  replicas  of  personal  letters, 
official  documents  and  the  like.  This  includes  the 
ability  to  open  and  re-seal  letters,  produce  false 
seals,  or  move  real  ones  to  a  forgery.  The  chance 
for this is MD + PC + (4 × Rank). If a spy is liter-
ate  with  the  language  of  the  document,  then  the 

language rank can be added to the base chance for 
forgery. 

Hiding:  a  spy  can  find unlikely  hiding  places  and 
conceal themselves for long periods of time, keep-
ing still and quiet. The spy must roll WP + EN + (5 
×  Rank)  to  maintain  this  for  extended  periods  of 
time. 

Imitation:  the  spy  can  study  and  copy  behaviour, 
mannerisms, and accents. This will allow the spy to 
maintain the roles of ordinary people and not stand 
out  as  a  foreigner.  As  a  guide,  this  takes  6  - 
(Rank/2)  hours  of  exposure  to  the  society,  with 
modifiers  based  on  the  apparent  familiarity  or 
strangeness of the society. 

Information: by using other skills and knowledge 
a  spy  has  an  increased  chance  of  getting  the  most 
relevant  information  they  are  seeking.  They  have 
an  increased  chance  of  noticing  disinformation 
about  their  area  of  knowledge,  and  distinguishing 
fabrication and pretence from fact and reality. The 
spy may recognise information as having worth to 
other  individuals  or  spies  in  their  network,  even 
though it is not useful to them. 

Lip  Reading:  a  spy  may  understand  spoken  con-
versations  outside  of  their  hearing  range.  A  clear 
line  of  sight  to the  targets’  face  and knowledge  of 
the language being spoken (min rank 6) is required. 
The  spy  may  garner  fragments  but  not  all  of  the 
conversation  if  they  are  not  fluent  with  the  lan-
guage,  can  only  see  one  of  the  participants,  or  an 
unlikely topic is being discussed. 

Pick  Locks:  while  using  appropriate  tools,  a  spy 
can spend (240 - (20 × Rank) ) seconds to attempt 
to pick a lock to either unlock or lock it. The base 
chance of the attempt is (MD + (4 × Rank) ) - (6 × 
Lock Rank). If the attempt fails, the lock resists the 
attempt.  A  catastrophic  failure  may  damage  the 
lock. 

Resist  Torture:  being  familiar  with  extraction 
techniques, a spy may add their Rank to their effec-
tive  WP  for  resisting  torture  attempts.  They  may 
also choose to apparently lower their WP by up to 
Rank to release false information. 

Shadowing:  a  spy  learns  the  skills  of  following 
individuals at a distance without being observed. In 
addition they have an increased chance of noticing 
when they themselves are being followed, and may 
attempt  to  lose  their  followers  if  the  appropriate 
terrain is available. To spot a tail the Spy must roll 
under  (2  ×  PC)  +  (5  ×  Rank)  -  5  times  opposing 
Spy’s  rank.  To  successfully  tail  another  spy  re-
quires  a  roll  of  PC  +  AG  +  (5  ×  Rank)  -  5  times 
opposing Spy’s rank. 

Sleight  of  Hand:  a  spy  can  palm,  swap  or  place 
small  objects,  without attracting  notice.  Removing 
objects from people requires the Pick Pockets skill 
from Thief. 

A  GM  may  give  the  following  abilities  to  a  Spy 
over  the  course  of  play,  or  they  may  be  requested 
by  a  player  as  part  of  their  character  knowledge 
and backgrounds. 

Network (area): the spy has joined a spy ring in a 
specific  area,  and  knows  the  specific  field  craft 
routines used by that ring. 

Spy Master (area): the spy has set up a spy ring in 
a  specific  area,  and  knows  where  to  place  spies, 
how  to  store  information  for  later  reference,  and 
how  to  manage  other  spies.  This  ability  requires 
Rank 8 plus. 

\end{Chapter}
