\begin{Chapter}{The College of Greater Summonings (Ver 1.0)}

This  is  a  non-player  college  only.  Player  Char-
acters  may  not  learn  it,  transfer  to  it,  or  learn 
any of the spells or rituals with the exception of 
the counterspells. 

The  College  of  Greater  Summonings  is  concerned 
exclusively with the summoning and controlling of 
entities  from other  dimensions.  All  such  summon-
ings  and  associated  magical  procedures  are  Ritual 
Magic.  Members  of  the  College  of  Greater  Sum-
monings  possess  no  Talent  or  Spell  Magic  as  a 
result  of  their  association  with  the  College.  Their 
power  lies  exclusively  in  their  ability  to  summon 
and  control  beings  via  the  performance  of  special 
rituals. 

All  summonings  of  this  College,  regardless  of  the 
type  of  being  they  are  designed  to  summon,  are 
performed in the same manner. First the summoner 
must  perform  a  Ritual  Cleansing  of  their  body 
(requiring  from  1  to  10  hours).  Then  they  must 
prepare  and  implement  the  proper  Ritual  of  Sum-
moning.  The  Ritual  of  Summoning  itself  requires 
only  one  hour  to  execute.  Once  the  Adept  has 
performed  the  Ritual  of  Summoning,  they  may 
wish to control the entity they have  summoned by 
implementing either a Ritual of Binding or a Ritual 
of  True  Speaking.  Binding  and  True  Speaking 
Rituals  each  require  one  hour  to  prepare  and  im-
plement.  If  the  summoner  does  not  implement 
these  rituals  immediately  upon  successfully  sum-
moning the entity who is the target, they will have 
to  engage  in  a  new  Ritual  of  Cleansing  before 
attempting to enact another ritual. 

When  performing  the  summoning  rituals  of  this 
College, the summoner occupies a Circle of Protec-
tion  which  also  contains  within  it  a  Pentacle  of 
Power.  Unless  the  summoner  and  companions 
occupy  this  protective  area,  there  is  a  chance  that 
whatever they summon will be able to turn on them 
and destroy them. Only the summoner occupies the 
Pentacle. Their companions occupy the Circle, but 
are outside the Pentacle. 

The summoned entity appears in the vicinity of the 
Circle of Protection, but outside it (unless the ritual 
backfires  and  the  Circle  is  broken).  In  order  to 
perform  (and  as  part  of  the  performance  of)  the 
Rituals  of  True  Speaking  and  Binding,  the  sum-
moned  entity  is  forced  into  the  Triangle  (outside 
the Circle of Protection). 

The  Pentacle,  Circle  of  Protection,  and  Triangle 
must all be drawn on the ground by the summoner 
before the rituals are performed. 

24.1 Restrictions 

Members of the College of Greater Summonings 
must  meet  requirements  of  time,  place,  equip-
ment,  knowledge,  and  circumstance  in  order  to 
perform their College’s magic. 

The  following  list  of  requirements  must be  met  in 
all  or  most  situations  wherein  the  arts  of  the  Col-
lege of Greater Summonings are practised: 

1.  The  Adept  must  know  the  spell  or  ritual  being 
employed. 

2. If a summoning of an entity other than an Incu-
bus  or  Succubus  is  being  performed,  the  Adept 
must know the name of the entity being summoned 
and speak it during the ritual. 

3.  Throughout  all  preparations  for  the  ritual  and 
throughout the ritual itself, the Adept must remain 
stationary. They may engage in no other activity. 

4. The space used for the performance of the ritual 
must  be  large  enough  to  accommodate  the  ritual 
symbols  and  the  entity  being  summoned  (usually, 
an  area  30’  ×  30’  and  12’  high  is  adequate)  and 
may not have been affected by an area counterspell 
in the last 12 hours. 

5.  If  the  magic  is  a  Special  Knowledge  Ritual,  it 
may  only  be  performed  on  certain  days  of  the 

month  and  at  certain  hours  of  the  day  (listed  in 
§24.7). 

roll is made for each companion if more than one is 
granted. 

6. The Adept must possess the necessary tools and 
equipment  to  perform  any  ritual  and  must  have 
access  to  such  substances  as  clear  water  for  the 
cleansing which precedes each ritual. 

If  the  Adept  fails  to  meet  any  of  these  require-
ments, they  may  not  attempt to  perform any  ritual 
of  the  College  of  Greater  Summoning.  They  may 
still  exercise  any  special  Talent  Magic  they  may 
possess and may cast counterspells. 

The MA requirement for this College is 9. 

24.2 Base Chance Modifiers 
The  following  numbers  are  added  to  the  Base 
Chance  of  successfully  performing  a  counterspell 
or ritual of the College of Greater Summoning: 

Each Rank the Adept has achieved with the 
spell or ritual 
Adept occupies a Mana-rich place 
+15 
The  following  numbers  are  added  to  the  Base 
Chance  of  successfully  performing  a  ritual  of  the 
College of Greater Summoning: 

+3 

+3 

+20 

Each hour (maximum of 10) Adept spends in 
Ritual Cleansing prior to performing the 
ritual 
Ritual is part of the College’s body of Gen-
eral Knowledge and is begun at midnight 
Ritual is part of the College’s body of Gen-
eral Knowledge and is performed while the 
moon Luna is 2, 4, 6, 8, 10, 12 or 14 days 
into its cycle 
Special Knowledge Rituals are affected differently 
from General Knowledge Rituals so far as the hour 
of  the  day  or  day  of  the  moon  in  which  they  are 
performed  is  concerned.  The  effects  of  the  hour 
and day on these rituals is discussed in §24.7. 

+20 

24.3 Talents 
There  are  no  Talent  Magics  granted  an  Adept 
simply because they are a member of this College. 
However, there is a possibility that the Adept may 
be  assigned  a  “companion”  by  a  demon  they  have 
summoned.  Once  an  Adept  has  been  assigned  a 
companion, they may automatically call forth (and 
later  dismiss)  that  companion.  Once  called  forth, 
the  companion  instantly  appears  to  do  the  bidding 
of  the  Adept.  In  this  sense,  there  is  some  Talent 
Magic available to members of this College. 

Upon  dismissing  a  demon  they  have  summoned, 
the  summoner  may  request  that  the  demon  assign 
them a companion from among its legions of lesser 
spirits.  The  Base  Chance  that  a  demon  will  grant 
this  request  is  listed  under  the  “Lesser  Spirits” 
heading  for  each  demon.  The  summoner  rolls 
D100,  and  if  the  resulting  number  is  less  than  or 
equal  to  the  Base  Chance,  the  demon  grants  a 
companion.  If  the  number  is  half  or  less  than  the 
Base Chance, the demon grants two companions. 

A  companion  will  be  either  a  lesser  Devil  or  an 
imp  who  is  constantly  “on  call”  to  the  summoner. 
The companion is available until one of the follow-
ing events occurs: 

1.  The  summoner  requests  a  companion  from  any 
other  Demon  except  the  Demon  who  granted  the 
summoner’s  current  companion  (regardless  of 
whether or not the request is granted). 

2. The companion is dispelled by having a General 
Knowledge Counterspell of the College of  Greater 
Summonings cast over it by the summoner. 

3. The companion is forced into its own dimension 
by a reduction in its Endurance sufficient to cause 
unconsciousness. 

4.  The  summoner  voluntarily  attacks  the  compan-
ion or quits the College of Greater Summonings. 

When a companion is granted, the summoner rolls 
D10.  If  the  result  is  1–4,  their  companion  will  be 
an imp. If it is 5–10, it will be a devil. A separate 

83 

24.4 General Knowledge Spells 
The  College  of  Greater  Summonings  General 
Knowledge  and  Special  Knowledge  Counterspells 
are the only spells usable by members of this Col-
lege.  See  §10.2  for  descriptions  of  the  nature  and 
working of counterspells. 

24.5 General Knowledge Rituals 
There  are  six  rituals  that  an  Adept  learns  upon 
becoming a member of this College. The first is the 
Ritual of Cleansing in which the Adept bathes their 
body and meditates as a prelude to attempting any 
other  ritual.  It  is  not  possible  to  attempt  any  other 
ritual  of  this  College  unless  one  has  first  engaged 
in  the  Ritual  of  Cleansing.  There  are  three  sum-
moning rituals that are part of the body of General 
Knowledge  of this College. They are the Ritual of 
Summoning  Succubi,  the  Ritual  of  Summoning 
Incubi,  and  the  Ritual  of  Summoning  Heroes. 
There  are  two  non-summoning  rituals  associated 
with  the  General  Knowledge  of  this  College  be-
sides  the  Ritual  of  Cleansing.  They  are  the  Ritual 
of Binding and the Ritual of True Speaking. These 
two rituals may only be employed after a summon-
ing  ritual  has  been  successfully  completed.  They 
are  sometimes  necessary  to  gain  the  required  ser-
vices of beings from other dimensions. 

Also as part of the study of the General Knowledge 
of  this  College,  the  Adept  receives  a  parcel  con-
taining  the  tools  necessary  to  their  magic.  The 
parcel consists of the following: 

Item 

Weight  Value 

2 lb 

1 lb 
6 lb 

250sp 
150sp 

1 lb 
5 lb 
3 lb 
3 lb 

50sp 
10sp 
20sp 
100sp 

1 sceptre of dogwood 
1 broadsword of silvered 
steel 
1 mitre 
1 cap 
1 robe of virgin linen 
1 girdle of lion’s skin im-
printed with symbols 
1 censor of silver and gems 
3000sp 
In  addition  to  these  tools,  the  Adept  is  provided 
with a supply of materials which together weigh 5 
pounds  and  which  must  be  replenished  on  the 
average  of  every  three  months  at  a  cost  of  1000 
Silver  Pennies.  These  items  include:  A  packet  of 
charcoal,  a  packet  of  powdered  agrimony  (for 
making  tea  used  in  the  Ritual  of  Cleansing),  a  3 
ounce  tin  of  myrrh,  a  vial  of  white  frankincense 
dissolved  in  white  wine,  a  6  ounce  pot  of  sandal-
wood  and  powdered  antimony,  a  2  ounce  tin  of 
ambergris salve, an 8 ounce box of multi-coloured 
chalks (for drawing the Triangle, Circle, Hexagram 
and Pentagram necessary for the various Rituals of 
Summoning). 

The  Adept  is  also  taught  the  Generic  True  Names 
of  all  things  that  occupy  dimensions  other  than 
their  own  and  may  be  summoned  to  this  dimen-
sion. They also learn the Individual True Names of 
all  heroes  of  other  dimensions  known  to  the  Col-
lege  and  of  the  72  great  Demons  of  the  Seventh 
Plane.  The  Adept  does  not  learn  the  rituals  neces-
sary  to  summon  Demons  from  the  Seventh  Plane 
— only the names of the Demons and how to iden-
tify  them.  A  member  of  this  College  may  learn 
additional  Individual  True  Names  of  entities  from 
other planes by first acquiring them from a Demon 
and  then  studying  the  True  Names  in  the  same 
manner  as  a  member  of  the  College  of  Naming 
Incantations. 

The  following  Rituals  constitute  the  major  part  of 
the  knowledge  the  Adept  acquires  in  their  general 
training. 

Cleansing (Q-1) 

Effects:  The  Ritual  of  Cleansing  must  be  per-
formed  prior  to  any  other  rituals  of  this  College. 
The  Adept  cleans  their  body  and  purifies  their 

24 COLLEGE OF GREATER SUMMONINGS 

mind  for  from 1  to 10  hours.  The  Base  Chance  of 
any succeeding Ritual being successful is increased 
by 3 for each hour spent in Ritual Cleansing. There 
is  no  Base  Chance  for  this  ritual  being  successful. 
The  Adept  states  the  number  of  hours  they  will 
expend on the ritual and at the end of that period of 
time, the Adept is cleansed. They may perform no 
other  activity  while  engaged  in  this  ritual.  If  their 
concentration  is  broken,  they  must  restart  the 
cleansing from the beginning or abandon the effort. 
Any  rituals  the  Adept  desires  to  perform  after  the 
cleansing  must  be  performed  immediately.  They 
may  perform  any  number  of  rituals  within  three 
hours  of  the  cleansing,  but  these  rituals  must  im-
mediately follow each other. Any time spent in any 
other  activity  destroys  the  effects  of  the  cleansing 
and  a new  cleansing  must be  accomplished  before 
another ritual can be enacted. 

Summoning Succubi (Q-2) 

Experience Multiple: 300 
Base Chance: 5% + 3 / Rank 
Effects: This ritual may be used to summon one or 
more  Succubi,  who  will  arrive  on  this  plane  fa-
vourably  disposed  towards  the  summoner  for  hav-
ing summoned them, and thus will not immediately 
need to be bound. Succubi arrive on this plane with 
only  one  goal:  the  enticement  and  mating  with 
humans.  Any  deed  the  summoner  desires  of  them 
which  will  further  their  goal  will  be  approved  and 
encouraged  by  the  Succubi.  If  the  Succubi  are 
summoned  to  perform  a  task  not  immediately 
related  to  their  goal,  they  will  have  to  be  Bound 
and forced to do the task. The Base Chance is 5% 
(+ 3 / Rank), and it is reduced by 10 for each Suc-
cubi  above  one  being  summoned  at  once.  If  the 
ritual  backfires,  the  Succubi  will  appear  and  at-
tempt to molest and then devour the summoner and 
his or her companions. 

Succubi remain on this plane and serve for a num-
ber of days equal to [D + 4]. The die is rolled indi-
vidually  for  each  Succubus  successfully  sum-
moned. The Succubi may also be forcibly returned 
to  their  own  plane  whenever  their  Endurance  is 
reduced  to  the  point  that  they  are  reduced  to  un-
consciousness  or  a  counterspell  is  cast  over  them 
by their summoner. 

Once  returned  to  their  own  plane,  they  may  not 
return to  the  aid  of the  summoner  unless the  sum-
moner  performs  another  Ritual  of  Summoning 
Succubi. 

Summoning Incubi (Q-3) 

Experience Multiple: 300 
Base Chance: 5% + 3 / Rank 
Effects: This ritual operates in the same manner of 
Q-2 (the Ritual of Summoning Succubi). Incubi are 
exactly  like  Succubi,  except  that  they  appear  only 
in the male form, rather than as females. They will 
tend  to  be  2–3  points  stronger  and  will  have  1–2 
points  less  Endurance  than  Succubi,  but  will  oth-
erwise be the same. 

Summoning Heroes (Q-4) 

Experience Multiple: 500 
Effects: The Adept may summon a great hero from 
another dimension to assist them. This hero may be 
any  character  from  the  body  of  fantasy  literature 
known  to  both  the  Adept  and  the  GM.  The  GM 
always  sets  the  characteristics  of  the  hero,  their 
weaponry  and  armour,  the  number  and  type  of 
companions  (if  any)  and  the  length  of  time  and 
terms under which they will remain in the Adept’s 
dimension and assist them. The GM  may limit the 
use  of  this  ritual  to  periods  when  various  celestial 
bodies are in conjunction (once a game year or so). 
The  GM  need  not inform the  Adept of  any  details 
concerning  the  results  of  the  summoning  until  it 
has been performed. 

The  Base  Chance  that  this  ritual  will  succeed  will 
vary according to the hero the Adept is attempting 
to  summon,  but  should  usually  be  set  at  less  than 
20%. 

True Speaking (Q-5) 

Experience Multiple: 250 
Base Chance: 50% 
Effects:  Whenever  any  entity  except  a  hero  is 
summoned,  the  summoner  may  wish  to  ask  them 
questions. In all cases, the truthfulness of the entity 
summoned  (Incubus,  Succubus,  Demon)  will  be 
less  than  100%.  Sometimes  it  will  be  very  near 
100%, but in the case of a Demon  who is particu-
larly  resentful  that  they  have  been  summoned  (or 
one that is an habitual liar) the chance of the entity 
telling  the  truth  may  be  as  low  as  5%.  The  only 
way to be sure that the entity will tell the truth is to 
perform a Ritual of True Speaking and then ask the 
entity  for  answers  to  the  desired  questions  during 
the  ritual.  The  entity  is  required  to  stand  within  a 
triangle and give answers to the summoner’s ques-
tions during  the  ritual.  The  ritual  lasts  for  one  full 
hour,  and  the  summoner  may  do nothing  else  dur-
ing  the  time  the  ritual  is  in  progress  except  ask 
questions.  The  effects  of  the  ritual  do  not  extend 
beyond  the  end  of  the  ritual.  The  Base  Chance  of 
forcing the entity into the triangle and successfully 
performing  the  ritual  is  50%.  The  GM  rolls  D100 
to  determine  the  success  of  the  ritual.  They  need 
not inform players of the result of the dice roll. The 
entity may passively resist the effects of the ritual. 

Binding (Q-6) 

Experience Multiple: 400 
Effects: An Adept may perform a Ritual of Binding 
in order to bind an entity (other than a hero) to this 
plane. They cannot normally control a Demon that 
is not bound, except while inside the Pentacle they 
occupied when they summoned the demon. Before 
they  can  leave  the  Pentacle  or  require  any  service 
of the demon except for conversation and teaching 
of skills, the summoner must in most cases perform 
a  Ritual  of  Binding.  The  Base  Chance  of  success-
fully  binding  a  particular  demon  is  equal  to  half 
(rounded  down)  the  unmodified  Base  Chance  to 
summon  that  demon  initially.  The  Demon  can 
actively resist the workings of a Ritual of Binding. 

Once bound, a demon will remain on this plane for 
a number of days equal to [D - 3] and will serve the 
summoner more or  less willingly during that time. 
In  some  cases,  special  conditions  must  be  met  to 
bind  the  demon.  Usually,  this  means  the  giving  of 
presents,  most  often  human  life.  If  the  demon 
resists such an offering, it will have also broken the 
summoning and may turn on the summoner despite 
the  fact  that  the  summoner  is  protected  by  the 
Pentacle.  When  this  happens,  the  summoner  rolls 
D100. If the result is less than or equal to the sum-
moner’s  Magic  Resistance,  the  demon  is  banished 
to  its  own  plane.  If  the  result  is  greater  than  the 
summoner’s  Magical  Resistance,  the  summoner  is 
overcome by the will of the demon and voluntarily 
breaks the circle of protection allowing the demon 
to enter the pentacle and attack them. 

Some demons  will serve more  willingly and faith-
fully  than  others.  The  details  of  which  demons 
serve willingly and which resist service even when 
bound  are  discussed  under  the  heading  dealing 
with  each  particular  demon.  A  summoner  may 
control  only  one  bound  demon  at  a  time,  though 
they  may  summon  other  demons  (or  non-demonic 
entities)  for  purposes  of  conversation.  They  may 
voluntarily  dismiss  a  demon  at  any  time  prior  to 
the end of its service and the demon is immediately 
returned to its own dimension unless it has already 
broken the binding and turned on the summoner. 

24.6 Special Knowledge Spells 
There  are  no  Special  Knowledge  Spells  for  this 
College. 

24.7 Special Knowledge Rituals 
There  are  six  separate  Special  Knowledge  Rituals. 
Each  is  designed  to  summon  a  particular  Rank  of 
demon.  There  are  six  Ranks  of  demons:  Duke, 
Prince,  President,  Earl,  Marquis,  and  King.  Each 
Rank  is  subject  to  certain  limitations  as  to  where 
and when it can be summoned. 

84 

All  Ranks  of  demons  are  summoned  in  the  same 
manner. The summoner announces the demon they 
are summoning and that demon’s Rank. They then 
perform  the  appropriate  Ritual  of  Summoning.  At 
the end of the ritual (that is after one hour), a check 
is made to see if the ritual has been effective. The 
summoner  rolls  D100.  If  the  result  is  equal  to  or 
less  than  the  Base  Chance  of  summoning  the  par-
ticular  demon  that  is  the  object  of  the  ritual,  the 
demon  is  summoned  and  appears  before  the  sum-
moner. Otherwise, the demon does not appear and 
the  summoner  may  not  make  a  further  attempt  to 
summon that demon that day. They may attempt to 
summon  another  demon  instead  but  must  first 
repeat  the  Ritual  of  Cleansing.  The  description  of 
each demon lists the Base Chance to summon that 
demon. 

Special Knowledge Rituals may only be performed 
on days 2, 4, 6, 8, 10, 12 and 14 of the cycle of the 
moon, Luna. They may be attempted on other days, 
but can never have any effect. There is a possibility 
that an Adept could lose track of time and attempt 
to  employ  a  Special  Knowledge  Ritual  on  a  day 
when it will not work.  In such cases, the GM may 
choose  not  to  inform  the  individual  that  the  ritual 
can have no effect and may allow them to perform 
it  anyway,  only  telling  the  player  why  they  have 
been unable to summon a demon at the end of the 
ritual. 

An  Adept  who  knows  a  particular  Ritual  of  Sum-
moning  may  summon  any  of  the  demons  of  that 
Rank  Each  description  includes:  the  demon’s 
name;  the  Base  Chance  of  summoning  (and  bind-
ing)  it;  the  percentage  chance  that  the  demon  will 
agree  to  grant  the  summoner  a  companion  from 
among  its  legions  of  lesser  spirits;  the  special  tal-
ents,  skills  and  magical  abilities  of  the  demon;  a 
quantification  (given  as  a  span  of  possible  num-
bers)  of  the  demon’s  characteristics;  the  demon’s 
natural armour (given under the heading NA as the 
number of hits absorbed for each Strike); the natu-
ral weapons of the demon (and any other weapons 
habitually  carried),  and  any  special  comments  on 
the  demon’s  nature  or  abilities.  Also  included  is  a 
short physical description of the demon. 

Once  a  demon  is  dismissed  (returned  to  its  own 
dimension)  it  cannot  return  to  this  plane  in  less 
than a day.  A demon  who has been  dispelled by a 
counterspell or rendered unconscious may return to 
this plane (by being re-summoned in a new ritual) 
only  after  one  full  month  in  its  own  plane  (spent 
reforming the scattered energy pulses that make up 
its being). 

Demons  may  be  controlled  while  the  summoner 
stands within the Pentacle which they must draw to 
perform  the  summoning  ritual.  They  will  speak  to 
the  summoner  under  this  circumstance  and  will 
sometimes  tell  the  truth.  To  ensure  absolute  hon-
esty, a Ritual of True Speaking must be performed. 
To  ensure  that  the  demon  will  not  destroy  the 
summoner once the ritual is over, a Ritual of Bind-
ing must be performed. A demon who is not bound 
must  be  dismissed  at  the  end  of  the  summoning 
(once  the  summoner  has  finished  speaking  to  the 
demon)  and  will  then  usually  return  to  its  own 
plane.  However,  a  demon  who  is  particularly  sav-
age or who has been offered a gift which it rejects 
will  attempt  to  devour  the  summoner  before  de-
parting.  The  summoner  must  then  make  a  check 
against  the  Willpower  to  determine  if  the  sum-
moner  breaks  the  magical  circle  protecting  them 
and  fights  the  demon  or  remains  safely  within  the 
Pentacle  (in  which  case,  the  frustrated  demon 
departs). 

Adepts  are  not  provided  with  a  Shield  when  they 
are  admitted  to  membership  in  the  College  of 
Greater Summoning, as a Shield is not necessary to 
perform  the  rituals  of  the  College.  However,  they 
may  wish  to  make  or  have  made  a  special  Shield 
(actually a disc of metal engraved with symbols of 
occult power) to protect them during the summon-
ing  of  demons.  This  shield  is  made  by  Adepts  of 
the  College  of  Shaping  Magics  —  see  Arcane 

Wisdom. A summoner who does not have a Shield 
can suffer a backfire. A summoner who is using the 
proper  Shield  cannot  suffer  a  backfire.  A  backfire 
from  a  Special  Knowledge  Ritual  consists  of  the 
appearance  of  the  demon  being  summoned,  but 
inside the circle of protection (Pentacle) so that the 
demon is free to attempt to destroy the summoner. 
In  such  cases,  the  demon  may  be  returned  to  its 
own  dimension by  a  counterspell  or  by  being  ren-
dered  unconscious,  but  will  otherwise  remain  on 
this  plane  and  freely  roam  about  attacking  and 
destroying  until  somehow  banished.  It  will  not, 
however  begin  roaming  the  earth  until  it  has  de-
stroyed its summoner. A backfire occurs whenever 
an Adept rolls a number which is 30 or more than 
the modified Base Chance of summoning a particu-
lar  demon  while  making  a  check  to  see  if  that 
demon  is  summoned.  A  backfire  is  treated  as  “no 
effect”  if  the  summoner  has  the  proper  Shield  in 
their hand during the ritual. 

The  individual  Shields  that  will  protect  a  sum-
moner  from  backfire  are  discussed  under  each 
individual  Ritual  of  Summoning.  Shields  only 
affect the summoning of demons. There is no back-
fire due to an ineffective attempt at employing any 
other Ritual of Summoning. Rituals of True Speak-
ing  and  Binding  may  backfire  (regardless  of 
whether or not a Shield is employed) and result in 
the  summoner  being  affected  by  their  own  ritual 
and  forced  either  to  answer  all  questions  of  the 
demon  as  truthfully  as  possible  or  to  serve  the 
demon so long as the demon remains on this plane. 
This backfire result may be passively resisted. 

Note:  The  interaction  between  a  player  and  the 
“demon  characters”  they  summon  is  the  most  im-
portant aspect of the workings of this College, and 
the GM should strive to keep players on their toes 
by  developing  the  demonic  character  as  fully  as 
possible  (making  it  cooperative  about  some  things 
and  uncooperative  about  others,  for  example). 
Some  demons  are  savage  in  the  extreme  and  will 
always be out to do what damage they can short of 
murdering  the  summoner  (and  sometimes  that  is 
not excluded). Others are milder and will pass up a 
golden opportunity to devour their summoner. The 
notes on individual demons are meant to serve as a 
guide  to  their  characters  as  well  as  their  abilities. 
Their  descriptions  also  give  the  forms  in  which 
they may appear. Often, these forms will be insub-
stantial  and  the  demon  will  have  no  power  (nor 
will anyone have power over it) while it is in those 
forms.  However,  the  information  is  included  as 
clues  for  the  GM  in  structuring  the  demonic  char-
acter. 

Summoning Demonic Dukes (R-1) 

Experience Multiple: 300 
Effects:  This  ritual  is  used  to  summon  the  follow-
ing  demons  from  the  seventh  plane:  Agares,  Aim, 
Alloces,  Amdusias,  Astaroth,  Barbatos,  Bathin, 
Berith,  Bune, Crocell,  Dantalion,  Eligos,  Furcalor, 
Furcas, Gremory, Gusion, Havres, Murmur, Sallos, 
Uvall,  Valefor,  Vapula,  Vephar  and  Zepar.  The 
demons  of  this  Rank  may  only  be  summoned  be-
tween sunrise and noon on days  when the weather 
is  clear  and  the  sun  can  be  seen.  Any  attempt  to 

perform this ritual at any other time will be totally 
ineffective. 

The only Shield that will protect against the possi-
bility of backfire while summoning Dukes is a disk 
of purest copper 

12  inches  across,  inscribed  with  the  names  and 
signs  of  all  the  Dukes  of  the  seventh  plane.  This 
shield  may  only  be  manufactured  by  a  Shaping 
mage.  It  weighs  2  pounds  and  the  average  cost  of 
manufacture will be 3000+ Silver Pennies. It takes 
about three months to manufacture. 

Summoning Demonic Princes (R-2) 

Experience Multiple: 350 
Effects:  This  ritual  is  used  to  summon  the  follow-
ing  demons  from  the  seventh  plane:  Gasp,  Ipos, 
Orobas,  Seir,  Sitri,  Siolas,  and  Vassago.  These 
demons may be summoned at any time of the day. 

The only Shield that will protect against the possi-
bility  of  backfire  while  summoning  Princes  is  a 
disk of hammered tin inscribed with the names and 
symbols  of  the  Princes  of  the  seventh  plane.  This 
Shield  may  only  be  manufactured  by  a  Shaping 
mage.  It  weighs  3  pounds  and  the  average  cost  of 
manufacture will be 3000+ Silver Pennies. It takes 
about three months to manufacture. 

Summoning Demonic Presidents (R-3) 

Experience Multiple: 350 
Effects:  This  ritual  is  used  to  summon  the  follow-
ing  demons  from  the  seventh  plane:  Avnas,  Buer, 
Camio,  Foras,  Haagenti,  Labolas,  Malphas,  Mar-
bas,  Volac  and  Voso.  These  demons  may  only  be 
summoned during daylight. 

The only shield that will protect against the possi-
bility of backfire while summoning Presidents is a 
disk  of  base  metal  (other  than  cold  iron)  coated 
with quicksilver. This shield may only be manufac-
tured  by  a  Shaping  mage.  It  weighs  3  pounds  and 
the  average  cost  to  manufacture  will  be  4000+ 
Silver Pennies. It takes about four months to manu-
facture. 

Summoning Demonic Earls (R-4) 

Experience Multiple: 500 
Effects:  This  ritual  is  used  to  summon  the  follow-
ing  demons  from  the  seventh  plane:  Andromalius, 
Bifrons, Botis, Furfur, Nlaithus, Marax, Raum and 
Renove. These demons may be summoned only in 
woods  and  hills  and  only  in  places  that  are  quiet. 
They may be summoned at any time of the day or 
night. 

The only shield that will protect against the possi-
bility of backfire  while summoning Earls is a disk 
of  hammered  bronze  inscribed  with  the  names  of 
the Earls of the seventh plane. This shield may be 
manufactured  by  Shaping  magicians.  It  weighs 
three  pounds  and  the  average  cost  to  manufacture 
will  be  5000+  Silver  Pennies.  It  will  take  about 
four months to manufacture. 

Summoning Demonic Marquis (R-5) 

Experience Multiple: 550 
Effects:  This  ritual  is  used  to  summon  the  follow-
ing  demons  from  the  seventh  plane:  Amon,  An-
dras,  Andrealphus,  Cimejus,  Decarabia,  Forneus, 

24 COLLEGE OF GREATER SUMMONINGS 

Leraje,  Marchosias,  Naberius,  Orias,  Phenex, 
Samieina,  Savnok  and  Shaz.  These  demons  may 
only be summoned between 3 in the afternoon and 
sunrise. 

The only Shield that will protect against the possi-
bility  of  backfire  while  summoning  Marquis  is  a 
disk  of  fine  silver  inscribed  with the  names  of  the 
Marquis of the seventh plane. This shield may only 
be  manufactured  by  a  Shaping  mage.  It  weighs  3 
pounds and the average cost of manufacture will be 
7000 Silver Pennies. It will take about five months 
to manufacture. 

Summoning Demonic Kings (R-6) 

Experience Multiple: 600 
Effects:  This  ritual  is  used  for  summoning  the 
following  demons  from  the  seventh  plane:  Asmo-
day,  Bael,  Balam,  Beleth,  Belial,  Palmon,  Purson, 
Vine and Zagan. These demons may only be sum-
moned  between  9  in  the  morning  and  noon  and 
between 3 in the afternoon and sunset. 

The only shield that will protect against the possi-
bility of backfire while summoning Kings is a disk 
of hammered gold inscribed with the names of the 
Kings  of  the  seventh  plane.  This  Shield  may  only 
be  manufactured  by  Shaping  magicians.  It  weighs 
three  pounds  and  the  average  cost  of  manufacture 
will be 15,000 Silver Pennies. It will take about six 
months to manufacture. 

24.8 Additional Notes 

Hexagram 

A Hexagram is a symbol drawn on a parchment of 
calf’s skin, covered with a cloth of fine white linen, 
and  draped  from  the  girdle  of  lion’s  skin  outside 
the  white  robe  worn  by  a  Greater  Summoner.  It 
helps  cause  the  demon  to  take  physical  form  and 
compels them to be obedient. 

Counterspells 

Counterspells,  as  they  relate  to  the  College  of 
Greater Summoning, are used as a form of banish-
ment. When  a  General  Knowledge  Counterspell  is 
cast  over  an  Imp,  Devil,  Succubi,  Incubi,  or  Hero 
by  the  Adept  who  either  was  granted  the  compan-
ion or summoned the entity, that entity or compan-
ion is banished back to the dimension from whence 
it  came.  A  counterspell  cast  by  anyone  else  will 
have no effect whatsoever. If a Special Knowledge 
Counterspell  is  cast  over  a  Summoned  Demon  by 
the Adept that Demon is banished back to the plane 
from whence it came. It should be emphasised here 
that  counterspells  can  only  be  passively  resisted, 
and  it  is  up  to  the  GM  to  decide  which  Demons, 
once summoned to this plane, will resist being sent 
back. 

Knowing the Past, Present or Future 

In  reference  to  Imps  and  Devils,  the  7%  Base 
Chance of knowing past, present and future events 
represents  the  chance  they  will  know  the  correct 
answer  to  a  specific  question  posed  to  them.  It  is 
up to  the  GM  to  perform the  roll  and  decide  what 
their  answer  (if  any)  will  be  if  the  result  is  above 
the Base Chance. 

\end{Chapter}
