\begin{Skill}[3.0]{warrior}{Warrior}

Warriors train in advanced combat and weapon techniques, and practice
the art of melee combat until it becomes a series of finely honed
instincts.  They practise on all terrains with all types of weapons
against different configurations and tactics, so that they may always
respond optimally without the need for thought.

Warriors usually join and train within one Warrior Guild (or school),
most guilds are exclusive and will not train those of other guilds,
many have strong ethical codes such as Chivalry and training is often
only available after indoctrination in the code.

\section{Restrictions}

A Warrior may never train in the Warrior skill without a training
partner of at least equal Warrior rank.

Assassin teaches a method of fighting that does not stack with
Warrior.  The benefits of Warrior OR Assassin (not both) may be
applied to any specific attack or defence.

A Warrior must have at least 3 weapons at rank 4 or higher including 1
melee rated weapon and 1 close rated weapon.

A Warrior requires various minimum weapons skills per Rank.  Before
being able to achieve any new Rank in the warrior skill, a Warrior
must achieve the Ranks summarised in table 49.5 below.

\section{Benefits}

A Warrior gains a bonus to defence in melee and close combat of 2\%
(+2\% per rank).

When using ranked melee or close rated weapons, or weapons from a
category where they have a higher ranked weapon, a Warrior:

\begin{Itemize}
\item Gains +1 Engaged IV per rank (no benefit at rank 0).
\item Gains +2\% (+2\% per rank) to Strike Chance.
\item Gains +1 damage at ranks 5 \& 10.
\item Gains +1 effective weapon rank for the Parry Calculation at ranks 4 \& 8. 
\end{Itemize}

For categories where the Warrior has a ranked weapon, they:

\begin{Itemize}

\item may use any weapon from the category at half the Rank of their
  maximum Ranked weapon in that category or at half their warrior
  Rank, whichever is lower.

\item may Rank weapons in the category without a tutor for an
  additional 10\% EP cost.

\item is treated as a merchant of their warrior Rank when attempting
  to buy or value non-magical weapons, with which they have ranks.  If
  they have a passing familiarity with the weapon (i.e.  has Ranks in
  the same category but not the actual weapon) they will act as a
  merchant of half their warrior Rank.

\end{Itemize}

A Rank 5 Warrior may start to acquire Warrior Special Abilities.

While some abilities are general combat abilities and only need to be
learned once, most are weapon specific and must be learned once per
weapon.  A warrior gains one special ability at ranks 5, 6, \& 7, two
at ranks 8, 9, \& 10.

Additional special abilities may be gained without increasing in rank
by the expenditure of 10,000 Experience Points and 8 weeks of training
per ability.  These costs are discounted by 25\% if the warrior has
reached rank 8, or by 50\% if they have reached rank 10.

\section{Warrior Special Abilities}

\begin{Description}
\item[Close-Evasion (General)] May Evade while in close combat.
  Standard defence bonuses for evasion and potential Parry.

\item[Close-Withdraw (General)] May add their Warrior Rank to their
  Strength for the purposes of Withdraw from Close.

\item[Pre-Engage (General)] May choose to act on their Engaged IV
  provided they are within half Warrior rank hexes of an engagement
  and they are joining the engagement as their action.

\item[Two Step (Weapon)] May move an additional 1 TMR before executing
  a Melee Attack, Special Attack, or Charge.

\item[Advanced Multi-Hex (Weapon)] Advanced training in striking multiple
opponents. Reduces Multi-Hex Strike SC\% penalty by 2\% per rank.

\item[Double-Hex Strike (Weapon)] May make a Special Attack (as per
  Multi-Hex Strike) against 2 adjacent opponents with a 1-handed B or
  C Class weapon.  May be combined with Advanced Multi-Hex to reduce
  the SC penalty.

\item[Multi-Hex Bash (Weapon)] May make a Special Attack (as per
  Multi-Hex Strike) with a 2-handed C Class weapon.  May be combined
  with Advanced Multi-Hex to reduce the SC penalty.

\item[Draw and Strike (Weapon)] May combine a Prepare Weapon Pass
  Action and Melee Attack as a Special Attack.

\item[Off-hand (Weapon)] Special training in using the weapon in
  either hand, treated as ambidextrous when using this weapon.

\item[Overstrike (Weapon)] May apply damage bonuses for both skill and
  strength as a Special Attack.

\item[Repulse (Weapon)] Add 2 to effective weapon rank when attempting
  to repulse a closing figure.

\item[Advanced Disarm (Weapon)] Advanced training in disarming
  opponents.  Reduces Disarm SC\% penalty by 2\% per rank.  In
  addition if the Disarm attempt is an EN hit then the chance of
  retaining the weapon is halved, on a Specific Grievous the target
  may not attempt to retain the weapon.

\item[Strong Guard (Weapon)] Add 2 to effective weapon rank when
  defending against a Parry/Riposte.  And may add 2\% per Warrior Rank
  to chance of resisting a Disarm.

\item[True Riposte (Weapon)] May add 1 per 2 full ranks in Warrior to
  the Parry/Riposte Calculation when evading.

\item[Shield Block] Shield an adjacent figure as a free action.  Loose
  all defence from their shield and contribute half of it to an
  adjacent target for the remainder of the pulse.

\end{Description}

\subsection{Quasi-Magical Warrior Abilities}

These abilities cost Fatigue to use (exertion/tiredness FT, not
damage).  They may only be performed if the FT cost can be paid.

\begin{Description}

\item[Fortitude] Costs 2 FT, must be declared at the start of the
  pulse.  Adds Warrior Rank to the amount of damage required to be
  Stunned and 2\% per rank bonus to fear checks for the remainder of
  the pulse.

\item[Unstoppable] Costs 2 FT.  May re-roll a failed Consciousness or
  Stun Recovery check.  Or, may make a 1 × WP check each pulse to
  stay conscious when between 0 and − Rank Endurance.

\item[Quick Stand] Costs 2 FT.  May leap to their feet from a Prone
  position as a free action.

\item[Solid Strike (Weapon)] Costs 1 FT per damage D10 rolled, must be
  declared before making the attack roll.  Any roll of less than half
  Warrior rank is increased to half Warrior rank, i.e. a damage dice
  roll of 2 by a rank 7 Warrior becomes 4.

\item[Lightning Strike (Weapon)] Costs 1 FT, must be declared at the
  start of the pulse.  May add twice Warrior Rank to engaged
  initiative for a pulse.

\item[Full Defence (Weapon)] Costs 1 FT, must be declared at the start
  of the pulse. Add Warrior rank to defence, or twice Warrior rank
  when evading.

\end{Description}

\begin{table*}
\section{Weapon Categories}

\medskip

Note: These definitions only apply to the warrior skill, not combat in general. 

\medskip

\begin{dqtblr}{measure=vbox,colspec={|X|X|X|X|},vlines={green7}}
1. Shortswords
\begin{Itemize}
\item Dagger
\item Main Gauche
\item Shortsword
\item Sai
\end{Itemize}

&

4. A Class Swords  
\begin{Itemize}
\item Rapier
\item Estoc
\end{Itemize}

&

7. Blunt
\begin{Itemize}
\item War Club  
\item Mace  
\item War Hammer  
\item War Pick  
\item Mattock 
\end{Itemize}

&

10. B Class Pole
\begin{Itemize}
\item Halberd  
\item Poleaxe
\item Glaive
\item Quarterstaff
\end{Itemize}
\\

2. Single Edged
\begin{Itemize}
\item Tulwar
\item Falchion
\item Scimitar
\item Sabre
\end{Itemize}

&

5. Oriental Swords
\begin{Itemize}
\item Katana/Wakizashi
\item O-Dachi
\item Ninjato
\end{Itemize}

&

8. Entangling / Chain
\begin{Itemize}
\item Flail
\item Morningstar  
\item Manriki Kusari
\item Bola
\item Net
\item Nunchaku
\end{Itemize}

&

11. Unarmed
\begin{Itemize}
\item Unarmed
\item Cestus
\end{Itemize}
\\

3. Double Edged
\begin{Itemize}
\item Hand \& a half
\item Two-Handed Sword
\item Claymore
\item Broadsword
\end{Itemize}

&

6. Axes
\begin{Itemize}
\item Hand Axe
\item Battle Axe  
\item Great Axe 
\end{Itemize}

&

9. A Class Pole
\begin{Itemize}
\item Javelin
\item Spear
\item Pike
\item Lance
\item Trident
\end{Itemize}

&

12. Shield
\begin{Itemize}
\item Tower  
\item Kite
\item Large Round
\item Small Round  
\item Buckler
\end{Itemize}
\\
\end{dqtblr}

\end{table*}

\begin{table*}
\section{Ranking Requirements}

\medskip

\begin{dqtblr}{colspec={|l|l|l|l|l|l|l|},vlines={green7}}
Warrior Rank & \# Categories & Minimum Rank & \# weapons $\geq$ Rank 4 & \# weapons $\geq$ Rank 5 & \# weapons $\geq$ Rank 6 & \# weapons $\geq$ Rank 7 \\
0	& 4	& 0	& 1	& 	& 	& \\
1	& 4	& 1	& 	& 1	& 	& \\
2	& 7	& 1	& 	& 1	& 	& \\
3	& 4	& 2	& 	& 2	& 	& \\
4	& 7	& 2	& 	& 2	& 	& \\
5	& 4	& 3	& 	& 	& 1	& \\
6	& 7	& 3	& 	& 	& 2	& \\
7	& 4	& 4	& 	& 	& 	& 1 \\
8	& 7	& 4	& 	& 	& 	& 2 \\
9	& 4	& 5	& 	& 	& 	& 3 \\
10	& 7	& 5	& 	& 	& 	& 4 \\
\end{dqtblr}
\end{table*}

\end{Skill}
