\begin{Chapter}{The College of Water Magics (Ver 1.3)}

The College of Water Magics is concerned with the 
shaping  of  the  element  of  water.  Water  Mages,  as 
Adepts of this college are known, have difficulty in 
practising  their  magics  on  land  without  an  abun-
dant source of their raw material and thus the col-
lege  has  fallen  somewhat  into  disuse  in  many  lo-
cales.  The  typical  vision  people  have  of  Water 
Mages is of an Adept standing on the sea or guid-
ing a ship around reefs to a safe harbour. While it 
is  true  that  some  Water  Mages  do  perform  such 
duties on ships, the majority either live beside large 
bodies of water or in more recent times have taken 
to  living  entirely  underwater,  thus  spending  all  of 
their  time  in  a  different  world  from  that  experi-
enced  by  normal  people.  In  general,  Water  Mages 
enjoy  good  relations  with  other  colleges  with  the 
notable  exceptions  of  the  elemental  College  of 
Fire, due to its opposing elemental nature. 

Traditional Colours 

The traditional colours worn by Water Mages have 
tended towards the greens and blues that symbolise 
the  sea.  Aquamarine  and  sea  blue  are  particularly 
popular amongst those who wish to advertise their 
college.  The  hermitic  type  of  Water  Mage  who 
spends  the  majority  of  their  time  underwater  usu-
ally  wear  very  little,  a  small  bathing  suit  at  the 
most,  as  most  denizens  of  the  deep  are  not  con-
cerned about nudity. 

Traditional Symbols 

Water  Mage  adornments  are  usually  unique  items 
found only upon the sea floor such as rare shells or 
pieces  of  coral,  also  wave,  fish  and  whale  motifs 
are popular or any animal or occurrence specific to 
the sea they live in. 

23.1 Restrictions 
Adepts  of  the  College  of  Water  Magics  may  only 
practice  their  arts  if  they  are  in  contact  with,  or 
near,  water.  With  the  exception  of  Dowsing,  they 
may  never  practice  their  arts  in  a  vacuum  or  in  a 
totally  arid  place.  They  may  not  summon  water-
dwelling  creatures  into  an  area  that  does  not  con-
tain  a  body  of  water  large  enough  for  the  water-
dwelling  creature  to  immerse  itself  entirely.  They 
may  use  their  magic  on  land  (in  a  non-arid  area) 
but  suffer  some  diminution  in  their  abilities.  A 
character  must  have  18  MA  to  become  a  water 
mage. 

23.2 Base Chance Modifiers 
The  Base  Chance  of  performing  any  Talent,  Spell 
or  Ritual  of  this  College  is  modified  by  the  addi-
tion  of  the  following  numbers.  Apply  one  of  the 
following: 

+20 
+15 
+10 
+5 
+5  

Adept is completely immersed in water 
Adept is in physical contact with water 
Adept is within 25 feet of water 
Adept is within 500 feet of water 
Adept is surrounded by mist or fog, or is 
standing in the rain 
Adept is on land and over 1 mile from 
water 
Adept is on land and over 10 miles from 
water 
This  modifier  may  be  applied  in  conjunction  with 
one of the above: 

-15 

-5 

All  water  in  the  vicinity  of  the  Adept  is 
frozen 

-
20 

Notes 

1. For the purpose of these modifiers “water” refers 
to a body of water such as a sea, lake, ocean, pond, 
river,  stream,  spring  or  other  feature  containing 
large  amounts  of  water  (1000  gallons  or  more)  or 
existing  as  part  of  a  larger  system  or  network  of 
waterways.  A  barrel,  bucket,  or  tun  of  water  does 
not qualify as a body of water. 

2. Apply the most advantageous modifier from the 
first  group  in  conjunction  with  the  frozen  water 
modifier. 

23.3 Talents  

Predict Weather (T-1)  

Range: 10 miles (+10 / Rank) 
Experience Multiple: 75 
Base Chance: MA + 5% / Rank 
Effects:  The  Adept  predicts  with  some  accuracy 
what  the  weather  will  be  like  over  the  ensuing 
three days in the area within the range of the talent. 
The Adept must be at sea (on an ocean, sea or large 
lake)  or  within  15  miles  of  an  ocean,  sea  or  large 
lake.  If  the  Adept  makes  a  successful  Predict 
Weather check, the GM rolls D100 and checks the 
accuracy  of  the  prediction.  The  following  results 
may occur: 

Dice 

Accuracy 

01–05 

Totally wrong (opposite of the prediction 
occurs) 

06–10  Generally incorrect (fairly wide diver-

gence) 

11–85  Generally correct (close but not totally 

accurate) 
Almost totally accurate (say within 1 
degree) 

86–
100 
Following the dice roll the GM delivers the predic-
tion as though it were generally correct. 

Resist: None 
Storage: Investment, Potion 
Target: Entity 
Effects:  This  spell  protects  the  target  from  the 
effects  of  temperature  down  to  0°C  -  2  /  Rank.  It 
will  totally  protect  the  target  from  the  effects  of 
Hypothermia.  In  addition,  the  target  takes  1  (+1/5 
Ranks  (round  up))  less  damage  from  magical  or 
non-magical cold-based attacks. 

Create Fog (G-3) 

Range: 20 feet + 20 / Rank 
Duration: 6 minutes + 6 / Rank 
Experience Multiple: 100 
Base Chance: 20% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Volume of Air 
Effects: This spell allows the Adept to create 1000 
cubic  feet  (+500  /  Rank)  of  fog  entirely  within 
range.  The  fog  must  be  continuous  and  connected 
to  the  surface  above  which  it  is conjured.  In  addi-
tion the conjured mist has the effect that all entities 
within  the  fog  have  their  Strike  Chances  reduced 
by 5% (+1 / Rank) due to limited visibility. Visibil-
ity is reduced to 1 hex. The spell of Water Breath-
ing completely negates the effect of this spell. 

Dowsing (T-2) 

Mage Current (G-4) 

Experience Multiple: 75 
Base Chance: 2 × MA + 4% / Rank 
Effects: The Adept is able to sense the presence of 
water  above  ground,  and  within  100  feet  (+  100  / 
Rank).  The  Adept  can determine  the direction and 
distance  to  the  nearest  source  of  water.  Detection 
of  underground  water  has  a  range  of  1  feet  (+  2  / 
Rank).  Assessment  of  freshness  increases  with 
increased  Rank,  e.g.  Rank  3  can  tell  salt  from 
fresh,  Rank  6  can  tell  alkali  from  clear,  etc.  This 
talent  may  be  exercised  when  the  Adept  is  in  a 
vacuum or  a  totally  arid  place.  In  general  the  GM 
will roll to determine the success of this talent. 

Aquatic Affinity (T-3) 

Experience Multiple: 100 
Effects:  Adepts  of  the  College  of  Water  Magic 
have a kinship with their element which gives them 
the following abilities: 

1. The Adept may modify their swimming rolls by 
4%  for  every  Rank  they  have  achieved  with  this 
talent. 

2. Above Rank 10, the Adept is considered to be an 
Aquatic  entity  for  the  purposes  of  calculating  un-
derwater defence. 

3. Movement in water is considered to be one row 
less  difficult  on  the  Encumbrance  table.  If  the 
Adept has more than Rank 10 in this talent then it 
is two levels less difficult. 

4. General knowledge of the Aquatic environment. 

23.4 General Knowledge Spells 

Buoyancy (G-1) 

Range: Touch 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 100 
Base Chance: 45% 
Resist: None 
Storage: Investment, Potion 
Target: Entity 
Effects:  This  spell  allows  the  target  to  float  at 
whatever  depth  they  choose.  If  they  want  to  float 
upon  the  surface  this  spell  enables  them  to  do  so 
without  using  any  float  actions.  The rate  of  ascent 
or descent is 5 feet (+2 additional / Rank) and this 
spell  also  protects  the  target  from  any  pressure 
effects related to being at extreme depths. 

Cold Resistance (G-2) 

Range: Touch 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 150 
Base Chance: 40% 

79 

Range: 10 feet + 10 / Rank 
Duration:  Concentration:  maximum  30  minutes  + 
30 / Rank 
Experience Multiple: 125 
Base Chance: 30% 
Resist: None 
Storage: Investment 
Target: Water 
Effects:  The  Adept  creates  a  current  in  the  water 
within  the  range  (a  volume  that  moves  with  the 
Adept)  of  5  miles  per  hour  (+1  /  Rank)  in  one 
direction. All objects/entities, etc. within the range 
that  are  submerged  in  the  water  or  floating  on  the 
water  are  carried  at  the  speed  of  the  current.  The 
Adept  may  freely  alter  the  direction  of  the  mage 
current during the spell (requires active concentra-
tion). Passive concentration is required to maintain 
this spell. 

Navigation (G-5) 

Range: Self 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 125 
Base Chance: 15% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  The  spell  attunes  the  Adept’s  mind  to  the 
sea  and  winds,  allowing  them  to  sail  with  less 
chance  of  mishap.  This  decreases  the  chance  of 
veering off course, running aground, etc., by 5% (+ 
1 / Rank). In addition the Adept may add 1 (+ 1 / 4 
full  Ranks),  to their  effective  Rank  in  the  Naviga-
tor Skill. 

Rehydration (G-6) 

Range: Touch 
Duration: Immediate 
Experience Multiple: 350 
Base Chance: 30% 
Resist: None 
Storage: Investment, Magical Trap 
Target: Object 
Effects:  This  spell  rehydrates  (restores  dried  sub-
stances  to  their  normal  water  content)  including 
dehydrated entities. 

Saturated Earth (G-7) 

Range: 100 feet + 50 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 250 
Base Chance: 25% 
Resist: None 
Storage: Investment, Ward 
Target: Area 

23 COLLEGE OF WATER MAGICS 

Effects:  The  Adept  summons  water  from  deep 
within  the  earth,  rising  to  the  surface  in  100  sec-
onds (-5 / Rank), causing the ground in an area of 
100  feet  (+10  /  Rank)  radius  to  become  totally 
sodden.  Earth  will  turn  to  mud.  Sand  has  a  30% 
chance  of  turning  into  quicksand  (swallowing  any 
trapped  individuals  in  20  seconds)  and  a  70% 
chance  of  turning  into  firm  hard  sand  facilitating 
easy  movement.  The  Adept  may  alter  these 
chances  by  1%  per  Rank.  Wetlands  will  be  unaf-
fected  by  this  spell,  as  will  solid  rocky  terrain. 
With this spell, all wells in an area can be caused to 
fill and all crops will be well watered. 

Ship Strength (G-8) 

Range: 90 feet + 15 / Rank 
Duration: 5 days + 1 / Rank 
Experience Multiple: 200 
Base Chance: 20% 
Resist: None 
Storage: Investment 
Target: Ship, boat, wood 
Effects:  This  spell  may  be  used  to  strengthen  the 
structure of any ship or boat, repair leaks or holes, 
restep  masts,  etc.  At  Rank  10  and  above,  a  small 
sail craft can be constructed out of available wood. 
At Rank 15 or above, a large craft (40 feet or more 
in  length)  may  be  constructed.  At  the  end  of  the 
spell’s  duration,  its  effects  are  undone;  a  con-
structed craft will fall apart (even in mid-ocean). In 
addition,  any  craft  which  is  affected  by  this  spell 
has its chance of encountering sea monsters, pirates 
or typhoons, decreased by 5% (+1 / Rank). 

Speak to Aquatic Life (G-9) 

Range: 15 feet + 15 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 150 
Base Chance: 40% 
Resist: None 
Storage: Potion 
Target: Self 
Effects:  This  spell  allows  the  Adept  to  communi-
cate with any one form of aquatic life within range. 
This  communication  usually  consists  of  both 
speech  and  gesture.  All  members  of  the  selected 
generic  type  of  aquatic  that  are  within  speaking 
range  can  be  understood  and  can  understand  the 
Adept. The Adept may have more than one spell on 
at  once,  as  long  as  they  are  for  different  generic 
types. 

Summon Aquatic Life (G-10) 

Range: Unlimited 
Duration: Immediate 
Experience Multiple: 100 
Base Chance: 20% 
Resist: None 
Storage: Investment, Potion 
Target: Aquatic Species 
Effects: The Adept may summon 1 aquatic creature 
(+  1  /  Rank).  It  may  take  up  to  30  minutes  (30 
seconds / Rank) for the aquatic life forms to arrive. 
The  species  must  be  native  to  the  area  to  be  sum-
moned and the Adept must be within 100 feet of a 
body of water. 

Water Breathing (G-11) 

Range: Touch 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 100 
Base Chance: 25% 
Resist: None 
Storage: Investment, Potion 
Target: Entity 
Effects:  This  spell  forms  a  set  of  gills  in  the  sub-
ject’s neck and covers their eyes with a transparent 
film.  This  allows  the  target  to  breathe  and  see 
equally  well  under  water  as  on  land.  The  target 
may  cast  spells  subject  to  the  restrictions  of  their 
College. The spell does not affect the target’s abil-
ity  to  operate  on  the  surface.  This  spell  allows 
vocal  communication  to  a  range  of  the  target’s 
perception in hexes. 

Water Creation (G-12) 

Range: Touch 

Duration: Immediate 
Experience Multiple: 150 
Base Chance: 35% 
Resist: None 
Storage: Investment 
Target: Water 
Effects:  The  Adept  must  touch  the  substance  from 
which  water  is  to  be  drawn.  The  spell  allows  the 
Adept  to  extract  moisture  from  the  air,  or  from 
plants  (providing  there  is  moisture  available  to  be 
extracted) to the amount of 1 pint (+ 1 / Rank). 

Waterproofing (G-13) 

Range: Touch 
Duration: Special 
Experience Multiple: 150 
Base Chance: 30% 
Resist: None 
Storage: Investment 
Target: Entity or Object 
Effects:  This  spell  protects  any  affected  entity  or 
object  completely  from  all  forms  of  non-magical 
water  damage.  It  works  on  any  item  less  than  or 
equal  to  5  pounds  (+  10  /  Rank)  in  weight.  Thus, 
written scrolls would not run, steel would not rust, 
potions  would  not  dilute,  clothing  would  not  get 
wet  etc.  The  duration  is  5  days  (+  5  /  Rank)  for 
objects, and 3 hours (+ 3 / Rank) for entities. 

Wave Control (G-14) 

Range: 90 feet + 90 / Rank 
Duration: 15 minutes + 15 / Rank 
Experience Multiple: 125 
Base Chance: 25% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Volume of Water 
Effects:  The  Adept  can  increase  or  decrease  the 
size of all  waves  within range by 5 feet (+ 1 addi-
tional foot / Rank). 

23.5 General Knowledge Rituals 

Binding Water (Q-1) 

Range: 10 feet + 15 / Rank 
Duration: 1 hour + 1 / Rank 
Experience Multiple: 750 
Base Chance: MA + 4% / Rank 
Resist: Special 
Effects:  The  Adept  can  bind  the  element  of  water 
while  maintaining  their  concentration.  It  takes  an 
hour  to  perform  this  ritual.  The  results  are  similar 
to  the  binding  of  all  other  elements.  The  Adept 
gains control of all facets of the element. They can, 
for  example,  create  an  intelligent  water  sprite 
(which  will  always  have  characteristics  several 
points  lower  than  the  Adept),  however,  its  every 
action would have to be directed by the Adept. At 
Rank  10  or  higher,  the  Adept  can  create  a  free-
willed  water  sprite  that  will  be  loyal  to  them  and 
not  require  its  every  move  be  directed.  Such  enti-
ties  will  never  leave  the  water.  The  will  not  be 
resisted  unless  the  area  contains  a  Water  Elemen-
tal, which may passively and actively resist. 

23.6 Special Knowledge Spells 

Control Aquatic Life (S-1) 

Range: 10 feet + 10 / Rank 
Duration: Concentration: no maximum 
Experience Multiple: 100 
Base Chance: 20% 
Resist: Passive 
Storage: Investment 
Target: Aquatic Entity 
Effects: This spell allows the caster to control 1 (+ 
1  /  2  Ranks)  aquatic  creatures  (which  must  be  of 
the  same  generic  type).  These  creatures  will  serve 
the caster so long as the caster concentrates or until 
they are told to go away (move out of range of the 
spell).  A  creature  that  is  no  longer  controlled  but 
still in the vicinity may attack its former master. If 
the target is of a particularly small, schooling spe-
cies, the Adept may be able to control the whole or 
part  of  the  school.  This  spell  cannot  be  used  to 
control  sentient  creatures.  Aquatic  Mammals  are 
usually non-sentient. 

80 

Dehydration (S-2) 

Range: 20 feet + 20 / Rank 
Duration: Immediate 
Experience Multiple: 500 
Base Chance: 1% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Object, Entity 
Effects:  The  target  of  this  spell  must  successfully 
resist  or  have  all  of  the  water  removed  from  their 
body,  resulting  in  instant  mummification.  The 
target’s  corpse  will  remain  preserved  indefinitely 
(if  kept  dry,  since  little  or  no  tissue  damage  oc-
curs). After rehydration the body will be able to be 
resurrected  as  if  it  had  just  died.  If  the  target  is 
formerly or never living, the Adept receives +40% 
on the Base Chance. When used in this manner the 
spell can be used to dehydrate such things as food 
for  preservation,  such  things  lasting  indefinitely  if 
kept dry. 

Flash Flood (S-3) 

Range: 600 feet + 600 / Rank 
Duration: 30 seconds + 30 / Rank 
Experience Multiple: 500 
Base Chance: 2% 
Resist: None 
Storage: Investment, Ward, Magical Trap 
Target: Watercourse 
Effects: The Adept causes a particular watercourse 
within  range  to  swell  and  burst  its  banks.  The 
watercourse  can  be  a  stream,  dry  river  bed,  small 
or  large  river  (i.e.  anywhere  that  might  be  subject 
to  such  an  occurrence  naturally,  including  drains 
and  sewers).  The  flood  will  occur  with  very  little 
warning.  After  30  seconds  of  low  rumblings,  the 
water  level  will suddenly rise, sweeping all before 
it.  The  flood  will  wipe  out  any  small  bridges  and 
dams  within  range,  wash  people  away,  unhorse 
riders,  wash  wagons  and  carts  away  (chance  of 
destruction  dependent  upon  construction).  The 
effects  are  most  noticeable  on  small  rivers  or  dry 
river beds. On a large river, the flood might appear 
as a large wave which would look rather innocuous 
at a distance  but  which  would  wreak  just as  much 
havoc.  After  the  spell  duration  expires,  the  water 
level  will drop just as quickly as it rose.  All those 
caught  in  the  flow  must  make  a  successful  swim-
ming  roll  to  avoid  drowning.  If  a  person  is  un-
horsed they must make a horsemanship roll to stay 
with  their  horse,  in  which  case  they  may  use  the 
horse’s  Rank  in  swimming  (generally  8)  to  deter-
mine drowning. 

Geyser (S-4) 

Range: 20 feet + 20 / Rank 
Duration: 30 seconds + 30 / Rank 
Experience Multiple: 350 
Base Chance: 15% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Ground 
Effects: The Adept calls forth from the ground 1 (+ 
1 / Rank) jets of hot steam and mud. Each jet must 
appear  in  an  adjacent hex  to  at  least one  other  jet. 
Any  Entity  occupying  an  affected  hex  must  resist 
or suffer [D - 4] (+ 1 / 2 or fraction Ranks), half if 
resisted  (round  down),  damage,  per  pulse.  If  a 
character resists they are allowed an automatic hex 
of  movement.  It  is  necessary  to  resist  for  each 
separate hex passed through, however only 1 resis-
tance  check  is  necessary  for  an  individual  hex 
regardless of the duration it is occupied for. Magi-
cal Waterproofing  or  Protection  from Normal  Fire 
will  shield  a  character  completely  from  any  dam-
age. 

Liquid Purification (S-5) 

Range: 15 feet 
Duration: Immediate 
Experience Multiple: 350 
Base Chance: 30% 
Resist: None 
Storage: Investment 
Target: Aqueous Liquid 

Effects: The Adept may turn 1 quart (+ 1 / Rank) of 
any aqueous liquid into drinkable water. 

storm front will take D10 × 3 (1 / Rank) minutes to 
arrive. 

Liquid Transmutation (S-6) 

Walk on Water (S-10) 

Range: Touch 
Duration: Immediate 
Experience Multiple: 350 
Base Chance: 30% 
Resist: None 
Storage: Investment 
Target: Potable water 
Effects:  The  Adept  may  turn  1  pint  of  drinkable 
water into any other common liquid of their choice. 
The quality of the transformed liquid is dependent 
upon  Rank  (Rank  0,  a  wine  might  be  made  that 
was just drinkable, Rank 10 a Silver or Gold Medal 
young  wine,  Rank  20,  an  unsurpassed  wine  of  its 
type). 

Maelstrom (S-7) 

Range: 30 feet + 30 / Rank 
Duration: 10 seconds + 10 / Rank 
Experience Multiple: 500 
Base Chance: 10% 
Resist: Passive 
Storage: Investment, Ward, Magical Trap 
Target: Volume of Water 
Effects:  The  Adept  creates  a  horrifying  watery 
vortex  with  a  diameter  of  10  feet  (+  10  /  Rank) 
which  exists  entirely  within  the  spell’s  range.  All 
objects  and  entities  within  20  feet  of  the  vortex 
must successfully resist or they are sucked into the 
eye of the vortex and down to the sea bottom. This 
spell will only be effective if cast over a large body 
of water (sea, ocean or lake). 

Rainstorm (S-8) 

Range: 20 feet + 20 / Rank 
Duration: 30 minutes + 30 / Rank 
Experience Multiple: 300 
Base Chance: 20% 
Resist: Active, Passive 
Storage: Investment, Ward, Magical Trap 
Target: Entity 
Effects: The caster summons a miniature rainstorm, 
radius  5  feet  (+  1  /  Rank),  that  appears  over  the 
target’s  head  (little  dark  rain  cloud,  tiny  lightning 
bolts and all). If the target fails to resist, the storm 
will  follow  them  for  the  duration  (even  inside 
buildings),  completely  soaking  everything  not 
waterproof  in  60  seconds  (5  /  Rank).  The  distrac-
tion  caused  by  the  storm will  add 2%  per Rank  to 
any  rolls  the  target  makes  that  are  of  a  mental 
nature (i.e. require thought or concentration) or 1% 
per Rank to other rolls. All characters in the area of 
effect  will  suffer  these  effects  although  the  effects 
are  halved  if  they  are  not  the  actual  target  of  the 
spell.  If  the  target  is  waterproofed  the  effects  will 
be minimal and if they are a Water Mage they will 
suffer  no  ill  effect,  but  will  have  their  Base 
Chances increased by 5%. If the Adept chooses to 
cast this  spell  on  himself  the  duration  is  increased 
to 24 hours. 

Storm Calling (S-9) 

Range: Works at any Range 
Duration: Special 
Experience Multiple: 200 
Base Chance: 40% 
Resist: None 
Storage: Investment 
Target: Storm Front 
Effects:  The  Adept  may  summon  any  storm  front 
which may exist anywhere in sight. Upon reaching 
the spot occupied by the Adept at the time of cast-
ing,  the  storm  front  will  slow  and  finally  cease 
moving  and  begin  to  downpour  (snow,  rain,  sleet, 
or whatever else the GM feels clouds may contain). 
Generally  a  storm  front  can  be  seen  for  20  to  30 
miles. If no front can be seen, the spell may still be 
cast  but  the  Base  Chance  is  reduced  by  20.  The 

Range: 10 feet + 10 / Rank 
Duration: 5 minutes + 5 / Rank 
Experience Multiple: 200 
Base Chance: 30% 
Resist: None 
Storage: Investment, Potion 
Target: Entity 
Effects: The target is able to walk on the surface of 
water  as  if  it  were  solid  and  non-slip.  They  may 
travel at their normal TMR. Reduce TMR by 1 for 
every foot above 3 in wave height. In rough waters, 
agility  rolls  may  be  deemed  necessary  (unless 
crawling). 

Waters of Healing (S-11) 

Range: Touch 
Duration: Special 
Experience Multiple: 400 
Base Chance: 30% 
Resist: None 
Storage: Special 
Target: Pint of water 
Effects:  The  Adept  can  turn 1  pint  of  water  into  a 
half  pint  healing  potion.  This  potion  will  immedi-
ately  neutralise  the  effects  of  any  venom,  plus  it 
will  cure  1  point  of  Endurance  loss  due  to  the 
venom.  Or,  if  there  are  no  venoms  present,  the 
Potion will cure [D - 5] (+ 1 / Rank) damage (En-
durance  then  Fatigue).  The  Potion  will  last  2  min-
utes (+ 2 / Rank). Or the Adept may spend an hour 
and  utilise  materials  costing  200sp  to  make  a  Po-
tion with the same effects that will last indefinitely. 
The  ingredients  are  used  up  regardless  of  success 
or failure. 

Waters of Strength (S-12) 

Range: Touch 
Duration: Special 
Experience Multiple: 350 
Base Chance: 30% 
Resist: None 
Storage: Special 
Target: Pint of water 
Effects:  The  Adept  can  turn 1  pint  of  water  into  a 
half  pint  strength potion.  This  potion  adds  [D  -  2] 
(+ 1 / Rank) to Physical Strength for 5 minutes (+ 5 
/  Rank).  Potions  are  not  cumulative.  The  Potion 
will last 2 minutes (+ 2 / Rank). Or the Adept may 
spend  an  hour  and  utilise  materials  costing  500sp 
to make a Potion with the same effects that will last 
indefinitely. The ingredients are used regardless of 
success or failure. 

Waters of Vision (S-13) 

Range: Touch 
Duration: 10 seconds 
Experience Multiple: 250 
Base Chance: 25% 
Resist: None 
Storage: Investment 
Target: Pool of water 
Effects: The Adept must touch a pool of water with 
their  hand.  They  may  then  view  visions  (usually 
precognitive  in  nature)  concocted  by  the  GM.  At 
Rank 5 and  above,  they  may  use  this  technique  to 
spy into an area to see what is going on there. The 
maximum  distance  from  the  character  to  the  area 
being spied into is 5 miles (+ 15 / Rank). The point 
of  view  of  the  Waters  cannot  be  changed.  If  the 
Adept  is  not  able  to  form a  clear  mental  image  of 
the  location  to  be  spied  upon  or  command  the 
waters to focus on an unambiguously defined point 
in  space,  then it is  the  GM’s  discretion  as to  what 
will be seen 

Waterspout (S-14) 

Range: 60 feet + 60 / Rank 

23 COLLEGE OF WATER MAGICS 

Duration: 30 seconds + 30 / Rank 
Experience Multiple: 750 
Base Chance: 5% 
Resist: Passive 
Storage: Investment, Ward, Magical Trap 
Target: Volume of water 
Effects: The Adept creates a tornado-like formation 
with a radius of 5 feet (+ 5 / Rank), over a body of 
water.  The  waterspout  kills  anyone  and  destroys 
anything  which  occupies  the  same  space  with  it 
unless  they  successfully  resist.  The  waterspout 
does  [D  -  1]  (+  1  /  Rank)  damage  to  a  character 
who  resists  successfully 
inflicting 
enough damage to kill them. A character need only 
resist the waterspout once. Boats and ships will be 
affected  at  the  GM’s  discretion.  Small  boats  may 
be  broken  into  match  sticks  and  will  almost  cer-
tainly capsize, large ships may suffer less damage, 
but  lose  rigging,  masts  etc.  If  the  Adept  concen-
trates they may move the  waterspout in any direc-
tion they desire at a rate of 10 mph (+ 5 / Rank). It 
may never be moved onto dry land. 

instead  of 

Wave Riding (S-15) 

Range: 10 feet + 10 / Rank 
Duration:  Concentration:  maximum  30  minutes  + 
30 / Rank 
Experience Multiple: 300 
Base Chance: 20% 
Resist: None 
Storage: Investment 
Target: Entity 
Effects: This spell forms a wave of water under the 
target  controlled  by  the  Adept,  allowing  travel  up 
to 10 mph (+ 2 / Rank). This spell will only form a 
wave  on  a  suitably  sized  body  of  water.  Speed  of 
current  will  add/subtract  to  speed,  wind  will  not. 
The  wave  requires  passive  concentration  to  main-
tain  and  to  keep  moving  in  the  required  direction. 
The  wave  dissipates  if  the  Adept  loses  concentra-
tion. Multiple casts of this spell may be maintained 
by  the  same  passive  concentration  action,  thus 
several  targets  may  ride  the  same  wave  (not  sepa-
rate ones). 

23.7 Special Knowledge Rituals  

Summoning and Binding Water Elemental 
(R-1)  

Range: 20 feet 
Duration: passive concentration 
Experience Multiple: 250 
Base Chance: MA +5% / Rank +1% WP above 15 
Resist: None 
Effects: The Adept may summon a Water Elemen-
tal and bind it to temporary service by performing 
this  ritual.  The  ritual takes  two  hours.  It  may  only 
be performed if the summoner is in contact with, or 
within 20 feet of, a large body of water (ocean, sea 
or  lake)  and the  summoner  remains  stationary  and 
takes  no  other  action  during  the  time  the  ritual  is 
being performed. At the end of the two hour ritual, 
the  player  makes  a  Cast  Check.  If  the  ritual  is 
successful  the  Elemental  is  summoned  and  con-
trolled. If the ritual backfires then the Elemental is 
summoned  but  not  controlled  and  will  attack  the 
summoner and their friends. 

A  Water  Elemental  always  appears  within  20  feet 
of the summoner. It has a combined Endurance and 
Fatigue  equal  to  15  (+15  /  Rank).  The  Elemental 
will  remain  until  it  is  sent  back  to  its  own  dimen-
sion  by  the  Adept  (with  a  Special  Knowledge 
Counterspell  of  the  College  of  Water  Magics)  or 
banished. If it is controlled by the summoner it will 
remain controlled until the summoner’s concentra-
tion is broken. 

\end{Chapter}
