\begin{Skill}[1.1]{weaponsmith}{Weaponsmith}

\section{Restrictions}

The skill is related to that of armourer, and a weaponsmith who is a
more skilled armourer expends only three-quarters of the necessary
Experience Points to acquire or improve this skill.  The reverse is
also true.

A weaponsmith’s progress in their skill is inhibited by a low Manual
Dexterity, and aided by a high Manual Dexterity.  A weaponsmith has an
increased Experience Point cost of 5\% for each point of Manual
Dexterity less than 16.  A weaponsmith decreases their Experience
point cost by 5\% for each point of Manual Dexterity greater than
20. A weaponsmith will have great difficulty passing their
apprenticeship if their Manual Dexterity is less than 10.

\section{Benefits}

A weaponsmith acquires one ability every two Ranks.  The character
begins with one of the following abilities at Rank 0.  All abilities
can be performed skillfully.

\begin{Itemize}
\item  Make and maintain swords including daggers. 
\item  Make and maintain hafted weapons. 
\item  Make and maintain thrown weapons. 
\item  Make and maintain pole weapons. 
\item  Make and maintain missile weapons. 
\item  Make and maintain entangling weapons.
\item  Make and maintain experimental weapons. 
\item  Make and maintain siege engines. 
\item  Make and maintain shields. 
\end{Itemize}

Additional abilities may be gained without increasing in rank by the
expenditure of 5,000 Experience Points and 4 weeks of training per
ability.  These costs are discounted by 25\% if the weaponsmith has
reached rank 8, or by 50\% if they have reached rank 10.

\precis{A weaponsmith can build increasingly more effective weapons as
  their Rank increases.}

For every Rank that a weaponsmith achieves, they may create weapons
that have an increased base chance of 1\%.  For every Rank divisible
by five that a weaponsmith achieves, they may create weapons that
cause an extra point of damage.  These two effects are not cumulative.

For example, a Rank 8 weaponsmith may construct a weapon with a Base
Chance increased by 3\% and a Damage Modifier increased by 1, or a
weapon with a BC increased by 8\% and no increase in DM.

Note: The weapons statistics as shown in the weapons chart are
manufactured at an effective Rank of 0 i.e. they are the mass-produced
variety.  They may have been manufactured by a weaponsmith of greater
Rank than this, but the skill level used was elementary.

The time and cost required for a weaponsmith to construct a weapon is
dependent on the Rank that is used, and the type of weapon.

\precis{A Weapon may be manufactured at any Rank up to the
  weaponsmith’s Rank.}

\begin{Enumerate}

\item The time required is (10 × (Effective Rank + DM)) hours, with a
  minimum of 10 hours.

\item The cost is 80\% of the Base Cost as shown in the weapons table
  × (1 + effective Rank + DM increase) silver pennies.

\item For every rank that a Weaponsmith has beyond the effective rank
  of the of the weapon, they reduce the time required, as given above,
  by 5\%.  For example, a Rank 8 weaponsmith churns out a Rank 0
  weapon in only 6 hours, or produces a +1 Damage sword, no BC
  modifier, in 51 hours rather than 60.

\end{Enumerate}

A weaponsmith is treated as a merchant of their weaponsmith Rank when
attempting to buy or value weapons which are part of their abilities.

If the equipment concerned is unfamiliar, then they operate as a
merchant of half their Rank (rounded down).

\section{Costs}

A weaponsmith, with the exception of some missile weapons, can only
perform their skill in a properly maintained workshop.

It costs 2000 silver pennies to construct a workshop and 500 silver
pennies per year to maintain it with tools and materials. A basic tool
kit will cost 100 + (100 × Rank) silver pennies.  A workshop may be
rented at a cost of 10 silver pennies per day.

\section{Silvering Weapons}

A weaponsmith may incorporate an additional metal during the
manufacture of a weapon.  This improves the appearance of the weapon,
and may provide other benefits as given elsewhere in the rules.

Cost is 80\% of Base Cost as shown in the weapons table × (metal
factor + effective rank + DM increase).  The metal factor is 1 for
cold iron weapons, 10 for silvered, 120 for gilded, and 180 for
truesilvered.

For example: a truesilvered rank 10 Hand \& a half with +2 DM costs
(80\% × 85) × (180 + 10 + 2) = 13,056sp or a gilded Rank 6 Great Axe
with +1 DM and +1\% SC costs (80\% × 30) × (120 + 6 + 1) = 3,048sp.

\end{Skill}
