\begin{Tables}{Weapons, Shields and Armour Charts}

\section{Weapons}

\subsection{Special Notes for Weapons}

\begin{multicols}{2}
Weapons are normally wielded one-handed, and the exceptions are noted
with a (2) after the name of the weapon.  Some may be used either one
or two-handed, and these are noted with a (1–2).  When weapons of this
type are wielded two handed, increase their Damage Modifier by 1.

\begin{Description}

\item[-] indicates that a weapon has no Class for purposes of Grievous
  Injuries; when a possible Grievous Injury is rolled, only damage
  affecting Endurance results.

\item[V] indicates that the characteristic is variable.

\item[A] When attacking a foe whose modified AG is between 12 and 9
  (inclusive) the weapon may be used to attack twice in one pulse
  without penalty; if the modified AG is 8 or less, the weapon may
  attack three times in a pulse.

\item[B] Only giants can use giant weapons.

\item[C] A torch is not actually a weapon, but may be used as such in
  emergencies.  Also, brandishing a burning torch in the face of an
  animal may cause it to flee. Any animal whose WP is 10 or less may
  be scared off if it fails a roll of 4 × WP.  A successful roll
  indicates the animal is not impressed. No Rank may ever be achieved
  with a torch.

\item[D] The sap may only be used to knock out targets wearing only
  leather, cloth or no armour. Used by an assassin, any hit knocks out
  the target; for anyone else, any hit stuns and 4 or more points of
  effective damage knocks out the target.  This will not work on
  targets larger than human size.

\item[E] The following weapons also function as thrown weapons:
  dagger, hand axe, battle axe, giant axe, crude club, war club, giant
  club, mace, war hammer, javelin, spear, giant spear, net, bola and
  rock.

\item[F] Up to three throwing darts may be thrown at one, two, or
  three targets in one action with no penalty.

\item[G] A boomerang returns to the thrower if it does not hit
  anything during its flight.

\item[H] A grenado is filled with any substance (manufactured by an
  alchemist) designed to burst into flames on impact.  These
  substances include greek fire, methane, and anything else the GM
  will allow.  It bursts on landing (it need not be thrown at a
  particular figure), and its effects are determined by the substance
  contained within. If a “miss” is rolled for the strike check, the GM
  should randomly determine whether the grenado landed short, long,
  left, or right of the target (or any combination thereof).

\item[I] A javelin functions as a thrown weapon unless it is launched
  by a spear thrower, in which case the spear thrower’s
  characteristics are used and it functions as a missile weapon.

\item[J] A pike may be used to melee attack any figure within two
  hexes; its melee zone extends into what would normally be the first
  hexes of the figure’s ranged zone.

\item[K] A lance may be used only by a mounted figure.

\item[L] All missile weapons must be loaded before firing; this action
  is in addition to preparing the weapon itself. A pass action must be
  taken in order to load the sling, any draw bow, the spear thrower,
  and the blowgun.  Two consecutive pass actions must be taken to load
  a crossbow (three if using a cranequin).

\item[M] Longbows may not be used while mounted, nor by small figures,
  including Dwarves \& Halflings.

\item[N] An entangling weapon may be used as a garotte in Close
  Combat.

\item[O] The whip may be used to entangle and do damage in the same
  pulse to the same target in melee combat.  In close combat, it
  functions as a garotte.  Once the target is entangled, the attacker
  may choose to leave them entangled (thus letting go of the whip), or
  disentangle the target themselves, and retain possession of the
  weapon.

\item[P] Cesti are worn on the hands and need not be prepared in order
  to be used.

\item[Q] The garotte is used to strangle the target and may only be
  used against human-sized or smaller victims.  When used by a trained
  assassin, once a successful hit has been scored, it will continue to
  do damage every pulse from then on until the victim is dead or the
  assassin has taken effective damage from either the victim or an
  outside source.  If the victim’s PS is greater than the assassin’s
  the GM may permit them to attempt to break the hold, similar to the
  attempt to restrain.  If the attempt is successful, the hold is
  broken and the assassin will have to make another successful strike
  check to continue the strangulation.  Some types of plate armour
  may, at the GM’s discretion, prevent the successful use of this
  weapon due to protection around the neck area.  A non-assassin has
  to roll a strike check every pulse to see if any damage can be done.

\item[R] Unarmed has a base chance of 2 × AG + PS over 15. The damage
  modifier is -4 (+ 1 for every 3 full points of PS over 15).

\item[S] All shot, darts, arrows, and quarrels come in appropriate
pouches or quivers of 20; the weight and cost of the pouch or quiver
is not included in the information given for the accessory.

\item[T] A cranequin is used to cock crossbows; it requires a PS of
  11 and two free hands.

\item[U] If a blowgun dart does any effective damage, the damage is
  not scored against their victim but rather they suffer the effects
  of the substance which coats the tip (poison, for instance; see
  Alchemist §30).
\end{Description}
\end{multicols}

\begingroup

\def\arraystretch{0.80}

\begin{tabularx}{\linewidth}{Xcccccccccc} \toprule
Swords			& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Dagger$^A$		& 10oz	& 7	& 10	& 40	& +0 	& 8	& A	& RMC	& 10	& 9	\\ \hline
Main gauche		& 1	& 8	& 15	& 45	& +1	& P	& A	& MC	& 20	& 10	\\ \hline
Shortsword		& 2	& 10	& 12	& 45	& +3	& P	& A	& M	& 40	& 6	\\ \hline
Falchion		& 4	& 12	& 11	& 50	& +2	& P	& B	& M	& 35	& 8	\\ \hline
Scimitar		& 4	& 11	& 15	& 50	& +3	& P	& B	& M	& 60	& 8	\\ \hline
Tulwar			& 4	& 13	& 15	& 50	& +4	& P	& B	& M	& 65	& 8	\\ \hline
Rapier			& 2	& 11	& 18	& 45	& +3	& P	& A	& M	& 35	& 10	\\ \hline
Sabre			& 3	& 14	& 15	& 60	& +3	& P	& B	& M	& 40	& 7	\\ \hline
Broadsword		& 3	& 15	& 15	& 55	& +4	& P	& B	& M	& 50	& 6	\\ \hline
Estoc			& 2	& 15	& 17	& 45	& +5	& P	& A	& M	& 65	& 9	\\ \hline
Hand \& a half (1-2)	& 6	& 17	& 16	& 60	& +5	& P	& B	& M	& 85	& 7	\\ \hline
Claymore (1-2)		& 5	& 16	& 13	& 50	& +4	& P	& B	& M	& 800	& 7	\\ \hline
Two-handed sword (2)	& 9	& 22	& 14	& 55	& +7	& P	& B	& M	& 100	& 5	\\ \midrule
Hafted weapons		& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Hand axe		& 2	& 8	& 11	& 40	& +1	& 8	& B	& RMC	& 15	& 4	\\ \hline
Battle axe (1-2)	& 5 	& 14	& 14	& 60	& +4	& 6	& B	& RM	& 20	& 7	\\ \hline
Great axe (2)		& 6	& 19	& 17	& 65	& +6	& P	& B	& M	& 30	& 7	\\ \hline
Giant axeB		& 25	& 29	& 12	& 65	& +10	& 6	& B	& RM	& 50	& 7	\\ \hline
Crude club		& 4	& 16	& 10	& 45	& +2	& 6	& C	& RM	& 3	& 2	\\ \hline
War club		& 3	& 14	& 10	& 50	& +2	& 7	& C	& RM	& 5	& 5	\\ \hline
Giant club		& 10	& 25	& 9	& 50	& +8	& 9	& C	& RM	& 10	& 5	\\ \hline
TorchC			& 3	& 8	& 12	& 40	& +1	& P	& C	& M	& 1	& -	\\ \hline
Mace			& 5	& 16	& 9	& 50	& +4	& 5	& C	& RM	& 15	& 5	\\ \hline
Giant mace$^B$		& 25	& 27	& 10	& 50	& +7	& 8	& C	& RM	& 40	& 5	\\ \hline
War hammer		& 5	& 15	& 13	& 45	& +3	& 6	& C	& RM	& 15	& 5	\\ \hline
War pick (1-2)		& 4	& 17	& 13	& 45	& +4	& P	& C	& M	& 20	& 5	\\ \hline
Flail			& 4	& 14	& 15	& 50	& +2	& P	& C	& M	& 15	& 5	\\ \hline
Morningstar (1-2)	& 5	& 18	& 15	& 60	& +4	& P	& C	& M	& 20	& 5	\\ \hline
Mattock (2)		& 6	& 19	& 14	& 55	& +6	& P	& C	& M	& 18	& 5	\\ \hline
Quarterstaff (2)	& 3	& 12	& 16	& 55	& +2	& P	& C	& M	& 3	& 9	\\ \hline
Thrown weapons$^E$		& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Throwing dart$^F$	& 3oz	& 9	& 15	& 40	& +0	& 12	& A	& R	& 1	& 10	\\ \hline
Boomerang$^G$		& 1	& 11	& 15	& 40	& +0	& 20	& C	& R	& 2	& 7	\\ \hline
Grenado$^H$		& 2	& 99	& 15	& 40	& V	& 15	& -	& R	& V	& 4	\\ \hline
Pole weapons		& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Javelin$^I$		& 3	& 12	& 15	& 45	& +2	& 12	& A	& RM	& 4	& 10	\\ \hline
Spear (1-2)		& 5	& 15	& 14	& 50	& +3	& 6	& A	& RM	& 10	& 5	\\ \hline
Giant spear(1-2)$^B$	& 15	& 22	& 16	& 55	& +7	& 12	& A	& RM	& 20	& 5	\\ \hline
Pike (2)$^J$		& 8	& 18	& 16	& 45	& +5	& P	& A	& M	& 15	& 5	\\ \hline
Lance$^K$		& 7	& 16	& 18	& 45	& +6	& P	& A	& M	& 4	& 5	\\ \hline
Halberd (2)		& 6	& 16	& 16	& 55	& +3	& P	& N	& M	& 15	& 5	\\ \hline
Poleaxe (2)		& 6	& 18	& 15	& 55	& +4	& P	& N	& M	& 20	& 5	\\ \hline
Trident (1-2)		& 5	& 14	& 16	& 45	& +2	& 5	& A	& RM	& 8	& 5	\\ \hline
Glaive (2)		& 7	& 16	& 18	& 55	& +5	& P	& B	& M	& 15	& 9	\\ \hline
Giant glaive (2)$^B$	& 14	& 22	& 18	& 65	& +9	& P	& B	& M	& 30	& 9	\\ \hline
Missile weapons$^L$	& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Sling (2)		& 1	& 7	& 15	& 40	& +1	& 60	& C	& R	& 1	& 8	\\ \hline
Self bow (2)		& 2	& 10	& 15	& 45	& +0	& 30	& A	& R	& 20	& 8	\\ \hline
Short bow (2)		& 4	& 14	& 15	& 45	& +2	& 60	& A	& R	& 20	& 8	\\ \hline
Long bow (2)$^M$	& 6	& 16	& 15	& 55	& +4	& 180	& A	& R	& 25	& 8	\\ \hline
Composite bow (2)	& 8	& 17	& 15	& 55	& +4	& 225	& A	& R	& 80	& 8	\\ \hline
Giant bow (2)$^B$	& 14	& 25	& 17	& 55	& +7	& 45	& A	& R	& 80	& 8	\\ \hline
Crossbow (2)		& 7	& 14	& 14	& 60	& +3	& 80	& A	& R	& 15	& 5	\\ \hline
Heavy crossbow (2)	& 10	& 20	& 14	& 60	& +5	& 90	& A	& R	& 20	& 5	\\ \hline
Spear thrower (2)	& 4	& 11	& 14	& 50	& +2	& 15	& A	& R	& 5	& 10	\\ \hline
Blowgun (2)$^U$		& 1	& 7	& 16	& 30	& -3	& 7	& -	& R	& 3	& 10	\\ \hline
Entangling weapons	& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
NetN			& 2	& 11	& 16	& 30	& -5	& 5	& -	& RM	& 4	& 4	\\ \hline
BolaN			& 2	& 11	& 15	& 35	& -3	& 10	& -	& R	& 5	& 6	\\ \hline
Whip$^N O$		& 3	& 10	& 16	& 40	& -3	& P	& -	& M	& 6	& 10	\\ \hline
Lasso (2)		& 4	& 12	& 18	& 40	& -4	& 6	& -	& RM	& 5	& 6	\\ \hline
Special weapons		& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Rock			& V	& 5	& 10	& 30	& -1	& 8	& C	& RMC	& -	& 6	\\ \hline
Cestus$^P$		& 3	& 12	& 14	& 35	& -1	& P	& C	& MC	& 15	& 9	\\ \hline
Garotte (2)$^Q$		& 1	& 12	& 15	& 30	& +3	& P	& -	& C	& 3	& 3	\\ \hline
Sap$^D$			& 1	& 9	& 11	& 40	& +1	& P	& C	& MC	& 2	& 3	\\ \hline
Shield			& V	& 10	& 12	& 40	& -2	& P	& C	& M	& V	& 4	\\ \hline
Unarmed			& Wt	& PS	& MD	& BC	& DM	& Range	& Class	& Use	& Cost	& Rk	\\ \midrule
Unarmed$^R$		&	& 	& 	& V	& -4	& P	& C	& MC	& 	& 10	\\ \bottomrule
\end{tabularx}

\endgroup


\begin{tabularx}{\linewidth}{Xcccl}
AccessoriesS	& Qty	& Wt	& Cost	& Notes \\
Shot		& 20	& 4	& 1	& Use in sling \\
Dart		& 20	& 2	& 5	& Use in blowgun \\
Arrows		& 20	& 2	& 10	& Use in draw bows \\
Quarrels	& 20	& 7	& 15	& Use in crossbows \\
Cranequin	& 1	& 3	& 10	& Use with crossbows \\
\end{tabularx}


\section{Shields}

\begin{tabularx}{\linewidth}{Xcccc}
Shield type	& Wt	& Def / Rank	& MD loss	& Cost \\
Buckler		& 3	& 2\%		& -		& 5 \\
Small round	& 5	& 3\%		& -2		& 8 \\
Large round	& 10	& 4\%		& -3		& 10 \\
Kite 		& 15	& 5\%		& -4		& 15 \\
Tower		& 25	& 6\%		& -6		& 20 \\
Main gauche†	& 1	& 2\%		& -		& 20 \\
\end{tabularx}

Weight The weight of the shield in pounds. 
 
Defence / Rank The percentage by which the figure’s defence is
increased per Rank while the shield is prepared (Rank 0 is counted as
a Rank†).
 
Manual Dexterity Loss The number of points the figure’s MD is reduced
by for all purposes while that shield is prepared.

 
Cost The cost in Silver Pennies for a shield of average workmanship.
 

When a shield is not prepared, it is considered slung on the back of
the figure carrying it.  All shields except the tower shield and main
gauche are constructed of wood and hides and do not affect the flow of
mana in regard to Adepts.

†The main gauche does not subtract its defence from any fire attack,
and cannot make a shield rush attack. The main gauche functions both
as a weapon and a shield, and only one EP expenditure is used to rise
in Rank in both.  No defence bonus is gained at Rank 0.


\section{Armour Chart with Guild Prices}

 
\begin{tabularx}{\linewidth}{Xccccc}
Armour Type		& Wt	& Prot	& AG mod& Cost	& Stealth \\ 
Cloth			& 1	& 1	& 0	& 30	& +5 \\
Heavy Furs		& 2	& 2	& 0	& 40	& +5 \\
Soft Leather		& 3	& 3	& 0	& 50	& 0 \\
Leather			& 3	& 4	& -1	& 50	& 0 \\
Scale†			& 4	& 5	& -3	& 600	& -5 \\
Full Scale†		& 6	& 5	& -2	& 750	& -10 \\
Chainmail†		& 7	& 6	& -2	& 1,200	& -10 \\
Partial Plate†		& 6	& 6	& -2	& 1,500	& -15 \\
Full Plate†		& 8	& 7	& -2	& 2,000	& -20 \\
Improved Plate†		& 7	& 8	& -3	& 2,850	& -20 \\
Heavy Plate†		& 8	& 9	& -3	& 3,500	& -25 \\
Jousting Armour†	& 9	& 10	& -4	& 5,000	& -30 \\
Heavy Jousting Armour†	& 15	& 15	& -8	& 12,000& -50 \\
Mithril			& 2	& 10	& -2	& Quest	& -10 \\
Dragon Skin		& 4	& ‡	& -1	& Quest	& 0 \\
\end{tabularx}

†Cold Iron  

‡As per Dragon - 3 

\begin{description}
\item[Weight] The number by which a figure’s size is multiplied to
  find the weight of the armour in pounds. Size number for the
  character races are: Halfling (3), Dwarf (4), Elf (5), Orc (6),
  Human (6), Hill Giant (9). For all others, the multiple is their
  height in feet (round up).  Females should subtract 0.5 from the
  multiples.

\item[Protection] The number of Damage Points the armour absorbs.

\item[Agility Loss] The number of points the figure’s AG is reduced
  for all purposes when the armour is worn. Does not include possible
  additional AG loss for the weight of the armour.

\item[Cost] The cost in Silver Pennies for the armour.  Costs assumes
  average workmanship and mansized armour; larger or smaller armour
  should cost proportionally more or less.

\item[Stealth Adjustment] The amount by which a figure wearing that
  type of armour has their stealth percentage adjusted.

\item[Note] Cloth armour is worn underneath all other armours and its
  protection and weight are factored into those armours.
\end{description}

\begin{itemize}
\item  Silvered Armour has the same protection – cost is + 30,000 sp.
Permits magic at -10\% to Base Chance.

\item  Truesilvered Armour has the same protection – cost is + 180,000
sp. Permits magic at no modification to Base Chance.

\item Bronze Armour has 2 points less protection – cost is the same.
Permits magic at no modification to Base Chance.
\end{itemize}

\end{Tables}
